### 2015-10-04 comporg/test1.tex
initial

\documentclass{article}
\usepackage{amsmath}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{listings}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\begin{document}
\begin{center}
\textbf{\textit{CPSC 330 Midterm Review}}\\
\end{center}

\section*{Hardware and Performance}
\hrule
\noindent\\
\indent The following equations are given:
\begin{enumerate}
\item $CPU\ Time = \frac{Clock\ Cycles}{Clock\ Rate}$
\item $CPU\ Time = Instructions \times \frac{Cycles}{Instruction} \times \frac{Seconds}{Cycle}$
\item $CPU\ Time = Clock\ Cycles \times Cycle\ Time$
\item $CPI = \frac{Clock\ Cycles}{Instructions}$
\item $CPU\ Time = Instruction\ Count \times CPI \times Cycle\ Time$
\item $CPU\ Time = Instruction\ Count \times \frac{CPI}{Clock\ Rate}$
\item $(Clock\ Cycles)_{Total} = \sum_{i = 1}^{n}\left(CPI_{i} \times Count_{i}\right)$
\item $MIPS = \frac{Instruction\ Count}{CPU\ Time \times 10^{6}}$
\item $Power = Capacitive\ Load \times (Voltage)^{2} \times Frequency$
\end{enumerate}

\hrule
\noindent \\
\indent \textit{\textbf{Question 1:}} How can we improve performance?\\
\noindent \textit{\textbf{Solution:}} There are several factors that can help to improve performance of a system. Consider the following equation:
\[
\frac{seconds}{program} = \frac{cycles}{program} \times \frac{seconds}{cycle} = \frac{cycles/program}{clock\ rate}
\]
\noindent From this equation, we can tell that there are a few ways to increase performance. For example, we can
\begin{enumerate}
\item We can \textbf{decrease} the number of cycles in a program. If we look at the right-most side of the equation, we can see that decreasing the numerator will decrease the $\frac{seconds}{program}$.
\item By a similar argument, we can say that \textbf{increasing} the clock rate will also decrease the $\frac{seconds}{program}$.
\item Increasing the cycles is the same as \textbf{decreasing} the clock cycle time.
\end{enumerate}

\hrule
\noindent \\
\indent \textit{\textbf{Question 2:}} Suppose we have two implementations of the same ISA on machines $A$ and $B$. Machine $A$ has a clock rate of 4GHz and a CPI of 2.0, and machine $B$ has a clock rate of 2GHz and a CPI of 1.2. If we run the same program on both machines, which machine is faster, and by how much?\\
\noindent \textit{\textbf{Solution:}} Recall the equations we have above. In our case, we will be looking at equation [5]. Since the program is the same on both machines, the number of instructions will cancel out. We want to find the ratio of the speeds, and then convert them into percentages.
\begin{align*}
\frac{CPU\ Time_{A}}{CPU\ Time_{B}} &= \frac{
Instruction\ Count \times \frac{CPI_{A}}{Clock\ Rate_{A}}
}{
Instruction\ Count \times \frac{CPI_{B}}{Clock\ Rate_{B}}
}\\
&= \frac{
\frac{CPI_{A}}{Clock\ Rate_{A}}
}{
\frac{CPI_{B}}{Clock\ Rate_{B}}
}\\
&= \frac{CPI_{A}\times Clock\ Rate_{B}}{Clock\ Rate_{A}\times CPI_{B}}\\
&= \frac{2.0 \times 2}{1.2 \times 4}\\
&= 83.3\%
\end{align*}
\noindent So, Machine A is $83.3\%$ faster than machine B.\\
\hrule
\newpage

\indent \textit{\textbf{Question 3:}} Two computers, $C_{1}$ and $C_{2}$ have the following metrics when running the same program:
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
& Computer $C_{1}$ & Computer $C_{2}$\\ \hline
\# Instructions & 10 billion & 8 billion\\ \hline
Clock Rate & 4GHz & 4GHz\\ \hline
CPI & 1.0 & 1.1\\ \hline
\end{tabular}
\end{center}
\noindent Which computer is faster, and which has higher MIPS?\\\\
\textit{\textbf{Solution:}} Let's address the first part of the question: \textit{Which computer is faster?}. To do this, we can use equation [6]. We will have the following:
\begin{alignat*}{3}
CPU\ Time_{1} &= 10e9 \times \frac{1.0}{4e9} \qquad\qquad CPU\ Time_{2} &&= 8e9 \times \frac{1.1}{4e9}\\
&= \frac{10}{4} &&= \frac{8.8}{4}\\
&= 2.5 &&= 2.2
\end{alignat*}
\noindent So we can see that $C_{2}$ is faster than $C_{2}$. Now lets answer the second part of the conversation. To find the MIPS, we need to use [8]:
\begin{alignat*}{3}
MIPS_{1} &= \frac{10e9}{4e9} \qquad\qquad MIPS_{2} &&= \frac{8e9}{4e9}\\
&= 2.5 &&= 2
\end{alignat*}
\noindent And we are done.\\

\hrule
\noindent \\
\indent \textit{\textbf{Question 4:}} A compiler designer is trying to decide between two code sequences for a particular machine. Based on the hardware implementation, there are three different classes of instructions: Class A, Class B, and Class C, and they require one, two, and three cycles, respectively.\\
\indent \indent The first code sequence has 5 instructions: 2 of A, 1 of B, and 2 of C\\
\indent \indent The second sequence has 6 instructions: 4 of A, 1 of B, and 1 of C\\
\noindent Which sequence will be faster, and by how much? What is the CPI for each sequence?\\\\
\textit{\textbf{Solution:}} Recall equation [4]. It tells us that the CPI will be equal to the clock cycles divided by the number of instructions. To solve this problem, we need to realize that the total number of cycles will be the sum of each cycle. In other words, we have
\begin{alignat*}{3}
CPI_{1} &= \frac{1 + 1 + 2 + 3 + 3}{5} \qquad\qquad CPI_{2} &&= \frac{1 + 1 + 1 + 1 + 2 + 3}{6}\\
&= \frac{10}{5} &&= \frac{9}{6}\\
&= 2 &&= 1.5
\end{alignat*}
\noindent So we have the CPI for each sequence. Since we have the same machine, the ratios for both [6] and [5] will have elements that cancel out. Let's use [5] in this case, as it will be simpler.
\begin{align*}
\frac{CPU\ Time_{1}}{CPU\ Time_{2}} &= \frac{5 \times 2 \times Cycle\ Time}{6 \times 1.5 \times Cycle\ Time}\\
&= \frac{10}{9}\\
&= 1.11
\end{align*}
\noindent This tells us that sequence 1 is faster than sequence 2, and in fact it is 11\% faster.\\
\hrule
\newpage

\indent At this point, we will introduce a table that shows us what factors affect the CPU Time. Recall from [5] that The number of instructions, the CPI, and the cycle time all influence the CPU Time. An ``X`` implies that the CPU time factor is directly affected, and a $\sim$ implies that it indirectly affects a factor of the CPU Time.
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
& \# Instructions & CPI & Clock Cycles \\ \hline
Algorithms & X & $\sim$ & \\ \hline
Language & X & $\sim$ & \\ \hline
Compiler & X & X & \\ \hline
ISA & X & X & X \\ \hline
Technology & & & X\\ \hline
\end{tabular}
\end{center}
\hrule
\noindent \\
\indent \textit{\textbf{Question 5:}} Consider the following table which describes the attributes of a certain machine:
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
Op & Frequency & CPI & Frequency $\times$ CPI\\ \hline
ALU & 50\% & 1 & 0.5\\ \hline
Load & 20\% & 5 & 1.0\\ \hline
Store & 10\% & 3 & 0.3\\ \hline
Branch & 20\% & 2 & 0.4\\ \hline
& & & $\sum = 2.2$\\ \hline 
\end{tabular}
\end{center}
\begin{enumerate}[\qquad\qquad a.]
\item How much faster would the machine be if a better data cache reduced the average load time to 2 cycles?
\item How does this compare with using branch prediction to shave a cycle off the branch time?
\item What if two ALU instructions could be executed at once?
\end{enumerate}
\noindent \textit{\textbf{Solution:}} Let's start with part \textit{a}. To solve this, we will have to recalculate the total sum of Frequency $\times$ CPI. Luckily, all the values will stay the same except of the CPI for the load, which is being changed to 2. The new value for Frequency $\times$ CPI is now $0.2 \times 2$, or 0.4. This makes the total sum 1.6, and the ratio of the old sum to the new sum is $2.2 : 1.6$, or 37.5\%.\\\\
\noindent We will do the same thing for part \textit{b}. In this case we will have one less cycle for the branch time, bringing the Frequency $\times$ CPI for the Branch operation to 0.2. Our new sum is 2.0, which means that we have a ratio of $2.2 : 2.0$, or 10\%\\\\
\noindent Part \textit{c} is a little different, but not too far off from \textit{a} and \textit{b}. In this case, it's not the CPI that's changing, but rather the frequency. We can process 2 ALU instructions at once, so the frequency will be half of what it currently is. In other words, we will have a frequency of 25\%, instead of 50\%. This makes the Frequency $\times$ CPI for the ALU 0.25, and the total sum is now 1.95. Now we have a ratio of $2.2 : 1.95$, or 12.8\%.\\
\hrule
\newpage

\section*{MIPS Assembly}
\hrule
\noindent\\
\indent \textit{\textbf{Question 1:}} Translate the following piece of pseudocode into MIPS instructions, with \$t0 and \$t1 as temporary registers.
\begin{center} 
\begin{tabular}{c}
\begin{lstlisting}
f = (g + h) - (i + j);
\end{lstlisting}
\end{tabular}
\end{center}
\noindent \textit{\textbf{Solution:}} This is simple. Just remember that R-Type MIPS instructions (e.g. add, sub, ...)  take three arguments: the register to store the value into, the left-hand operand, and the right-hand operand. We have as follows:
\begin{center} 
\begin{tabular}{c}
\begin{lstlisting}
add $t0, g, h
add $t1, i, j
sub f, $t0, $t1
\end{lstlisting}
\end{tabular}
\end{center}
\hrule
\noindent \\

\indent \textit{\textbf{Question 2:}} Assume A is an array of 100 words. Variables g and h are associated with registers \$s1 and \$s2 and the starting (base) address of the array is in \$s3. Use \$t0 as a temporary register. Translate the following pseudocode into MIPS assembly.
\begin{center} 
\begin{tabular}{c}
\begin{lstlisting}
g = h + A[8];
\end{lstlisting}
\end{tabular}
\end{center}
\noindent \textit{\textbf{Solution:}} The trick here is to remember that in a 32-bit system, each address in memory is separated by 4 bytes. In other words, if you want to access the 9th index of A, you need to move four spaces in memory, rather than just one. We also need to remember the syntax for an offset. We have two ways of doing this: we can use the add instruction on the address, or we can specify an offset. Since the base address of the array is stored in \$s3, we need to offset by $8 \times 4$, or 32. We will have the following:
\begin{center} 
\begin{tabular}{c}
\begin{lstlisting}
lw $t0, 32($s3)
add $s1, $s2, $t0
\end{lstlisting}
\end{tabular}
\end{center}
\hrule
\noindent \\

\indent Before we continue, lets talk about MIPS instructions in more detail. MIPS is built for a 32-bit system, and as a result, MIPS instructions are all 32 bits. We have three types of instructions, R-Type, I-Type, and J-Type. R-Type instructions deal with arithmetic operations, such as add, sub, mult, slt, and the like. I-Type instructions deal with memory, and include instructions like lw, sw, la, and others. I-Type also deals with instructions that load values immediate J-Type instructions deal with branching and jumps, which includes instructions such as j, beq, and bne. All MIPS instructions have the same number of bits, which is best illustrated by the R-Type format:
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
op - 6 bits & rs - 5 bits & rt - 5 bits & rd - 5 bits & shamt - 5 bits & funct - 6 bits \\ \hline
\end{tabular}
\end{center}
\noindent Note that the instruction type changes the meaning of the bits. For example, consider an I-Type instruction as follows:
\begin{center} 
\begin{tabular}{c}
\begin{lstlisting}
lw $t0, 400($t1)
\end{lstlisting}
\end{tabular}
\end{center}
\noindent The opcode for this instruction is $(35)_{10}$ (you would have to look this up in a table). The values for rs and rt are $9_{10}$ and $8_{10}$, respectively. However, we won't use the last 16 bits the same way we would use it for an R-Type instruction. For an I-Type instruction, the last 16 bits are used for the offset. In this case, they 32 bit representation of this instruction would look like:
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
$(35)_{10}$ - 6 bits & $(9)_{10}$ - 5 bits & $(8)_{10}$ - 5 bits & \multicolumn{3}{|c|}{$(400)_{10}$  - 16 bits} \\ \hline
\end{tabular}
\end{center}
\noindent Now consider the R-Type add instruction:
\begin{center} 
\begin{tabular}{c}
\begin{lstlisting}
add $t0, $s2, $t0
\end{lstlisting}
\end{tabular}
\end{center}
\noindent Which would result in the following bits: 
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
$(0)_{10}$ - 6 bits & $(18)_{10}$ - 5 bits & $(8)_{10}$ - 5 bits & $(8)_{10}$ - 5 bits & $(0)_{10}$ - 5 bits & $(32)_{10}$ - 6 bits \\ \hline
\end{tabular}
\end{center}
\noindent As you can see, an R-Type instruction uses each section, while the I-Type does not. Let's look at a J-Type instruction, which will only have two sections
\begin{center} 
\begin{tabular}{c}
\begin{lstlisting}
j branchname
\end{lstlisting}
\end{tabular}
\end{center}
\noindent This instruction results in the following table:
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
$(2)_{10}$ - 6 bits &  \multicolumn{5}{|c|}{$(24)_{10}$ - 26 bits}\\ \hline
\end{tabular}
\end{center}
\noindent It's worth mentioning that there are also FR-Type and FI-Type instructions. The difference between FR/FI-Type and R/I-Type is that FR/FI-Type deal with floating point numbers. The instruction format is the same, but rather than rs, rt, rd, and shamt, we have fmt, ft, fs, and fd. \\
\hrule
\noindent \\

\indent \textit{\textbf{Question 3:}} Given the following MIPS code, create a trace table:
\begin{center} 
\begin{tabular}{c}
\begin{lstlisting}
0x400:a addi 	$t0, $zero, 3
0x404:b add 	$v0, $t0, $t0 
0x408:c addi 	$t0, $t0, -1
0x40c:d bne 	$t0, $zero, b
0x410:e j 	e
\end{lstlisting}
\end{tabular}
\end{center}
\noindent \textit{\textbf{Solution:}} For each instruction, we will analyze the values of the program counter and the \$t0 and \$v0 registers. Let's take it instruction by instruction:
\begin{enumerate}[a)]
\item 
\begin{lstlisting}
0x400:a addi $t0, $zero, 3
\end{lstlisting}
\noindent In this case, there is nothing in \$v0, so we just forget about it for now. The program counter is the element that holds the current address in memory, so in this step it contains 0x400. Finally, we are adding $3 + 0$ and storing the value in \$t0, which makes \$t0 = 3.
\item
\begin{lstlisting}
0x404:b add $v0, $t0, $t0 
\end{lstlisting}
\noindent Now, we are adding \$t0 to itself and storing the value in \$v0. We know from \textit{a)} that \$t0 contains 3, so \$v0 = 6. The program counter has increased by 4.
\item 
\begin{lstlisting}
0x408:c addi $t0, $t0, -1
\end{lstlisting}
\noindent We are essentially decrementing \$t0 here, so \$t0 will contain 2. \$v0 is unchanged, and the program counter is incremented by 4.
\item 
\begin{lstlisting}
0x40c:d bne $t0, $zero, b
\end{lstlisting}
\noindent Here is where things get interesting. Recall that a ben instruction says that if (in this case) \$t0 is not equal to \$zero, then goto b. Since \$t0 is not 0, we jump back up to b. The program counter in this step is incremented by 4, and \$v0 and \$t0 do not change.
\end{enumerate}
\noindent Now we can continue to fill out the table. We know that it will continue to branch until \$t0 holds the value of \$zero. The table is as follows:
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|}
\hline
\multicolumn{12}{| l |}{Trace Table}\\ \hline
Step & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11\\ \hline
pc & 0x400 & 0x404 & 0x408 & 0x40c & 0x404 & 0x408 & 0x40c & 0x404 & 0x408 & 0x40c & 0x410 \\ \hline
\$t0 & 3 & 3 & 2 & 2 & 2 & 1 & 1 & 1 & 0 & 0 & 0\\ \hline
\$v0 & - & 6 & 6 & 6 & 4 & 4 & 4 & 2 & 2 & 2 & 2\\ \hline
\end{tabular}
\end{center}
\hrule
\newpage

\section*{The ALU in Hardware}
\hrule
\noindent\\
\indent Now we will talk about the ALU in terms of hardware. For the next few problems, we will consider the following diagram, which is one ALU unit. In a 32-bit system, there would be (you guessed it) 32 of these units.\\
\begin{center}
\includegraphics[scale=0.35]{alu2}
\end{center}
\noindent We actually can observe quite a lot about the ALU pictured above. For example, lets imagine that we have $a = 1$, and $A\ invert = 1$. Then we know that the value from the \textit{MUX} that handles the $a$ input will output 0, since the select line choses the inverted $a$. The same can be said for the \textit{MUX} that handles the $b$ signal. There is also another \textit{MUX} that handles the resulting output of the ALU.\\
\indent You can imagine that in a 32-bit system, this arrangement would not be very efficient. We won't go into too much detail, but the strategy to implement the ALU in a 32-bit (or even 64-bit) is \textbf{not} to align each of the ALU's in series. Rather, the hardware uses a \textit{carry lookahead} to help make the system more efficient Let's talk about how different instructions are implemented in the ALU.\\
\hrule
\noindent \\

\indent \textit{\textbf{Question 1:}} Describe how the following instruction could be implemented in the ALU:
\begin{center}
\begin{tabular}{c}
\begin{lstlisting}
slt $t0, $t1, $t2
\end{lstlisting}
\end{tabular}
\end{center}
\noindent \textit{\textbf{Solution:}} Recall what the $slt$ instruction does. If \$t1 $<$ \$t2, then \$t0 receives the value of 1; otherwise, \$t0 gets 0. So the question we are really asking is ``How can the ALU determine if a value is less than another value?'' We can tell when two values are equal by subtracting them and seeing if the value is zero, so that takes care of the case where the values are equal. Now we need a way to determine if \$t1 $>$ \$t2. To do this, we can subtract \$t2 from \$t1, and if the value is greater than $0$, then we know that \$t1 $>$ \$t2. If neither of these are true, then we set all the bits in \$t0 to 0. The hardware is implemented using adders. For the hardware to take the 2's compliment, it must invert the and then add 1 to the result. For example, we can try the following:
\begin{enumerate}[a)]
\item Let $a = 0011110$ and $b = 0101110$. Determine the value of $f$ given the following:
\begin{center}
\begin{tabular}{c}
\begin{lstlisting}
slt f, a, b
\end{lstlisting}
\end{tabular}
\end{center}
\end{enumerate}
\noindent Remember that in order to subtract binary numbers, we need to find the 2's compliment of the second number and add it into the first. Alternatively, we can convert the numbers into base 10 and then back into base 2. We will do it the second way here. Changing both into decimal give us:
\begin{alignat*}{3}
(0011110)_{2} &= (0 + 0 + 2^{4} + 2^{3} + 2^{2} + 2 + 0)_{10} \qquad\qquad (0101110)_{2} &&= (0 + 2^{5} + 0 + 2^{3} + 2^{2} + 2 + 0)_{10}\\
&= (30)_{10} &&= (46)_{10}
\end{alignat*}
\noindent It's very easy to see here that $a < b$, so we know that $f = 1$. Now let's try it with the 2's compliment, since the hardware cannot change binary into decimal. We only need to change $a$ into the 2's compliment.
\[a:\ 0011110 \to 1100001\]
\begin{center}
\begin{tabular}{c r}
& 1100001 \\
+ & 1 \\ \hline
& 1100010
\end{tabular}
\end{center}
\noindent Now we add $b$ to the 2's compliment of $a$:
\begin{center}
\begin{tabular}{c r}
& 1100010 \\
+ & 0101110 \\ \hline
& 10010000
\end{tabular}
\end{center}
\noindent Note that we have a signed int here, so we can just ignore the signed bit. This gives us what we expected: that is $a < b$, or that $b - a > 0$. Thus, f will be set to 00000001.\\
\hrule
\noindent\\

\indent \textit{\textbf{Question 2:}} Describe how the following instruction could be implemented in the ALU:
\begin{center}
\begin{tabular}{c}
\begin{lstlisting}
beq $t0, $t1, label 
\end{lstlisting}
\end{tabular}
\end{center}
\noindent \textit{\textbf{Solution:}} This is a very simple problem. We already looked at the $slt$ instruction, and we stated that we can tell if $\$t0 == \$t1$ if their difference is zero, and the program counter will change to the address of \textit{label}.
\end{document}

### 2015-10-04 pde/pdetest1.tex
initial

\documentclass{article}
\usepackage{amssymb,amsmath}
\usepackage{mathtools}
\usepackage{graphicx}
\topmargin=-0.3in
\textheight=9.2in
\textwidth=168mm
\oddsidemargin=-0.2in
\evensidemargin=-0.2in
\begin{document}
\section{Second Order Partial Differential Equations}
\hrule
\noindent\\
\indent We will begin these notes in a reverse order; that is, we will start by solving a second order PDE, which is probably the hardest problem we will be presented with. Let's go ahead and look at the heat equation, which is a nice one to solve.\\
\noindent \textbf{\textit{Ex:}} Solve the following PDE
\[
\begin{cases}
\frac{\partial u}{\partial t} = k \frac{\partial^{2} u}{\partial x^{2}}\\
u(x,0) = x\\
u(0,t) = u(L,t) = 0\\
\end{cases}
\]
\indent \textbf{\textit{Solution:}} We know that out solution will be in the form of $ u(x,t) = X(x)T(t) $. If we take this fact at face value, we can use the separation of variables method to solve this. First, we substitute in our solution into the PDE:
\begin{align*}
\frac{\partial}{\partial t}(X(x)T(t)) &= k \frac{\partial^{2}}{\partial x^{2}}(X(x)T(t))\\
X(x)\frac{dT}{dt} &= k\frac{d^{2}X}{dx^{2}}T(t)\\
\frac{dT}{dt}\frac{1}{T(t)} &= \frac{d^{2}X}{dx^{2}}\frac{k}{X(x)}
\end{align*}
\noindent The only way that our function of x and our function of t can be equal is if they are equal to a constant;
thus we now have:
\[
\frac{dT}{dt}\frac{1}{T(t)} = \frac{d^{2}X}{dx^{2}}\frac{k}{X(x)} = -\lambda
\]
\noindent And now we have a first order ODE for the time and a second order ODE for the position:
\[
\frac{dT}{dt} = -k \lambda T(t) \qquad\text{and}\qquad \frac{d^{2}X}{dx^{2}} = -\lambda X(x)
\]
\noindent Our next step is to find the Eigenvalues and Eigenfunctions of the second order ODE; we will come back to the time ODE when we have values for $\lambda$.\\ 
\indent To solve for these values, we have to find non-trivial eigenvalues that correspond to the boundary values of the ODE. In this case, we know that $X(0) = 0$ and $X(L) = 0$. These boundary conditions come from the BC on the PDE which states $u(0,t) = u(L,t) = 0$. To find the eigenvalues/functions, we need to find the characteristic equation corresponding to the ODE in question. Let's go ahead and find it:
\begin{gather*}
\frac{d^{2}X}{dx^{2}} = -\lambda \frac{X(x)}{k}\\
\frac{d^{2}X}{dx^{2}}  + \lambda \frac{X(x)}{k} = 0\\
r^{2} + \lambda = 0\\
r^{2} = -\lambda\\
r = \pm \sqrt{- \lambda}
\end{gather*}
\indent Now, we can substitute values of $\lambda$ to find out solutions. Let's start with $\lambda > 0$. Since we have a $-\lambda$ under the radical, we will have $r = \pm \sqrt{\lambda} \mathrm{i}$. This will yield the solution of 
\[
y(x) = c_{1}\cos{\sqrt{\lambda}x} + c_{2}\sin{\sqrt{\lambda}x}
\]
\noindent Applying our boundary conditions gives us:
\[
y(0) = c_{1}\cos{0} + c_{2}\sin{0} = 0
\]
\noindent Now we know that $c_{1} = 0$. This result came from the fact that $\sin{0} = 0$, so the $c_{2}$ term disappears. Now we are left with $\cos{0} = 1$, which tells us that $c_{1} = 0$. We can apply this to our second boundary condition like so:
\begin{gather*}
y(L) = 0 = c_{2}\sin{\sqrt{\lambda}L}\\
%\sqrt{\lambda}L = \sin^{-1}{0}
\end{gather*}
\noindent It looks like there will only be trivial solutions, but we know this isn't true due to the Jacobian matrix. Recall that if the determinant of the Jacobian $\neq 0$, then we only have trivial solutions, and it the determinant of the Jacobian $= 0$, then we do not have the trivial solution. Let's take a look at the Jacobian for this case: 
\begin{gather*}
\left| 
\begin{array}{cc}
1 & 0\\
\cos{\sqrt{\lambda}L} & \sin{\sqrt{\lambda}L}
\end{array}
\right| = 0\\
\sin{\sqrt{\lambda}L} = 0\\
\sqrt{\lambda}L = \arcsin{0}\\
\sqrt{\lambda} = \frac{\arcsin{0}}{L}\\
\lambda = \left(\frac{\arcsin{0}}{L}\right)^{2}\\
\lambda = \left(\frac{n\pi}{L}\right)^{2}
\end{gather*}
\noindent Notice that we replaced the $\arcsin{0}$ term with $n\pi$. Recall from trig that the $\sin$ function is $0$ at every $n\pi$, starting with $n = 0$. Next, we just need to find our corresponding eigenfunction. Our $c_{1}$ was $0$, so we can ignore the entire $\cos$ term. Thus, we have an eigenfunction from plugging in our $\lambda$ that looks like:
\[
X(x) = \sin{\frac{n\pi x}{L}}
\]
\indent Unfortunately, we aren't done yet. Now we have to find the corresponding eigenvalues/functions for $\lambda = 0$. This isn't too hard, and can be done in a few lines if we recall the characteristic equation. Plugging in $0$ for $\lambda$ gives us that $r = 0$. Using this piece of information, we now know that $y(x) = c_{1} + xc_{2}$. Applying our boundary conditions gives us $c_{1} = 0, c_{2} = 0$. I'll leave you to do the math (try pluggin into the Jacobian; you'll see that there is no way to force the determinant to be 0) , but we can see that we have only trivial solutions for $\lambda = 0$.\\
\indent Our final case is for $\lambda < 0$. I'll go ahead and tell you that this will yield another trivial solution, but we can take a look at the work behind it. Since $\lambda > 0$, we know that our $r$ from the characteristic equation will be $r = \pm \sqrt{\lambda}$. From here we are going to do something a little unconventional that will help us. Usually, we would have a solution of the form $y(x) = c_{1}e^{\sqrt{\lambda}x} + c_{2}x e^{-\sqrt{\lambda}x}$. In our case, it will be more convenient to use hyperbolic trig functions. It's not too far off from our solution from $\lambda > 0$:
\begin{align*}
y(x) &= c_{1}e^{\sqrt{\lambda}x} + c_{2}e^{-\sqrt{\lambda}x}\\
&= c_{1}\cosh{\sqrt{\lambda}x} + c_{2}\sinh{\sqrt{\lambda}x}
\end{align*}
\noindent I'll leave you to apply the boundary conditions (remember that $\sinh{0} = 0$ and $\cosh{0} = 1$), but you will find that you have only trivial solutions.\\
\indent So, we are done with the second order ODE, and now we turn our eye to the first order time equation. We know our $\lambda$, so it isn't too hard to solve:
\begin{gather*}
\frac{dT}{dt} = -\lambda T(t)\\
T(t) = ce^{-k\left(\frac{n\pi}{L}\right)^{2}t}
\end{gather*}
\indent Finally we can put it all together. Our penultimate solution will be 
\[
u_{n}(x,t) = B_{n}\sin{\left(\frac{n\pi x}{L}\right)}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}
\]
\noindent We can analyze this solution and tell where each piece comes from. The $B_{n}$ is a coefficient that results from our $c's$ in the general solutions of the differential equations. The $sin$ part results from out eigen vector. Remember, we get a different value for each $n$, so there may or may not be $0$ terms in the series. Finally, we have the exponential term. This term comes from t he solution to our time equation. When we put it all together, we can see that this is, of course a series; for every $n$, we get a different solution. From here we can find the Fourier Series (which will be a Fourier Sine Series). We won't worry about that yet, but we will come back to it.

\newpage
\section{First Order Partial Differential Equations}
\hrule
\noindent\\
\noindent\textbf{\textit{Ex:}} Solve the following PDE
\[
\begin{cases*}
yu_{x} + xu_{y} = 0\\
u(0,y) = e^{-y^{2}}
\end{cases*}
\]
\indent \textbf{\textit{Solution:}} This is a nice PDE; there are two ways to solve it, but we will focus on the easier one. We can see that we have a coefficient on both $u_{x}$ and $u_{y}$. This tells us that when $u$ was differentiated, there was a constant $y$ and $x$. So now we can say that
\[
\frac{dy}{dx} = \frac{x}{y}
\]
\noindent This is a simple ODE to solve. We can use separation of variables:
\begin{align*}
xdx &= ydy\\
\int{x dx} &= \int{y dy}\\
\frac{1}{2}x^{2} &= \frac{1}{2}y^{2} + c\\
x^{2} - y^{2} &= c_{0}\\
y^{2} - x^{2} &= c_{1}
\end{align*}
\noindent You'll notice that we multiplied both sides by $-1$. This is for convenience. Now we can express the answer as
\[
u(x,y) = f(y^{2} - x^{2})
\]
\noindent And that will suffice.

\newpage
\noindent\textbf{\textit{Ex:}} Solve the following PDE
\[
u_{x} + u_{y} + u = x + y
\]
\indent \textbf{\textit{Solution:}} In this case, we have a non-zero term on the right hand side of the equation. We will use substitution to solve this. For the time being, we will let $T$ be $x + y$. To find our $S$, we can use a similar method to the one that we used above. We know that the coefficients on the $u_{x}$ and $u_{y}$ are $1$ in this case. Now we have:
\begin{align*}
\frac{dy}{dx} &= 1\\
dy &= dx\\
\int dy &= \int dx\\
x &= y + c\\
x - y &= c
\end{align*}
\noindent Now we can say that $S = x - y$. Next we have to actually use the substitution. We need to express $u_{x}$ and $u_{y}$ in terms of $S$ and $T$. The following equations come from the chain rule; when the $S$ and $T$ terms are differentiated, we get $u_{S}$ and $u_{T}$ left behind.
\begin{alignat*}{3}
u_{x} &= u_{S}\frac{\partial S}{\partial x} + u_{T}\frac{\partial T}{\partial x} 
\qquad\qquad &&u_{y} = u_{S}\frac{\partial S}{\partial y} + u_{T}\frac{\partial T}{\partial y}\\
u_{x} &= -u_{S} + u_{T}  \qquad\qquad &&u_{y} = u_{S} + u_{T}
\end{alignat*}
\noindent Now we have our substitution for $u_{x}$ and $u_{y}$. We can now plug in our values into the original equation. Note that we didn't do anything with $u$; it gets left alone. In addition, if we needed $u_{xx}$ or $u_{yy}$, we can just differentiate again. We will see this later when we talk about classifications. Now, back to plugging in:
\begin{gather*}
(-u_{S} + u_{T}) + (u_{S} + u_{t}) + u = T\\
2u_{T} + u = T
\end{gather*}
\noindent This is just a first order ODE. Since it's not homogeneous, we need to solve for the general solution, and then the particular solution. Let's start with the general.
\begin{gather*}
2y^{'} + y = 0\\
2r + 1 = 0\\
r = -\frac{1}{2}\\
y = ce^{\left(-\frac{1}{2}\right)x}
\end{gather*}
\noindent In this case, we know that our $x$ is really $T$. So we finally have $y = ce^{\left(-\frac{1}{2}\right)T}$. We will back substitute soon, but let's find the particular solution before we do that.
\noindent Recall from ODE that we have several cases for the particular solution. In our case, we have a linear equation on the right hand side of our ODE. This lets us say:
\begin{gather*}
y = AT + B\\
y^{'} = A\\
\end{gather*}
\noindent Now we can substitute again into our non-homogeneous ODE as follows:
\[
2y^{'} + y = T \equiv 2A + AT + B = T
\]
\noindent And we have that $A = 1$, and $B = -2$. Now we have a final solution of:
\begin{gather*}
y = ce^{\left(-\frac{1}{2}\right)T} + T - 2\\
y = ce^{\left(-\frac{1}{2}\right)(x+y)} + (x + y) - 2
\end{gather*}
\noindent We aren't quite done yet, as we have a solution to an ODE, not a PDE. It isn't too difficult to switch back; The $c$ is really the only difference. In our ODE, the $c$ is a constant term, but to our PDE, the $c$ is a function. The $c$ now becomes $f(s)$. We will go ahead and back substitute, which will give us $f(y-x)$. Now we have our final solution:
\[
u(x,y) = f(y-x)e^{\left(-\frac{1}{2}\right)(x+y)} + (x + y) - 2
\]
\noindent And we are done.

\newpage
\section{Classification of Partial Differential Equations}
\hrule
\noindent\\
\indent We mentioned earlier the topic of classification of PDE's. Since we only have this and Fourier Series to talk about, we will go ahead and knock out classification. There's some information that we need to go ahead and put out here. This will be explained in the problem, but it's handy to have it for reference: 
\begin{align*}
b^{2} - ac < 0 &\implies Elliptic\ with\ complex\ roots\\
b^{2} - ac = 0 &\implies Parabolic\ with\ the\ same\ root\ twice\\
b^{2} - ac > 0 &\implies Hyperbolic\ with\ two\ distinct\ real\ roots
\end{align*}
\noindent Another helpful piece of information is what's known as the \textit{classical trinity} of PDEs.
\begin{align*}
u_{t} - ku_{xx} &= 0 \qquad The\ heat\ equation\ (parabolic)\\
u_{tt} - c^{2}u_{xx} &= 0 \qquad The\ wave\ equation\ (hyperbolic)\\
u_{xx} + u_{yy} &= 0 \qquad The\ Laplace\ equation\ (elliptic)
\end{align*}
Let's go ahead and look at an example.\\
\indent \textbf{Ex. }Classify the following PDE:
\[yu_{xx} + u_{yy} = 0\]
\indent \textbf{\textit{Solution:}} The first thing we need to identify are the coefficients on the $u_{xx}$, $u_{yy}$, and $u_{xy}$ or $u_{yx}$ terms. Note that the $u_{xy}$ and $u_{yx}$ are interchangeable. We will assign these coefficients to $a$, $b$, and $c$ corresponding to $u_{xx}$, $u_{xy}$, and $u_{yy}$, respectively. Looking at the PDE, we can see that
\[
a = y \qquad b = 0 \qquad c = 1
\]
\noindent Now we look at $b^{2} - ac$. It should be noted that we are really dealing with $b = \frac{b_{0}}{2}$. In other words, we need to divide our $b$ value by $2$. In the case of the first problem, we are dealing with $b = 0$, so it's not evident. So, now we can go ahead and classify our PDE as follows:
\[
b^{2} - ac = 0 - y < 0 \implies Elliptic
\]
Note that we assume $y > 0$. So our PDE is elliptic. Now we need to make it look like it's corresponding equation from the classical trinity. In this case we have an Elliptic equation, so we want to make it look like Laplace's equation. We need to find values for $u_{xx}$ and $u_{yy}$, and then substitute those into the equation. This is similar to how we solved the first order PDE's; let's look at our characteristic equation:
\begin{gather*}
am^{2} + bm + c = 0\\
ym^{2} + 1 = 0\\
m = \pm \frac{i}{\sqrt{y}}
\end{gather*}
\noindent So, now we have two solutions, and they are both imaginary as expected. Next we need to find values for $S$ and $T$ that we can substitute into the equation. We will do both of them at the same time here. Notice that we will drop the $i$; we don't really care about the imaginary part:
\begin{gather*}
\frac{dy}{dx} = \frac{1}{\sqrt{y}} \qquad\qquad\qquad \frac{dy}{dx} = -\frac{1}{\sqrt{y}}\\
\int\frac{dy}{dx} =\int\frac{1}{\sqrt{y}} \qquad\qquad\qquad \int\frac{dy}{dx} = -\int\frac{1}{\sqrt{y}}\\
S = \frac{2}{3}y^{\left(\frac{3}{2}\right)} - x \qquad\qquad\qquad T = \frac{2}{3}y^{\left(\frac{3}{2}\right)} + x
\end{gather*}
\noindent Now we can substitute into our equation. Unfortunately, we will need to take the partial derivative twice in order to get the substitutions we need. Let's start with $u_{xx}$.
\begin{align*}
u_{x} &= u_{S}\frac{\partial S}{\partial x} + u_{T}\frac{\partial T}{\partial x}\\
&= -u_{S} + u_{T}\\
u_{xx} &= \frac{\partial}{\partial x}(-u_{S} + u_{T})\\
&= u_{SS} + u_{TT}
\end{align*}
\noindent That wasn't too bad. We will skip the math for finding $u_{yy}$, which would give us
\[
u_{yy} = yu_{SS} + yu_{TT}
\]
\noindent Now we plug everything in. Remember that we were under the assumption that $y > 0$:
\begin{gather*}
y(u_{SS} + u_{TT}) + yu_{SS} + yu_{TT} = 0\\
2yu_{SS} + 2yu_{TT} = 0\\
u_{SS} + u_{TT} = 0
\end{gather*}
\noindent And we are finished.

\newpage
\section{Fourier Series and Partial Differential Equations}
\hrule
\noindent\\
\indent It's finally time to talk about Fourier Series. Before we start though, it may be helpful to look at some common patterns amongst the Fourier Series. We have the following information that will help us:\\\\
\centerline{
\begin{tabular}{|c|c|c|c|}
\hline
& Cosine & Sine & Full Series \\ \hline
Interval & $[0,L]$ & $[0,L]$ & $[-L,L]$\\ \hline
Basis & $\{1,\cos{\left(\frac{n\pi x}{L}\right)}\}$ &
$\{\sin{\left(\frac{n\pi x}{L}\right)}\}$ &
$\{1,\sin{\left(\frac{n\pi x}{L}\right)},\cos{\left(\frac{n\pi x}{L}\right)}\}$\\ \hline
Series & 
$\frac{b_{0}}{2} + \sum_{n=1}^{\infty}[b_{n}\cos{\left(\frac{n\pi x}{L}\right)}]$& 
$\sum_{n=1}^{\infty}[a_{n}\sin{\left(\frac{n\pi x}{L}\right)}]$&
$\frac{b_{0}}{2} + \sum_{n=1}^{\infty} [a_{n}\sin{\left(\frac{n\pi x}{L}\right)} + b_{n}\cos{\left(\frac{n\pi x}{L}\right)}]$\\ \hline
\end{tabular}
}\\\\
\noindent To solve for the coefficients, we have:
\begin{align*}
Sine:  \qquad\qquad &a_{n} = \frac{2}{L}\int_{0}^{L}f(x)\sin{\left(\frac{n\pi x}{L}\right)}dx\\
Cosine: \qquad\qquad &b_{n} = \frac{2}{L}\int_{0}^{L}f(x)\cos{\left(\frac{n\pi x}{L}\right)}dx\\
&b_{0} =  \frac{1}{L}\int_{0}^{L}f(x)dx\\
Full: \qquad\qquad &a_{n} = \frac{2}{L}\int_{-L}^{L}f(x)\sin{\left(\frac{n\pi x}{L}\right)}dx\\
&b_{n} = \frac{2}{L}\int_{-L}^{L}f(x)\cos{\left(\frac{n\pi x}{L}\right)}dx\\
&b_{0} =  \frac{1}{L}\int_{-L}^{L}f(x)dx
\end{align*}
The first thing we will look at is the Full Fourier Series. Note that this series is just a combination of a Fourier Sine Series and a Fourier Cosine Series. A typical problem would look something like this:\\\\
\textbf{\textit{Ex:}} Find the corresponding Fourier Series for
\begin{gather*}
f(x)=
\begin{cases*}
L \qquad -L \leq x \leq 0\\
2x \qquad 0 \leq x \leq L
\end{cases*}
\end{gather*}
\indent \textbf{\textit{Solution:}} So we have a piecewise function here, but that shouldn't throw us off too much; all we need to do is combine the integrals for each part of the coefficient solution. Let's start by finding the terms for the $cos$ series.
\begin{align*}
B_{0} &= \frac{1}{2L}\left[\int_{-L}^{0}Ldx + \int_{0}^{L}2xdx\right]\\
&= \frac{1}{2L}[2L^{2}]\\
&= L
\end{align*}
\noindent That wasn't too bad, was it? Now We need to find $B_{n}$, and then we will have our cosine part of the series. The integration can get pretty messy, so we will skip some steps.
\begin{align*}
B_{n} &= \frac{1}{L}\left[\int_{-L}^{0}L\cos{\left(\frac{n\pi x}{L}\right)}dx + \int_{0}^{L}2x\cos{\left(\frac{n\pi x}{L}\right)}dx\right]\\
&= \frac{1}{L}\left[0 + \left(\frac{2L^{2}}{n^{2}\pi^{2}}\right)((-1)^{n} - 1)\right]\\
&= \frac{2L}{(n\pi)^{2}}((-1)^{n} - 1)
\end{align*}
\noindent Note that the sequence will be 0 for all $n = 2k$. This allows us to substitute $2k+1$ for $n$, since we don't care about the $0$ terms. For our cosine part of the series, we now have:
\begin{align*}
B_{n} &= \frac{2L}{(2k+1)^{2}\pi^{2}}((-1)^{2k+1}-1)\\
&= \frac{2L}{(2k+1)^{2}\pi^{2}}(-2)\\
&= \frac{-4L}{(2k+1)^{2}\pi^{2}}
\end{align*}
\noindent So, now we need to find the $\sin$ part of the series. This is easier, since we only need to find $A_{n}$.
\begin{align*}
A_{n} &= \frac{1}{L}\left[\int_{-L}^{0}L\sin{\left(\frac{n\pi x}{L}\right)}dx + \int_{0}^{L}2x\sin{\left(\frac{n\pi x}{L}\right)}dx \right]\\
&= \frac{L}{n\pi}\left[-1-(-1)^{n}\right]
\end{align*}
\noindent Note that we can do a similar substitution here as we did with the cosine series. This time, however, we will have $0$ for $n = 2k+1$ We now have:
\begin{align*}
A_{n} &= \frac{L}{2k\pi}\left[-1-(-1)^{2k}\right]\\
&= \frac{L}{2k\pi}\left[-1 - 1\right]\\
&= \frac{L}{2k\pi}\left[-2\right]\\
&= \frac{-2L}{2k\pi}
\end{align*}
\noindent Now we have all the parts that we need. Our final answer will be essentially the sum of all the parts:
\begin{align*}
f(x)&\sim \frac{L}{2} + \sum_{k = 0}^{\infty}\left[\frac{-4L}{(2k+1)^{2}\pi^{2}}\cos{\left(\frac{(2k + 1)\pi x}{L}\right)} \right] + \sum_{k = 0}^{\infty}\left[\frac{-2L}{2k\pi}\sin{\left(\frac{2k\pi x}{L}\right)} \right]\\
&\sim \frac{L}{2} + \sum_{k = 0}^{\infty}\left[\frac{-4L}{(2k+1)^{2}\pi^{2}}\cos{\left(\frac{(2k + 1)\pi x}{L}\right)} + \frac{-2L}{2k\pi}\sin{\left(\frac{2k\pi x}{L}\right)} \right]
\end{align*}
\noindent And we are done.\\

\newpage
\indent Now lets take a look at a slightly more theoretical Fourier Series.\\
\noindent\textbf{\textit{Ex:}} Find the corresponding Fourier Series for\\
\[f(x)=\left|\cos{x}\right|\qquad -\pi\leq x \leq \pi\]
\indent\textbf{\textit{Solution:}} Before we dive into calculations here, let's think about the function at hand. We have the absolute value of $\cos{x}$. Usually, $\cos$ would be an odd function, which would imply that our $\cos$ series component to the whole Fourier Series would be $0$. However, the absolute value makes our function even, and as such the $\sin$ component of the series will be $0$. That's good news for us, because now we don't have to find $a_{n}$; we actually only have a Fourier Cosine Series here. Recall that we have a formula for $b_{n}$ and $b_{0}$, and that our general solution will have the form of
\[\frac{b_{0}}{2} + \sum_{n=1}^{\infty}[b_{n}\cos{\left(\frac{n\pi x}{L}\right)}]\]
\noindent Our function is bounded above by $\pi$, so we really have $L= \pi$. This gives us:
\[\frac{b_{0}}{2} + \sum_{n=1}^{\infty}[b_{n}\cos{\left(nx\right)}]\]
\noindent Let's go ahead and find our $b_{0}$:
\begin{align*}
b_{0} &= \frac{1}{\pi}\int_{-\pi}^{\pi}\left|\cos{x}\right|dx\\
&= \frac{2}{\pi}\int_{0}^{\pi}\left|\cos{x}\right|dx\\
&= \frac{4}{\pi}\int_{0}^{\frac{\pi}{2}}\left|\cos{x}\right|dx\\
&= \frac{4}{\pi}\left[\sin{\frac{\pi}{2}} - \sin{0}\right]\\
&= \frac{4}{\pi}
\end{align*}
\noindent Let's take a second and discuss some the moves we made here. First of all, we know that since the function is even, we can say that the the integral of the entire range is really twice the integral of half the range. Since we are dealing with the cosine function, we actually have four times a quarter of the range. If we take a look at a graph, this may make a little more sense:\\
\begin{center}
\begin{tabular}{c c}
\includegraphics[scale=0.5]{cos_abs} & \includegraphics[scale=0.5]{cos}\\
$f(x) = |\cos{x}|$ on $[0,2\pi]$ & $f(x) = \cos{x}$ on $[0,2\pi]$
\end{tabular}
\end{center}
\noindent\\ You can see that the section of the graph from $[\frac{\pi}{2},\frac{3 \pi}{2}]$ is positive for the $|\cos{x}|$, but negative for $\cos{x}$. If we simply integrated the $|\cos{x}|$ over $[0,\pi]$, the result would be 0, but graphically this result would not make any sense. Now let's take a look at the $b_{n}$ term. Before we jump into the integration, let's think about this function:
\[\left|\cos{x}\right| = 
\begin{cases*}
\cos(x) \qquad if\ 0\leq x \leq \frac{\pi}{2}\\
-\cos(x)\quad if\ \frac{\pi}{2}\ \leq x \leq \pi
\end{cases*}
\]
\noindent Now we can set up the formula for the $b_{n}$ term. We have as follows:
\begin{align*}
b_{n} &= \frac{2}{\pi}\left[ \int_{0}^{\frac{\pi}{2}} \cos{x}\cos{nx}dx - \int_{\frac{\pi}{2}}^{\pi} \cos{x}\cos{nx}dx \right]\\
&= \frac{2}{\pi} \left[ \frac{-2 \cos{\left(\frac{n\pi}{2}\right)} + \sin{\left(n\pi\right)}}{n^{2} - 1}\right]\\
\end{align*}
\noindent Note that (as we have seen before) that the $\sin$ term disappears; however, the more interesting part is what happens with our $\cos$ term. The value of $\cos{\left(\frac{n\pi}{2}\right)}$ will be 0 at all odd $n$, or that is to say for all $n = 2k + 1$. So, we will let our $n = 2k$. This gives us:
\begin{align*}
b_{n} &= \frac{2}{\pi} \left[ \frac{-2 \cos{\left(k \pi \right)}}{(2k)^{2} - 1}\right]\\
&= -\frac{4}{\pi}\left[\frac{(-1)^{k}}{(2k)^{2}-1}\right]
\end{align*}
\noindent And that leaves us with the final result
\[
f(x)\sim \frac{2}{\pi} - \sum_{k=0}^{\infty}\left[\frac{4}{\pi}\left[\frac{(-1)^{k}}{(2k)^{2}-1}\right] \cos{\left(\frac{2k\pi}{2}\right)}\right]
\]
\newpage

\indent Let's recall the 2nd order PDE that we found a series solution for earlier. Now we want to apply the initial condition of $u(x,0) = x$.\\
\noindent\textbf{\textit{Ex:}} Find the particular solution for the following PDE:
\[
\begin{cases}
\frac{\partial u}{\partial t} = k \frac{\partial^{2} u}{\partial x^{2}}\\
u(x,0) = x\\
u(0,t) = u(L,t) = 0\\
\end{cases}
\]
\indent\textbf{\textit{Solution:}} Earlier, we found a general solution of the form
\begin{align*}
u_{n}(x,t) &= a_{n}\sin{\left(\frac{n\pi x}{L}\right)}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}\\
u(x,t) &= \sum_{n = 1}^{\infty}\left[a_{n}\sin{\left(\frac{n\pi x}{L}\right)}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}\right]
\end{align*}
\noindent Without loss of generality, let's let $L = \pi$. This would change our boundary condition to $u(\pi,t) = 0$, but this wouldn't change the solution that we have here in any way other than replacing the $L$'s in the equation with $\pi$. We will also let $k=1$. Again this doesn't change anything other than the fact that we will replace $k$ with 1. We now have:
\[
u(x,t) = \sum_{n = 1}^{\infty}\left[a_{n}\sin{\left(nx\right)}e^{-\left(n\right)^{2}t}\right]
\]
\noindent This is a much nicer sequence for us to deal with. Our goal now is to find $a_{n}$, which will give us a particular solution. We know that our answer for the particular solution will be in the form of a series, since our initial condition is not represented in the form of terms of a series (we will see an example of this later). Let's go ahead and apply our initial condition. Doing so will yield:
\[
x = \sum_{n = 0}^{\infty}\left[a_{n}\sin{nx}\right]
\]
\noindent Now, we need to come up with an $a_{n}$. We have a Fourier Sine Series here, so we can use the formula for finding $a_{n}$ that corresponds to our case.
\begin{align*}
a_{n} &= \frac{2}{\pi}\int_{0}^{\pi} x\sin{\left(nx\right)}dx\\
&= -\frac{2n\pi \cos{\left(n\pi\right)} + 2\sin{\left(n\pi\right)}}{n^{2}\pi}\\
&= -\frac{2(-1)^{n}}{n}
\end{align*}
\noindent So our particular solution is:
\[
u(x,t) = \sum_{n = 1}^{\infty}\left[-\frac{2(-1)^{n}}{n}\sin{\left(n x\right)}e^{-\left(n\right)^{2}t}\right]
\]
\noindent And we are done. We haven't discussed the convergence of series yet, and it turns out that if the series converges, it will converge to the solution. We won't worry about it right now. Instead, let's look at another example

\newpage
\noindent\textbf{\textit{Ex:}} Solve the following boundary value problem given the general solution:
\[
\begin{cases}
u_{tt} - u_{xx} = 0\\
u(0,t) = 0\qquad u(1,t) = 0\\
u(x,0) = f(x)\qquad u_{t}(x,0) = g(x)
\end{cases}\\
\]
\[f(x) = 2\sin{(\pi x)} + 3\sin{(2\pi x)}\qquad g(x) = 4\sin{(3\pi x)} - 7\sin{(5\pi x)}\]
\[u(x,t) = \sum_{n=1}^{\infty}\left[\left(a_{n}\cos{[n\pi t]} + b_{n}\sin{[n\pi t]}\right)\sin{[n\pi x]}\right]\]
\indent\textbf{\textit{Solution:}} There's a lot going on here, so let's break it down. First of all, notice that we have two initial conditions, one of which is a derivative. This should make sense since the equation have two second order partial derivatives included in it. Next notice that we have a Fourier Sine Series, despite both the $\sin{[n\pi t]}$ and $\cos{[n\pi t]}$ terms. If we had worked out the entire problem, we would see that our eigenvalue/function for this equation takes the form of 
\[
\{\lambda_{n} = n\pi, \sin{(n\pi x)}\}
\]
\noindent which tells us that the corresponding series is a Fourier Sine Series. We also see the initial conditions on our equation are already representative of a series of sine functions, which is not just a conincidence. To apply the initial conditions, we will need both the general solution and the derivative with respect to $t$ of the general solution. Let's go ahead and apply the first initial condition, $u(x,0) = f(x)$. Letting $t= 0$ gives us:
\[
u(x,0) = 2\sin{(\pi x)} + 3\sin{(2\pi x)} = \sum_{n=1}^{\infty}\left[a_{n}\sin{(n\pi x)}\right]
\]
\noindent We see that the term with $b_{n}$ has dropped out, since $\sin{(0)} = 0$. Now we just have to match up the coefficients on $a_{n}$. We can see that when we have $n=1$, $a_{n} = 2$, and when $n=2$, $a_{n} = 3$. For all other $n$, we have $a_{n} = 0$, so we ignore those terms. Next, we can apply the initial condition for the derivative of the general solution with respect to $t$. We expect that all the terms related to $a_{n}$ will disappear, and in fact they will. Let's take a look:
\begin{align*}
u_{t}(x,t) &= \sum_{n=1}^{\infty}\left[(-a_{n}n\pi\sin{[n\pi t]} + b_{n}n\pi\cos{[n\pi t]})\sin{n\pi x}\right]\\
u_{t}(x,0) &= 4\sin{(3\pi x)} - 7\sin{(5\pi x)} = \sum_{n=1}^{\infty}\left[n\pi b_{n}\sin{(n\pi x)}\right]
\end{align*}
\noindent We are going to do the same thing here that we did to find our $a_{n}$ terms, but notice that we don't just have $b_{n}$; we actually have $n\pi b_{n}$. So we have the following:
\begin{alignat*}{3}
&3\pi b_{3} = 4 \qquad \qquad \qquad &&5\pi b_{5} = -7\\
&b_{3} = \frac{4}{3\pi} &&b_{5} = -\frac{7}{5\pi}
\end{alignat*}
\noindent Much like the solutions for $a_{n}$, all the other $b_{n}$ terms are 0. Now we can construct a particular solution. In this case, the answer will not be in the form of a series, since we have definite terms. Our solution is as follows:
\[
u(x,t) = 2\cos{(\pi t)}\sin{(\pi x)} + 3\cos{(2\pi t)}\sin{(2\pi x)} + \frac{4}{3\pi}\sin{(3\pi t)}\sin{(3\pi x)} - \frac{7}{5\pi}\sin{(5\pi t)}\sin{(5\pi x)}
\]
\noindent And we are done.
\end{document}

### 2015-10-05 osi/os_test1.tex
added os class test

\documentclass{article}
\usepackage{amsmath}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{listings}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\begin{document}
\section{UNIX Pipes}
\hrule
\noindent\\
\indent We will start off this guide by talking about UNIX pipes. Pipes are used by the operating system in order to transfer data from from one process to another during execution. We can imagine that a pipe acts like a file, where one process can read from the file, and the other process can write to the file. Pipes tend to follow a FIFO structure. Here are some common features of a UNIX pipe:
\begin{enumerate}
\item Pipes are limited to 10 logical blocks, with each block containing 512 bytes of data.
\item Generally, one process will read from the pipe while the other will write.
\item Data is written to one end of the pipe by a process and read from the other end of the pipe by another process.
\item There is no file pointer associated with a pipe, as there is with a normal file in UNIX.
\end{enumerate}
\indent Pipes are generally created in the kernel of the OS, and generally each process will have it's own pipe that it writes to and reads from. In other words, if \textit{Process A} is reading from a pipe that \textit{Process B} is writing to, the will also be a pipe to which \textit{Process A} can write and \textit{Process B} can read. Let's take a look at the system call to write to a file:
\begin{center} 
\begin{tabular}{c}
\begin{lstlisting}
size_t write(int file_des, const void *buffer, size_t bytes)
\end{lstlisting}
\end{tabular}
\end{center}
\indent This function will write \textit{bytes} amount of \textit{buffer} to the file descriptor given by \textit{file\_des}. It returns -1 if the call fails, or the number of bytes written otherwise. Similarly we have the \textit{read} system call:
\begin{center} 
\begin{tabular}{c}
\begin{lstlisting}
size_t read(int file_des, const void *buffer, size_t  num_bytes)
\end{lstlisting}
\end{tabular}
\end{center}
\indent The \textit{read} function will number of bytes read if successful, or -1 otherwise. We should note that we actually have two types of pipes, unnamed and named. From a programatic standpoint, we see no difference between the file descriptors for the two pipes besides the call used to make them. Unnamed pipes are used between related processes, such as parent/child, whereas named pipes are created with a directory entry and read/write permissions. Unnamed and named pipes are made with the \textit{pipe} and \textit{mknod} system calls, respectively. Let's take a look at the \textit{pipe} system call.
\begin{center} 
\begin{tabular}{c}
\begin{lstlisting}
int pipe (int file_des[2])
\end{lstlisting}
\end{tabular}
\end{center}
\indent This system call takes in an array of ints that, upon success, will contain the file descriptors for reading and writing. It will also return 0 upon success, and -1 upon failure.\\
\indent You can image that pipes could be used for I/O between processes. We have two system calls that we can use to accomplish this feat:
 \begin{center} 
\begin{tabular}{c}
\begin{lstlisting}
FILE *popen (const char *command,  const char *type)
\end{lstlisting}\\
\begin{lstlisting}
int pclose (FILE *stream)
\end{lstlisting}
\end{tabular}
\end{center}

\end{document}

### 2015-10-05 pde/pdetest1.tex
update ode test

@@
 \centerline{
+\bgroup
+\def\arraystretch{1.5}
 \begin{tabular}{|c|c|c|c|}
@@
 \end{tabular}
+\egroup
 }\\\\
@@
 \end{align*}
-The first thing we will look at is the Full Fourier Series.
+\indent It will also help us to quickly review the concept of odd and even functions, as they will help us to simplify the process of computing Fourier series later.
+Specifically, let's look at the graphs for $\sin$ and $\cos$ on $[-\pi,\pi]$, as well as another function that will turn out to be neither even nor odd..
+\begin{center}
+\begin{tabular}{c c c}
+\includegraphics[scale=0.4]{cos_eo} & \includegraphics[scale=0.4]{sin_eo} & \includegraphics[scale=0.4]{cube_eo}\\
+$f(x) = \cos{x}$ & $f(x) = \sin{x}$ & $f(x) = x^{3}-x +1$
+\end{tabular}
+\end{center}
+\noindent\\
+\indent We should recall that a function is \textit{even} when $f(x) = f(-x)$ for all $x$ in the domain of $f$.
+Graphically speaking, an \textit{even} function is reflected across only the $y$-axis.
+By looking at the above graphs, we can see that $f(x) = \cos{x}$ is an even function.\\
+\indent A function is \textit{odd} when $-f(x) = f(-x)$ for all x in the domain of $f$.
+Graphically speaking, the function is reflected across the $x$-axis and the $y$-axis.
+From the above graphs, we can say that $f(x) = \sin{x}$ is an odd function.
+Note that we can have a function that is neither odd nor even, but we won't see much of this in the following problems.
+The reason that we want to discuss odd and even functions here is because the integration that we will be seeing later will be affected by the symmetry of a function.
+Let's take a look at how even and odd functions effect integration:\\\\
+\textbf{\textit{Ex:}} Compute the following integral:
+\[
+\int_{0}^{\pi}\cos{x}dx
+\]
+\indent \textbf{\textit{Solution:}} Looking at this integral, we can see from the graph of the $\cos$ above that this will be 0.
+That answer is meaningless to us however.
+What we are really looking for is
+\[
+2\int_{0}^{\frac{\pi}{2}}\cos{x}dx
+\]
+\noindent We are allowed to do this since the $\cos$ function is even. Essentially, we have turned the negative half of the cosine into a positive half. Similarly, we could ask for something like
+\[
+\int_{0}^{\pi}|\cos{x}|dx
+\]
+\noindent Since the absolute value of a function is no integrable, we drop the absolute value signs.
+This will leave us with our initial problem, which we will have to change anyways.
+Our integral becomes:
+\[
+\int_{0}^{\pi}|\cos{x}|dx = 2\int_{0}^{\frac{\pi}{2}}\cos{x}dx = 1
+\]
+\noindent And we are done.
+\newpage
+
+\indent The first type of series we will look at is the Full Fourier Series.
 Note that this series is just a combination of a Fourier Sine Series and a Fourier Cosine Series.
@@
 \noindent And we are done.
+\newpage
+
+\indent Up to this point, we have found a Fourier series and left our solutions at that. We will now look into the convergence of a Fourier series. This topic includes two questions:
+\begin{enumerate}
+\item Does the series converge?
+\item If the series converges, to what value does it converge?
+\end{enumerate}
+\noindent Both of these questions are highly theoretical. We will discuss the answers to them, but first we need to go over some terms:
+\begin{enumerate}
+\item \textit{\textbf{Jump Discontinuity:}} To say that $f(x)$ has a jump discontinuity at $x = a$ is to say that
+\[
+\lim_{x \to a^{-}}f(x) \neq \lim_{x \to a^{+}}f(x)
+\]
+Where both the limit from the left and the limit from the right exist.
+\item \textit{\textbf{Piecewise Smooth:}} A function $f(x)$ is said to be piecewise smooth if the function can be broken into two distinct intervals, and on each interval both $f(x)$ and $f^{'}(x)$ are continuous.
+\item \textit{\textbf{Periodic Extension:}} The periodic extension of $f(x)$ on $[-L,L]$ is the repetition of the function on intervals on periods to the left and right of $[-L,L]$.
+\end{enumerate}
+\indent With these definitions, we can lay out a general rule for convergence of a Fourier series.
+We won't give a proof of these rules for the time being.
+We just need them to lay out some groundwork for the rest of the information we will present on Fourier series.
+We have the following:
+\begin{quote}
+Suppose $f(x)$ is a function, and that $f(x)$ is piecewise smooth on the interval $[-L,L]$. The corresponding Fourier series for $f(x)$ converges to either
+\begin{enumerate}
+\item the periodic extension of $f(x)$, if $f(x)$ is continuous, or
+\item the average of the two one-sided limits, if the periodic extension of $f(x)$ has a jump discontinuity at $x = a$.
+\end{enumerate}
+\end{quote}
+\noindent We should note that the second condition can be expressed by the formula
+\[\frac{1}{2}\left[\lim_{x \to a^{-}}f(x) + \lim_{x \to a^{+}}f(x)\right]\]
+which gives us a nice way to calculate the value to which the series converges if we happen to meet the criteria in (2). Let's take a look at some examples.\\\\
+
+\noindent\textbf{\textit{Ex:}} Find the Fourier series for the following equation, and determine if the series converges. If it does, tell where the series converges to.
+\[ f(x) =
+\begin{cases}
+L \qquad\qquad -L \leq x \leq 0\\
+2x \qquad\qquad 0 \leq x \leq L
+\end{cases}
+\]
+\indent\textbf{\textit{Solution:}} We will go ahead and skip the details of finding the Fourier series for this function.
+If you do want to find it, you will have to find a full piecewise Fourier series.
+For this particular $f(x)$, we have:
+\begin{align*}
+f(x) &\sim L + \sum_{n=1}^{\infty}\left[\frac{2L}{n^{2}\pi^{2}}([-1]^{n}-1)\cos{\left(\frac{n\pi x}{L}\right)}\right] - \sum_{n=1}^{\infty}\left[\frac{L}{n\pi}(1 + [-1]^{n})\sin{\left(\frac{n\pi x}{L}\right)}\right]
+\end{align*}
+\noindent You'll notice that we didn't simplify this solution to a nicer form.
+If we evaluate this series, we will run across terms where the series is 0, but that is okay for now.
+You should also notice that this function is defined as a piecewise function, and on both intervals ($[-L,0]$ and $[0,L]$), $f(x)$ is continuous.
+In other words, $f(x)$ is piecewise smooth, so we know that this series converges.
+Now let's take a look at the graph of $f(x)$ on an arbitrary interval of $[-1,1]$.
+\begin{center}
+\begin{tabular}{c}
+\includegraphics[scale=0.6]{pw_graph_01}\\
+$
+f(x) =
+\begin{cases}
+1 \qquad\qquad -1 \leq x \leq 0\\
+2x \qquad\qquad 0 \leq x \leq 1
+\end{cases}
+$
+\end{tabular}
+\end{center}
+\noindent We can see that there is a jump in the graph at $x = 0$.
+In other words, there is a jump discontinuity at $x = 0$, and that means that we need to use (2) in order to find the value of convergence.
+We will let $\sum$ stand in for our entire series.
+We have:
+\begin{align*}
+\sum &\to \frac{1}{2}\left[\lim_{x\to 0}L + \lim_{x\to0}2x\right]\\
+&\to \frac{1}{2}[L + 0]\\
+&\to \frac{L}{2}
+\end{align*}
+\noindent So, at $x=0$, the series converges to $\frac{L}{2}$.
+We are not done yet, however.
+Now, we need to consider the periodic extension of the series.
+The function will repeat every $[-L,L]$, so we need to look at what the convergence of the series looks like at $-L$ and $L$.
+Let's take a look:
+\begin{alignat*}{3}
+\sum&\to\frac{1}{2}\left[\lim_{x\to-L^{-}}f(x) + \lim_{x\to-L^{+}}f(x)\right] \qquad\qquad \sum&&\to \frac{1}{2}\left[\lim_{x\to L^{-}}f(x) + \lim_{x\to L^{+}}f(x)\right]\\
+&\to \frac{1}{2}[2L + L] &&\to \frac{1}{2}[2L + L]\\
+&\to\frac{3L}{2} &&\to\frac{3L}{2}
+\end{alignat*}
+\noindent And we are done.
+\newpage
+
 \end{document}

### 2015-10-20 pde/pdetest1.tex
added a lot of PDE stuff

@@
 \begin{document}
-\section{Second Order Partial Differential Equations}
-\hrule
-\noindent\\
-\indent We will begin these notes in a reverse order; that is, we will start by solving a second order PDE, which is probably the hardest problem we will be presented with.
-Let's go ahead and look at the heat equation, which is a nice one to solve.\\
-\noindent \textbf{\textit{Ex:}} Solve the following PDE
-\[
-\begin{cases}
-\frac{\partial u}{\partial t} = k \frac{\partial^{2} u}{\partial x^{2}}\\
-u(x,0) = x\\
-u(0,t) = u(L,t) = 0\\
-\end{cases}
-\]
-\indent \textbf{\textit{Solution:}} We know that out solution will be in the form of $ u(x,t) = X(x)T(t) $.
-If we take this fact at face value, we can use the separation of variables method to solve this.
-First, we substitute in our solution into the PDE:
-\begin{align*}
-\frac{\partial}{\partial t}(X(x)T(t)) &= k \frac{\partial^{2}}{\partial x^{2}}(X(x)T(t))\\
-X(x)\frac{dT}{dt} &= k\frac{d^{2}X}{dx^{2}}T(t)\\
-\frac{dT}{dt}\frac{1}{T(t)} &= \frac{d^{2}X}{dx^{2}}\frac{k}{X(x)}
-\end{align*}
-\noindent The only way that our function of x and our function of t can be equal is if they are equal to a constant;
-thus we now have:
-\[
-\frac{dT}{dt}\frac{1}{T(t)} = \frac{d^{2}X}{dx^{2}}\frac{k}{X(x)} = -\lambda
-\]
-\noindent And now we have a first order ODE for the time and a second order ODE for the position:
-\[
-\frac{dT}{dt} = -k \lambda T(t) \qquad\text{and}\qquad \frac{d^{2}X}{dx^{2}} = -\lambda X(x)
-\]
-\noindent Our next step is to find the Eigenvalues and Eigenfunctions of the second order ODE; we will come back to the time ODE when we have values for $\lambda$.\\ 
-\indent To solve for these values, we have to find non-trivial eigenvalues that correspond to the boundary values of the ODE.
-In this case, we know that $X(0) = 0$ and $X(L) = 0$.
-These boundary conditions come from the BC on the PDE which states $u(0,t) = u(L,t) = 0$.
-To find the eigenvalues/functions, we need to find the characteristic equation corresponding to the ODE in question.
-Let's go ahead and find it:
-\begin{gather*}
-\frac{d^{2}X}{dx^{2}} = -\lambda \frac{X(x)}{k}\\
-\frac{d^{2}X}{dx^{2}}  + \lambda \frac{X(x)}{k} = 0\\
-r^{2} + \lambda = 0\\
-r^{2} = -\lambda\\
-r = \pm \sqrt{- \lambda}
-\end{gather*}
-\indent Now, we can substitute values of $\lambda$ to find out solutions.
-Let's start with $\lambda > 0$.
-Since we have a $-\lambda$ under the radical, we will have $r = \pm \sqrt{\lambda} \mathrm{i}$.
-This will yield the solution of
-\[
-y(x) = c_{1}\cos{\sqrt{\lambda}x} + c_{2}\sin{\sqrt{\lambda}x}
-\]
-\noindent Applying our boundary conditions gives us:
-\[
-y(0) = c_{1}\cos{0} + c_{2}\sin{0} = 0
-\]
-\noindent Now we know that $c_{1} = 0$.
-This result came from the fact that $\sin{0} = 0$, so the $c_{2}$ term disappears.
-Now we are left with $\cos{0} = 1$, which tells us that $c_{1} = 0$.
-We can apply this to our second boundary condition like so:
-\begin{gather*}
-y(L) = 0 = c_{2}\sin{\sqrt{\lambda}L}\\
-%\sqrt{\lambda}L = \sin^{-1}{0}
-\end{gather*}
-\noindent It looks like there will only be trivial solutions, but we know this isn't true due to the Jacobian matrix.
-Recall that if the determinant of the Jacobian $\neq 0$, then we only have trivial solutions, and it the determinant of the Jacobian $= 0$, then we do not have the trivial solution.
-Let's take a look at the Jacobian for this case:
-\begin{gather*}
-\left| 
-\begin{array}{cc}
-1 & 0\\
-\cos{\sqrt{\lambda}L} & \sin{\sqrt{\lambda}L}
-\end{array}
-\right| = 0\\
-\sin{\sqrt{\lambda}L} = 0\\
-\sqrt{\lambda}L = \arcsin{0}\\
-\sqrt{\lambda} = \frac{\arcsin{0}}{L}\\
-\lambda = \left(\frac{\arcsin{0}}{L}\right)^{2}\\
-\lambda = \left(\frac{n\pi}{L}\right)^{2}
-\end{gather*}
-\noindent Notice that we replaced the $\arcsin{0}$ term with $n\pi$.
-Recall from trig that the $\sin$ function is $0$ at every $n\pi$, starting with $n = 0$.
-Next, we just need to find our corresponding eigenfunction.
-Our $c_{1}$ was $0$, so we can ignore the entire $\cos$ term.
-Thus, we have an eigenfunction from plugging in our $\lambda$ that looks like:
-\[
-X(x) = \sin{\frac{n\pi x}{L}}
-\]
-\indent Unfortunately, we aren't done yet.
-Now we have to find the corresponding eigenvalues/functions for $\lambda = 0$.
-This isn't too hard, and can be done in a few lines if we recall the characteristic equation.
-Plugging in $0$ for $\lambda$ gives us that $r = 0$.
-Using this piece of information, we now know that $y(x) = c_{1} + xc_{2}$.
-Applying our boundary conditions gives us $c_{1} = 0, c_{2} = 0$.
-I'll leave you to do the math (try pluggin into the Jacobian; you'll see that there is no way to force the determinant to be 0) , but we can see that we have only trivial solutions for $\lambda = 0$.\\
-\indent Our final case is for $\lambda < 0$.
-I'll go ahead and tell you that this will yield another trivial solution, but we can take a look at the work behind it.
-Since $\lambda > 0$, we know that our $r$ from the characteristic equation will be $r = \pm \sqrt{\lambda}$.
-From here we are going to do something a little unconventional that will help us.
-Usually, we would have a solution of the form $y(x) = c_{1}e^{\sqrt{\lambda}x} + c_{2}x e^{-\sqrt{\lambda}x}$.
-In our case, it will be more convenient to use hyperbolic trig functions.
-It's not too far off from our solution from $\lambda > 0$:
-\begin{align*}
-y(x) &= c_{1}e^{\sqrt{\lambda}x} + c_{2}e^{-\sqrt{\lambda}x}\\
-&= c_{1}\cosh{\sqrt{\lambda}x} + c_{2}\sinh{\sqrt{\lambda}x}
-\end{align*}
-\noindent I'll leave you to apply the boundary conditions (remember that $\sinh{0} = 0$ and $\cosh{0} = 1$), but you will find that you have only trivial solutions.\\
-\indent So, we are done with the second order ODE, and now we turn our eye to the first order time equation. We know our $\lambda$, so it isn't too hard to solve:
-\begin{gather*}
-\frac{dT}{dt} = -\lambda T(t)\\
-T(t) = ce^{-k\left(\frac{n\pi}{L}\right)^{2}t}
-\end{gather*}
-\indent Finally we can put it all together. Our penultimate solution will be 
-\[
-u_{n}(x,t) = B_{n}\sin{\left(\frac{n\pi x}{L}\right)}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}
-\]
-\noindent We can analyze this solution and tell where each piece comes from.
-The $B_{n}$ is a coefficient that results from our $c's$ in the general solutions of the differential equations.
-The $sin$ part results from out eigen vector.
-Remember, we get a different value for each $n$, so there may or may not be $0$ terms in the series.
-Finally, we have the exponential term.
-This term comes from t he solution to our time equation.
-When we put it all together, we can see that this is, of course a series; for every $n$, we get a different solution.
-From here we can find the Fourier Series (which will be a Fourier Sine Series).
-We won't worry about that yet, but we will come back to it.
+%\indent Let's look at another second order PDE: the wave equation. This equation will be a little harder to solve due to the fact that it contains two second order derivates. Let's take a look:\\
+%\noindent\textbf{\textit{Ex:}} Solve the following PDE
+%\[
+%\begin{cases*}
+%u_{tt} - u_{xx} = 0\\
+%u(0,t)=u(L,t) = 0\\
+%u(x,0) = f(x),\ u_{t}(x,0) = g(x)
+%\end{cases*}
+%\]
+%\indent Let's take a second before we jump into solving this equation to analyze it.
+We have a homogeneous, second order PDE.
+We also have boundary conditions, much like with the heat equation.
+The initial conditions are a little strange, but we won't get into them until we discuss Fourier series.\\
+%\noindent We will use separation of variables to solve this equation. With this technique, we know that we are looking for a solution in the form of 
+%\[
+%u(x,t) = \xi(x)\phi(t).
+%\]
+%\noindent So, we substitute into the original PDE, and we get as follows:
+%\begin{gather*}
+%u_{tt} - u_{xx} = 0 =\phi^{''}\xi - \xi^{''}\phi\\
+%\frac{\phi^{''}}{\phi} = \frac{\xi^{''}}{\xi} = \lambda
+%\end{gather*}
+%\noindent Now we need to find our eigenvalues and eigenfunctions.
+There's a trick here, however.
+In the heat equation, we had the liberty of solving the time equation whenever we wanted to.
+Here, we don't have that ability just yet, since $\phi(t)$ is ``trapped'' in a second order ODE.
+The only real choice we have here is to do what we did earlier, and apply our boundary conditions to the $\xi(x)$ ODE.
+We won't delve into the process for finding the eigenvalues, so here's the only non-trivial pair:
+%\[
+%\{\lambda_{n} = (n\pi)^{2},\ \sin{(n\pi x)}\}
+%\]
+%\indent Now let's take a second to imagine that we had had boundary conditions for our second order ODE.
+We would have had the same setup for finding the eigenvalues, but our eigenfunction would have involved $t$ instead of $x$.
 
@@
 \]
-\noindent And that will suffice.
+\noindent And we are done.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Second Order Partial Differential Equations}
+\hrule
+\noindent\\
+\subsection{The Heat and Wave Equations}
+\indent Let's take a look at the second order PDE.
+We will start with a homogeneous equation, which we will tackle with the separation of variables method.
+The general strategy is that we want to take our PDE and convert it to an ODE, which we know how to solve.
+The problems presented here also involve boundary conditions, which will require us to solve for eigenvalues/functions.
+We will also be dealing with series (specifically Fourier Series) to help us solve these problems.
+Let's start by looking at the homogeneous heat equation.\\
+\noindent \textbf{\textit{Ex:}} Solve the following PDE
+\[
+\begin{cases}
+\frac{\partial u}{\partial t} = k \frac{\partial^{2} u}{\partial x^{2}}\\
+u(x,0) = x\\
+u(0,t) = u(L,t) = 0\\
+\end{cases}
+\]
+\indent \textbf{\textit{Solution:}} We will use the separation of variable technique to solve this problem.
+We know to use separation of variable because we are dealing with a homogeneous PDE with homogenous boundary conditions.
+Separation of variables tells us that our solution will be in the form of $ u(x,t) = X(x)T(t) $.
+If we take this fact at face value, we can use the separation of variables method to solve this.
+First, we substitute in our solution into the PDE:
+\begin{align*}
+\frac{\partial}{\partial t}(X(x)T(t)) &= k \frac{\partial^{2}}{\partial x^{2}}(X(x)T(t))\\
+X(x)\frac{dT}{dt} &= k\frac{d^{2}X}{dx^{2}}T(t)\\
+\frac{dT}{dt}\frac{1}{T(t)} &= \frac{d^{2}X}{dx^{2}}\frac{k}{X(x)}
+\end{align*}
+\noindent The only way that our function of x and our function of t can be equal is if they are equal to a constant;
+thus we now have:
+\[
+\frac{dT}{dt}\frac{1}{T(t)} = \frac{d^{2}X}{dx^{2}}\frac{k}{X(x)} = -\lambda
+\]
+\noindent And now we have a first order ODE for the time and a second order ODE for the position:
+\[
+\frac{dT}{dt} = -k \lambda T(t) \qquad\text{and}\qquad \frac{d^{2}X}{dx^{2}} = -\lambda X(x)
+\]
+\noindent Our next step is to find the Eigenvalues and Eigenfunctions of the second order ODE; we will come back to the time ODE when we have values for $\lambda$.\\ 
+\indent To solve for these values, we have to find non-trivial eigenvalues that correspond to the boundary values of the ODE.
+In this case, we know that $X(0) = 0$ and $X(L) = 0$.
+These boundary conditions come from the BC on the PDE which states $u(0,t) = u(L,t) = 0$.
+To find the eigenvalues/functions, we need to find the characteristic equation corresponding to the ODE in question.
+Let's go ahead and find it:
+\begin{gather*}
+\frac{d^{2}X}{dx^{2}} = -\lambda \frac{X(x)}{k}\\
+\frac{d^{2}X}{dx^{2}}  + \lambda \frac{X(x)}{k} = 0\\
+r^{2} + \lambda = 0\\
+r^{2} = -\lambda\\
+r = \pm \sqrt{- \lambda}
+\end{gather*}
+\indent Now, we can substitute values of $\lambda$ to find out solutions.
+Let's start with $\lambda > 0$.
+Since we have a $-\lambda$ under the radical, we will have $r = \pm \sqrt{\lambda} \mathrm{i}$.
+This will yield the solution of
+\[
+y(x) = c_{1}\cos{\sqrt{\lambda}x} + c_{2}\sin{\sqrt{\lambda}x}
+\]
+\noindent Applying our boundary conditions gives us:
+\[
+y(0) = c_{1}\cos{0} + c_{2}\sin{0} = 0
+\]
+\noindent Now we know that $c_{1} = 0$.
+This result came from the fact that $\sin{0} = 0$, so the $c_{2}$ term disappears.
+Now we are left with $\cos{0} = 1$, which tells us that $c_{1} = 0$.
+We can apply this to our second boundary condition like so:
+\begin{gather*}
+y(L) = 0 = c_{2}\sin{\sqrt{\lambda}L}\\
+\end{gather*}
+\noindent It looks like there will only be trivial solutions, but we know this isn't true due to the Jacobian matrix.
+Recall that if the determinant of the Jacobian $\neq 0$, then we only have trivial solutions, and it the determinant of the Jacobian $= 0$, then we do not have the trivial solution.
+Let's take a look at the Jacobian for this case:
+\begin{gather*}
+\left| 
+\begin{array}{cc}
+1 & 0\\
+\cos{\sqrt{\lambda}L} & \sin{\sqrt{\lambda}L}
+\end{array}
+\right| = 0\\
+\sin{\sqrt{\lambda}L} = 0\\
+\sqrt{\lambda}L = \arcsin{0}\\
+\sqrt{\lambda} = \frac{\arcsin{0}}{L}\\
+\lambda = \left(\frac{\arcsin{0}}{L}\right)^{2}\\
+\lambda = \left(\frac{n\pi}{L}\right)^{2}
+\end{gather*}
+\noindent Notice that we replaced the $\arcsin{0}$ term with $n\pi$.
+Recall from trig that the $\sin$ function is $0$ at every $n\pi$, starting with $n = 0$.
+Next, we just need to find our corresponding eigenfunction.
+Our $c_{1}$ was $0$, so we can ignore the entire $\cos$ term.
+Thus, we have an eigenfunction from plugging in our $\lambda$ that looks like:
+\[
+X(x) = \sin{\frac{n\pi x}{L}}
+\]
+\indent Unfortunately, we aren't done yet.
+Now we have to find the corresponding eigenvalues/functions for $\lambda = 0$.
+This isn't too hard, and can be done in a few lines if we recall the characteristic equation.
+Plugging in $0$ for $\lambda$ gives us that $r = 0$.
+Using this piece of information, we now know that $y(x) = c_{1} + xc_{2}$.
+Applying our boundary conditions gives us $c_{1} = 0, c_{2} = 0$.
+I'll leave you to do the math (try pluggin into the Jacobian; you'll see that there is no way to force the determinant to be 0) , but we can see that we have only trivial solutions for $\lambda = 0$.\\
+\indent Our final case is for $\lambda < 0$.
+I'll go ahead and tell you that this will yield another trivial solution, but we can take a look at the work behind it.
+Since $\lambda > 0$, we know that our $r$ from the characteristic equation will be $r = \pm \sqrt{\lambda}$.
+From here we are going to do something a little unconventional that will help us.
+Usually, we would have a solution of the form $y(x) = c_{1}e^{\sqrt{\lambda}x} + c_{2}x e^{-\sqrt{\lambda}x}$.
+In our case, it will be more convenient to use hyperbolic trig functions.
+It's not too far off from our solution from $\lambda > 0$:
+\begin{align*}
+y(x) &= c_{1}e^{\sqrt{\lambda}x} + c_{2}e^{-\sqrt{\lambda}x}\\
+&= c_{1}\cosh{\sqrt{\lambda}x} + c_{2}\sinh{\sqrt{\lambda}x}
+\end{align*}
+\noindent I'll leave you to apply the boundary conditions (remember that $\sinh{0} = 0$ and $\cosh{0} = 1$), but you will find that you have only trivial solutions.\\
+\indent So, we are done with the second order ODE, and now we turn our eye to the first order time equation. We know our $\lambda$, so it isn't too hard to solve:
+\begin{gather*}
+\frac{dT}{dt} = -\lambda T(t)\\
+T(t) = ce^{-k\left(\frac{n\pi}{L}\right)^{2}t}
+\end{gather*}
+\indent Finally we can put it all together. Our penultimate solution will be 
+\[
+u_{n}(x,t) = B_{n}\sin{\left(\frac{n\pi x}{L}\right)}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}
+\]
+\noindent We can analyze this solution and tell where each piece comes from.
+The $B_{n}$ is a coefficient that results from our $c's$ in the general solutions of the differential equations.
+The $sin$ part results from out eigen vector.
+Remember, we get a different value for each $n$, so there may or may not be $0$ terms in the series.
+Finally, we have the exponential term.
+This term comes from t he solution to our time equation.
+When we put it all together, we can see that this is, of course a series; for every $n$, we get a different solution.
+From here we can find the Fourier Series (which will be a Fourier Sine Series).
+We won't worry about that yet, but we will come back to it.
+\newpage
+
+\indent It will benefit us to take a moment and discuss how boundary conditions effect our Fourier series.
+We have three different combinations of boundary conditions, where each one will produce a different result in our series solution.
+\begin{enumerate}
+\item \textit{Dirichlet Boundary Conditions}\\
+Dirichlet boundary conditions involve the function for which we are looking.
+In other words, the boundary conditions are presented in terms of the functions we are trying to find, rather than the derivative of that function.
+Typically, this type of boundary condition will take the form of (in the homogeneous case)
+\[u(0,t) = u(L,t) = 0.\]
+In regards to our Fourier series, we expect to see $\sin{(\frac{n\pi x}{L})}$. Our basis will look like
+\[
+\left\{ \lambda_{n} = \left(\frac{n\pi}{2}\right)^{2}, \phi_{n}(x) = \sin{\left(\frac{n\pi x}{L}\right)} \right\}
+\]
+\item \textit{Neumann Boundary Conditions}\\
+Neumann boundary conditions involve the derivatives of the function for which we are solving. Neumann boundary conditions will look like
+\[u_{x}(0,t) = u_{x}(L,t) = 0.\]
+This will cause our Fourier series to contain $\cos{(\frac{n\pi x}{L})}$. Out basis will look like
+\[
+\left\{ \lambda_{n} = \left(\frac{n\pi}{2}\right)^{2}, \phi_{n}(x) = \cos{\left(\frac{n\pi x}{L}\right)} \right\}
+\]
+\item \textit{Robin Boundary Conditions}\\
+Robin boundary conditions will contain a mixture of the function and it's derivative. Typically, we will see
+\[ u_{x}(0,t) = u(L,t) = 0\qquad \text{or}\qquad u(0,t) = u_{x}(L,0) = 0.\]
+These boundary conditions will give us two different results, depending on whether we have the derivative first or second. We will have
+\begin{align*}
+u_{x}(0,t) = u(L,t) = 0 &\implies \left\{\lambda_{n} = \left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}\right)^{2}, \phi_{n}(x) = \cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}x\right)}\right\}\\
+u(0,t) = u_{x}(L,t) = 0 &\implies \left\{\lambda_{n} = \left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}\right)^{2}, \phi_{n}(x) =\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}x\right)}\right\}
+\end{align*}
+\end{enumerate}
+
+\indent Now we can take a look at a slightly harder PDE.
+We still have the heat equation, but we will not have homogeneous boundary conditions.
+Our goal here is to find a way to transform our boundary conditions into a homogenous form and take it from there.\\\\
+\noindent \textbf{\textit{Ex:}} Solve the following PDE
+\[
+\begin{cases*}
+u_{t} - u_{xx} = 0\\
+u(0,t) = 0,\ u(1,t) = t\\
+u(x,0) = x^{2}
+\end{cases*}
+\]
+\indent \textbf{\textit{Solution:}} We would like to just jump into this PDE and solve it, but it's not quite that simple.
+We have to find a way to express our boundary conditions in a homogeneous manner.
+We can do this through a clever substitution; we will create a new PDE, denoted $\Upsilon$.
+To find $\Upsilon$, we use the following formula:
+\[ \Upsilon = B_{1} + \frac{x}{l}\Delta B, \]
+\noindent where $B$ is a boundary condition. In our case, we have as follows:
+\begin{align*}
+\Upsilon &= B_{1} + \frac{x}{l}\Delta B\\
+&= 0 + \frac{x}{1}(t - 0)\\
+&= xt
+\end{align*}
+\noindent Now, we find a new PDE, denoted by $\upsilon$. We have another formula that tells us
+\[ \upsilon(x,t) = u(x,t) - \Upsilon(x,t) \]
+\noindent Applying this formula gives us
+\begin{align*}
+\upsilon(x,t) &= u(x,t) - \Upsilon\\
+&= u(x,t) - xt
+\end{align*}
+\noindent Let's take a moment to analyze the result.
+We wanted to find a new PDE in terms of our old one that would allow us to have homogeneous boundary conditions.
+We found a solution, called $\upsilon$, in terms of our original PDE, called $u$.
+Our original PDE was $u_{t} - u_{xx} = 0$, so to properly perform our substitution, we need to find $\upsilon_{t}$ and $\upsilon_{xx}$.
+Through some careful differentiation, we have
+\[ \upsilon_{t} = u_{t} - x \qquad\text{and}\qquad \upsilon_{xx} = u_{xx}\]
+\noindent To find our new PDE, we simply evaluate.
+\[
+\upsilon_{t} - \upsilon_{xx} = u_{t} - x - u_{xx} = u_{t} - u_{xx} - x = 0 - x = -x
+\]
+\noindent So we finally have a new PDE that looks like
+\[
+\begin{cases*}
+\upsilon_{t} - \upsilon_{xx} = -x\\
+\upsilon(0,t) = \upsilon(1,t) = 0\\
+\upsilon(x,0) = x^{2}
+\end{cases*}
+\]
+\noindent You may wonder how this has made anything simpler, since the PDE itself is no longer homogeneous.
+By forcing the boundary conditions to be homogeneous, we can now use a Fourier series to find our solution.
+Recall that earlier we discussed the boundary condition and their resulting eigenvalue/eigenfunction.
+Here, we have Dirichlet boundary conditions, which tells us that we have a Fourier sine series.
+We have that
+\begin{align*}
+\upsilon(x,t) &= \sum_{n = 1}^{\infty}\left[a_{n}(t)\sin{(n\pi x)}\right]\\
+\upsilon_{t}(x,t) &= \sum_{n = 1}^{\infty}\left[a_{n}'(t)\sin{(n\pi x)}\right]\\
+\upsilon_{xx}(x,t) &= -\sum_{n = 1}^{\infty}\left[a_{n}(t)(n\pi)^{2}\sin{(n\pi x)}\right]\\ 
+\end{align*}
+\noindent Which tells us that
+\[
+\upsilon_{t} - \upsilon_{xx} = \sum_{n = 1}^{\infty}\left[(a_{n}'(t) + (n\pi)^{2}a_{n}(t))\sin{(n\pi x)}\right] = -x
+\]
+\noindent This may look rather complicated, but all we have is a function's Fourier sine series. Essentially, we have
+\[
+-x = \sum_{n = 1}^{\infty}\left[ b_{n}\sin{(n\pi x)} \right]
+\]
+\noindent To find $b_{n}$, we use the formula that was discussed in the earlier section about Fourier series. We have that
+\begin{align*}
+a_{n}'(t) + (n\pi)^{2}a_{n}(t) &= \int_{0}^{1}-x\sin{(n\pi x)}dx\\
+&= \frac{2x}{n\pi}\cos{(n\pi x)}\bigg\vert_{0}^{1}\\
+&= \frac{2}{n\pi}
+\end{align*}
+\noindent We now have an ODE which we can solve for $a_{n}$.
+Note that from out PDE, we do have an initial condition for the ODE, but it is not as simple as just saying $a_{n}(0) = x^{2}$.
+At $t = 0$, we have the following series:
+\[
+\sum_{n=1}^{\infty} a_{n}(0)\sin{(n\pi x)}
+\]
+\noindent We got this by taking our original $\upsilon(x,t)$ and plugging in $0$ for $t$. We know how to get $a_{n}(0)$, since this is just a simple Fourier sine series. It is given by
+\begin{align*}
+a_{n}(0) &= 2\int_{0}^{1}x^{2}\sin{(n\pi x)}dx\\
+&= \frac{(-1)^{(n + 1)}}{n\pi} + \frac{2}{(n\pi)^{3}}\left[ (-1)^{n} - 1 \right]
+\end{align*}
+\noindent We now have all we need to solve the ODE. We will omit the work for the solution, but the final answer for $a_{n}$ is 
+\[
+a_{n}(t) = \frac{4\left[ (-1)^{n} - 1 \right]}{(n\pi)^{3}}e^{-(n\pi)^{2}t} + \frac{2(-1)^{n}}{(n\pi)^{3}}
+\]
+\noindent Now all that's left to do is go back to $u$ from $\upsilon$.
+This may seem hard, but we saw earlier that $\upsilon = u - \Upsilon$.
+Solving for $u$ gives us $u = \upsilon + \Upsilon$.
+Recall that $\Upsilon$ was $xt$, and $\upsilon$ was the series for which we just found $a_{n}$.
+Putting it all together gives us
+\[
+u(x,t) = xt + \sum_{n=1}^{\infty}\left[\frac{4\left[ (-1)^{n} - 1 \right]}{(n\pi)^{3}}e^{-(n\pi)^{2}t} + \frac{2(-1)^{n}}{(n\pi)^{3}}\right]\sin{(n\pi x)}
+\]
+\noindent And we are done.
+\newpage
+
+\noindent \textbf{\textit{Ex:}} Solve the following PDE
+\[
+\begin{cases*}
+u_{tt} - u_{xx} = 1 + t\cos{(2\pi x)}\\
+u_{x}(0,t) = u_{x}(1,t) = 0\\
+u(x,0) = 2\cos{(2\pi x)},\quad u_{t}(x,0) = 1 + \cos{(\pi x)} - \cos{(2\pi x)}
+\end{cases*}
+\]
+\indent \textbf{\textit{Solution:}} This problem is an example of the wave equation with a non-homogeneous solution.
+We are lucky this time; the boundary conditions are already homogeneous, so there is no need to change our variable.
+We know that since both of the boundary conditions are derivatives, we are dealing with a Fourier cosine series.
+This yields the eigenvalue/function of
+\[
+\left\{\lambda_{n} = (n\pi)^{2}, \cos{(n\pi x)}\right\}
+\]
+Since we are dealing with a cosine series, we also have a leading term, $a_{0}$. Our full series looks like
+\[
+u(x,t) = a_{0}(t) + \sum_{n=1}^{\infty} a_{n}(t)\cos{(n\pi x)}
+\]
+We need $u_{tt}$ and $u_{xx}$, so we have:
+\[
+u_{tt}(x,t) = a_{0}''(t) + \sum_{n=1}^{\infty} a_{n}''(t)\cos{(n\pi x)} 
+\qquad \text{and} \qquad 
+u_{xx}(x,t) = -\sum_{n=1}^{\infty} (n\pi)^{2}a_{n}(t)\cos{(n\pi x)}
+\]
+All we have to do now is simply plug into our original equation. This gives us
+\begin{align*}
+a_{0}''(t) + \sum_{n=1}^{\infty} a_{n}''(t)\cos{(n\pi x)} + \sum_{n=1}^{\infty} (n\pi)^{2}a_{n}(t)\cos{(n\pi x)} &= 1 + t\cos{(2\pi x)}\\
+a_{0}''(t) + \sum_{n=1}^{\infty}(a_{n}''(t) + (n\pi)^{2}a_{n}(t))\cos{(n\pi x)} &= 1 + t\cos{(2\pi x)}
+\end{align*}
+Now we can play the matching game. We have a $constant + series\ term$ on the right side, and $constant + series$ on the left, so we can match up the constants. This gives us 
+\[
+a_{0}''(t) = 1
+\]
+Now we match up the series term. On the right hand side, we have $\cos{(2\pi x)}$, so we know that for $n= 2$, we have
+\[
+a_{2}''(t) + 4\pi^{2}a_{2}(t) = t
+\]
+$a_{0}$ and $a_{2}$ are the only terms that appear, so we can assume for now that $\forall n \neq 2\ \text{and}\ 3, a_{n} = 0$.
+We are left with two ODE's, so we now need to find the corresponding initial conditions.
+Let's apply the first initial condition to our PDE.
+This gives us
+\[a_{0}(0) + \sum_{n=1}^{\infty}a_{n}(0)\cos{(n\pi x)} = 2\cos{(2\pi t)}\]
+Which tells us
+\[a_{0} = 0 \qquad\qquad \qquad\qquad a_{2} = 2\]
+Now we can apply the second initial condition. After taking the partial derivative of $u$ with respect to $t$, we have
+\[
+a_{0}'(0) + \sum_{n=1}^{\infty}a_{n}'(0)\cos{(n\pi x)} = 1 + \cos{(\pi x)} - \cos{(2\pi x)}
+\]
+Some more clever pattern matching give us
+\[
+a_{0}'(0) = 1 \qquad\qquad a_{1}'(0) = 1 \qquad\qquad a_{2}'(0) = -1
+\]
+Notice that we have a term $a_{1}'$.
+We cannot simply ignore this value; rather we have to take it into account when we find our solution.
+We know that we have an ODE for $a_{1}(t)$, but it may not be evident how to express it.
+Recall that earlier we said $\forall n \neq 2\ \text{and}\ 3, a_{n} = 0$.
+This means that $a_{1} = 0$, and since $a_{1}$ is a coefficient in the Fourier series, we know that $a_{1}''(t) + \pi^{2}a_{1}(t) = 0$.
+We now have three different ODE's that we must solve:
+\begin{alignat*}{3}
+&\begin{cases}
+a_{0}''(t) = 1\\
+a_{0}'(0) = 1,\quad a_{0}(0) = 1
+\end{cases}
+&\begin{cases}
+a_{1}''(t) + \pi^{2}a_{1}(t) = 0\\
+a_{1}'(0) = 1,\quad a_{1}(0) = 0
+\end{cases}
+&&\begin{cases}
+a_{2}''(t) + 4\pi^{2}a_{2}(t) = t\\
+a_{2}'(0) = 1,\quad a_{2}(0) = 2
+\end{cases}
+\end{alignat*}
+Luckily, Mathematica is here to save the day. We will skip the work and let the computer do the thinking for us here. We get that
+\begin{align*}
+a_{0}(t) &= \frac{1}{2}t^{2} + t\\
+a_{1}(t) &= \frac{\pi\cos{(t\pi)} + \sin{(t\pi)}}{\pi}\\
+a_{2}(t) &= \frac{2\pi t + 16\pi^{3}\cos{(2\pi t)} - \sin{(2\pi t)} + 4\pi^{2}\sin{(2\pi t)}}{8\pi^{3}}
+\end{align*}
+Now all we need to do is plug back in to our $u$. Our final answer becomes
+\[
+u(x,t) = \frac{1}{2}t^{2} + t + \frac{\pi\cos{(t\pi)} + \sin{(t\pi)}}{\pi}\cos{(\pi x)} +  \frac{2\pi t + 16\pi^{3}\cos{(2\pi t)} - \sin{(2\pi t)} + 4\pi^{2}\sin{(2\pi t)}}{8\pi^{3}}\cos{(2\pi x)}
+\]
+And we are done.
+\newpage
+
+\subsection{The Laplace Equation}
+\noindent \\
+\indent Up to this point, we have only looked at the wave equation and the heat equation.
+There is, however, one other equation in the classical trinity: the Laplace Equation.
+We haven't dealt with this equation yet because we have to be very careful in our approach to solving it.
+When we looked at the homogeneous wave and heat equations, we used the general approach called \textit{Separation of Variables}.
+In the case of the homogeneous Laplace equation, we may not be able to tackle the problem using separation of variables.
+The problem arises from the various types of boundaries that effect the domain.
+We can either have a 'rectangular' bound, which will be separable, or a 'circular' bound which is \textbf{not} separable.
+Imagine that we have the following two problems:
+\begin{align*}
+&A =\begin{cases}
+u_{xx} + u_{yy} = 0\quad \text{on}\ [0,1]\times[0,2]\\
+u(x,0) = u(x,2) = 0\\
+u(0,y) = y, u(1,y) = 0
+\end{cases}
+&B =\begin{cases}
+u_{xx} + u_{yy} = 0\quad \text{on}\ x^{2} + y^{2} < a^{2}\\
+u(x,y)\ \big |_{x^{2} + y^{2} = a^{2}} = f(x,y)
+\end{cases}
+\end{align*}
+$A$ is an example of a separable Laplace equation.
+We can see graphically that the domain $[0,1] \times [0,2]$ is rectangular.
+$B$'s domain is not separable, since it is circular.
+Let's take a look $A$.\\\\
+\noindent \textbf{\textit{Ex:}} Solve the following PDE
+\[
+\begin{cases*}
+u_{xx} + u_{yy} = 0\quad \text{on}\ [0,1]\times[0,2]\\
+u(x,0) = u(x,2) = 0\\
+u(0,y) = y, u(1,y) = 0
+\end{cases*}
+\]
+\indent \textbf{\textit{Solution:}} Earlier, we discussed the different boundary conditions we can encounter and their corresponding eigenvalues/functions.
+In this problem, we have Dirichlet boundary conditions.
+This tells us that
+\[
+\left\{\lambda_{n} = \left(\frac{n\pi}{2}\right)^{2}, \phi_{n}(y) = \sin{\left(\frac{n\pi y}{2}\right)}\right\}
+\]
+Notice that we used $y$ here instead of $x$.
+We did this because the boundary conditions are defined in terms of $y$, not $x$.
+Now we can continue to solve our problem.
+Recall that our separation of variables gives us two second order ODE's,
+\[
+X'' - \lambda X = 0 \qquad\qquad \text{and} \qquad\qquad \lambda Y - Y'' = 0
+\]
+We already solved for $Y(y)$ by finding the corresponding eigenvalues/vectors. Now we plug in our $\lambda_{n}$ into our $X$ ODE. Solving the resulting ODE will give us a basis of
+\[
+\left\{ e^{\frac{n\pi}{2}x}, e^{-\frac{n\pi}{2}x} \right\}
+\]
+Now we could continue at this point, but if we think ahead, we can see that when we eventually apply our initial conditions, nothing will cancel out.
+At this point, we want to transform our basis.
+We have as follows
+\[
+\left\{ e^{\frac{n\pi}{2}x}, e^{-\frac{n\pi}{2}x} \right\} \implies \left\{ \cosh{\left(\frac{n\pi x}{2}\right)}, \sinh{\left(\frac{n\pi x}{2}\right)} \right\}
+\]
+All we did here is apply an identity. Now we have everything we need to set up a solution. Our $u(x,y)$ will be
+\[
+u(x,y) = \sum_{n=1}^{\infty}\left[a_{n}\cosh{\left(\frac{n\pi x}{2}\right)} + b_{n}\sinh{\left(\frac{n\pi x}{2}\right)}\right]\sin{\left(\frac{n\pi y}{2}\right)}
+\]
+We can apply our initial conditions, which gives us $b_{n}\ \text{and}\ a_{n}$:
+\begin{align*}
+a_{n} &= \frac{4(-1)^{n+1}}{n\pi}\\
+b_{n} &= -\coth{\left(\frac{n \pi}{2}\right)}\frac{4(-1)^{n+1}}{n\pi}
+\end{align*}
+This yields a final solution of
+\[
+u(x,y) = \sum_{n=1}^{\infty}\left[ \frac{4(-1)^{n+1}}{n\pi} \cosh{\left(\frac{n\pi x}{2}\right)} -\coth{\left(\frac{n \pi}{2}\right)}\frac{4(-1)^{n+1}}{n\pi}\sinh{\left(\frac{n\pi x}{2}\right)}\right]\sin{\left(\frac{n\pi y}{2}\right)}
+\]
+And we are done
+\newpage
+
+\indent Now we turn our attention to the Laplace Equation with a circular domain.
+As we discussed earlier, we can't simply say that our solution will be in the form of $X(x)Y(y)$.
+However, we can force the domain to be rectangular if we use polar coordinates.
+The translation is not too difficult.
+Let's consider $B$ from above.\\\\
+
+\noindent \textbf{\textit{Ex:}} Solve the following PDE
+\[
+\begin{cases*}
+u_{xx} + u_{yy} = 0\quad \text{on}\ x^{2} + y^{2} < a^{2}\\
+u(x,y)\ \big |_{x^{2} + y^{2} = a^{2}} = f(x,y)
+\end{cases*}
+\]
+\indent \textbf{\textit{Solution:}} We want to express the partial derivatives here in terms of $r\ \text{and}\ \theta$. Recall that we have
+\[
+x = r\cos{\theta} \qquad\qquad \text{and} \qquad\qquad y = r\sin{\theta}
+\]
+First, we will find the partial derivatives of $x$ and $y$ with respect to $r$ and $\theta$. Since $x = r\cos{\theta}$, 
+\[
+\frac{\partial x}{\partial r} = \cos{\theta} \qquad\qquad \frac{\partial x}{\partial \theta} = -r\sin{\theta}
+\]
+We can compute the derivatives for $y$ in a similar fashion:
+\[
+\frac{\partial y}{\partial r} = \sin{\theta} \qquad\qquad \frac{\partial y}{\partial \theta} = r\cos{\theta}
+\]
+Now we need to find $u_{rr}$ and $u_{\theta\theta}$. Let's focus on $u_{rr}$ first. We will need to start by finding $u_{r}$. We have as follows
+\begin{align*}
+\frac{\partial u}{\partial r}& = \frac{\partial u}{\partial x}\frac{\partial x}{\partial r} + \frac{\partial u}{\partial y}\frac{\partial y}{\partial r}\\
+&= \cos{\theta}\frac{\partial u}{\partial x} + \sin{\theta}\frac{\partial u}{\partial y}
+\end{align*}
+We have $u_{r}$, so now we need to compute $u_{rr}$. We have as follows:
+\begin{align*}
+\frac{\partial^{2} u}{\partial r^{2}} &= \cos{\theta}\frac{\partial }{\partial r}\frac{\partial u}{\partial x} + \sin{\theta}\frac{\partial }{\partial r}\frac{\partial u}{\partial y}\\
+&= \cos^{2}{\theta}\frac{\partial^{2} u}{\partial x^{2}} + 2\cos{\theta}\sin{\theta}\frac{\partial^{2}u}{\partial x\partial y} + \sin^{2}{\theta}\frac{\partial^{2} u}{\partial y^{2}}
+\end{align*}
+We won't delve into finding $u_{\theta\theta}$, as the partials get a little out of hand. However, we end up with
+\begin{align*}
+\frac{\partial^{2}u}{\partial r^{2}} + \frac{1}{r^{2}}\frac{\partial^{2}u}{\partial\theta^{2}} &= -\frac{1}{r}\frac{\partial u}{\partial r} + \frac{\partial^{2} u}{\partial x^{2}} + \frac{\partial^{2} u}{\partial y^{2}}\\
+u_{xx} + u_{yy} &= u_{rr} + \frac{1}{r}u_{r} + \frac{1}{r^{2}}u_{\theta\theta} = 0
+\end{align*}
+We now have the Laplace equation represented in the polar coordinate system.
+Note that the translation of the boundary condition is simple.
+It becomes a function of only $\theta$.
+We have the following system:
+\[
+\begin{cases*}
+u_{rr} + \frac{1}{r}u_{r} + \frac{1}{r^{2}}u_{\theta\theta} = 0\quad \text{on}\ 0 < r < a\\
+u(\theta) = f(\theta)\quad \text{on}\ 0 < \theta < 2\pi
+\end{cases*}
+\]
+Now we can use separation of variables to solve the PDE. The steps are the same (although this time we will use $R(r)\ \text{and}\ T(\theta)$). We will end up with the following equivalence:
+\[
+\frac{r^{2}R'' + rR'}{R} = \frac{T''}{T} = \lambda
+\]
+Since we have a Dirichlet boundary conditions, we know that
+\[
+\lambda_{n} = \frac{n^{2}}{4}
+\]
+Plugging this into our equation for $R$ gives us
+\[
+r^{2}R'' + rR' - \frac{n^{2}}{4}R = 0 
+\]
 \end{document}

### 2015-10-26 pde/pdetest1.tex
added laplace

@@
 \usepackage{graphicx}
+\usepackage{multicol}
 \topmargin=-0.3in
@@
 \begin{document}
-%\indent Let's look at another second order PDE: the wave equation. This equation will be a little harder to solve due to the fact that it contains two second order derivates. Let's take a look:\\
-%\noindent\textbf{\textit{Ex:}} Solve the following PDE
-%\[
-%\begin{cases*}
-%u_{tt} - u_{xx} = 0\\
-%u(0,t)=u(L,t) = 0\\
-%u(x,0) = f(x),\ u_{t}(x,0) = g(x)
-%\end{cases*}
-%\]
-%\indent Let's take a second before we jump into solving this equation to analyze it.
-We have a homogeneous, second order PDE.
-We also have boundary conditions, much like with the heat equation.
-The initial conditions are a little strange, but we won't get into them until we discuss Fourier series.\\
-%\noindent We will use separation of variables to solve this equation. With this technique, we know that we are looking for a solution in the form of 
-%\[
-%u(x,t) = \xi(x)\phi(t).
-%\]
-%\noindent So, we substitute into the original PDE, and we get as follows:
-%\begin{gather*}
-%u_{tt} - u_{xx} = 0 =\phi^{''}\xi - \xi^{''}\phi\\
-%\frac{\phi^{''}}{\phi} = \frac{\xi^{''}}{\xi} = \lambda
-%\end{gather*}
-%\noindent Now we need to find our eigenvalues and eigenfunctions.
-There's a trick here, however.
-In the heat equation, we had the liberty of solving the time equation whenever we wanted to.
-Here, we don't have that ability just yet, since $\phi(t)$ is ``trapped'' in a second order ODE.
-The only real choice we have here is to do what we did earlier, and apply our boundary conditions to the $\xi(x)$ ODE.
-We won't delve into the process for finding the eigenvalues, so here's the only non-trivial pair:
-%\[
-%\{\lambda_{n} = (n\pi)^{2},\ \sin{(n\pi x)}\}
-%\]
-%\indent Now let's take a second to imagine that we had had boundary conditions for our second order ODE.
-We would have had the same setup for finding the eigenvalues, but our eigenfunction would have involved $t$ instead of $x$.
-
-\newpage
 \section{First Order Partial Differential Equations}
@@
 \end{alignat*}
-Luckily, Mathematica is here to save the day. We will skip the work and let the computer do the thinking for us here. We get that
+Luckily, Mathematica is here to save the day. We will skip the work and let the computer do the thinking for us. We get that
 \begin{align*}
@@
 \]
-And we are done.
+Just for interest, the plot of this solution looks like
+\begin{center}
+\includegraphics[scale=0.6]{pg23_plot} 
+\end{center}
+On $-\pi \leq \theta \leq \pi$ and $-\pi \leq x \leq \pi$; and we are done.
 \newpage
@@
 $B$'s domain is not separable, since it is circular.
-Let's take a look $A$.\\\\
+Let's take a look at $A$.\\\\
 \noindent \textbf{\textit{Ex:}} Solve the following PDE
@@
 However, we can force the domain to be rectangular if we use polar coordinates.
-The translation is not too difficult.
-Let's consider $B$ from above.\\\\
+Let's consider a variation of $B$ from above.\\\\
 
@@
 u_{xx} + u_{yy} = 0\quad \text{on}\ x^{2} + y^{2} < a^{2}\\
-u(x,y)\ \big |_{x^{2} + y^{2} = a^{2}} = f(x,y)
+u(a,\theta) = U_{1}\quad \text{on}\ 0 \leq \theta \leq \pi\\
+u(a,\theta) = U_{2}\quad \text{on}\ \pi \leq \theta \leq 2\pi
 \end{cases*}
@@
 \end{align*}
-We now have the Laplace equation represented in the polar coordinate system.
-Note that the translation of the boundary condition is simple.
-It becomes a function of only $\theta$.
-We have the following system:
-\[
-\begin{cases*}
-u_{rr} + \frac{1}{r}u_{r} + \frac{1}{r^{2}}u_{\theta\theta} = 0\quad \text{on}\ 0 < r < a\\
-u(\theta) = f(\theta)\quad \text{on}\ 0 < \theta < 2\pi
-\end{cases*}
-\]
-Now we can use separation of variables to solve the PDE. The steps are the same (although this time we will use $R(r)\ \text{and}\ T(\theta)$). We will end up with the following equivalence:
-\[
-\frac{r^{2}R'' + rR'}{R} = \frac{T''}{T} = \lambda
-\]
-Since we have a Dirichlet boundary conditions, we know that
-\[
-\lambda_{n} = \frac{n^{2}}{4}
-\]
-Plugging this into our equation for $R$ gives us
-\[
-r^{2}R'' + rR' - \frac{n^{2}}{4}R = 0 
-\]
+Lets go ahead and formalize it:
+\[
+\nabla^{2}u = u_{rr} + \frac{1}{r}u_{r} + \frac{1}{r^{2}}u_{\theta\theta}
+\]
+Now we can treat this as a 2D problem; in other words, we can now use separation of variables to solve for $u$. As before, we know our solution will look like
+\[
+u(r,\theta) = R(r)T(\theta)
+\]
+Plugging everything in and realizing by the usual trick that everything s equal to a constant gives us
+\[
+-\frac{T''}{T} = \frac{R'' + \frac{R'}{r}}{\frac{R}{r^{2}}} = \lambda
+\]
+Since we are dealing in polar coordinates, we will go ahead and walk through the eigenvalues/functions.
+We will be solving for $T(\theta)$, since our boundary conditions are in terms of $\theta$.
+We also should note that the solutions will repeat themselves every $2\pi$, since we are in the polar system.
+In addition, our $u$ should be finite at $r=0$.
+Let's take a look at the eigenvalues for the Laplace equation.\\\\
+\begin{minipage}[t]{0.32\textwidth}
+\underline{$\lambda = 0$}\\
+When $\lambda = 0$, we have \[T(\theta) = a\theta + b.\] This also tells us that \[R(r) = c\ln{(r)} + d.\] Putting everything together gives us \[ u(r,\theta) = (a\theta + b)(c\ln{r} + d).\] Since the solution is periodic on $2\pi$, we know that $a=0$.
+Since the solution has to remain finite at $r=0$, we know that $c = 0$.
+This gives us that $u = bd$, or that $u$ is equal to a constant.
+\end{minipage}
+\hspace{0.01\textwidth}
+\begin{minipage}[t]{0.32\textwidth}
+\underline{$\lambda > 0$}\\
+With $\lambda > 0$, we have our characteristic equation $(m = \pm\ \sqrt{\lambda})$ yielding a real valued solution. This tells us that
+\[
+T(\theta) = c_{1}\cosh{(\sqrt{\lambda}\theta)} + c_{2}\sinh{(\sqrt{\lambda}\theta)}.
+\]
+The solution is periodic on $2\pi$, so we can say that $c_{1} = c_{2} = 0$. Therefore, we have a trivial solution.
+\end{minipage}
+\hspace{0.01\textwidth}
+\begin{minipage}[t]{0.32\textwidth}
+\underline{$\lambda < 0$}\\
+$\lambda < 0$ gives us 
+\[
+T(\theta) = c_{1}\cos{(\sqrt{\lambda}\theta)} + c_{2}\sin{(\sqrt{\lambda}\theta)}
+\]
+This solution is periodic with $2\pi$, and it yields a trivial eigenvector.
+\end{minipage}
+\noindent\\\\\\
+\noindent So we know that the only $\lambda$ that yields eigenvalues is $\lambda = 0$, and we know that this particular eigenvalue yields a constant solution.
+If we solve the differential equation for $R$, we will find that the only solution we care about is $r^{n}$.
+We can finally say that our solution will look like
+\[
+u(r,\theta) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}r^{n}[a_{n}\cos{(n\theta)} + b_{n}\sin{(n\theta)}]
+\]
+This solution is worth memorizing, as it will generally be the same for all variations of the Laplace equation similar to this one.
+Now we can apply the conditions for $\theta$.
+We have two conditions, but it is really just a piecewise function which we will call $f(\theta)$.
+We also know that our bound on the domain is $a$, or the length of the radius.
+This is enough information to find all the various variables in our $u(r,\theta)$.
+We have as follows:
+\[
+u(a,\theta) = f(\theta) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}a^{n}[a_{n}\cos{(n\theta)} + b_{n}\sin{(n\theta)}]
+\]
+Note that the $a$ which is the length of the radius is \textit{not} the same $a$ as $a_{n}$. Now we can solve for $a_{0}$ and $a_{n}$.
+\begin{alignat*}{3}
+a_{0} &= \frac{1}{\pi}\int_{0}^{2\pi}f(\theta)d\theta \qquad\qquad a_{n} &&= \frac{1}{\pi}\left[\int_{0}^{\pi}U_{1}\cos{(n\theta)}d\theta + \int_{\pi}^{2\pi}U_{2}\cos{(n\theta)}d\theta\right]\\
+&= U_{1} + U_{2} &&=0
+\end{alignat*}
+For $b_{n}$, we have
+\begin{align*}
+a^{n}b_{n} &= \frac{1}{\pi}\left[\int_{0}^{\pi}U_{1}\sin{(n\theta)}d\theta + \int_{\pi}^{2\pi}U_{2}\sin{(n\theta)}d\theta\right]\\
+&= \frac{1 - (-1)^{n}}{n\pi}(U_{1} - U_{2})\\
+b_{n} &= \frac{\left[1 - (-1)^{n}\right](U_{1} - U_{2})}{n\pi a^{n}}\\
+&= \frac{2(U_{1} - U_{2})}{(2k+1)\pi a^{2k+1}}
+\end{align*}
+Putting it all together gives us a final solution of
+\[
+u(r,\theta) = \frac{U_{1} + U_{2}}{2} + \sum_{k=0}^{\infty}r^{k}\frac{2(U_{1} - U_{2})}{(2k+1)\pi a^{2k+1}}
+\]
+And we are done.
+\newpage
+
+\indent Now let's look at another Laplace equation in polar coordinates.
+Again, we will be dealing with Dirichlet boundary conditions, so we will be able to jump straight into the solution.
+This one won't take us as long to solve, as we have already done a lot of the work.
+We are asked the following:\\\\
+\noindent \textbf{\textit{Ex:}} Solve the following PDE
+\[
+\begin{cases*}
+\nabla^{2}u = 0\\
+u(3,\theta) =
+\begin{cases*}
+1 \qquad\qquad 0 \leq \theta \leq \pi\\
+\sin^{2}{\theta}\qquad \pi \leq \theta \leq 2\pi 
+\end{cases*}
+\end{cases*}
+\]
+\indent \textbf{\textit{Solution:}} We already know what our separable equation will look like:
+\[
+\frac{r^{2}R'' + rR'}{R} = -\frac{T''}{T} = -\lambda
+\]
+This yields an eigenvalue of $\lambda = n^{2}$. We know this because our function $T$ must be periodic on $2\pi$, and as such we get the following piecewise function:
+\[
+T(\theta) =
+\begin{cases*}
+a_{0}\qquad\qquad\qquad\qquad\qquad\quad\ \text{for}\ n=0\\
+a_{n}\cos{(n\theta)} + b_{n}\sin{(n\theta)}\qquad \text{for}\ n > 0
+\end{cases*}
+\]
+The solution to our equation for $R$ yields
+\[
+R(r) = 
+\begin{cases*}
+c_{1} + c_{2}\ln{r} \qquad\qquad \text{for}\ n=0\\
+c_{1}r^{n} + c_{2}r^{-n}\qquad\quad \text{for}\ n > 0
+\end{cases*}
+\]
+We can disregard $c_{2}$ for this equation, since it would cause our solution to go to $\infty$ as $r \to 0$, at the center of the disk. Putting everything together gives us
+\begin{align*}
+u(r,\theta) &= R(r)T(\theta)\\
+&=
+\begin{cases*}
+a_{0}\qquad\qquad\qquad\qquad\qquad\qquad\quad\ \text{for}\ n=0\\
+a_{n}r^{n}\cos{(n\theta)} + b_{n}r^{n}\sin{(n\theta)}\qquad \text{for}\ n > 0
+\end{cases*}
+\end{align*}
+We should probably construct a series out of this, since that is what we usually do. The corresponding series is
+\[
+u(r,\theta) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}a_{n}r^{n}\cos{(n\theta)} + b_{n}r^{n}\sin{(n\theta)}
+\]
+And applying the boundary conditions gives us
+\[
+u(3,\theta) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}a_{n}3^{n}\cos{(n\theta)} + b_{n}3^{n}\sin{(n\theta)}
+\]
+For $a_{0}$, we have
+\begin{align*}
+a_{0} &= \frac{1}{\pi}\left[\int_{0}^{\pi}1 d\theta + \int_{\pi}^{2\pi}sin^{2}{\theta}d\theta\right]\\
+&= \frac{3}{2}
+\end{align*}
+Solving for $a_{n}$ gives us
+\begin{align*}
+a_{n} &= \frac{1}{3^{n}\pi}\left[\int_{0}^{\pi}\cos{(n\theta)}d\theta + \int_{\pi}^{2\pi}\cos{(n\theta)}\sin^{2}{\theta}d\theta\right]\\
+&= \frac{[n^{2} - 2 - 4\cos{(n\pi)}]\sin{(n\pi)}}{3^{n}n(n^{2}-4)\pi}\\
+&= 0
+\end{align*}
+Our $b_{n}$ term is
+\begin{align*}
+b_{n} &= \frac{1}{3^{n}\pi}\left[\int_{0}^{\pi}\sin{(n\theta)}d\theta + \int_{\pi}^{2\pi}\sin{(n\theta)}\sin^{2}{\theta}d\theta\right]\\
+&= \frac{2\left[n^{2} - 6 - 4\cos{(n\pi)}\right]\sin^{2}{\left(\frac{n\pi}{2}\right)}}{3^{n}n\pi(n^{2} - 4)}\\
+&= \frac{2\left[n^{2} - 6 - 4(-1)^{n}\right]\sin^{2}{\left(\frac{n\pi}{2}\right)}}{3^{n}n\pi(n^{2} - 4)}\\
+&=
+\begin{cases*}
+0 \qquad\qquad\qquad\qquad\quad\ \ n \in \{ 2k\ |\ k \in \mathbb{Z} \}\\
+\frac{1}{3^{n}\pi}\left[\frac{2}{n} + \frac{4}{n(n^{2} - 4)}\right]\qquad n \in \{ 2k + 1\ |\ k \in \mathbb{Z} \}
+\end{cases*}
+\end{align*}
+We essentially have our solution. All we have to do is combine everything:
+\[
+u(r,\theta) = \frac{3}{4} + \sum_{k=0}^{\infty}\frac{r^{2k + 1}}{3^{2k+1}\pi}\frac{2(4k^{2} + 4k -1)}{(2k + 3)(4k^{2} -1)}\sin{\left[(2k +1)\theta\right]}
+\]
+Let's take a look at some of the plots for this equation.\\
+\begin{center}
+\begin{tabular}{c c c}
+\includegraphics[scale=0.3]{lap_01_3} & \includegraphics[scale=0.3]{lap_01_30} & \includegraphics[scale=0.3]{lap_01_50}\\
+\textit{The first 3 terms} & \textit{The first 30 terms} & \textit{The first 50 terms}
+\end{tabular}
+\end{center}
+Plugging this infinite sum into mathematica tells us that it does not converge; and we are done.
 \end{document}

### 2015-10-29 pde/pdetest1.tex
added hear sheet

@@
 \usepackage{multicol}
+\usepackage{array}
 \topmargin=-0.3in
@@
 \hrule
-\noindent\\
+\noindent\\\\
+\indent We will begin by solving the typical first order PDE.
+The strategy here is to break it down into and ODE or several ODEs that can be solved by methods we already have.
+It does not matter whether the PDE is homogeneous or not; we will most likely be able to \\\\
 \noindent\textbf{\textit{Ex:}} Solve the following PDE
@@
 \hrule
-\noindent\\
+\noindent\\\\
 \indent We mentioned earlier the topic of classification of PDE's.
 Since we only have this and Fourier Series to talk about, we will go ahead and knock out classification.
-There's some information that we need to go ahead and put out here.
+There's some information that we need to go ahead and review here.
 This will be explained in the problem, but it's handy to have it for reference:
@@
 \noindent\\
+\subsection{Finding Fourier Series}
 \indent It's finally time to talk about Fourier Series.
@@
 \noindent And we are done.
-We haven't discussed the convergence of series yet, and it turns out that if the series converges, it will converge to the solution.
-We won't worry about it right now.
-Instead, let's look at another example
 
@@
 
+\subsection{Convergence of Fourier Series}
 \indent Up to this point, we have found a Fourier series and left our solutions at that. We will now look into the convergence of a Fourier series. This topic includes two questions:
@@
 \end{align*}
-Lets go ahead and formalize it:
-\[
-\nabla^{2}u = u_{rr} + \frac{1}{r}u_{r} + \frac{1}{r^{2}}u_{\theta\theta}
+Let's go ahead and formalize it:
+\[
+\nabla^{2}u = u_{rr} + \frac{1}{r}u_{r} + \frac{1}{r^{2}}u_{\theta\theta} = 0
 \]
@@
 \end{center}
-Plugging this infinite sum into mathematica tells us that it does not converge; and we are done.
+Plugging this infinite sum into mathematica tells us that it does not converge; the values are heading towards $\infty$ as the sum grows. We will leave our answer as a series, and we are done.
+
+\newpage
+\subsection{Wrapping Up}
+\indent Before we move onto the next topic, we will present some common patterns amongst the second order PDEs.
+Below, you will find each of the three equations in the classical trinity, and their corresponding solutions based on the boundary conditions.
+\subsubsection{The Heat Equation}
+\[
+\begin{array}{c c}
+\begin{cases}
+u_{t} - ku_{xx} = 0\\
+u(0,t) = u(L,t) = 0\\
+u(x,0) = f(x)
+\end{cases}
+&
+\begin{cases}
+u_{t} - ku_{xx} = 0\\
+u_{x}(0,t) = u_{x}(L,t) = 0\\
+u(x,0) = g(x)
+\end{cases}\\
+&\\
+u(x,t) = \sum_{n=1}^{\infty}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}\sin{\left(\frac{n\pi x}{L}\right)}
+& u(x,t) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}\cos{\left(\frac{n\pi x}{L}\right)}\\
+&\\
+\begin{cases}
+u_{t} - ku_{xx} = 0\\
+u(0,t) = u_{x}(L,t) = 0\\
+u(x,0) = h(x)
+\end{cases}
+&
+\begin{cases}
+u_{t} - ku_{xx} = 0\\
+u_{x}(0,t) = u(L,t) = 0\\
+u(x,0) = k(x)
+\end{cases}\\
+&\\
+u(x,t) =  \sum_{n=1}^{\infty}e^{-k\left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}\right)^{2}t}\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}
+& u(x,t) =  \sum_{n=1}^{\infty}e^{-k\left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}\right)^{2}t}\cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}
+\end{array}
+\]
+\noindent\\
+\subsubsection{The Wave Equation}
+\[
+\begin{array}{c c}
+\begin{cases}
+u_{tt} - c^{2}u_{xx} = 0\\
+u(0,t) = u(L,t) = 0\\
+u(x,0) = f(x) \quad u_{t}(x,0) = g(x)
+\end{cases}
+&
+\begin{cases}
+u_{tt} - c^{2}u_{xx} = 0\\
+u_{x}(0,t) = u_{x}(L,t) = 0\\
+u(x,0) = f(x) \quad u_{t}(x,0) = g(x)
+\end{cases}\\
+&\\
+u(x,t) = \sum_{n=1}^{\infty}\left[a_{n}\cos{\left(\frac{n \pi t c}{L}\right)} + b_{n}\sin{\left(\frac{n \pi t c}{L}\right)}\right]\sin{\left(\frac{n \pi x}{L}\right)}
+&u(x,t) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}\left[a_{n}\cos{\left(\frac{n \pi t c}{L}\right)} + b_{n}\sin{\left(\frac{n \pi t c}{L}\right)}\right]\cos{\left(\frac{n \pi x}{L}\right)}\\
+&\\
+\end{array}
+\]
+\[
+\begin{array}{c}
+\begin{cases}
+u_{tt} - c^{2}u_{xx} = 0\\
+u(0,t) = u_{x}(L,t) = 0\\
+u(x,0) = f(x) \quad u_{t}(x,0) = g(x)
+\end{cases}\\
+\\
+u(x,t) = a_{0} + b_{0}t + \sum_{n=1}^{\infty}\left[ a_{n}\cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi t}{L}\right)} + b_{n}\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}\right]\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}\\
+\\
+\begin{cases}
+u_{tt} - c^{2}u_{xx} = 0\\
+u_{x}(0,t) = u(L,t) = 0\\
+u(x,0) = f(x) \quad u_{t}(x,0) = g(x)
+\end{cases}\\
+\\
+u (x,t) = a_{0} + b_{0}t + \sum_{n=1}^{\infty}\left[ a_{n}\cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi t}{L}\right)} + b_{n}\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}\right]\cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}
+\end{array}
+\]
+\subsubsection{The Laplace Equation}
+\paragraph{Cartesian Coordinates}
+\[
+\begin{array}{c c}
+\begin{cases}
+u_{xx} + u_{yy} = 0\\
+u(0,y) = u(L,y) = 0
+\end{cases}
+&
+\begin{cases}
+u_{xx} + u_{yy} = 0\\
+u_{x}(0,y) = u_{x}(L,y) = 0
+\end{cases}\\
+&\\
+u(x,y) = \sum_{n=1}^{\infty}\left[a_{n}\sinh{\left(\frac{n \pi y}{L}\right)} + b_{n}\cosh{\left(\frac{n\pi y}{L}\right)}\right]\sin{\left(\frac{n\pi x}{L}\right)}
+& u(x,y) = \sum_{n=1}^{\infty}\left[a_{n}\sinh{\left(\frac{n \pi y}{L}\right)} + b_{n}\cosh{\left(\frac{n\pi y}{L}\right)}\right]\cos{\left(\frac{n\pi x}{L}\right)}
+\end{array}
+\]
+\\
+\[
+\begin{array}{c}
+\begin{cases}
+u_{xx} + u_{yy} = 0\\
+u_{x}(0,y) = u(L,y) = 0
+\end{cases}\\
+\\
+u(x,t) = \sum_{n=1}^{\infty}\left[a_{n}\sinh{\left(\left[n + \frac{1}{2}\right]\frac{\pi y}{L}\right)} + b_{n}\cosh{\left(\left[n + \frac{1}{2}\right]\frac{\pi y}{L}\right)}\right]\cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi y}{L}\right)}\\
+\\
+\begin{cases}
+u_{xx} + u_{yy} = 0\\
+u(0,y) = u_{x}(L,y) = 0
+\end{cases}\\
+\\
+u(x,t) = \sum_{n=1}^{\infty}\left[a_{n}\sinh{\left(\left[n + \frac{1}{2}\right]\frac{\pi y}{L}\right)} + b_{n}\cosh{\left(\left[n + \frac{1}{2}\right]\frac{\pi y}{L}\right)}\right]\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi y}{L}\right)}
+\end{array}
+\]
+\paragraph{Polar Coordinates}
+\begin{gather*}
+\begin{cases}
+u_{rr} + \frac{1}{r}u_{r} + \frac{1}{r^{2}}u_{\theta\theta} = 0\\
+u(h,\theta) = f(\theta)\quad \text{on}\ x < \theta < L
+\end{cases}\\
+u(r,\theta) = a_{0} + \sum_{n=1}^{\infty}a_{n}r^{n}\cos{\left(\frac{n \pi \theta}{L}\right)} + b_{n}r^{n}\sin{\left(\frac{n \pi \theta}{L}\right)}
+\end{gather*}
 \end{document}

### 2015-10-31 pde/pdetest1.tex
ode and paper

@@
 \end{gather*}
+\newpage
+
+\section{The Cauchy Problem}
+\hrule
+\noindent \\\\
+\indent Up until now, we have always been given our PDE's with boundary conditions and initial values; but imagine that we did not have a boundary?
+The traditional method that we have been using would fail at this point.
+Since we would not be able to find any eigenvalues, we would not be able to form a series solution by using the \textit{Separation of Variables} technique.
+Let's consider the familiar heat equation:
+\[
+\begin{cases}
+\xi_{t} - k\xi_{xx} = 0\\
+\xi(x,0) = f(x)
+\end{cases}
+\]
+Through much careful derivation and theory, we can say that the general solution for this problem will be
+\[
+\xi(x,t) = \frac{1}{\sqrt{4 \pi k t}}\int_{-\infty}^{\infty}e^{-\frac{(x-y)^{2}}{4kt}}f(y)dy.
+\]
+The question now should be ``How can we solve this?'' The first step we will take is letting
+\[
+u = \frac{y-x}{\sqrt{4kt}}.
+\]
+We are going to perform a \textit{u substitution} to try and put this integral into a nicer form. This yields
+\begin{align*}
+y &= x + u\sqrt{4kt}\\
+dy &= \sqrt{4kt}\ du.
+\end{align*}
+This changes our original problem into
+\[
+\xi(x,t) = \frac{1}{\sqrt{\pi}}\int_{-\infty}^{\infty}e^{-u^{2}}f(x+u\sqrt{4kt})du.
+\]
+Let's look at an example\\\\
+\noindent \textbf{\textit{Ex:}} Solve the following PDE
+\[
+\begin{cases}
+\mu_{t} - k\mu_{xx} = 0\\
+\mu(x,0) = \cos{x}
+\end{cases}
+\]
+\indent \textbf{\textit{Solution:}} From above, we know what our general solution will look like. All we need to do is substitute in our initial condition. We have the following:
+\begin{align*}
+\mu(x,t) &= \frac{1}{\sqrt{\pi}}\int_{-\infty}^{\infty}e^{-u^{2}}\cos{(x + u\sqrt{4kt})}du\\
+&= \frac{1}{\sqrt{\pi}}\int_{-\infty}^{\infty}e^{-u^{2}}\left[\cos{(x)}\cos{(u\sqrt{4kt})} - \sin{(x)}\sin{(u\sqrt{4kt})}\right]du\\
+&=  \frac{2}{\sqrt{\pi}}\cos{(x)}\int_{0}^{\infty}e^{-u^{2}}\cos{(u\sqrt{4kt})}du.
+\end{align*}
+You may be wondering what happened to the $\sin$ terms.
+Recall that $\sin$ is an odd function, so it will ultimately be $0$ on the interval $[-\infty,\infty]$; because of this fact, we can remove it now.
+We also use the face that the $\cos$ is even, so it is okay to halve the interval and multiply by two.
+Now we need to evaluate this integral.
+We will substitute $\alpha = \sqrt{4kt}$, and assign a function, $\Upsilon(\alpha)$, to the integral.
+This gives us the following system:
+\[
+\begin{cases}
+\Upsilon(\alpha) = \int_{0}^{\infty}e^{-u^{2}}\cos{(\alpha u)}du\\
+\Upsilon(0) = \int_{0}^{\infty}e^{-u^{2}}du\\
+\end{cases}
+\]
+Believe it or not, we actually know the answer to $\Upsilon(0)$. To find it, we can assign a function to the integral in a similar fashion as we have already done. Calling it $\Psi$, we have
+\begin{align*}
+\Psi &= \int_{0}^{\infty}e^{-u^{2}}du\\
+\Psi^{2} &= \int_{0}^{\infty}e^{-u^{2}}du \int_{0}^{\infty}e^{-v^{2}}dv\\
+\Psi^{2} &= \int_{0}^{\infty}\int_{0}^{\infty}e^{-u^{2}-v^{2}}dudv\\
+\Psi^{2} &= \int_{0}^{\pi/2}\int_{0}^{\infty}re^{-r^{2}}drd\theta\\
+\Psi^{2} &= \frac{\pi}{2}\left[-\frac{1}{2}e^{-r^{2}}\right]\Bigg |_{r=0}^{r=\infty}\\
+\Psi^{2} &= \frac{\pi}{4}\\
+\Psi &= \frac{\sqrt{\pi}}{2}.
+\end{align*}
+So we know that $\Upsilon(0) = \frac{\sqrt{\pi}}{2}$. Plugging this into our system gives us
+\[
+\begin{cases}
+\Upsilon(\alpha) = \int_{0}^{\infty}e^{-u^{2}}\cos{(\alpha u)}du\\
+\Upsilon(0) = \frac{\sqrt{\pi}}{2}\\
+\end{cases}
+\]
+Now, we will take the derivative of $\Upsilon$. This gives us
+\[
+\Upsilon'(\alpha) = -\int_{0}^{\infty}ue^{-u^{2}}\sin{(\alpha u)}du.
+\]
+Integration by parts gives us
+\[
+\frac{1}{2}e^{-u^{2}}\sin{(\alpha u)}\Big |_{0}^{\infty} - \alpha\int_{0}^{\infty}\frac{1}{2}e^{-u^{2}}\cos{(\alpha u)}du.
+\]
+The value from $0$ to $\infty$ is zero, since the $\sin$ function odd. This leaves us with following relationship:
+\[
+\Upsilon'(\alpha) = -\frac{\alpha}{2}\Upsilon(\alpha) \alpha.
+\]
+We can solve this as an ODE. It's separable, so we get
+\begin{align*}
+\frac{d\Upsilon}{\Upsilon} &= -\frac{\alpha}{2}d\alpha\\
+\ln{(\Upsilon)} &= -\frac{\alpha^{2}}{4} + c_{1}\\
+\Upsilon &= c_{2}e^{-\frac{\alpha^{2}}{4}}.
+\end{align*}
+Applying the initial condition gives us
+\[
+\Upsilon = \frac{\sqrt{\pi}}{2}e^{-\frac{\alpha^{2}}{4}}.
+\]
+Substituting back in gives us a final answer of
+\begin{align*}
+\mu(x,t) &= \frac{2}{\sqrt{\pi}}\cos{x}\Upsilon(\sqrt{4kt})\\
+&= \cos{x}e^{-kt},
+\end{align*}
+and we are done.
 \end{document}

### 2015-10-31 wim/paper.tex
ode and paper

\documentclass{article}
\usepackage{setspace}
\usepackage{amssymb,amsmath}
\usepackage{mathtools}
\usepackage[margin=1in]{geometry}
\doublespacing
\title{The Count Distinct Problem}
\author{Steven Rosendahl}
\date{}
\begin{document}
\maketitle
\indent Today's modern world is completely connected; nearly everyone has access to the internet at some point 
during the day. This constant connection to the internet means that data is constantly being transferred and 
collected. Companies such as Google, Facebook, and Twitter often collect the data that they
\\\\
\indent The problem of processing data has not always been a problem. Before the internet was widely used by
the population, storing data was expensive, and as a result most companies did not want to spend money on the 
the hardware and space that would be required to store large amounts of data. In addition, data was not 
transferred and gathered at a fast enough rate to warrant finding a way to count the data; even if companies wanted to count the data, the computers at the time were not powerful enough to perform the necessary calculations. The problem of counting big data did not arise until the 1990's, when the internet began to facilitate the easy transfer and collection of data. Today, more than 25 exabytes of data are created a day, and out of that data, nearly 50\% of it is stored for later use. The main factor that contributes most to the problem is that gathering and storing data is cheap, but analyzing it is not. The volume of the data is what leads to the count distinct problem that we will analyze.
\\\\
\indent The first solution to the count distinct problem is called the \textit{Hash Table}. The algorithm is as follows:
\begin{center}
Let $\mathbb{S}$ be a set of random elements. In order to count the number of distinct elements in $\mathbb{S}$,
\begin{enumerate}
\item Apply a bijection, $h(n)$ to all $n\in\mathbb{S}$, and store the result in a set $\mathbb{V}$.
\item Count the number of elements in $\mathbb{V}$.
\end{enumerate}
\end{center}
Since $\mathbb{V}$ is a bijection, it is surjective. As a result, if we take any $x\in\mathbb{S}\ \text{and}\ y\in\mathbb{S}$, then if $x = y$, they will hash to the same value. For example, Say we want to count the number of distinct names in the set $\mathbb{S} = \{\text{``Alice''},\text{``Emily''},\text{``Alice''}\}$. We can create a simple hash function where we map each letter to its corresponding value in the alphabet, sum up each value in a name, and mod it with the number of letters in the name. Applying the function on each element of $\mathbb{S}$ yields a set $\mathbb{V} = \{0,4\}$. $||\mathbb{V}|| = 2$, so there were two distinct elements in $\mathbb{S}$. In this example, the hash function we chose was bijective. However, this may not always be the case.
\\\\
\indent Imagine that we are solving the twitter problem where we want to count the number of distinct hashtags on twitter on a given day. Our set $\mathbb{S}$ will now contain every hashtag that was made on a given day. We know that we want a hash function that is surjective, but this is an unrealistic, since we could have some $x \in \mathbb{S}$ and $y\in\mathbb{S}$ where $x \neq y$ but $h(x) = h(y)$. To model this issue, we can consider the following scenario:
\begin{quote}
On October 30th, 2015, Mojang released a special Halloween cape to all players of the popular video game Minecraft. As a result, ``\#cape'' began trending on Twitter. On the same day, Target announced that all their \textit{Bob's Burgers} costumes would be on sale for the low price of \$12.95, and ``\#bob'' began trending on twitter.
\end{quote}
Our hash function from earlier will not work in this case, since both ``\#bob'' and ``\#cape'' hash to the same value. On a larger scale, it is easy to see that we would need to either handle collisions or create a better hash function. Creating a way to handle collisions would not benefit us in this case, since the data is being processed in real time. We would not be able to tell if we were hashing a value we already have seen or hashing a new value that happens to hash to the same value as another element we have already seen. Since the collision policy would not help us, we need to come up with a better hashing function.
\\\\
\indent More complex hash function, such as SHA-2 (Secure Hash Algorithm 2) or MD5 (Message Digest 5) offer a solution to the problems at which we are looking. SHA-2 is used by the NSA to encrypt data, and has yet to produce any collisions when hashing values. However, this is not a viable solution to the count distinct problem either. The problem arises from the volume of the data being processed. In both the Twitter and Facebook problems, we are analyzing exabytes of data. The SHA-2 function outputs a 512 byte long integer for every input, which means that the resulting size of the set $\mathbb{V}$ would be in the worst case $512 \times ||\mathbb{S}||$. All of $\mathbb{V}$ would have to be held in memory, which means that analyzing 4 billion hashtags (which is a generously low estimate) could possibly result in $(4 \times 10^{13}) \times 512$ bytes of data being stored at one time, assuming every hashtag was 10 bytes long. MD5 outputs smaller sized integers (128 in this case), but the collision probability is much higher. The hash method is not a viable solution to the count distinct problem when we are dealing with a large volume of data. Hashing, however is used in the other solutions that we will discuss.
\\\\
\indent The most beneficial aspect of the hash is that it has a 0\% error rate. In other words, the cardinality of $\mathbb{V}$ will be exactly how many distinct elements are in $\mathbb{S}$, assuming the hash function is surjective. However, if we are looking for a \textit{good enough} estimate of the number of distinct elements in $\mathbb{S}$, we can use an algorithm called a Linear Probabilistic Counter. With the LPC, the user can specify how large they want the error to be. This method uses a data structure called a bit-map. Bit-maps are a matrix where either a 1 or 0 is stored in a cell, which allows for the map to be incredibly space efficient. The LPC also uses a hash, but it only needs to temporarily produce a position and then it can ``forget'' the value it just produced. The number of distinct elements is determined by the number of 1's in the bit-map.
\\\\
\indent The LPC produces an estimate of the number of elements in the 
\end{document}

### 2015-11-04 wim/paper.tex
working on final paper

@@
 \usepackage{amssymb,amsmath}
+\usepackage{url}
+\usepackage{fancyvrb}
 \usepackage{mathtools}
+\usepackage[square,sort,comma,numbers]{natbib}
 \usepackage[margin=1in]{geometry}
@@
 \maketitle
-\indent Today's modern world is completely connected; nearly everyone has access to the internet at some point 
-during the day. This constant connection to the internet means that data is constantly being transferred and 
-collected. Companies such as Google, Facebook, and Twitter often collect the data that they
+\indent Processing data efficiently has always been a goal of organizations that collect data; companies like Facebook and Google have been collecting and storing data for as long as they have existed.
+The count-distinct problem poses the question of how many distinct elements are in a given set. 
+The three problems that we will analyze are
+\begin{enumerate}
+\item \underline{The Twitter Problem}: How many unique hashtags are created on any given day?
+\item \underline{The Facebook Problem}: How many unique Facebook app installs are made on any given day?
+\item \underline{The Pok\'emon Problem}: How many unique Pok\'emon encounters will a player face in a given play through of a game from each generation.
+\end{enumerate}
+Each of these problems can be solved by using one of three different solutions to the count distinct problem: the hash table, the linear probabilistic counter, and the hyperloglog algorithm.
 \\\\
-\indent The problem of processing data has not always been a problem. Before the internet was widely used by
-the population, storing data was expensive, and as a result most companies did not want to spend money on the 
-the hardware and space that would be required to store large amounts of data. In addition, data was not 
-transferred and gathered at a fast enough rate to warrant finding a way to count the data; even if companies wanted to count the data, the computers at the time were not powerful enough to perform the necessary calculations.
+\indent The count-distinct problem has not always been a problem. 
+Before the internet was widely used by the population, storing data was expensive, and as a result most companies did not want to spend money on the hardware and space that would be required to store large amounts of data.
+In addition, data was not transferred and gathered at a fast enough rate to warrant finding a way to count the data; even if companies wanted to count the data, the computers at the time were not powerful enough to perform the necessary calculations.
 The problem of counting big data did not arise until the 1990's, when the internet began to facilitate the easy transfer and collection of data. 
-Today, more than 25 exabytes of data are created a day, and out of that data, nearly 50\% of it is stored for later use.
+Today, more than 25 exabytes of data are created and stored a day. 
 The main factor that contributes most to the problem is that gathering and storing data is cheap, but analyzing it is not.
-The volume of the data is what leads to the count distinct problem that we will analyze.
+The volume of the data leads to the count distinct problem that we will analyze \cite[pp.
+1]{Yousra}.
 \\\\
-\indent The first solution to the count distinct problem is called the \textit{Hash Table}. The algorithm is as follows:
+\indent The first solution to the count distinct problem is called the \textit{Hash Table}. 
+A hash-table is a matrix that is stored in memory where each entry was uniquely obtained from another value. 
+The algorithm is as follows:
 \begin{center}
@@
 \item Apply a bijection, $h(n)$ to all $n\in\mathbb{S}$, and store the result in a set $\mathbb{V}$.
-\item Count the number of elements in $\mathbb{V}$.
+\item Count the number of elements in $\mathbb{V}$. $||\mathbb{V}||$ will be the number of unique elements in $\mathbb{S}$.
 \end{enumerate}
@@
 Since $\mathbb{V}$ is a bijection, it is surjective. 
-As a result, if we take any $x\in\mathbb{S}\ \text{and}\ y\in\mathbb{S}$, then if $x = y$, they will hash to the same value.
+As a result, if we take any $x\in\mathbb{S}\ \text{and}\ y\in\mathbb{S}$, then if $x = y$, they will hash to the same value, and we can ignore it \cite[p. 6]{Maurer}. 
 For example, Say we want to count the number of distinct names in the set $\mathbb{S} = \{\text{``Alice''},\text{``Emily''},\text{``Alice''}\}$. 
-We can create a simple hash function where we map each letter to its corresponding value in the alphabet, sum up each value in a name, and mod it with the number of letters in the name.
+We can create a simple hash function where we map each letter to its corresponding value in the alphabet, sum up total of all the letter values in a name, and mod that value with the number of letters in the name.
 Applying the function on each element of $\mathbb{S}$ yields a set $\mathbb{V} = \{0,4\}$.
@@
 In this example, the hash function we chose was bijective.
-However, this may not always be the case.
+However, this may not always be the case. In fact, it is practically impossible to create a hash function that is truly surjective for huge amounts of data \cite[p. 6]{Maurer}. 
+For smaller sets of data, however, we can create a hash function that is surjective. 
+The pok\'emon problem provides a good example of the benefits of a hash table.
 \\\\
-\indent Imagine that we are solving the twitter problem where we want to count the number of distinct hashtags on twitter on a given day.
+\indent The most beneficial aspect of the hash is that it has a 0\% error rate. 
+In other words, the cardinality of $\mathbb{V}$ will be exactly how many distinct elements are in $\mathbb{S}$, assuming the hash function is surjective. 
+We know that for very large $\mathbb{S}$, the hash table method is incredibly inefficient, but for the Pok\'emon problem, we are dealing with $\mathbb{S}$ small enough to apply the hash table. 
+In this particular case, 721 pok\'emon can be encountered \cite{Pokemon}, which means that the maximum size needed for our set $\mathbb{V}$ is 721. 
+Typically, a player will encounter in the neighborhood of 1000 wild pok\'emon a game, and with a total of 6 generations of games, we have a set $\mathbb{S}$ of size 6000. 
+The hashing function we created earlier will not be good enough here; instead we can create a new hashing function where we sum up the value of the name as before, then add the pok\'emon's unique id, and finally mod that value with 721.
+The following Java code can be used to model our hash table:\\\\
+\begin{tabular}{c c} 
+\begin{BVerbatim}
+public int getHashValue(){
+    char[] array = nameOfPokemon.toCharArray();
+    int total = 0;
+    for(char c : array){
+	total += (int)c;
+    }
+    total += type;
+    total %= 721;
+    return total;
+}
+\end{BVerbatim}
+&
+\begin{BVerbatim}
+...
+for(Pokemon pm : S){
+    int hashValue = pm.getHashValue();
+    storeHashValue();
+}
+calculateAndPrintUniqueValues();
+\end{BVerbatim}
+\end{tabular}
+Running this code will give us a final answer of 458 unique encounters with wild pok\'emon. 
+In the case of the pok\'emon problem, a simple hash function sufficed, since the only collisions that were encountered were pok\'emon where the name was the same. 
+If we consider a larger data set, however, we may not be able to create a surjective hash function.
+\\\\
+\indent Imagine that we are solving the Twitter problem where we want to count the number of distinct hashtags on twitter on a given day. 
 Our set $\mathbb{S}$ will now contain every hashtag that was made on a given day. 
@@
 Our hash function from earlier will not work in this case, since both ``\#bob'' and ``\#cape'' hash to the same value. 
-On a larger scale, it is easy to see that we would need to either handle collisions or create a better hash function.
-Creating a way to handle collisions would not benefit us in this case, since the data is being processed in real time.
-We would not be able to tell if we were hashing a value we already have seen or hashing a new value that happens to hash to the same value as another element we have already seen.
+On a larger scale, such as the Facebook or Twitter problem, we would need to either handle collisions or create a better hash function. 
+Creating a way to handle collisions would not be beneficial, since the data is being processed in real time. 
+We would not be able to tell if we were hashing a value we already have seen or hashing a new value that happens to hash to the same value as another element we have already seen; as a result we would have no way to know which keys were different and which keys were the same when looking at $\mathbb{V}$ \cite[p.
+6]{Maurer}.
 Since the collision policy would not help us, we need to come up with a better hashing function. 
 \\\\
-\indent More complex hash function, such as SHA-2 (Secure Hash Algorithm 2) or MD5 (Message Digest 5) offer a solution to the problems at which we are looking.
-SHA-2 is used by the NSA to encrypt data, and has yet to produce any collisions when hashing values.
-However, this is not a viable solution to the count distinct problem either.
-The problem arises from the volume of the data being processed.
+\indent More complex hash functions, such as SHA-2 (Secure Hash Algorithm 2) or MD5 (Message Digest 5) offer a solution to the collision problem.
+SHA-2 is used by organizations such as the NSA to encrypt data, and has yet to produce any collisions when hashing values \cite[p.
+301]{SHA-Algo}.
+However, this is not a viable solution to the count distinct problem either due to the volume of the data being processed.
 In both the Twitter and Facebook problems, we are analyzing exabytes of data.
 The SHA-2 function outputs a 512 byte long integer for every input, which means that the resulting size of the set $\mathbb{V}$ would be in the worst case $512 \times ||\mathbb{S}||$.
-All of $\mathbb{V}$ would have to be held in memory, which means that analyzing 4 billion hashtags (which is a generously low estimate) could possibly result in $(4 \times 10^{13}) \times 512$ bytes of data being stored at one time, assuming every hashtag was 10 bytes long.
-MD5 outputs smaller sized integers (128 in this case), but the collision probability is much higher.
-The hash method is not a viable solution to the count distinct problem when we are dealing with a large volume of data.
-Hashing, however is used in the other solutions that we will discuss.
+All of $\mathbb{V}$ would have to be held in memory, which means that analyzing 50 million hashtags (which is a low estimate)\cite{Twitter} could possibly result in $(50 \times 10^{10}) \times 512$ bytes of data being stored at one time, assuming every hashtag was 10 bytes long.
+A hash table of this size could not be loaded all at once into main memory; it would have to be stored into secondary memory which would exponentially increase the amount of time required to read the data \cite[p.
+209]{Whang}.
+MD5 outputs smaller sized integers (128 in this case), but the collision probability is much higher \cite[p.
+22 - 23]{Break-MD5}.
+The hash method is not a viable solution to the count distinct problem when we are dealing with a large volume of data, but hash functions are still used in the other solutions to the count-distinct problem.
 \\\\
-\indent The most beneficial aspect of the hash is that it has a 0\% error rate.
-In other words, the cardinality of $\mathbb{V}$ will be exactly how many distinct elements are in $\mathbb{S}$, assuming the hash function is surjective.
-However, if we are looking for a \textit{good enough} estimate of the number of distinct elements in $\mathbb{S}$, we can use an algorithm called a Linear Probabilistic Counter.
+\indent The hash-table method provided a 0\% error, but was not efficient in handling large sets of data.
+However, if we are looking for a \textit{good enough} estimate (within the neighborhood of 1\% to 5\% error) of the number of distinct elements in $\mathbb{S}$, we can use an algorithm called a Linear Probabilistic Counter.
 With the LPC, the user can specify how large they want the error to be.
@@
 Bit-maps are a matrix where either a 1 or 0 is stored in a cell, which allows for the map to be incredibly space efficient.
-The LPC also uses a hash, but it only needs to temporarily produce a position and then it can ``forget'' the value it just produced.
-The number of distinct elements is determined by the number of 1's in the bit-map.
+The LPC also uses a hash, but it only needs to temporarily produce a position and then it can ``forget'' the value it just produced, which in turn reduces the overhead memory needed to store the elements that was required by the hash table method \cite{Whang}.
+The number of distinct elements is determined by the number of 1's in the bit-map when the algorithm is finished.
 \\\\
-\indent The LPC produces an estimate of the number of elements in the 
+\indent In order to either increase or decrease the error, the user of the LPC algorithm can either increase or decrease the size of the bit-map that is held in main memory.
+The following equation gives us a relationship between the size of the bitmap and the error that will be produced:
+\[
+bias\left(\frac{\hat{n}}{n}\right) = \frac{e^{t} - t - 1}{2n},
+\]
+where $bias\left(\frac{\hat{n}}{n}\right)$ is the error.
+The value $t$ is the ratio of the number of distinct elements determined by the algorithm (denoted by $n$) to the size of the bitmap \cite{Whang}.
+If $t$ is close to 0, then the error will be very small; if $t$ is much greater than 1, the error will be very large.
+In other words, if the hash function being used produces a large amount of collisions, then the error will be high.
+Luckily, an algorithm such as SHA-2 will be okay to use, since the LPC never stores the hashed value in main memory.
+We can also calculate the number of distinct elements by
+\[
+\hat{n} = -m \ln{(\Lambda_{n})},
+\]
+where $m$ is the size of the bit-map and $\Lambda_{n}$ is the ratio of the number of 0's in the bit map to the size of the bit map \cite[p. 212]{Whang}. 
+\\\\
+\indent The Facebook problem can be solved efficiently by using the LPC algorithm.
+We would have set $\mathbb{S}$ of size 20 million \cite{Facebook}, and we can choose an arbitrary sized bit map $\mathbb{V}$ of 10 million bits, or roughly 1.25 Mb, which lies on the upper limit of the efficiency of the LPC.
+If the memory required for the bit map increases beyond 1.25Mb, most operating systems will move the I/O operations to secondary memory, which defeats the purpose of the LPC algorithm \cite[p.
+211]{Whang}; as a result, the Twitter problem would not be efficiently solved by this method, since the amount of hashtags created in a day far exceeds 20 million.
+In the case of the Facebook problem, we can apply the LPC algorithm analytically, rather than actually processing 20 million application installs.
+\\\\
+%%%%%%%%FINISH THIS SECTION%%%%%%%%%
+\indent We know that on a given day, Facebook users make 20 million app installs, and that there are currently about 7 million apps on Facebook \cite{Facebook}. 
+However, not all 7 million applications are installed every day; in fact, the amount of installs decays exponentially starting from the number one most installed application on Facebook, which is Spotify as of August 2015 \cite{Facebook-2}.
+We would expect that apps such as Spotify would dominate the market, whereas apps in the lower spectrum of the downloads would be negligible.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\\\\
+\indent The LPC and the hash table algorithms handled relatively small data sets well, but the Twitter problem deals with a data set of nearly 200,000,000 tweets a day \cite{Twitter-2}.
+As a result, the LPC would be too inefficient to process all the data, but the HyperLogLog algorithm allows for data to be processed more efficiently than the LPC or the hash table. 
+The HLL algorithm actually uses parts of the hash table and LPC algorithms; values are still hashed to a bit map held in memory. 
+The hash function hashes values to binary numbers, and the algorithm decides whether or not that value has already been seen the value before by analyzing the leading 0's in the binary number \cite[pp.
+685, 689]{Heule}.
+In other words, the algorithm keeps a record of the first location of a 1 in a binary string, called $\rho$ \cite[pp. 130]{Flaj}.
+If the sequence of 0's is identical to another sequence of 0's, the probability that the original values were the same is very high, and the algorithm will ignore what it finds to be a duplicate value.
+The amount of memory that the HLL algorithm requires per bitmap is expressed by
+\[
+\text{Memory Required}\, = \log_{2}{\left(\log_{2}{\left(M \right)}\right)},
+\]
+where $M$ is the size of the original set of data \cite[pp. 129]{Flaj}.
+Finally, the algorithm takes the harmonic average of all the totals of the separate bit maps, which allows the algorithm to increase in accuracy as the size of the original set grows \cite{Yousra}.
+\\\\
+\indent We can solve the Twitter problem with the HLL algorithm. 
+We will make the assumption that all 200,000,000 tweets made will have hashtags, which means we will have a set $\mathbb{S}$ of size 200,000,000, assuming every tweet made in a day contains a hashtag.
+By the formula for above, we get that the total required memory per bit map will be
+\begin{align*}
+\text{Memory} &=\log_{2}{\left(\log_{2}{\left( 200,000,000\right)}\right)}\\
+&\approx  4.78531\text{b},
+\end{align*}
+The LPC algorithm would take at least ten times the amount of space just to hold the bit map in main memory.
+In addition to the memory efficiency of the HLL, the algorithm also has a relatively smaller error than the LPC, expressed by
+\[
+\text{Error}\ =\frac{\sqrt{3\log{(2)} - 1}}{\sqrt{m}}.
+\]
+Finally, the HyperLogLog finds a harmonic mean of the hashed values in the bit map, rather than a geometric mean, which increases the accuracy of the HLL even more. 
+\\\\
+To solve the Twitter problem,
+
+\newpage
+\begin{thebibliography}{9}
+\bibitem{Facebook} Facebook Statistics. (2015, September 20). Retrieved October 30, 2015, from \url{http://www.statisticbrain.com/facebook-statistics/}
+\bibitem{Whang} Whang, Kyu-Young, Vander-Zanden, Brad T. \& Taylor, Howard M. (1990). A Linear-time Probabilistic Counting Algorithm for Database Applications. ACM Trans. Database Syst., 15, 208-229.
+\bibitem{Maurer} Maurer, W. D. \& Lewis, T. G. (1975). Hash Table Methods. ACM Comput. Surv., 7, 5-19.
+\bibitem{SHA-Algo} Chaves, Ricardo, Kuzmanov, Georgi, Sousa, Leonel \& Vassiliadis, Stamatis (2006). Improving SHA-2 Hardware Implementations. , 4249, 298-310.
+\bibitem{Break-MD5} Wang, Xiaoyun \& Yu, Hongbo (2005). How to Break MD5 and Other Hash Functions. , 3494, 19-35.
+\bibitem{Heule} Heule, S., Nunkesser, M., \& Hall, A.
+(2013).
+HyperLogLog in Practice: Algorithmic Engineering of a State of the Art Cardinality Estimation Algorithm.
+In \textit{Proceedings of the 16th International Conference on Extending Database Technology} (Vol.
+EDBT '13, pp.
+683-692).
+New York, NY, Genoa: ACM.
+\bibitem{Twitter} \#numbers. (2011, March 14). Retrieved October 29, 2015, from \url{https://blog.twitter.com/2011/numbers}
+\bibitem{Pokemon} How many Pokemon are there altogether, before and after Omega Ruby and Alpha Sapphire?
+(2014, October 27).
+Retrieved October 29, 2015, from \url{http://pokemondb.net/pokebase/219332/pokemon-there-altogether-before-after-omega-alpha-sapphire}
+\bibitem{Facebook-2} Most popular Facebook apps 2015 | Statistic.
+(2015, August 1).
+Retrieved November 3, 2015, from \url{http://www.statista.com/statistics/276371/most-popular-facebook-apps-by-monthly-active-users/}
+\bibitem{Twitter-2} Twitter Statistics. (2015, September 15). Retrieved November 3, 2015, from \url{http://www.statisticbrain.com/twitter-statistics/}
+\bibitem{Yousra}Yousra Chabchoub, Georges H\'ebrail. \textit{Sliding HyperLogLog: Estimating cardinality in a data stream.} 2010. 
+\bibitem{Flaj} Flajolet, P., Fusy, \'E., Gandouet, O., \& Meunier, F.
+(n.d.).
+HyperLogLog: The Analysis of a Near-Optimal Cardinality Estimation Algorithm.
+2007 \textit{Conference on Analysis of Algorithms, 07,} 127-143.
+Retrieved October 27, 2015, from \url{http://algo.inria.fr/flajolet/Publications/FlFuGaMe07.pdf}
+\end{thebibliography}
 \end{document}

### 2015-11-04 wim/pkmn_problem/Hasher.java

@@ -1,0 +1,77 @@
+import java.util.*;
+import java.io.*;
+
+public class Hasher {
+	static Pokemon[] pkmn = new Pokemon[6000];
+	static int[] hvs = new int[721];
+
+	private static void readInPokemon(String filename){
+		try {
+			Scanner in = new Scanner(new File(filename));
+			Scanner lineReader;
+			List<Pokemon> tmp = new ArrayList<>();
+			while(in.hasNextLine()){
+				lineReader = new Scanner(in.nextLine()).useDelimiter(",");
+				lineReader.next();
+				Pokemon adder = new Pokemon(lineReader.next(), lineReader.nextInt());
+				tmp.add(adder);
+			}
+			for(int i = 0; i < 6000; i++){
+				Pokemon p = tmp.get((int)(Math.random() * 712));
+				// System.out.println("Pokemon: " + p + " Hash Value: " + p.getHashValue());
+				pkmn[i] = p;
+			}
+		} catch(Exception ex){
+			ex.printStackTrace();
+		}
+	}
+
+	public static void main(String[] args){
+		for(int i = 0; i < 712; i++){
+			hvs[i] = -1;
+		}
+		readInPokemon("pokemon_list.txt");
+		int collisions = 0;
+		for(Pokemon pm : pkmn){
+			int hv = pm.getHashValue();
+			if(hvs[hv] == 0){
+				System.out.println("Collision! HashValue: " + hv);
+				collisions++;
+			}
+			hvs[pm.getHashValue()] = 0;
+		}
+		int unique = 0;
+		for(int i : hvs){
+			if (i == 0){
+				unique++;
+			}
+		}
+		System.out.println("Total # of collisions: " + collisions);
+		System.out.println("The total is " + unique);
+	}
+
+	private static class Pokemon{
+		private String name;
+		private int type;
+
+		public Pokemon(String n, int t){
+			name = n;
+			type = t;
+		}
+
+		public int getHashValue(){
+			char[] array = name.toCharArray();
+			int total = 0;
+			for(char c : array){
+				total += (int)c;
+			}
+			total += type;
+			total %= 721;
+			return total;
+		}
+
+		public String toString(){
+			return name;
+		}
+	}
+}

### 2015-11-04 wim/twitter_problem/twitter_stream.rb

@@ -1,0 +1,21 @@
+require 'twitter'
+client = Twitter::REST::Client.new do |config|
+  config.consumer_key        = "JtwwG40EW2O6CKBNp1qeD57IH"
+  config.consumer_secret     = "NUSZ0jPyeWwTKPAlUDYLnoeatyDCsPp5waIJsYnQM6LJ5oDjDJ"
+  config.access_token        = "3494713935-hoqlE0ocEarpGUbSosIEi3YJvHqrJH3j60x7Z74"
+  config.access_token_secret = "fKTi6UiCESwluB9UbBOsdcBVjHHPfUOxr6PbSWy4UCUi8"
+end
+
+while true do
+	open('twitter_data.txt','a') do |f|
+		client.search("#", result_type: 'recent', lang: 'en').take(1000).each do |tweet|
+			f.puts tweet.text
+		end
+	end
+	str = "Gathered Sample at " + Time.now.inspect
+	open('tw_log.txt','a') do |file|
+		file.puts str
+		puts str
+	end
+	sleep 1200
+end

### 2015-11-04 wim/twitter_problem/const.rb
what

@@ -1,0 +1,6 @@
+class Constants
+	@@oh_one = "JtwwG40EW2O6CKBNp1qeD57IH"
+	@@oh_two = "NUSZ0jPyeWwTKPAlUDYLnoeatyDCsPp5waIJsYnQM6LJ5oDjDJ"
+	@@oh_three = "fKTi6UiCESwluB9UbBOsdcBVjHHPfUOxr6PbSWy4UCUi8"
+	@@oh_four = "3494713935-hoqlE0ocEarpGUbSosIEi3YJvHqrJH3j60x7Z74"
+end

### 2015-11-04 wim/twitter_problem/const.rb
fixing things one thing at a time

@@ -1,6 +1,0 @@
-class Constants
-	@@oh_one = "JtwwG40EW2O6CKBNp1qeD57IH"
-	@@oh_two = "NUSZ0jPyeWwTKPAlUDYLnoeatyDCsPp5waIJsYnQM6LJ5oDjDJ"
-	@@oh_three = "fKTi6UiCESwluB9UbBOsdcBVjHHPfUOxr6PbSWy4UCUi8"
-	@@oh_four = "3494713935-hoqlE0ocEarpGUbSosIEi3YJvHqrJH3j60x7Z74"
-end

### 2015-11-04 wim/twitter_problem/twitter_stream.rb

@@ -1,9 +1,10 @@
 require 'twitter'
+require_relative 'const.rb'
 client = Twitter::REST::Client.new do |config|
-  config.consumer_key        = "JtwwG40EW2O6CKBNp1qeD57IH"
-  config.consumer_secret     = "NUSZ0jPyeWwTKPAlUDYLnoeatyDCsPp5waIJsYnQM6LJ5oDjDJ"
-  config.access_token        = "3494713935-hoqlE0ocEarpGUbSosIEi3YJvHqrJH3j60x7Z74"
-  config.access_token_secret = "fKTi6UiCESwluB9UbBOsdcBVjHHPfUOxr6PbSWy4UCUi8"
+  config.consumer_key        = Constants.oh_one
+  config.consumer_secret     = Constants.oh_two
+  config.access_token        = Constants.oh_three
+  config.access_token_secret = Constants.oh_four
 end
 
 while true do
@@ -12,10 +13,10 @@ while true do
 			f.puts tweet.text
 		end
 	end
-	str = "Gathered Sample at " + Time.now.inspect
+	str = "Gathered Sample at #{Time.now.inspect}"
 	open('tw_log.txt','a') do |file|
 		file.puts str
 		puts str
 	end
-	sleep 1200
+	sleep 120
 end

### 2015-11-04 wim/twitter_problem/hll.rb
stream parsed more efficiently

@@ -1,0 +1,19 @@
+require 'hyperll'
+
+mhll = Hyperll::HyperLogLog.new()
+
+# hll = Hyperll::HyperLogLog.new(10)
+# hll.offer(1)
+# hll.offer(2)
+# hll.offer(3)
+# #hll.cardinality # => 3
+
+# hll2 = Hyperll::HyperLogLog.new(10)
+# hll2.offer(3)
+# hll2.offer(4)
+# hll2.offer(5)
+# #hll.cardinality # => 3
+
+# merged = Hyperll::HyperLogLog.new(10)
+# merged.merge(hll, hll2)
+# puts merged.cardinality # => 5

### 2015-11-04 wim/twitter_problem/twitter_stream.rb

@@ -1,22 +1,36 @@
 require 'twitter'
 require_relative 'const.rb'
-client = Twitter::REST::Client.new do |config|
-  config.consumer_key        = Constants.oh_one
-  config.consumer_secret     = Constants.oh_two
-  config.access_token        = Constants.oh_three
-  config.access_token_secret = Constants.oh_four
+
+def routine
+	tags = 0
+	
+	client = Twitter::REST::Client.new do |config|
+	  config.consumer_key        = Constants.oh_one
+	  config.consumer_secret     = Constants.oh_two
+	  config.access_token        = Constants.oh_three
+	  config.access_token_secret = Constants.oh_four
+	end
+
+	while true do
+		open('twitter_data.txt','a') do |f|
+			client.search("#", result_type: 'recent', lang: 'en').take(2000).each do |tweet|
+				twt = tweet.text
+				array = twt.split(' ')
+				array.each do |val|
+					if val[0] == "#" and val.length > 1 then
+						f.puts val
+						tags += 1
+					end
+				end
+			end
+		end
+		str = "Gathered Sample at #{Time.now.inspect}\n\tCurrent total number of tags: #{tags}"
+		open('tw_log.txt','a') do |file|
+			file.puts str
+			puts str
+		end
+		sleep 60
+	end
 end
 
-while true do
-	open('twitter_data.txt','a') do |f|
-		client.search("#", result_type: 'recent', lang: 'en').take(1000).each do |tweet|
-			f.puts tweet.text
-		end
-	end
-	str = "Gathered Sample at #{Time.now.inspect}"
-	open('tw_log.txt','a') do |file|
-		file.puts str
-		puts str
-	end
-	sleep 120
-end
+routine

### 2015-11-05 wim/paper.tex
fixes

@@
 \usepackage{setspace}
+\usepackage{listings}
+\usepackage{float}
 \usepackage{amssymb,amsmath}
+\usepackage{wrapfig}
 \usepackage{url}
+\usepackage{graphicx}
 \usepackage{fancyvrb}
 \usepackage{mathtools}
+\usepackage{floatflt}
 \usepackage[square,sort,comma,numbers]{natbib}
@@
 \indent Processing data efficiently has always been a goal of organizations that collect data; companies like Facebook and Google have been collecting and storing data for as long as they have existed.
+One important statistic that companies can use is the number of distinct elements in a large set of data.
 The count-distinct problem poses the question of how many distinct elements are in a given set. 
-The three problems that we will analyze are
-\begin{enumerate}
-\item \underline{The Twitter Problem}: How many unique hashtags are created on any given day?
-\item \underline{The Facebook Problem}: How many unique Facebook app installs are made on any given day?
-\item \underline{The Pok\'emon Problem}: How many unique Pok\'emon encounters will a player face in a given play through of a game from each generation.
-\end{enumerate}
-Each of these problems can be solved by using one of three different solutions to the count distinct problem: the hash table, the linear probabilistic counter, and the hyperloglog algorithm.
-\\\\
-\indent The count-distinct problem has not always been a problem. 
-Before the internet was widely used by the population, storing data was expensive, and as a result most companies did not want to spend money on the hardware and space that would be required to store large amounts of data.
-In addition, data was not transferred and gathered at a fast enough rate to warrant finding a way to count the data; even if companies wanted to count the data, the computers at the time were not powerful enough to perform the necessary calculations.
-The problem of counting big data did not arise until the 1990's, when the internet began to facilitate the easy transfer and collection of data. 
-Today, more than 25 exabytes of data are created and stored a day. 
-The main factor that contributes most to the problem is that gathering and storing data is cheap, but analyzing it is not.
+There are several algorithms that exist that can solve the count distinct problem; we will consider three of them:
+the hash table, the Linear Probabilistic Counter, and the HyperLogLog. 
+In order to analyze these algorithms, we will apply them to real world problems, namely (1) The Pok\'emon Problem, which poses the question ``How many unique Pok\'emon encounters will a player face in a given play through of a game from each generation?,'' (2) The Facebook Problem, which asks ``How many unique Facebook app installs are made on any given day?'', and (3) The Twitter Problem, where we will determine how many unique hashtags are created on any given day.
+We can use the hash table, the Linear Probabilistic Counter, and the HyperLogLog to solve each problem, respectively.
+%\\\\
+%\indent The count-distinct problem has not always been a problem. 
+%Before the internet was widely used by the population, storing data was expensive, and as a result most companies did not want to spend money on the hardware and space that would be required to store large amounts of data.
+%In addition, data was not transferred and gathered at a fast enough rate to warrant finding a way to count the data; even if companies wanted to count the data, the computers at the time were not powerful enough to perform the necessary calculations.
+%The problem of counting big data did not arise until the 1990's, when the internet began to facilitate the easy transfer and collection of data. 
+%Today, more than 25 exabytes of data are created and stored a day. 
+%The main factor that contributes most to the problem is that gathering and storing data is cheap, but analyzing it is not.
 The volume of the data leads to the count distinct problem that we will analyze \cite[pp.
@@
 \indent The first solution to the count distinct problem is called the \textit{Hash Table}. 
-A hash-table is a matrix that is stored in memory where each entry was uniquely obtained from another value. 
+A hash-table is a matrix that is stored in memory where each entry was uniquely obtained from another value, called a key. 
 The algorithm is as follows:
@@
 However, if we are looking for a \textit{good enough} estimate (within the neighborhood of 1\% to 5\% error) of the number of distinct elements in $\mathbb{S}$, we can use an algorithm called a Linear Probabilistic Counter.
-With the LPC, the user can specify how large they want the error to be.
-This method uses a data structure called a bit-map.
-Bit-maps are a matrix where either a 1 or 0 is stored in a cell, which allows for the map to be incredibly space efficient.
-The LPC also uses a hash, but it only needs to temporarily produce a position and then it can ``forget'' the value it just produced, which in turn reduces the overhead memory needed to store the elements that was required by the hash table method \cite{Whang}.
-The number of distinct elements is determined by the number of 1's in the bit-map when the algorithm is finished.
+With the Linear Probabilistic Counter, the user can specify how large they want the error to be.
+This method uses a data structure called a bitmap.
+bitmaps are a matrix where either a 1 or 0 is stored in a cell, which allows for the map to be incredibly space efficient.
+The Linear Probabilistic Counter also uses a hash, but it only needs to temporarily produce a position and then it can ``forget'' the value it just produced, which in turn reduces the overhead memory needed to store the elements that was required by the hash table method \cite{Whang}.
+The number of distinct elements is determined by the number of 1's in the bitmap when the algorithm is finished.
 \\\\
-\indent In order to either increase or decrease the error, the user of the LPC algorithm can either increase or decrease the size of the bit-map that is held in main memory.
+\indent In order to either increase or decrease the error, the user of the Linear Probabilistic Counter algorithm can either increase or decrease the size of the bitmap that is held in main memory.
 The following equation gives us a relationship between the size of the bitmap and the error that will be produced:
@@
 In other words, if the hash function being used produces a large amount of collisions, then the error will be high.
-Luckily, an algorithm such as SHA-2 will be okay to use, since the LPC never stores the hashed value in main memory.
+Luckily, an algorithm such as SHA-2 will be okay to use, since the Linear Probabilistic Counter never stores the hashed value in main memory.
 We can also calculate the number of distinct elements by
@@
 \]
-where $m$ is the size of the bit-map and $\Lambda_{n}$ is the ratio of the number of 0's in the bit map to the size of the bit map \cite[p. 212]{Whang}. 
+where $m$ is the size of the bitmap and $\Lambda_{n}$ is the ratio of the number of 0's in the bitmap to the size of the bitmap \cite[p. 212]{Whang}. 
 \\\\
-\indent The Facebook problem can be solved efficiently by using the LPC algorithm.
-We would have set $\mathbb{S}$ of size 20 million \cite{Facebook}, and we can choose an arbitrary sized bit map $\mathbb{V}$ of 10 million bits, or roughly 1.25 Mb, which lies on the upper limit of the efficiency of the LPC.
-If the memory required for the bit map increases beyond 1.25Mb, most operating systems will move the I/O operations to secondary memory, which defeats the purpose of the LPC algorithm \cite[p.
+\indent The Facebook problem can be solved efficiently by using the Linear Probabilistic Counter algorithm.
+We would have set $\mathbb{S}$ of size 20 million \cite{Facebook}, and we can choose an arbitrary sized bitmap $\mathbb{V}$ of 10 million bits, or roughly 1.25 Mb, which lies on the upper limit of the efficiency of the Linear Probabilistic Counter.
+If the memory required for the bitmap increases beyond 1.25Mb, most operating systems will move the I/O operations to secondary memory, which defeats the purpose of the Linear Probabilistic Counter algorithm \cite[p.
 211]{Whang}; as a result, the Twitter problem would not be efficiently solved by this method, since the amount of hashtags created in a day far exceeds 20 million.
-In the case of the Facebook problem, we can apply the LPC algorithm analytically, rather than actually processing 20 million application installs.
+In the case of the Facebook problem, we can apply the Linear Probabilistic Counter algorithm analytically, rather than actually processing 20 million application installs.
 \\\\
@@
 \\\\
-\indent The LPC and the hash table algorithms handled relatively small data sets well, but the Twitter problem deals with a data set of nearly 200,000,000 tweets a day \cite{Twitter-2}.
-As a result, the LPC would be too inefficient to process all the data, but the HyperLogLog algorithm allows for data to be processed more efficiently than the LPC or the hash table. 
-The HLL algorithm actually uses parts of the hash table and LPC algorithms; values are still hashed to a bit map held in memory. 
+\indent The Linear Probabilistic Counter and the hash table algorithms handled relatively small data sets well, but the Twitter problem deals with a data set of nearly 200,000,000 tweets a day \cite{Twitter-2}.
+As a result, the Linear Probabilistic Counter would be too inefficient to process all the data, but the HyperLogLog algorithm allows for data to be processed more efficiently than the Linear Probabilistic Counter or the hash table.
+The HyperLogLog algorithm actually uses parts of the hash table and Linear Probabilistic Counter algorithms; values are still hashed to a bitmap held in memory. 
 The hash function hashes values to binary numbers, and the algorithm decides whether or not that value has already been seen the value before by analyzing the leading 0's in the binary number \cite[pp.
@@
 If the sequence of 0's is identical to another sequence of 0's, the probability that the original values were the same is very high, and the algorithm will ignore what it finds to be a duplicate value.
-The amount of memory that the HLL algorithm requires per bitmap is expressed by
+The amount of memory that the HyperLogLog algorithm requires per bitmap is expressed by
 \[
@@
 where $M$ is the size of the original set of data \cite[pp. 129]{Flaj}.
-Finally, the algorithm takes the harmonic average of all the totals of the separate bit maps, which allows the algorithm to increase in accuracy as the size of the original set grows \cite{Yousra}.
+The error amount can be found by
+\[
+\text{Error}\ = \frac{\sqrt{3\log{(2)} - 1}}{\sqrt{m}},
+\]
+where $m$ is the number of spaces in the bitmap.
+Finally, the algorithm takes the harmonic average of all the totals of the separate bitmaps, given by
+\[
+H = \frac{n}{\sum_{i=1}^{n}\frac{1}{x_{i}}},
+\] 
+which allows the algorithm to increase in accuracy as the size of the original set grows \cite{Yousra}.
 \\\\
-\indent We can solve the Twitter problem with the HLL algorithm. 
-We will make the assumption that all 200,000,000 tweets made will have hashtags, which means we will have a set $\mathbb{S}$ of size 200,000,000, assuming every tweet made in a day contains a hashtag.
-By the formula for above, we get that the total required memory per bit map will be
+\indent We can solve the Twitter problem with the HyperLogLog algorithm.
+We will use the code in \textit{Figure 1} to aggregate a random sampling of the most recent 2000 tweets that contain a ``\#'' every two minutes for 24 hours.
+This will allow for a large random sample of tweets that will closely model the real problem that we are trying to solve. 
+We also need a hashing function that will produce a binary string of 1's and 0's given any key; luckily this hash function already exists.
+A murmur hash will produce a 32 bit binary string, which is exactly what we need.
+Finally, we actually need to implement the algorithm. 
+We will create a class to perform the calculation that is initialized with the amount of memory we are willing to spend on the bitmap for the HyperLogLog algorithm. 
+We know that the HyperLogLog can process nearly 1 billion elements with only 1.5 kilobytes of space to an accuracy of 2\% \cite{Flaj}.
+\begin{floatingfigure}[r]{8cm}
+
+\centering
+\includegraphics[scale=0.3]{twitter_problem/ruby_code_01}
+\caption{An infinite while loop can be useful}
+\end{floatingfigure}
+\noindent This is given by the equation we have for memory.
+Using a similar argument, we can come up with the amount of memory we will need. 
+If we want the error to be 1\%, then we can calculate the number of spaces required in the bitmap by using the equation we have for the error.
 \begin{align*}
-\text{Memory} &=\log_{2}{\left(\log_{2}{\left( 200,000,000\right)}\right)}\\
-&\approx  4.78531\text{b},
+0.01 &= \frac{\sqrt{3\log{(2)} - 1}}{\sqrt{m}}\\
+0.01\,\sqrt{m} &= \sqrt{3\log{(2)} - 1}\\
+\sqrt{m}\, &= \frac{\sqrt{3\log{(2)} - 1}}{0.01}\\
+m &= \frac{\sqrt{3\log{(2)} - 1}}{0.01^{2}}\\
+m &\approx 10,794.4\ \text{spaces in the bitmap}.
 \end{align*}
-The LPC algorithm would take at least ten times the amount of space just to hold the bit map in main memory.
-In addition to the memory efficiency of the HLL, the algorithm also has a relatively smaller error than the LPC, expressed by
-\[
-\text{Error}\ =\frac{\sqrt{3\log{(2)} - 1}}{\sqrt{m}}.
-\]
-Finally, the HyperLogLog finds a harmonic mean of the hashed values in the bit map, rather than a geometric mean, which increases the accuracy of the HLL even more. 
+Now we can find the total memory required for our logarithm. If we let the program from \textit{Figure 1} run for 24 hours, at worst it will parse 1,440,000 hashtags.
+We can now calculate the memory required by
+\begin{align*}
+\text{Memory}\, &= \log_{2}{(\log_{2}{(1,440,000)})}\\
+&= 4.35457 \, \text{bytes per entry}.
+\end{align*}
+Multiplying this number by the total number of entires gives us a grand total of 47005 bytes, or 4.7005 kilobytes of required memory.
 \\\\
-To solve the Twitter problem,
+\indent We now have all the necessary information to parse the data.
+We will implement the HyperLogLog in an object oriented manner; that is, we will create a Ruby class called HyperLogLog that we can initialize with the number of bytes per entry.
+We also need a method to calculate the unique elements, and we need a way to actually call this routine.
+\begin{center}
+\begin{tabular}{c c c}
+\includegraphics[scale=0.3]{twitter_problem/hll_init}
+&
+\includegraphics[scale=0.4]{twitter_problem/hll_count}
+&
+\includegraphics[scale=0.4]{twitter_problem/hll_imp}
+\\
+\textit{Initialization} & \textit{Counting} & \textit{Implementing}
+\end{tabular}
+\end{center}
+\noindent Note that we settled with 4 as the size of each entry in the bit-map. This value is close enough to 4.35 to safely say that the error will be just about 1\%.
+Finally, in our main program, we will loop infinitely (it will be up to us as to when to kill the process), and read in $\sim 2000$ tweets every two minutes. 
+After 24 hours, the HyperLogLog algorithm outputs 
+%%%%%%%%RESULTS GO HERE%%%%%%%%%%
+\\\\
+\indent
 

### 2015-11-05 wim/fb_problem/facebook_stream.rb

@@ -1,0 +1,13 @@
+require_relative "const.rb"
+require "facebook_oauth"
+
+def routine
+	client = FacebookOauth::Client.new(
+		:application_id => Constants.oh_one,
+		:application_secret => Constants.oh_two,
+		:callback => Constants.oh_three
+	)
+	client.authorize_url => 
+		"https://graph.facebook.com/oauth/authorize?scope=SCOPE&client_id=ID&type=web_server&redirect_uri=CALLBACK"
+		
+end

### 2015-11-05 wim/twitter_problem/hll.rb

@@ -1,19 +1,17 @@
 require 'hyperll'
 
-mhll = Hyperll::HyperLogLog.new()
-
-# hll = Hyperll::HyperLogLog.new(10)
-# hll.offer(1)
-# hll.offer(2)
-# hll.offer(3)
-# #hll.cardinality # => 3
-
-# hll2 = Hyperll::HyperLogLog.new(10)
-# hll2.offer(3)
-# hll2.offer(4)
-# hll2.offer(5)
-# #hll.cardinality # => 3
-
-# merged = Hyperll::HyperLogLog.new(10)
-# merged.merge(hll, hll2)
-# puts merged.cardinality # => 5
+module Worker
+	def Worker.calculate(len)
+		mhll = Hyperll::HyperLogLog.new(4)
+		File.open("twitter_data.txt","r") do |file|
+			file.each_line do |line|
+				mhll.offer line
+			end
+		end
+		str = "\tUnique Elements: #{mhll.cardinality}"
+		File.open("tw_log.txt",'a') do |file|
+			file.puts str
+			puts str
+		end
+	end
+end

### 2015-11-05 wim/twitter_problem/twitter_stream.rb

@@ -1,5 +1,7 @@
+#!/usr/bin/ruby
 require 'twitter'
 require_relative 'const.rb'
+require_relative 'hll.rb'
 
 def routine
 	tags = 0
@@ -29,7 +31,10 @@ def routine
 			file.puts str
 			puts str
 		end
-		sleep 60
+		Thread.new {
+			Worker.calculate tags
+		}
+		sleep 120
 	end
 end
 

### 2015-11-05 wim/twitter_problem/hll.rb
exceptions

@@ -2,7 +2,7 @@ require 'hyperll'
 
 module Worker
 	def Worker.calculate(len)
-		mhll = Hyperll::HyperLogLog.new(4)
+		mhll = Hyperll::HyperLogLog.new(10)
 		File.open("twitter_data.txt","r") do |file|
 			file.each_line do |line|
 				mhll.offer line

### 2015-11-05 wim/twitter_problem/twitter_stream.rb

@@ -14,27 +14,42 @@ def routine
 	end
 
 	while true do
-		open('twitter_data.txt','a') do |f|
-			client.search("#", result_type: 'recent', lang: 'en').take(2000).each do |tweet|
-				twt = tweet.text
-				array = twt.split(' ')
-				array.each do |val|
-					if val[0] == "#" and val.length > 1 then
-						f.puts val
-						tags += 1
+		begin
+			open('twitter_data.txt','a') do |f|
+				client.search("#", result_type: 'recent', lang: 'en').take(2000).each do |tweet|
+					twt = tweet.text
+					array = twt.split(' ')
+					array.each do |val|
+						if val[0] == "#" and val.length > 1 then
+							f.puts val
+							tags += 1
+						end
 					end
 				end
 			end
+			str = "Gathered Sample at #{Time.now.inspect}\n\tCurrent total number of tags: #{tags}"
+			open('tw_log.txt','a') do |file|
+				file.puts str
+				puts str
+			end
+			Thread.new {
+				Worker.calculate tags
+			}
+			sleep 120
+		rescue Twitter::Error::TooManyRequests
+			str = "Request Limit Reached at #{Time.now.inspect}. Sleeping for 16 minutes..."
+			open('tw_log.txt','a') do |file|
+				file.puts str
+				puts str
+			end
+			sleep 960
+			str = "Waking up at #{Time.now.inspect}."
+			open('tw_log.txt','a') do |file|
+				file.puts str
+				puts str
+			end
+			retry
 		end
-		str = "Gathered Sample at #{Time.now.inspect}\n\tCurrent total number of tags: #{tags}"
-		open('tw_log.txt','a') do |file|
-			file.puts str
-			puts str
-		end
-		Thread.new {
-			Worker.calculate tags
-		}
-		sleep 120
 	end
 end
 

### 2015-11-05 wim/paper.tex
this paper is destroying my life

@@
 \usepackage{floatflt}
+\usepackage{framed}
 \usepackage[square,sort,comma,numbers]{natbib}
@@
 We can use the hash table, the Linear Probabilistic Counter, and the HyperLogLog to solve each problem, respectively.
-%\\\\
-%\indent The count-distinct problem has not always been a problem. 
-%Before the internet was widely used by the population, storing data was expensive, and as a result most companies did not want to spend money on the hardware and space that would be required to store large amounts of data.
-%In addition, data was not transferred and gathered at a fast enough rate to warrant finding a way to count the data; even if companies wanted to count the data, the computers at the time were not powerful enough to perform the necessary calculations.
-%The problem of counting big data did not arise until the 1990's, when the internet began to facilitate the easy transfer and collection of data. 
-%Today, more than 25 exabytes of data are created and stored a day. 
-%The main factor that contributes most to the problem is that gathering and storing data is cheap, but analyzing it is not.
-The volume of the data leads to the count distinct problem that we will analyze \cite[pp.
-1]{Yousra}.
 \\\\
-\indent The first solution to the count distinct problem is called the \textit{Hash Table}. 
+\indent The first solution to the count distinct problem is called a hash table. 
 A hash-table is a matrix that is stored in memory where each entry was uniquely obtained from another value, called a key. 
@@
 The hashing function we created earlier will not be good enough here; instead we can create a new hashing function where we sum up the value of the name as before, then add the pok\'emon's unique id, and finally mod that value with 721.
-The following Java code can be used to model our hash table:\\\\
-\begin{tabular}{c c} 
-\begin{BVerbatim}
-public int getHashValue(){
-    char[] array = nameOfPokemon.toCharArray();
-    int total = 0;
-    for(char c : array){
-	total += (int)c;
-    }
-    total += type;
-    total %= 721;
-    return total;
-}
-\end{BVerbatim}
-&
-\begin{BVerbatim}
-...
-for(Pokemon pm : S){
-    int hashValue = pm.getHashValue();
-    storeHashValue();
-}
-calculateAndPrintUniqueValues();
-\end{BVerbatim}
-\end{tabular}
-Running this code will give us a final answer of 458 unique encounters with wild pok\'emon. 
+\begin{floatingfigure}[H!]{8cm}
+\centering
+\begin{framed}
+\includegraphics[scale=0.4]{pkmn_problem/hash_01}
+\caption{Hashing Pok\'emon based on name and type}
+\end{framed}
+\end{floatingfigure}
+\noindent Running the code in \textit{Figure 1} will give us a final answer of 461 unique encounters with wild pok\'emon. 
 In the case of the pok\'emon problem, a simple hash function sufficed, since the only collisions that were encountered were pok\'emon where the name was the same. 
-If we consider a larger data set, however, we may not be able to create a surjective hash function.
-\\\\
+If we consider a larger data set, however, we may not be able to create a surjective hash function.\\
 \indent Imagine that we are solving the Twitter problem where we want to count the number of distinct hashtags on twitter on a given day. 
@@
 On the same day, Target announced that all their \textit{Bob's Burgers} costumes would be on sale for the low price of \$12.95, and ``\#bob'' began trending on twitter.
-\end{quote}
+\end{quote}  \clearpage
 Our hash function from earlier will not work in this case, since both ``\#bob'' and ``\#cape'' hash to the same value. 
@@
 6]{Maurer}.
-Since the collision policy would not help us, we need to come up with a better hashing function. 
-\\\\
+Since the collision policy would not help us, we need to come up with a better hashing function.\\
 \indent More complex hash functions, such as SHA-2 (Secure Hash Algorithm 2) or MD5 (Message Digest 5) offer a solution to the collision problem.
@@
 22 - 23]{Break-MD5}.
-The hash method is not a viable solution to the count distinct problem when we are dealing with a large volume of data, but hash functions are still used in the other solutions to the count-distinct problem.
-\\\\
+The hash method is not a viable solution to the count distinct problem when we are dealing with a large volume of data, but hash functions are still used in the other solutions to the count-distinct problem.\\
 \indent The hash-table method provided a 0\% error, but was not efficient in handling large sets of data.
@@
 The Linear Probabilistic Counter also uses a hash, but it only needs to temporarily produce a position and then it can ``forget'' the value it just produced, which in turn reduces the overhead memory needed to store the elements that was required by the hash table method \cite{Whang}.
-The number of distinct elements is determined by the number of 1's in the bitmap when the algorithm is finished.
-\\\\
-\indent In order to either increase or decrease the error, the user of the Linear Probabilistic Counter algorithm can either increase or decrease the size of the bitmap that is held in main memory.
-The following equation gives us a relationship between the size of the bitmap and the error that will be produced:
-\[
-bias\left(\frac{\hat{n}}{n}\right) = \frac{e^{t} - t - 1}{2n},
-\]
-where $bias\left(\frac{\hat{n}}{n}\right)$ is the error.
-The value $t$ is the ratio of the number of distinct elements determined by the algorithm (denoted by $n$) to the size of the bitmap \cite{Whang}.
-If $t$ is close to 0, then the error will be very small; if $t$ is much greater than 1, the error will be very large.
-In other words, if the hash function being used produces a large amount of collisions, then the error will be high.
-Luckily, an algorithm such as SHA-2 will be okay to use, since the Linear Probabilistic Counter never stores the hashed value in main memory.
-We can also calculate the number of distinct elements by
+The number of distinct elements is determined by the number of 1's in the bitmap when the algorithm is finished.\\
+\indent The LPC algorithm is divided into several steps.
+In the first step, the algorithm creates an arbitrarily sized bitmap in memory, and then hashes the values into the bitmap by determining the address in the bitmap at which the corresponding bit should be changed to a 1.
+In the next step, the algorithm counts the number of 0's in the bitmap, and stores the result.
+Finally, the algorithm estimates the cardinality of the unique set by dividing the count of 0's by the size of the bitmap, and then plugging in the result to
 \[
@@
 \]
-where $m$ is the size of the bitmap and $\Lambda_{n}$ is the ratio of the number of 0's in the bitmap to the size of the bitmap \cite[p. 212]{Whang}. 
-\\\\
-\indent The Facebook problem can be solved efficiently by using the Linear Probabilistic Counter algorithm.
-We would have set $\mathbb{S}$ of size 20 million \cite{Facebook}, and we can choose an arbitrary sized bitmap $\mathbb{V}$ of 10 million bits, or roughly 1.25 Mb, which lies on the upper limit of the efficiency of the Linear Probabilistic Counter.
-If the memory required for the bitmap increases beyond 1.25Mb, most operating systems will move the I/O operations to secondary memory, which defeats the purpose of the Linear Probabilistic Counter algorithm \cite[p.
-211]{Whang}; as a result, the Twitter problem would not be efficiently solved by this method, since the amount of hashtags created in a day far exceeds 20 million.
-In the case of the Facebook problem, we can apply the Linear Probabilistic Counter algorithm analytically, rather than actually processing 20 million application installs.
-\\\\
-%%%%%%%%FINISH THIS SECTION%%%%%%%%%
-\indent We know that on a given day, Facebook users make 20 million app installs, and that there are currently about 7 million apps on Facebook \cite{Facebook}. 
-However, not all 7 million applications are installed every day; in fact, the amount of installs decays exponentially starting from the number one most installed application on Facebook, which is Spotify as of August 2015 \cite{Facebook-2}.
-We would expect that apps such as Spotify would dominate the market, whereas apps in the lower spectrum of the downloads would be negligible.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\\\\
+where $m$ is the size of the bitmap and $\Lambda_{n}$ is the ratio \cite{Whang}. 
+Figure 2 shows the Ruby implementation of the LPC algorithm.
+\begin{floatingfigure}[R]{6cm}
+\begin{framed}
+\centering
+\includegraphics[scale=0.4]{fb_problem/lpc}
+\caption{Determining the cardinality}
+\end{framed}
+\end{floatingfigure}
+\noindent The LPC algorithm also allows us to specify the amount of error we are willing to tolerate.
+This is accomplished by changing the size of the bitmap; if the bitmap is large, or close to the size of the original set $\mathbb{S}$, then the error will be small.
+Alternatively, if we decrease the size of the bitmap, we use less memory, but the error grows exponentially \cite{Hoff}.\\
+\indent We can use the LPC algorithm to solve the Facebook problem; unfortunately, we have no way of accessing the actual data we need, so we will have to do some guessing. 
+We know that Facebook users make about 7,000,000 app installs a dat \cite{Facebook-2}.
+This means that the size of the set $\mathbb{S}$ will be 7,000,000.
+Now we need to make some guesses; we expect that the applications that are most installed will appear a lot in our set, and that the applications that are installed less will appear less in $\mathbb{S}$.
+We do have access the to the top 20 Facebook application installs in August 2015, so we can plot those points and look for a trend in the data.
+Doing so tells us that the data follows an exponential curve \cite{Facebook-2}, which means that Spotify (the top application at the time) will appear frequently in $\mathbb{S}$, whereas LINE, the $20^{\text{th}}$ most popular Facebook app, will appear sparingly.
+This means that we will need a bitmap relatively close in size to the original set, since only a select few of the application have a high appearance rate in $\mathbb{S}$.
+In other words, we expect a lot of distinct elements in $\mathbb{S}$.\\
+\indent We can model the exponential function by choosing random numbers with a certain weight towards different numbers.
+We will let the each number correspond to a different application.
+For example, if we let 200 be the number for Spotify, we would expect a high appearance rate of 200 in $\mathbb{S}$.\\
+%%%Need to figure out how to do this
 \indent The Linear Probabilistic Counter and the hash table algorithms handled relatively small data sets well, but the Twitter problem deals with a data set of nearly 200,000,000 tweets a day \cite{Twitter-2}.
@@
 \] 
-which allows the algorithm to increase in accuracy as the size of the original set grows \cite{Yousra}.
-\\\\
+which allows the algorithm to increase in accuracy as the size of the original set grows \cite{Yousra}.\\
 \indent We can solve the Twitter problem with the HyperLogLog algorithm.
-We will use the code in \textit{Figure 1} to aggregate a random sampling of the most recent 2000 tweets that contain a ``\#'' every two minutes for 24 hours.
+We will use the code in \textit{Figure 2} to aggregate a random sampling of the most recent 2000 tweets that contain a ``\#'' every two minutes for 24 hours.
 This will allow for a large random sample of tweets that will closely model the real problem that we are trying to solve. 
@@
 Finally, we actually need to implement the algorithm. 
-We will create a class to perform the calculation that is initialized with the amount of memory we are willing to spend on the bitmap for the HyperLogLog algorithm. 
-We know that the HyperLogLog can process nearly 1 billion elements with only 1.5 kilobytes of space to an accuracy of 2\% \cite{Flaj}.
-\begin{floatingfigure}[r]{8cm}
-
+\begin{floatingfigure}[hr]{8cm}
 \centering
+\begin{framed}
 \includegraphics[scale=0.3]{twitter_problem/ruby_code_01}
 \caption{An infinite while loop can be useful}
+\end{framed}
 \end{floatingfigure}
-\noindent This is given by the equation we have for memory.
+We will create a class to perform the calculation that is initialized with the amount of memory we are willing to spend on the bitmap for the HyperLogLog algorithm. 
+\noindent We know that the HyperLogLog can process nearly 1 billion elements with only 1.5 kilobytes of space to an accuracy of 2\% \cite{Flaj}.\clearpage
+This is given by the equation we have for memory.
 Using a similar argument, we can come up with the amount of memory we will need. 
@@
 \end{align*}
-Multiplying this number by the total number of entires gives us a grand total of 47005 bytes, or 4.7005 kilobytes of required memory.
-\\\\
+Multiplying this number by the total number of entires gives us a grand total of 47005 bytes, or 4.7005 kilobytes of required memory.\\
 \indent We now have all the necessary information to parse the data.
@@
 \end{center}
-\noindent Note that we settled with 4 as the size of each entry in the bit-map. This value is close enough to 4.35 to safely say that the error will be just about 1\%.
+\noindent \\
+Note that we settled with 4 as the size of each entry in the bit-map. This value is close enough to 4.35 to safely say that the error will be just about 1\%.
 Finally, in our main program, we will loop infinitely (it will be up to us as to when to kill the process), and read in $\sim 2000$ tweets every two minutes. 
-After 24 hours, the HyperLogLog algorithm outputs 
+After running the algorithm for 24 hours, we will have a\\ 
 %%%%%%%%RESULTS GO HERE%%%%%%%%%%
-\\\\
-\indent
+\indent We have looked at three different approaches to the count distinct problem. 
+We used the HyperLogLog to solve only the Twitter problem, but we could have used the algorithm to solve all of the problems. 
+The count distinct problem
 
@@
 \begin{thebibliography}{9}
+\bibitem{Hoff} Hoff, T.
+(2012, April 5).
+Big Data Counting: How to count a billion distinct objects using only 1.5KB of Memory - High Scalability -.
+Retrieved October 27, 2015, from \url{http://highscalability.com/blog/2012/4/5/big-data-counting-how-to-count-a-billion-distinct-objects-us.html}
 \bibitem{Facebook} Facebook Statistics. (2015, September 20). Retrieved October 30, 2015, from \url{http://www.statisticbrain.com/facebook-statistics/}

### 2015-11-05 wim/fb_problem/lpc.rb

@@ -1,0 +1,18 @@
+def count(m, s)
+	bitmap = Array.new(m)
+	for i in 0..m do
+		bitmap[i] = 0
+	end
+	s.each do |value|
+		address = m_hash(value)
+		bitmap[address] = 1
+	end
+	zeros = 0
+	bitmap.each do |one_or_zero|
+		if one_or_zero == 0 then
+			zeros = zeros + 1
+		end
+	end
+	lambda = zeros / m
+	cardinality = (-1 * m) * Math.log(lambda)
+end

### 2015-11-05 wim/pkmn_problem/hash.rb

@@ -1,0 +1,55 @@
+def main
+	array = Array.new
+	pkmn_list = Array.new(721)
+	hash_table = Array.new(721)
+	for i in 0..721 do
+		hash_table[i] = -1
+	end
+	File.open("pokemon_list.txt",'r') do |file|
+		file.each_line do |line|
+			str_arr = line.split ','
+			pkmn_list.push Pokemon.new(str_arr[1],str_arr[2].to_i)
+		end
+	end
+	for i in 0..6000 do
+		pokee = pkmn_list.sample
+		while pokee == nil do
+			pokee = pokee = pkmn_list.sample
+		end
+		puts "pokee is #{pokee}"
+		array.push pokee
+	end
+	array.each do |pokemon|
+		hash_table[pokemon.custom_hash] = pokemon.custom_hash
+	end
+	total = 0
+	hash_table.each do |val|
+		if val != -1
+			total = total + 1
+		end
+	end
+	puts "The total number of unique encounters is #{total}"
+end
+
+class Pokemon
+	def initialize(name="", type=0)
+		@name = name
+		@type = type
+	end
+
+	def custom_hash
+		total = 0
+		@name.each_byte do |car|
+			total = total + car
+		end
+		total = total + @type
+		total = total % 721
+		return total
+	end
+
+	def to_s
+		@name
+	end
+end
+
+main

### 2015-11-06 wim/paper.tex
conclusion

@@
 In order to analyze these algorithms, we will apply them to real world problems, namely (1) The Pok\'emon Problem, which poses the question ``How many unique Pok\'emon encounters will a player face in a given play through of a game from each generation?,'' (2) The Facebook Problem, which asks ``How many unique Facebook app installs are made on any given day?'', and (3) The Twitter Problem, where we will determine how many unique hashtags are created on any given day.
-We can use the hash table, the Linear Probabilistic Counter, and the HyperLogLog to solve each problem, respectively.
-\\\\
+We can use the hash table, the Linear Probabilistic Counter, and the HyperLogLog to solve each problem, respectively.\\
 \indent The first solution to the count distinct problem is called a hash table. 
@@
 For smaller sets of data, however, we can create a hash function that is surjective. 
-The pok\'emon problem provides a good example of the benefits of a hash table.
-\\\\
+The pok\'emon problem provides a good example of the benefits of a hash table.\\
 \indent The most beneficial aspect of the hash is that it has a 0\% error rate. 
@@
 The number of distinct elements is determined by the number of 1's in the bitmap when the algorithm is finished.\\
-\indent The LPC algorithm is divided into several steps.
+\indent The Linear Probabilistic Counter algorithm is divided into several steps.
 In the first step, the algorithm creates an arbitrarily sized bitmap in memory, and then hashes the values into the bitmap by determining the address in the bitmap at which the corresponding bit should be changed to a 1.
@@
 where $m$ is the size of the bitmap and $\Lambda_{n}$ is the ratio \cite{Whang}. 
-Figure 2 shows the Ruby implementation of the LPC algorithm.
+Figure 2 shows the Ruby implementation of the Linear Probabilistic Counter algorithm.
 \begin{floatingfigure}[R]{6cm}
@@
 \end{floatingfigure}
-\noindent The LPC algorithm also allows us to specify the amount of error we are willing to tolerate.
+\noindent The Linear Probabilistic Counter algorithm also allows us to specify the amount of error we are willing to tolerate.
 This is accomplished by changing the size of the bitmap; if the bitmap is large, or close to the size of the original set $\mathbb{S}$, then the error will be small.
 Alternatively, if we decrease the size of the bitmap, we use less memory, but the error grows exponentially \cite{Hoff}.\\
-\indent We can use the LPC algorithm to solve the Facebook problem; unfortunately, we have no way of accessing the actual data we need, so we will have to do some guessing. 
+\indent We can use the Linear Probabilistic Counter algorithm to solve the Facebook problem; unfortunately, we have no way of accessing the actual data we need, so we will have to do some guessing. 
 We know that Facebook users make about 7,000,000 app installs a dat \cite{Facebook-2}.
@@
 We also need a hashing function that will produce a binary string of 1's and 0's given any key; luckily this hash function already exists. 
-A murmur hash will produce a 32 bit binary string, which is exactly what we need.
-Finally, we actually need to implement the algorithm. 
 \begin{floatingfigure}[hr]{8cm}
@@
 \end{floatingfigure}
+A murmur hash will produce a 32 bit binary string, which is exactly what we need.
+Finally, we actually need to implement the algorithm. 
 We will create a class to perform the calculation that is initialized with the amount of memory we are willing to spend on the bitmap for the HyperLogLog algorithm. 
@@
 We used the HyperLogLog to solve only the Twitter problem, but we could have used the algorithm to solve all of the problems. 
-The count distinct problem
+The count distinct problem is heavily applicable in the real world; companies are able to turn the numbers that are determined by algorithms such as the HyperLogLog to increase profits, target ads, and keep more data.
+Finding an efficient way to count large sets of data may never be competely accurate, but with algorithms like the HyperLogLog and the Linear Probabilistic counter, we can efficiently and with nearly 100\% accuracy solve the count distinct problem.
 

### 2015-11-07 wim/paper.tex
almost solved Twitter problem

@@
 We can use the hash table, the Linear Probabilistic Counter, and the HyperLogLog to solve each problem, respectively.\\
+
 \indent The first solution to the count distinct problem is called a hash table. 
@@
 The pok\'emon problem provides a good example of the benefits of a hash table.\\
+
 \indent The most beneficial aspect of the hash is that it has a 0\% error rate. 
@@
 If we consider a larger data set, however, we may not be able to create a surjective hash function.\\
+
 \indent Imagine that we are solving the Twitter problem where we want to count the number of distinct hashtags on twitter on a given day. 
@@
 Since the collision policy would not help us, we need to come up with a better hashing function.\\
+
 \indent More complex hash functions, such as SHA-2 (Secure Hash Algorithm 2) or MD5 (Message Digest 5) offer a solution to the collision problem.
@@
 The hash method is not a viable solution to the count distinct problem when we are dealing with a large volume of data, but hash functions are still used in the other solutions to the count-distinct problem.\\
+
 \indent The hash-table method provided a 0\% error, but was not efficient in handling large sets of data.
@@
 In the next step, the algorithm counts the number of 0's in the bitmap, and stores the result.
-Finally, the algorithm estimates the cardinality of the unique set by dividing the count of 0's by the size of the bitmap, and then plugging in the result to
+Finally, the algorithm estimates the cardinality of the unique set by dividing the count of 0's by the size of the bitmap, and then substituting in the result to
 \[
@@
 We will let the each number correspond to a different application.
-For example, if we let 200 be the number for Spotify, we would expect a high appearance rate of 200 in $\mathbb{S}$.\\
-%%%Need to figure out how to do this
+For example, if we let 200 be the number for Spotify, we would expect a high appearance rate of 200 in $\mathbb{S}$.
+We can use an algorithm (courtesy of GitHub user Danieth5x) that will generate a pseudo-random sample based on a function, which will be an exponential function in our case.
+We can argue that it is not important as to which exponential function we choose, since the distribution will be more or less the same; in our case we will pick $e^{x}$.
+Running the pseudo-random generator produces a set $\mathbb{S}$ of size 7,000,000, and we can now apply the Linear Probabilistic Counter to the data set.
+For our purposes, we will use the hash function supplied by Ruby; the only change we need to make is to mod the result of the Ruby hash with the number of spaces in the bitmap, which ensures that we will never produce a  hashed value out of the bounds of our bitmap.
+Finally, running the algorithm tells us that the total number of unique application installs is \underline{RESULT GOES HERE}.\\
+
 \indent The Linear Probabilistic Counter and the hash table algorithms handled relatively small data sets well, but the Twitter problem deals with a data set of nearly 200,000,000 tweets a day \cite{Twitter-2}.
@@
 \indent We can solve the Twitter problem with the HyperLogLog algorithm.
-We will use the code in \textit{Figure 2} to aggregate a random sampling of the most recent 2000 tweets that contain a ``\#'' every two minutes for 24 hours.
+We will use the code in \textit{Figure 3} to aggregate a random sampling of the most recent 2000 tweets that contain a ``\#'' every two minutes for 24 hours.
 This will allow for a large random sample of tweets that will closely model the real problem that we are trying to solve.
 We also need a hashing function that will produce a binary string of 1's and 0's given any key; luckily this hash function already exists.
+A murmur hash will produce a 32 bit binary string, which is exactly what we need.
+Finally, we actually need to implement the algorithm. 
+We will create a class to perform the calculation that is initialized with the amount of memory we are willing to spend on the bitmap for the HyperLogLog algorithm. 
+We know that the HyperLogLog can process nearly 1 billion elements with only 1.5 kilobytes of space to an accuracy of 2\% \cite{Flaj}.
+This is given by the equation we have for memory.
+Using a similar argument, we can determine the amount of memory we will need.
 \begin{floatingfigure}[hr]{8cm}
@@
 \end{floatingfigure}
-A murmur hash will produce a 32 bit binary string, which is exactly what we need.
-Finally, we actually need to implement the algorithm. 
-We will create a class to perform the calculation that is initialized with the amount of memory we are willing to spend on the bitmap for the HyperLogLog algorithm. 
-\noindent We know that the HyperLogLog can process nearly 1 billion elements with only 1.5 kilobytes of space to an accuracy of 2\% \cite{Flaj}.\clearpage
-This is given by the equation we have for memory.
-Using a similar argument, we can come up with the amount of memory we will need. 
-If we want the error to be 1\%, then we can calculate the number of spaces required in the bitmap by using the equation we have for the error.
+\noindent If we want the error to be 1\%, then we can calculate the number of spaces required in the bitmap by using the equation we have for the error.
 \begin{align*}
@@
 We also need a method to calculate the unique elements, and we need a way to actually call this routine.
+\clearpage
 \begin{center}
@@
 Finally, in our main program, we will loop infinitely (it will be up to us as to when to kill the process), and read in $\sim 2000$ tweets every two minutes. 
-After running the algorithm for 24 hours, we will have a\\ 
+After running the algorithm for 24 hours, we will have a set of data that represents the number of unique hashtags found (by using the HyperLogLog algorithm) as the number of tweets being analyzed increases.
+The data grows in a linear fashion; this is expected since we are gathering a random sample from all over the world, which means that at any given time, we expect to see more or less the same amount of unique hashtags.
+Plotting the data and applying a best fit line using Mathematica gives us
+\[
+Equation\ Goes\ Here.
+\]
+From this data, we can make an estimate that at 200,000,000 tweets, we will have
+\[
+Results\ From\ Plugging\ In\ 200,000,000,
+\]
+which means that we will have roughly \underline{INSERT RESULT HERE} unique hashtags by the HyperLogLog algorithm.\\ 
 %%%%%%%%RESULTS GO HERE%%%%%%%%%%
@@
 \begin{thebibliography}{9}
-\bibitem{Hoff} Hoff, T.
-(2012, April 5).
-Big Data Counting: How to count a billion distinct objects using only 1.5KB of Memory - High Scalability -.
-Retrieved October 27, 2015, from \url{http://highscalability.com/blog/2012/4/5/big-data-counting-how-to-count-a-billion-distinct-objects-us.html}
+\bibitem{Twitter} \#numbers. (2011, March 14). Retrieved October 29, 2015, from \url{https://blog.twitter.com/2011/numbers}
+\bibitem{SHA-Algo} Chaves, Ricardo, Kuzmanov, Georgi, Sousa, Leonel \& Vassiliadis, Stamatis (2006). Improving SHA-2 Hardware Implementations. , 4249, 298-310.
 \bibitem{Facebook} Facebook Statistics. (2015, September 20). Retrieved October 30, 2015, from \url{http://www.statisticbrain.com/facebook-statistics/}
-\bibitem{Whang} Whang, Kyu-Young, Vander-Zanden, Brad T. \& Taylor, Howard M. (1990). A Linear-time Probabilistic Counting Algorithm for Database Applications. ACM Trans. Database Syst., 15, 208-229.
-\bibitem{Maurer} Maurer, W. D. \& Lewis, T. G. (1975). Hash Table Methods. ACM Comput. Surv., 7, 5-19.
-\bibitem{SHA-Algo} Chaves, Ricardo, Kuzmanov, Georgi, Sousa, Leonel \& Vassiliadis, Stamatis (2006). Improving SHA-2 Hardware Implementations. , 4249, 298-310.
-\bibitem{Break-MD5} Wang, Xiaoyun \& Yu, Hongbo (2005). How to Break MD5 and Other Hash Functions. , 3494, 19-35.
+\bibitem{Flaj} Flajolet, P., Fusy, \'E., Gandouet, O., \& Meunier, F.
+(n.d.).
+HyperLogLog: The Analysis of a Near-Optimal Cardinality Estimation Algorithm.
+2007 \textit{Conference on Analysis of Algorithms, 07,} 127-143.
+Retrieved October 27, 2015, from \url{http://algo.inria.fr/flajolet/Publications/FlFuGaMe07.pdf}
 \bibitem{Heule} Heule, S., Nunkesser, M., \& Hall, A.
@@
 New York, NY, Genoa: ACM.
-\bibitem{Twitter} \#numbers. (2011, March 14). Retrieved October 29, 2015, from \url{https://blog.twitter.com/2011/numbers}
+\bibitem{Hoff} Hoff, T.
+(2012, April 5).
+Big Data Counting: How to count a billion distinct objects using only 1.5KB of Memory - High Scalability -.
+Retrieved October 27, 2015, from \url{http://highscalability.com/blog/2012/4/5/big-data-counting-how-to-count-a-billion-distinct-objects-us.html}
 \bibitem{Pokemon} How many Pokemon are there altogether, before and after Omega Ruby and Alpha Sapphire?
@@
 Retrieved October 29, 2015, from \url{http://pokemondb.net/pokebase/219332/pokemon-there-altogether-before-after-omega-alpha-sapphire}
+\bibitem{Maurer} Maurer, W. D. \& Lewis, T. G. (1975). Hash Table Methods. ACM Comput. Surv., 7, 5-19.
 \bibitem{Facebook-2} Most popular Facebook apps 2015 | Statistic.
@@
 \bibitem{Twitter-2} Twitter Statistics. (2015, September 15). Retrieved November 3, 2015, from \url{http://www.statisticbrain.com/twitter-statistics/}
+\bibitem{Break-MD5} Wang, Xiaoyun \& Yu, Hongbo (2005). How to Break MD5 and Other Hash Functions. , 3494, 19-35.
+\bibitem{Whang} Whang, Kyu-Young, Vander-Zanden, Brad T. \& Taylor, Howard M. (1990). A Linear-time Probabilistic Counting Algorithm for Database Applications. ACM Trans. Database Syst., 15, 208-229.
 \bibitem{Yousra}Yousra Chabchoub, Georges H\'ebrail. \textit{Sliding HyperLogLog: Estimating cardinality in a data stream.} 2010. 
-\bibitem{Flaj} Flajolet, P., Fusy, \'E., Gandouet, O., \& Meunier, F.
-(n.d.).
-HyperLogLog: The Analysis of a Near-Optimal Cardinality Estimation Algorithm.
-2007 \textit{Conference on Analysis of Algorithms, 07,} 127-143.
-Retrieved October 27, 2015, from \url{http://algo.inria.fr/flajolet/Publications/FlFuGaMe07.pdf}
 \end{thebibliography}

### 2015-11-07 wim/fb_problem/daniel.rb

@@ -1,0 +1,62 @@
+require "pp"
+class Distribution
+    attr_accessor :x_min
+    attr_accessor :x_max
+    attr_accessor :counts
+    attr_accessor :distribution
+    attr_accessor :types
+
+    def initialize(types, f, x_min = 1, x_max = 100)
+        self.x_min        = x_min
+        self.x_max        = x_max
+        self.counts       = Array.new(types)
+        self.distribution = Array.new(types + 1)
+        self.types        = types
+
+        # Sum distribution for normalization
+        m = (x_max - x_min).fdiv(types)
+        sum_of_distribution = 0;
+
+        self.distribution[0] = f.call(x_min)
+        sum_of_distribution += self.distribution[0]
+
+        (1...types).each do |c|
+            self.distribution[c] = f.call(x_min + m * c)
+            sum_of_distribution += self.distribution[c]
+            self.distribution[c] += self.distribution[c-1]
+        end
+
+        (0...types).each do |c|
+            self.distribution[c] = self.distribution[c].fdiv(sum_of_distribution)
+            puts self.distribution[c]
+        end
+
+        # Normalize distribution and
+        # Manipulate such that c*(n) = c(n) + c(n-1) + ... + c(1) + c(0)
+    end   
+
+    def count(n)
+        res = Array.new(n)
+        (0...n).each do |i|
+            r = rand()
+            res[i] = self.distribution.find_index { |x| x >= r }
+        end
+        res
+    end
+
+    def bucket_count(n)
+        res = Array.new(n)
+        n.times do
+            r = rand()
+            res[self.distribution.find_index { |x| x >= r }] += 1
+        end
+        res
+    end
+end
+
+f = lambda do |x|
+  x ** 2
+end
+
+distribution = Distribution.new(1000, f, 1, 100)
+p distribution.count(100000)

### 2015-11-07 wim/fb_problem/facebook_stream.rb

@@ -1,13 +1,0 @@
-require_relative "const.rb"
-require "facebook_oauth"
-
-def routine
-	client = FacebookOauth::Client.new(
-		:application_id => Constants.oh_one,
-		:application_secret => Constants.oh_two,
-		:callback => Constants.oh_three
-	)
-	client.authorize_url => 
-		"https://graph.facebook.com/oauth/authorize?scope=SCOPE&client_id=ID&type=web_server&redirect_uri=CALLBACK"
-		
-end

### 2015-11-11 pde/pdetest1.tex
flip flop

@@
 &\\
-u(x,t) = \sum_{n=1}^{\infty}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}\sin{\left(\frac{n\pi x}{L}\right)}
-& u(x,t) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}\cos{\left(\frac{n\pi x}{L}\right)}\\
+u(x,t) = \sum_{n=1}^{\infty}b_{n}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}\sin{\left(\frac{n\pi x}{L}\right)}
+& u(x,t) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}a_{n}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}\cos{\left(\frac{n\pi x}{L}\right)}\\
 &\\
@@
 &\\
-u(x,t) =  \sum_{n=1}^{\infty}e^{-k\left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}\right)^{2}t}\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}
-& u(x,t) =  \sum_{n=1}^{\infty}e^{-k\left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}\right)^{2}t}\cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}
+u(x,t) =  \sum_{n=1}^{\infty}a_{n}e^{-k\left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}\right)^{2}t}\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}
+& u(x,t) =  \sum_{n=1}^{\infty}a_{n}e^{-k\left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}\right)^{2}t}\cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}
 \end{array}
@@
 and we are done.
+
+\newpage
+\indent Now we will move on to the Cauchy wave equation. We will be presented with a similar problem: a wave equation on an unbounded domain. Consider the following problem:
+\noindent \\\\
+\noindent \textbf{\textit{Ex:}} Solve the following PDE
+\[
+\begin{cases}
+u_{tt} - c^{2}u_{xx} = 0\\
+u(x,0) = \sin{x}\\
+u_{t}(x,0) = \cos{x}
+\end{cases}
+\]
+\indent \textbf{\textit{Solution:}} Just as with the heat equation, we have a formal solution for the homogeneous wave equation.
+We will do the derivation, but I will tell you that the formal solution is
+\[
+u(x,t) = \frac{1}{2}\left[f(x+ct) + f(x-ct)\right] + \frac{1}{2c}\int_{x-ct}^{x+ct}g(s)\mathop{ds}.
+\]
+Now let's derive the formal solution. We want a solution in the form of
+\[
+u(x,t) = F(x+ct) + G(x-ct).
+\]
+Now we can say that $u_{t} = cF'(x+ct) - cG'(x-ct)$, and $u_{t}(x,0) = cF'(x) - cG'(x)$. This means
+\begin{align*}
+F(x) + G(x) &= f(x)\\
+F(x) - G(x) &= \frac{1}{c}\int_{0}^{x}g(s)\mathop{ds} + c_{1}.
+\end{align*}
+Solving the system yields
+\begin{align*}
+u(x,t) &= \frac{1}{2} f(x+ct) + \frac{1}{2c}\int_{0}^{x+ct}g(s)\mathop{ds} + \frac{c_{1}}{2} + \frac{1}{2}f(x-ct) -\frac{1}{2c}\int_{0}^{x-ct}g(s)\mathop{ds} - \frac{c_{1}}{2}\\
+&= \frac{1}{2}\left[f(x+ct) + f(x-ct)\right] + \frac{1}{2c}\int_{x-ct}^{x+ct}g(s)\mathop{ds},
+\end{align*}
+and that is the derivation. This problem is simply plugging into the formal solution. We have the following:
+\begin{align*}
+u(x,t) &= \frac{1}{2}\left[\sin{(x+ct)} + \sin{(x-ct)}\right] + \frac{1}{2c}\int_{x-ct}^{x+ct}\cos{(s)}\mathop{ds}\\
+&= \frac{\sin{(x+ct)} + \sin{(x-ct)}}{2} + \frac{\cos{(x)}\sin{(ct)}}{c}\\
+&=  \frac{\cos{(x)}\sin{(ct)}}{c} + \cos{(ct)}\sin{(x)},
+\end{align*}
+and we are done.
+
+\newpage
+\indent So far, we have only dealt with homogeneous Cauchy problems.
+We will now look at the case where we have a non-homogeneous problem.
+The first variation of this problem is the \textit{half-line} problem.
+In this problem, we will have a bound on our equation, and we will generalize the solution so that we are dealing with an unbounded Cauchy problem.
+Let's look at an example.
+\noindent\\\\ \textbf{\textit{Ex:}} Solve the following PDE
+\[
+\begin{cases}
+u_{t} - ku_{xx} = 0\\
+u(x,0) = \phi(x)\\
+u(0,t) = 0
+\end{cases}
+\]
+\indent \textbf{\textit{Solution:}} We have both an initial condition and a boundary condition here.
+This problem is saying that at $x=0$, there is an impassable boundary.
+This forces our domain to be $[0,\infty)$.
+To solve a Cauchy problem, however, we want a domain of $(-\infty,\infty)$.
+We can force this domain by extending our initial condition, $\phi(x)$, over the boundary.
+We have the following rule:
+\[
+\phi_{0}(x) =
+\begin{cases}
+\phi(x)\quad\quad x>0\\
+-\phi(-x)\ x<0.
+\end{cases}
+\]
+Now we are dealing with the system
+\[
+\begin{cases}
+u_{t} - ku_{xx} = 0\\
+u(x,0) = \phi_{0}(x),
+\end{cases}
+\]
+which we can solve using the formal solution to the heat equation. Recall that
+\[
+u(x,t) = \frac{1}{\sqrt{4k\pi t}}\int_{-\infty}^{\infty}e^{-\frac{(x-y)^{2}}{4kt}}\phi_{0}(y)\mathop{dy}.
+\]
+Substituting into the equation gives us
+\begin{align*}
+u(x,t) &= \frac{1}{\sqrt{4k\pi t}}\left[\int_{0}^{\infty}e^{-\frac{(x-y)^{2}}{4kt}}\phi(y)\mathop{dy} + \int_{-\infty}^{0}e^{-\frac{(x-y)^{2}}{4kt}}(-\phi(-y))\mathop{dy} \right]\\
+&= \frac{1}{\sqrt{4k\pi t}}
+\end{align*}
 \end{document}

### 2015-11-11 wim/presentation.tex
flip flop

\documentclass{beamer}
\mode<presentation>
\usetheme{boxes}
\usepackage{amsmath}
\AtBeginSection[]{
\begin{frame}<beamer>
\frametitle{Summary}
\tableofcontents[currentsection]
\end{frame}
}
\title{The Count Distinct Problem}
\author{Steven Rosendahl}

\begin{document}
\begin{frame}
\titlepage
\end{frame}

\section{The Problem}
\begin{frame}{The Question}
\pause
\begin{itemize}
\item What is the problem?
\pause
\item Image we have a set called $\mathbb{V}$ that contains a billion elements of the same type.
\pause
\item How many \textit{unique} elements are in $\mathbb{V}$?
\end{itemize}
\end{frame}

\begin{frame}{Questions}
\begin{enumerate}
\pause
\item \underline{Pok\'emon Problem:} How many unique Pok\'emon will a player encounter in a given playthrough of all the games?
\pause
\item \underline{Facebook Problem:} How many unique application installs are made a day on Facebook?
\pause
\item \underline{Twitter Problem:} How many unique hashtags are made a day on Twitter?
\pause
\end{enumerate}
We will use $\mathbb{S}$ to represent the set of all the data, and $\mathbb{V}$ to represent the set of unique elements. 
\end{frame}

\section{The Hash Table}
\begin{frame}{Hashing}
\begin{itemize}
\pause
\item What is \textit{hashing}?
\pause
\item Applying a function $h(x)$ to every element in $\mathbb{S}$, and storing the result in $\mathbb{V}$.
\pause
\item Ideally, $h(x)$ is
\begin{enumerate}
\pause 
\item Onto (surjective)
\pause
\item One-to-one (injective)
\end{enumerate}
\pause
\item We can ignore the duplicate values in $\mathbb{V}$.
\end{itemize}
\end{frame}

\begin{frame}{Solving the Pok\'emon Problem}
\begin{itemize}
\pause
\item How can we solve the Pok\'emon Problem using a hash?
\pause
\item Create a hash function that turns a given Pok\'emon into a numerical value
\pause
\item Store the result in $\mathbb{V}$ if it is not already there.
\pause
\item The hash function:
\begin{enumerate}
\item Sum up the ASCII value of each character in a Pok\'emon's name. Call this $n$.
\pause
\item Add $n$ to the Pok\'emon's corresponding National Pok\'edex number. Call this $m$.
\pause
\item Find $m\ mod\ 721$.
\pause
\end{enumerate}
\end{itemize}
\end{frame}

\begin{frame}{Problems With The Hash Table}
\begin{itemize}
\pause
\item Memory Intensive
\begin{itemize}
\item Pok\'emon problem only dealt with a set $\mathbb{S}$ of size 6000
\pause
\item Twitter Problem deals with $\mathbb{S}$ of size 200,000,000.
\pause
\item Collisions and collision policies also add to the amount of memory required.
\end{itemize}
\end{itemize}
\end{frame}

\section{The HyperLogLog}
\begin{frame}{The Algorithm}
\begin{enumerate}
\pause
\item Create a 
\end{enumerate}
\end{frame}
\end{document}

### 2015-11-11 wim/fb_problem/daniel.rb

@@ -28,7 +28,7 @@ class Distribution
 
         (0...types).each do |c|
             self.distribution[c] = self.distribution[c].fdiv(sum_of_distribution)
-            puts self.distribution[c]
+            # puts self.distribution[c]
         end
 
         # Normalize distribution and
@@ -54,9 +54,9 @@ class Distribution
     end
 end
 
-f = lambda do |x|
-  x ** 2
-end
+# f = lambda do |x|
+#   x ** 2
+# end
 
-distribution = Distribution.new(1000, f, 1, 100)
-p distribution.count(100000)
+# distribution = Distribution.new(1000, f, 1, 100)
+# p distribution.count(100000)

### 2015-11-11 wim/fb_problem/lpc.rb

@@ -1,10 +1,14 @@
+require_relative 'daniel.rb'
+
 def count(m, s)
 	bitmap = Array.new(m)
-	for i in 0..m do
+	i = 0
+	while i < m do
 		bitmap[i] = 0
+		i += 1
 	end
 	s.each do |value|
-		address = m_hash(value)
+		address = m_hash(value, m)
 		bitmap[address] = 1
 	end
 	zeros = 0
@@ -16,3 +20,20 @@ def count(m, s)
 	lambda = zeros / m
 	cardinality = (-1 * m) * Math.log(lambda)
 end
+
+def m_hash(value, size)
+	value.hash % size
+end
+
+def main
+	f = lambda do |x|
+		# Math::E ** x
+		x + 1
+	end
+	distribution = Distribution.new(3000, f)
+	set = distribution.count 100000
+	puts set.inspect
+	# puts count(1000000, set)
+end
+
+main

### 2015-11-11 wim/twitter_problem/hll.rb

@@ -1,17 +1,0 @@
-require 'hyperll'
-
-module Worker
-	def Worker.calculate(len)
-		mhll = Hyperll::HyperLogLog.new(10)
-		File.open("twitter_data.txt","r") do |file|
-			file.each_line do |line|
-				mhll.offer line
-			end
-		end
-		str = "\tUnique Elements: #{mhll.cardinality}"
-		File.open("tw_log.txt",'a') do |file|
-			file.puts str
-			puts str
-		end
-	end
-end

### 2015-11-11 wim/twitter_problem/twitter_stream.rb

@@ -1,56 +1,88 @@
 #!/usr/bin/ruby
 require 'twitter'
+require 'hyperll'
 require_relative 'const.rb'
-require_relative 'hll.rb'
 
-def routine
-	tags = 0
-	
-	client = Twitter::REST::Client.new do |config|
-	  config.consumer_key        = Constants.oh_one
-	  config.consumer_secret     = Constants.oh_two
-	  config.access_token        = Constants.oh_three
-	  config.access_token_secret = Constants.oh_four
-	end
+module TwitterProblem
+	def TwitterProblem.routine
+		tags = 0
+		
+		client = Twitter::REST::Client.new do |config|
+		  config.consumer_key        = Constants.oh_one
+		  config.consumer_secret     = Constants.oh_two
+		  config.access_token        = Constants.oh_three
+		  config.access_token_secret = Constants.oh_four
+		end
 
-	while true do
-		begin
-			open('twitter_data.txt','a') do |f|
-				client.search("#", result_type: 'recent', lang: 'en').take(2000).each do |tweet|
-					twt = tweet.text
-					array = twt.split(' ')
-					array.each do |val|
-						if val[0] == "#" and val.length > 1 then
-							f.puts val
-							tags += 1
+		while true do
+			begin
+				File.open('twitter_data.txt','a') do |f|
+					client.search("#", result_type: 'recent', lang: 'en').take(2000).each do |tweet|
+						twt = tweet.text
+						array = twt.split(' ')
+						array.each do |val|
+							if val[0] == "#" and val.length > 1 then
+								f.puts val
+								tags += 1
+							end
 						end
 					end
 				end
+				str = "Gathered Sample at #{Time.now.inspect}\n\tCurrent total number of tags: #{tags}"
+				File.open('tw_log.txt','a') do |file|
+					file.puts str
+					puts str
+				end
+				Thread.new {
+					calculate
+				}
+				sleep 120
+			rescue Twitter::Error::TooManyRequests
+				str = "Request Limit Reached at #{Time.now.inspect}. Sleeping for 16 minutes..."
+				File.open('tw_log.txt','a') do |file|
+					file.puts str
+					puts str
+				end
+				sleep 960
+				str = "Waking up at #{Time.now.inspect}."
+				File.open('tw_log.txt','a') do |file|
+					file.puts str
+					puts str
+				end
+				retry
 			end
[…]

### 2015-11-13 wim/paper.tex
flip flop
so many thing

@@
 For our purposes, we will use the hash function supplied by Ruby; the only change we need to make is to mod the result of the Ruby hash with the number of spaces in the bitmap, which ensures that we will never produce a  hashed value out of the bounds of our bitmap.
-Finally, running the algorithm tells us that the total number of unique application installs is \underline{RESULT GOES HERE}.\\
+Finally, running the algorithm tells us that the total number of unique application installs is 3216.\\
 
@@
 \\
-\textit{Initialization} & \textit{Counting} & \textit{Implementing}
+Initialization & Counting & Implementing
 \end{tabular}
@@
 \[
-Equation\ Goes\ Here.
+f(x) = 0.284356 x + 7,361.39.
 \]
 From this data, we can make an estimate that at 200,000,000 tweets, we will have
-\[
-Results\ From\ Plugging\ In\ 200,000,000,
-\]
-which means that we will have roughly \underline{INSERT RESULT HERE} unique hashtags by the HyperLogLog algorithm.\\ 
-%%%%%%%%RESULTS GO HERE%%%%%%%%%%
+\begin{align*}
+f(200,000,000) &= 0.284356(200,000,000) + 7,361.39\\
+&= 56,878,500,
+\end{align*}
+which means that we will have roughly 56 million unique hashtags by the HyperLogLog algorithm.\\ 
 \indent We have looked at three different approaches to the count distinct problem. 

### 2015-11-13 wim/presentation.tex
so many thing

@@
 \pause
-\item Create a 
+\item Create a bitmap in memory. We will call this $\mathbb{V}$.
+\pause
+\item For each value $s$ is $\mathbb{S}$, hash $s$ to a binary number.
+\pause
+%\item
+\begin{itemize}
+\item We will use a \textit{Murmur Hash} to do this.
+\end{itemize}
+\pause
+\item Analyze the first sequence of 0's in the binary value, and store the number of leading 0's into the bitmap.
+\begin{itemize}
+\item If the sequence of 0's has been seen before, there is a high probability that the term is a duplicate
+\end{itemize}
+\pause
+\item Take the harmonic average of all the totals in the bitmap.
 \end{enumerate}
 \end{frame}
+
+\begin{frame}{The Math}
+\pause
+\begin{itemize}
+\item How much memory is required?
+\pause
+\[
+\text{Memory Required} = \log_{2}{(\log_{2}{(M)})}
+\]
+\pause
+\item $M$ is the size of the original set of data ($\mathbb{S}$).
+\pause
+\item What is the predicted error?
+\pause
+\[
+\text{Error} = \frac{\sqrt{3\log{(2)} - 1}}{m}
+\]
+\pause
+\item $m$ is the number of spaces in the bitmap ($\mathbb{V}$).
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Solving the Twitter Problem}
+\pause
+\begin{itemize}
+\item How can we process count 200,000,000 hashtags on an average computer?
+\pause
+\item We can lower the sample size and apply a best fit line to the data.
+\begin{enumerate}
+\pause
+\item For 24 hours, gather 2000 tweets containing ``\#'' every 2 minutes.
+\pause
+\item Using the HyperLogLog, determine the unique number of total hashtags every time a new sample is gathered.
+\end{enumerate}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Results}
+\pause
+\begin{figure}
+\includegraphics[scale=0.5]{twitter_problem/resultspng}
+\caption{$0.284356 x + 7361.39$}
+\end{figure}
+\pause
+\begin{itemize}
+\item Plugging in 200,000,000 gives us
+\pause
+\item $5.68785\times10^7$ unique hashtags.
+\end{itemize}
+\end{frame}
 \end{document}

### 2015-11-13 wim/fb_problem/lpc.rb

@@ -3,22 +3,28 @@ require_relative 'daniel.rb'
 def count(m, s)
 	bitmap = Array.new(m)
 	i = 0
-	while i < m do
+	# while i < m do
+	# 	bitmap[i] = 0
+	# 	i += 1
+	# end
+	for i in 0..m do
 		bitmap[i] = 0
-		i += 1
 	end
 	s.each do |value|
 		address = m_hash(value, m)
 		bitmap[address] = 1
 	end
-	zeros = 0
+	zeros = 0.0
 	bitmap.each do |one_or_zero|
 		if one_or_zero == 0 then
-			zeros = zeros + 1
+			zeros += 1.0
 		end
 	end
-	lambda = zeros / m
-	cardinality = (-1 * m) * Math.log(lambda)
+	# puts "zeros: #{zeros}"
+	# puts "m: #{m}"
+	lamb = (zeros / m)
+	# puts "lambda: #{lamb}"
+	cardinality = (-1 * m) * Math.log(lamb)
 end
 
 def m_hash(value, size)
@@ -28,12 +34,13 @@ end
 def main
 	f = lambda do |x|
 		# Math::E ** x
-		x + 1
+		Math::E ** x
 	end
-	distribution = Distribution.new(3000, f)
-	set = distribution.count 100000
-	puts set.inspect
-	# puts count(1000000, set)
+	amnt = 7000000
+	distribution = Distribution.new(30000, f)
+	set = distribution.count amnt
+	#puts set.inspect
+	puts count(amnt, set)
 end
 
 main

### 2015-11-13 wim/paper.tex
working on paper

@@
 \end{floatingfigure}
-\noindent Running the code in \textit{Figure 1} will give us a final answer of 461 unique encounters with wild pok\'emon. 
+\noindent In order to come up with a set $\mathbb{S}$, we can randomly populate an array with 6000 pok\'emon; we are assuming that a player has an equal chance to run into any pok\'emon, and that they can encounter all pok\'emon more than once.
+Running the code in \textit{Figure 1} gives us a final answer of 461 unique encounters with wild pok\'emon. 
 In the case of the pok\'emon problem, a simple hash function sufficed, since the only collisions that were encountered were pok\'emon where the name was the same. 
@@
 On the same day, Target announced that all their \textit{Bob's Burgers} costumes would be on sale for the low price of \$12.95, and ``\#bob'' began trending on twitter.
-\end{quote}  \clearpage
+\end{quote}
 Our hash function from earlier will not work in this case, since both ``\#bob'' and ``\#cape'' hash to the same value. 

### 2015-11-13 wim/fb_problem/lpc.rb

@@ -3,10 +3,6 @@ require_relative 'daniel.rb'
 def count(m, s)
 	bitmap = Array.new(m)
 	i = 0
-	# while i < m do
-	# 	bitmap[i] = 0
-	# 	i += 1
-	# end
 	for i in 0..m do
 		bitmap[i] = 0
 	end
@@ -20,10 +16,7 @@ def count(m, s)
 			zeros += 1.0
 		end
 	end
-	# puts "zeros: #{zeros}"
-	# puts "m: #{m}"
 	lamb = (zeros / m)
-	# puts "lambda: #{lamb}"
 	cardinality = (-1 * m) * Math.log(lamb)
 end
 

### 2015-11-15 wim/pkmn_problem/Hasher.java
remove pkmn hash

@@ -1,77 +1,0 @@
-import java.util.*;
-import java.io.*;
-
-public class Hasher {
-	static Pokemon[] pkmn = new Pokemon[6000];
-	static int[] hvs = new int[721];
-
-	private static void readInPokemon(String filename){
-		try {
-			Scanner in = new Scanner(new File(filename));
-			Scanner lineReader;
-			List<Pokemon> tmp = new ArrayList<>();
-			while(in.hasNextLine()){
-				lineReader = new Scanner(in.nextLine()).useDelimiter(",");
-				lineReader.next();
-				Pokemon adder = new Pokemon(lineReader.next(), lineReader.nextInt());
-				tmp.add(adder);
-			}
-			for(int i = 0; i < 6000; i++){
-				Pokemon p = tmp.get((int)(Math.random() * 712));
-				// System.out.println("Pokemon: " + p + " Hash Value: " + p.getHashValue());
-				pkmn[i] = p;
-			}
-		} catch(Exception ex){
-			ex.printStackTrace();
-		}
-	}
-
-	public static void main(String[] args){
-		for(int i = 0; i < 712; i++){
-			hvs[i] = -1;
-		}
-		readInPokemon("pokemon_list.txt");
-		int collisions = 0;
-		for(Pokemon pm : pkmn){
-			int hv = pm.getHashValue();
-			if(hvs[hv] == 0){
-				System.out.println("Collision! HashValue: " + hv);
-				collisions++;
-			}
-			hvs[pm.getHashValue()] = 0;
-		}
-		int unique = 0;
-		for(int i : hvs){
-			if (i == 0){
-				unique++;
-			}
-		}
-		System.out.println("Total # of collisions: " + collisions);
-		System.out.println("The total is " + unique);
-	}
-
-	private static class Pokemon{
-		private String name;
-		private int type;
-
-		public Pokemon(String n, int t){
-			name = n;
-			type = t;
-		}
-
-		public int getHashValue(){
-			char[] array = name.toCharArray();
-			int total = 0;
-			for(char c : array){
-				total += (int)c;
-			}
-			total += type;
-			total %= 721;
-			return total;
-		}
-
-		public String toString(){
-			return name;
-		}
-	}
-}

### 2015-11-16 cheatsheets/ode_formal_slns.tex
Added ODE stuff

--- 2015-10-29 pde/pdetest1.tex
@@
 \usepackage{mathtools}
-\usepackage{graphicx}
-\usepackage{multicol}
 \usepackage{array}
@@
 \evensidemargin=-0.2in
+\title{Formal Solutions To ODE's}
+\date{}
 \begin{document}
-\section{First Order Partial Differential Equations}
-\hrule
-\noindent\\\\
-\indent We will begin by solving the typical first order PDE.
-The strategy here is to break it down into and ODE or several ODEs that can be solved by methods we already have.
-It does not matter whether the PDE is homogeneous or not; we will most likely be able to \\\\
-\noindent\textbf{\textit{Ex:}} Solve the following PDE
-\[
-\begin{cases*}
-yu_{x} + xu_{y} = 0\\
-u(0,y) = e^{-y^{2}}
-\end{cases*}
-\]
-\indent \textbf{\textit{Solution:}} This is a nice PDE; there are two ways to solve it, but we will focus on the easier one.
-We can see that we have a coefficient on both $u_{x}$ and $u_{y}$.
-This tells us that when $u$ was differentiated, there was a constant $y$ and $x$.
-So now we can say that
-\[
-\frac{dy}{dx} = \frac{x}{y}
-\]
-\noindent This is a simple ODE to solve. We can use separation of variables:
-\begin{align*}
-xdx &= ydy\\
-\int{x dx} &= \int{y dy}\\
-\frac{1}{2}x^{2} &= \frac{1}{2}y^{2} + c\\
-x^{2} - y^{2} &= c_{0}\\
-y^{2} - x^{2} &= c_{1}
-\end{align*}
-\noindent You'll notice that we multiplied both sides by $-1$. This is for convenience. Now we can express the answer as
-\[
-u(x,y) = f(y^{2} - x^{2})
-\]
-\noindent And we are done.
+\maketitle
 
-\newpage
-\noindent\textbf{\textit{Ex:}} Solve the following PDE
-\[
-u_{x} + u_{y} + u = x + y
-\]
-\indent \textbf{\textit{Solution:}} In this case, we have a non-zero term on the right hand side of the equation.
-We will use substitution to solve this.
-For the time being, we will let $T$ be $x + y$.
-To find our $S$, we can use a similar method to the one that we used above.
-We know that the coefficients on the $u_{x}$ and $u_{y}$ are $1$ in this case.
-Now we have:
-\begin{align*}
-\frac{dy}{dx} &= 1\\
-dy &= dx\\
-\int dy &= \int dx\\
-x &= y + c\\
-x - y &= c
-\end{align*}
-\noindent Now we can say that $S = x - y$.
-Next we have to actually use the substitution.
-We need to express $u_{x}$ and $u_{y}$ in terms of $S$ and $T$.
-The following equations come from the chain rule; when the $S$ and $T$ terms are differentiated, we get $u_{S}$ and $u_{T}$ left behind.
-\begin{alignat*}{3}
-u_{x} &= u_{S}\frac{\partial S}{\partial x} + u_{T}\frac{\partial T}{\partial x} 
-\qquad\qquad &&u_{y} = u_{S}\frac{\partial S}{\partial y} + u_{T}\frac{\partial T}{\partial y}\\
-u_{x} &= -u_{S} + u_{T}  \qquad\qquad &&u_{y} = u_{S} + u_{T}
-\end{alignat*}
-\noindent Now we have our substitution for $u_{x}$ and $u_{y}$.
-We can now plug in our values into the original equation.
-Note that we didn't do anything with $u$; it gets left alone.
-In addition, if we needed $u_{xx}$ or $u_{yy}$, we can just differentiate again.
-We will see this later when we talk about classifications.
-Now, back to plugging in:
-\begin{gather*}
-(-u_{S} + u_{T}) + (u_{S} + u_{t}) + u = T\\
-2u_{T} + u = T
-\end{gather*}
-\noindent This is just a first order ODE. Since it's not homogeneous, we need to solve for the general solution, and then the particular solution. Let's start with the general.
-\begin{gather*}
-2y^{'} + y = 0\\
-2r + 1 = 0\\
-r = -\frac{1}{2}\\
-y = ce^{\left(-\frac{1}{2}\right)x}
-\end{gather*}
-\noindent In this case, we know that our $x$ is really $T$.
-So we finally have $y = ce^{\left(-\frac{1}{2}\right)T}$.
-We will back substitute soon, but let's find the particular solution before we do that.
-\noindent Recall from ODE that we have several cases for the particular solution. In our case, we have a linear equation on the right hand side of our ODE. This lets us say:
-\begin{gather*}
-y = AT + B\\
-y^{'} = A\\
-\end{gather*}
-\noindent Now we can substitute again into our non-homogeneous ODE as follows:
-\[
-2y^{'} + y = T \equiv 2A + AT + B = T
-\]
-\noindent And we have that $A = 1$, and $B = -2$. Now we have a final solution of:
-\begin{gather*}
-y = ce^{\left(-\frac{1}{2}\right)T} + T - 2\\
-y = ce^{\left(-\frac{1}{2}\right)(x+y)} + (x + y) - 2
-\end{gather*}
-\noindent We aren't quite done yet, as we have a solution to an ODE, not a PDE.
-It isn't too difficult to switch back; The $c$ is really the only difference.
-In our ODE, the $c$ is a constant term, but to our PDE, the $c$ is a function.
-The $c$ now becomes $f(s)$.
-We will go ahead and back substitute, which will give us $f(y-x)$.
-Now we have our final solution:
-\[
-u(x,y) = f(y-x)e^{\left(-\frac{1}{2}\right)(x+y)} + (x + y) - 2
-\]
-\noindent And we are done.
+\section{Homogeneous Solutions}
 
-\newpage
-\section{Classification of Partial Differential Equations}
-\hrule
-\noindent\\\\
-\indent We mentioned earlier the topic of classification of PDE's.
-Since we only have this and Fourier Series to talk about, we will go ahead and knock out classification.
-There's some information that we need to go ahead and review here.
-This will be explained in the problem, but it's handy to have it for reference:
-\begin{align*}
-b^{2} - ac < 0 &\implies Elliptic\ with\ complex\ roots\\
-b^{2} - ac = 0 &\implies Parabolic\ with\ the\ same\ root\ twice\\
-b^{2} - ac > 0 &\implies Hyperbolic\ with\ two\ distinct\ real\ roots
-\end{align*}
-\noindent Another helpful piece of information is what's known as the \textit{classical trinity} of PDEs.
-\begin{align*}
-u_{t} - ku_{xx} &= 0 \qquad The\ heat\ equation\ (parabolic)\\
-u_{tt} - c^{2}u_{xx} &= 0 \qquad The\ wave\ equation\ (hyperbolic)\\
-u_{xx} + u_{yy} &= 0 \qquad The\ Laplace\ equation\ (elliptic)
-\end{align*}
-Let's go ahead and look at an example.\\
-\indent \textbf{Ex. }Classify the following PDE:
-\[yu_{xx} + u_{yy} = 0\]
-\indent \textbf{\textit{Solution:}} The first thing we need to identify are the coefficients on the $u_{xx}$, $u_{yy}$, and $u_{xy}$ or $u_{yx}$ terms.
-Note that the $u_{xy}$ and $u_{yx}$ are interchangeable.
-We will assign these coefficients to $a$, $b$, and $c$ corresponding to $u_{xx}$, $u_{xy}$, and $u_{yy}$, respectively.
-Looking at the PDE, we can see that
-\[
-a = y \qquad b = 0 \qquad c = 1
-\]
-\noindent Now we look at $b^{2} - ac$.
-It should be noted that we are really dealing with $b = \frac{b_{0}}{2}$.
-In other words, we need to divide our $b$ value by $2$.
-In the case of the first problem, we are dealing with $b = 0$, so it's not evident.
-So, now we can go ahead and classify our PDE as follows:
-\[
-b^{2} - ac = 0 - y < 0 \implies Elliptic
-\]
-Note that we assume $y > 0$.
-So our PDE is elliptic.
-Now we need to make it look like it's corresponding equation from the classical trinity.
-In this case we have an Elliptic equation, so we want to make it look like Laplace's equation.
-We need to find values for $u_{xx}$ and $u_{yy}$, and then substitute those into the equation.
-This is similar to how we solved the first order PDE's; let's look at our characteristic equation:
-\begin{gather*}
-am^{2} + bm + c = 0\\
-ym^{2} + 1 = 0\\
-m = \pm \frac{i}{\sqrt{y}}
-\end{gather*}
-\noindent So, now we have two solutions, and they are both imaginary as expected.
-Next we need to find values for $S$ and $T$ that we can substitute into the equation.
-We will do both of them at the same time here.
-Notice that we will drop the $i$; we don't really care about the imaginary part:
-\begin{gather*}
-\frac{dy}{dx} = \frac{1}{\sqrt{y}} \qquad\qquad\qquad \frac{dy}{dx} = -\frac{1}{\sqrt{y}}\\
-\int\frac{dy}{dx} =\int\frac{1}{\sqrt{y}} \qquad\qquad\qquad \int\frac{dy}{dx} = -\int\frac{1}{\sqrt{y}}\\
-S = \frac{2}{3}y^{\left(\frac{3}{2}\right)} - x \qquad\qquad\qquad T = \frac{2}{3}y^{\left(\frac{3}{2}\right)} + x
-\end{gather*}
-\noindent Now we can substitute into our equation. Unfortunately, we will need to take the partial derivative twice in order to get the substitutions we need. Let's start with $u_{xx}$.
-\begin{align*}
-u_{x} &= u_{S}\frac{\partial S}{\partial x} + u_{T}\frac{\partial T}{\partial x}\\
-&= -u_{S} + u_{T}\\
-u_{xx} &= \frac{\partial}{\partial x}(-u_{S} + u_{T})\\
-&= u_{SS} + u_{TT}
-\end{align*}
-\noindent That wasn't too bad. We will skip the math for finding $u_{yy}$, which would give us
-\[
-u_{yy} = yu_{SS} + yu_{TT}
-\]
-\noindent Now we plug everything in. Remember that we were under the assumption that $y > 0$:
-\begin{gather*}
-y(u_{SS} + u_{TT}) + yu_{SS} + yu_{TT} = 0\\
-2yu_{SS} + 2yu_{TT} = 0\\
-u_{SS} + u_{TT} = 0
-\end{gather*}
-\noindent And we are finished.
+\subsection{First Order}
 
-\newpage
-\section{Fourier Series and Partial Differential Equations}
-\hrule
-\noindent\\
-\subsection{Finding Fourier Series}
-\indent It's finally time to talk about Fourier Series.
-Before we start though, it may be helpful to look at some common patterns amongst the Fourier Series.
-We have the following information that will help us:\\\\
-\centerline{
-\bgroup
-\def\arraystretch{1.5}
-\begin{tabular}{|c|c|c|c|}
-\hline
-& Cosine & Sine & Full Series \\ \hline
-Interval & $[0,L]$ & $[0,L]$ & $[-L,L]$\\ \hline
-Basis & $\{1,\cos{\left(\frac{n\pi x}{L}\right)}\}$ &
-$\{\sin{\left(\frac{n\pi x}{L}\right)}\}$ &
-$\{1,\sin{\left(\frac{n\pi x}{L}\right)},\cos{\left(\frac{n\pi x}{L}\right)}\}$\\ \hline
-Series & 
-$\frac{b_{0}}{2} + \sum_{n=1}^{\infty}[b_{n}\cos{\left(\frac{n\pi x}{L}\right)}]$& 
-$\sum_{n=1}^{\infty}[a_{n}\sin{\left(\frac{n\pi x}{L}\right)}]$&
-$\frac{b_{0}}{2} + \sum_{n=1}^{\infty} [a_{n}\sin{\left(\frac{n\pi x}{L}\right)} + b_{n}\cos{\left(\frac{n\pi x}{L}\right)}]$\\ \hline
-\end{tabular}
-\egroup
-}\\\\
-\noindent To solve for the coefficients, we have:
-\begin{align*}
-Sine:  \qquad\qquad &a_{n} = \frac{2}{L}\int_{0}^{L}f(x)\sin{\left(\frac{n\pi x}{L}\right)}dx\\
-Cosine: \qquad\qquad &b_{n} = \frac{2}{L}\int_{0}^{L}f(x)\cos{\left(\frac{n\pi x}{L}\right)}dx\\
-&b_{0} =  \frac{1}{L}\int_{0}^{L}f(x)dx\\
-Full: \qquad\qquad &a_{n} = \frac{2}{L}\int_{-L}^{L}f(x)\sin{\left(\frac{n\pi x}{L}\right)}dx\\
-&b_{n} = \frac{2}{L}\int_{-L}^{L}f(x)\cos{\left(\frac{n\pi x}{L}\right)}dx\\
-&b_{0} =  \frac{1}{L}\int_{-L}^{L}f(x)dx
-\end{align*}
-\indent It will also help us to quickly review the concept of odd and even functions, as they will help us to simplify the process of computing Fourier series later.
-Specifically, let's look at the graphs for $\sin$ and $\cos$ on $[-\pi,\pi]$, as well as another function that will turn out to be neither even nor odd..
-\begin{center}
-\begin{tabular}{c c c}
-\includegraphics[scale=0.4]{cos_eo} & \includegraphics[scale=0.4]{sin_eo} & \includegraphics[scale=0.4]{cube_eo}\\
-$f(x) = \cos{x}$ & $f(x) = \sin{x}$ & $f(x) = x^{3}-x +1$
-\end{tabular}
-\end{center}
-\noindent\\
-\indent We should recall that a function is \textit{even} when $f(x) = f(-x)$ for all $x$ in the domain of $f$.
-Graphically speaking, an \textit{even} function is reflected across only the $y$-axis.
-By looking at the above graphs, we can see that $f(x) = \cos{x}$ is an even function.\\
-\indent A function is \textit{odd} when $-f(x) = f(-x)$ for all x in the domain of $f$.
-Graphically speaking, the function is reflected across the $x$-axis and the $y$-axis.
-From the above graphs, we can say that $f(x) = \sin{x}$ is an odd function.
-Note that we can have a function that is neither odd nor even, but we won't see much of this in the following problems.
-The reason that we want to discuss odd and even functions here is because the integration that we will be seeing later will be affected by the symmetry of a function.
-Let's take a look at how even and odd functions effect integration:\\\\
-\textbf{\textit{Ex:}} Compute the following integral:
-\[
-\int_{0}^{\pi}\cos{x}dx
-\]
-\indent \textbf{\textit{Solution:}} Looking at this integral, we can see from the graph of the $\cos$ above that this will be 0.
-That answer is meaningless to us however.
-What we are really looking for is
-\[
-2\int_{0}^{\frac{\pi}{2}}\cos{x}dx
-\]
-\noindent We are allowed to do this since the $\cos$ function is even. Essentially, we have turned the negative half of the cosine into a positive half. Similarly, we could ask for something like
-\[
-\int_{0}^{\pi}|\cos{x}|dx
-\]
-\noindent Since the absolute value of a function is no integrable, we drop the absolute value signs.
-This will leave us with our initial problem, which we will have to change anyways.
-Our integral becomes:
-\[
-\int_{0}^{\pi}|\cos{x}|dx = 2\int_{0}^{\frac{\pi}{2}}\cos{x}dx = 1
-\]
-\noindent And we are done.
-\newpage
-
-\indent The first type of series we will look at is the Full Fourier Series.
-Note that this series is just a combination of a Fourier Sine Series and a Fourier Cosine Series.
-A typical problem would look something like this:\\\\
-\textbf{\textit{Ex:}} Find the corresponding Fourier Series for
-\begin{gather*}
-f(x)=
-\begin{cases*}
-L \qquad -L \leq x \leq 0\\
-2x \qquad 0 \leq x \leq L
-\end{cases*}
-\end{gather*}
-\indent \textbf{\textit{Solution:}} So we have a piecewise function here, but that shouldn't throw us off too much; all we need to do is combine the integrals for each part of the coefficient solution.
-Let's start by finding the terms for the $cos$ series.
-\begin{align*}
-B_{0} &= \frac{1}{2L}\left[\int_{-L}^{0}Ldx + \int_{0}^{L}2xdx\right]\\
-&= \frac{1}{2L}[2L^{2}]\\
-&= L
-\end{align*}
-\noindent That wasn't too bad, was it? Now We need to find $B_{n}$, and then we will have our cosine part of the series. The integration can get pretty messy, so we will skip some steps.
-\begin{align*}
-B_{n} &= \frac{1}{L}\left[\int_{-L}^{0}L\cos{\left(\frac{n\pi x}{L}\right)}dx + \int_{0}^{L}2x\cos{\left(\frac{n\pi x}{L}\right)}dx\right]\\
-&= \frac{1}{L}\left[0 + \left(\frac{2L^{2}}{n^{2}\pi^{2}}\right)((-1)^{n} - 1)\right]\\
-&= \frac{2L}{(n\pi)^{2}}((-1)^{n} - 1)
-\end{align*}
-\noindent Note that the sequence will be 0 for all $n = 2k$. This allows us to substitute $2k+1$ for $n$, since we don't care about the $0$ terms. For our cosine part of the series, we now have:
-\begin{align*}
-B_{n} &= \frac{2L}{(2k+1)^{2}\pi^{2}}((-1)^{2k+1}-1)\\
-&= \frac{2L}{(2k+1)^{2}\pi^{2}}(-2)\\
-&= \frac{-4L}{(2k+1)^{2}\pi^{2}}
-\end{align*}
-\noindent So, now we need to find the $\sin$ part of the series. This is easier, since we only need to find $A_{n}$.
-\begin{align*}
-A_{n} &= \frac{1}{L}\left[\int_{-L}^{0}L\sin{\left(\frac{n\pi x}{L}\right)}dx + \int_{0}^{L}2x\sin{\left(\frac{n\pi x}{L}\right)}dx \right]\\
-&= \frac{L}{n\pi}\left[-1-(-1)^{n}\right]
-\end{align*}
-\noindent Note that we can do a similar substitution here as we did with the cosine series. This time, however, we will have $0$ for $n = 2k+1$ We now have:
-\begin{align*}
-A_{n} &= \frac{L}{2k\pi}\left[-1-(-1)^{2k}\right]\\
-&= \frac{L}{2k\pi}\left[-1 - 1\right]\\
-&= \frac{L}{2k\pi}\left[-2\right]\\
-&= \frac{-2L}{2k\pi}
-\end{align*}
-\noindent Now we have all the parts that we need. Our final answer will be essentially the sum of all the parts:
-\begin{align*}
-f(x)&\sim \frac{L}{2} + \sum_{k = 0}^{\infty}\left[\frac{-4L}{(2k+1)^{2}\pi^{2}}\cos{\left(\frac{(2k + 1)\pi x}{L}\right)} \right] + \sum_{k = 0}^{\infty}\left[\frac{-2L}{2k\pi}\sin{\left(\frac{2k\pi x}{L}\right)} \right]\\
-&\sim \frac{L}{2} + \sum_{k = 0}^{\infty}\left[\frac{-4L}{(2k+1)^{2}\pi^{2}}\cos{\left(\frac{(2k + 1)\pi x}{L}\right)} + \frac{-2L}{2k\pi}\sin{\left(\frac{2k\pi x}{L}\right)} \right]
-\end{align*}
-\noindent And we are done.\\
-
-\newpage
-\indent Now lets take a look at a slightly more theoretical Fourier Series.\\
-\noindent\textbf{\textit{Ex:}} Find the corresponding Fourier Series for\\
-\[f(x)=\left|\cos{x}\right|\qquad -\pi\leq x \leq \pi\]
-\indent\textbf{\textit{Solution:}} Before we dive into calculations here, let's think about the function at hand.
-We have the absolute value of $\cos{x}$.
-Usually, $\cos$ would be an odd function, which would imply that our $\cos$ series component to the whole Fourier Series would be $0$.
-However, the absolute value makes our function even, and as such the $\sin$ component of the series will be $0$.
-That's good news for us, because now we don't have to find $a_{n}$; we actually only have a Fourier Cosine Series here.
-Recall that we have a formula for $b_{n}$ and $b_{0}$, and that our general solution will have the form of
-\[\frac{b_{0}}{2} + \sum_{n=1}^{\infty}[b_{n}\cos{\left(\frac{n\pi x}{L}\right)}]\]
-\noindent Our function is bounded above by $\pi$, so we really have $L= \pi$. This gives us:
-\[\frac{b_{0}}{2} + \sum_{n=1}^{\infty}[b_{n}\cos{\left(nx\right)}]\]
-\noindent Let's go ahead and find our $b_{0}$:
-\begin{align*}
-b_{0} &= \frac{1}{\pi}\int_{-\pi}^{\pi}\left|\cos{x}\right|dx\\
-&= \frac{2}{\pi}\int_{0}^{\pi}\left|\cos{x}\right|dx\\
-&= \frac{4}{\pi}\int_{0}^{\frac{\pi}{2}}\left|\cos{x}\right|dx\\
-&= \frac{4}{\pi}\left[\sin{\frac{\pi}{2}} - \sin{0}\right]\\
-&= \frac{4}{\pi}
-\end{align*}
-\noindent Let's take a second and discuss some the moves we made here.
-First of all, we know that since the function is even, we can say that the the integral of the entire range is really twice the integral of half the range.
-Since we are dealing with the cosine function, we actually have four times a quarter of the range.
-If we take a look at a graph, this may make a little more sense:\\
-\begin{center}
-\begin{tabular}{c c}
-\includegraphics[scale=0.5]{cos_abs} & \includegraphics[scale=0.5]{cos}\\
-$f(x) = |\cos{x}|$ on $[0,2\pi]$ & $f(x) = \cos{x}$ on $[0,2\pi]$
-\end{tabular}
-\end{center}
-\noindent\\ You can see that the section of the graph from $[\frac{\pi}{2},\frac{3 \pi}{2}]$ is positive for the $|\cos{x}|$, but negative for $\cos{x}$.
-If we simply integrated the $|\cos{x}|$ over $[0,\pi]$, the result would be 0, but graphically this result would not make any sense.
-Now let's take a look at the $b_{n}$ term.
-Before we jump into the integration, let's think about this function:
-\[\left|\cos{x}\right| = 
-\begin{cases*}
-\cos(x) \qquad if\ 0\leq x \leq \frac{\pi}{2}\\
--\cos(x)\quad if\ \frac{\pi}{2}\ \leq x \leq \pi
-\end{cases*}
-\]
-\noindent Now we can set up the formula for the $b_{n}$ term. We have as follows:
-\begin{align*}
-b_{n} &= \frac{2}{\pi}\left[ \int_{0}^{\frac{\pi}{2}} \cos{x}\cos{nx}dx - \int_{\frac{\pi}{2}}^{\pi} \cos{x}\cos{nx}dx \right]\\
-&= \frac{2}{\pi} \left[ \frac{-2 \cos{\left(\frac{n\pi}{2}\right)} + \sin{\left(n\pi\right)}}{n^{2} - 1}\right]\\
-\end{align*}
-\noindent Note that (as we have seen before) that the $\sin$ term disappears; however, the more interesting part is what happens with our $\cos$ term.
-The value of $\cos{\left(\frac{n\pi}{2}\right)}$ will be 0 at all odd $n$, or that is to say for all $n = 2k + 1$.
-So, we will let our $n = 2k$.
-This gives us:
-\begin{align*}
-b_{n} &= \frac{2}{\pi} \left[ \frac{-2 \cos{\left(k \pi \right)}}{(2k)^{2} - 1}\right]\\
-&= -\frac{4}{\pi}\left[\frac{(-1)^{k}}{(2k)^{2}-1}\right]
-\end{align*}
-\noindent And that leaves us with the final result
-\[
-f(x)\sim \frac{2}{\pi} - \sum_{k=0}^{\infty}\left[\frac{4}{\pi}\left[\frac{(-1)^{k}}{(2k)^{2}-1}\right] \cos{\left(\frac{2k\pi}{2}\right)}\right]
-\]
-\newpage
-
-\indent Let's recall the 2nd order PDE that we found a series solution for earlier. Now we want to apply the initial condition of $u(x,0) = x$.\\
-\noindent\textbf{\textit{Ex:}} Find the particular solution for the following PDE:
+We have the following system
 \[
 	\begin{cases}
-\frac{\partial u}{\partial t} = k \frac{\partial^{2} u}{\partial x^{2}}\\
-u(x,0) = x\\
-u(0,t) = u(L,t) = 0\\
+	y'+ay=0\\
+	y(0)=\alpha
 	\end{cases}
 \]
-\indent\textbf{\textit{Solution:}} Earlier, we found a general solution of the form
-\begin{align*}
-u_{n}(x,t) &= a_{n}\sin{\left(\frac{n\pi x}{L}\right)}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}\\
-u(x,t) &= \sum_{n = 1}^{\infty}\left[a_{n}\sin{\left(\frac{n\pi x}{L}\right)}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}\right]
-\end{align*}
-\noindent Without loss of generality, let's let $L = \pi$.
-This would change our boundary condition to $u(\pi,t) = 0$, but this wouldn't change the solution that we have here in any way other than replacing the $L$'s in the equation with $\pi$.
-We will also let $k=1$.
-Again this doesn't change anything other than the fact that we will replace $k$ with 1.
-We now have:
-\[
-u(x,t) = \sum_{n = 1}^{\infty}\left[a_{n}\sin{\left(nx\right)}e^{-\left(n\right)^{2}t}\right]
-\]
-\noindent This is a much nicer sequence for us to deal with.
-Our goal now is to find $a_{n}$, which will give us a particular solution.
-We know that our answer for the particular solution will be in the form of a series, since our initial condition is not represented in the form of terms of a series (we will see an example of this later).
-Let's go ahead and apply our initial condition.
-Doing so will yield:
-\[
-x = \sum_{n = 0}^{\infty}\left[a_{n}\sin{nx}\right]
-\]
-\noindent Now, we need to come up with an $a_{n}$. We have a Fourier Sine Series here, so we can use the formula for finding $a_{n}$ that corresponds to our case.
-\begin{align*}
-a_{n} &= \frac{2}{\pi}\int_{0}^{\pi} x\sin{\left(nx\right)}dx\\
-&= -\frac{2n\pi \cos{\left(n\pi\right)} + 2\sin{\left(n\pi\right)}}{n^{2}\pi}\\
-&= -\frac{2(-1)^{n}}{n}
-\end{align*}
-\noindent So our particular solution is:
-\[
-u(x,t) = \sum_{n = 1}^{\infty}\left[-\frac{2(-1)^{n}}{n}\sin{\left(n x\right)}e^{-\left(n\right)^{2}t}\right]
-\]
-\noindent And we are done.
 
-\newpage
-\noindent\textbf{\textit{Ex:}} Solve the following boundary value problem given the general solution:
-\[
-\begin{cases}
-u_{tt} - u_{xx} = 0\\
-u(0,t) = 0\qquad u(1,t) = 0\\
-u(x,0) = f(x)\qquad u_{t}(x,0) = g(x)
-\end{cases}\\
-\]
-\[f(x) = 2\sin{(\pi x)} + 3\sin{(2\pi x)}\qquad g(x) = 4\sin{(3\pi x)} - 7\sin{(5\pi x)}\]
-\[u(x,t) = \sum_{n=1}^{\infty}\left[\left(a_{n}\cos{[n\pi t]} + b_{n}\sin{[n\pi t]}\right)\sin{[n\pi x]}\right]\]
-\indent\textbf{\textit{Solution:}} There's a lot going on here, so let's break it down.
-First of all, notice that we have two initial conditions, one of which is a derivative.
-This should make sense since the equation have two second order partial derivatives included in it.
-Next notice that we have a Fourier Sine Series, despite both the $\sin{[n\pi t]}$ and $\cos{[n\pi t]}$ terms.
-If we had worked out the entire problem, we would see that our eigenvalue/function for this equation takes the form of
-\[
-\{\lambda_{n} = n\pi, \sin{(n\pi x)}\}
-\]
-\noindent which tells us that the corresponding series is a Fourier Sine Series.
-We also see the initial conditions on our equation are already representative of a series of sine functions, which is not just a conincidence.
-To apply the initial conditions, we will need both the general solution and the derivative with respect to $t$ of the general solution.
-Let's go ahead and apply the first initial condition, $u(x,0) = f(x)$.
-Letting $t= 0$ gives us:
-\[
-u(x,0) = 2\sin{(\pi x)} + 3\sin{(2\pi x)} = \sum_{n=1}^{\infty}\left[a_{n}\sin{(n\pi x)}\right]
-\]
-\noindent We see that the term with $b_{n}$ has dropped out, since $\sin{(0)} = 0$.
-Now we just have to match up the coefficients on $a_{n}$.
-We can see that when we have $n=1$, $a_{n} = 2$, and when $n=2$, $a_{n} = 3$.
-For all other $n$, we have $a_{n} = 0$, so we ignore those terms.
-Next, we can apply the initial condition for the derivative of the general solution with respect to $t$.
-We expect that all the terms related to $a_{n}$ will disappear, and in fact they will.
-Let's take a look:
-\begin{align*}
-u_{t}(x,t) &= \sum_{n=1}^{\infty}\left[(-a_{n}n\pi\sin{[n\pi t]} + b_{n}n\pi\cos{[n\pi t]})\sin{n\pi x}\right]\\
-u_{t}(x,0) &= 4\sin{(3\pi x)} - 7\sin{(5\pi x)} = \sum_{n=1}^{\infty}\left[n\pi b_{n}\sin{(n\pi x)}\right]
-\end{align*}
-\noindent We are going to do the same thing here that we did to find our $a_{n}$ terms, but notice that we don't just have $b_{n}$; we actually have $n\pi b_{n}$. So we have the following:
-\begin{alignat*}{3}
-&3\pi b_{3} = 4 \qquad \qquad \qquad &&5\pi b_{5} = -7\\
-&b_{3} = \frac{4}{3\pi} &&b_{5} = -\frac{7}{5\pi}
-\end{alignat*}
-\noindent Much like the solutions for $a_{n}$, all the other $b_{n}$ terms are 0.
-Now we can construct a particular solution.
-In this case, the answer will not be in the form of a series, since we have definite terms.
-Our solution is as follows:
-\[
-u(x,t) = 2\cos{(\pi t)}\sin{(\pi x)} + 3\cos{(2\pi t)}\sin{(2\pi x)} + \frac{4}{3\pi}\sin{(3\pi t)}\sin{(3\pi x)} - \frac{7}{5\pi}\sin{(5\pi t)}\sin{(5\pi x)}
-\]
-\noindent And we are done.
-\newpage
-
-\subsection{Convergence of Fourier Series}
-\indent Up to this point, we have found a Fourier series and left our solutions at that. We will now look into the convergence of a Fourier series. This topic includes two questions:
-\begin{enumerate}
-\item Does the series converge?
-\item If the series converges, to what value does it converge?
-\end{enumerate}
-\noindent Both of these questions are highly theoretical. We will discuss the answers to them, but first we need to go over some terms:
-\begin{enumerate}
-\item \textit{\textbf{Jump Discontinuity:}} To say that $f(x)$ has a jump discontinuity at $x = a$ is to say that
-\[
-\lim_{x \to a^{-}}f(x) \neq \lim_{x \to a^{+}}f(x)
-\]
-Where both the limit from the left and the limit from the right exist.
-\item \textit{\textbf{Piecewise Smooth:}} A function $f(x)$ is said to be piecewise smooth if the function can be broken into two distinct intervals, and on each interval both $f(x)$ and $f^{'}(x)$ are continuous.
-\item \textit{\textbf{Periodic Extension:}} The periodic extension of $f(x)$ on $[-L,L]$ is the repetition of the function on intervals on periods to the left and right of $[-L,L]$.
-\end{enumerate}
-\indent With these definitions, we can lay out a general rule for convergence of a Fourier series.
-We won't give a proof of these rules for the time being.
-We just need them to lay out some groundwork for the rest of the information we will present on Fourier series.
-We have the following:
-\begin{quote}
-Suppose $f(x)$ is a function, and that $f(x)$ is piecewise smooth on the interval $[-L,L]$. The corresponding Fourier series for $f(x)$ converges to either
-\begin{enumerate}
-\item the periodic extension of $f(x)$, if $f(x)$ is continuous, or
-\item the average of the two one-sided limits, if the periodic extension of $f(x)$ has a jump discontinuity at $x = a$.
-\end{enumerate}
-\end{quote}
-\noindent We should note that the second condition can be expressed by the formula
-\[\frac{1}{2}\left[\lim_{x \to a^{-}}f(x) + \lim_{x \to a^{+}}f(x)\right]\]
-which gives us a nice way to calculate the value to which the series converges if we happen to meet the criteria in (2). Let's take a look at some examples.\\\\
-
-\noindent\textbf{\textit{Ex:}} Find the Fourier series for the following equation, and determine if the series converges. If it does, tell where the series converges to.
-\[ f(x) =
-\begin{cases}
-L \qquad\qquad -L \leq x \leq 0\\
-2x \qquad\qquad 0 \leq x \leq L
-\end{cases}
-\]
-\indent\textbf{\textit{Solution:}} We will go ahead and skip the details of finding the Fourier series for this function.
-If you do want to find it, you will have to find a full piecewise Fourier series.
-For this particular $f(x)$, we have:
-\begin{align*}
-f(x) &\sim L + \sum_{n=1}^{\infty}\left[\frac{2L}{n^{2}\pi^{2}}([-1]^{n}-1)\cos{\left(\frac{n\pi x}{L}\right)}\right] - \sum_{n=1}^{\infty}\left[\frac{L}{n\pi}(1 + [-1]^{n})\sin{\left(\frac{n\pi x}{L}\right)}\right]
-\end{align*}
-\noindent You'll notice that we didn't simplify this solution to a nicer form.
-If we evaluate this series, we will run across terms where the series is 0, but that is okay for now.
-You should also notice that this function is defined as a piecewise function, and on both intervals ($[-L,0]$ and $[0,L]$), $f(x)$ is continuous.
-In other words, $f(x)$ is piecewise smooth, so we know that this series converges.
-Now let's take a look at the graph of $f(x)$ on an arbitrary interval of $[-1,1]$.
-\begin{center}
-\begin{tabular}{c}
-\includegraphics[scale=0.6]{pw_graph_01}\\
-$
-f(x) =
-\begin{cases}
-1 \qquad\qquad -1 \leq x \leq 0\\
-2x \qquad\qquad 0 \leq x \leq 1
-\end{cases}
-$
-\end{tabular}
-\end{center}
-\noindent We can see that there is a jump in the graph at $x = 0$.
-In other words, there is a jump discontinuity at $x = 0$, and that means that we need to use (2) in order to find the value of convergence.
-We will let $\sum$ stand in for our entire series.
-We have:
-\begin{align*}
-\sum &\to \frac{1}{2}\left[\lim_{x\to 0}L + \lim_{x\to0}2x\right]\\
-&\to \frac{1}{2}[L + 0]\\
-&\to \frac{L}{2}
-\end{align*}
-\noindent So, at $x=0$, the series converges to $\frac{L}{2}$.
-We are not done yet, however.
-Now, we need to consider the periodic extension of the series.
-The function will repeat every $[-L,L]$, so we need to look at what the convergence of the series looks like at $-L$ and $L$.
-Let's take a look:
-\begin{alignat*}{3}
-\sum&\to\frac{1}{2}\left[\lim_{x\to-L^{-}}f(x) + \lim_{x\to-L^{+}}f(x)\right] \qquad\qquad \sum&&\to \frac{1}{2}\left[\lim_{x\to L^{-}}f(x) + \lim_{x\to L^{+}}f(x)\right]\\
-&\to \frac{1}{2}[2L + L] &&\to \frac{1}{2}[2L + L]\\
-&\to\frac{3L}{2} &&\to\frac{3L}{2}
-\end{alignat*}
-\noindent And we are done.
-\newpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Second Order Partial Differential Equations}
-\hrule
-\noindent\\
-\subsection{The Heat and Wave Equations}
-\indent Let's take a look at the second order PDE.
-We will start with a homogeneous equation, which we will tackle with the separation of variables method.
-The general strategy is that we want to take our PDE and convert it to an ODE, which we know how to solve.
-The problems presented here also involve boundary conditions, which will require us to solve for eigenvalues/functions.
-We will also be dealing with series (specifically Fourier Series) to help us solve these problems.
-Let's start by looking at the homogeneous heat equation.\\
-\noindent \textbf{\textit{Ex:}} Solve the following PDE
-\[
-\begin{cases}
-\frac{\partial u}{\partial t} = k \frac{\partial^{2} u}{\partial x^{2}}\\
-u(x,0) = x\\
-u(0,t) = u(L,t) = 0\\
-\end{cases}
-\]
-\indent \textbf{\textit{Solution:}} We will use the separation of variable technique to solve this problem.
-We know to use separation of variable because we are dealing with a homogeneous PDE with homogenous boundary conditions.
-Separation of variables tells us that our solution will be in the form of $ u(x,t) = X(x)T(t) $.
-If we take this fact at face value, we can use the separation of variables method to solve this.
-First, we substitute in our solution into the PDE:
-\begin{align*}
-\frac{\partial}{\partial t}(X(x)T(t)) &= k \frac{\partial^{2}}{\partial x^{2}}(X(x)T(t))\\
-X(x)\frac{dT}{dt} &= k\frac{d^{2}X}{dx^{2}}T(t)\\
-\frac{dT}{dt}\frac{1}{T(t)} &= \frac{d^{2}X}{dx^{2}}\frac{k}{X(x)}
-\end{align*}
-\noindent The only way that our function of x and our function of t can be equal is if they are equal to a constant;
-thus we now have:
-\[
-\frac{dT}{dt}\frac{1}{T(t)} = \frac{d^{2}X}{dx^{2}}\frac{k}{X(x)} = -\lambda
-\]
-\noindent And now we have a first order ODE for the time and a second order ODE for the position:
-\[
-\frac{dT}{dt} = -k \lambda T(t) \qquad\text{and}\qquad \frac{d^{2}X}{dx^{2}} = -\lambda X(x)
-\]
-\noindent Our next step is to find the Eigenvalues and Eigenfunctions of the second order ODE; we will come back to the time ODE when we have values for $\lambda$.\\ 
-\indent To solve for these values, we have to find non-trivial eigenvalues that correspond to the boundary values of the ODE.
-In this case, we know that $X(0) = 0$ and $X(L) = 0$.
-These boundary conditions come from the BC on the PDE which states $u(0,t) = u(L,t) = 0$.
-To find the eigenvalues/functions, we need to find the characteristic equation corresponding to the ODE in question.
-Let's go ahead and find it:
-\begin{gather*}
-\frac{d^{2}X}{dx^{2}} = -\lambda \frac{X(x)}{k}\\
-\frac{d^{2}X}{dx^{2}}  + \lambda \frac{X(x)}{k} = 0\\
-r^{2} + \lambda = 0\\
-r^{2} = -\lambda\\
-r = \pm \sqrt{- \lambda}
-\end{gather*}
-\indent Now, we can substitute values of $\lambda$ to find out solutions.
-Let's start with $\lambda > 0$.
-Since we have a $-\lambda$ under the radical, we will have $r = \pm \sqrt{\lambda} \mathrm{i}$.
-This will yield the solution of
-\[
-y(x) = c_{1}\cos{\sqrt{\lambda}x} + c_{2}\sin{\sqrt{\lambda}x}
-\]
-\noindent Applying our boundary conditions gives us:
-\[
-y(0) = c_{1}\cos{0} + c_{2}\sin{0} = 0
-\]
-\noindent Now we know that $c_{1} = 0$.
-This result came from the fact that $\sin{0} = 0$, so the $c_{2}$ term disappears.
-Now we are left with $\cos{0} = 1$, which tells us that $c_{1} = 0$.
-We can apply this to our second boundary condition like so:
-\begin{gather*}
-y(L) = 0 = c_{2}\sin{\sqrt{\lambda}L}\\
-\end{gather*}
-\noindent It looks like there will only be trivial solutions, but we know this isn't true due to the Jacobian matrix.
-Recall that if the determinant of the Jacobian $\neq 0$, then we only have trivial solutions, and it the determinant of the Jacobian $= 0$, then we do not have the trivial solution.
-Let's take a look at the Jacobian for this case:
-\begin{gather*}
-\left| 
-\begin{array}{cc}
-1 & 0\\
-\cos{\sqrt{\lambda}L} & \sin{\sqrt{\lambda}L}
-\end{array}
-\right| = 0\\
-\sin{\sqrt{\lambda}L} = 0\\
-\sqrt{\lambda}L = \arcsin{0}\\
-\sqrt{\lambda} = \frac{\arcsin{0}}{L}\\
-\lambda = \left(\frac{\arcsin{0}}{L}\right)^{2}\\
-\lambda = \left(\frac{n\pi}{L}\right)^{2}
-\end{gather*}
-\noindent Notice that we replaced the $\arcsin{0}$ term with $n\pi$.
-Recall from trig that the $\sin$ function is $0$ at every $n\pi$, starting with $n = 0$.
-Next, we just need to find our corresponding eigenfunction.
-Our $c_{1}$ was $0$, so we can ignore the entire $\cos$ term.
-Thus, we have an eigenfunction from plugging in our $\lambda$ that looks like:
-\[
-X(x) = \sin{\frac{n\pi x}{L}}
-\]
-\indent Unfortunately, we aren't done yet.
-Now we have to find the corresponding eigenvalues/functions for $\lambda = 0$.
-This isn't too hard, and can be done in a few lines if we recall the characteristic equation.
-Plugging in $0$ for $\lambda$ gives us that $r = 0$.
-Using this piece of information, we now know that $y(x) = c_{1} + xc_{2}$.
-Applying our boundary conditions gives us $c_{1} = 0, c_{2} = 0$.
-I'll leave you to do the math (try pluggin into the Jacobian; you'll see that there is no way to force the determinant to be 0) , but we can see that we have only trivial solutions for $\lambda = 0$.\\
-\indent Our final case is for $\lambda < 0$.
-I'll go ahead and tell you that this will yield another trivial solution, but we can take a look at the work behind it.
-Since $\lambda > 0$, we know that our $r$ from the characteristic equation will be $r = \pm \sqrt{\lambda}$.
-From here we are going to do something a little unconventional that will help us.
-Usually, we would have a solution of the form $y(x) = c_{1}e^{\sqrt{\lambda}x} + c_{2}x e^{-\sqrt{\lambda}x}$.
-In our case, it will be more convenient to use hyperbolic trig functions.
-It's not too far off from our solution from $\lambda > 0$:
-\begin{align*}
-y(x) &= c_{1}e^{\sqrt{\lambda}x} + c_{2}e^{-\sqrt{\lambda}x}\\
-&= c_{1}\cosh{\sqrt{\lambda}x} + c_{2}\sinh{\sqrt{\lambda}x}
-\end{align*}
-\noindent I'll leave you to apply the boundary conditions (remember that $\sinh{0} = 0$ and $\cosh{0} = 1$), but you will find that you have only trivial solutions.\\
-\indent So, we are done with the second order ODE, and now we turn our eye to the first order time equation. We know our $\lambda$, so it isn't too hard to solve:
-\begin{gather*}
-\frac{dT}{dt} = -\lambda T(t)\\
-T(t) = ce^{-k\left(\frac{n\pi}{L}\right)^{2}t}
-\end{gather*}
-\indent Finally we can put it all together. Our penultimate solution will be 
-\[
-u_{n}(x,t) = B_{n}\sin{\left(\frac{n\pi x}{L}\right)}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}
-\]
-\noindent We can analyze this solution and tell where each piece comes from.
-The $B_{n}$ is a coefficient that results from our $c's$ in the general solutions of the differential equations.
-The $sin$ part results from out eigen vector.
-Remember, we get a different value for each $n$, so there may or may not be $0$ terms in the series.
-Finally, we have the exponential term.
-This term comes from t he solution to our time equation.
-When we put it all together, we can see that this is, of course a series; for every $n$, we get a different solution.
-From here we can find the Fourier Series (which will be a Fourier Sine Series).
-We won't worry about that yet, but we will come back to it.
-\newpage
-
-\indent It will benefit us to take a moment and discuss how boundary conditions effect our Fourier series.
-We have three different combinations of boundary conditions, where each one will produce a different result in our series solution.
-\begin{enumerate}
-\item \textit{Dirichlet Boundary Conditions}\\
-Dirichlet boundary conditions involve the function for which we are looking.
-In other words, the boundary conditions are presented in terms of the functions we are trying to find, rather than the derivative of that function.
-Typically, this type of boundary condition will take the form of (in the homogeneous case)
-\[u(0,t) = u(L,t) = 0.\]
-In regards to our Fourier series, we expect to see $\sin{(\frac{n\pi x}{L})}$. Our basis will look like
-\[
-\left\{ \lambda_{n} = \left(\frac{n\pi}{2}\right)^{2}, \phi_{n}(x) = \sin{\left(\frac{n\pi x}{L}\right)} \right\}
-\]
-\item \textit{Neumann Boundary Conditions}\\
-Neumann boundary conditions involve the derivatives of the function for which we are solving. Neumann boundary conditions will look like
-\[u_{x}(0,t) = u_{x}(L,t) = 0.\]
-This will cause our Fourier series to contain $\cos{(\frac{n\pi x}{L})}$. Out basis will look like
-\[
-\left\{ \lambda_{n} = \left(\frac{n\pi}{2}\right)^{2}, \phi_{n}(x) = \cos{\left(\frac{n\pi x}{L}\right)} \right\}
-\]
-\item \textit{Robin Boundary Conditions}\\
-Robin boundary conditions will contain a mixture of the function and it's derivative. Typically, we will see
-\[ u_{x}(0,t) = u(L,t) = 0\qquad \text{or}\qquad u(0,t) = u_{x}(L,0) = 0.\]
-These boundary conditions will give us two different results, depending on whether we have the derivative first or second. We will have
-\begin{align*}
-u_{x}(0,t) = u(L,t) = 0 &\implies \left\{\lambda_{n} = \left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}\right)^{2}, \phi_{n}(x) = \cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}x\right)}\right\}\\
-u(0,t) = u_{x}(L,t) = 0 &\implies \left\{\lambda_{n} = \left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}\right)^{2}, \phi_{n}(x) =\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}x\right)}\right\}
-\end{align*}
-\end{enumerate}
-
-\indent Now we can take a look at a slightly harder PDE.
-We still have the heat equation, but we will not have homogeneous boundary conditions.
-Our goal here is to find a way to transform our boundary conditions into a homogenous form and take it from there.\\\\
-\noindent \textbf{\textit{Ex:}} Solve the following PDE
-\[
-\begin{cases*}
-u_{t} - u_{xx} = 0\\
-u(0,t) = 0,\ u(1,t) = t\\
-u(x,0) = x^{2}
-\end{cases*}
-\]
-\indent \textbf{\textit{Solution:}} We would like to just jump into this PDE and solve it, but it's not quite that simple.
-We have to find a way to express our boundary conditions in a homogeneous manner.
-We can do this through a clever substitution; we will create a new PDE, denoted $\Upsilon$.
-To find $\Upsilon$, we use the following formula:
-\[ \Upsilon = B_{1} + \frac{x}{l}\Delta B, \]
-\noindent where $B$ is a boundary condition. In our case, we have as follows:
-\begin{align*}
-\Upsilon &= B_{1} + \frac{x}{l}\Delta B\\
-&= 0 + \frac{x}{1}(t - 0)\\
-&= xt
-\end{align*}
-\noindent Now, we find a new PDE, denoted by $\upsilon$. We have another formula that tells us
-\[ \upsilon(x,t) = u(x,t) - \Upsilon(x,t) \]
-\noindent Applying this formula gives us
-\begin{align*}
-\upsilon(x,t) &= u(x,t) - \Upsilon\\
-&= u(x,t) - xt
-\end{align*}
-\noindent Let's take a moment to analyze the result.
-We wanted to find a new PDE in terms of our old one that would allow us to have homogeneous boundary conditions.
-We found a solution, called $\upsilon$, in terms of our original PDE, called $u$.
-Our original PDE was $u_{t} - u_{xx} = 0$, so to properly perform our substitution, we need to find $\upsilon_{t}$ and $\upsilon_{xx}$.
-Through some careful differentiation, we have
-\[ \upsilon_{t} = u_{t} - x \qquad\text{and}\qquad \upsilon_{xx} = u_{xx}\]
-\noindent To find our new PDE, we simply evaluate.
-\[
-\upsilon_{t} - \upsilon_{xx} = u_{t} - x - u_{xx} = u_{t} - u_{xx} - x = 0 - x = -x
-\]
-\noindent So we finally have a new PDE that looks like
-\[
-\begin{cases*}
-\upsilon_{t} - \upsilon_{xx} = -x\\
-\upsilon(0,t) = \upsilon(1,t) = 0\\
-\upsilon(x,0) = x^{2}
-\end{cases*}
-\]
-\noindent You may wonder how this has made anything simpler, since the PDE itself is no longer homogeneous.
-By forcing the boundary conditions to be homogeneous, we can now use a Fourier series to find our solution.
-Recall that earlier we discussed the boundary condition and their resulting eigenvalue/eigenfunction.
-Here, we have Dirichlet boundary conditions, which tells us that we have a Fourier sine series.
-We have that
-\begin{align*}
-\upsilon(x,t) &= \sum_{n = 1}^{\infty}\left[a_{n}(t)\sin{(n\pi x)}\right]\\
-\upsilon_{t}(x,t) &= \sum_{n = 1}^{\infty}\left[a_{n}'(t)\sin{(n\pi x)}\right]\\
-\upsilon_{xx}(x,t) &= -\sum_{n = 1}^{\infty}\left[a_{n}(t)(n\pi)^{2}\sin{(n\pi x)}\right]\\ 
-\end{align*}
-\noindent Which tells us that
-\[
-\upsilon_{t} - \upsilon_{xx} = \sum_{n = 1}^{\infty}\left[(a_{n}'(t) + (n\pi)^{2}a_{n}(t))\sin{(n\pi x)}\right] = -x
-\]
-\noindent This may look rather complicated, but all we have is a function's Fourier sine series. Essentially, we have
-\[
--x = \sum_{n = 1}^{\infty}\left[ b_{n}\sin{(n\pi x)} \right]
-\]
-\noindent To find $b_{n}$, we use the formula that was discussed in the earlier section about Fourier series. We have that
-\begin{align*}
-a_{n}'(t) + (n\pi)^{2}a_{n}(t) &= \int_{0}^{1}-x\sin{(n\pi x)}dx\\
-&= \frac{2x}{n\pi}\cos{(n\pi x)}\bigg\vert_{0}^{1}\\
-&= \frac{2}{n\pi}
-\end{align*}
-\noindent We now have an ODE which we can solve for $a_{n}$.
-Note that from out PDE, we do have an initial condition for the ODE, but it is not as simple as just saying $a_{n}(0) = x^{2}$.
-At $t = 0$, we have the following series:
-\[
-\sum_{n=1}^{\infty} a_{n}(0)\sin{(n\pi x)}
-\]
-\noindent We got this by taking our original $\upsilon(x,t)$ and plugging in $0$ for $t$. We know how to get $a_{n}(0)$, since this is just a simple Fourier sine series. It is given by
-\begin{align*}
-a_{n}(0) &= 2\int_{0}^{1}x^{2}\sin{(n\pi x)}dx\\
-&= \frac{(-1)^{(n + 1)}}{n\pi} + \frac{2}{(n\pi)^{3}}\left[ (-1)^{n} - 1 \right]
-\end{align*}
-\noindent We now have all we need to solve the ODE. We will omit the work for the solution, but the final answer for $a_{n}$ is 
-\[
-a_{n}(t) = \frac{4\left[ (-1)^{n} - 1 \right]}{(n\pi)^{3}}e^{-(n\pi)^{2}t} + \frac{2(-1)^{n}}{(n\pi)^{3}}
-\]
-\noindent Now all that's left to do is go back to $u$ from $\upsilon$.
-This may seem hard, but we saw earlier that $\upsilon = u - \Upsilon$.
-Solving for $u$ gives us $u = \upsilon + \Upsilon$.
-Recall that $\Upsilon$ was $xt$, and $\upsilon$ was the series for which we just found $a_{n}$.
-Putting it all together gives us
-\[
-u(x,t) = xt + \sum_{n=1}^{\infty}\left[\frac{4\left[ (-1)^{n} - 1 \right]}{(n\pi)^{3}}e^{-(n\pi)^{2}t} + \frac{2(-1)^{n}}{(n\pi)^{3}}\right]\sin{(n\pi x)}
-\]
-\noindent And we are done.
-\newpage
-
-\noindent \textbf{\textit{Ex:}} Solve the following PDE
-\[
-\begin{cases*}
-u_{tt} - u_{xx} = 1 + t\cos{(2\pi x)}\\
-u_{x}(0,t) = u_{x}(1,t) = 0\\
-u(x,0) = 2\cos{(2\pi x)},\quad u_{t}(x,0) = 1 + \cos{(\pi x)} - \cos{(2\pi x)}
-\end{cases*}
-\]
-\indent \textbf{\textit{Solution:}} This problem is an example of the wave equation with a non-homogeneous solution.
-We are lucky this time; the boundary conditions are already homogeneous, so there is no need to change our variable.
-We know that since both of the boundary conditions are derivatives, we are dealing with a Fourier cosine series.
-This yields the eigenvalue/function of
-\[
-\left\{\lambda_{n} = (n\pi)^{2}, \cos{(n\pi x)}\right\}
-\]
-Since we are dealing with a cosine series, we also have a leading term, $a_{0}$. Our full series looks like
-\[
-u(x,t) = a_{0}(t) + \sum_{n=1}^{\infty} a_{n}(t)\cos{(n\pi x)}
-\]
-We need $u_{tt}$ and $u_{xx}$, so we have:
-\[
-u_{tt}(x,t) = a_{0}''(t) + \sum_{n=1}^{\infty} a_{n}''(t)\cos{(n\pi x)} 
-\qquad \text{and} \qquad 
-u_{xx}(x,t) = -\sum_{n=1}^{\infty} (n\pi)^{2}a_{n}(t)\cos{(n\pi x)}
-\]
-All we have to do now is simply plug into our original equation. This gives us
-\begin{align*}
-a_{0}''(t) + \sum_{n=1}^{\infty} a_{n}''(t)\cos{(n\pi x)} + \sum_{n=1}^{\infty} (n\pi)^{2}a_{n}(t)\cos{(n\pi x)} &= 1 + t\cos{(2\pi x)}\\
-a_{0}''(t) + \sum_{n=1}^{\infty}(a_{n}''(t) + (n\pi)^{2}a_{n}(t))\cos{(n\pi x)} &= 1 + t\cos{(2\pi x)}
-\end{align*}
-Now we can play the matching game. We have a $constant + series\ term$ on the right side, and $constant + series$ on the left, so we can match up the constants. This gives us 
-\[
-a_{0}''(t) = 1
-\]
-Now we match up the series term. On the right hand side, we have $\cos{(2\pi x)}$, so we know that for $n= 2$, we have
-\[
-a_{2}''(t) + 4\pi^{2}a_{2}(t) = t
-\]
-$a_{0}$ and $a_{2}$ are the only terms that appear, so we can assume for now that $\forall n \neq 2\ \text{and}\ 3, a_{n} = 0$.
-We are left with two ODE's, so we now need to find the corresponding initial conditions.
-Let's apply the first initial condition to our PDE.
-This gives us
-\[a_{0}(0) + \sum_{n=1}^{\infty}a_{n}(0)\cos{(n\pi x)} = 2\cos{(2\pi t)}\]
-Which tells us
-\[a_{0} = 0 \qquad\qquad \qquad\qquad a_{2} = 2\]
-Now we can apply the second initial condition. After taking the partial derivative of $u$ with respect to $t$, we have
-\[
-a_{0}'(0) + \sum_{n=1}^{\infty}a_{n}'(0)\cos{(n\pi x)} = 1 + \cos{(\pi x)} - \cos{(2\pi x)}
-\]
-Some more clever pattern matching give us
-\[
-a_{0}'(0) = 1 \qquad\qquad a_{1}'(0) = 1 \qquad\qquad a_{2}'(0) = -1
-\]
-Notice that we have a term $a_{1}'$.
-We cannot simply ignore this value; rather we have to take it into account when we find our solution.
-We know that we have an ODE for $a_{1}(t)$, but it may not be evident how to express it.
-Recall that earlier we said $\forall n \neq 2\ \text{and}\ 3, a_{n} = 0$.
-This means that $a_{1} = 0$, and since $a_{1}$ is a coefficient in the Fourier series, we know that $a_{1}''(t) + \pi^{2}a_{1}(t) = 0$.
-We now have three different ODE's that we must solve:
-\begin{alignat*}{3}
-&\begin{cases}
-a_{0}''(t) = 1\\
-a_{0}'(0) = 1,\quad a_{0}(0) = 1
-\end{cases}
-&\begin{cases}
-a_{1}''(t) + \pi^{2}a_{1}(t) = 0\\
-a_{1}'(0) = 1,\quad a_{1}(0) = 0
-\end{cases}
-&&\begin{cases}
-a_{2}''(t) + 4\pi^{2}a_{2}(t) = t\\
-a_{2}'(0) = 1,\quad a_{2}(0) = 2
-\end{cases}
-\end{alignat*}
-Luckily, Mathematica is here to save the day. We will skip the work and let the computer do the thinking for us. We get that
-\begin{align*}
-a_{0}(t) &= \frac{1}{2}t^{2} + t\\
-a_{1}(t) &= \frac{\pi\cos{(t\pi)} + \sin{(t\pi)}}{\pi}\\
-a_{2}(t) &= \frac{2\pi t + 16\pi^{3}\cos{(2\pi t)} - \sin{(2\pi t)} + 4\pi^{2}\sin{(2\pi t)}}{8\pi^{3}}
-\end{align*}
-Now all we need to do is plug back in to our $u$. Our final answer becomes
-\[
-u(x,t) = \frac{1}{2}t^{2} + t + \frac{\pi\cos{(t\pi)} + \sin{(t\pi)}}{\pi}\cos{(\pi x)} +  \frac{2\pi t + 16\pi^{3}\cos{(2\pi t)} - \sin{(2\pi t)} + 4\pi^{2}\sin{(2\pi t)}}{8\pi^{3}}\cos{(2\pi x)}
-\]
-Just for interest, the plot of this solution looks like
-\begin{center}
-\includegraphics[scale=0.6]{pg23_plot} 
-\end{center}
-On $-\pi \leq \theta \leq \pi$ and $-\pi \leq x \leq \pi$; and we are done.
-\newpage
-
-\subsection{The Laplace Equation}
-\noindent \\
-\indent Up to this point, we have only looked at the wave equation and the heat equation.
-There is, however, one other equation in the classical trinity: the Laplace Equation.
-We haven't dealt with this equation yet because we have to be very careful in our approach to solving it.
-When we looked at the homogeneous wave and heat equations, we used the general approach called \textit{Separation of Variables}.
-In the case of the homogeneous Laplace equation, we may not be able to tackle the problem using separation of variables.
-The problem arises from the various types of boundaries that effect the domain.
-We can either have a 'rectangular' bound, which will be separable, or a 'circular' bound which is \textbf{not} separable.
-Imagine that we have the following two problems:
-\begin{align*}
-&A =\begin{cases}
-u_{xx} + u_{yy} = 0\quad \text{on}\ [0,1]\times[0,2]\\
-u(x,0) = u(x,2) = 0\\
-u(0,y) = y, u(1,y) = 0
-\end{cases}
-&B =\begin{cases}
-u_{xx} + u_{yy} = 0\quad \text{on}\ x^{2} + y^{2} < a^{2}\\
-u(x,y)\ \big |_{x^{2} + y^{2} = a^{2}} = f(x,y)
-\end{cases}
-\end{align*}
-$A$ is an example of a separable Laplace equation.
-We can see graphically that the domain $[0,1] \times [0,2]$ is rectangular.
-$B$'s domain is not separable, since it is circular.
-Let's take a look at $A$.\\\\
-\noindent \textbf{\textit{Ex:}} Solve the following PDE
-\[
-\begin{cases*}
-u_{xx} + u_{yy} = 0\quad \text{on}\ [0,1]\times[0,2]\\
-u(x,0) = u(x,2) = 0\\
-u(0,y) = y, u(1,y) = 0
-\end{cases*}
-\]
-\indent \textbf{\textit{Solution:}} Earlier, we discussed the different boundary conditions we can encounter and their corresponding eigenvalues/functions.
-In this problem, we have Dirichlet boundary conditions.
-This tells us that
-\[
-\left\{\lambda_{n} = \left(\frac{n\pi}{2}\right)^{2}, \phi_{n}(y) = \sin{\left(\frac{n\pi y}{2}\right)}\right\}
-\]
-Notice that we used $y$ here instead of $x$.
-We did this because the boundary conditions are defined in terms of $y$, not $x$.
-Now we can continue to solve our problem.
-Recall that our separation of variables gives us two second order ODE's,
-\[
-X'' - \lambda X = 0 \qquad\qquad \text{and} \qquad\qquad \lambda Y - Y'' = 0
-\]
-We already solved for $Y(y)$ by finding the corresponding eigenvalues/vectors. Now we plug in our $\lambda_{n}$ into our $X$ ODE. Solving the resulting ODE will give us a basis of
-\[
-\left\{ e^{\frac{n\pi}{2}x}, e^{-\frac{n\pi}{2}x} \right\}
-\]
-Now we could continue at this point, but if we think ahead, we can see that when we eventually apply our initial conditions, nothing will cancel out.
-At this point, we want to transform our basis.
-We have as follows
-\[
-\left\{ e^{\frac{n\pi}{2}x}, e^{-\frac{n\pi}{2}x} \right\} \implies \left\{ \cosh{\left(\frac{n\pi x}{2}\right)}, \sinh{\left(\frac{n\pi x}{2}\right)} \right\}
-\]
-All we did here is apply an identity. Now we have everything we need to set up a solution. Our $u(x,y)$ will be
-\[
-u(x,y) = \sum_{n=1}^{\infty}\left[a_{n}\cosh{\left(\frac{n\pi x}{2}\right)} + b_{n}\sinh{\left(\frac{n\pi x}{2}\right)}\right]\sin{\left(\frac{n\pi y}{2}\right)}
-\]
-We can apply our initial conditions, which gives us $b_{n}\ \text{and}\ a_{n}$:
-\begin{align*}
-a_{n} &= \frac{4(-1)^{n+1}}{n\pi}\\
-b_{n} &= -\coth{\left(\frac{n \pi}{2}\right)}\frac{4(-1)^{n+1}}{n\pi}
-\end{align*}
-This yields a final solution of
-\[
-u(x,y) = \sum_{n=1}^{\infty}\left[ \frac{4(-1)^{n+1}}{n\pi} \cosh{\left(\frac{n\pi x}{2}\right)} -\coth{\left(\frac{n \pi}{2}\right)}\frac{4(-1)^{n+1}}{n\pi}\sinh{\left(\frac{n\pi x}{2}\right)}\right]\sin{\left(\frac{n\pi y}{2}\right)}
-\]
-And we are done
-\newpage
-
-\indent Now we turn our attention to the Laplace Equation with a circular domain.
-As we discussed earlier, we can't simply say that our solution will be in the form of $X(x)Y(y)$.
-However, we can force the domain to be rectangular if we use polar coordinates.
-Let's consider a variation of $B$ from above.\\\\
-
-\noindent \textbf{\textit{Ex:}} Solve the following PDE
-\[
-\begin{cases*}
-u_{xx} + u_{yy} = 0\quad \text{on}\ x^{2} + y^{2} < a^{2}\\
-u(a,\theta) = U_{1}\quad \text{on}\ 0 \leq \theta \leq \pi\\
-u(a,\theta) = U_{2}\quad \text{on}\ \pi \leq \theta \leq 2\pi
-\end{cases*}
-\]
-\indent \textbf{\textit{Solution:}} We want to express the partial derivatives here in terms of $r\ \text{and}\ \theta$. Recall that we have
-\[
-x = r\cos{\theta} \qquad\qquad \text{and} \qquad\qquad y = r\sin{\theta}
-\]
-First, we will find the partial derivatives of $x$ and $y$ with respect to $r$ and $\theta$. Since $x = r\cos{\theta}$, 
-\[
-\frac{\partial x}{\partial r} = \cos{\theta} \qquad\qquad \frac{\partial x}{\partial \theta} = -r\sin{\theta}
-\]
-We can compute the derivatives for $y$ in a similar fashion:
-\[
-\frac{\partial y}{\partial r} = \sin{\theta} \qquad\qquad \frac{\partial y}{\partial \theta} = r\cos{\theta}
-\]
-Now we need to find $u_{rr}$ and $u_{\theta\theta}$. Let's focus on $u_{rr}$ first. We will need to start by finding $u_{r}$. We have as follows
-\begin{align*}
-\frac{\partial u}{\partial r}& = \frac{\partial u}{\partial x}\frac{\partial x}{\partial r} + \frac{\partial u}{\partial y}\frac{\partial y}{\partial r}\\
-&= \cos{\theta}\frac{\partial u}{\partial x} + \sin{\theta}\frac{\partial u}{\partial y}
-\end{align*}
-We have $u_{r}$, so now we need to compute $u_{rr}$. We have as follows:
-\begin{align*}
-\frac{\partial^{2} u}{\partial r^{2}} &= \cos{\theta}\frac{\partial }{\partial r}\frac{\partial u}{\partial x} + \sin{\theta}\frac{\partial }{\partial r}\frac{\partial u}{\partial y}\\
-&= \cos^{2}{\theta}\frac{\partial^{2} u}{\partial x^{2}} + 2\cos{\theta}\sin{\theta}\frac{\partial^{2}u}{\partial x\partial y} + \sin^{2}{\theta}\frac{\partial^{2} u}{\partial y^{2}}
-\end{align*}
-We won't delve into finding $u_{\theta\theta}$, as the partials get a little out of hand. However, we end up with
-\begin{align*}
-\frac{\partial^{2}u}{\partial r^{2}} + \frac{1}{r^{2}}\frac{\partial^{2}u}{\partial\theta^{2}} &= -\frac{1}{r}\frac{\partial u}{\partial r} + \frac{\partial^{2} u}{\partial x^{2}} + \frac{\partial^{2} u}{\partial y^{2}}\\
-u_{xx} + u_{yy} &= u_{rr} + \frac{1}{r}u_{r} + \frac{1}{r^{2}}u_{\theta\theta} = 0
-\end{align*}
-Let's go ahead and formalize it:
-\[
-\nabla^{2}u = u_{rr} + \frac{1}{r}u_{r} + \frac{1}{r^{2}}u_{\theta\theta} = 0
-\]
-Now we can treat this as a 2D problem; in other words, we can now use separation of variables to solve for $u$. As before, we know our solution will look like
-\[
-u(r,\theta) = R(r)T(\theta)
-\]
-Plugging everything in and realizing by the usual trick that everything s equal to a constant gives us
-\[
--\frac{T''}{T} = \frac{R'' + \frac{R'}{r}}{\frac{R}{r^{2}}} = \lambda
-\]
-Since we are dealing in polar coordinates, we will go ahead and walk through the eigenvalues/functions.
-We will be solving for $T(\theta)$, since our boundary conditions are in terms of $\theta$.
-We also should note that the solutions will repeat themselves every $2\pi$, since we are in the polar system.
-In addition, our $u$ should be finite at $r=0$.
-Let's take a look at the eigenvalues for the Laplace equation.\\\\
-\begin{minipage}[t]{0.32\textwidth}
-\underline{$\lambda = 0$}\\
-When $\lambda = 0$, we have \[T(\theta) = a\theta + b.\] This also tells us that \[R(r) = c\ln{(r)} + d.\] Putting everything together gives us \[ u(r,\theta) = (a\theta + b)(c\ln{r} + d).\] Since the solution is periodic on $2\pi$, we know that $a=0$.
-Since the solution has to remain finite at $r=0$, we know that $c = 0$.
-This gives us that $u = bd$, or that $u$ is equal to a constant.
-\end{minipage}
-\hspace{0.01\textwidth}
-\begin{minipage}[t]{0.32\textwidth}
-\underline{$\lambda > 0$}\\
-With $\lambda > 0$, we have our characteristic equation $(m = \pm\ \sqrt{\lambda})$ yielding a real valued solution. This tells us that
-\[
-T(\theta) = c_{1}\cosh{(\sqrt{\lambda}\theta)} + c_{2}\sinh{(\sqrt{\lambda}\theta)}.
-\]
-The solution is periodic on $2\pi$, so we can say that $c_{1} = c_{2} = 0$. Therefore, we have a trivial solution.
-\end{minipage}
-\hspace{0.01\textwidth}
-\begin{minipage}[t]{0.32\textwidth}
-\underline{$\lambda < 0$}\\
-$\lambda < 0$ gives us 
-\[
-T(\theta) = c_{1}\cos{(\sqrt{\lambda}\theta)} + c_{2}\sin{(\sqrt{\lambda}\theta)}
-\]
-This solution is periodic with $2\pi$, and it yields a trivial eigenvector.
-\end{minipage}
-\noindent\\\\\\
-\noindent So we know that the only $\lambda$ that yields eigenvalues is $\lambda = 0$, and we know that this particular eigenvalue yields a constant solution.
-If we solve the differential equation for $R$, we will find that the only solution we care about is $r^{n}$.
-We can finally say that our solution will look like
-\[
-u(r,\theta) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}r^{n}[a_{n}\cos{(n\theta)} + b_{n}\sin{(n\theta)}]
-\]
-This solution is worth memorizing, as it will generally be the same for all variations of the Laplace equation similar to this one.
-Now we can apply the conditions for $\theta$.
-We have two conditions, but it is really just a piecewise function which we will call $f(\theta)$.
-We also know that our bound on the domain is $a$, or the length of the radius.
-This is enough information to find all the various variables in our $u(r,\theta)$.
-We have as follows:
-\[
-u(a,\theta) = f(\theta) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}a^{n}[a_{n}\cos{(n\theta)} + b_{n}\sin{(n\theta)}]
-\]
-Note that the $a$ which is the length of the radius is \textit{not} the same $a$ as $a_{n}$. Now we can solve for $a_{0}$ and $a_{n}$.
-\begin{alignat*}{3}
-a_{0} &= \frac{1}{\pi}\int_{0}^{2\pi}f(\theta)d\theta \qquad\qquad a_{n} &&= \frac{1}{\pi}\left[\int_{0}^{\pi}U_{1}\cos{(n\theta)}d\theta + \int_{\pi}^{2\pi}U_{2}\cos{(n\theta)}d\theta\right]\\
-&= U_{1} + U_{2} &&=0
-\end{alignat*}
-For $b_{n}$, we have
-\begin{align*}
-a^{n}b_{n} &= \frac{1}{\pi}\left[\int_{0}^{\pi}U_{1}\sin{(n\theta)}d\theta + \int_{\pi}^{2\pi}U_{2}\sin{(n\theta)}d\theta\right]\\
-&= \frac{1 - (-1)^{n}}{n\pi}(U_{1} - U_{2})\\
-b_{n} &= \frac{\left[1 - (-1)^{n}\right](U_{1} - U_{2})}{n\pi a^{n}}\\
-&= \frac{2(U_{1} - U_{2})}{(2k+1)\pi a^{2k+1}}
-\end{align*}
-Putting it all together gives us a final solution of
-\[
-u(r,\theta) = \frac{U_{1} + U_{2}}{2} + \sum_{k=0}^{\infty}r^{k}\frac{2(U_{1} - U_{2})}{(2k+1)\pi a^{2k+1}}
-\]
-And we are done.
-\newpage
-
-\indent Now let's look at another Laplace equation in polar coordinates.
-Again, we will be dealing with Dirichlet boundary conditions, so we will be able to jump straight into the solution.
-This one won't take us as long to solve, as we have already done a lot of the work.
-We are asked the following:\\\\
-\noindent \textbf{\textit{Ex:}} Solve the following PDE
-\[
-\begin{cases*}
-\nabla^{2}u = 0\\
-u(3,\theta) =
-\begin{cases*}
-1 \qquad\qquad 0 \leq \theta \leq \pi\\
-\sin^{2}{\theta}\qquad \pi \leq \theta \leq 2\pi 
-\end{cases*}
-\end{cases*}
-\]
-\indent \textbf{\textit{Solution:}} We already know what our separable equation will look like:
-\[
-\frac{r^{2}R'' + rR'}{R} = -\frac{T''}{T} = -\lambda
-\]
-This yields an eigenvalue of $\lambda = n^{2}$. We know this because our function $T$ must be periodic on $2\pi$, and as such we get the following piecewise function:
-\[
-T(\theta) =
-\begin{cases*}
-a_{0}\qquad\qquad\qquad\qquad\qquad\quad\ \text{for}\ n=0\\
-a_{n}\cos{(n\theta)} + b_{n}\sin{(n\theta)}\qquad \text{for}\ n > 0
-\end{cases*}
-\]
-The solution to our equation for $R$ yields
-\[
-R(r) = 
-\begin{cases*}
-c_{1} + c_{2}\ln{r} \qquad\qquad \text{for}\ n=0\\
-c_{1}r^{n} + c_{2}r^{-n}\qquad\quad \text{for}\ n > 0
-\end{cases*}
-\]
-We can disregard $c_{2}$ for this equation, since it would cause our solution to go to $\infty$ as $r \to 0$, at the center of the disk. Putting everything together gives us
-\begin{align*}
-u(r,\theta) &= R(r)T(\theta)\\
-&=
-\begin{cases*}
-a_{0}\qquad\qquad\qquad\qquad\qquad\qquad\quad\ \text{for}\ n=0\\
-a_{n}r^{n}\cos{(n\theta)} + b_{n}r^{n}\sin{(n\theta)}\qquad \text{for}\ n > 0
-\end{cases*}
-\end{align*}
-We should probably construct a series out of this, since that is what we usually do. The corresponding series is
-\[
-u(r,\theta) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}a_{n}r^{n}\cos{(n\theta)} + b_{n}r^{n}\sin{(n\theta)}
-\]
-And applying the boundary conditions gives us
-\[
-u(3,\theta) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}a_{n}3^{n}\cos{(n\theta)} + b_{n}3^{n}\sin{(n\theta)}
-\]
-For $a_{0}$, we have
-\begin{align*}
-a_{0} &= \frac{1}{\pi}\left[\int_{0}^{\pi}1 d\theta + \int_{\pi}^{2\pi}sin^{2}{\theta}d\theta\right]\\
-&= \frac{3}{2}
-\end{align*}
-Solving for $a_{n}$ gives us
-\begin{align*}
-a_{n} &= \frac{1}{3^{n}\pi}\left[\int_{0}^{\pi}\cos{(n\theta)}d\theta + \int_{\pi}^{2\pi}\cos{(n\theta)}\sin^{2}{\theta}d\theta\right]\\
-&= \frac{[n^{2} - 2 - 4\cos{(n\pi)}]\sin{(n\pi)}}{3^{n}n(n^{2}-4)\pi}\\
-&= 0
-\end{align*}
-Our $b_{n}$ term is
-\begin{align*}
-b_{n} &= \frac{1}{3^{n}\pi}\left[\int_{0}^{\pi}\sin{(n\theta)}d\theta + \int_{\pi}^{2\pi}\sin{(n\theta)}\sin^{2}{\theta}d\theta\right]\\
-&= \frac{2\left[n^{2} - 6 - 4\cos{(n\pi)}\right]\sin^{2}{\left(\frac{n\pi}{2}\right)}}{3^{n}n\pi(n^{2} - 4)}\\
-&= \frac{2\left[n^{2} - 6 - 4(-1)^{n}\right]\sin^{2}{\left(\frac{n\pi}{2}\right)}}{3^{n}n\pi(n^{2} - 4)}\\
-&=
-\begin{cases*}
-0 \qquad\qquad\qquad\qquad\quad\ \ n \in \{ 2k\ |\ k \in \mathbb{Z} \}\\
-\frac{1}{3^{n}\pi}\left[\frac{2}{n} + \frac{4}{n(n^{2} - 4)}\right]\qquad n \in \{ 2k + 1\ |\ k \in \mathbb{Z} \}
-\end{cases*}
-\end{align*}
-We essentially have our solution. All we have to do is combine everything:
-\[
-u(r,\theta) = \frac{3}{4} + \sum_{k=0}^{\infty}\frac{r^{2k + 1}}{3^{2k+1}\pi}\frac{2(4k^{2} + 4k -1)}{(2k + 3)(4k^{2} -1)}\sin{\left[(2k +1)\theta\right]}
-\]
-Let's take a look at some of the plots for this equation.\\
-\begin{center}
-\begin{tabular}{c c c}
-\includegraphics[scale=0.3]{lap_01_3} & \includegraphics[scale=0.3]{lap_01_30} & \includegraphics[scale=0.3]{lap_01_50}\\
-\textit{The first 3 terms} & \textit{The first 30 terms} & \textit{The first 50 terms}
-\end{tabular}
-\end{center}
-Plugging this infinite sum into mathematica tells us that it does not converge; the values are heading towards $\infty$ as the sum grows. We will leave our answer as a series, and we are done.
-
-\newpage
-\subsection{Wrapping Up}
-\indent Before we move onto the next topic, we will present some common patterns amongst the second order PDEs.
-Below, you will find each of the three equations in the classical trinity, and their corresponding solutions based on the boundary conditions.
-\subsubsection{The Heat Equation}
-\[
-\begin{array}{c c}
-\begin{cases}
-u_{t} - ku_{xx} = 0\\
-u(0,t) = u(L,t) = 0\\
-u(x,0) = f(x)
-\end{cases}
-&
-\begin{cases}
-u_{t} - ku_{xx} = 0\\
-u_{x}(0,t) = u_{x}(L,t) = 0\\
-u(x,0) = g(x)
-\end{cases}\\
-&\\
-u(x,t) = \sum_{n=1}^{\infty}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}\sin{\left(\frac{n\pi x}{L}\right)}
-& u(x,t) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}\cos{\left(\frac{n\pi x}{L}\right)}\\
-&\\
-\begin{cases}
-u_{t} - ku_{xx} = 0\\
-u(0,t) = u_{x}(L,t) = 0\\
-u(x,0) = h(x)
-\end{cases}
-&
-\begin{cases}
-u_{t} - ku_{xx} = 0\\
-u_{x}(0,t) = u(L,t) = 0\\
-u(x,0) = k(x)
-\end{cases}\\
-&\\
-u(x,t) =  \sum_{n=1}^{\infty}e^{-k\left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}\right)^{2}t}\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}
-& u(x,t) =  \sum_{n=1}^{\infty}e^{-k\left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}\right)^{2}t}\cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}
-\end{array}
-\]
-\noindent\\
-\subsubsection{The Wave Equation}
-\[
-\begin{array}{c c}
-\begin{cases}
-u_{tt} - c^{2}u_{xx} = 0\\
-u(0,t) = u(L,t) = 0\\
-u(x,0) = f(x) \quad u_{t}(x,0) = g(x)
-\end{cases}
-&
-\begin{cases}
-u_{tt} - c^{2}u_{xx} = 0\\
-u_{x}(0,t) = u_{x}(L,t) = 0\\
-u(x,0) = f(x) \quad u_{t}(x,0) = g(x)
-\end{cases}\\
-&\\
-u(x,t) = \sum_{n=1}^{\infty}\left[a_{n}\cos{\left(\frac{n \pi t c}{L}\right)} + b_{n}\sin{\left(\frac{n \pi t c}{L}\right)}\right]\sin{\left(\frac{n \pi x}{L}\right)}
-&u(x,t) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}\left[a_{n}\cos{\left(\frac{n \pi t c}{L}\right)} + b_{n}\sin{\left(\frac{n \pi t c}{L}\right)}\right]\cos{\left(\frac{n \pi x}{L}\right)}\\
-&\\
-\end{array}
-\]
-\[
-\begin{array}{c}
-\begin{cases}
-u_{tt} - c^{2}u_{xx} = 0\\
-u(0,t) = u_{x}(L,t) = 0\\
-u(x,0) = f(x) \quad u_{t}(x,0) = g(x)
-\end{cases}\\
-\\
-u(x,t) = a_{0} + b_{0}t + \sum_{n=1}^{\infty}\left[ a_{n}\cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi t}{L}\right)} + b_{n}\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}\right]\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}\\
-\\
-\begin{cases}
-u_{tt} - c^{2}u_{xx} = 0\\
-u_{x}(0,t) = u(L,t) = 0\\
-u(x,0) = f(x) \quad u_{t}(x,0) = g(x)
-\end{cases}\\
-\\
-u (x,t) = a_{0} + b_{0}t + \sum_{n=1}^{\infty}\left[ a_{n}\cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi t}{L}\right)} + b_{n}\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}\right]\cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}
-\end{array}
-\]
-\subsubsection{The Laplace Equation}
-\paragraph{Cartesian Coordinates}
-\[
-\begin{array}{c c}
-\begin{cases}
-u_{xx} + u_{yy} = 0\\
-u(0,y) = u(L,y) = 0
-\end{cases}
-&
-\begin{cases}
-u_{xx} + u_{yy} = 0\\
-u_{x}(0,y) = u_{x}(L,y) = 0
-\end{cases}\\
-&\\
-u(x,y) = \sum_{n=1}^{\infty}\left[a_{n}\sinh{\left(\frac{n \pi y}{L}\right)} + b_{n}\cosh{\left(\frac{n\pi y}{L}\right)}\right]\sin{\left(\frac{n\pi x}{L}\right)}
-& u(x,y) = \sum_{n=1}^{\infty}\left[a_{n}\sinh{\left(\frac{n \pi y}{L}\right)} + b_{n}\cosh{\left(\frac{n\pi y}{L}\right)}\right]\cos{\left(\frac{n\pi x}{L}\right)}
-\end{array}
-\]
-\\
-\[
-\begin{array}{c}
-\begin{cases}
-u_{xx} + u_{yy} = 0\\
-u_{x}(0,y) = u(L,y) = 0
-\end{cases}\\
-\\
-u(x,t) = \sum_{n=1}^{\infty}\left[a_{n}\sinh{\left(\left[n + \frac{1}{2}\right]\frac{\pi y}{L}\right)} + b_{n}\cosh{\left(\left[n + \frac{1}{2}\right]\frac{\pi y}{L}\right)}\right]\cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi y}{L}\right)}\\
-\\
-\begin{cases}
-u_{xx} + u_{yy} = 0\\
-u(0,y) = u_{x}(L,y) = 0
-\end{cases}\\
-\\
-u(x,t) = \sum_{n=1}^{\infty}\left[a_{n}\sinh{\left(\left[n + \frac{1}{2}\right]\frac{\pi y}{L}\right)} + b_{n}\cosh{\left(\left[n + \frac{1}{2}\right]\frac{\pi y}{L}\right)}\right]\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi y}{L}\right)}
-\end{array}
-\]
-\paragraph{Polar Coordinates}
-\begin{gather*}
-\begin{cases}
-u_{rr} + \frac{1}{r}u_{r} + \frac{1}{r^{2}}u_{\theta\theta} = 0\\
-u(h,\theta) = f(\theta)\quad \text{on}\ x < \theta < L
-\end{cases}\\
-u(r,\theta) = a_{0} + \sum_{n=1}^{\infty}a_{n}r^{n}\cos{\left(\frac{n \pi \theta}{L}\right)} + b_{n}r^{n}\sin{\left(\frac{n \pi \theta}{L}\right)}
-\end{gather*}
 \end{document}

### 2015-11-16 cheatsheets/ode_formal_slns.tex
Added homogeneous solutions

@@
 \usepackage{array}
+\usepackage{hyperref}
 \topmargin=-0.3in
@@
 
-We have the following system
+We have the following differential equation
 \[
-	\begin{cases}
-	y'+ay=0\\
-	y(0)=\alpha
-	\end{cases}
+	y'+ay=0.
+\]
+Then
+\[
+	y=c e^{-at},
+\]
+where $c$ is determined by the initial condition.
+
+\subsection{Second Order}
+
+We have the following differential equation
+\begin{equation}
+	y''+ay'+by=0. \label{eqn1}
+\end{equation}
+Then we solve the following equation for $r$,
+\[
+	r^2+ar+b=0.
 \]
 
+If $r=\alpha\pm i\beta$, then the general solution is
+\[
+	y=c_1 e^{\alpha x}\cos(\beta x)+c_2 e^{\alpha x}\sin(\beta x)
+\]
+where $c_1$ and $c_2$ are determined by the initial conditions.
+
+If \autoref{eqn1} has two real solutions, $r_1$ and $r_2$, then
+\[
+	y=c_1e^{r_1 x}+c_2 e^{r_2x}
+\]
+where $c_1$ and $c_2$ are determined by the initial conditions.
+
+If \autoref{eqn1} has a real repeated root, $r$, then
+\[
+	y=c_1e^{rx}+c_2xe^{rx}
+\]
+where $c_1$ and $c_2$ are determined by the initial conditions.
+
 \end{document}

### 2015-11-16 cheatsheets/ode_formal_slns.tex
Added nonhomogeneous solutions

@@
 
+\section{Nonhomogeneous}
+
+Suppose we have the following differential equation.
+\begin{equation}\label{eqn2}
+	ay'' + by' +cy = f(x)
+\end{equation}
+
+We begin by finding the solution to the homogeneous equation,
+\[
+	ay'' + by' +cy = 0.
+\]
+Call this solution the general solution or $y_g$.
+Note, do \textbf{NOT} apply any initial conditions for this solution.
+
+We now have a table for $f(x)$ and the form of the particular solution.
+\[
+\begin{array}{|c|c|}
+\hline
+f & y_p\\\hline
+x & Ax+B\\\hline
+x^2 & Ax+B+C\\\hline
+e^x & Ae^x\\\hline
+\cos(n\pi x) \text{ or } \sin(n\pi x) & A\cos(n\pi x) + B\sin(n\pi x)\\\hline
+x\cos(n\pi x) \text{ or } \sin(n\pi x) & (Ax+B)\cos(n\pi x) + (Cx +D)\sin(n\pi x)\\\hline
+\end{array}
+\]
+To find the $A,B,C,D$, plug $y_p$ back into \autoref{eqn2}, and match terms with $f(x)$.
+The solution to \autoref{eqn2} is $y=y_g+y_p$.
+Now apply your initial conditions to determine your as of now undetermined coefficients.
+
 \end{document}

### 2015-11-16 cheatsheets/ode_formal_slns.tex
Fixed error

@@
 f & y_p\\\hline
+k & A\\\hline
 x & Ax+B\\\hline
-x^2 & Ax+B+C\\\hline
+x^2 & Ax^2+Bx+C\\\hline
 e^x & Ae^x\\\hline
@@
 \]
+If your $f$ is a linear combination of the above, then youur $y_p$ is a linear combination of the applicable entries above.
 To find the $A,B,C,D$, plug $y_p$ back into \autoref{eqn2}, and match terms with $f(x)$.

### 2015-11-17 wim/pkmn_problem/hash.rb
working on presentation

@@ -16,7 +16,6 @@ def main
 		while pokee == nil do
 			pokee = pokee = pkmn_list.sample
 		end
-		puts "pokee is #{pokee}"
 		array.push pokee
 	end
 	array.each do |pokemon|

### 2015-11-18 contract/contact.tex
it would really help if i could spell

\documentclass{article}
\usepackage{amssymb,amsmath}
\usepackage{multicol}
\topmargin=-0.3in
\textheight=9.2in
\textwidth=168mm
\oddsidemargin=-0.2in
\evensidemargin=-0.2in
\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{\textbackslash newline}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}
\title{Independent Contractor Agreement}
\date{}
\author{}
\begin{document}
%\maketitle
\begin{center}
{\huge Independent Contractor Agreement}
\end{center}
\vspace{1cm}
\noindent\\
\indent The following contract is a specification of the requirements for the first release (v1.0.0) of the WhensIt application. By signing this contract, both parties agree to uphold the responsibilities mentioned in the following sections.
\section{Specification of Features Qualifying for First Release}
\indent The following features must be implemented by the software developer in full in order to qualify the application for release:
\begin{multicols}{2}
\begin{enumerate}
\item The application will compile and run on any sever supporting the Meteor.js 1.2.0.2 framework.
\item The application will run in all Android 4.4.x+ , iOS 8+ , Macintosh OS X 10.7+, and Windows 7+ systems.
\item The application will have a working login screen that allows users to 1) create an account, or 2) log in with an existing account.
\item The application will present the user with a main page that includes the following features:
\begin{itemize}
\item The user's chosen profile picture, or the default profile picture,
\item The user's chosen name,
\item The user's current score,
\item A localized map, powered by Google Maps, that will display nearby events and exhibits.
\end{itemize}
\item The application will provide an interface to the quiz portion of the application. Quizzes will be structured as follows:
\begin{itemize}
\item Quizzes will be composed of Questions that will be created by the owner(s) of the application.
\item Questions will contain four answers, one of which \textit{must} be the correct answer.
\item Upon choosing an answer, the user will be notified of the correctness of their answer through either a sound, a color change, or an in-app alert.
\item Quizzes will continue until the user has answered all Questions in the Quiz.
\item Upon finishing a Quiz, a user will be notified of their total score, the Questions that were answered incorrectly, and the Questions that were answered correctly.
\item Users will be able to retake Quizzes as many times as they wish.
\end{itemize}
\item Users will have the ability to delete their account. Upon deletion, \textit{all} of the user's corresponding data will be deleted from the database.
\item The application will make the distinction between Events and Exhibits, which is as follows:
\begin{itemize}
\item Events will be time sensitive; Events will only be accessible during specific times. These times will be determined by the owner(s) of the application.
\item Exhibits will be at all times until the owner(s) of the application determine that the Exhibit should be removed.
\item Events and Exhibits will be accessible through both a \textit{unique} three character code and a QR code that will be available for the user to scan.
\end{itemize}
\item The application will provide an administrator interface, to which \textit{only} the owner(s) of the application will have access, unless the owner(s) decide to provide administrator privileges to another user. The administrator interface will allow the user to access and edit all items in the database with the exception of the database representation of a User.
\item All API keys \textit{must} remain under control of \textit{only} the owner(s) of the application. API Keys that need to remain secret are as follows:
\begin{itemize}
\item Parse Application Key
\item Parse Client Key
\item Parse Javascript Key
\item Parse REST API Key
\item Parse Master Key
\item Google Developers Application ID
\item Google Maps Javascript API Key
\end{itemize}
\item The application will follow the stylistic guidelines as agreed upon by the contractor and the developer.
\end{enumerate}
\end{multicols}
\vspace{1cm}
\section{Agreement of Ownership}
\noindent\\
\indent Upon completion of the application, ownership of the finished application will be transferred to the contractor who may then defer temporary ownership to the developer when necessary.
Ownership includes the rights to the database, all the code as permitted by the licensing agreement and this contract, and the ability to make and request changes to the code and the end product.
The code must remain open source and licensed under the No License license, which states that the code may not be modified and then distributed by anyone other than the owner.
The No License license permits private use of the code for non-proprietary purposes.\\
\indent At the time when the developer leaves the project, the developer agrees to transfer ownership of the repository, ownership of the database, all application keys, and all releases of the application to whomever the owner appoints.
\vspace{1cm}
\section{Agreement of Payment}
\noindent\\
\indent Upon completion of the features mentioned in \textbf{\textit{Specification of Features Qualifying for First Release}}, the contractor agrees to pay a one time flat fee of \underline{\hspace{1em}\$1,200.00\hspace{1em}} for the finished product version, v1.0.0.
For any work done after the release of v1.0.0, the contractor agrees to pay an hourly rate of \underline{\hspace{1em}\$10.00 per hour\hspace{1em}} for every hour of work put towards the development of the new release.
The developer agrees to only charge for work done \textit{specifically} for the new release of the application and will provide proof of work through commits to the repository where the application code is located.

\vspace{4cm}
\underline{\hspace{15em}} \hspace{4em} \underline{\hspace{15em}}\\

\hspace{1em}Developer\hspace{15em}Contractor\\

\underline{\hspace{15em}}\\

\hspace{1em}Date
\end{document}

### 2015-11-21 prettyproofs/proofs.tex
presentation

\documentclass{article}
\usepackage{amssymb,amsmath}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{array}
\topmargin=-0.3in
\textheight=9.2in
\textwidth=168mm
\oddsidemargin=-0.2in
\evensidemargin=-0.2in

\newcommand{\NN}{\mathbb{N}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\CC}{\mathbb{C}}

\newcommand{\Proof}{\textit{\textbf{Proof: }}}
\newcommand{\scc}{\begin{cases*}}
\newcommand{\ecc}{\end{cases*}}

\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{axiom}{Axiom}
\newtheorem{proof}{Proof}

\title{Proofs}
\date{}
\begin{document}
\maketitle
\hrule
\noindent\\\\
\begin{axiom}
The number $0.99999... = 1.00000...$.
\end{axiom}
\begin{theorem}
There is no bijection between the natural numbers $\NN$ and the set $(0,1)$.
\end{theorem}
\noindent
\indent \Proof Suppose $|\NN| = |(0,1)|$. 
Let $f : \NN \to (0,1)$ be a bijection. 
We will let $a_{i} = f(i)$ and $a_{ij}$ be the $j^{\text{th}}$ decimal value of $a_{i}$. 
We can construct the bijection by
\begin{alignat*}{3}
a_{1} &= .a_{11}a_{12}a_{13}a_{14}\dots\ &&= f(a_{1})\\
a_{2} &= .a_{21}a_{22}a_{23}a_{24}\dots &&= f(a_{2})\\
a_{3} &= .a_{31}a_{32}a_{33}a_{34}\dots &&= f(a_{3})\\
a_{4} &= .a_{41}a_{42}a_{43}a_{44}\dots &&= f(a_{4})\\
\dots
\end{alignat*}
Now we can take an element $b \in (0,1)$ such that $b_{i}$ is the $i^{\text{th}}$ digit of $b$. 
We want to construct the digits of $b$ such that each space along the $a_{ii}$ diagonal is not equal to a digit in $b$. We can express this by
\[
b_{i} = 
\scc
n,\qquad a_{i} \neq n\\
n+1,\ a_{i} = n\\
8,\qquad\ a_{i} = 9
\ecc
\]
\end{document}

### 2015-11-21 wim/presentation.tex
working on presentation
presentation

@@
 \mode<presentation>
-\usetheme{boxes}
+\usetheme{Luebeck}
+
 \usepackage{amsmath}
-\AtBeginSection[]{
-\begin{frame}<beamer>
-\frametitle{Summary}
-\tableofcontents[currentsection]
-\end{frame}
+\usepackage{listings}
+\usepackage{verbatim}
+
+\lstset{%
+showstringspaces=false,
+numbers=left,
+numbersep=10pt,
+language=Ruby,
 }
@@
 \author{Steven Rosendahl}
-
+\date{}
 \begin{document}
@@
 \begin{frame}{The Question}
-\pause
 \begin{itemize}
@@
 \begin{enumerate}
-\pause
 \item \underline{Pok\'emon Problem:} How many unique Pok\'emon will a player encounter in a given playthrough of all the games?
-\pause
-\item \underline{Facebook Problem:} How many unique application installs are made a day on Facebook?
 \pause
@@
 \begin{itemize}
-\pause
 \item What is \textit{hashing}?
@@
 \begin{itemize}
-\pause
 \item How can we solve the Pok\'emon Problem using a hash?
@@
 \pause
+\item Results:
+\begin{itemize}
+\pause
+\item On average: 461 unique encounters
+\end{itemize}
 \end{enumerate}
@@
 
+\begin{frame}[fragile]{Implementation}
+\begin{lstlisting}[basicstyle=\small]
+array.each do |pokemon|
+     hash_table[pokemon.hash] = pokemon.hash
+end
+total = 0
+hash_table.each do |val|
+     if val != -1
+          total = total + 1
+     end
+end
+puts "The total number of unique encounters is #{total}"
+\end{lstlisting}
+\end{frame}
+
 \begin{frame}{Problems With The Hash Table}
 \begin{itemize}
-\pause
 \item Memory Intensive
@@
 \begin{enumerate}
-\pause
 \item Create a bitmap in memory. We will call this $\mathbb{V}$.
@@
 \pause
-%\item
 \begin{itemize}
@@
 \begin{frame}{The Math}
-\pause
 \begin{itemize}
@@
 \item $M$ is the size of the original set of data ($\mathbb{S}$).
+\pause
+\item For the Twitter Problem:
+\pause
+\begin{align*}
+\text{Memory Required} &\approx \log_{2}{(\log_{2}{(200,000,000 \times 10)})}\\
+&\approx 4.94 kb
+\end{align*}
 \pause
@@
 \begin{frame}{Solving the Twitter Problem}
-\pause
 \begin{itemize}
@@
 
+\begin{frame}[fragile]{Implementation}
+\begin{lstlisting}[language=Ruby]
+mhll = Hyperll::HyperLogLog.new(10)
+File.open("twitter_data.txt","r") do |file|
+     file.each_line do |line|
+          mhll.offer line
+     end
+end
+str = "Unique Elements: #{mhll.cardinality}"
+puts str
+\end{lstlisting}
+\end{frame}
+
 \begin{frame}{Results}
-\pause
 \begin{figure}
-\includegraphics[scale=0.5]{twitter_problem/resultspng}
+\includegraphics[scale=0.7]{twitter_problem/resultspng}
 \caption{$0.284356 x + 7361.39$}

### 2015-12-01 comporg/homework_5.tex
so much pain

\documentclass{article}
\usepackage{amssymb,amsmath}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{multicol}
\topmargin=-0.3in
\textheight=9.2in
\textwidth=168mm
\oddsidemargin=-0.2in
\evensidemargin=-0.2in

\newcommand{\SLN}{\textit{\textbf{Solution: }}}

\begin{document}
\noindent \textit{\textbf{Steven Rosendahl\\CPSC 330}}
\vspace{2cm}
\begin{enumerate}
\item Assume an instruction cache miss rate for a program is 2\% and a data cache miss rate is 4\%. 
If the processor has a CPI of 2 without any memory stalls and the miss penalty is 100 cycles for all misses, 
determine how much faster a processor would run with a perfect cache that never misses. 
The frequency of all loads and stores is 36\% for SPECint2000.
\begin{quote}
\indent \SLN We want to find the ratio of the performance of the perfect processor over the stalling processor.
We have the following:
\begin{align*}
\frac{\text{Perfect Performance}}{\text{Stall Performance}} &= \frac{\text{CPU Stall}}{\text{CPU Perfect}}\\
&= \frac{\text{Instructions} \times \text{CPI}_{\text{Stall}} \times \text{Clock Cycles}}
{\text{Instructions} \times \text{CPI}_{\text{Perfect}} \times \text{Clock Cycles}}\\ 
&= \frac{\text{CPI}_{\text{Stall}}}{\text{CPI}_{\text{Perfect}}}\\
&= \frac{2 + 2.00 + 1.44}{2}\\
&= 2.72
\end{align*}
\end{quote}
\item Find the AMAT for a processor with a 2ns clock, 
a miss penalty of 20 clock cycles, 
a miss rate of 0.05 misses per instruction, 
and a cache access time (including hit detection) of 1 clock cycle.
Assume that the read and write miss penalties are the same and ignore other write stalls.
\begin{quote}
\indent \SLN We have that the equation for \textit{AMAT} is given by
\[
\text{AMAT} = \text{Hit Time} + \text{Miss Rate} \times \text{Miss Penalty}.
\]
Plugging into the equation gives us
\begin{align*}
\text{AMAT} &= 2ns + 0.05 \times (20 \times 2ns)\\
&= 4ns.
\end{align*}
\end{quote}
\item Suppose we can improve the miss rate to 0.03 missed per reference by doubling the cache size. 
This causes the cache access time to increase to 1.2 clock cycles.
Using AMAT as a metric, determine if this is a good trade off.
\begin{quote}
\indent \SLN In this problem, we need to recompute the AMAT for the new processor and compare it to the result in (2).
The AMAT is again given by
\begin{align*}
\text{AMAT} &= (1.2 \times 2ns) + (0.03 \times 20 \times 2ns)\\
&= 3.6ns.
\end{align*}
This result is less than 4ns, so it is a good trade off.
\end{quote}
\item (a) Compute the size of the cache shown below.
(b) Compute the total number of bits required to implement this cache for the Intrisity FastMATH embedded fast microprocessor 
(this number represents the total amount of memory needed for storing all data, tags, and valid bits).
\begin{center}
\includegraphics[scale=0.7]{hw5_graphic}
\end{center}
\begin{quote}
\indent \SLN\\
\hspace{3cm} (a) In order to determine the cache size, we need to first determin the number of entries in the cache.
From the diagram, we can see that we have $2^{8}$ entries, or 256 bytes.
Next, we know that there are 16 slots of size 32 bits in each entry in the cache.
We have that
\[
\text{Size} = (0.256)\frac{16\times 32}{8} = 16.384\ kb.
\]
(b) The total number of bytes required will be given by
\[
\text{Size Required} = (1 + 18 + 512)256 = 135936\ bytes.
\]
\end{quote}
\item Describe the general characteristics of a program that would exhibit very little temporal 
and spatial locality with regards to data accesses.
\begin{quote}
\indent \SLN With very little spacial locality, a program would not be able to access contiguous blocks of memory quickly.
This would result in a slower access time for the program. 
With very little temporal locality, the program would need to continually access different blocks of memory over a short
period of time, which would result in a slower program.
\end{quote}
\item Describe the general characteristics of a program that would exhibit very high amounts of temporal locality 
but very little special locality with regards to data accesses.
\begin{quote}
\indent \SLN In this case, the program would not be accessing blocks of memory that are close to each other,
which would cause the program to run slowly. However, the program would be keeping track of the frequently
accessed blocks of memory, which allows for it to run faster than in (5). 
\end{quote}
\item Describe the general characteristics of a program that would exhibit very little temporal locality
but very high amounts of special locality with regards to data accesses.
\begin{quote}
\indent \SLN With very little temporal locality, the program would continually access different blocks of memory
for the same data. 
However, the program would be accessing the same blocks of memory, which would make the program a little faster.
\end{quote}
\item 
Consider a memory hierarchy using one of the three organizations for main memory shown below. 
Assume that the cache block size is 16 words, that the width of organization (b) of the figure is four words, 
and that the number of banks in organization (b) of the figure is four. 
If the main memory latency for a new access is 10 memory bus clock cycles and the transfer time is 1 memory bus clock cycle, 
what are the miss penalties (in clock cycles) for each of these organizations?
\begin{center}
\includegraphics[scale=0.5]{hw5_graphic2}
\end{center}
\begin{quote}
\begin{enumerate}
\item $\text{Miss Penalty} = 1 + 16 \cdot 10 + 16 \cdot 1 = 177$
\item $\text{Miss Penalty} = 1 + 10\cdot\frac{16}{4} + 4\cdot 1 = 45$
\item $\text{Miss Penalty} = 1 + 10\cdot\frac{16}{4} + 16\cdot 1 = 57$
\end{enumerate}
\end{quote}
\item Caches are important to providing a high performance memory hierarchy to processors. 
Given list of 32-bit memory address references given as word addresses: 3,180,43,2,191,88,190,14,181,44,186,253. 
For each of these references, identify the binary address, the tag and the index given in a direct-mapped 
cache with two-word blocks and a total size of 8 bocks. Also, list if each reference is a hit or a miss, 
assuming the cache is initially empty.
\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\hline
Word Address & Binary Address & Tag & Index & Hit/Miss\\
\hline
3 & 0000 0011 & 0 & 1 & Miss\\
\hline
180 & 1011 0100 & 11 & 2 & Miss\\
\hline 
43 & 0010 1011 & 2 & 5 & Miss\\
\hline
2 & 0000 0010 & 0 & 1 & Hit\\
\hline
191 & 1011 1111 & 11 & 7 & Miss\\
\hline
88 & 0101 1000 & 5 & 4 & Miss\\ 
\hline
190 & 1011 1110 & 11 & 7 & Hit\\
\hline
14 & 0000 1110 & 0 & 7 & Miss\\
\hline
181 & 1011 0101 & 11 & 2 & Hit\\
\hline
44 & 0010 1100 & 2 & 6 & Miss\\
\hline
186 & 1011 1010 & 11 & 10 & Miss\\
\hline
253 & 1111 1101 & 15 & 6 & Miss\\
\hline
\end{tabular}
\end{center}
\end{enumerate}
\end{document}

### 2015-12-01 proofs/portfolio.tex
so much pain

\documentclass{article}
\usepackage{amssymb,amsmath}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{array}
\topmargin=-0.3in
\textheight=9.2in
\textwidth=168mm
\oddsidemargin=-0.2in
\evensidemargin=-0.2in

\newcommand{\Proof}{\textit{\textbf{Proof: }}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\QQ}{\mathbb{Q}}

\begin{document}
\begin{enumerate}
\item
\begin{enumerate}
%%%QUESTION 1%%%
\item Let $a,b$ and $c$ be natural numbers with $a$ odd.
Prove that if $a|(b -c)$ and $a|(b+c)$, then $a|b$ and $a|c$.
\begin{quote}
\indent \Proof Let $a|(b-c)$ and $a|(b + c)$.
Then $b-c = ax$ and $b + c = ay$ for some $x,y\in \NN$.
We have that $b = ax + c$.
Then 
\begin{gather*}
ax + c + c = ay\\
ax + 2c = ay\\
2c = ay - ax\\
2c = a(y-x).
\end{gather*}
Since $a$ is odd, $y-x$ must be even since $2c$ is even.
Then $2 | (y-x)$, and we have $c = a\frac{y-x}{2}$, or $c = az$ for $z\in\NN$.
Therefore $a | c$.
If we let $c = b-ax$, we have
\begin{gather*}
b + b - ax = ay\\
2b - ax = ay\\
2b = ay + ax\\
2b = a(y + x)
\end{gather*}
Since $2b$ is even and $a$ is odd, $y-x$ must be even, or $2 | (y+x)$.
Therefore $b = a\frac{y+x}{2}$, or $b = aj$ for $j\in\NN$.
Therefore $a | b$.
\end{quote}
$\triangle$
\item Using a truth table, show that $\neg(P\wedge Q)$ and $(\neg P \vee \neg Q)$ are logically equivalent.\\
\begin{quote}
\begin{center}
\begin{tabular}{c | c | c | c}
$P$ & $Q$ & $\neg(P\wedge Q)$ & $(\neg P \vee \neg Q)$\\
\hline
0 & 0 & 1 & 1\\
0 & 1 & 1 & 1\\
1 & 0 & 1 & 1\\
1 & 1 & 0 & 0
\end{tabular}
\end{center}
\end{quote}
\end{enumerate}
\newpage
%%%QUESTION 2%%%
\item
\begin{enumerate}
\item Establish the following identity using induction.
\[
\sum_{i = 1}^{n}i^{3} = \left(\frac{n(n+1)}{2}\right)^{2}
\]
\begin{quote}
\Proof \\
\textit{\textbf{Base Case: }} $n = 1$
\begin{align*}
\sum_{i = 1}^{1}i^{3} &= \left(\frac{1(1+1)}{2}\right)^{2}\\
1 ^{3} &= \left(\frac{2}{2}\right)^{2}\\
1 &= 1
\end{align*}
\textit{\textbf{Assume: }} $$\sum_{i = 1}^{n}i^{3} = \left(\frac{n(n+1)}{2}\right)^{2}$$
\textit{\textbf{Prove: }}  
\begin{align*}
\sum_{i = 1}^{n + 1}i^{3} &= \left(\frac{(n + 1)(n+2)}{2}\right)^{2}\\
\sum_{i = 1}^{n}i^{3} + \sum_{i = n}^{n + 1}i^{3} &= \left(\frac{(n + 1)(n+2)}{2}\right)^{2}\\
\left(\frac{n(n+1)}{2}\right)^{2} + (n + 1)^{3} &= \left(\frac{(n + 1)(n+2)}{2}\right)^{2}\ \text{By Induction Hypothesis}\\
n^{2}(n+1)^{2} + 4(n+1)^{3} &= (n+1)^{2}(n+2)^{2}\\
(n+1)^{2}(n^{2} + 4n + 4) &= (n+1)^{2}(n+2)^{2}\\
(n+2)^{2} &= (n+2)^{2}
\end{align*}
\end{quote}
$\triangle$
\item Prove that if $n^{3}$ is odd, then $n$ is odd.
\begin{quote}
\Proof Assume that if $n$ is even, then $n^{3}$ is even.
Then $n = 2k$ for $k\in \ZZ$, which mean that $n^{3} = (2k)^{3}$.
$(2k)^{3} = 2(2^{2}k^{3})$ where $(2^{2}k^{3}) \in \ZZ$. Therefore $n^{3}$ is even.
\end{quote}
$\triangle$
\end{enumerate}
\newpage
%%%QUESTION 3%%%
\item
\begin{enumerate}
\item Using the definition, prove that $f: \QQ \to \QQ$ given by $f(x) = 3x+2$ is bijective.
\begin{quote}
\Proof Let $x,y \in \QQ$ such that $f(x) = f(y)$. Then
\begin{align*}
3x + 2 &= 3y + 2\\
3x &= 3y\\
x &= y
\end{align*}
Therefore $f$ is injective.\\
Let $y\in \QQ$ such that $y = \frac{x-2}{3}$. Then
\begin{align*}
f(y) &= 3\left(\frac{x-2}{3}\right) + 2\\
&= x - 2 + 2\\
&= x
\end{align*}
Therefore $f$ is surjective.\\
Therefore $f$ is bijective.
\end{quote}
$\triangle$
\item Define a relation on $\NN \times \NN$ by $(a,b) \sim (c,d)$ if $a + d = b + c$. 
Prove that $\sim$ is an equivalence relation.
\begin{quote}
\textbf{\textit{Symmetric:}} Let $(a,b) \sim (c,d)$. Then
\begin{align*}
a + d &= b + c\\
-b - c &= -a - d\\
(-1)(b + c) &= (-1)(a + d)\\
b + c &= a + d
\end{align*}
Therefore $(c,d) \sim (a,b)$, and $\sim$ is symmetric.\\\\
\textbf{\textit{Reflexive:}} Let $(a,b) \in\NN\times\NN$. 
If $(a,b) \sim (a,b)$, then $a + b = a + b$.
Therefore $(a,b) \sim (a,b)$, and $\sim$ is reflexive.\\\\
\textbf{\textit{Transitive:}} Let $(a,b) \sim (c,d)$ and $(c,d) \sim (e,f)$.
Then $a + d = b + c$ and $c + f = d + e$. 
Solving for $c$ yields $c = d + e - f$, and substituting gives us $a + d = b + d + e - f$.
Then $a + f = e + b$, and $(a,b) \sim (e,f)$.
Therefore, $\sim$ is transitive.\\\\
Therefore $\sim$ is an equivalence relation.
\end{quote}
$\triangle$
\end{enumerate}
%%%QUESTION 4%%%
\newpage
\item
\begin{enumerate}
\item Let $\mathcal{X}$ be a finite set with cardinality $n$.
Prove that the power set, $\mathcal{P}(\mathcal{X})$, has cardinality $2^{n}$.
\begin{quote}
\Proof\\
\textit{\textbf{Base Case:}} A set of size 1.\\
Let $\mathcal{X}$ be the set $\{x\}$.
Then $\mathcal{P}(\mathcal{X})$ is $\{\emptyset,\{x\}\}$, which has a cardinality of $2^{||\mathcal{X}||}$, or $2 ^{1}$.\\\\
\textit{\textbf{Assume:}} $||\mathcal{P}(\{x_{0}, x_{1}, x_{2}, \ldots, x_{n}\})|| = 2^{||\mathcal{X}||}$. \\\\
\textit{\textbf{Prove:}} $||\mathcal{P}(\{x_{0}, x_{1}, x_{2}, \ldots,x_{n}, x_{n+1}\})|| = 2^{||\mathcal{X}|| + 1}$.\\
We know that $\mathcal{P}$ is the set of all subsets of $\mathcal{X}$. 
If we count the number of subsets of $\{x_{0}, x_{1}, x_{2},\ldots,x_{n},x_{n+1}\}$, we know that the subset will either contain $x_{n+1}$, or it will not contain $x_{n+1}$.
If the subset $\gamma$ does not contain $x_{n+1}$, then $\gamma \subseteq \{x_{0},x_{1},x_{2},\ldots,x_{n}\}$, and there are $2^{||\mathcal{X}||}$ $\gamma$ by the induction hypothesis.
If the subset $\lambda$ contains $x_{n+1}$, then it is the result of some set $\gamma \cup \lambda$.
Since $\gamma \subseteq\{x_{0},x_{1},x_{2},\ldots,x_{n}\}$, we only need $\gamma \cup \{x_{n+1}\}$ to account for all possible sets.
Therefore $||\mathcal{P}(\gamma \cup \{x_{n+1}\})||$ is $||\mathcal{P}(\gamma)|| \cdot ||\mathcal{P}(\{x_{n+1}\})||$, or $2^{||\mathcal{X}||} \cdot 2^{||\{x_{n+1}\}||}$.
This is equivalent to $2^{||\mathcal{X}||}\cdot2^{1}$, or $2^{||\mathcal{X}|| + 1}$.
\end{quote}
$\triangle$
\item Let $n \geq 2$ be an integer.
Prove that $a \equiv b(\text{mod}\ n)$ if $n|(b-a)$ is an equivalence relation on $\ZZ$.
\begin{quote}
\Proof Let $n \geq 2$ and $n | (b - a)$.
Therefore $b - a  = nk,\ k\in\ZZ$.
Then $b - a \equiv 0\ mod\ n$.
It follows that $b \equiv a$, and since $mod\ n$ is an equivalence relation, $a \equiv b$ by symmetry.
\end{quote}
$\triangle$
\item Let $A$ and $B$ be sets.
Prove that $\overline{A \cup B} = \overline{A} \cap \overline{B}$.
\begin{quote}
\Proof\\ 
{$\overline{A \cup B} \subseteq \overline{A} \cap \overline{B}$:} 
Let $x\in\overline{A \cup B}$.
Then $x \notin A$ or $B$.
Since $x \notin A$, $x \in \overline{A}$.
Since $x \notin B$, $x \in \overline{B}$.
Therefore $x \in \overline{A}$ and $x\in\overline{B}$, or $x \in \overline{A} \cap \overline{B}$.\\\\
{$\overline{A \cup B} \supseteq \overline{A} \cap \overline{B}$:} 
Let $x\in\overline{A} \cap \overline{B}$.
Then $x \notin A$ and $x \notin B$.
Therefore $x \notin A \cup B$, or $x \in \overline{A \cup B}$.
\end{quote}
$\triangle$
\item Prove that $\sqrt{5}$ is irrational.
\begin{quote}
\Proof Assume $\sqrt{5}$ is rational.
Then $\sqrt{5} = \frac{p}{q}$ for some $p,q \in \ZZ$ with $q \neq 0$, and $p$ and $q$ have opposite parity.
It follows that $5 = \frac{p^{2}}{q^{2}}$, or $5q^{2} = p^{2}$.
If $q$ is odd, then $p$ must be even.
However, an odd number squared and multiplied by an odd number is odd, which means that $5q^{2}\in\{2k+1|k\in\ZZ\}$. 
This implies that $p^{2}$ is odd, but $p$ is even, so $p^{2}$ is even.
Therefore by contradiction, $\sqrt{5}$ must be irrational.
\end{quote}
$\triangle$
\end{enumerate}
\end{enumerate}
\end{document}

### 2015-12-02 proofs/portfolio.tex
semester is almost over!
finished portfolio

@@
 \begin{document}
+\noindent\textit{\textbf{Steven Rosendahl}}\\
+\textit{\textbf{Proofs Homework}}
 \begin{enumerate}
@@
 \textit{\textbf{Prove: }}  
+\[
+\sum_{i = 1}^{n + 1}i^{3} = \left(\frac{(n + 1)(n+2)}{2}\right)^{2}
+\]
 \begin{align*}
-\sum_{i = 1}^{n + 1}i^{3} &= \left(\frac{(n + 1)(n+2)}{2}\right)^{2}\\
-\sum_{i = 1}^{n}i^{3} + \sum_{i = n}^{n + 1}i^{3} &= \left(\frac{(n + 1)(n+2)}{2}\right)^{2}\\
-\left(\frac{n(n+1)}{2}\right)^{2} + (n + 1)^{3} &= \left(\frac{(n + 1)(n+2)}{2}\right)^{2}\ \text{By Induction Hypothesis}\\
-n^{2}(n+1)^{2} + 4(n+1)^{3} &= (n+1)^{2}(n+2)^{2}\\
-(n+1)^{2}(n^{2} + 4n + 4) &= (n+1)^{2}(n+2)^{2}\\
-(n+2)^{2} &= (n+2)^{2}
+\therefore \sum_{i = 1}^{n}i^{3} + \sum_{i = n + 1}^{n + 1}i^{3} &= \left(\frac{n(n+1)}{2}\right)^{2} + (n + 1)^{3}\ \ \text{By the Induction Hypothesis}\\
+&= \frac{n^{2}(n+1)^{2}}{4} + (n+1)^{3}\\
+&= \frac{n^{2}(n+1)^{2}}{4} + \frac{4(n+1)^{3}}{4}\\
+&= \frac{n^{2}(n+1)^{2} + 4(n+1)^{3}}{4}\\
+&= \frac{(n+1)^{2}(n^{2} + 4n + 4)}{4}\\
+&= \frac{(n+1)^{2}(n+2)^{2}}{2^{2}}\\
+&= \left(\frac{(n+1)(n+2)}{2}\right)^{2}
 \end{align*}
@@
 \begin{quote}
-\Proof Assume that if $n$ is even, then $n^{3}$ is even.
+\Proof Assume the contrapositive: if $n$ is even, then $n^{3}$ is even.
 Then $n = 2k$ for $k\in \ZZ$, which mean that $n^{3} = (2k)^{3}$.
@@
 \item Let $n \geq 2$ be an integer.
-Prove that $a \equiv b(\text{mod}\ n)$ if $n|(b-a)$ is an equivalence relation on $\ZZ$.
+Prove that $a \equiv b(\text{mod}\ n)$ is an equivalence relation on $\ZZ$.
 \begin{quote}
-\Proof Let $n \geq 2$ and $n | (b - a)$.
-Therefore $b - a  = nk,\ k\in\ZZ$.
-Then $b - a \equiv 0\ mod\ n$.
-It follows that $b \equiv a$, and since $mod\ n$ is an equivalence relation, $a \equiv b$ by symmetry.
+Let $R$ be the relation $a\equiv b\ mod\ n$.\\
+\textit{\textbf{Symmetric:}} Let $aRb$.
+Then $a \equiv b(\text{mod}\ n)$, or $a - b = nk$ for $k\in\ZZ$.
+It follows that
+\begin{align*}
+a &= nk + b\\
+-nk &= b-a\\
+nj &= b-a,\in\ZZ\\
+\end{align*}
+Therefore $b \equiv a\ mod\ n$, and $R$ is symmetric.\\\\
+\textit{\textbf{Reflexive:}} Let $aRa$.
+Then $a \equiv a\ mod\ n$.
+It follows that $n|(a-a)$, or $n|0$.
+Since $n \geq 2$, $n|0$, and $R$ is reflexive.\\\\
+\textit{\textbf{Transitive:}} Let $aRb$ and $bRc$.
+Then $a\equiv b\ mod\ n$ and $b\equiv c\ mod\ n$.
+By definition, $a-b = nk,\ k\in\ZZ$ and $b - c = nj,\ j\in\ZZ$.
+Then $a-b = a-nj-c = nk$.
+It follows that $a-c = nk + nj$, or $a-c = n(k + j)$.
+Therefore $a\equiv b\ mod\ n$, and $R$ is transitive\\\\
+Therefore $R$ is an equivalence relation.
+%\Proof Let $n \geq 2$ and $n | (b - a)$.
+%Therefore $b - a  = nk,\ k\in\ZZ$.
+%Then $b - a \equiv 0\ mod\ n$.
+%It follows that $b \equiv a$, and since $mod\ n$ is an equivalence relation, $a \equiv b$ by symmetry.
 \end{quote}
@@
 \begin{quote}
-\Proof Assume $\sqrt{5}$ is rational.
-Then $\sqrt{5} = \frac{p}{q}$ for some $p,q \in \ZZ$ with $q \neq 0$, and $p$ and $q$ have opposite parity.
-It follows that $5 = \frac{p^{2}}{q^{2}}$, or $5q^{2} = p^{2}$.
-If $q$ is odd, then $p$ must be even.
-However, an odd number squared and multiplied by an odd number is odd, which means that $5q^{2}\in\{2k+1|k\in\ZZ\}$. 
-This implies that $p^{2}$ is odd, but $p$ is even, so $p^{2}$ is even.
-Therefore by contradiction, $\sqrt{5}$ must be irrational.
+\Proof Let $p$ be a prime number, and assume that $\sqrt{p}$ is rational.
+Then $\sqrt{p} = \frac{n}{m}$ for $n,m \in \NN$.
+It follows that $n^{2} p = m^{2}$.
+We know that there are two factors of $p$, namely 1 and $p$, and that a squared number will have an even number of prime factors, since it has double the prime factors as its root.
+Then $n^{2}p$ will have an odd number of prime factors, since its prime factors are the prime factors of $n^{2}$ and the number $p$.
+Since $n^{2}p = m^{2}$, $m^{2}$ must also have an odd number of prime factors.
+However, $m^{2}$ has an even number of prime factors.
+Therefore, by contradiction, the root of a prime number is irrational.
+Therefore, since 5 is prime, $\sqrt{5}$ is irrational. $\triangle$
 \end{quote}
-$\triangle$
 \end{enumerate}

### 2015-12-03 cheatsheets/calctwo.tex
i dont even know what i'm doing anymore

\documentclass{article}
\usepackage{amssymb,amsmath}
\usepackage{mathtools}
\usepackage{array}
\topmargin=-0.3in
\textheight=9.2in
\textwidth=168mm
\oddsidemargin=-0.2in
\evensidemargin=-0.2in
\title{Calculus Two Review}
\date{}

\newcommand{\beq}{\begin{quote}}
\newcommand{\eeq}{\end{quote}}
\newcommand{\dx}{\mathop{dx}}

\DeclareMathOperator{\arcsec}{arcsec}

\begin{document}
\section{Logarithmic Equations}
\begin{minipage}[t]{0.3\textwidth}
It can often be helpful to express a logarithm in terms of base $e$.
The following relationship can help us get to the natural log, $\ln$.
\[\log_{n}x = \frac{\ln{x}}{\ln{n}}\]
\end{minipage}
\hspace{0.5cm}
\begin{minipage}[t]{0.3\textwidth}
We can also use the properties of logarithmic functions to simplify exponential functions.
\[n^{x} = e^{\ln{x}}\]
\end{minipage}
\hspace{0.5cm}
\begin{minipage}[t]{0.3\textwidth}
Applying calculus to logarithmic and exponential functions can yield interesting results.
\[\frac{d}{dx}(\log_{n}{(x)}) = \frac{1}{x\ln{(n)}}\]
\[\int n^{x}dx = \frac{n^{x}}{\ln{(n)}}\]
\end{minipage}

\section{Inverse Trigonometry Functions}
The derivatives of inverse trig functions are useful for integration.
We have the following derivatives:
\[
\frac{d}{dx}(\arcsin{(x)}) = \frac{1}{\sqrt{1-x^{2}}}\qquad\qquad
\frac{d}{dx}(\arctan{(x)}) = \frac{1}{1+x^{2}}\qquad\qquad
\frac{d}{dx}(\arcsec{(x)}) = \frac{1}{|x|\sqrt{1+x^{2}}}
\]

\section{Integration Techniques}
The main focus of Calculus II is integration.
In Calculus I, we learned how to integrate by using a $u$ substitution, but that technique is often not powerful enough to handle more complex integrals.
Luckily, we have several different methods that allow us to integrate.
\subsection{Integration By Parts}
\[\int u dv = uv = \int v du\]
This technique is useful for integrating two functions, $f(x)$ and $g(x)$.
Suppose we want to find the integral $\int xe^{x^{2}}dx$, or $\int x\sin{\pi x}dx$, or even $\int (3t + 5)\ln{(t/5)}dt$.
All of these examples are set up to be solved by integration by parts.
We want to split the integrand into two separate functions.
We then differentiate one ($u$), and integrate the other ($dv$).
Applying our formula will often simplify the original integral into something that we know how to easily integrate.
Typically, when we choose our $u$ and $dv$, we want to choose the function whose integral we know for $dv$, and the other one for $u$ (remember, it is often easier to take a derivative than integrate).

\subsection{Trig Combination}
Ultimately, this technique will end up using integration by parts in most cases.
There are a few trig identities that we need to know first:
\[
\sin^{2}{(x)} + \cos^{2}{(x)} = 1\qquad\qquad
\sin^{2}{(x)} = \frac{1}{2}(1-\cos{(2x)})\qquad\qquad
\cos^{2}{(x)} = \frac{1}{2}(1+\cos{(2x)})
\]
There are some common patterns that often appear in integrals that we can solve using theis method.
\begin{enumerate}
\item If we see a $\cos{x}$ to an odd power, then we should let $u = \sin{x}$.
\item If we see a $\sin{x}$ to an odd power, then we should let $u = \cos{x}$.
\item If we see a $\sec{x}$ to an odd power, then we should let $u = \tan{x}$.
\item If we have a $\tan{x}$ to an odd power, and we have $\sec{x}$ present, then we should let $u=\sec{x}$.
\item If we have $\sec{x}$ to an odd power and $\tan{x}$ to an even power, then convert the $\tan{x}$ to $\sec{x}$.
\item If we \textbf{only} have $\sec{x}$ to an odd power, then we need to use integration by parts.
\item If we have a $\tan{x}$ and do \textbf{not} have a $\sec{x}$ present, then we need to convert the $\tan{x}$ into $\sec{x}$.
\item If we see a trig function to an even power, then we need to use the half angle formulas first.
\end{enumerate}

\subsection{Trig Substitution}
This technique for integration is applicable mostly to integrals that contain functions that look like the derivatives of inverse trig functions.
We have a few special cases that tell us what substitution to make.
\begin{alignat*}{3}
\sqrt{a^{2} - u^{2}} &\to 1 - \sin^{2}{\theta} = \cos^{2}{\theta} &&\to u = a\sin{\theta}\\
\sqrt{a^{2} + u^{2}} &\to 1 + \tan^{2}{\theta} = \sec^{2}{\theta} &&\to u = a\tan{\theta}\\
\sqrt{u^{2} + a^{2}} &\to \sec^{2}{\theta} - 1 = \tan^{2}{\theta} &&\to u = a\sec{\theta}\\
\end{alignat*}

\subsection{Partial Fractions}
We want to use this technique to solve fraction of functions:
\[
\int \frac{\mathcal{P}(x)}{\mathcal{Q}(x)}dx\ \text{Where the degree of }\mathcal{P}\ \text{is less than the degree of }\mathcal{Q}
\]
To use this method, we need to
\begin{enumerate}
\item Factor the denomiator
\item Determine the decomposition
\item Determine the unknown values of the numerator
\end{enumerate}

\section{Series and Sequences}
\subsection{Telescoping Series}
A telescoping series takes the form of
\[
\sum_{n=1}^{\infty} \frac{1}{n} - \frac{1}{n+1}
\]
The series will expand in such a way that all the terms will cancel except for the first and last terms.
This allows us to say that the series will converge at
\[
\lim_{n\to\infty} 1 - \frac{1}{n+1}
\]

\subsection{Geometric Series}
A geometric series has the form
\[
\sum_{n=1}^{\infty}ar^{n-1}
\]
where $a\neq 0$ and $r$ is constant.
We know that
\[
|r| \geq 1 \implies\ \text{Divergence}\qquad\qquad |r| < 1\implies\ \text{Convergence}
\]
If the sequence converges, it converges to
\[
\frac{s_{0}}{1 - r}
\]
where $s_{0}$ is the first term in the series.

\subsection{P Series}
\[
\sum_{n=1}^{\infty}\frac{1}{n^{p}}
\]
If $p \leq 1$, then the series diverges. If $p > 1$, then the series converges.

\subsection{Tests of Convergence}
\subsubsection{The Divergence Test}
If
\[
\lim_{n\to\infty}a_{n}\neq 0
\]
then the series $\sum a_{n}$ diverges.
Note that this test \textbf{only} determines when the series diverges.
It says nothing about the series converging.

\subsubsection{Comparison Test}
Let $\sum a_{n}$ and $\sum b_{n}$ be positive termed series.
\begin{itemize}
\item If $b_{n}$ converges, $a_{n} < b_{n} \implies a_{n}$ converges.
\item If $b_{n}$ diverges, $a_{n} > b_{n} \implies a_{n}$ diverges.
\end{itemize}

\subsubsection{Limit Comparison Test}
Let $\sum a_{n}$ and $\sum b_{n}$ be positive termed series.
If
\[
\lim_{n\to\infty}\frac{a_{n}}{b_{n}} = \mathcal{L}
\]
then
\begin{enumerate}
\item If $\mathcal{L} = 0$ and $\sum b_{n}$ converges, then $\sum a_{n}$ converges.
\item If $\mathcal{L} \to \infty$, and $\sum b_{n}$ diverges, then $\sum a_{n}$ diverges.
\item If $\mathcal{L} > 0$, then both series either both converge or both diverge.
\end{enumerate}

\subsubsection{Ratio and Root Test}
\[
\lim_{n\to\infty}\frac{a_{n+1}}{a_{n}} = \mathcal{L}\qquad\qquad\qquad
\lim_{n\to\infty}\sqrt[\leftroot{-2}\uproot{2}n]{a_{n}} = \mathcal{L}
\]
\begin{enumerate}
\item $\mathcal{L} < 1 \implies a_{n}$ converges
\item $\mathcal{L} > 1 \implies a_{n}$ diverges
\item $\mathcal{L} = 1$ is inconclusive
\end{enumerate}

\subsubsection{Alternating Series Test}
The alternating series
\[
\sum (-1)^{n+1}a_{n}
\]
converges if
\begin{enumerate}
\item $a_{n+1} \leq a_{n}$
\item $\lim_{n\to\infty}a_{n} = 0$
\end{enumerate}

\subsubsection{Integral Test}
Given the series
\[\sum_{n=j}^{\infty} a_{n}\]
let $f(n) = a_{n}$.
The series will converge if and only if
\[
\int_{j}^{\infty}f(n)dn
\]
exists.

\subsection{Types of Convergence}
\subsubsection{Absolute Convergence}
A series converges absolutely if both $\sum a_{n}$ and $\sum |a_{n}|$ converge.

\subsubsection{Conditional Convergence}
A series converges conditionally if $\sum a_{n}$ converges and $\sum |a_{n}|$ diverges.

\subsection{Special Series}
\subsubsection{Power Series}
A power series is a special type of p-series that takes the form
\[
\sum_{n=0}^{\infty} a_{n}(x-x_{0})^{n}
\]
All power series converge for $x = x_{0}$.
With power series, we often want to know the radius of convergence of the series.
The radius of convergence, $\mathcal{R}$, can be determined by finding the distance that $x$ can move from $x_{0}$ while the series still converges.
The interval $\mathcal{I} = [x_{0} - \mathcal{R}, x_{0} + \mathcal{R}]$ is called the interval of convergence.

\subsubsection{Taylor Series}
The Taylor series representation of a function can be given by
\[
\sum_{n=0}^{\infty}\frac{f^{(n)}(x_{0})}{n!}(x-x_{0})^{n}
\]
where $f^{(n)}$ is the $n^{\text{th}}$ derivative of $f$.
Taylor series provide an appromation of a function.

\subsubsection{Binomial Series}
A binomial series is given by
\[
\sum_{k=0}^{n}\frac{n!}{k!(n-k)!}x^{k}
\]
\end{document}

### 2015-12-03 comporg/review.tex
i dont even know what i'm doing anymore

--- 2015-11-16 cheatsheets/ode_formal_slns.tex
@@
 \usepackage{mathtools}
+\usepackage[table]{xcolor}
 \usepackage{array}
@@
 \evensidemargin=-0.2in
-\title{Formal Solutions To ODE's}
-\date{}
+
 \begin{document}
-\maketitle
-
-\section{Homogeneous Solutions}
-
-\subsection{First Order}
-
-We have the following differential equation
-\[
-	y'+ay=0.
-\]
-Then
-\[
-	y=c e^{-at},
-\]
-where $c$ is determined by the initial condition.
-
-\subsection{Second Order}
-
-We have the following differential equation
-\begin{equation}
-	y''+ay'+by=0. \label{eqn1}
-\end{equation}
-Then we solve the following equation for $r$,
-\[
-	r^2+ar+b=0.
-\]
-
-If $r=\alpha\pm i\beta$, then the general solution is
-\[
-	y=c_1 e^{\alpha x}\cos(\beta x)+c_2 e^{\alpha x}\sin(\beta x)
-\]
-where $c_1$ and $c_2$ are determined by the initial conditions.
-
-If \autoref{eqn1} has two real solutions, $r_1$ and $r_2$, then
-\[
-	y=c_1e^{r_1 x}+c_2 e^{r_2x}
-\]
-where $c_1$ and $c_2$ are determined by the initial conditions.
-
-If \autoref{eqn1} has a real repeated root, $r$, then
-\[
-	y=c_1e^{rx}+c_2xe^{rx}
-\]
-where $c_1$ and $c_2$ are determined by the initial conditions.
 

### 2015-12-05 comporg/review.tex
cleaning up

@@
 \usepackage{mathtools}
-\usepackage[table]{xcolor}
 \usepackage{array}
-\usepackage{hyperref}
+\usepackage{color}
+\usepackage{graphicx}
+\usepackage{minted}
+% \pagecolor{black}
+% \color{white}
 \topmargin=-0.3in
@@
 \evensidemargin=-0.2in
-
 \begin{document}
-
+\begin{enumerate}
+\item Identify the data hazards in the following code, and fix them using $nop$ (software).
+\begin{center}
+\begin{minted}[linenos=true]{nasm}
+sub $2, $1, $3
+and $12, $2, $5
+or  $13, $6, $2
+add $14, $2, $2
+sw  $15, 100($2)
+\end{minted}
+\end{center}
+\indent The data hazard is present in the $\$2$ register.
+The $sub$ instruction will take three cycles to write to the $\$2$ register, but the $and$ and $or$ instructions are attempting to access $\$2$ before the register has a value in it.
+As a result, two $nop$ instructions should be inserted after the $sub$ instruction to offset the data access.
+\begin{minted}[linenos=true]{nasm}
+sub $2, $1, $3
+nop
+nop
+and $12, $2, $5
+or  $13, $6, $2
+add $14, $2, $2
+sw  $15, 100($2)
+\end{minted}
+\item Identify and fix the data hazards shown below by showing the fowarding paths (hardware).
+\begin{minted}[linenos=true]{nasm}
+add $3, $4, $6
+sub $5, $3, $2
+lw  $7, 100($5)
+%add $8, $7, $2
+\end{minted}
+There are two data hazards in this set of instructions.
+In line two, register $\$3$ is being accessed by the $sub$ instruction before it is stored by the $add$ instruction.
+The other hazard is with the $\$7$ register, which is used by the $add$ instruction before a value is loaded by the $lw$ instruction.
+\item Compare the performance for the single-cycle, multi-cycle, and piplined control using the following instruction mix.
+\begin{center}
+\begin{tabular}{|l|c|c|c|c|c|}
+\hline
+& Loads & Stores & Branches & Jumps & ALU\\
+\hline
+Instruction Mix & 25\% & 10\% & 11\% & 2\% & 52\%\\
+\hline
+Clock Cycles & 5 & 4 & 3 & 3 & 4\\
+\hline
+\end{tabular}
+\end{center}
+The performance for a single cycle machine is 200ps for memory access, 100ps for ALU operations, and 50ps for register read and write.
+\begin{enumerate}
+\item What is the clock cycle time for a single-cycle data path?
+\item What is the average CPI for the multi-cycle design?
+\item What is the average CPI for the piplined design?
+\item What are the average instruction times for each designs?
+\end{enumerate}
+\item Assume an instruction cache miss rate for a program is 2\% and a data cache miss rate is 4\%.
+If the processor has a CPI of 2 without any memory stalls and the miss penalty if 100 cycles for all misses, determine how much faster a processor with a perfect cache that never misses would be.
+The frequency of all loads and stores is 36\% for the SPECint2000.
+\end{enumerate}
 \end{document}

### 2015-12-05 comporg/review.tex
all day studying

@@
 \evensidemargin=-0.2in
+
+\newcommand{\sqr}{$\square \qquad|$ & $\square \qquad|$ & $>\qquad|$ & $\square\qquad|$ & $\square\ \qquad$ }
+
 \begin{document}
+\begin{enumerate}
+\item $CPU\ Time = \frac{Clock\ Cycles}{Clock\ Rate}$
+\item $CPU\ Time = Instructions \times \frac{Cycles}{Instruction} \times \frac{Seconds}{Cycle}$
+\item $CPU\ Time = Clock\ Cycles \times Cycle\ Time$
+\item $CPI = \frac{Clock\ Cycles}{Instructions}$
+\item $CPU\ Time = Instruction\ Count \times CPI \times Cycle\ Time$
+\item $CPU\ Time = Instruction\ Count \times \frac{CPI}{Clock\ Rate}$
+\item $(Clock\ Cycles)_{Total} = \sum_{i = 1}^{n}\left(CPI_{i} \times Count_{i}\right)$
+\item $MIPS = \frac{Instruction\ Count}{CPU\ Time \times 10^{6}}$
+\item $Power = Capacitive\ Load \times (Voltage)^{2} \times Frequency$
+\item $AMAT = Hit\ Time + Miss\ Rate \times Miss\ Penalty$
+\end{enumerate}
+\hrule
 \begin{enumerate}
@@
 In line two, register $\$3$ is being accessed by the $sub$ instruction before it is stored by the $add$ instruction.
-The other hazard is with the $\$7$ register, which is used by the $add$ instruction before a value is loaded by the $lw$ instruction.
+The other hazard is with the $\$7$ register, which is used by the $add$ instruction before a value is loaded by the $lw$ instruction.\\
+% \begin{tabular}{l c c c c c c c c c}
+% $add$ & \sqr & & & &\\
+% $sub$ & & \sqr & & &\\
+% $lw$  & & & \sqr & &\\
+% $nop$ & & & & \sqr &\\
+% $add$ & & & &  &\sqr\\
+% \end{tabular}
 \item Compare the performance for the single-cycle, multi-cycle, and piplined control using the following instruction mix.
@@
 The performance for a single cycle machine is 200ps for memory access, 100ps for ALU operations, and 50ps for register read and write.
+In a piplined implementation, loads, stores, and ALU operations take 1 clock cycle.
+Branches take 1 clock cycle when predicted correctly, and 2 when not;
+Jumps take 2 clock cycles.
 \begin{enumerate}
-\item What is the clock cycle time for a single-cycle data path?
-\item What is the average CPI for the multi-cycle design?
-\item What is the average CPI for the piplined design?
+\item What is the clock cycle time for a single-cycle data path?\\
+\indent In order calculate the clock cycle time, we need to know how long each instruction takes to execute.
+We can create the following table to calculate the total time:\\
+
+\hspace*{-1.7cm}
+\begin{tabular}{|c|c|c|c|c|c|c|}
+\hline
+Instruction Class & Instruction Memory & Register Read & ALU Op & Data Memory & Register Write & Total\\
+\hline
+R-Type & 200 & 50 & 100 & 0 & 50 & 400\\
+\hline
+Load Word & 200 & 50 & 100 & 200 & 50 & 600\\
+\hline
+Store Word & 200 & 50 & 100 & 200 & 0 & 550\\
+\hline
+Branch & 200 & 50 & 100 & 0 & 0 & 350\\
+\hline
+Jump & 200 & 0 & 0 & 0 & 0 & 200\\
+\hline
+\end{tabular}
+\vspace{0.1cm}
+
+The single cycle implementation is bottlenecked by the longest instruction, which happens to be the Load Word instruction.
+Therefore, the clock cycle time is 600ps.
+\item What is the average CPI for the multi-cycle design?\\
+\indent The CPI for the multi-cycle design is dependant on the number of types of instruction that appear in the program.
+To calculate the CPI, we have
+\begin{align*}
+CPI &= (0.25)(5) + (0.1)(4) + (0.11)(3) + (0.02)(3) + (0.52)(4)\\
+&= 4.12\ \text{cycles per instruction}
+\end{align*}
+\item What is the average CPI for the piplined design?\\
+\indent The pipelined design's CPI, like the multi-cycle, is determined by the number of types of instructions that appear in the program.
+However, the piplined cycle uses less clock cycles per instruction in this implementation.
+In addition, the branch instructions have a $50\%$ of being correct, so we need to find the average clock cycle time for the branches.
+We can find the CPI for the pipelined design as follows:
+\begin{align*}
+CPI &= (0.25)(1) + (0.1)(1) + (0.11)\left(\frac{1 + 2}{2}\right) + (0.02)(2) + (0.52)(1)\\
+&= 1.075\ \text{cycles per instruction}
+\end{align*}
 \item What are the average instruction times for each designs?
+\indent I don't know.
 \end{enumerate}
 \item Assume an instruction cache miss rate for a program is 2\% and a data cache miss rate is 4\%.
-If the processor has a CPI of 2 without any memory stalls and the miss penalty if 100 cycles for all misses, determine how much faster a processor with a perfect cache that never misses would be.
-The frequency of all loads and stores is 36\% for the SPECint2000.
+If the processor has a CPI of 2 without any memory stalls and the miss penalty of 100 cycles for all misses, determine how much faster a processor with a perfect cache that never misses would be.
+The frequency of all loads and stores is 36\% for the SPECint2000.\\
+\indent We will start by letting the instruction count $= \mathcal{I}$.
+We know how to calculate the number of memory miss cycles:
+\[
+\mathcal{I}\times 0.02\times 100\text{ cycles} = 2\mathcal{I}.
+\]
+This result follows from the fact that the \textbf{cache} miss rate is 2\% with a penalty of 100 clock cycles.
+We now need to calculate the number of memory miss cycles for data references.
+Again, we know the frequency an penalty for \textbf{data cache} misses:
+\[
+\mathcal{I}\times 0.04\times 0.36\times 100 = 1.44\mathcal{I}.
+\]
+As a result, the total number of misses is $1.44\mathcal{I} + 2\mathcal{I}$, or $3.44\mathcal{I}$.
+To obtain the CPI with stalls, we simply add the number of misses divided by the instruction count to the perfect CPI.
+\[
+\text{CPI}_{\text{Misses}} = 3.44 + 2 = 5.44.
+\]
+The ratio of the imperfect cache to the perfect cache is
+\[
+\frac{5.44}{2} = 2.72,
+\]
+so the perfect cache is 2.72 times faster than the imperfect cache.
+\item For the following cache, answer the following questions.
+\begin{center}
+\includegraphics[scale=0.7]{hw5_graphic}
+\end{center}
+\begin{enumerate}
+\item Compute the total size of the cache.\\
+If we had not been given the number of entries in the cache, we could compute it as $2^{\text{index}}$, which is $2^{8}$, or $256$ in this case.
+Each entry in the cache occupies 512 bits of memory, so the total size of the cache is $256 * 512 = 131072$.
+\item Compute the total number of bits required to implement this cache for the Intrinsity FastMATH embedded microprocessor.\\
+The total number of bits needed is determined by the total size of each entry in the cache.
+There are 512 bits of memory in each entry, but there is also a 1 bit data entry and an 18 bit tag, which means each entry in the cache is $512 + 1 + 18 = 531$ bits.
+With 256 entries, we have $531 * 256 = 135936$ bits.
+\end{enumerate}
+\item Consider a memory hierarchy using one of the three organizations for main memory shown below.
+Assume that the cache block size is 16 words, that the width of organization (b) of the figure is four words,
+and that the number of banks in organization (b) of the figure is four.
+If the main memory latency for a new access is 10 memory bus clock cycles and the transfer time is 1 memory bus clock cycle, what are the miss penalties (in clock cycles) for each of these organizations?
+\begin{center}
+\includegraphics[scale=0.5]{hw5_graphic2}
+\end{center}
+\begin{enumerate}
+\item For each organization, we will always have to go through the bus to access memory.
+We know that we only have a miss if we have to move from the cache into memory and return the fetched memory back to the CPU.
+For this particular computer, it takes 10 clock cycles to retrieve the memory.
+As a result, we will have the following time for a miss penalty in the cache
+\[
+\text{Miss} = 1 + (10)(16) + (16)(1) = 177.
+\]
+\item For this organization, we have changed the amount of memory we have to search through; specifically, we now only need to access a quarter of the amount of memory that we have had to in (a).
+We do still have to go though the bus, so we have
+\[
+\text{Miss} = 1 + (10)\left(\frac{16}{4}\right) + \left(\frac{16}{4}\right)(1) = 45.
+\]
+\item Finally, in this organization our memory has been split into four separate cells.
+We still have to travese through the bus, so we know that we will have to wait the usual clock cycle.
+Similar to (b), we will have to only search through a quarter of the memory to find the address that we need; we will however still have to transmit 16 words through the bus back to the cache.
+This setup yields
+\[
+\text{Miss} = 1 + (10)\left(\frac{16}{4}\right) + (16)(1) = 57.
+\]
+\end{enumerate}
 \end{enumerate}

### 2015-12-14 contract/contact.tex
added linear algebra submodule

@@
 \end{multicols}
+The application will undergo testing before release and during times when the application is being upgraded, and will be compleated in a timely fashion as agreed upon by the developer and the contractor.
 \vspace{1cm}
@@
 \indent At the time when the developer leaves the project, the developer agrees to transfer ownership of the repository, ownership of the database, all application keys, and all releases of the application to whomever the owner appoints.
+In addition, the developer must be allowed to claim credit for the devlopment of the application both before and after the application is released.
 \vspace{1cm}

### 2015-12-14 updater.sh
added updater and update gitignore

@@ -1,0 +1,15 @@
+MESSAGE=$1
+
+if [ -z "$MESSAGE" ]
+then
+	echo "No Commit Message Provided!"
+	echo "Exiting..."
+else
+	cd linear-algebra-notes
+	git checkout master
+	git pull
+	cd ..
+	git add .
+	git commit -m "$MESSAGE"
+	git push
+fi

### 2015-12-20 proofs/defs.tex
create proofs defs

\documentclass{article}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{array}
\usepackage{amssymb}
\topmargin=-0.3in
\textheight=9.2in
\textwidth=168mm
\oddsidemargin=-0.2in
\evensidemargin=-0.2in

\newcommand{\deff}[1]{\textbf{\textit{\underline{#1}}}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\PS}{\mathcal{P}}
\newcommand{\US}{\mathcal{U}}
\newcommand{\rel}{\mathcal{R}}

\newcommand{\seta}{\mathbb{A}}
\newcommand{\setb}{\mathbb{B}}
\newcommand{\setc}{\mathbb{C}}
\newcommand{\setd}{\mathbb{D}}

\title{Proofs Definitions}
\date{}
\begin{document}
\maketitle
\begin{enumerate}
\item \deff{Set:} A set is a collection of objects.
The items in a set are called elements.
\item \deff{The Empty Set:} The empty set, given by $\emptyset$, is the set with
no elements.
\item \deff{Subset:} Let $\seta$ and $\setb$ be sets. Then $\seta\subset\setb$ if
$\forall x\in\seta$, $x\in\setb$.
\item \deff{Power Set:} Let $\seta$ be a set. Then the Power Set of $\seta$,
noted as $\PS(\seta)$, is the set of all subsets of $\seta$.
\item \deff{The Universal Set:} The universal set $\US$ is the set of which all
sets are subsets.
\item \deff{Intersection:} Let $\seta$ and $\setb$ be sets. The intersection of
$\seta$ and $\setb$, given by $\seta\cap\setb$ is
$\{x|x\in\seta\land x\in\setb\}$.
\item \deff{Union:} Let $\seta$ and $\setb$ be sets. The union of
$\seta$ and $\setb$, given by $\seta\cup\setb$ is
$\{x|x\in\seta\lor x\in\setb\}$.
\item \deff{Trivial Intersection (Disjoint):} Let $\seta$ and $\setb$ be sets. If $\seta$
and $\setb$ have a trivial intersection, then $\seta\cap\setb = \emptyset$.
\item \deff{Set Difference:} Let $\seta$ and $\setb$ be sets. The set difference
of $\seta$ and $\setb$, noted as $\seta - \setb$, is
$\{x|x\in\seta\land x\notin\setb\}$.
\item \deff{Cartesian Product:} Let $\seta$ and $\setb$ be sets. Then the product
of $\seta$ and $\setb$, given by $\seta\times\setb$, is
$\{(a,b)|a\in\seta\land b\in\setb\}$.
\item \deff{Well Ordered:} A set $\seta$ is well ordered if for every non-empty
set $\setb\subset\seta$, $\setb$ has a least element.
\item \deff{Compliment:} The compliment of a set $\seta$ in regards to $\US$ is
the set difference of $\US$ and $\seta$.
\item \deff{Negation:} The negation of a statement $\rho$, given by $\neg\rho$,
is the statement that has the opposite truth values of $\rho$.
\item \deff{Disjunction:} The disjunction of statements $\rho$ and $\varphi$ is
the statement where either $\rho$ and $\varphi$ is true, or both are true, given
by $\rho\lor\varphi$.
\item \deff{Conjunction:} The conjunction of statements $\rho$ and $\varphi$ is
the statement where $\rho$ and $\varphi$ are both true, given by $\rho\land\varphi$.
\item \deff{Implication:} The implication of $\rho$ and $\varphi$, noted as
$\rho\implies\varphi$, is the statement where if $\rho$ then $\varphi$ is true.
\begin{center}
\begin{tabular}{c|c|c|c|c}
$\rho$ & $\varphi$ & $\rho\land\varphi$ & $\rho\lor\varphi$ & $\rho\implies\varphi$\\
\hline
0 & 0 & 0 & 0 & 1\\
0 & 1 & 0 & 1 & 1\\
1 & 0 & 0 & 1 & 0\\
1 & 1 & 1 & 1 & 1
\end{tabular}
\end{center}
\item \deff{Logically Equivalent:} Two statements $\rho$ and $\varphi$ are
logically equivalent, noted by $\rho\equiv\varphi$, if they have the same truth
values.
\item \deff{Tautology:} A statement $\rho$ is a tautology if it is true for all
possible truth values.
\item \deff{Contradiction:} A statement $\rho$ is a contradiction if it is false
for all possible truth values.
\item \deff{Universal Quantifier:} The universal quantifier, $\forall$, is a
quantifier that asserts that a given statement $\rho$ holds for all elements in
the specified domain.
\item \deff{Existential Quantifier:} The existential quantifier, $\exists$, is a
quantifier that assets that a given statement $\rho$ holds for at least one
element in the specified domain.
\item \deff{Axiom:} An axiom is a statement that is accepted as true without
proof.
\item \deff{Theorem:} A theorem is a statement that is proved to be true.
\item \deff{Lemma:} A lemma is a statement that serves as an intermediate step
in a proof.
\item \deff{Corollary:} A corollary is a statement that follows from an eariler
result.
\item \deff{Vacuous Statement:} A vacuous statement is a statement in which the
assumption is always false. The statement is always true.
\item \deff{Division:} Let $a,b\in\RR$. To say $a$ divides $b$, noted as $a|b$,
implies that $b=ax$ for some $x\in\ZZ$.
\item \deff{Induction:} For all $n\in\NN$, let $\rho(n)$ be a statement. If
$\rho_{0}$ is true and $\rho(n)\implies\rho(n + 1)$, then $\rho(n)$ is true for
all $n$.
\item \deff{Strong Induction:} For all $n\in\NN$, let $\rho(n)$ be a statement.
If $\rho_{0}$ is true and $\rho(i)\implies\rho(n+1)$ for all $i\in\NN$, then
$\rho(n)$ is true for all $n\in\NN$.
\item \deff{Relation:} A relation $\rel$ from set $\seta$ to set $\setb$ is a
subset of $\seta\times\setb$. Set $\seta$ is related to set $\setb$, noted as
$\seta\rel\setb$ if $(a,b)\in\rel$ for $a\in\seta$ and $b\in\setb$.
\item \deff{Inverse Relation:} Given a relation $\rel$ from $\seta$ to $\setb$,
$\rel$ inverse, noted as $\rel^{-1}$ or $\rel^{\text{opp}}$, is
$\{(b,a)|(a,b)\in\rel\}$.
\item \deff{Reflixivity:} A relation $\rel$ is reflexive on a set $\seta$ if
$a\rel a$ for all $a\in\seta$.
\item \deff{Symmetry:} A relation $\rel$ is symmetric on a set $\seta$ if
$a\rel\eta\implies\eta\rel a$ for all $a,\eta\in\rel$.
\item \deff{Transitivity:} A relation $\rel$ is transitive on a set $\seta$ if
$a\rel\eta$ and $\eta\rel\delta\implies a\rel\delta$ for $a,\eta,\delta\in\seta$.
\item \deff{Equivalence Relation:} A relation $\rel$ is an equivalence relation
if it is reflexive, symmetric, and transitive.
\item \deff{Equivalence Class:} Let $\seta$ be a non-empty set with elements
$a$ and $\eta$. The equivalence class of $a$, noted as $[a]$, is
$\{\eta|\eta\rel a\}$.
\item \deff{Partition:} Let $\seta$ be a non-empty set. A patition of $\seta$,
given by $\mathbb{P}$, is a set of subsets of $\seta$ where
$\cup_{i\in\mathbb{I}}\seta = \mathbb{P}$ and $\seta_{i}\cap\seta_{j}\neq\emptyset$
for all $i\neq j$.
\item \deff{Function:} A function $f$ from $\seta\to\setb$ is a relation from
$\seta\to\setb$ where if $(a,b)$ and $(a,c)$ are in $f$, then $b=c$, and
$\forall x\in\seta,\exists y\in\setb$ such that $(x,y)\in f$.
\item \deff{Image:} The image of a function $f:\seta\to\setb$ is
$\{f(a)|a\in\seta\}$.
\item \deff{Preimage:} Let $f$ be a function such that $f:\seta\to\setb$ and
$\setd\subset\setb$. Then the preimage of $\setd$, noted as $f^{-1}(\setd)$, is
$\{a\in\seta|f(a)\in\setd\}$.
\item \deff{Injection:} A function $f:\seta\to\setb$ is injective if
$f(a) = f(b)\implies a = b$ for all $a,b\in\seta$.
\item \deff{Surjection:} A function $f:\seta\to\setb$ is surjective if
$\forall b\in\setb, \exists a\in\seta$ such that $f(a) = b$.
\item \deff{Bijection:} A function $f:\seta\to\setb$ is bijective if it is
injective and surjective.
\item \deff{Composition:} Let $f:\seta\to\setb$ and $g:\setb\to\setd$ be
functions. Then the composition of $g$ and $f$, given by $g\circ f$, is
defined as $(g\circ f)(a) = g(f(a))$. 
\end{enumerate}
\end{document}

### 2015-12-20 pde/pdetest1.tex
edit PDE

@@
 \evensidemargin=-0.2in
+
+\title{Partial Differential Equations}
+\date{}
 \begin{document}
+\maketitle
+\newpage
 \section{First Order Partial Differential Equations}
@@
 \noindent\\\\
-\indent We will begin by solving the typical first order PDE.
-The strategy here is to break it down into and ODE or several ODEs that can be solved by methods we already have.
-It does not matter whether the PDE is homogeneous or not; we will most likely be able to \\\\
 \noindent\textbf{\textit{Ex:}} Solve the following PDE
 \[
-\begin{cases*}
-yu_{x} + xu_{y} = 0\\
-u(0,y) = e^{-y^{2}}
-\end{cases*}
-\]
-\indent \textbf{\textit{Solution:}} This is a nice PDE; there are two ways to solve it, but we will focus on the easier one.
-We can see that we have a coefficient on both $u_{x}$ and $u_{y}$.
-This tells us that when $u$ was differentiated, there was a constant $y$ and $x$.
-So now we can say that
-\[
-\frac{dy}{dx} = \frac{x}{y}
-\]
-\noindent This is a simple ODE to solve. We can use separation of variables:
-\begin{align*}
-xdx &= ydy\\
-\int{x dx} &= \int{y dy}\\
-\frac{1}{2}x^{2} &= \frac{1}{2}y^{2} + c\\
-x^{2} - y^{2} &= c_{0}\\
-y^{2} - x^{2} &= c_{1}
-\end{align*}
-\noindent You'll notice that we multiplied both sides by $-1$. This is for convenience. Now we can express the answer as
-\[
-u(x,y) = f(y^{2} - x^{2})
-\]
-\noindent And we are done.
-
-\newpage
-\noindent\textbf{\textit{Ex:}} Solve the following PDE
-\[
 u_{x} + u_{y} + u = x + y
 \]
-\indent \textbf{\textit{Solution:}} In this case, we have a non-zero term on the right hand side of the equation.
-We will use substitution to solve this.
-For the time being, we will let $T$ be $x + y$.
-To find our $S$, we can use a similar method to the one that we used above.
-We know that the coefficients on the $u_{x}$ and $u_{y}$ are $1$ in this case.
-Now we have:
+\indent \textbf{\textit{Solution:}} In this case, we have a non-zero term on the
+right hand side of the equation. We will use substitution to solve this. For the
+time being, we will let $T$ be $x + y$. To find our $S$, we can use a similar
+method to the one that we used above. We know that the coefficients on the
+$u_{x}$ and $u_{y}$ are $1$ in this case. Now we have:
 \begin{align*}
@@
 \end{align*}
-\noindent Now we can say that $S = x - y$.
-Next we have to actually use the substitution.
-We need to express $u_{x}$ and $u_{y}$ in terms of $S$ and $T$.
-The following equations come from the chain rule; when the $S$ and $T$ terms are differentiated, we get $u_{S}$ and $u_{T}$ left behind.
+\noindent Now we can say that $S = x - y$; however, we need to check our choice
+for $T$ to make sure this substitution will not just yield a trivial solution.
+We can do this by Next we have to actually use the substitution. We need to
+determine whether the coefficient matrix, or the Jacobian, of our substitutions
+has a determinant of zero. If it does, then we need to choose a different $T$.
+\[
+\left|
+\begin{array}{c c}
+1 & 1\\
+1 & -1
+\end{array}
+\right| = 2 \neq 0.
+\]
+Our choices of variables will work here, so now we need to express $u_{x}$ and
+$u_{y}$ in terms of $S$ and $T$. The following equations come from the chain
+rule; when the $S$ and $T$ terms are differentiated, we get $u_{S}$ and $u_{T}$
+left behind.
 \begin{alignat*}{3}
@@
 \end{alignat*}
-\noindent Now we have our substitution for $u_{x}$ and $u_{y}$.
-We can now plug in our values into the original equation.
-Note that we didn't do anything with $u$; it gets left alone.
-In addition, if we needed $u_{xx}$ or $u_{yy}$, we can just differentiate again.
-We will see this later when we talk about classifications.
-Now, back to plugging in:
+\noindent We can now plug in our values into the original equation. Note that we
+didn't do anything with $u$; it gets left alone. In addition, if we needed
+$u_{xx}$ or $u_{yy}$, we can just differentiate again. Plugging in yields:
 \begin{gather*}
@@
 \end{gather*}
-\noindent This is just a first order ODE. Since it's not homogeneous, we need to solve for the general solution, and then the particular solution. Let's start with the general.
+\noindent This is just a first order ODE. Since it's not homogeneous, we need to
+solve for the general solution, and then the particular solution. Let's start
+with the general.
 \begin{gather*}
@@
 \end{gather*}
-\noindent In this case, we know that our $x$ is really $T$.
-So we finally have $y = ce^{\left(-\frac{1}{2}\right)T}$.
-We will back substitute soon, but let's find the particular solution before we do that.
-\noindent Recall from ODE that we have several cases for the particular solution. In our case, we have a linear equation on the right hand side of our ODE. This lets us say:
+\noindent In this case, we know that our $x$ is really $T$. So we finally have
+$y = ce^{\left(-\frac{1}{2}\right)T}$. We will back substitute soon, but let's
+find the particular solution before we do that.
+\noindent Recall from ODE that we have several cases for the particular
+solution. In our case, we have a linear equation on the right hand side of our
+ODE. This lets us say:
 \begin{gather*}
@@
 \noindent We aren't quite done yet, as we have a solution to an ODE, not a PDE.
-It isn't too difficult to switch back; The $c$ is really the only difference.
-In our ODE, the $c$ is a constant term, but to our PDE, the $c$ is a function.
-The $c$ now becomes $f(s)$.
-We will go ahead and back substitute, which will give us $f(y-x)$.
-Now we have our final solution:
+It isn't too difficult to switch back; The $c$ is really the only difference. In
+our ODE, the $c$ is a constant term, but to our PDE, the $c$ is a function. The
+$c$ now becomes $f(s)$. We will go ahead and back substitute, which will give us
+$f(y-x)$. Now we have our final solution:
 \[
@@
 \noindent\\\\
-\indent We mentioned earlier the topic of classification of PDE's.
-Since we only have this and Fourier Series to talk about, we will go ahead and knock out classification.
-There's some information that we need to go ahead and review here.
-This will be explained in the problem, but it's handy to have it for reference:
-\begin{align*}
-b^{2} - ac < 0 &\implies Elliptic\ with\ complex\ roots\\
-b^{2} - ac = 0 &\implies Parabolic\ with\ the\ same\ root\ twice\\
-b^{2} - ac > 0 &\implies Hyperbolic\ with\ two\ distinct\ real\ roots
-\end{align*}
-\noindent Another helpful piece of information is what's known as the \textit{classical trinity} of PDEs.
-\begin{align*}
-u_{t} - ku_{xx} &= 0 \qquad The\ heat\ equation\ (parabolic)\\
-u_{tt} - c^{2}u_{xx} &= 0 \qquad The\ wave\ equation\ (hyperbolic)\\
-u_{xx} + u_{yy} &= 0 \qquad The\ Laplace\ equation\ (elliptic)
-\end{align*}
-Let's go ahead and look at an example.\\
+\indent We mentioned earlier the topic of classification of PDE's. Since we only
+have this and Fourier Series to talk about, we will go ahead and knock out
+classification. There's some information that we need to go ahead and review
+here. This will be explained in the problem, but it's handy to have it for
+reference:
+\begin{align*}
+b^{2} - ac < 0 &\implies \text{Elliptic with complex roots}\\
+b^{2} - ac = 0 &\implies \text{Parabolic with the same root twice}\\
+b^{2} - ac > 0 &\implies \text{Hyperbolic with two distinct real roots}
+\end{align*}
+\noindent Another helpful piece of information is what's known as the
+\textit{classical trinity} of PDEs.
+\begin{align*}
+u_{t} - ku_{xx} &= 0 \qquad \text{The heat equation (parabolic)}\\
+u_{tt} - c^{2}u_{xx} &= 0 \qquad \text{The wave equation (hyperbolic)}\\
+u_{xx} + u_{yy} &= 0 \qquad \text{The Laplace equation (elliptic)}
+\end{align*}
+Let's go ahead and look at an example.\\\\
 \indent \textbf{Ex. }Classify the following PDE:
 \[yu_{xx} + u_{yy} = 0\]
-\indent \textbf{\textit{Solution:}} The first thing we need to identify are the coefficients on the $u_{xx}$, $u_{yy}$, and $u_{xy}$ or $u_{yx}$ terms.
-Note that the $u_{xy}$ and $u_{yx}$ are interchangeable.
-We will assign these coefficients to $a$, $b$, and $c$ corresponding to $u_{xx}$, $u_{xy}$, and $u_{yy}$, respectively.
-Looking at the PDE, we can see that
+\indent \textbf{\textit{Solution:}} The first thing we need to identify are the
+coefficients on the $u_{xx}$, $u_{yy}$, and $u_{xy}$ or $u_{yx}$ terms. Note
+that the $u_{xy}$ and $u_{yx}$ are interchangeable. We will assign these
+coefficients to $a$, $b$, and $c$ corresponding to $u_{xx}$, $u_{xy}$, and
+$u_{yy}$, respectively. Looking at the PDE, we can see that
 \[
@@
 \]
-\noindent Now we look at $b^{2} - ac$.
-It should be noted that we are really dealing with $b = \frac{b_{0}}{2}$.
-In other words, we need to divide our $b$ value by $2$.
-In the case of the first problem, we are dealing with $b = 0$, so it's not evident.
-So, now we can go ahead and classify our PDE as follows:
-\[
-b^{2} - ac = 0 - y < 0 \implies Elliptic
-\]
-Note that we assume $y > 0$.
-So our PDE is elliptic.
-Now we need to make it look like it's corresponding equation from the classical trinity.
-In this case we have an Elliptic equation, so we want to make it look like Laplace's equation.
-We need to find values for $u_{xx}$ and $u_{yy}$, and then substitute those into the equation.
-This is similar to how we solved the first order PDE's; let's look at our characteristic equation:
+\noindent Now we look at $b^{2} - ac$. We are really dealing with
+$b = \frac{b_{0}}{2}$. In other words, we need to divide our $b$ value by $2$.
+In the case of the first problem, we are dealing with $b = 0$. Now we can
+classify our PDE as follows:
+\[
+b^{2} - ac = 0 - y < 0 \implies \text{Elliptic}
+\]
+Note that we assume $y > 0$. So our PDE is elliptic. Now we need to make it look
+like it's corresponding equation from the classical trinity. In this case we
+have an Elliptic equation, so we want to make it look like Laplace's equation.
+We need to find values for $u_{xx}$ and $u_{yy}$, and then substitute those into
+the equation. This is similar to how we solved the first order PDE's; let's look
+at our characteristic equation:
 \begin{gather*}
@@
 \noindent So, now we have two solutions, and they are both imaginary as expected.
-Next we need to find values for $S$ and $T$ that we can substitute into the equation.
-We will do both of them at the same time here.
-Notice that we will drop the $i$; we don't really care about the imaginary part:
+Next we need to find values for $S$ and $T$ that we can substitute into the
+equation. We will do both of them at the same time here. Notice that we will
+drop the $i$:
 \begin{gather*}
@@
 \end{gather*}
-\noindent Now we can substitute into our equation. Unfortunately, we will need to take the partial derivative twice in order to get the substitutions we need. Let's start with $u_{xx}$.
+\noindent Now we can substitute into our equation. Unfortunately, we will need
+to take the partial derivative twice in order to get the substitutions we need.
+Let's start with $u_{xx}$.
 \begin{align*}
@@
 \end{align*}
-\noindent That wasn't too bad. We will skip the math for finding $u_{yy}$, which would give us
+\noindent We will skip the math for finding $u_{yy}$, which would give us
 \[
@@
 \]
-\noindent Now we plug everything in. Remember that we were under the assumption that $y > 0$:
+\noindent Now we plug everything in. Remember that we were under the assumption
+that $y > 0$:
 \begin{gather*}
@@
 \subsection{Finding Fourier Series}
-\indent It's finally time to talk about Fourier Series.
-Before we start though, it may be helpful to look at some common patterns amongst the Fourier Series.
-We have the following information that will help us:\\\\
+\indent It's finally time to talk about Fourier Series. Before we start though,
+it may be helpful to look at some common patterns amongst the Fourier Series.\\\\
 \centerline{
@@
 \end{align*}
-\indent It will also help us to quickly review the concept of odd and even functions, as they will help us to simplify the process of computing Fourier series later.
-Specifically, let's look at the graphs for $\sin$ and $\cos$ on $[-\pi,\pi]$, as well as another function that will turn out to be neither even nor odd..
+\indent It will also help us to quickly review the concept of odd and even
+functions, as they will help us to simplify the process of computing Fourier
+series later. Specifically, let's look at the graphs for $\sin$ and $\cos$ on
+$[-\pi,\pi]$, as well as another function that will turn out to be neither even
+nor odd.
 \begin{center}
@@
 \noindent\\
-\indent We should recall that a function is \textit{even} when $f(x) = f(-x)$ for all $x$ in the domain of $f$.
-Graphically speaking, an \textit{even} function is reflected across only the $y$-axis.
-By looking at the above graphs, we can see that $f(x) = \cos{x}$ is an even function.\\
-\indent A function is \textit{odd} when $-f(x) = f(-x)$ for all x in the domain of $f$.
-Graphically speaking, the function is reflected across the $x$-axis and the $y$-axis.
-From the above graphs, we can say that $f(x) = \sin{x}$ is an odd function.
-Note that we can have a function that is neither odd nor even, but we won't see much of this in the following problems.
-The reason that we want to discuss odd and even functions here is because the integration that we will be seeing later will be affected by the symmetry of a function.
-Let's take a look at how even and odd functions effect integration:\\\\
-\textbf{\textit{Ex:}} Compute the following integral:
-\[
-\int_{0}^{\pi}\cos{x}dx
-\]
-\indent \textbf{\textit{Solution:}} Looking at this integral, we can see from the graph of the $\cos$ above that this will be 0.
-That answer is meaningless to us however.
-What we are really looking for is
-\[
-2\int_{0}^{\frac{\pi}{2}}\cos{x}dx
-\]
-\noindent We are allowed to do this since the $\cos$ function is even. Essentially, we have turned the negative half of the cosine into a positive half. Similarly, we could ask for something like
-\[
-\int_{0}^{\pi}|\cos{x}|dx
-\]
-\noindent Since the absolute value of a function is no integrable, we drop the absolute value signs.
-This will leave us with our initial problem, which we will have to change anyways.
-Our integral becomes:
-\[
-\int_{0}^{\pi}|\cos{x}|dx = 2\int_{0}^{\frac{\pi}{2}}\cos{x}dx = 1
-\]
-\noindent And we are done.
-\newpage
-
+\indent We should recall that a function is \textit{even} when $f(x) = f(-x)$
+for all $x$ in the domain of $f$. Graphically speaking, an \textit{even}
+function is reflected across only the $y$-axis. By looking at the above graphs,
+we can see that $f(x) = \cos{x}$ is an even function.\\
+\indent A function is \textit{odd} when $-f(x) = f(-x)$ for all x in the domain
+of $f$. Graphically speaking, the function is reflected across the $x$-axis and
+the $y$-axis. From the above graphs, we can say that $f(x) = \sin{x}$ is an odd
+function. Note that we can have a function that is neither odd nor even, but we
+won't see much of this in the following problems. The reason that we want to
+discuss odd and even functions here is because the integration that we will be
+seeing later will be affected by the symmetry.\\\\
 \indent The first type of series we will look at is the Full Fourier Series.
-Note that this series is just a combination of a Fourier Sine Series and a Fourier Cosine Series.
-A typical problem would look something like this:\\\\
+Note that this series is just a combination of a Fourier Sine Series and a
+Fourier Cosine Series. A typical problem would look something like this:\\\\
 \textbf{\textit{Ex:}} Find the corresponding Fourier Series for
@@
 \end{gather*}
-\indent \textbf{\textit{Solution:}} So we have a piecewise function here, but that shouldn't throw us off too much; all we need to do is combine the integrals for each part of the coefficient solution.
-Let's start by finding the terms for the $cos$ series.
+\indent \textbf{\textit{Solution:}} So we have a piecewise function here, but
+that shouldn't throw us off too much; all we need to do is combine the integrals
+for each part of the coefficient solution. Let's start by finding the terms for
+the $cos$ series.
 \begin{align*}
@@
 \end{align*}
-\noindent That wasn't too bad, was it? Now We need to find $B_{n}$, and then we will have our cosine part of the series. The integration can get pretty messy, so we will skip some steps.
+\noindent Now We need to find $B_{n}$, and then we will have our cosine part of
+the series. The integration can get pretty messy, so we will skip some steps.
 \begin{align*}
@@
 \end{align*}
-\noindent Note that the sequence will be 0 for all $n = 2k$. This allows us to substitute $2k+1$ for $n$, since we don't care about the $0$ terms. For our cosine part of the series, we now have:
+\noindent Note that the sequence will be 0 for all $n = 2k$. This allows us to
+substitute $2k+1$ for $n$, since we don't care about the $0$ terms. For our
+cosine part of the series, we now have:
 \begin{align*}
@@
 \end{align*}
-\noindent So, now we need to find the $\sin$ part of the series. This is easier, since we only need to find $A_{n}$.
-\begin{align*}
-A_{n} &= \frac{1}{L}\left[\int_{-L}^{0}L\sin{\left(\frac{n\pi x}{L}\right)}dx + \int_{0}^{L}2x\sin{\left(\frac{n\pi x}{L}\right)}dx \right]\\
+\noindent Now we need to find the sine part of the series. This is easier,
+since we only need to find $A_{n}$.
+\begin{align*}
+A_{n} &= \frac{1}{L}\left[\int_{-L}^{0}L\sin{\left(\frac{n\pi x}{L}\right)}dx +
+\int_{0}^{L}2x\sin{\left(\frac{n\pi x}{L}\right)}dx \right]\\
 &= \frac{L}{n\pi}\left[-1-(-1)^{n}\right]
 \end{align*}
-\noindent Note that we can do a similar substitution here as we did with the cosine series. This time, however, we will have $0$ for $n = 2k+1$ We now have:
+\noindent Note that we can do a similar substitution here as we did with the
+cosine series. This time, however, we will have $0$ for $n = 2k+1$ We now have:
 \begin{align*}
@@
 \end{align*}
-\noindent Now we have all the parts that we need. Our final answer will be essentially the sum of all the parts:
-\begin{align*}
-f(x)&\sim \frac{L}{2} + \sum_{k = 0}^{\infty}\left[\frac{-4L}{(2k+1)^{2}\pi^{2}}\cos{\left(\frac{(2k + 1)\pi x}{L}\right)} \right] + \sum_{k = 0}^{\infty}\left[\frac{-2L}{2k\pi}\sin{\left(\frac{2k\pi x}{L}\right)} \right]\\
-&\sim \frac{L}{2} + \sum_{k = 0}^{\infty}\left[\frac{-4L}{(2k+1)^{2}\pi^{2}}\cos{\left(\frac{(2k + 1)\pi x}{L}\right)} + \frac{-2L}{2k\pi}\sin{\left(\frac{2k\pi x}{L}\right)} \right]
+\noindent Now we have all the parts that we need. Our final answer will be
+the sum of all the parts:
+\begin{align*}
+f(x)&\sim \frac{L}{2} + \sum_{k = 0}^{\infty}\left[\frac{-4L}{(2k+1)^{2}\pi^{2}}
+\cos{\left(\frac{(2k + 1)\pi x}{L}\right)} \right] + \sum_{k =
+0}^{\infty}\left[\frac{-2L}{2k\pi}\sin{\left(\frac{2k\pi x}{L}\right)} \right]\\
+&\sim \frac{L}{2} + \sum_{k = 0}^{\infty}\left[\frac{-4L}{(2k+1)^{2}\pi^{2}}
+\cos{\left(\frac{(2k + 1)\pi x}{L}\right)} + \frac{-2L}{2k\pi}\sin{
+\left(\frac{2k\pi x}{L}\right)} \right]
 \end{align*}
@@
 \[f(x)=\left|\cos{x}\right|\qquad -\pi\leq x \leq \pi\]
-\indent\textbf{\textit{Solution:}} Before we dive into calculations here, let's think about the function at hand.
-We have the absolute value of $\cos{x}$.
-Usually, $\cos$ would be an odd function, which would imply that our $\cos$ series component to the whole Fourier Series would be $0$.
-However, the absolute value makes our function even, and as such the $\sin$ component of the series will be $0$.
-That's good news for us, because now we don't have to find $a_{n}$; we actually only have a Fourier Cosine Series here.
-Recall that we have a formula for $b_{n}$ and $b_{0}$, and that our general solution will have the form of
+\indent\textbf{\textit{Solution:}} Before we dive into calculations here, let's
+analyze the function at hand. We have the absolute value of $\cos{x}$. Usually,
+$\cos$ would be an odd function, which would imply that our $\cos$ series
+component to the whole Fourier Series would be $0$. However, the absolute value
+makes our function even, and as such the $\sin$ component of the series will be
+$0$. That's good news for us, because now we don't have to find $a_{n}$; we
+actually only have a Fourier Cosine Series here. Recall that we have a formula
+for $b_{n}$ and $b_{0}$, and that our general solution will have the form of
 \[\frac{b_{0}}{2} + \sum_{n=1}^{\infty}[b_{n}\cos{\left(\frac{n\pi x}{L}\right)}]\]
-\noindent Our function is bounded above by $\pi$, so we really have $L= \pi$. This gives us:
+\noindent Our function is bounded by $\pi$, so we really have $L= \pi$. This gives us:
 \[\frac{b_{0}}{2} + \sum_{n=1}^{\infty}[b_{n}\cos{\left(nx\right)}]\]
@@
 \end{align*}
-\noindent Let's take a second and discuss some the moves we made here.
-First of all, we know that since the function is even, we can say that the the integral of the entire range is really twice the integral of half the range.
-Since we are dealing with the cosine function, we actually have four times a quarter of the range.
-If we take a look at a graph, this may make a little more sense:\\
+\noindent We know that since the function is even, we can say that the the
+integral of the entire range is really twice the integral of half the range.
+Since we are dealing with the cosine function, we actually have four times a
+quarter of the range. If we take a look at a graph, this may make more sense:\\
 \begin{center}
@@
 \end{center}
-\noindent\\ You can see that the section of the graph from $[\frac{\pi}{2},\frac{3 \pi}{2}]$ is positive for the $|\cos{x}|$, but negative for $\cos{x}$.
-If we simply integrated the $|\cos{x}|$ over $[0,\pi]$, the result would be 0, but graphically this result would not make any sense.
-Now let's take a look at the $b_{n}$ term.
-Before we jump into the integration, let's think about this function:
+\noindent\\ You can see that the section of the graph from
+$[\frac{\pi}{2},\frac{3 \pi}{2}]$ is positive for the $|\cos{x}|$, but negative
+for $\cos{x}$. If we simply integrated the $|\cos{x}|$ over $[0,\pi]$, the
+result would be 0, but graphically this result would not make any sense. Now
+let's take a look at the $b_{n}$ term. Before we jump into the integration,
+let's think about this function:
 \[\left|\cos{x}\right| =
 \begin{cases*}
-\cos(x) \qquad if\ 0\leq x \leq \frac{\pi}{2}\\
--\cos(x)\quad if\ \frac{\pi}{2}\ \leq x \leq \pi
+\cos(x) \qquad 0\leq x \leq \frac{\pi}{2}\\
+-\cos(x)\quad  \frac{\pi}{2}\ \leq x \leq \pi
 \end{cases*}
@@
 \end{align*}
-\noindent Note that (as we have seen before) that the $\sin$ term disappears; however, the more interesting part is what happens with our $\cos$ term.
-The value of $\cos{\left(\frac{n\pi}{2}\right)}$ will be 0 at all odd $n$, or that is to say for all $n = 2k + 1$.
-So, we will let our $n = 2k$.
-This gives us:
+\noindent Note that (as we have seen before) the sine term disappears; however,
+the more interesting part is what happens with our cosine term. The value of
+$\cos{\left(\frac{n\pi}{2}\right)}$ will be 0 at all odd $n$, or for all
+$n = 2k + 1$. So, we will let our $n = 2k$. This gives us:
 \begin{align*}
@@
 \[
-f(x)\sim \frac{2}{\pi} - \sum_{k=0}^{\infty}\left[\frac{4}{\pi}\left[\frac{(-1)^{k}}{(2k)^{2}-1}\right] \cos{\left(\frac{2k\pi}{2}\right)}\right]
-\]
-\newpage
-
-\indent Let's recall the 2nd order PDE that we found a series solution for earlier. Now we want to apply the initial condition of $u(x,0) = x$.\\
-\noindent\textbf{\textit{Ex:}} Find the particular solution for the following PDE:
-\[
-\begin{cases}
-\frac{\partial u}{\partial t} = k \frac{\partial^{2} u}{\partial x^{2}}\\
-u(x,0) = x\\
-u(0,t) = u(L,t) = 0\\
-\end{cases}
-\]
-\indent\textbf{\textit{Solution:}} Earlier, we found a general solution of the form
-\begin{align*}
-u_{n}(x,t) &= a_{n}\sin{\left(\frac{n\pi x}{L}\right)}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}\\
-u(x,t) &= \sum_{n = 1}^{\infty}\left[a_{n}\sin{\left(\frac{n\pi x}{L}\right)}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}\right]
-\end{align*}
-\noindent Without loss of generality, let's let $L = \pi$.
-This would change our boundary condition to $u(\pi,t) = 0$, but this wouldn't change the solution that we have here in any way other than replacing the $L$'s in the equation with $\pi$.
-We will also let $k=1$.
-Again this doesn't change anything other than the fact that we will replace $k$ with 1.
-We now have:
-\[
-u(x,t) = \sum_{n = 1}^{\infty}\left[a_{n}\sin{\left(nx\right)}e^{-\left(n\right)^{2}t}\right]
-\]
-\noindent This is a much nicer sequence for us to deal with.
-Our goal now is to find $a_{n}$, which will give us a particular solution.
-We know that our answer for the particular solution will be in the form of a series, since our initial condition is not represented in the form of terms of a series (we will see an example of this later).
-Let's go ahead and apply our initial condition.
-Doing so will yield:
-\[
-x = \sum_{n = 0}^{\infty}\left[a_{n}\sin{nx}\right]
-\]
-\noindent Now, we need to come up with an $a_{n}$. We have a Fourier Sine Series here, so we can use the formula for finding $a_{n}$ that corresponds to our case.
-\begin{align*}
-a_{n} &= \frac{2}{\pi}\int_{0}^{\pi} x\sin{\left(nx\right)}dx\\
-&= -\frac{2n\pi \cos{\left(n\pi\right)} + 2\sin{\left(n\pi\right)}}{n^{2}\pi}\\
-&= -\frac{2(-1)^{n}}{n}
-\end{align*}
-\noindent So our particular solution is:
-\[
-u(x,t) = \sum_{n = 1}^{\infty}\left[-\frac{2(-1)^{n}}{n}\sin{\left(n x\right)}e^{-\left(n\right)^{2}t}\right]
-\]
+f(x)\sim \frac{2}{\pi} - \sum_{k=0}^{\infty}\left[\frac{4}{\pi}\left[\frac{(-1)^{k}}{(2k)^{2}-1}\right]
+\cos{\left(\frac{2k\pi}{2}\right)}\right]
+\]
+And we are done.
+\newpage
+
+\subsection{Convergence of Fourier Series}
+\indent Up to this point, we have found a Fourier series and left our solutions
+at that. We will now look into the convergence of a Fourier series. This topic
+includes two questions:
+\begin{enumerate}
+\item Does the series converge?
+\item If the series converges, to what value does it converge?
+\end{enumerate}
+\noindent Before we jump into convergence, we need to review some ideas.
+\begin{enumerate}
+\item \textit{\textbf{Jump Discontinuity:}} To say that $f(x)$ has a jump
+discontinuity at $x = a$ is to say that
+\[
+\lim_{x \to a^{-}}f(x) \neq \lim_{x \to a^{+}}f(x)
+\]
+Where both the limit from the left and the limit from the right exist.
+\item \textit{\textbf{Piecewise Smooth:}} A function $f(x)$ is said to be
+piecewise smooth if the function can be broken into two distinct intervals, and
+on each interval both $f(x)$ and $f^{'}(x)$ are continuous.
+\item \textit{\textbf{Periodic Extension:}} The periodic extension of $f(x)$ on
+$[-L,L]$ is the repetition of the function on intervals on periods to the left
+and right of $[-L,L]$.
+\end{enumerate}
+\indent With these definitions, we can lay out a general rule for convergence of
+a Fourier series. We won't give a proof of these rules for the time being. We
+just need them to lay out some groundwork for the rest of the information we
+will present on Fourier series. We have the following:
+\begin{quote}
+Suppose $f(x)$ is a function, and that $f(x)$ is piecewise smooth on the interval
+$[-L,L]$. The corresponding Fourier series for $f(x)$ converges to either
+\begin{enumerate}
+\item the periodic extension of $f(x)$, if $f(x)$ is continuous, or
+\item the average of the two one-sided limits, if the periodic extension of
+$f(x)$ has a jump discontinuity at $x = a$.
+\end{enumerate}
+\end{quote}
+\noindent We should note that the second condition can be expressed by the formula
+\[\frac{1}{2}\left[\lim_{x \to a^{-}}f(x) + \lim_{x \to a^{+}}f(x)\right]\]
+which gives us a nice way to calculate the value to which the series converges
+if we happen to meet the criteria in (2). Let's take a look at some examples.\\\\
+
+\noindent\textbf{\textit{Ex:}} Find the Fourier series for the following
+equation, and determine if the series converges. If it does, tell where the
+series will converge.
+\[ f(x) =
+\begin{cases}
+L \qquad\qquad -L \leq x \leq 0\\
+2x \qquad\qquad 0 \leq x \leq L
+\end{cases}
+\]
+\indent\textbf{\textit{Solution:}} We will go ahead and skip the details of
+finding the Fourier series for this function. If you do want to find it, you
+will have to find a full piecewise Fourier series. For this particular $f(x)$,
+we have:
+\begin{align*}
+f(x) &\sim L + \sum_{n=1}^{\infty}\left[\frac{2L}{n^{2}\pi^{2}}([-1]^{n}-1)\cos{\left(\frac{n\pi
+x}{L}\right)}\right] - \sum_{n=1}^{\infty}\left[\frac{L}{n\pi}(1 + [-1]^{n})\sin{\left(\frac{n\pi
+x}{L}\right)}\right]
+\end{align*}
+\noindent Notice that we didn't simplify this solution to a nicer form.
+If we evaluate this series, we will run across terms where the series is 0, but
+that is okay for now. This function is also defined as a piecewise function,
+and on both intervals ($[-L,0]$ and $[0,L]$), $f(x)$ is continuous. In other
+words, $f(x)$ is piecewise smooth, so we know that this series converges. Now
+let's take a look at the graph of $f(x)$ on an arbitrary interval of $[-1,1]$.
+\begin{center}
+\begin{tabular}{c}
+\includegraphics[scale=0.6]{pw_graph_01}\\
+$
+f(x) =
+\begin{cases}
+1 \qquad\qquad -1 \leq x \leq 0\\
+2x \qquad\qquad 0 \leq x \leq 1
+\end{cases}
+$
+\end{tabular}
+\end{center}
+\noindent We can see that there is a jump in the graph at $x = 0$. In other
+words, there is a jump discontinuity at $x = 0$, and that means that we need to
+use (2) in order to find the value of convergence. We will let $\sum$ stand in
+for our entire series. We have:
+\begin{align*}
+\sum &\to \frac{1}{2}\left[\lim_{x\to 0}L + \lim_{x\to0}2x\right]\\
+&\to \frac{1}{2}[L + 0]\\
+&\to \frac{L}{2}
+\end{align*}
+\noindent So, at $x=0$, the series converges to $\frac{L}{2}$. We are not done
+yet, however. Now, we need to consider the periodic extension of the series. The
+function will repeat every $[-L,L]$, so we need to look at what the convergence
+of the series looks like at $-L$ and $L$. Let's take a look:
+\begin{alignat*}{3}
+\sum&\to\frac{1}{2}\left[\lim_{x\to-L^{-}}f(x) + \lim_{x\to-L^{+}}f(x)\right] \qquad\qquad \sum&&\to
+\frac{1}{2}\left[\lim_{x\to L^{-}}f(x) + \lim_{x\to L^{+}}f(x)\right]\\
+&\to \frac{1}{2}[2L + L] &&\to \frac{1}{2}[2L + L]\\
+&\to\frac{3L}{2} &&\to\frac{3L}{2}
+\end{alignat*}
 \noindent And we are done.
-
-\newpage
-\noindent\textbf{\textit{Ex:}} Solve the following boundary value problem given the general solution:
+\newpage
+
+\section{Second Order Partial Differential Equations}
+\hrule
+\noindent\\
+\subsection{The Heat and Wave Equations}
+\indent Let's take a look at the second order PDE. We will start with a
+homogeneous equation, which we will tackle with the separation of variables
+method. The general strategy is that we want to take our PDE and convert it to
+an ODE, which we know how to solve. The problems presented here also involve
+boundary conditions, which will require us to solve for eigenvalues/functions.
+We will also be dealing with series (specifically Fourier Series) to help us
+solve these problems. Let's look at an example:\\
+
+\newpage
+\noindent\textbf{\textit{Ex:}} Solve the following boundary value problem:
 \[
@@
 \[f(x) = 2\sin{(\pi x)} + 3\sin{(2\pi x)}\qquad g(x) = 4\sin{(3\pi x)} - 7\sin{(5\pi x)}\]
-\[u(x,t) = \sum_{n=1}^{\infty}\left[\left(a_{n}\cos{[n\pi t]} + b_{n}\sin{[n\pi t]}\right)\sin{[n\pi x]}\right]\]
-\indent\textbf{\textit{Solution:}} There's a lot going on here, so let's break it down.
-First of all, notice that we have two initial conditions, one of which is a derivative.
-This should make sense since the equation have two second order partial derivatives included in it.
-Next notice that we have a Fourier Sine Series, despite both the $\sin{[n\pi t]}$ and $\cos{[n\pi t]}$ terms.
-If we had worked out the entire problem, we would see that our eigenvalue/function for this equation takes the form of
+given the general solution:
+\[u(x,t) = \sum_{n=1}^{\infty}\left[\left(a_{n}\cos{[n\pi t]} + b_{n}\sin{[n\pi t]}\right)\sin{[n\pi
+x]}\right]\]
+\indent\textbf{\textit{Solution:}} There's a lot going on here, so let's break
+it down. First of all, notice that we have two initial conditions, one of which
+is a derivative. This should make sense since the equation have two second order
+partial derivatives included in it. Next notice that we have a Fourier Sine Series,
+despite both the $\sin{[n\pi t]}$ and $\cos{[n\pi t]}$ terms. If we had worked
+out the entire problem, we would see that our eigenvalue/function for this
+equation takes the form of
 \[
@@
 \noindent which tells us that the corresponding series is a Fourier Sine Series.
-We also see the initial conditions on our equation are already representative of a series of sine functions, which is not just a conincidence.
-To apply the initial conditions, we will need both the general solution and the derivative with respect to $t$ of the general solution.
-Let's go ahead and apply the first initial condition, $u(x,0) = f(x)$.
-Letting $t= 0$ gives us:
+We also see the initial conditions on our equation are already representative
+of a series of sine functions, which is not just a conincidence. To apply the
+initial conditions, we will need both the general solution and the derivative
+with respect to $t$ of the general solution. Let's go ahead and apply the first
+initial condition, $u(x,0) = f(x)$. Letting $t= 0$ gives us:
 \[
@@
 \]
-\noindent We see that the term with $b_{n}$ has dropped out, since $\sin{(0)} = 0$.
-Now we just have to match up the coefficients on $a_{n}$.
-We can see that when we have $n=1$, $a_{n} = 2$, and when $n=2$, $a_{n} = 3$.
-For all other $n$, we have $a_{n} = 0$, so we ignore those terms.
-Next, we can apply the initial condition for the derivative of the general solution with respect to $t$.
-We expect that all the terms related to $a_{n}$ will disappear, and in fact they will.
-Let's take a look:
-\begin{align*}
-u_{t}(x,t) &= \sum_{n=1}^{\infty}\left[(-a_{n}n\pi\sin{[n\pi t]} + b_{n}n\pi\cos{[n\pi t]})\sin{n\pi x}\right]\\
-u_{t}(x,0) &= 4\sin{(3\pi x)} - 7\sin{(5\pi x)} = \sum_{n=1}^{\infty}\left[n\pi b_{n}\sin{(n\pi x)}\right]
-\end{align*}
-\noindent We are going to do the same thing here that we did to find our $a_{n}$ terms, but notice that we don't just have $b_{n}$; we actually have $n\pi b_{n}$. So we have the following:
+\noindent We see that the term with $b_{n}$ has dropped out, since
+$\sin{(0)} = 0$. Now we just have to match up the coefficients on $a_{n}$. We
+can see that when we have $n=1$, $a_{n} = 2$, and when $n=2$, $a_{n} = 3$. For
+all other $n$, we have $a_{n} = 0$, so we ignore those terms. Next, we can apply
+the initial condition for the derivative of the general solution with respect to
+$t$. We expect that all the terms related to $a_{n}$ will disappear, and in fact
+they will. Let's take a look:
+\begin{align*}
+u_{t}(x,t) &= \sum_{n=1}^{\infty}\left[(-a_{n}n\pi\sin{[n\pi t]} + b_{n}n\pi\cos{[n\pi t]})\sin{n\pi
+x}\right]\\
+u_{t}(x,0) &= 4\sin{(3\pi x)} - 7\sin{(5\pi x)} = \sum_{n=1}^{\infty}\left[n\pi b_{n}\sin{(n\pi
+x)}\right]
+\end{align*}
+\noindent We are going to do the same thing here that we did to find our $a_{n}$
+terms, but notice that we don't just have $b_{n}$; we actually have $n\pi b_{n}$.
+We have the following:
 \begin{alignat*}{3}
@@
 \noindent Much like the solutions for $a_{n}$, all the other $b_{n}$ terms are 0.
-Now we can construct a particular solution.
-In this case, the answer will not be in the form of a series, since we have definite terms.
-Our solution is as follows:
-\[
-u(x,t) = 2\cos{(\pi t)}\sin{(\pi x)} + 3\cos{(2\pi t)}\sin{(2\pi x)} + \frac{4}{3\pi}\sin{(3\pi t)}\sin{(3\pi x)} - \frac{7}{5\pi}\sin{(5\pi t)}\sin{(5\pi x)}
+Now we can construct a particular solution. In this case, the answer will not be
+in the form of a series, since we have definite terms. Our solution is as follows:
+\[
+u(x,t) = 2\cos{(\pi t)}\sin{(\pi x)} + 3\cos{(2\pi t)}\sin{(2\pi x)} + \frac{4}{3\pi}\sin{(3\pi
+t)}\sin{(3\pi x)} - \frac{7}{5\pi}\sin{(5\pi t)}\sin{(5\pi x)}
 \]
@@
 
-\subsection{Convergence of Fourier Series}
-\indent Up to this point, we have found a Fourier series and left our solutions at that. We will now look into the convergence of a Fourier series. This topic includes two questions:
-\begin{enumerate}
-\item Does the series converge?
-\item If the series converges, to what value does it converge?
-\end{enumerate}
-\noindent Both of these questions are highly theoretical. We will discuss the answers to them, but first we need to go over some terms:
-\begin{enumerate}
-\item \textit{\textbf{Jump Discontinuity:}} To say that $f(x)$ has a jump discontinuity at $x = a$ is to say that
-\[
-\lim_{x \to a^{-}}f(x) \neq \lim_{x \to a^{+}}f(x)
-\]
-Where both the limit from the left and the limit from the right exist.
-\item \textit{\textbf{Piecewise Smooth:}} A function $f(x)$ is said to be piecewise smooth if the function can be broken into two distinct intervals, and on each interval both $f(x)$ and $f^{'}(x)$ are continuous.
-\item \textit{\textbf{Periodic Extension:}} The periodic extension of $f(x)$ on $[-L,L]$ is the repetition of the function on intervals on periods to the left and right of $[-L,L]$.
-\end{enumerate}
-\indent With these definitions, we can lay out a general rule for convergence of a Fourier series.
-We won't give a proof of these rules for the time being.
-We just need them to lay out some groundwork for the rest of the information we will present on Fourier series.
-We have the following:
-\begin{quote}
-Suppose $f(x)$ is a function, and that $f(x)$ is piecewise smooth on the interval $[-L,L]$. The corresponding Fourier series for $f(x)$ converges to either
-\begin{enumerate}
-\item the periodic extension of $f(x)$, if $f(x)$ is continuous, or
-\item the average of the two one-sided limits, if the periodic extension of $f(x)$ has a jump discontinuity at $x = a$.
-\end{enumerate}
-\end{quote}
-\noindent We should note that the second condition can be expressed by the formula
-\[\frac{1}{2}\left[\lim_{x \to a^{-}}f(x) + \lim_{x \to a^{+}}f(x)\right]\]
-which gives us a nice way to calculate the value to which the series converges if we happen to meet the criteria in (2). Let's take a look at some examples.\\\\
-
-\noindent\textbf{\textit{Ex:}} Find the Fourier series for the following equation, and determine if the series converges. If it does, tell where the series converges to.
-\[ f(x) =
-\begin{cases}
-L \qquad\qquad -L \leq x \leq 0\\
-2x \qquad\qquad 0 \leq x \leq L
-\end{cases}
-\]
-\indent\textbf{\textit{Solution:}} We will go ahead and skip the details of finding the Fourier series for this function.
-If you do want to find it, you will have to find a full piecewise Fourier series.
-For this particular $f(x)$, we have:
-\begin{align*}
-f(x) &\sim L + \sum_{n=1}^{\infty}\left[\frac{2L}{n^{2}\pi^{2}}([-1]^{n}-1)\cos{\left(\frac{n\pi x}{L}\right)}\right] - \sum_{n=1}^{\infty}\left[\frac{L}{n\pi}(1 + [-1]^{n})\sin{\left(\frac{n\pi x}{L}\right)}\right]
-\end{align*}
-\noindent You'll notice that we didn't simplify this solution to a nicer form.
-If we evaluate this series, we will run across terms where the series is 0, but that is okay for now.
-You should also notice that this function is defined as a piecewise function, and on both intervals ($[-L,0]$ and $[0,L]$), $f(x)$ is continuous.
-In other words, $f(x)$ is piecewise smooth, so we know that this series converges.
-Now let's take a look at the graph of $f(x)$ on an arbitrary interval of $[-1,1]$.
-\begin{center}
-\begin{tabular}{c}
-\includegraphics[scale=0.6]{pw_graph_01}\\
-$
-f(x) =
-\begin{cases}
-1 \qquad\qquad -1 \leq x \leq 0\\
-2x \qquad\qquad 0 \leq x \leq 1
-\end{cases}
-$
-\end{tabular}
-\end{center}
-\noindent We can see that there is a jump in the graph at $x = 0$.
-In other words, there is a jump discontinuity at $x = 0$, and that means that we need to use (2) in order to find the value of convergence.
-We will let $\sum$ stand in for our entire series.
-We have:
-\begin{align*}
-\sum &\to \frac{1}{2}\left[\lim_{x\to 0}L + \lim_{x\to0}2x\right]\\
-&\to \frac{1}{2}[L + 0]\\
-&\to \frac{L}{2}
-\end{align*}
-\noindent So, at $x=0$, the series converges to $\frac{L}{2}$.
-We are not done yet, however.
-Now, we need to consider the periodic extension of the series.
-The function will repeat every $[-L,L]$, so we need to look at what the convergence of the series looks like at $-L$ and $L$.
-Let's take a look:
-\begin{alignat*}{3}
-\sum&\to\frac{1}{2}\left[\lim_{x\to-L^{-}}f(x) + \lim_{x\to-L^{+}}f(x)\right] \qquad\qquad \sum&&\to \frac{1}{2}\left[\lim_{x\to L^{-}}f(x) + \lim_{x\to L^{+}}f(x)\right]\\
-&\to \frac{1}{2}[2L + L] &&\to \frac{1}{2}[2L + L]\\
-&\to\frac{3L}{2} &&\to\frac{3L}{2}
-\end{alignat*}
-\noindent And we are done.
-\newpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Second Order Partial Differential Equations}
-\hrule
-\noindent\\
-\subsection{The Heat and Wave Equations}
-\indent Let's take a look at the second order PDE.
-We will start with a homogeneous equation, which we will tackle with the separation of variables method.
-The general strategy is that we want to take our PDE and convert it to an ODE, which we know how to solve.
-The problems presented here also involve boundary conditions, which will require us to solve for eigenvalues/functions.
-We will also be dealing with series (specifically Fourier Series) to help us solve these problems.
-Let's start by looking at the homogeneous heat equation.\\
 \noindent \textbf{\textit{Ex:}} Solve the following PDE
@@
 \]
-\indent \textbf{\textit{Solution:}} We will use the separation of variable technique to solve this problem.
-We know to use separation of variable because we are dealing with a homogeneous PDE with homogenous boundary conditions.
-Separation of variables tells us that our solution will be in the form of $ u(x,t) = X(x)T(t) $.
-If we take this fact at face value, we can use the separation of variables method to solve this.
-First, we substitute in our solution into the PDE:
+\indent \textbf{\textit{Solution:}} We will use the separation of variable
+technique to solve this problem. We know to use separation of variable because
+we are dealing with a homogeneous PDE with homogenous boundary conditions.
+Separation of variables tells us that our solution will be in the form of
+$u(x,t) = X(x)T(t)$. If we take this fact at face value, we can use the
+separation of variables method to solve this. First, we substitute in our solution
+into the PDE:
 \begin{align*}
@@
 \end{align*}
-\noindent The only way that our function of x and our function of t can be equal is if they are equal to a constant;
-thus we now have:
+\noindent The only way that our function of x and our function of t can be equal
+is if they are equal to a constant; thus we now have:
 \[
@@
 \]
-\noindent And now we have a first order ODE for the time and a second order ODE for the position:
+\noindent And now we have a first order ODE for the time and a second order ODE
+for the position:
 \[
@@
 \]
-\noindent Our next step is to find the Eigenvalues and Eigenfunctions of the second order ODE; we will come back to the time ODE when we have values for $\lambda$.\\ 
-\indent To solve for these values, we have to find non-trivial eigenvalues that correspond to the boundary values of the ODE.
-In this case, we know that $X(0) = 0$ and $X(L) = 0$.
-These boundary conditions come from the BC on the PDE which states $u(0,t) = u(L,t) = 0$.
-To find the eigenvalues/functions, we need to find the characteristic equation corresponding to the ODE in question.
-Let's go ahead and find it:
+\noindent Our next step is to find the Eigenvalues and Eigenfunctions of the
+second order ODE; we will come back to the time ODE when we have values for
+$\lambda$.\\
+\indent To solve for these values, we have to find non-trivial eigenvalues that
+correspond to the boundary values of the ODE. In this case, we know that
+$X(0) = 0$ and $X(L) = 0$. These boundary conditions come from the BC on the PDE
+which states $u(0,t) = u(L,t) = 0$. To find the eigenvalues/functions, we need
+to find the characteristic equation corresponding to the ODE in question. Let's
+go ahead and find it:
 \begin{gather*}
@@
 \end{gather*}
-\indent Now, we can substitute values of $\lambda$ to find out solutions.
-Let's start with $\lambda > 0$.
-Since we have a $-\lambda$ under the radical, we will have $r = \pm \sqrt{\lambda} \mathrm{i}$.
-This will yield the solution of
+\indent Now, we can substitute values of $\lambda$ to find out solutions. Let's
+start with $\lambda > 0$. Since we have a $-\lambda$ under the radical, we will
+have $r = \pm \sqrt{\lambda} \mathrm{i}$. This will yield the solution of
 \[
@@
 \]
-\noindent Now we know that $c_{1} = 0$.
-This result came from the fact that $\sin{0} = 0$, so the $c_{2}$ term disappears.
-Now we are left with $\cos{0} = 1$, which tells us that $c_{1} = 0$.
-We can apply this to our second boundary condition like so:
+\noindent Now we know that $c_{1} = 0$. This result came from the fact that
+$\sin{0} = 0$, so the $c_{2}$ term disappears. Now we are left with $\cos{0} = 1$,
+which tells us that $c_{1} = 0$. We can apply this to our second boundary
+condition like so:
 \begin{gather*}
@@
 \end{gather*}
-\noindent It looks like there will only be trivial solutions, but we know this isn't true due to the Jacobian matrix.
-Recall that if the determinant of the Jacobian $\neq 0$, then we only have trivial solutions, and it the determinant of the Jacobian $= 0$, then we do not have the trivial solution.
-Let's take a look at the Jacobian for this case:
+\noindent It looks like there will only be trivial solutions, but we know this
+isn't true due to the Jacobian matrix. Recall that if the determinant of the
+Jacobian $\neq 0$, then we only have trivial solutions, and if the determinant
+of the Jacobian $= 0$, then we do not have the trivial solution. Let's take a
+look at the Jacobian for this case:
 \begin{gather*}
@@
 \end{gather*}
-\noindent Notice that we replaced the $\arcsin{0}$ term with $n\pi$.
-Recall from trig that the $\sin$ function is $0$ at every $n\pi$, starting with $n = 0$.
-Next, we just need to find our corresponding eigenfunction.
-Our $c_{1}$ was $0$, so we can ignore the entire $\cos$ term.
-Thus, we have an eigenfunction from plugging in our $\lambda$ that looks like:
+\noindent Notice that we replaced the $\arcsin{0}$ term with $n\pi$. Recall from
+trig that the $\sin$ function is $0$ at every $n\pi$, starting with $n = 0$.
+Next, we just need to find our corresponding eigenfunction. Our $c_{1}$ was $0$,
+so we can ignore the entire $\cos$ term. Thus, we have an eigenfunction from
+plugging in our $\lambda$ that looks like:
 \[
@@
 \]
-\indent Unfortunately, we aren't done yet.
-Now we have to find the corresponding eigenvalues/functions for $\lambda = 0$.
-This isn't too hard, and can be done in a few lines if we recall the characteristic equation.
-Plugging in $0$ for $\lambda$ gives us that $r = 0$.
-Using this piece of information, we now know that $y(x) = c_{1} + xc_{2}$.
-Applying our boundary conditions gives us $c_{1} = 0, c_{2} = 0$.
-I'll leave you to do the math (try pluggin into the Jacobian; you'll see that there is no way to force the determinant to be 0) , but we can see that we have only trivial solutions for $\lambda = 0$.\\
-\indent Our final case is for $\lambda < 0$.
-I'll go ahead and tell you that this will yield another trivial solution, but we can take a look at the work behind it.
-Since $\lambda > 0$, we know that our $r$ from the characteristic equation will be $r = \pm \sqrt{\lambda}$.
-From here we are going to do something a little unconventional that will help us.
-Usually, we would have a solution of the form $y(x) = c_{1}e^{\sqrt{\lambda}x} + c_{2}x e^{-\sqrt{\lambda}x}$.
-In our case, it will be more convenient to use hyperbolic trig functions.
-It's not too far off from our solution from $\lambda > 0$:
+\indent Unfortunately, we aren't done yet. Now we have to find the corresponding
+eigenvalues/functions for $\lambda = 0$. This isn't too hard, and can be done in
+a few lines if we recall the characteristic equation. Plugging in $0$ for
+$\lambda$ gives us that $r = 0$. Using this piece of information, we now know
+that $y(x) = c_{1} + xc_{2}$. Applying our boundary conditions gives us
+$c_{1} = 0, c_{2} = 0$. We can see that we have only trivial solutions for
+$\lambda = 0$.\\
+\indent Our final case is for $\lambda < 0$. I'll go ahead and tell you that
+this will yield another trivial solution, but we can take a look at the work
+behind it. Since $\lambda > 0$, we know that our $r$ from the characteristic
+equation will be $r = \pm \sqrt{\lambda}$. From here we are going to do something
+ a little unconventional that will help us. Usually, we would have a solution of
+the form $y(x) = c_{1}e^{\sqrt{\lambda}x} + c_{2}x e^{-\sqrt{\lambda}x}$. In our
+case, it will be more convenient to use hyperbolic trig functions. It's not too
+far off from our solution from $\lambda > 0$:
 \begin{align*}
@@
 \end{align*}
-\noindent I'll leave you to apply the boundary conditions (remember that $\sinh{0} = 0$ and $\cosh{0} = 1$), but you will find that you have only trivial solutions.\\
-\indent So, we are done with the second order ODE, and now we turn our eye to the first order time equation. We know our $\lambda$, so it isn't too hard to solve:
+\noindent If we apply the boundary conditions (remember that
+$\sinh{0} = 0$ and $\cosh{0} = 1$), we will find that we have only trivial
+solutions.\\
+\indent So, we are done with the second order ODE, and now we turn our eye to
+the first order time equation. We know our $\lambda$, so it isn't too hard to solve:
 \begin{gather*}
@@
 \noindent We can analyze this solution and tell where each piece comes from.
-The $B_{n}$ is a coefficient that results from our $c's$ in the general solutions of the differential equations.
-The $sin$ part results from out eigen vector.
-Remember, we get a different value for each $n$, so there may or may not be $0$ terms in the series.
-Finally, we have the exponential term.
-This term comes from t he solution to our time equation.
-When we put it all together, we can see that this is, of course a series; for every $n$, we get a different solution.
-From here we can find the Fourier Series (which will be a Fourier Sine Series).
-We won't worry about that yet, but we will come back to it.
-\newpage
-
-\indent It will benefit us to take a moment and discuss how boundary conditions effect our Fourier series.
-We have three different combinations of boundary conditions, where each one will produce a different result in our series solution.
+The $B_{n}$ is a coefficient that results from our $c's$ in the general solutions
+of the differential equations. The $sin$ part results from out eigen vector.
+Remember, we get a different value for each $n$, so there may or may not be $0$
+terms in the series. Finally, we have the exponential term. This term comes from
+the solution to our time equation. When we put it all together, we can see that
+this is, of course a series; for every $n$, we get a different solution. From
+here we can find the Fourier Series (which will be a Fourier Sine Series). We
+won't worry about that yet, but we will come back to it.
+\newpage
+
+\indent It will benefit us to take a moment and discuss how boundary conditions
+effect our Fourier series. We have three different combinations of boundary
+conditions, where each one will produce a different result in our series solution.
 \begin{enumerate}
 \item \textit{Dirichlet Boundary Conditions}\\
-Dirichlet boundary conditions involve the function for which we are looking.
-In other words, the boundary conditions are presented in terms of the functions we are trying to find, rather than the derivative of that function.
-Typically, this type of boundary condition will take the form of (in the homogeneous case)
+Dirichlet boundary conditions involve the function for which we are looking. In
+other words, the boundary conditions are presented in terms of the functions we
+are trying to find, rather than the derivative of that function. Typically, this
+type of boundary condition will take the form of (in the homogeneous case)
 \[u(0,t) = u(L,t) = 0.\]
-In regards to our Fourier series, we expect to see $\sin{(\frac{n\pi x}{L})}$. Our basis will look like
+In regards to our Fourier series, we expect to see $\sin{(\frac{n\pi x}{L})}$.
+Our basis will look like
 \[
@@
 \item \textit{Neumann Boundary Conditions}\\
-Neumann boundary conditions involve the derivatives of the function for which we are solving. Neumann boundary conditions will look like
+Neumann boundary conditions involve the derivatives of the function for which we
+are solving. Neumann boundary conditions will look like
 \[u_{x}(0,t) = u_{x}(L,t) = 0.\]
-This will cause our Fourier series to contain $\cos{(\frac{n\pi x}{L})}$. Out basis will look like
+This will cause our Fourier series to contain $\cos{(\frac{n\pi x}{L})}$. Our
+basis will look like
 \[
@@
 \item \textit{Robin Boundary Conditions}\\
-Robin boundary conditions will contain a mixture of the function and it's derivative. Typically, we will see
+Robin boundary conditions will contain a mixture of the function and it's
+derivative. Typically, we will see
 \[ u_{x}(0,t) = u(L,t) = 0\qquad \text{or}\qquad u(0,t) = u_{x}(L,0) = 0.\]
-These boundary conditions will give us two different results, depending on whether we have the derivative first or second. We will have
+These boundary conditions will give us two different results, depending on whether
+we have the derivative first or second. We will have
 \begin{align*}
@@
 
-\indent Now we can take a look at a slightly harder PDE.
-We still have the heat equation, but we will not have homogeneous boundary conditions.
-Our goal here is to find a way to transform our boundary conditions into a homogenous form and take it from there.\\\\
+\indent Now we can take a look at a slightly harder PDE. We still have the heat
+equation, but we will not have homogeneous boundary conditions. Our goal here is
+to find a way to transform our boundary conditions into a homogenous form and take
+it from there.\\\\
 \noindent \textbf{\textit{Ex:}} Solve the following PDE
@@
 \]
-\indent \textbf{\textit{Solution:}} We would like to just jump into this PDE and solve it, but it's not quite that simple.
-We have to find a way to express our boundary conditions in a homogeneous manner.
-We can do this through a clever substitution; we will create a new PDE, denoted $\Upsilon$.
-To find $\Upsilon$, we use the following formula:
+\indent \textbf{\textit{Solution:}} We would like to just jump into this PDE and
+solve it, but it's not quite that simple. We have to find a way to express our
+boundary conditions in a homogeneous manner. We can do this through a clever
+substitution; we will create a new PDE, denoted $\Upsilon$. To find $\Upsilon$,
+we use the following formula:
 \[ \Upsilon = B_{1} + \frac{x}{l}\Delta B, \]

### 2015-12-22 pde/pdetest1.tex
update pde

@@
 \usepackage{array}
+\usepackage{mathrsfs}
+\usepackage{minted}
+\usepackage{graphics}
+\usepackage{setspace}
 \topmargin=-0.3in
@@
 
+\newcommand{\Lap}{\mathscr{L}}
+\newcommand{\Four}{\mathscr{F}}
+
+\newcounter{mathematicapage}
+
 \title{Partial Differential Equations}
@@
 \newpage
-\section{First Order Partial Differential Equations}
-\hrule
-\noindent\\\\
-\noindent\textbf{\textit{Ex:}} Solve the following PDE
-\[
-u_{x} + u_{y} + u = x + y
-\]
-\indent \textbf{\textit{Solution:}} In this case, we have a non-zero term on the
-right hand side of the equation. We will use substitution to solve this. For the
-time being, we will let $T$ be $x + y$. To find our $S$, we can use a similar
-method to the one that we used above. We know that the coefficients on the
-$u_{x}$ and $u_{y}$ are $1$ in this case. Now we have:
-\begin{align*}
-\frac{dy}{dx} &= 1\\
-dy &= dx\\
-\int dy &= \int dx\\
-x &= y + c\\
-x - y &= c
-\end{align*}
-\noindent Now we can say that $S = x - y$; however, we need to check our choice
-for $T$ to make sure this substitution will not just yield a trivial solution.
-We can do this by Next we have to actually use the substitution. We need to
-determine whether the coefficient matrix, or the Jacobian, of our substitutions
-has a determinant of zero. If it does, then we need to choose a different $T$.
-\[
-\left|
-\begin{array}{c c}
-1 & 1\\
-1 & -1
-\end{array}
-\right| = 2 \neq 0.
-\]
-Our choices of variables will work here, so now we need to express $u_{x}$ and
-$u_{y}$ in terms of $S$ and $T$. The following equations come from the chain
-rule; when the $S$ and $T$ terms are differentiated, we get $u_{S}$ and $u_{T}$
-left behind.
-\begin{alignat*}{3}
-u_{x} &= u_{S}\frac{\partial S}{\partial x} + u_{T}\frac{\partial T}{\partial x}
-\qquad\qquad &&u_{y} = u_{S}\frac{\partial S}{\partial y} + u_{T}\frac{\partial T}{\partial y}\\
-u_{x} &= -u_{S} + u_{T}  \qquad\qquad &&u_{y} = u_{S} + u_{T}
-\end{alignat*}
-\noindent We can now plug in our values into the original equation. Note that we
-didn't do anything with $u$; it gets left alone. In addition, if we needed
-$u_{xx}$ or $u_{yy}$, we can just differentiate again. Plugging in yields:
-\begin{gather*}
-(-u_{S} + u_{T}) + (u_{S} + u_{t}) + u = T\\
-2u_{T} + u = T
-\end{gather*}
-\noindent This is just a first order ODE. Since it's not homogeneous, we need to
-solve for the general solution, and then the particular solution. Let's start
-with the general.
-\begin{gather*}
-2y^{'} + y = 0\\
-2r + 1 = 0\\
-r = -\frac{1}{2}\\
-y = ce^{\left(-\frac{1}{2}\right)x}
-\end{gather*}
-\noindent In this case, we know that our $x$ is really $T$. So we finally have
-$y = ce^{\left(-\frac{1}{2}\right)T}$. We will back substitute soon, but let's
-find the particular solution before we do that.
-\noindent Recall from ODE that we have several cases for the particular
-solution. In our case, we have a linear equation on the right hand side of our
-ODE. This lets us say:
-\begin{gather*}
-y = AT + B\\
-y^{'} = A\\
-\end{gather*}
-\noindent Now we can substitute again into our non-homogeneous ODE as follows:
-\[
-2y^{'} + y = T \equiv 2A + AT + B = T
-\]
-\noindent And we have that $A = 1$, and $B = -2$. Now we have a final solution of:
-\begin{gather*}
-y = ce^{\left(-\frac{1}{2}\right)T} + T - 2\\
-y = ce^{\left(-\frac{1}{2}\right)(x+y)} + (x + y) - 2
-\end{gather*}
-\noindent We aren't quite done yet, as we have a solution to an ODE, not a PDE.
-It isn't too difficult to switch back; The $c$ is really the only difference. In
-our ODE, the $c$ is a constant term, but to our PDE, the $c$ is a function. The
-$c$ now becomes $f(s)$. We will go ahead and back substitute, which will give us
-$f(y-x)$. Now we have our final solution:
-\[
-u(x,y) = f(y-x)e^{\left(-\frac{1}{2}\right)(x+y)} + (x + y) - 2
-\]
-\noindent And we are done.
-
+\include{first_order}
 \newpage
-\section{Classification of Partial Differential Equations}
-\hrule
-\noindent\\\\
-\indent We mentioned earlier the topic of classification of PDE's. Since we only
-have this and Fourier Series to talk about, we will go ahead and knock out
-classification. There's some information that we need to go ahead and review
-here. This will be explained in the problem, but it's handy to have it for
-reference:
-\begin{align*}
-b^{2} - ac < 0 &\implies \text{Elliptic with complex roots}\\
-b^{2} - ac = 0 &\implies \text{Parabolic with the same root twice}\\
-b^{2} - ac > 0 &\implies \text{Hyperbolic with two distinct real roots}
-\end{align*}
-\noindent Another helpful piece of information is what's known as the
-\textit{classical trinity} of PDEs.
-\begin{align*}
-u_{t} - ku_{xx} &= 0 \qquad \text{The heat equation (parabolic)}\\
-u_{tt} - c^{2}u_{xx} &= 0 \qquad \text{The wave equation (hyperbolic)}\\
-u_{xx} + u_{yy} &= 0 \qquad \text{The Laplace equation (elliptic)}
-\end{align*}
-Let's go ahead and look at an example.\\\\
-\indent \textbf{Ex. }Classify the following PDE:
-\[yu_{xx} + u_{yy} = 0\]
-\indent \textbf{\textit{Solution:}} The first thing we need to identify are the
-coefficients on the $u_{xx}$, $u_{yy}$, and $u_{xy}$ or $u_{yx}$ terms. Note
-that the $u_{xy}$ and $u_{yx}$ are interchangeable. We will assign these
-coefficients to $a$, $b$, and $c$ corresponding to $u_{xx}$, $u_{xy}$, and
-$u_{yy}$, respectively. Looking at the PDE, we can see that
-\[
-a = y \qquad b = 0 \qquad c = 1
-\]
-\noindent Now we look at $b^{2} - ac$. We are really dealing with
-$b = \frac{b_{0}}{2}$. In other words, we need to divide our $b$ value by $2$.
-In the case of the first problem, we are dealing with $b = 0$. Now we can
-classify our PDE as follows:
-\[
-b^{2} - ac = 0 - y < 0 \implies \text{Elliptic}
-\]
-Note that we assume $y > 0$. So our PDE is elliptic. Now we need to make it look
-like it's corresponding equation from the classical trinity. In this case we
-have an Elliptic equation, so we want to make it look like Laplace's equation.
-We need to find values for $u_{xx}$ and $u_{yy}$, and then substitute those into
-the equation. This is similar to how we solved the first order PDE's; let's look
-at our characteristic equation:
-\begin{gather*}
-am^{2} + bm + c = 0\\
-ym^{2} + 1 = 0\\
-m = \pm \frac{i}{\sqrt{y}}
-\end{gather*}
-\noindent So, now we have two solutions, and they are both imaginary as expected.
-Next we need to find values for $S$ and $T$ that we can substitute into the
-equation. We will do both of them at the same time here. Notice that we will
-drop the $i$:
-\begin{gather*}
-\frac{dy}{dx} = \frac{1}{\sqrt{y}} \qquad\qquad\qquad \frac{dy}{dx} = -\frac{1}{\sqrt{y}}\\
-\int\frac{dy}{dx} =\int\frac{1}{\sqrt{y}} \qquad\qquad\qquad \int\frac{dy}{dx} = -\int\frac{1}{\sqrt{y}}\\
-S = \frac{2}{3}y^{\left(\frac{3}{2}\right)} - x \qquad\qquad\qquad T = \frac{2}{3}y^{\left(\frac{3}{2}\right)} + x
-\end{gather*}
-\noindent Now we can substitute into our equation. Unfortunately, we will need
-to take the partial derivative twice in order to get the substitutions we need.
-Let's start with $u_{xx}$.
-\begin{align*}
-u_{x} &= u_{S}\frac{\partial S}{\partial x} + u_{T}\frac{\partial T}{\partial x}\\
-&= -u_{S} + u_{T}\\
-u_{xx} &= \frac{\partial}{\partial x}(-u_{S} + u_{T})\\
-&= u_{SS} + u_{TT}
-\end{align*}
-\noindent We will skip the math for finding $u_{yy}$, which would give us
-\[
-u_{yy} = yu_{SS} + yu_{TT}
-\]
-\noindent Now we plug everything in. Remember that we were under the assumption
-that $y > 0$:
-\begin{gather*}
-y(u_{SS} + u_{TT}) + yu_{SS} + yu_{TT} = 0\\
-2yu_{SS} + 2yu_{TT} = 0\\
-u_{SS} + u_{TT} = 0
-\end{gather*}
-\noindent And we are finished.
-
+\include{classify}
 \newpage
-\section{Fourier Series and Partial Differential Equations}
-\hrule
-\noindent\\
-\subsection{Finding Fourier Series}
-\indent It's finally time to talk about Fourier Series. Before we start though,
-it may be helpful to look at some common patterns amongst the Fourier Series.\\\\
-\centerline{
-\bgroup
-\def\arraystretch{1.5}
-\begin{tabular}{|c|c|c|c|}
-\hline
-& Cosine & Sine & Full Series \\ \hline
-Interval & $[0,L]$ & $[0,L]$ & $[-L,L]$\\ \hline
-Basis & $\{1,\cos{\left(\frac{n\pi x}{L}\right)}\}$ &
-$\{\sin{\left(\frac{n\pi x}{L}\right)}\}$ &
-$\{1,\sin{\left(\frac{n\pi x}{L}\right)},\cos{\left(\frac{n\pi x}{L}\right)}\}$\\ \hline
-Series &
-$\frac{b_{0}}{2} + \sum_{n=1}^{\infty}[b_{n}\cos{\left(\frac{n\pi x}{L}\right)}]$&
-$\sum_{n=1}^{\infty}[a_{n}\sin{\left(\frac{n\pi x}{L}\right)}]$&
-$\frac{b_{0}}{2} + \sum_{n=1}^{\infty} [a_{n}\sin{\left(\frac{n\pi x}{L}\right)} + b_{n}\cos{\left(\frac{n\pi x}{L}\right)}]$\\ \hline
-\end{tabular}
-\egroup
-}\\\\
-\noindent To solve for the coefficients, we have:
-\begin{align*}
-Sine:  \qquad\qquad &a_{n} = \frac{2}{L}\int_{0}^{L}f(x)\sin{\left(\frac{n\pi x}{L}\right)}dx\\
-Cosine: \qquad\qquad &b_{n} = \frac{2}{L}\int_{0}^{L}f(x)\cos{\left(\frac{n\pi x}{L}\right)}dx\\
-&b_{0} =  \frac{1}{L}\int_{0}^{L}f(x)dx\\
-Full: \qquad\qquad &a_{n} = \frac{2}{L}\int_{-L}^{L}f(x)\sin{\left(\frac{n\pi x}{L}\right)}dx\\
-&b_{n} = \frac{2}{L}\int_{-L}^{L}f(x)\cos{\left(\frac{n\pi x}{L}\right)}dx\\
-&b_{0} =  \frac{1}{L}\int_{-L}^{L}f(x)dx
-\end{align*}
-\indent It will also help us to quickly review the concept of odd and even
-functions, as they will help us to simplify the process of computing Fourier
-series later. Specifically, let's look at the graphs for $\sin$ and $\cos$ on
-$[-\pi,\pi]$, as well as another function that will turn out to be neither even
-nor odd.
-\begin{center}
-\begin{tabular}{c c c}
-\includegraphics[scale=0.4]{cos_eo} & \includegraphics[scale=0.4]{sin_eo} & \includegraphics[scale=0.4]{cube_eo}\\
-$f(x) = \cos{x}$ & $f(x) = \sin{x}$ & $f(x) = x^{3}-x +1$
-\end{tabular}
-\end{center}
-\noindent\\
-\indent We should recall that a function is \textit{even} when $f(x) = f(-x)$
-for all $x$ in the domain of $f$. Graphically speaking, an \textit{even}
-function is reflected across only the $y$-axis. By looking at the above graphs,
-we can see that $f(x) = \cos{x}$ is an even function.\\
-\indent A function is \textit{odd} when $-f(x) = f(-x)$ for all x in the domain
-of $f$. Graphically speaking, the function is reflected across the $x$-axis and
-the $y$-axis. From the above graphs, we can say that $f(x) = \sin{x}$ is an odd
-function. Note that we can have a function that is neither odd nor even, but we
-won't see much of this in the following problems. The reason that we want to
-discuss odd and even functions here is because the integration that we will be
-seeing later will be affected by the symmetry.\\\\
-\indent The first type of series we will look at is the Full Fourier Series.
-Note that this series is just a combination of a Fourier Sine Series and a
-Fourier Cosine Series. A typical problem would look something like this:\\\\
-\textbf{\textit{Ex:}} Find the corresponding Fourier Series for
-\begin{gather*}
-f(x)=
-\begin{cases*}
-L \qquad -L \leq x \leq 0\\
-2x \qquad 0 \leq x \leq L
-\end{cases*}
-\end{gather*}
-\indent \textbf{\textit{Solution:}} So we have a piecewise function here, but
-that shouldn't throw us off too much; all we need to do is combine the integrals
-for each part of the coefficient solution. Let's start by finding the terms for
-the $cos$ series.
-\begin{align*}
-B_{0} &= \frac{1}{2L}\left[\int_{-L}^{0}Ldx + \int_{0}^{L}2xdx\right]\\
-&= \frac{1}{2L}[2L^{2}]\\
-&= L
-\end{align*}
-\noindent Now We need to find $B_{n}$, and then we will have our cosine part of
-the series. The integration can get pretty messy, so we will skip some steps.
-\begin{align*}
-B_{n} &= \frac{1}{L}\left[\int_{-L}^{0}L\cos{\left(\frac{n\pi x}{L}\right)}dx + \int_{0}^{L}2x\cos{\left(\frac{n\pi x}{L}\right)}dx\right]\\
-&= \frac{1}{L}\left[0 + \left(\frac{2L^{2}}{n^{2}\pi^{2}}\right)((-1)^{n} - 1)\right]\\
-&= \frac{2L}{(n\pi)^{2}}((-1)^{n} - 1)
-\end{align*}
-\noindent Note that the sequence will be 0 for all $n = 2k$. This allows us to
-substitute $2k+1$ for $n$, since we don't care about the $0$ terms. For our
-cosine part of the series, we now have:
-\begin{align*}
-B_{n} &= \frac{2L}{(2k+1)^{2}\pi^{2}}((-1)^{2k+1}-1)\\
-&= \frac{2L}{(2k+1)^{2}\pi^{2}}(-2)\\
-&= \frac{-4L}{(2k+1)^{2}\pi^{2}}
-\end{align*}
-\noindent Now we need to find the sine part of the series. This is easier,
-since we only need to find $A_{n}$.
-\begin{align*}
-A_{n} &= \frac{1}{L}\left[\int_{-L}^{0}L\sin{\left(\frac{n\pi x}{L}\right)}dx +
-\int_{0}^{L}2x\sin{\left(\frac{n\pi x}{L}\right)}dx \right]\\
-&= \frac{L}{n\pi}\left[-1-(-1)^{n}\right]
-\end{align*}
-\noindent Note that we can do a similar substitution here as we did with the
-cosine series. This time, however, we will have $0$ for $n = 2k+1$ We now have:
-\begin{align*}
-A_{n} &= \frac{L}{2k\pi}\left[-1-(-1)^{2k}\right]\\
-&= \frac{L}{2k\pi}\left[-1 - 1\right]\\
-&= \frac{L}{2k\pi}\left[-2\right]\\
-&= \frac{-2L}{2k\pi}
-\end{align*}
-\noindent Now we have all the parts that we need. Our final answer will be
-the sum of all the parts:
-\begin{align*}
-f(x)&\sim \frac{L}{2} + \sum_{k = 0}^{\infty}\left[\frac{-4L}{(2k+1)^{2}\pi^{2}}
-\cos{\left(\frac{(2k + 1)\pi x}{L}\right)} \right] + \sum_{k =
-0}^{\infty}\left[\frac{-2L}{2k\pi}\sin{\left(\frac{2k\pi x}{L}\right)} \right]\\
-&\sim \frac{L}{2} + \sum_{k = 0}^{\infty}\left[\frac{-4L}{(2k+1)^{2}\pi^{2}}
-\cos{\left(\frac{(2k + 1)\pi x}{L}\right)} + \frac{-2L}{2k\pi}\sin{
-\left(\frac{2k\pi x}{L}\right)} \right]
-\end{align*}
-\noindent And we are done.\\
-
+\include{fourier_series}
 \newpage
-\indent Now lets take a look at a slightly more theoretical Fourier Series.\\
-\noindent\textbf{\textit{Ex:}} Find the corresponding Fourier Series for\\
-\[f(x)=\left|\cos{x}\right|\qquad -\pi\leq x \leq \pi\]
-\indent\textbf{\textit{Solution:}} Before we dive into calculations here, let's
-analyze the function at hand. We have the absolute value of $\cos{x}$. Usually,
-$\cos$ would be an odd function, which would imply that our $\cos$ series
-component to the whole Fourier Series would be $0$. However, the absolute value
-makes our function even, and as such the $\sin$ component of the series will be
-$0$. That's good news for us, because now we don't have to find $a_{n}$; we
-actually only have a Fourier Cosine Series here. Recall that we have a formula
-for $b_{n}$ and $b_{0}$, and that our general solution will have the form of
-\[\frac{b_{0}}{2} + \sum_{n=1}^{\infty}[b_{n}\cos{\left(\frac{n\pi x}{L}\right)}]\]
-\noindent Our function is bounded by $\pi$, so we really have $L= \pi$. This gives us:
-\[\frac{b_{0}}{2} + \sum_{n=1}^{\infty}[b_{n}\cos{\left(nx\right)}]\]
-\noindent Let's go ahead and find our $b_{0}$:
-\begin{align*}
-b_{0} &= \frac{1}{\pi}\int_{-\pi}^{\pi}\left|\cos{x}\right|dx\\
-&= \frac{2}{\pi}\int_{0}^{\pi}\left|\cos{x}\right|dx\\
-&= \frac{4}{\pi}\int_{0}^{\frac{\pi}{2}}\left|\cos{x}\right|dx\\
-&= \frac{4}{\pi}\left[\sin{\frac{\pi}{2}} - \sin{0}\right]\\
-&= \frac{4}{\pi}
-\end{align*}
-\noindent We know that since the function is even, we can say that the the
-integral of the entire range is really twice the integral of half the range.
-Since we are dealing with the cosine function, we actually have four times a
-quarter of the range. If we take a look at a graph, this may make more sense:\\
-\begin{center}
-\begin{tabular}{c c}
-\includegraphics[scale=0.5]{cos_abs} & \includegraphics[scale=0.5]{cos}\\
-$f(x) = |\cos{x}|$ on $[0,2\pi]$ & $f(x) = \cos{x}$ on $[0,2\pi]$
-\end{tabular}
-\end{center}
-\noindent\\ You can see that the section of the graph from
-$[\frac{\pi}{2},\frac{3 \pi}{2}]$ is positive for the $|\cos{x}|$, but negative
-for $\cos{x}$. If we simply integrated the $|\cos{x}|$ over $[0,\pi]$, the
-result would be 0, but graphically this result would not make any sense. Now
-let's take a look at the $b_{n}$ term. Before we jump into the integration,
-let's think about this function:
-\[\left|\cos{x}\right| =
-\begin{cases*}
-\cos(x) \qquad 0\leq x \leq \frac{\pi}{2}\\
--\cos(x)\quad  \frac{\pi}{2}\ \leq x \leq \pi
-\end{cases*}
-\]
-\noindent Now we can set up the formula for the $b_{n}$ term. We have as follows:
-\begin{align*}
-b_{n} &= \frac{2}{\pi}\left[ \int_{0}^{\frac{\pi}{2}} \cos{x}\cos{nx}dx - \int_{\frac{\pi}{2}}^{\pi} \cos{x}\cos{nx}dx \right]\\
-&= \frac{2}{\pi} \left[ \frac{-2 \cos{\left(\frac{n\pi}{2}\right)} + \sin{\left(n\pi\right)}}{n^{2} - 1}\right]\\
-\end{align*}
-\noindent Note that (as we have seen before) the sine term disappears; however,
-the more interesting part is what happens with our cosine term. The value of
-$\cos{\left(\frac{n\pi}{2}\right)}$ will be 0 at all odd $n$, or for all
-$n = 2k + 1$. So, we will let our $n = 2k$. This gives us:
-\begin{align*}
-b_{n} &= \frac{2}{\pi} \left[ \frac{-2 \cos{\left(k \pi \right)}}{(2k)^{2} - 1}\right]\\
-&= -\frac{4}{\pi}\left[\frac{(-1)^{k}}{(2k)^{2}-1}\right]
-\end{align*}
-\noindent And that leaves us with the final result
-\[
-f(x)\sim \frac{2}{\pi} - \sum_{k=0}^{\infty}\left[\frac{4}{\pi}\left[\frac{(-1)^{k}}{(2k)^{2}-1}\right]
-\cos{\left(\frac{2k\pi}{2}\right)}\right]
-\]
-And we are done.
+\include{second_order}
 \newpage
-
-\subsection{Convergence of Fourier Series}
-\indent Up to this point, we have found a Fourier series and left our solutions
-at that. We will now look into the convergence of a Fourier series. This topic
-includes two questions:
-\begin{enumerate}
-\item Does the series converge?
-\item If the series converges, to what value does it converge?
-\end{enumerate}
-\noindent Before we jump into convergence, we need to review some ideas.
-\begin{enumerate}
-\item \textit{\textbf{Jump Discontinuity:}} To say that $f(x)$ has a jump
-discontinuity at $x = a$ is to say that
-\[
-\lim_{x \to a^{-}}f(x) \neq \lim_{x \to a^{+}}f(x)
-\]
-Where both the limit from the left and the limit from the right exist.
-\item \textit{\textbf{Piecewise Smooth:}} A function $f(x)$ is said to be
-piecewise smooth if the function can be broken into two distinct intervals, and
-on each interval both $f(x)$ and $f^{'}(x)$ are continuous.
-\item \textit{\textbf{Periodic Extension:}} The periodic extension of $f(x)$ on
-$[-L,L]$ is the repetition of the function on intervals on periods to the left
-and right of $[-L,L]$.
-\end{enumerate}
-\indent With these definitions, we can lay out a general rule for convergence of
-a Fourier series. We won't give a proof of these rules for the time being. We
-just need them to lay out some groundwork for the rest of the information we
-will present on Fourier series. We have the following:
-\begin{quote}
-Suppose $f(x)$ is a function, and that $f(x)$ is piecewise smooth on the interval
-$[-L,L]$. The corresponding Fourier series for $f(x)$ converges to either
-\begin{enumerate}
-\item the periodic extension of $f(x)$, if $f(x)$ is continuous, or
-\item the average of the two one-sided limits, if the periodic extension of
-$f(x)$ has a jump discontinuity at $x = a$.
-\end{enumerate}
-\end{quote}
-\noindent We should note that the second condition can be expressed by the formula
-\[\frac{1}{2}\left[\lim_{x \to a^{-}}f(x) + \lim_{x \to a^{+}}f(x)\right]\]
-which gives us a nice way to calculate the value to which the series converges
-if we happen to meet the criteria in (2). Let's take a look at some examples.\\\\
-
-\noindent\textbf{\textit{Ex:}} Find the Fourier series for the following
-equation, and determine if the series converges. If it does, tell where the
-series will converge.
-\[ f(x) =
-\begin{cases}
-L \qquad\qquad -L \leq x \leq 0\\
-2x \qquad\qquad 0 \leq x \leq L
-\end{cases}
-\]
-\indent\textbf{\textit{Solution:}} We will go ahead and skip the details of
-finding the Fourier series for this function. If you do want to find it, you
-will have to find a full piecewise Fourier series. For this particular $f(x)$,
-we have:
-\begin{align*}
-f(x) &\sim L + \sum_{n=1}^{\infty}\left[\frac{2L}{n^{2}\pi^{2}}([-1]^{n}-1)\cos{\left(\frac{n\pi
-x}{L}\right)}\right] - \sum_{n=1}^{\infty}\left[\frac{L}{n\pi}(1 + [-1]^{n})\sin{\left(\frac{n\pi
-x}{L}\right)}\right]
-\end{align*}
-\noindent Notice that we didn't simplify this solution to a nicer form.
-If we evaluate this series, we will run across terms where the series is 0, but
-that is okay for now. This function is also defined as a piecewise function,
-and on both intervals ($[-L,0]$ and $[0,L]$), $f(x)$ is continuous. In other
-words, $f(x)$ is piecewise smooth, so we know that this series converges. Now
-let's take a look at the graph of $f(x)$ on an arbitrary interval of $[-1,1]$.
-\begin{center}
-\begin{tabular}{c}
-\includegraphics[scale=0.6]{pw_graph_01}\\
-$
-f(x) =
-\begin{cases}
-1 \qquad\qquad -1 \leq x \leq 0\\
-2x \qquad\qquad 0 \leq x \leq 1
-\end{cases}
-$
-\end{tabular}
-\end{center}
-\noindent We can see that there is a jump in the graph at $x = 0$. In other
-words, there is a jump discontinuity at $x = 0$, and that means that we need to
-use (2) in order to find the value of convergence. We will let $\sum$ stand in
-for our entire series. We have:
-\begin{align*}
-\sum &\to \frac{1}{2}\left[\lim_{x\to 0}L + \lim_{x\to0}2x\right]\\
-&\to \frac{1}{2}[L + 0]\\
-&\to \frac{L}{2}
-\end{align*}
-\noindent So, at $x=0$, the series converges to $\frac{L}{2}$. We are not done
-yet, however. Now, we need to consider the periodic extension of the series. The
-function will repeat every $[-L,L]$, so we need to look at what the convergence
-of the series looks like at $-L$ and $L$. Let's take a look:
-\begin{alignat*}{3}
-\sum&\to\frac{1}{2}\left[\lim_{x\to-L^{-}}f(x) + \lim_{x\to-L^{+}}f(x)\right] \qquad\qquad \sum&&\to
-\frac{1}{2}\left[\lim_{x\to L^{-}}f(x) + \lim_{x\to L^{+}}f(x)\right]\\
-&\to \frac{1}{2}[2L + L] &&\to \frac{1}{2}[2L + L]\\
-&\to\frac{3L}{2} &&\to\frac{3L}{2}
-\end{alignat*}
-\noindent And we are done.
+\include{cauchy}
 \newpage
-
-\section{Second Order Partial Differential Equations}
-\hrule
-\noindent\\
-\subsection{The Heat and Wave Equations}
-\indent Let's take a look at the second order PDE. We will start with a
-homogeneous equation, which we will tackle with the separation of variables
-method. The general strategy is that we want to take our PDE and convert it to
-an ODE, which we know how to solve. The problems presented here also involve
-boundary conditions, which will require us to solve for eigenvalues/functions.
-We will also be dealing with series (specifically Fourier Series) to help us
-solve these problems. Let's look at an example:\\
-
+\include{transforms}
 \newpage
-\noindent\textbf{\textit{Ex:}} Solve the following boundary value problem:
-\[
-\begin{cases}
-u_{tt} - u_{xx} = 0\\
-u(0,t) = 0\qquad u(1,t) = 0\\
-u(x,0) = f(x)\qquad u_{t}(x,0) = g(x)
-\end{cases}\\
-\]
-\[f(x) = 2\sin{(\pi x)} + 3\sin{(2\pi x)}\qquad g(x) = 4\sin{(3\pi x)} - 7\sin{(5\pi x)}\]
-given the general solution:
-\[u(x,t) = \sum_{n=1}^{\infty}\left[\left(a_{n}\cos{[n\pi t]} + b_{n}\sin{[n\pi t]}\right)\sin{[n\pi
-x]}\right]\]
-\indent\textbf{\textit{Solution:}} There's a lot going on here, so let's break
-it down. First of all, notice that we have two initial conditions, one of which
-is a derivative. This should make sense since the equation have two second order
-partial derivatives included in it. Next notice that we have a Fourier Sine Series,
-despite both the $\sin{[n\pi t]}$ and $\cos{[n\pi t]}$ terms. If we had worked
-out the entire problem, we would see that our eigenvalue/function for this
-equation takes the form of
-\[
-\{\lambda_{n} = n\pi, \sin{(n\pi x)}\}
-\]
-\noindent which tells us that the corresponding series is a Fourier Sine Series.
-We also see the initial conditions on our equation are already representative
-of a series of sine functions, which is not just a conincidence. To apply the
-initial conditions, we will need both the general solution and the derivative
-with respect to $t$ of the general solution. Let's go ahead and apply the first
-initial condition, $u(x,0) = f(x)$. Letting $t= 0$ gives us:
-\[
-u(x,0) = 2\sin{(\pi x)} + 3\sin{(2\pi x)} = \sum_{n=1}^{\infty}\left[a_{n}\sin{(n\pi x)}\right]
-\]
-\noindent We see that the term with $b_{n}$ has dropped out, since
-$\sin{(0)} = 0$. Now we just have to match up the coefficients on $a_{n}$. We
-can see that when we have $n=1$, $a_{n} = 2$, and when $n=2$, $a_{n} = 3$. For
-all other $n$, we have $a_{n} = 0$, so we ignore those terms. Next, we can apply
-the initial condition for the derivative of the general solution with respect to
-$t$. We expect that all the terms related to $a_{n}$ will disappear, and in fact
-they will. Let's take a look:
-\begin{align*}
-u_{t}(x,t) &= \sum_{n=1}^{\infty}\left[(-a_{n}n\pi\sin{[n\pi t]} + b_{n}n\pi\cos{[n\pi t]})\sin{n\pi
-x}\right]\\
-u_{t}(x,0) &= 4\sin{(3\pi x)} - 7\sin{(5\pi x)} = \sum_{n=1}^{\infty}\left[n\pi b_{n}\sin{(n\pi
-x)}\right]
-\end{align*}
-\noindent We are going to do the same thing here that we did to find our $a_{n}$
-terms, but notice that we don't just have $b_{n}$; we actually have $n\pi b_{n}$.
-We have the following:
-\begin{alignat*}{3}
-&3\pi b_{3} = 4 \qquad \qquad \qquad &&5\pi b_{5} = -7\\
-&b_{3} = \frac{4}{3\pi} &&b_{5} = -\frac{7}{5\pi}
-\end{alignat*}
-\noindent Much like the solutions for $a_{n}$, all the other $b_{n}$ terms are 0.
-Now we can construct a particular solution. In this case, the answer will not be
-in the form of a series, since we have definite terms. Our solution is as follows:
-\[
-u(x,t) = 2\cos{(\pi t)}\sin{(\pi x)} + 3\cos{(2\pi t)}\sin{(2\pi x)} + \frac{4}{3\pi}\sin{(3\pi
-t)}\sin{(3\pi x)} - \frac{7}{5\pi}\sin{(5\pi t)}\sin{(5\pi x)}
-\]
-\noindent And we are done.
+\include{pde_formal_slns}
 \newpage
-
-\noindent \textbf{\textit{Ex:}} Solve the following PDE
-\[
-\begin{cases}
-\frac{\partial u}{\partial t} = k \frac{\partial^{2} u}{\partial x^{2}}\\
-u(x,0) = x\\
-u(0,t) = u(L,t) = 0\\
-\end{cases}
-\]
-\indent \textbf{\textit{Solution:}} We will use the separation of variable
-technique to solve this problem. We know to use separation of variable because
-we are dealing with a homogeneous PDE with homogenous boundary conditions.
-Separation of variables tells us that our solution will be in the form of
-$u(x,t) = X(x)T(t)$. If we take this fact at face value, we can use the
-separation of variables method to solve this. First, we substitute in our solution
-into the PDE:
-\begin{align*}
-\frac{\partial}{\partial t}(X(x)T(t)) &= k \frac{\partial^{2}}{\partial x^{2}}(X(x)T(t))\\
-X(x)\frac{dT}{dt} &= k\frac{d^{2}X}{dx^{2}}T(t)\\
-\frac{dT}{dt}\frac{1}{T(t)} &= \frac{d^{2}X}{dx^{2}}\frac{k}{X(x)}
-\end{align*}
-\noindent The only way that our function of x and our function of t can be equal
-is if they are equal to a constant; thus we now have:
-\[
-\frac{dT}{dt}\frac{1}{T(t)} = \frac{d^{2}X}{dx^{2}}\frac{k}{X(x)} = -\lambda
-\]
-\noindent And now we have a first order ODE for the time and a second order ODE
-for the position:
-\[
-\frac{dT}{dt} = -k \lambda T(t) \qquad\text{and}\qquad \frac{d^{2}X}{dx^{2}} = -\lambda X(x)
-\]
-\noindent Our next step is to find the Eigenvalues and Eigenfunctions of the
-second order ODE; we will come back to the time ODE when we have values for
-$\lambda$.\\
-\indent To solve for these values, we have to find non-trivial eigenvalues that
-correspond to the boundary values of the ODE. In this case, we know that
-$X(0) = 0$ and $X(L) = 0$. These boundary conditions come from the BC on the PDE
-which states $u(0,t) = u(L,t) = 0$. To find the eigenvalues/functions, we need
-to find the characteristic equation corresponding to the ODE in question. Let's
-go ahead and find it:
-\begin{gather*}
-\frac{d^{2}X}{dx^{2}} = -\lambda \frac{X(x)}{k}\\
-\frac{d^{2}X}{dx^{2}}  + \lambda \frac{X(x)}{k} = 0\\
-r^{2} + \lambda = 0\\
-r^{2} = -\lambda\\
-r = \pm \sqrt{- \lambda}
-\end{gather*}
-\indent Now, we can substitute values of $\lambda$ to find out solutions. Let's
-start with $\lambda > 0$. Since we have a $-\lambda$ under the radical, we will
-have $r = \pm \sqrt{\lambda} \mathrm{i}$. This will yield the solution of
-\[
-y(x) = c_{1}\cos{\sqrt{\lambda}x} + c_{2}\sin{\sqrt{\lambda}x}
-\]
-\noindent Applying our boundary conditions gives us:
-\[
-y(0) = c_{1}\cos{0} + c_{2}\sin{0} = 0
-\]
-\noindent Now we know that $c_{1} = 0$. This result came from the fact that
-$\sin{0} = 0$, so the $c_{2}$ term disappears. Now we are left with $\cos{0} = 1$,
-which tells us that $c_{1} = 0$. We can apply this to our second boundary
-condition like so:
-\begin{gather*}
-y(L) = 0 = c_{2}\sin{\sqrt{\lambda}L}\\
-\end{gather*}
-\noindent It looks like there will only be trivial solutions, but we know this
-isn't true due to the Jacobian matrix. Recall that if the determinant of the
-Jacobian $\neq 0$, then we only have trivial solutions, and if the determinant
-of the Jacobian $= 0$, then we do not have the trivial solution. Let's take a
-look at the Jacobian for this case:
-\begin{gather*}
-\left|
-\begin{array}{cc}
-1 & 0\\
-\cos{\sqrt{\lambda}L} & \sin{\sqrt{\lambda}L}
-\end{array}
-\right| = 0\\
-\sin{\sqrt{\lambda}L} = 0\\
-\sqrt{\lambda}L = \arcsin{0}\\
-\sqrt{\lambda} = \frac{\arcsin{0}}{L}\\
-\lambda = \left(\frac{\arcsin{0}}{L}\right)^{2}\\
-\lambda = \left(\frac{n\pi}{L}\right)^{2}
-\end{gather*}
-\noindent Notice that we replaced the $\arcsin{0}$ term with $n\pi$. Recall from
-trig that the $\sin$ function is $0$ at every $n\pi$, starting with $n = 0$.
-Next, we just need to find our corresponding eigenfunction. Our $c_{1}$ was $0$,
-so we can ignore the entire $\cos$ term. Thus, we have an eigenfunction from
-plugging in our $\lambda$ that looks like:
-\[
-X(x) = \sin{\frac{n\pi x}{L}}
-\]
-\indent Unfortunately, we aren't done yet. Now we have to find the corresponding
-eigenvalues/functions for $\lambda = 0$. This isn't too hard, and can be done in
-a few lines if we recall the characteristic equation. Plugging in $0$ for
-$\lambda$ gives us that $r = 0$. Using this piece of information, we now know
-that $y(x) = c_{1} + xc_{2}$. Applying our boundary conditions gives us
-$c_{1} = 0, c_{2} = 0$. We can see that we have only trivial solutions for
-$\lambda = 0$.\\
-\indent Our final case is for $\lambda < 0$. I'll go ahead and tell you that
-this will yield another trivial solution, but we can take a look at the work
-behind it. Since $\lambda > 0$, we know that our $r$ from the characteristic
-equation will be $r = \pm \sqrt{\lambda}$. From here we are going to do something
- a little unconventional that will help us. Usually, we would have a solution of
-the form $y(x) = c_{1}e^{\sqrt{\lambda}x} + c_{2}x e^{-\sqrt{\lambda}x}$. In our
-case, it will be more convenient to use hyperbolic trig functions. It's not too
-far off from our solution from $\lambda > 0$:
-\begin{align*}
-y(x) &= c_{1}e^{\sqrt{\lambda}x} + c_{2}e^{-\sqrt{\lambda}x}\\
-&= c_{1}\cosh{\sqrt{\lambda}x} + c_{2}\sinh{\sqrt{\lambda}x}
-\end{align*}
-\noindent If we apply the boundary conditions (remember that
-$\sinh{0} = 0$ and $\cosh{0} = 1$), we will find that we have only trivial
-solutions.\\
-\indent So, we are done with the second order ODE, and now we turn our eye to
-the first order time equation. We know our $\lambda$, so it isn't too hard to solve:
-\begin{gather*}
-\frac{dT}{dt} = -\lambda T(t)\\
-T(t) = ce^{-k\left(\frac{n\pi}{L}\right)^{2}t}
-\end{gather*}
-\indent Finally we can put it all together. Our penultimate solution will be
-\[
-u_{n}(x,t) = B_{n}\sin{\left(\frac{n\pi x}{L}\right)}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}
-\]
-\noindent We can analyze this solution and tell where each piece comes from.
-The $B_{n}$ is a coefficient that results from our $c's$ in the general solutions
-of the differential equations. The $sin$ part results from out eigen vector.
-Remember, we get a different value for each $n$, so there may or may not be $0$
-terms in the series. Finally, we have the exponential term. This term comes from
-the solution to our time equation. When we put it all together, we can see that
-this is, of course a series; for every $n$, we get a different solution. From
-here we can find the Fourier Series (which will be a Fourier Sine Series). We
-won't worry about that yet, but we will come back to it.
-\newpage
-
-\indent It will benefit us to take a moment and discuss how boundary conditions
-effect our Fourier series. We have three different combinations of boundary
-conditions, where each one will produce a different result in our series solution.
-\begin{enumerate}
-\item \textit{Dirichlet Boundary Conditions}\\
-Dirichlet boundary conditions involve the function for which we are looking. In
-other words, the boundary conditions are presented in terms of the functions we
-are trying to find, rather than the derivative of that function. Typically, this
-type of boundary condition will take the form of (in the homogeneous case)
-\[u(0,t) = u(L,t) = 0.\]
-In regards to our Fourier series, we expect to see $\sin{(\frac{n\pi x}{L})}$.
-Our basis will look like
-\[
-\left\{ \lambda_{n} = \left(\frac{n\pi}{2}\right)^{2}, \phi_{n}(x) = \sin{\left(\frac{n\pi x}{L}\right)} \right\}
-\]
-\item \textit{Neumann Boundary Conditions}\\
-Neumann boundary conditions involve the derivatives of the function for which we
-are solving. Neumann boundary conditions will look like
-\[u_{x}(0,t) = u_{x}(L,t) = 0.\]
-This will cause our Fourier series to contain $\cos{(\frac{n\pi x}{L})}$. Our
-basis will look like
-\[
-\left\{ \lambda_{n} = \left(\frac{n\pi}{2}\right)^{2}, \phi_{n}(x) = \cos{\left(\frac{n\pi x}{L}\right)} \right\}
-\]
-\item \textit{Robin Boundary Conditions}\\
-Robin boundary conditions will contain a mixture of the function and it's
-derivative. Typically, we will see
-\[ u_{x}(0,t) = u(L,t) = 0\qquad \text{or}\qquad u(0,t) = u_{x}(L,0) = 0.\]
-These boundary conditions will give us two different results, depending on whether
-we have the derivative first or second. We will have
-\begin{align*}
-u_{x}(0,t) = u(L,t) = 0 &\implies \left\{\lambda_{n} = \left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}\right)^{2}, \phi_{n}(x) = \cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}x\right)}\right\}\\
-u(0,t) = u_{x}(L,t) = 0 &\implies \left\{\lambda_{n} = \left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}\right)^{2}, \phi_{n}(x) =\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}x\right)}\right\}
-\end{align*}
-\end{enumerate}
-
-\indent Now we can take a look at a slightly harder PDE. We still have the heat
-equation, but we will not have homogeneous boundary conditions. Our goal here is
-to find a way to transform our boundary conditions into a homogenous form and take
-it from there.\\\\
-\noindent \textbf{\textit{Ex:}} Solve the following PDE
-\[
-\begin{cases*}
-u_{t} - u_{xx} = 0\\
-u(0,t) = 0,\ u(1,t) = t\\
-u(x,0) = x^{2}
-\end{cases*}
-\]
-\indent \textbf{\textit{Solution:}} We would like to just jump into this PDE and
-solve it, but it's not quite that simple. We have to find a way to express our
-boundary conditions in a homogeneous manner. We can do this through a clever
-substitution; we will create a new PDE, denoted $\Upsilon$. To find $\Upsilon$,
-we use the following formula:
-\[ \Upsilon = B_{1} + \frac{x}{l}\Delta B, \]
-\noindent where $B$ is a boundary condition. In our case, we have as follows:
-\begin{align*}
-\Upsilon &= B_{1} + \frac{x}{l}\Delta B\\
-&= 0 + \frac{x}{1}(t - 0)\\
-&= xt
-\end{align*}
-\noindent Now, we find a new PDE, denoted by $\upsilon$. We have another formula that tells us
-\[ \upsilon(x,t) = u(x,t) - \Upsilon(x,t) \]
-\noindent Applying this formula gives us
-\begin{align*}
-\upsilon(x,t) &= u(x,t) - \Upsilon\\
-&= u(x,t) - xt
-\end{align*}
-\noindent Let's take a moment to analyze the result.
-We wanted to find a new PDE in terms of our old one that would allow us to have homogeneous boundary conditions.
-We found a solution, called $\upsilon$, in terms of our original PDE, called $u$.
-Our original PDE was $u_{t} - u_{xx} = 0$, so to properly perform our substitution, we need to find $\upsilon_{t}$ and $\upsilon_{xx}$.
-Through some careful differentiation, we have
-\[ \upsilon_{t} = u_{t} - x \qquad\text{and}\qquad \upsilon_{xx} = u_{xx}\]
-\noindent To find our new PDE, we simply evaluate.
-\[
-\upsilon_{t} - \upsilon_{xx} = u_{t} - x - u_{xx} = u_{t} - u_{xx} - x = 0 - x = -x
-\]
-\noindent So we finally have a new PDE that looks like
-\[
-\begin{cases*}
-\upsilon_{t} - \upsilon_{xx} = -x\\
-\upsilon(0,t) = \upsilon(1,t) = 0\\
-\upsilon(x,0) = x^{2}
-\end{cases*}
-\]
-\noindent You may wonder how this has made anything simpler, since the PDE itself is no longer homogeneous.
-By forcing the boundary conditions to be homogeneous, we can now use a Fourier series to find our solution.
-Recall that earlier we discussed the boundary condition and their resulting eigenvalue/eigenfunction.
-Here, we have Dirichlet boundary conditions, which tells us that we have a Fourier sine series.
-We have that
-\begin{align*}
-\upsilon(x,t) &= \sum_{n = 1}^{\infty}\left[a_{n}(t)\sin{(n\pi x)}\right]\\
-\upsilon_{t}(x,t) &= \sum_{n = 1}^{\infty}\left[a_{n}'(t)\sin{(n\pi x)}\right]\\
-\upsilon_{xx}(x,t) &= -\sum_{n = 1}^{\infty}\left[a_{n}(t)(n\pi)^{2}\sin{(n\pi x)}\right]\\
-\end{align*}
-\noindent Which tells us that
-\[
-\upsilon_{t} - \upsilon_{xx} = \sum_{n = 1}^{\infty}\left[(a_{n}'(t) + (n\pi)^{2}a_{n}(t))\sin{(n\pi x)}\right] = -x
-\]
-\noindent This may look rather complicated, but all we have is a function's Fourier sine series. Essentially, we have
-\[
--x = \sum_{n = 1}^{\infty}\left[ b_{n}\sin{(n\pi x)} \right]
-\]
-\noindent To find $b_{n}$, we use the formula that was discussed in the earlier section about Fourier series. We have that
-\begin{align*}
-a_{n}'(t) + (n\pi)^{2}a_{n}(t) &= \int_{0}^{1}-x\sin{(n\pi x)}dx\\
-&= \frac{2x}{n\pi}\cos{(n\pi x)}\bigg\vert_{0}^{1}\\
-&= \frac{2}{n\pi}
-\end{align*}
-\noindent We now have an ODE which we can solve for $a_{n}$.
-Note that from out PDE, we do have an initial condition for the ODE, but it is not as simple as just saying $a_{n}(0) = x^{2}$.
-At $t = 0$, we have the following series:
-\[
-\sum_{n=1}^{\infty} a_{n}(0)\sin{(n\pi x)}
-\]
-\noindent We got this by taking our original $\upsilon(x,t)$ and plugging in $0$ for $t$. We know how to get $a_{n}(0)$, since this is just a simple Fourier sine series. It is given by
-\begin{align*}
-a_{n}(0) &= 2\int_{0}^{1}x^{2}\sin{(n\pi x)}dx\\
-&= \frac{(-1)^{(n + 1)}}{n\pi} + \frac{2}{(n\pi)^{3}}\left[ (-1)^{n} - 1 \right]
-\end{align*}
-\noindent We now have all we need to solve the ODE. We will omit the work for the solution, but the final answer for $a_{n}$ is
-\[
-a_{n}(t) = \frac{4\left[ (-1)^{n} - 1 \right]}{(n\pi)^{3}}e^{-(n\pi)^{2}t} + \frac{2(-1)^{n}}{(n\pi)^{3}}
-\]
-\noindent Now all that's left to do is go back to $u$ from $\upsilon$.
-This may seem hard, but we saw earlier that $\upsilon = u - \Upsilon$.
-Solving for $u$ gives us $u = \upsilon + \Upsilon$.
-Recall that $\Upsilon$ was $xt$, and $\upsilon$ was the series for which we just found $a_{n}$.
-Putting it all together gives us
-\[
-u(x,t) = xt + \sum_{n=1}^{\infty}\left[\frac{4\left[ (-1)^{n} - 1 \right]}{(n\pi)^{3}}e^{-(n\pi)^{2}t} + \frac{2(-1)^{n}}{(n\pi)^{3}}\right]\sin{(n\pi x)}
-\]
-\noindent And we are done.
-\newpage
-
-\noindent \textbf{\textit{Ex:}} Solve the following PDE
-\[
-\begin{cases*}
-u_{tt} - u_{xx} = 1 + t\cos{(2\pi x)}\\
-u_{x}(0,t) = u_{x}(1,t) = 0\\
-u(x,0) = 2\cos{(2\pi x)},\quad u_{t}(x,0) = 1 + \cos{(\pi x)} - \cos{(2\pi x)}
-\end{cases*}
-\]
-\indent \textbf{\textit{Solution:}} This problem is an example of the wave equation with a non-homogeneous solution.
-We are lucky this time; the boundary conditions are already homogeneous, so there is no need to change our variable.
-We know that since both of the boundary conditions are derivatives, we are dealing with a Fourier cosine series.
-This yields the eigenvalue/function of
-\[
-\left\{\lambda_{n} = (n\pi)^{2}, \cos{(n\pi x)}\right\}
-\]
-Since we are dealing with a cosine series, we also have a leading term, $a_{0}$. Our full series looks like
-\[
-u(x,t) = a_{0}(t) + \sum_{n=1}^{\infty} a_{n}(t)\cos{(n\pi x)}
-\]
-We need $u_{tt}$ and $u_{xx}$, so we have:
-\[
-u_{tt}(x,t) = a_{0}''(t) + \sum_{n=1}^{\infty} a_{n}''(t)\cos{(n\pi x)}
-\qquad \text{and} \qquad
-u_{xx}(x,t) = -\sum_{n=1}^{\infty} (n\pi)^{2}a_{n}(t)\cos{(n\pi x)}
-\]
-All we have to do now is simply plug into our original equation. This gives us
-\begin{align*}
-a_{0}''(t) + \sum_{n=1}^{\infty} a_{n}''(t)\cos{(n\pi x)} + \sum_{n=1}^{\infty} (n\pi)^{2}a_{n}(t)\cos{(n\pi x)} &= 1 + t\cos{(2\pi x)}\\
-a_{0}''(t) + \sum_{n=1}^{\infty}(a_{n}''(t) + (n\pi)^{2}a_{n}(t))\cos{(n\pi x)} &= 1 + t\cos{(2\pi x)}
-\end{align*}
-Now we can play the matching game. We have a $constant + series\ term$ on the right side, and $constant + series$ on the left, so we can match up the constants. This gives us
-\[
-a_{0}''(t) = 1
-\]
-Now we match up the series term. On the right hand side, we have $\cos{(2\pi x)}$, so we know that for $n= 2$, we have
-\[
-a_{2}''(t) + 4\pi^{2}a_{2}(t) = t
-\]
-$a_{0}$ and $a_{2}$ are the only terms that appear, so we can assume for now that $\forall n \neq 2\ \text{and}\ 3, a_{n} = 0$.
-We are left with two ODE's, so we now need to find the corresponding initial conditions.
-Let's apply the first initial condition to our PDE.
-This gives us
-\[a_{0}(0) + \sum_{n=1}^{\infty}a_{n}(0)\cos{(n\pi x)} = 2\cos{(2\pi t)}\]
-Which tells us
-\[a_{0} = 0 \qquad\qquad \qquad\qquad a_{2} = 2\]
-Now we can apply the second initial condition. After taking the partial derivative of $u$ with respect to $t$, we have
-\[
-a_{0}'(0) + \sum_{n=1}^{\infty}a_{n}'(0)\cos{(n\pi x)} = 1 + \cos{(\pi x)} - \cos{(2\pi x)}
-\]
-Some more clever pattern matching give us
-\[
-a_{0}'(0) = 1 \qquad\qquad a_{1}'(0) = 1 \qquad\qquad a_{2}'(0) = -1
-\]
-Notice that we have a term $a_{1}'$.
-We cannot simply ignore this value; rather we have to take it into account when we find our solution.
-We know that we have an ODE for $a_{1}(t)$, but it may not be evident how to express it.
-Recall that earlier we said $\forall n \neq 2\ \text{and}\ 3, a_{n} = 0$.
-This means that $a_{1} = 0$, and since $a_{1}$ is a coefficient in the Fourier series, we know that $a_{1}''(t) + \pi^{2}a_{1}(t) = 0$.
-We now have three different ODE's that we must solve:
-\begin{alignat*}{3}
-&\begin{cases}
-a_{0}''(t) = 1\\
-a_{0}'(0) = 1,\quad a_{0}(0) = 1
-\end{cases}
-&\begin{cases}
-a_{1}''(t) + \pi^{2}a_{1}(t) = 0\\
-a_{1}'(0) = 1,\quad a_{1}(0) = 0
-\end{cases}
-&&\begin{cases}
-a_{2}''(t) + 4\pi^{2}a_{2}(t) = t\\
-a_{2}'(0) = 1,\quad a_{2}(0) = 2
-\end{cases}
-\end{alignat*}
-Luckily, Mathematica is here to save the day. We will skip the work and let the computer do the thinking for us. We get that
-\begin{align*}
-a_{0}(t) &= \frac{1}{2}t^{2} + t\\
-a_{1}(t) &= \frac{\pi\cos{(t\pi)} + \sin{(t\pi)}}{\pi}\\
-a_{2}(t) &= \frac{2\pi t + 16\pi^{3}\cos{(2\pi t)} - \sin{(2\pi t)} + 4\pi^{2}\sin{(2\pi t)}}{8\pi^{3}}
-\end{align*}
-Now all we need to do is plug back in to our $u$. Our final answer becomes
-\[
-u(x,t) = \frac{1}{2}t^{2} + t + \frac{\pi\cos{(t\pi)} + \sin{(t\pi)}}{\pi}\cos{(\pi x)} +  \frac{2\pi t + 16\pi^{3}\cos{(2\pi t)} - \sin{(2\pi t)} + 4\pi^{2}\sin{(2\pi t)}}{8\pi^{3}}\cos{(2\pi x)}
-\]
-Just for interest, the plot of this solution looks like
-\begin{center}
-\includegraphics[scale=0.6]{pg23_plot}
-\end{center}
-On $-\pi \leq \theta \leq \pi$ and $-\pi \leq x \leq \pi$; and we are done.
-\newpage
-
-\subsection{The Laplace Equation}
-\noindent \\
-\indent Up to this point, we have only looked at the wave equation and the heat equation.
-There is, however, one other equation in the classical trinity: the Laplace Equation.
-We haven't dealt with this equation yet because we have to be very careful in our approach to solving it.
-When we looked at the homogeneous wave and heat equations, we used the general approach called \textit{Separation of Variables}.
-In the case of the homogeneous Laplace equation, we may not be able to tackle the problem using separation of variables.
-The problem arises from the various types of boundaries that effect the domain.
-We can either have a 'rectangular' bound, which will be separable, or a 'circular' bound which is \textbf{not} separable.
-Imagine that we have the following two problems:
-\begin{align*}
-&A =\begin{cases}
-u_{xx} + u_{yy} = 0\quad \text{on}\ [0,1]\times[0,2]\\
-u(x,0) = u(x,2) = 0\\
-u(0,y) = y, u(1,y) = 0
-\end{cases}
-&B =\begin{cases}
-u_{xx} + u_{yy} = 0\quad \text{on}\ x^{2} + y^{2} < a^{2}\\
-u(x,y)\ \big |_{x^{2} + y^{2} = a^{2}} = f(x,y)
-\end{cases}
-\end{align*}
-$A$ is an example of a separable Laplace equation.
-We can see graphically that the domain $[0,1] \times [0,2]$ is rectangular.
-$B$'s domain is not separable, since it is circular.
-Let's take a look at $A$.\\\\
-\noindent \textbf{\textit{Ex:}} Solve the following PDE
-\[
-\begin{cases*}
-u_{xx} + u_{yy} = 0\quad \text{on}\ [0,1]\times[0,2]\\
-u(x,0) = u(x,2) = 0\\
-u(0,y) = y, u(1,y) = 0
-\end{cases*}
-\]
-\indent \textbf{\textit{Solution:}} Earlier, we discussed the different boundary conditions we can encounter and their corresponding eigenvalues/functions.
-In this problem, we have Dirichlet boundary conditions.
-This tells us that
-\[
-\left\{\lambda_{n} = \left(\frac{n\pi}{2}\right)^{2}, \phi_{n}(y) = \sin{\left(\frac{n\pi y}{2}\right)}\right\}
-\]
-Notice that we used $y$ here instead of $x$.
-We did this because the boundary conditions are defined in terms of $y$, not $x$.
-Now we can continue to solve our problem.
-Recall that our separation of variables gives us two second order ODE's,
-\[
-X'' - \lambda X = 0 \qquad\qquad \text{and} \qquad\qquad \lambda Y - Y'' = 0
-\]
-We already solved for $Y(y)$ by finding the corresponding eigenvalues/vectors. Now we plug in our $\lambda_{n}$ into our $X$ ODE. Solving the resulting ODE will give us a basis of
-\[
-\left\{ e^{\frac{n\pi}{2}x}, e^{-\frac{n\pi}{2}x} \right\}
-\]
-Now we could continue at this point, but if we think ahead, we can see that when we eventually apply our initial conditions, nothing will cancel out.
-At this point, we want to transform our basis.
-We have as follows
-\[
-\left\{ e^{\frac{n\pi}{2}x}, e^{-\frac{n\pi}{2}x} \right\} \implies \left\{ \cosh{\left(\frac{n\pi x}{2}\right)}, \sinh{\left(\frac{n\pi x}{2}\right)} \right\}
-\]
-All we did here is apply an identity. Now we have everything we need to set up a solution. Our $u(x,y)$ will be
-\[
-u(x,y) = \sum_{n=1}^{\infty}\left[a_{n}\cosh{\left(\frac{n\pi x}{2}\right)} + b_{n}\sinh{\left(\frac{n\pi x}{2}\right)}\right]\sin{\left(\frac{n\pi y}{2}\right)}
-\]
-We can apply our initial conditions, which gives us $b_{n}\ \text{and}\ a_{n}$:
-\begin{align*}
-a_{n} &= \frac{4(-1)^{n+1}}{n\pi}\\
-b_{n} &= -\coth{\left(\frac{n \pi}{2}\right)}\frac{4(-1)^{n+1}}{n\pi}
-\end{align*}
-This yields a final solution of
-\[
-u(x,y) = \sum_{n=1}^{\infty}\left[ \frac{4(-1)^{n+1}}{n\pi} \cosh{\left(\frac{n\pi x}{2}\right)} -\coth{\left(\frac{n \pi}{2}\right)}\frac{4(-1)^{n+1}}{n\pi}\sinh{\left(\frac{n\pi x}{2}\right)}\right]\sin{\left(\frac{n\pi y}{2}\right)}
-\]
-And we are done
-\newpage
-
-\indent Now we turn our attention to the Laplace Equation with a circular domain.
-As we discussed earlier, we can't simply say that our solution will be in the form of $X(x)Y(y)$.
-However, we can force the domain to be rectangular if we use polar coordinates.
-Let's consider a variation of $B$ from above.\\\\
-
-\noindent \textbf{\textit{Ex:}} Solve the following PDE
-\[
-\begin{cases*}
-u_{xx} + u_{yy} = 0\quad \text{on}\ x^{2} + y^{2} < a^{2}\\
-u(a,\theta) = U_{1}\quad \text{on}\ 0 \leq \theta \leq \pi\\
-u(a,\theta) = U_{2}\quad \text{on}\ \pi \leq \theta \leq 2\pi
-\end{cases*}
-\]
-\indent \textbf{\textit{Solution:}} We want to express the partial derivatives here in terms of $r\ \text{and}\ \theta$. Recall that we have
-\[
-x = r\cos{\theta} \qquad\qquad \text{and} \qquad\qquad y = r\sin{\theta}
-\]
-First, we will find the partial derivatives of $x$ and $y$ with respect to $r$ and $\theta$. Since $x = r\cos{\theta}$,
-\[
-\frac{\partial x}{\partial r} = \cos{\theta} \qquad\qquad \frac{\partial x}{\partial \theta} = -r\sin{\theta}
-\]
-We can compute the derivatives for $y$ in a similar fashion:
-\[
-\frac{\partial y}{\partial r} = \sin{\theta} \qquad\qquad \frac{\partial y}{\partial \theta} = r\cos{\theta}
-\]
-Now we need to find $u_{rr}$ and $u_{\theta\theta}$. Let's focus on $u_{rr}$ first. We will need to start by finding $u_{r}$. We have as follows
-\begin{align*}
-\frac{\partial u}{\partial r}& = \frac{\partial u}{\partial x}\frac{\partial x}{\partial r} + \frac{\partial u}{\partial y}\frac{\partial y}{\partial r}\\
-&= \cos{\theta}\frac{\partial u}{\partial x} + \sin{\theta}\frac{\partial u}{\partial y}
-\end{align*}
-We have $u_{r}$, so now we need to compute $u_{rr}$. We have as follows:
-\begin{align*}
-\frac{\partial^{2} u}{\partial r^{2}} &= \cos{\theta}\frac{\partial }{\partial r}\frac{\partial u}{\partial x} + \sin{\theta}\frac{\partial }{\partial r}\frac{\partial u}{\partial y}\\
-&= \cos^{2}{\theta}\frac{\partial^{2} u}{\partial x^{2}} + 2\cos{\theta}\sin{\theta}\frac{\partial^{2}u}{\partial x\partial y} + \sin^{2}{\theta}\frac{\partial^{2} u}{\partial y^{2}}
-\end{align*}
-We won't delve into finding $u_{\theta\theta}$, as the partials get a little out of hand. However, we end up with
-\begin{align*}
-\frac{\partial^{2}u}{\partial r^{2}} + \frac{1}{r^{2}}\frac{\partial^{2}u}{\partial\theta^{2}} &= -\frac{1}{r}\frac{\partial u}{\partial r} + \frac{\partial^{2} u}{\partial x^{2}} + \frac{\partial^{2} u}{\partial y^{2}}\\
-u_{xx} + u_{yy} &= u_{rr} + \frac{1}{r}u_{r} + \frac{1}{r^{2}}u_{\theta\theta} = 0
-\end{align*}
-Let's go ahead and formalize it:
-\[
-\nabla^{2}u = u_{rr} + \frac{1}{r}u_{r} + \frac{1}{r^{2}}u_{\theta\theta} = 0
-\]
-Now we can treat this as a 2D problem; in other words, we can now use separation of variables to solve for $u$. As before, we know our solution will look like
-\[
-u(r,\theta) = R(r)T(\theta)
-\]
-Plugging everything in and realizing by the usual trick that everything s equal to a constant gives us
-\[
--\frac{T''}{T} = \frac{R'' + \frac{R'}{r}}{\frac{R}{r^{2}}} = \lambda
-\]
-Since we are dealing in polar coordinates, we will go ahead and walk through the eigenvalues/functions.
-We will be solving for $T(\theta)$, since our boundary conditions are in terms of $\theta$.
-We also should note that the solutions will repeat themselves every $2\pi$, since we are in the polar system.
-In addition, our $u$ should be finite at $r=0$.
-Let's take a look at the eigenvalues for the Laplace equation.\\\\
-\begin{minipage}[t]{0.32\textwidth}
-\underline{$\lambda = 0$}\\
-When $\lambda = 0$, we have \[T(\theta) = a\theta + b.\] This also tells us that \[R(r) = c\ln{(r)} + d.\] Putting everything together gives us \[ u(r,\theta) = (a\theta + b)(c\ln{r} + d).\] Since the solution is periodic on $2\pi$, we know that $a=0$.
-Since the solution has to remain finite at $r=0$, we know that $c = 0$.
-This gives us that $u = bd$, or that $u$ is equal to a constant.
-\end{minipage}
-\hspace{0.01\textwidth}
-\begin{minipage}[t]{0.32\textwidth}
-\underline{$\lambda > 0$}\\
-With $\lambda > 0$, we have our characteristic equation $(m = \pm\ \sqrt{\lambda})$ yielding a real valued solution. This tells us that
-\[
-T(\theta) = c_{1}\cosh{(\sqrt{\lambda}\theta)} + c_{2}\sinh{(\sqrt{\lambda}\theta)}.
-\]
-The solution is periodic on $2\pi$, so we can say that $c_{1} = c_{2} = 0$. Therefore, we have a trivial solution.
-\end{minipage}
-\hspace{0.01\textwidth}
-\begin{minipage}[t]{0.32\textwidth}
-\underline{$\lambda < 0$}\\
-$\lambda < 0$ gives us
-\[
-T(\theta) = c_{1}\cos{(\sqrt{\lambda}\theta)} + c_{2}\sin{(\sqrt{\lambda}\theta)}
-\]
-This solution is periodic with $2\pi$, and it yields a trivial eigenvector.
-\end{minipage}
-\noindent\\\\\\
-\noindent So we know that the only $\lambda$ that yields eigenvalues is $\lambda = 0$, and we know that this particular eigenvalue yields a constant solution.
-If we solve the differential equation for $R$, we will find that the only solution we care about is $r^{n}$.
-We can finally say that our solution will look like
-\[
-u(r,\theta) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}r^{n}[a_{n}\cos{(n\theta)} + b_{n}\sin{(n\theta)}]
-\]
-This solution is worth memorizing, as it will generally be the same for all variations of the Laplace equation similar to this one.
-Now we can apply the conditions for $\theta$.
-We have two conditions, but it is really just a piecewise function which we will call $f(\theta)$.
-We also know that our bound on the domain is $a$, or the length of the radius.
-This is enough information to find all the various variables in our $u(r,\theta)$.
-We have as follows:
-\[
-u(a,\theta) = f(\theta) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}a^{n}[a_{n}\cos{(n\theta)} + b_{n}\sin{(n\theta)}]
-\]
-Note that the $a$ which is the length of the radius is \textit{not} the same $a$ as $a_{n}$. Now we can solve for $a_{0}$ and $a_{n}$.
-\begin{alignat*}{3}
-a_{0} &= \frac{1}{\pi}\int_{0}^{2\pi}f(\theta)d\theta \qquad\qquad a_{n} &&= \frac{1}{\pi}\left[\int_{0}^{\pi}U_{1}\cos{(n\theta)}d\theta + \int_{\pi}^{2\pi}U_{2}\cos{(n\theta)}d\theta\right]\\
-&= U_{1} + U_{2} &&=0
-\end{alignat*}
-For $b_{n}$, we have
-\begin{align*}
-a^{n}b_{n} &= \frac{1}{\pi}\left[\int_{0}^{\pi}U_{1}\sin{(n\theta)}d\theta + \int_{\pi}^{2\pi}U_{2}\sin{(n\theta)}d\theta\right]\\
-&= \frac{1 - (-1)^{n}}{n\pi}(U_{1} - U_{2})\\
-b_{n} &= \frac{\left[1 - (-1)^{n}\right](U_{1} - U_{2})}{n\pi a^{n}}\\
-&= \frac{2(U_{1} - U_{2})}{(2k+1)\pi a^{2k+1}}
-\end{align*}
-Putting it all together gives us a final solution of
-\[
-u(r,\theta) = \frac{U_{1} + U_{2}}{2} + \sum_{k=0}^{\infty}r^{k}\frac{2(U_{1} - U_{2})}{(2k+1)\pi a^{2k+1}}
-\]
-And we are done.
-\newpage
-
-\indent Now let's look at another Laplace equation in polar coordinates.
-Again, we will be dealing with Dirichlet boundary conditions, so we will be able to jump straight into the solution.
-This one won't take us as long to solve, as we have already done a lot of the work.
-We are asked the following:\\\\
-\noindent \textbf{\textit{Ex:}} Solve the following PDE
-\[
-\begin{cases*}
-\nabla^{2}u = 0\\
-u(3,\theta) =
-\begin{cases*}
-1 \qquad\qquad 0 \leq \theta \leq \pi\\
-\sin^{2}{\theta}\qquad \pi \leq \theta \leq 2\pi
-\end{cases*}
-\end{cases*}
-\]
-\indent \textbf{\textit{Solution:}} We already know what our separable equation will look like:
-\[
-\frac{r^{2}R'' + rR'}{R} = -\frac{T''}{T} = -\lambda
-\]
-This yields an eigenvalue of $\lambda = n^{2}$. We know this because our function $T$ must be periodic on $2\pi$, and as such we get the following piecewise function:
-\[
-T(\theta) =
-\begin{cases*}
-a_{0}\qquad\qquad\qquad\qquad\qquad\quad\ \text{for}\ n=0\\
-a_{n}\cos{(n\theta)} + b_{n}\sin{(n\theta)}\qquad \text{for}\ n > 0
-\end{cases*}
-\]
-The solution to our equation for $R$ yields
-\[
-R(r) =
-\begin{cases*}
-c_{1} + c_{2}\ln{r} \qquad\qquad \text{for}\ n=0\\
-c_{1}r^{n} + c_{2}r^{-n}\qquad\quad \text{for}\ n > 0
-\end{cases*}
-\]
-We can disregard $c_{2}$ for this equation, since it would cause our solution to go to $\infty$ as $r \to 0$, at the center of the disk. Putting everything together gives us
-\begin{align*}
-u(r,\theta) &= R(r)T(\theta)\\
-&=
-\begin{cases*}
-a_{0}\qquad\qquad\qquad\qquad\qquad\qquad\quad\ \text{for}\ n=0\\
-a_{n}r^{n}\cos{(n\theta)} + b_{n}r^{n}\sin{(n\theta)}\qquad \text{for}\ n > 0
-\end{cases*}
-\end{align*}
-We should probably construct a series out of this, since that is what we usually do. The corresponding series is
-\[
-u(r,\theta) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}a_{n}r^{n}\cos{(n\theta)} + b_{n}r^{n}\sin{(n\theta)}
-\]
-And applying the boundary conditions gives us
-\[
-u(3,\theta) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}a_{n}3^{n}\cos{(n\theta)} + b_{n}3^{n}\sin{(n\theta)}
-\]
-For $a_{0}$, we have
-\begin{align*}
-a_{0} &= \frac{1}{\pi}\left[\int_{0}^{\pi}1 d\theta + \int_{\pi}^{2\pi}sin^{2}{\theta}d\theta\right]\\
-&= \frac{3}{2}
-\end{align*}
-Solving for $a_{n}$ gives us
-\begin{align*}
-a_{n} &= \frac{1}{3^{n}\pi}\left[\int_{0}^{\pi}\cos{(n\theta)}d\theta + \int_{\pi}^{2\pi}\cos{(n\theta)}\sin^{2}{\theta}d\theta\right]\\
-&= \frac{[n^{2} - 2 - 4\cos{(n\pi)}]\sin{(n\pi)}}{3^{n}n(n^{2}-4)\pi}\\
-&= 0
-\end{align*}
-Our $b_{n}$ term is
-\begin{align*}
-b_{n} &= \frac{1}{3^{n}\pi}\left[\int_{0}^{\pi}\sin{(n\theta)}d\theta + \int_{\pi}^{2\pi}\sin{(n\theta)}\sin^{2}{\theta}d\theta\right]\\
-&= \frac{2\left[n^{2} - 6 - 4\cos{(n\pi)}\right]\sin^{2}{\left(\frac{n\pi}{2}\right)}}{3^{n}n\pi(n^{2} - 4)}\\
-&= \frac{2\left[n^{2} - 6 - 4(-1)^{n}\right]\sin^{2}{\left(\frac{n\pi}{2}\right)}}{3^{n}n\pi(n^{2} - 4)}\\
-&=
-\begin{cases*}
-0 \qquad\qquad\qquad\qquad\quad\ \ n \in \{ 2k\ |\ k \in \mathbb{Z} \}\\
-\frac{1}{3^{n}\pi}\left[\frac{2}{n} + \frac{4}{n(n^{2} - 4)}\right]\qquad n \in \{ 2k + 1\ |\ k \in \mathbb{Z} \}
-\end{cases*}
-\end{align*}
-We essentially have our solution. All we have to do is combine everything:
-\[
-u(r,\theta) = \frac{3}{4} + \sum_{k=0}^{\infty}\frac{r^{2k + 1}}{3^{2k+1}\pi}\frac{2(4k^{2} + 4k -1)}{(2k + 3)(4k^{2} -1)}\sin{\left[(2k +1)\theta\right]}
-\]
-Let's take a look at some of the plots for this equation.\\
-\begin{center}
-\begin{tabular}{c c c}
-\includegraphics[scale=0.3]{lap_01_3} & \includegraphics[scale=0.3]{lap_01_30} & \includegraphics[scale=0.3]{lap_01_50}\\
-\textit{The first 3 terms} & \textit{The first 30 terms} & \textit{The first 50 terms}
-\end{tabular}
-\end{center}
-Plugging this infinite sum into mathematica tells us that it does not converge; the values are heading towards $\infty$ as the sum grows. We will leave our answer as a series, and we are done.
-
-\newpage
-\subsection{Wrapping Up}
-\indent Before we move onto the next topic, we will present some common patterns amongst the second order PDEs.
-Below, you will find each of the three equations in the classical trinity, and their corresponding solutions based on the boundary conditions.
-\subsubsection{The Heat Equation}
-\[
-\begin{array}{c c}
-\begin{cases}
-u_{t} - ku_{xx} = 0\\
-u(0,t) = u(L,t) = 0\\
-u(x,0) = f(x)
-\end{cases}
-&
-\begin{cases}
-u_{t} - ku_{xx} = 0\\
-u_{x}(0,t) = u_{x}(L,t) = 0\\
-u(x,0) = g(x)
-\end{cases}\\
-&\\
-u(x,t) = \sum_{n=1}^{\infty}b_{n}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}\sin{\left(\frac{n\pi x}{L}\right)}
-& u(x,t) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}a_{n}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}\cos{\left(\frac{n\pi x}{L}\right)}\\
-&\\
-\begin{cases}
-u_{t} - ku_{xx} = 0\\
-u(0,t) = u_{x}(L,t) = 0\\
-u(x,0) = h(x)
-\end{cases}
-&
-\begin{cases}
-u_{t} - ku_{xx} = 0\\
-u_{x}(0,t) = u(L,t) = 0\\
-u(x,0) = k(x)
-\end{cases}\\
-&\\
-u(x,t) =  \sum_{n=1}^{\infty}a_{n}e^{-k\left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}\right)^{2}t}\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}
-& u(x,t) =  \sum_{n=1}^{\infty}a_{n}e^{-k\left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}\right)^{2}t}\cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}
-\end{array}
-\]
-\noindent\\
-\subsubsection{The Wave Equation}
-\[
-\begin{array}{c c}
-\begin{cases}
-u_{tt} - c^{2}u_{xx} = 0\\
-u(0,t) = u(L,t) = 0\\
-u(x,0) = f(x) \quad u_{t}(x,0) = g(x)
-\end{cases}
-&
-\begin{cases}
-u_{tt} - c^{2}u_{xx} = 0\\
-u_{x}(0,t) = u_{x}(L,t) = 0\\
-u(x,0) = f(x) \quad u_{t}(x,0) = g(x)
-\end{cases}\\
-&\\
-u(x,t) = \sum_{n=1}^{\infty}\left[a_{n}\cos{\left(\frac{n \pi t c}{L}\right)} + b_{n}\sin{\left(\frac{n \pi t c}{L}\right)}\right]\sin{\left(\frac{n \pi x}{L}\right)}
-&u(x,t) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}\left[a_{n}\cos{\left(\frac{n \pi t c}{L}\right)} + b_{n}\sin{\left(\frac{n \pi t c}{L}\right)}\right]\cos{\left(\frac{n \pi x}{L}\right)}\\
-&\\
-\end{array}
-\]
-\[
-\begin{array}{c}
-\begin{cases}
-u_{tt} - c^{2}u_{xx} = 0\\
-u(0,t) = u_{x}(L,t) = 0\\
-u(x,0) = f(x) \quad u_{t}(x,0) = g(x)
-\end{cases}\\
-\\
-u(x,t) = a_{0} + b_{0}t + \sum_{n=1}^{\infty}\left[ a_{n}\cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi t}{L}\right)} + b_{n}\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}\right]\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}\\
-\\
-\begin{cases}
-u_{tt} - c^{2}u_{xx} = 0\\
-u_{x}(0,t) = u(L,t) = 0\\
-u(x,0) = f(x) \quad u_{t}(x,0) = g(x)
-\end{cases}\\
-\\
-u (x,t) = a_{0} + b_{0}t + \sum_{n=1}^{\infty}\left[ a_{n}\cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi t}{L}\right)} + b_{n}\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}\right]\cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi x}{L}\right)}
-\end{array}
-\]
-\subsubsection{The Laplace Equation}
-\paragraph{Cartesian Coordinates}
-\[
-\begin{array}{c c}
-\begin{cases}
-u_{xx} + u_{yy} = 0\\
-u(0,y) = u(L,y) = 0
-\end{cases}
-&
-\begin{cases}
-u_{xx} + u_{yy} = 0\\
-u_{x}(0,y) = u_{x}(L,y) = 0
-\end{cases}\\
-&\\
-u(x,y) = \sum_{n=1}^{\infty}\left[a_{n}\sinh{\left(\frac{n \pi y}{L}\right)} + b_{n}\cosh{\left(\frac{n\pi y}{L}\right)}\right]\sin{\left(\frac{n\pi x}{L}\right)}
-& u(x,y) = \sum_{n=1}^{\infty}\left[a_{n}\sinh{\left(\frac{n \pi y}{L}\right)} + b_{n}\cosh{\left(\frac{n\pi y}{L}\right)}\right]\cos{\left(\frac{n\pi x}{L}\right)}
-\end{array}
-\]
-\\
-\[
-\begin{array}{c}
-\begin{cases}
-u_{xx} + u_{yy} = 0\\
-u_{x}(0,y) = u(L,y) = 0
-\end{cases}\\
-\\
-u(x,t) = \sum_{n=1}^{\infty}\left[a_{n}\sinh{\left(\left[n + \frac{1}{2}\right]\frac{\pi y}{L}\right)} + b_{n}\cosh{\left(\left[n + \frac{1}{2}\right]\frac{\pi y}{L}\right)}\right]\cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi y}{L}\right)}\\
-\\
-\begin{cases}
-u_{xx} + u_{yy} = 0\\
-u(0,y) = u_{x}(L,y) = 0
-\end{cases}\\
-\\
-u(x,t) = \sum_{n=1}^{\infty}\left[a_{n}\sinh{\left(\left[n + \frac{1}{2}\right]\frac{\pi y}{L}\right)} + b_{n}\cosh{\left(\left[n + \frac{1}{2}\right]\frac{\pi y}{L}\right)}\right]\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi y}{L}\right)}
-\end{array}
-\]
-\paragraph{Polar Coordinates}
-\begin{gather*}
-\begin{cases}
-u_{rr} + \frac{1}{r}u_{r} + \frac{1}{r^{2}}u_{\theta\theta} = 0\\
-u(h,\theta) = f(\theta)\quad \text{on}\ x < \theta < L
-\end{cases}\\
-u(r,\theta) = a_{0} + \sum_{n=1}^{\infty}a_{n}r^{n}\cos{\left(\frac{n \pi \theta}{L}\right)} + b_{n}r^{n}\sin{\left(\frac{n \pi \theta}{L}\right)}
-\end{gather*}
-\newpage
-
-\section{The Cauchy Problem}
-\hrule
-\noindent \\\\
-\indent Up until now, we have always been given our PDE's with boundary conditions and initial values; but imagine that we did not have a boundary?
-The traditional method that we have been using would fail at this point.
-Since we would not be able to find any eigenvalues, we would not be able to form a series solution by using the \textit{Separation of Variables} technique.
-Let's consider the familiar heat equation:
-\[
-\begin{cases}
-\xi_{t} - k\xi_{xx} = 0\\
-\xi(x,0) = f(x)
-\end{cases}
-\]
-Through much careful derivation and theory, we can say that the general solution for this problem will be
-\[
-\xi(x,t) = \frac{1}{\sqrt{4 \pi k t}}\int_{-\infty}^{\infty}e^{-\frac{(x-y)^{2}}{4kt}}f(y)dy.
-\]
-The question now should be ``How can we solve this?'' The first step we will take is letting
-\[
-u = \frac{y-x}{\sqrt{4kt}}.
-\]
-We are going to perform a \textit{u substitution} to try and put this integral into a nicer form. This yields
-\begin{align*}
-y &= x + u\sqrt{4kt}\\
-dy &= \sqrt{4kt}\ du.
-\end{align*}
-This changes our original problem into
-\[
-\xi(x,t) = \frac{1}{\sqrt{\pi}}\int_{-\infty}^{\infty}e^{-u^{2}}f(x+u\sqrt{4kt})du.
-\]
-Let's look at an example\\\\
-\noindent \textbf{\textit{Ex:}} Solve the following PDE
-\[
-\begin{cases}
-\mu_{t} - k\mu_{xx} = 0\\
-\mu(x,0) = \cos{x}
-\end{cases}
-\]
-\indent \textbf{\textit{Solution:}} From above, we know what our general solution will look like. All we need to do is substitute in our initial condition. We have the following:
-\begin{align*}
-\mu(x,t) &= \frac{1}{\sqrt{\pi}}\int_{-\infty}^{\infty}e^{-u^{2}}\cos{(x + u\sqrt{4kt})}du\\
-&= \frac{1}{\sqrt{\pi}}\int_{-\infty}^{\infty}e^{-u^{2}}\left[\cos{(x)}\cos{(u\sqrt{4kt})} - \sin{(x)}\sin{(u\sqrt{4kt})}\right]du\\
-&=  \frac{2}{\sqrt{\pi}}\cos{(x)}\int_{0}^{\infty}e^{-u^{2}}\cos{(u\sqrt{4kt})}du.
-\end{align*}
-You may be wondering what happened to the $\sin$ terms.
-Recall that $\sin$ is an odd function, so it will ultimately be $0$ on the interval $[-\infty,\infty]$; because of this fact, we can remove it now.
-We also use the face that the $\cos$ is even, so it is okay to halve the interval and multiply by two.
-Now we need to evaluate this integral.
-We will substitute $\alpha = \sqrt{4kt}$, and assign a function, $\Upsilon(\alpha)$, to the integral.
-This gives us the following system:
-\[
-\begin{cases}
-\Upsilon(\alpha) = \int_{0}^{\infty}e^{-u^{2}}\cos{(\alpha u)}du\\
-\Upsilon(0) = \int_{0}^{\infty}e^{-u^{2}}du\\
-\end{cases}
-\]
-Believe it or not, we actually know the answer to $\Upsilon(0)$. To find it, we can assign a function to the integral in a similar fashion as we have already done. Calling it $\Psi$, we have
-\begin{align*}
-\Psi &= \int_{0}^{\infty}e^{-u^{2}}du\\
-\Psi^{2} &= \int_{0}^{\infty}e^{-u^{2}}du \int_{0}^{\infty}e^{-v^{2}}dv\\
-\Psi^{2} &= \int_{0}^{\infty}\int_{0}^{\infty}e^{-u^{2}-v^{2}}dudv\\
-\Psi^{2} &= \int_{0}^{\pi/2}\int_{0}^{\infty}re^{-r^{2}}drd\theta\\
-\Psi^{2} &= \frac{\pi}{2}\left[-\frac{1}{2}e^{-r^{2}}\right]\Bigg |_{r=0}^{r=\infty}\\
-\Psi^{2} &= \frac{\pi}{4}\\
-\Psi &= \frac{\sqrt{\pi}}{2}.
-\end{align*}
-So we know that $\Upsilon(0) = \frac{\sqrt{\pi}}{2}$. Plugging this into our system gives us
-\[
-\begin{cases}
-\Upsilon(\alpha) = \int_{0}^{\infty}e^{-u^{2}}\cos{(\alpha u)}du\\
-\Upsilon(0) = \frac{\sqrt{\pi}}{2}\\
-\end{cases}
-\]
-Now, we will take the derivative of $\Upsilon$. This gives us
-\[
-\Upsilon'(\alpha) = -\int_{0}^{\infty}ue^{-u^{2}}\sin{(\alpha u)}du.
-\]
-Integration by parts gives us
-\[
-\frac{1}{2}e^{-u^{2}}\sin{(\alpha u)}\Big |_{0}^{\infty} - \alpha\int_{0}^{\infty}\frac{1}{2}e^{-u^{2}}\cos{(\alpha u)}du.
-\]
-The value from $0$ to $\infty$ is zero, since the $\sin$ function odd. This leaves us with following relationship:
-\[
-\Upsilon'(\alpha) = -\frac{\alpha}{2}\Upsilon(\alpha) \alpha.
-\]
-We can solve this as an ODE. It's separable, so we get
-\begin{align*}
-\frac{d\Upsilon}{\Upsilon} &= -\frac{\alpha}{2}d\alpha\\
-\ln{(\Upsilon)} &= -\frac{\alpha^{2}}{4} + c_{1}\\
-\Upsilon &= c_{2}e^{-\frac{\alpha^{2}}{4}}.
-\end{align*}
-Applying the initial condition gives us
-\[
-\Upsilon = \frac{\sqrt{\pi}}{2}e^{-\frac{\alpha^{2}}{4}}.
-\]
-Substituting back in gives us a final answer of
-\begin{align*}
-\mu(x,t) &= \frac{2}{\sqrt{\pi}}\cos{x}\Upsilon(\sqrt{4kt})\\
-&= \cos{x}e^{-kt},
-\end{align*}
-and we are done.
-
-\newpage
-\indent Now we will move on to the Cauchy wave equation. We will be presented with a similar problem: a wave equation on an unbounded domain. Consider the following problem:
-\noindent \\\\
-\noindent \textbf{\textit{Ex:}} Solve the following PDE
-\[
-\begin{cases}
-u_{tt} - c^{2}u_{xx} = 0\\
-u(x,0) = \sin{x}\\
-u_{t}(x,0) = \cos{x}
-\end{cases}
-\]
-\indent \textbf{\textit{Solution:}} Just as with the heat equation, we have a formal solution for the homogeneous wave equation.
-We will do the derivation, but I will tell you that the formal solution is
-\[
-u(x,t) = \frac{1}{2}\left[f(x+ct) + f(x-ct)\right] + \frac{1}{2c}\int_{x-ct}^{x+ct}g(s)\mathop{ds}.
-\]
-Now let's derive the formal solution. We want a solution in the form of
-\[
-u(x,t) = F(x+ct) + G(x-ct).
-\]
-Now we can say that $u_{t} = cF'(x+ct) - cG'(x-ct)$, and $u_{t}(x,0) = cF'(x) - cG'(x)$. This means
-\begin{align*}
-F(x) + G(x) &= f(x)\\
-F(x) - G(x) &= \frac{1}{c}\int_{0}^{x}g(s)\mathop{ds} + c_{1}.
-\end{align*}
-Solving the system yields
-\begin{align*}
-u(x,t) &= \frac{1}{2} f(x+ct) + \frac{1}{2c}\int_{0}^{x+ct}g(s)\mathop{ds} + \frac{c_{1}}{2} + \frac{1}{2}f(x-ct) -\frac{1}{2c}\int_{0}^{x-ct}g(s)\mathop{ds} - \frac{c_{1}}{2}\\
-&= \frac{1}{2}\left[f(x+ct) + f(x-ct)\right] + \frac{1}{2c}\int_{x-ct}^{x+ct}g(s)\mathop{ds},
-\end{align*}
-and that is the derivation. This problem is simply plugging into the formal solution. We have the following:
-\begin{align*}
-u(x,t) &= \frac{1}{2}\left[\sin{(x+ct)} + \sin{(x-ct)}\right] + \frac{1}{2c}\int_{x-ct}^{x+ct}\cos{(s)}\mathop{ds}\\
-&= \frac{\sin{(x+ct)} + \sin{(x-ct)}}{2} + \frac{\cos{(x)}\sin{(ct)}}{c}\\
-&=  \frac{\cos{(x)}\sin{(ct)}}{c} + \cos{(ct)}\sin{(x)},
-\end{align*}
-and we are done.
-
-\newpage
-\indent So far, we have only dealt with homogeneous Cauchy problems.
-We will now look at the case where we have a non-homogeneous problem.
-The first variation of this problem is the \textit{half-line} problem.
-In this problem, we will have a bound on our equation, and we will generalize the solution so that we are dealing with an unbounded Cauchy problem.
-Let's look at an example.
-\noindent\\\\ \textbf{\textit{Ex:}} Solve the following PDE
-\[
-\begin{cases}
-u_{t} - ku_{xx} = 0\\
-u(x,0) = \phi(x)\\
-u(0,t) = 0
-\end{cases}
-\]
-\indent \textbf{\textit{Solution:}} We have both an initial condition and a boundary condition here.
-This problem is saying that at $x=0$, there is an impassable boundary.
-This forces our domain to be $[0,\infty)$.
-To solve a Cauchy problem, however, we want a domain of $(-\infty,\infty)$.
-We can force this domain by extending our initial condition, $\phi(x)$, over the boundary.
-We have the following rule:
-\[
-\phi_{0}(x) =
-\begin{cases}
-\phi(x)\quad\quad x>0\\
--\phi(-x)\ x<0.
-\end{cases}
-\]
-Now we are dealing with the system
-\[
-\begin{cases}
-u_{t} - ku_{xx} = 0\\
-u(x,0) = \phi_{0}(x),
-\end{cases}
-\]
-which we can solve using the formal solution to the heat equation. Recall that
-\[
-u(x,t) = \frac{1}{\sqrt{4k\pi t}}\int_{-\infty}^{\infty}e^{-\frac{(x-y)^{2}}{4kt}}\phi_{0}(y)\mathop{dy}.
-\]
-Substituting into the equation gives us
-\begin{align*}
-u(x,t) &= \frac{1}{\sqrt{4k\pi t}}\left[\int_{0}^{\infty}e^{-\frac{(x-y)^{2}}{4kt}}\phi(y)\mathop{dy} + \int_{-\infty}^{0}e^{-\frac{(x-y)^{2}}{4kt}}(-\phi(-y))\mathop{dy} \right]\\
-&= \frac{1}{\sqrt{4k\pi t}}
-\end{align*}
+\include{mathe_code}
 \end{document}

### 2015-12-22 pde/transforms.tex
update pde

\section{Transforms}
\hrule
\noindent \\\\
\indent Much like when we solve ODE's, we can use transforms to solve PDE's. When we use a Laplace
transform to solve an ODE, we are transforming the ODE into an algebraic equation that is much
easier to solve. When we transform a PDE, we get an ODE that we can solve using techniques that
we already know. In this section, we will look at two transforms: the Laplace Transform and the
Fourier Transform.\\
\subsection{The Laplace Transform}
\indent Let's begin by looking at an exaple use of a Laplace Transform to solve an ODE.\\\\
\noindent \textbf{\textit{Ex:}} Solve the following ODE using a Laplace Transform.
\[
\begin{cases*}
y'' - 5y' + 6y = 0\\
y(0) = 2\\
y'(0) = 2
\end{cases*}
\]
\indent \textbf{\textit{Solution:}} We should begin by recalling the definition of the Laplace operator:
\[
\Lap[f(t)] =\int_{0}^{\infty}f(t)e^{-st}dt
\]
We also know that the Laplace transforms of a derivative and a second derivative:
\[
\Lap[f'(t)] = s\Lap[f(t)] - f(0)\qquad\qquad \Lap[f''(t)] = s^{2}\Lap[f(t)]-sf(0)-f'(0)
\]
This is really all the information we need. Applying the transform to both sides of the equation yields
\begin{gather*}
\Lap[y'' - 5y' + 6y ] = \Lap[0]\\
s^{2}\gamma(s)-sy(0)-y'(0) - s\gamma(s)-y(0) = 0\\
s^{2}\gamma(s)-2s-2-s\gamma(s)-2 = 0\\
s^{2}\gamma(s)-s\gamma(s)-2s-4 = 0\\
\gamma(s) = \frac{2s - 8}{s^{2} -5s + 6}\\
\gamma(s) = \frac{4}{s-2}-\frac{2}{s-3}
\end{gather*}
We now need to transform our equation back to its ODE form; we do in fact have an inverse Laplace
operator, but it involves complex variables. If we consult a table of transforms, we will see that
$\gamma(s)$ becomes
\[
y(t) = 4e^{2t} +- 2e^{3t}
\]
And we are done.
\newpage

\indent Now we want to apply the Laplace transform to a PDE. The idea is the same, but our PDE involves
two variables: usually $x$, representative of position and $t$, representative of time. The Laplace
transform \textit{only} transforms with respect to our $t$ variable. Let's take a look at an
example:\\\\
\noindent \textbf{\textit{Ex:}} Solve the following PDE using a Laplace Transform.
\[
\begin{cases*}
u_{x} + u_{t} = x\\
u(0,t) = 0\\
u(x,0) = 0
\end{cases*}
\]
\indent \textbf{\textit{Solution:}} In this problem we are dealing with a first order PDE. Much
like an ODE, we simply need to determine what the Laplace transforms of $u_{x}$ and $u_{t}$ are.
We will let $\Lap[u(x,t)] = \Upsilon(x,s)$. Let's begin with $u_{t}(x,t)$.
\begin{align*}
\Lap[u_{t}] &= \int_{0}^{\infty}e^{-st}u_{t}(x,t)dt\\
&= e^{-st}u(x,t)\bigg|_{0}^{\infty} + s\int_{0}^{\infty}e^{-st}u(x,t)dt\\
&= s\Upsilon(x,s)-u(x,0)
\end{align*}
In a similar fashion, we could find $\Lap[u_{tt}(x,t)]$. We won't derive it here, but it ends up
looking very similar to our second derivative transform for our ODE:
\[
\Lap[u_{tt}(x,t)] = s^{2}\Upsilon(x,s)-su(x,0)-u_{t}(x,0)
\]
We also need to determine the transform for $u_{x}(x,t)$. Recall that the Laplace transform only
deals with the time variable, so our transforms for both $u_{x}(x,t)$ and $u_{xx}(x,t)$ become
$\Upsilon_{x}(x,s)$ and $\Upsilon_{xx}(x,s)$, respectivly. We now have all the information we need
to continue. Applying our transform yields
\begin{gather*}
\Lap[u_{x} + u_{t}] = \Lap[x]\\
\Upsilon'(x,s) + s\Upsilon(x,s)-u(x,0) = \frac{x}{s}\\
\Upsilon'(x,s) + s\Upsilon(x,s) =\frac{x}{s}
\end{gather*}
This is just a first order ODE. The only thing we are missing now is the initial condition, which
we can determine by transforming the initial condition on the PDE. In this case, $u(0,t) = 0$, so
the initial condition is still 0 after applying the transform. Now we have
\[
\begin{cases*}
\Upsilon'(x,s) + s\Upsilon(x,s) =\frac{x}{s}\\
\Upsilon(0,s) = 0
\end{cases*}
\]
We can solve this ODE by using the integrating factor method. The integrating factor in this case
is
\begin{align*}
\rho &= e^{\int s\mathop{dx}}\\
&= e^{sx}
\end{align*}
Our ODE now becomes
\[
\frac{d}{dx}(e^{sx}\Upsilon(x,t)) = \frac{e^{xs}x}{s}
\]
Integration, simplification, and applying our initial condition yields
\[
\Upsilon(x,t) = \frac{x}{s^{2}}-\frac{1}{s^{3}} + \frac{e^{-sx}}{s^{3}}
\]
Again, we won't dive into the complexity of the inverse Laplace transform here, but if we do take
the inverse, we get
\[
u(x,t) = xt - \frac{t^{2}}{2} - H(t-x)\frac{(t-x)^{2}}{2}
\]
The function $H(t-x)$ is the Heavyside Step Function, and is defined as
\[
H(x) =
\begin{cases*}
0,\quad x<0\\
1,\quad x\geq 0
\end{cases*}
\]
And we are done.
\newpage

\subsection{The Fourier Transform}
\indent The Fourier transform also allows us to take a PDE and transform it into an ODE. In this case,
however, we will be transforming with respect to the position variable. There are several variations to
the Fourier transform; we will look at the regular Fourier transform, the Fourier Sine transform, and the
Fourier Cosine transform.
\subsubsection{The Full Fourier Transfrom}
The Fourier transform of a function $f(x)$ is given by
\[
\Four[f(x)] = \frac{1}{2\pi}\int_{-\infty}^{\infty}e^{-i\omega x}f(x)\mathop{dx}
\]
Just like the Laplace transform, we will need to determine the Fourier transform for the derivatives
of functions as well. Recall that with the Laplace transfrom, $\Lap[u_{x}]$ became $\Upsilon_{x}$.
Using the same argument, we know that
\begin{gather*}
\Four[u_{t}(x,t)] = \Lambda_{t}(\omega,t)\\
\Four[u_{tt}(x,t)] = \Lambda_{tt}(\omega,t)
\end{gather*}
When we apply the Fourier transfrom to our $x$ derivatives, we get
\begin{gather*}
\Four[u_{x}(x,t)] = i\omega\Lambda(\omega,t)\\
\Four[u_{xx}(x,t)] = -\omega\Lambda(\omega,t)
\end{gather*}
Let's consider the following example:\\\\
\noindent \textbf{\textit{Ex:}} Solve the following PDE using a Fourier Transform.
\[
\begin{cases*}
2u_{x} + 3u_{t} = 0\\
u(x,0) = 3
\end{cases*}
\]
\indent \textbf{\textit{Solution:}} We are again dealing with a first order PDE here. We will
begin by taking the Fourier transform of both sides of the PDE. The Fourier transform of 0 is
still 0, so we get
\begin{gather*}
\Four[2u_{x} + 3u_{t}] = \Four[0]\\
2i\omega\Lambda(\omega,t) + 3\Lambda_{t}(\omega,t) = 0
\end{gather*}
This is just a first order ODE with respect to $t$. We also need to transform our inital condition.
Doing so yields
\[
\Four[3] = 3\delta(\omega)
\]
The delta function is defined as the derivative of the Heavyside Step Function. We now have an
ODE represented by
\[
\begin{cases*}
2i\omega\Lambda(\omega,t) + 3\Lambda_{t}(\omega,t) = 0\\
\Lambda(\omega,0) = 3\delta(\omega)
\end{cases*}
\]
Solving this ODE and applying the initial condition yields
\[
\Lambda(\omega,t) = 3\delta(\omega)e^{-\frac{2}{3}i\omega t}
\]
From this step, we would transform the equation back to the $x,t$ coordinate system. We won't do that
here, so we are done.
\newpage

\subsubsection{The Fourier Sine Transform}

### 2015-12-26 pde/high_order.tex
working on PDE

\section{High Order Partial Differential Equations}
\hrule
\noindent \\\\
This section is concerned with high order PDE's. We will be focusing on PDE's that are still linear,
but contain third, fourth, and higher derivatives. Before we continue, we will need to discuss Green's
theorem.

\subsection{Green's Theorem}
Green's theorem stems from the harmonic function that we know as Laplace's equation. We have seen the
operator $\nabla$, which by definition means
\[
\nabla f = <f_{x},f_{y},f_{z}>
\]
When we dealt with Laplace's equation, we saw the operator used as
\[
\nabla^{2}u = u_{xx} + u_{yy} = 0
\]
We ignored the $u_{zz}$ term because we were only concerned with the 2D plane. Green's function provides
a universal solution to harmonic equations in higher orders than the Laplace equation which we are used to
dealing with. Before we continue, we need to discuss Green's Identities.
\subsubsection{Green's First Identity}
Green's first identity is represented by the equation
\[
\int_{D}\nabla\phi\cdot\nabla\psi\mathop{dV} + \int_{D}\phi\nabla^{2}\psi\mathop{dV} =
\int_{C}\phi\nabla\psi\cdot\mathbf{n}\mathop{dS}
\]
This result follows from the \textit{Divergence Theorem} and the product rule for partial derivatives.
We have $D$, which is the closed surface over which we are integrating, and two function $\phi$ and
$\psi$. Green's first identity provides us with a way to relate $n$ integrals to $n - 1$ integrals of the
same function (it can be extended to show that the identity works for $n$ and $n-1$ integrals).
% We can consider the Neumann style problem as an application of Green's first identity. If we let
% $\psi = 1$, Green's first identity becomes
% \[
% \iint_{\partial D}\frac{\partial\phi}{\partial n}\mathop{dS} = \iiint_{D}\nabla^{2}\phi\mathop{dx}
% \]
\subsubsection{Green's Second Identity}
We can derive Green's second identity from the first identity realtively simply. We simply need to
interchange $\psi$ and $\phi$, then subtract across. This yields
% G = \phi
% u = \psi
% dA = dV
\[
\int_{D} \psi\nabla^{2}\phi\mathop{dV} - \int_{D}\phi\nabla^{2}\psi\mathop{dV} =
\int_{C}\psi\nabla\phi\cdot\mathbf{n}\mathop{dS} - \int_{C}\phi\nabla\psi\cdot\mathbf{n}\mathop{dS}
\]
Again, this identity relates an $n$ dimensional surface integral to an $n-1$ dimesional surface integral
of the same functions.
\subsubsection{The Delta Function}
In addition to Green's identities, we also need to discuss the delta function, $\delta(x)$. We have
seen this function before (recall $\Four[1]$). Looking at the function in depth, we have the following
definition:
\[
\delta(x)=
\begin{cases*}
0,\quad x\neq0\\
\infty\quad x=0
\end{cases*}
\]
We also have the following properties of the delta function:
\begin{gather*}
\int_{-\infty}^{\infty}\delta(x)\mathop{dx} = 1\\
\int_{-\infty}^{\infty}f(x)\delta(x-a)\mathop{dx} = f(a)
\end{gather*}
The second property, called the sifting property, is of particular interest, as we will see soon.
The delta function can also be extended to the 2D case:
\[
\delta(x,y) =
\begin{cases*}
0,\qquad(x,y)\neq0\\
\infty,\qquad(x,y)=0
\end{cases*}
\]
The sifting property from the 1D case becomes
\[
\iint f(x,y)\delta(x-a,y-b)\mathop{dA} = f(a,b)
\]
where $A$ is a surface over which we are integrating.
\subsubsection{Green's Functions}
Finally, we need to discuss the notion of a Green's Function. A Green Function is formally defined
as
\[
G(x,s) = \mathcal{L}^{-1}\delta(x-s)
\]
For some linear differential operator $\mathcal{L}$. If we take a function $f(x)$ and multiply and
integrate, we get the following identity:
\[
u(x) = \int G(x,s)f(s)\mathop{ds}
\]
Green's Functions can be used to solve many different types of equations;
For any general ODE, we require two constraints to be fulfilled:
\begin{enumerate}
\item A solution to the ODE exists, and
\item the solutions are linearly independant.
\end{enumerate}
Recall that we can check for linear independance of solution via the Wronskian, $\mathcal{W}$.
We can derive a Green's Function for a differential equation by
\[
G(x,s) =
\begin{cases*}
\frac{y_{1}(s)y_{2}(x)}{\mathcal{W}(y_{1},y_{2})(s)},\qquad a\leq s\leq x\leq b\\
\frac{y_{1}(x)y_{2}(s)}{\mathcal{W}(y_{1},y_{2})(s)},\qquad a\leq x\leq s\leq b
\end{cases*}
\]
which yields the solution
\[
y(x) = \int_{a}^{b}G(x,s)f(s)\mathop{ds}
\]
Let's consider an example.\\

\noindent\textbf{\textit{Ex:}} Solve the following Boundary Value Problem:
\[
\begin{cases*}
y''(x) = x^{2}\\
y(0) = 0\\
y(1) = 0
\end{cases*}
\]
\indent\textbf{\textit{Solution:}} To solve this problem, we need to first find the Green Function
associated with the problem. Let's consider the homogeneous ODE $y'' = 0$. This has the solution
\[
y(x) = c_{1} + c_{2}x
\]
We can take any arbitrary $c$ as long as our choice does not yield a trivial solution and it satisfies
the boundary conditions. Let's consider
\[
y_{1}(x) = x\qquad\qquad\text{and}\qquad\qquad y_{2}(x) = 1 - x
\]
We can see that this choice satisfies the boundary conditions, and that
\[
\left|
\begin{array}{c c}
0 & 1\\
1 & 1
\end{array}\right|
= -1 \neq 0
\]
which implies that the choice we made is linearly independant. We can now derive a Green's function
for our ODE as follows:
\begin{align*}
G(x,s) &=
\begin{cases*}
\frac{s(1-x)}{-1}\\
\frac{x(1-s)}{-1}
\end{cases*}\\
&=
\begin{cases*}
x-s\\
s-x
\end{cases*}
\end{align*}
We are bounded by $[0,1]$ in this case, so we really have
\[
G(x,s) =
\begin{cases*}
x-s,\qquad 0\leq s\leq x\\
s-x,\qquad x\leq s\leq 1
\end{cases*}
\]
We said earlier that the solution $y(x)$ can be found by
\begin{align*}
y(x) &= \int_{a}^{b}G(x,s)f(s)\mathop{ds}\\
&= \int_{0}^{x}(x-s)s^{2}\mathop{ds} + \int_{x}^{1}(s-x)(s^{2})\mathop{ds}\\
&= \frac{1}{12}(x^{x} - x)
\end{align*}
And we are done.
\newpage

Green's Functions are useful for more than ODE's. We want to extend the idea to PDE's, which can
be doe simply enough. Before we continue, we will mention Poisson's Equation, which is a
generalized form of Laplace's Equation:
\[
\nabla^{2}\psi = -4\pi\rho
\]
If we take $\rho = 0$, we get back Laplace's Equation. Let's consider the following example:\\

\noindent\textbf{\textit{Ex:}} Solve the Poisson equation
\[
\begin{cases*}
\nabla^{2}u = -f(x,y)\qquad \text{on }\Omega = \{(x,y)|0<x<\pi,0<y<\pi\}\\
u = 0,\qquad\qquad\qquad\text{on } \partial\Omega
\end{cases*}
\]
\indent\textbf{\textit{Solution:}} We know the formal solution to this PDE:
\[
u(x,y) = \sum_{n=1}^{\infty}b_{n}(y)\sin{(nx)}
\]

### 2015-12-26 pde/transforms.tex
working on PDE

@@
 
-\subsubsection{The Fourier Sine Transform}
+\subsubsection{The Fourier Sine and Cosine Transform}
+Much like the Fourier series, the Fourier transform can also be split into a sine and cosine
+transform. The rules for the transformation still hold; we are still only concerned with the
+position variable $x$. The transform for the positional derivatives do change however. We have
+the following:
+\[
+\Four_{s}[f(x)] = \frac{2}{\pi}\int_{0}^{\infty}\sin{(\omega x)}f(x)\mathop{dx}
+\qquad\qquad\qquad
+\Four_{c}[f(x)] = \frac{2}{\pi}\int_{0}^{\infty}\cos{(\omega x)}f(x)\mathop{dx}
+\]
+Applying these definitions to our positional derivatives yields
+\begin{alignat*}{3}
+&\Four_{s}[u_{x}(x,t)] = -\omega\Four_{c}[u(x,t)] \qquad\qquad\qquad\qquad\qquad
+&&\Four_{c}[u_{x}(x,t)] = -\frac{2}{\pi}u(0,t)+\omega\Four_{s}[u(x,t)]\\
+&\Four_{s}[u_{xx}(x,t)] = \frac{2\omega u(0,t)}{\pi} + \omega^{2}\Four_{s}[u(x,t)]
+&&\Four_{c}[u_{xx}(x,t)] = -\frac{2}{\pi}u_{x}(0,t)-\omega^{2}\Four_{c}[u(x,t)]
+\end{alignat*}
+Notice that with the transform of the first derivative, we end up with the opposite transform of the
+original function. Let's look at an example.
+
+\noindent \textbf{\textit{Ex:}} Solve the following PDE using a Fourier Transform.
+\[
+% \begin{cases*}
+2u_{x} + 3u_{t} = 0
+% u(0,t) = 3
+% \end{cases*}
+\]
+\indent \textbf{\textit{Solution:}} We will start with the sine transform. Applying the sine transform to
+both sides of the equation yields
+\begin{gather*}
+2\Four_{s}[u_{x}] + 3\Four_{s}[u_{t}] = \Four_{s}[0]\\
+-2\omega\Lambda(\omega,t) + 3\Lambda_{t}(\omega,t) = 0\\
+-2\omega\Lambda + 3\Lambda' = 0
+\end{gather*}
+This is a first order ODE that we can easily solve. We get a final solution of
+\[
+\Lambda(\omega,t) = f(t)e^{-\frac{2}{3}\omega}
+\]
+And we are done.

### 2015-12-28 chil/texts/pippi_longstocking.tex
setting up things

\documentclass{standard}
\title{Pippi Longstocking}
\author{Astrid Lindgren}
\date{}

\begin{document}
\maketitle
\newpage

\section{Pippi Moves into Villa Villekulla}
\noindent\\

Way out at the end of a tiny little town was an old overgrown garden, and in the garden was an old house, and in the house lived Pippi Longstocking. She was nine years old, and she lived there all alone. She had no mother and no father, and that was of course very nice because there was no one to tell her to go to bed just when she was having the most fun, and no one who could make her take cod liver oil when she much preferred caramel candy.\\

Once upon a time Pippi had had a father of whom she was extremely fond. Naturally she had had a mother too, but that was so long ago that Pippi didn’t remember her at all. Her mother had died when Pippi was just a tiny baby and lay in a cradle and howled so that nobody could go anywhere near her. Pippi was sure that her mother was now up in Heaven, watching her little girl through a peephole in the sky, and Pippi often waved up at her and called, ``Don’t you worry about me. I’ll always come out on top.”\\

Pippi had not forgotten her father. He was a sea captain who sailed on the great ocean, and Pippi had sailed with him in his ship until one day her father was blown overboard in a storm and disappeared. But Pippi was absolutely certain that he would come back. She would never believe that he had drowned; she was sure he had floated until he landed on an island inhabited by cannibals. And she thought he had become the king of all the cannibals and went around with a golden crown on his head all day long.\\

``My papa is a cannibal king; it certainly isn’t every child who has such a stylish papa,” Pippi used to say with satisfaction. ``And as soon as my papa has built himself a boat he will come and get me, and I’ll be a cannibal princess. Heigh- ho, won’t that be exciting!”\\

Her father had bought the old house in the garden many years ago. He thought he would live there with Pippi when he grew old and couldn’t sail the seas any longer. And then this annoying thing had to happen, that he was blown into the ocean, and while Pippi was waiting for him to come back she went straight home to Villa Villekulla. That was the name of the house. It stood there ready and waiting for her. One lovely summer evening she had said good-by to all the sailors on her father’s boat. They were all fond of Pippi, and she of them.\\

``So long, boys.” she said and kissed each one on the forehead. ``Don’t you worry about me. I’ll always come out on top.”\\

Two things she took with her from the ship: a little monkey whose name was Mr. Nilsson--he was a present from her father--and a big suitcase full of gold pieces. The sailors stood upon the deck and watched as long as they could see her. She walked straight ahead without looking back at all, with Mr. Nilsson on her shoulder and her suitcase in her hand.\\

``A remarkable child,” said one of the sailors as Pippi disappeared in the distance.\\

He was right. Pippi was indeed a remarkable child. The most remarkable thing about her was that she was so strong. She was so very strong that in the whole wide world there was not a single police officer as strong as she. Why, she could lift a whole horse if she wanted to! And she wanted to. She had a horse of her own that she had bought with one of her many gold pieces the day she came home to Villa Villekulla. She had always longed for a horse, and now here he was, living on the porch. When Pippi wanted to drink her afternoon coffee there, she simply lifted him down into the garden.\\

Beside Villa Villekulla was another garden and another house. In that house lived a father and mother and two charming children, a boy and a girl. The boy’s name was Tommy and the girl’s Annika. They were good, well brought up, and obedient children. Tommy would never think of biting his nails, and he always did exactly what his mother told him to do. Annika never fussed when she didn’t get her own way, and she always looked pretty in her little well- ironed cotton dresses; she took the greatest care not to get than dirty. Tommy and Annika played nicely with each other in their garden, but they had often wished for a playmate. While Pippi was still sailing on the ocean with her father, they often used to hang over the fence and say to each other, ``Isn’t it silly that nobody ever moves into that house. Somebody ought to live there somebody with children.”\\

On that lovely summer evening when Pippi for the first time stepped over the threshold of Villa Villekulla, Tommy and Annika were not at home. They had gone to visit their grandmother for a week; and so they had no idea that anybody had moved into the house next door. On the first day after they came home again they stood by the gate, looking out onto the street, and even then they didn’t know that there actually was a playmate so near. Just as they were standing there considering what they should do and wondering whether anything exciting was likely to happen or whether it was going to be one of those dull days when they couldn’t think of anything to play--just then the gate of Villa Villekulla opened and a little girl stepped out. She was the most remarkable girl Tommy and Annika had ever seen. She was Miss Pippi Longstocking out for her morning promenade. This is the way she looked:\\

Her hair, the color of a carrot, was braided in two tight braids that stuck straight out. Her nose was the shape of a very small potato and was dotted all over with freckles. It must be admitted that the mouth under this nose was a very wide one, with strong white teeth. Her dress was rather unusual. Pippi herself had made it. She had meant it to be blue, but there wasn’t quite enough blue cloth, so Pippi had sewed little red pieces on it here and there. On her long thin legs she wore a pair of long stockings, one brown and the other black, and she had on a pair of black shoes that were exactly twice as long as her feet. These shoes her father had bought for her in South America so that Pippi would have some- thing to grow into, and she never wanted to wear any others.\\

But the thing that made Tommy and Annika open their eyes widest of all were the monkey sitting on the strange girl’s shoulder. It was a little monkey, dressed in blue pants, yellow jacket, and a white straw hat.\\

Pippi walked along the street with one foot on the sidewalk and the other in the gutter. Tommy and Annika watched as long as they could see her. In a little while she came back, and now she was walking backward. That was because she didn’t want to turn around to get home. When she reached Tommy’s and Annika’s gate she stopped.\\

The children looked at each other in silence. At last Tommy spoke. ``Why did you walk backward?”\\

``Why did I walk backwards?” said Pippi. ``Isn’t this a free country? Can’t a person walk any way she wants to? For that matter, let me tell you that in Egypt everybody walks that way, and nobody thinks it’s the least bit strange.”\\

``How do you know?” asked Tommy. ``You’ve never been in Egypt, have you?”\\

``I’ve never been in Egypt? Indeed I have. That’s one thing you can be sure of. I have been all over the world and seen many things stranger than people walking backward. I wonder what you would have said if I had come along walking on my hands the way they do in Farthest India.”\\

``Now you must be lying,” said Tommy.\\

Pippi thought a moment. ``You’re right,” she said sadly, ``I am lying.”\\

``It’s wicked to lie,” said Annika, who had at last gathered up enough courage to speak.\\

``Yes, it’s very wicked to lie,” said Pippi even more sadly. ``But I forget it now and then. And how can you expect a little child whose mother is an angel and whose father is king of a cannibal island and who herself has sailed on the ocean all her life -- how can you expect her to tell the truth always? And for that matter,” she continued, her whole freckled face lighting up, ``let me tell you that in the Congo there is not a single person who tells the truth. They lie all day long. Begin at seven in the morning and keep on until sundown. So if I should happen to lie now and then, you must try to excuse me and to remember that it is only because I stayed in the Congo a little too long. We can be friends anyway, can’t we?”\\

``Oh, sure,” said Tommy and realized suddenly that this was not going to be one of those dull days.\\

``By the way, why couldn’t you come and have breakfast with me?” asked Pippi.\\

``Why not?” said Tommy. ``Come on, let’s go.” ``Oh, yes, let’s,” said Annika.\\

``But first 1 must introduce you to Mr. Nilsson,” said Pippi, and the little monkey took off his cap and bowed politely.\\

Then they all went in through Villa Villekulla’s tumbledown garden gate, along the gravel path, bordered with old moss-covered trees--really good climbing trees they seemed to be--up to the house, and onto the porch. There stood the horse, munching oats out of a soup bowl.\\

``Why do you have a horse on the porch?” asked Tommy. All horses he knew lived in stables.\\

``Well,” said Pippi thoughtfully, “he’d be in the way in the kitchen, and he doesn’t like the parlor.\\

Tommy and Annika patted the horse and then went on into the house. It had a kitchen, a parlor, and a bedroom. But it certainly looked as if Pippi had forgotten to do her Friday cleaning that week. Tommy and Annika looked around cautiously just in case the king of the Cannibal Isles might be sitting in a corner somewhere. They had never seen a cannibal king in all their lives. But there was no father to be seen, nor any mother either.\\

Annika said anxiously, ``Do you live here all alone?”\\

``Of course not!” said Pippi. ``Mr. Nilsson and the horse live here too.”\\

``Yes, but I mean don’t you have any mother or father here?”\\

``No, not the least little tiny bit of a one,” said Pippi happily.\\

``But who tells you when to go to bed at night and things like that?” asked Annika.\\

``I tell myself,” said Pippi. First I tell myself in a nice friendly way; and then, if I don’t mind, I tell myself again more sharply; and if I still don’t mind, then I’m in for a spanking--see?”\\

Tommy and Annika didn’t see at all, but they thought maybe it was a good way. Meanwhile they had come out into the kitchen, and Pippi cried,\\

Now we’re going to make a pancake, Now there’s going to be a pancake. Now we’re going to fry a pancake.\\

Then she took three eggs and threw them up in the air. One fell down on her head and broke so that the yolk ran into her eyes, but the others she caught skillfully in a bowl, where they smashed to pieces.\\

``I always did hear that egg yolk was good far the hair,” said Pippi, wiping her eyes. ``You wait and see--mine will soon begin to grow so fast it will crackle. As a matter of fact, in Brazil all the people go about with eggs in their hair. And there are no bald-headed people. Only once was there a man who was so foolish that he ate his eggs instead of rubbing them on his hair. He became completely bald, and when he showed himself on the street there was such a riot that the police were called out.”\\

While she was speaking Pippi had neatly picked the eggshells out of the bowl with her fingers. Now she took a bath brush that hung on the wall and began to beat the pancake batter so hard that it splashed all over the walls. At last she poured what was left onto a griddle that stood on the stove.\\

When the pancake was brown on one side she tossed it halfway up to the ceiling, so that it turned right around in the air, and then she caught it on the griddle again. And when it was ready she threw it straight across the kitchen right onto a plate that stood on the table.\\

``Eat!” she cried. ``Eat before it gets cold!” And Tommy and Annika ate and thought it a very good pancake.\\

Afterward Pippi invited them to step into the parlor. There was only one piece of furniture in there. It was a huge chest with many tiny drawers. Pippi opened the drawers and showed Tommy and Annika all the treasures she kept there. There were wonderful birds’ eggs, strange shells and stones, pretty little boxes, lovely silver mirrors, pearl necklaces, and many other things that Pippi and her father had bought on their journeys around the world. Pippi gave each of her new playmates a little gift to remember her by. Tommy got a dagger with a shimmering mother-of-pearl handle and Annika, a little box with a cover decorated with pink shells. In the box there was a ring with a green stone.\\

``Suppose you go home now,” said Pippi, ``so that you can come back tomorrow. Because if you don’t go home you can’t come back, and that would be a shame.”\\

Tommy and Annika agreed that it would indeed. So they went home--past the horse, who had now eaten up all the oats, and out through the gate of Villa Villekulla. Mr. Nilsson waved his hat at them as they left.
\newpage

\section{Pippi is a Thing-Finder and Gets Into a Fight}

Annika woke up early the next morning. She jumped out of bed and ran over to Tommy.\\

``Wake up, Tommy,” she cried, pulling him by the arm. ``Wake up and let’s go and see that funny girl with the big shoes.”\\

Tommy was wide-awake in an instant.\\

``I knew, even while I was sleeping, that something exciting was going to happen today, but I didn’t remember what it was,” he said as he yanked off his pajama jacket. Off they went to the bathroom, washed themselves and brushed their teeth much faster than usual, had their clothes on in a twinkling, and a whole hour before their mother expected them came sliding down the banister and landed at the breakfast table. Down they sat and announced that they wanted their hot chocolate right at that very moment.\\

``What’s going to happen today that you’re in such a hurry?” asked their mother. ``We’re going to see the new girl next door,” said Tommy.\\

``We may stay all day,” said Annika.\\

That morning Pippi was busy making pepparkakor--a kind of Swedish cookie. She had made an enormous amount of dough and rolled it out on the kitchen floor.\\

``Because,” said Pippi to her little monkey, ``what earthly use is a baking board when one plans to make at least five hundred cookies?”\\

And there she lay on the floor, cutting out cookie hearts for dear life.\\

``Stop climbing around in the dough, Mr. Nilsson.” she said crossly just as the doorbell rang.\\

Pippi ran and opened the door. She was white as a miller from top to toe, and when she shook hands heartily with Tommy and Annika a whole cloud of flour blew over them.\\

``So nice you called,” she said and shook her apron - so there came another cloud of flour. Tommy and Annika got so much in their throats that they could not help coughing.\\

``What are you doing?” asked Tommy.

``Well, if I say that I’m sweeping the chimney, you won’t believe me, you’re so clever,” said Pippi. ``Fact is, I’m baking. But I’ll soon be done. You can sit on the wood-box for a while.”\\

Pippi could work fast, she could. Tommy and Annika sat and watched how she went through the dough, how she threw the cookies onto the cookie pans, and swung the pans into the oven. They thought it was as good as a circus.\\

``Done!” said Pippi at last and shut the oven door on the last pans with a bang. ``What are we going to do now?” asked Tommy.\\

``I don’t know what you are going to do,” said Pippi, ``but I know I can’t lie around and be lay. I am a Thing-Finder, and when you’re a Thing-Finder you don’t have a minute to spare.”\\

``What did you say you are?” asked Annika.\\

``A Thing-Finder.” ``What’s that?” asked Tommy.\\

``Somebody who hunts for things, naturally. What else could it be?” said Pippi as she swept all the flour left on the floor into a little pile.\\

``The whole world is full of things, and somebody has to look for them. And that’s just what a Thing-Finder does,” she finished.\\

``What kind of things?” asked Annika.\\

``Oh, all kinds,” said Pippi. ``Lumps of gold, ostrich feathers, dead rats, candy snap crackers, little tiny screws, and things like that.”\\

Tommy and Annika thought it sounded as if it would be fun and wanted very much to be Thing-Finders too, although Tommy did say he hoped he’d find a lump of gold and not a little tiny screw.\\

``We shall see what we shall see,” said Pippi. ``One always finds something. But we’ve got to hurry up and get going so that other Thing-Finders don’t pick up all the lumps of gold around here before we get them.”\\

All three Thing-Finders now set out. They decided that it would be best to begin hunting around the houses in the neighborhood, because Pippi said that although it could perfectly well happen that one might find a little screw deep in the woods, still the very best things were usually found where people were living.\\

``Though, for that matter,” she said, ``I’ve seen it the other way around too. I remember once when I was out hunting for things in the jungles of Borneo. Right in the heart of the forest, where no human being had ever before set foot, what do you suppose I found? Why, a very fine wooden leg! I gave it away later to a one-legged old man, and he said that a wooden leg like that wasn’t to be had for love or money.”\\

Tommy and Annika looked at Pippi to see just how a Thing-Finder acted. Pippi ran from one side of the road to the other, shaded her eyes with her hand, and hunted and hunted. Sometimes she crawled about on her hands and knees, stuck her hands in between the pickets of a fence, and then said in a disappointed tone, ``Oh, dear! I was sure I saw a lump of gold.”\\

``May we really take everything we find?” asked Annika.\\

``Yes, everything that is lying on the ground,” said Pippi.\\

Presently they came to an old man lying asleep on the lawn outside his cottage.\\

``There,” said Pippi, ``that man is lying on the ground and we have found him. Well take him!”\\

Tommy and Annika were utterly horrified.\\

``No, no, Pippi, we can’t take an old gentleman! We couldn’t possibly,” said Tommy. ``Anyway, whatever would we do with him?”\\

``What would we do with him? Oh, there are plenty of things we could do with him. We could keep him in a little rabbit hutch instead of a rabbit and feed him on dandelions. But if you don’t want to, I don’t care. Though it does bother me to think that some other Thing-Finder may come along and grab him.”\\

They went on. Suddenly Pippi gave a terrific yell. ``Well, I never saw the like,” she cried, as she picked up a large, rusty old tin can from the grass. ``What a find! What a find! Gang, that’s something you can never have too many of.”\\

Tommy looked at the can doubtfully. ``What can you use it for?”\\

``Oh, you can use it in all sorts of ways,” said Pippi. ``One way is to put cookies in it. Then it becomes a delightful Jar with Cookies. Another way is not to put cookies in it. Then it becomes a Jar without Cookies. That certainly isn’t quite so delightful, but still that’s good too.”\\

She examined the can, which was indeed rusty and had a hole in the bottom.\\

``It looks almost as if this were a jar without cookies,” she said thoughtfully. ``But you can put it over your head and pretend that it is midnight.”\\

And that is just what she did. With the can on her head she wandered around the block like a little metal tower and never stopped until she stumbled over a low wire fence and fell flat on her stomach. There was a big crash when the tin can hit the ground.\\

``Now, see that!” said Pippi and took off the can! If I hadn’t had this thing on me, I’d have fallen flat on my face and hurt myself terribly!”
``Yes,” said Annika, ``but if you hadn’t had the can on your head, then you wouldn’t have tripped on the wire fence in the first place.”\\

Before she had finished speaking there was another triumphant cry from Pippi, who was holding up an empty spool of thread.\\

``This seems to be my lucky day,” she said.\\

``Such a sweet, sweet little spool to blow soap bubbles with or to hang around my neck for a necklace. I’ll go home and make one this very minute.”\\

However, just at that moment the gate of one of the cottages nearby opened and a boy came rushing out. He looked scared, and that was no wonder, because head over heels after him came five other boys. They soon caught him and pushed him against the fence, and all five began to punch and hit him. He cried and held his arms in front of his face to protect himself.\\

``Give it to him! Give it to him!” cried the oldest and strongest of the boys, ``so that he’ll never dare to show himself on this street again.”\\

``Oh,” said Annika, ``it’s Willie they’re hurting. Oh, how can they be so mean?”\\

``It’s that awful Bengt. He’s always in a fight,” said Tommy. ``And five against one--what cowards!”\\

Pippi went up to the boys and tapped Bengt on the back with her forefinger. ``Hello, there,” she said. ``What’s the idea? Are you trying to make hash out of little Willie with all five of you jumping on him at once?”\\

Bengt turned around and saw a little girl he had never seen before: a wild- looking little stranger who dared to touch him. For a while he stood and gaped at her in astonishment; then a broad grin spread over his face. ``Boys,” he said, ``boys, let Willie alone and take a look at this girl. What a babe!”\\

He slapped his knees and laughed, and in an instant they had all docked around Pippi, all except Willie, who wiped away his tears and walked cautiously over to stand beside Tommy.\\

``Have you ever seen hair like hers? Red as fire! And such shoes,” Bengt continued. ``Can’t I borrow one?”\\

``I’d like to go out rowing and I haven’t any boat.” He took hold of one of Pippi’s braids but dropped it instantly and cried, ``Ouch, I burned myself.”\\

Then all five boys joined hands around Pippi, jumping up and down and screaming, ``Redhead! Redhead!”\\

Pippi stood in the middle of the ring and smiled in the friendliest way. Bengt had hoped she would get mad and begin to cry. At least she ought to have looked scared. When nothing happened he gave her a push.\\

``I don’t think you have a very nice way with ladies,” said Pippi. And she lifted him in her strong arms--high in the air--and carried him to a birch tree and hung him over a branch. Then she took the next boy and hung him over another branch. The next one she set on a gatepost out side a cottage, and the next she threw right over a fence so that he landed in a flowerbed. The last of the fighters she put in a tiny toy cart that stood by the road. Then Pippi, Tommy, Annika, and Willie stood and looked at the boys for a while. The boys were absolutely speechless with fright.\\

And Pippi said, ``You are cowards. Five of you attacking one boy! That’s cowardly. Then you begin to push a helpless little girl around. Oh, how mean!”\\

``Come now, well go home,” she said to Tommy and Annika. And to Willie, ``If they try to hurt you again, you come and tell me.” And to Bengt, who sat up in the tree and didn’t dare to stir, she said, ``is there anything else you have to say about my hair or my shoes? If so, you’d better say it now before I go home.”\\

But Bengt had nothing more to say about Pippi’s shoes or about her hair either. So Pippi took her can in one hand and her spool in the other and went away, followed by Tommy and Annika.\\

When they were back home in Pippi’s garden, Pippi said, ``Dear me, how awful! Here I found two beautiful things and you didn’t get anything. You must hunt a little more. Tommy, why don’t you look in that old hollow tree? Old trees are usually about the best places of all for Thing-Finders.”\\

Tommy said that he didn’t believe he and Annika would ever find anything, but to please Pippi he put his hand slowly down into the hollow tree trunk.\\

``Goodness!” he cried, utterly amazed, and pulled out his hand. In it he held a little notebook with a leather cover. In a special loop there was a little silver pencil.\\

``Well, that’s queer,” said Tommy.\\

``Now, see that!” said Pippi. ``There’s nothing so nice as being a Thing-Finder. It’s a wonder there aren’t more people that take it up. They’ll be tailors and shoemakers and chimney sweeps, and such like--but Thing-Finders, no indeed, that isn’t good enough for them!”\\

And then she said to Annika, ``Why don’t you feel in that old tree stump? One practically always finds things in old tree stumps.”\\

Annika stuck her hand down in the stump and almost immediately got hold of a red coral neck-lace. She and Tommy stood open-mouthed for a long time, they were so astonished. They thought that hereafter they would be Thing-Finders every single day.\\

Pippi had been up half the night before, playing ball, and now she suddenly felt sleepy. ``I think I’ll have to go and take a nap,” she said. ``Can’t you came with me and tuck me in!”\\

When Pippi was sitting on the edge of the bed, taking off her shoes, she looked at them thoughtfully and said, ``He was going out rowing, he said, that old Bengt.” She snorted disdainfully. ``I’ll teach him to row, indeed I will. Another time.”\\

``Say, Pippi,” said Tommy respectfully, ``why do you wear such big shoes?”\\

``So I can wiggle my toes, of course,” she answered. Then she crept into bed. She always slept with her feet on the pillow and her head way down under the quilt. ``That’s the way they sleep in Guatemala,” she announced. ``And it’s the only real way to sleep. See, like this, I can wiggle my toes when I’m sleeping too.”\\

``Can you go to sleep without a lullaby?” she went on. ``I always have to sing to myself for a while; otherwise I can’t sleep a wink.”\\

Tommy and Annika heard a humming sound under the quilt; it was Pippi singing herself to sleep. Quietly and cautiously they tiptoed out so that they would not disturb her. In the doorway they turned to take a last look toward the bed. They could see nothing of Pippi except her feet resting on the pillow. There she lay, wiggling her toes emphatically.\\

Tommy and Annika ran home. Annika held her coral necklace tightly in her hand.
``That certainly was strange,” she said. ``Tommy, you don’t suppose-do you suppose that Pippi had put these things in place beforehand?”\\

``You never can tell,” said Tommy. ``You just never can tell about anything when it comes to Pippi.”
\newpage

\section{Pippi Plays Tag With Some Policemen}
It soon became known throughout the little town that a nine-year-old girl was living all by herself in Villa Villekulla, and all the ladies and gentle-men in the town thought this would never do. All children must have someone to advise them, and all children must go to school to learn the multiplication tables. So the ladies and gentle-men decided that the little girl in Villa Villekulla must immediately be placed in a children’s home.\\

One lovely afternoon Pippi had invited Tommy and Annika over for afternoon coffee and pepparkakor. She had spread the party out on the front steps. It was so sunny and beautiful there, and the air was filled with the fragrance of the flowers in Pippi’s garden. Mr. Nilsson climbed around on the porch railing, and every now and then the horse stuck out his head so that he’d be invited to have a cookie.\\

“Oh, isn’t it glorious to be alive?” said Pippi, stretching out her legs as far as she could reach.\\

Just at that moment two police officers in full uniform came in through the gate.\\

“Hurray!” said Pippi. “This must be my lucky day too! Policemen are the very best things I know. Next to rhubarb pudding.” And with her face beaming she went to meet them.\\

“Is this the girl who has moved into Villa Villekulla?” asked one of the policemen.\\

“Quite the contrary,” said Pippi. “This is a tiny little auntie who lives on the third floor at the other end of the town.”\\

She said that only because she wanted to have a little fun with the policemen, but they didn’t think it was funny at all.\\

They said she shouldn’t be such a smarty. And then they went on to tell her that some nice people in the town were arranging for her to get into a children’s home.\\

“I already have a place in a children’s home,” said Pippi.\\

“What?” asked one of the policemen. “Has it been arranged already then? What children’s home?”\\

“This one,” said Pippi haughtily. “I am a child and this is my home: therefore it is a children’s home, and I have room enough here, plenty of room.”\\

“Dear child,” said the policeman, smiling, “you don’t understand. You must get into a real children’s home and have someone look after you.”\\

“Is one allowed to bring horses to your children’s home?” asked Pippi.\\

“No, of course not,” said the policeman. “That’s what I thought,” said Pippi sadly.\\

“Well, what about monkeys?”\\

“Of course not. You ought to realize that.”\\

“Well then,” said Pippi, “you’ll have to get kids for your children’s home somewhere else. I certainly don’t intend to move there.”\\

“But don’t you understand that you must go to school?”\\

“Why?”\\

“To learn things, of course.” “What sort of things?” asked Pippi.\\

“All sorts,” said the policeman. “Lots of useful things, the multiplication tables, for instance.”
“I have got along fine without any pluttifikation tables for nine years,” said Pippi, “and I guess I’ll get along without it from now on, too.”\\

“Yes, but just think how embarrassing it will be for you to be so ignorant. Imagine when you grow up and somebody asks you what the capital of Portugal is and you can’t answer!”\\

“Oh, I can answer all right,” said Pippi. “I’ll answer like this: ‘If you are so bound and determined to find out what the capital of Portugal is, then, for goodness’ sakes, write directly to Portugal and ask.”\\

“Yes, but don’t you think that you would be sorry not to know it yourself?”\\

“Oh, probably,” said Pippi. “No doubt I should lie awake nights and wonder and wonder, “What in the world is the capital of Portugal?” But one can’t be having fun all the time,” she continued, bending over and standing on her hands for a change. “For that matter, I’ve been in Lisbon with my papa,” she added, still standing upside down, for she could talk that way too.\\

But then one of the policemen said that Pippi certainly didn’t need to think she could do just as she pleased. She must come to the children’s home, and immediately. He went up to her and took hold of her arm, but Pippi freed herself quickly, touched him lightly, and said, “Tag!” Before he could wink an eye she had climbed up on the porch railing and from there onto the balcony above the porch. The policemen couldn’t quite see themselves getting up the same way, and so they rushed into the house and up the stairs, but by the time they had reached the balcony Pippi was halfway up the roof. She climbed up the shingles almost as if she were a little monkey herself. In a moment she was up on the ridgepole and from there jumped easily to the chimney. Down on the balcony stood the two policemen, scratching their heads, and on the lawn stood Tommy and Annika, staring at Pippi.\\

“Isn’t it fun to play tag!” cried Pippi. “And weren’t you nice to come over. It certainly is my lucky day today too.”\\

After the policemen had stood there a while wondering what to do, they went and got a ladder, leaned it against one of the gables of the house and then climbed up, first one policeman and then the other, to get Pippi down. They looked a little scared when they climbed out on the ridgepole and, carefully balancing themselves went step by step, toward Pippi.\\

“Don’t be scared,” cried Pippi. “There’s nothing to be afraid of. It’s just fun.”\\

When the policemen were a few steps away from Pippi, down she jumped from the chimney and, screeching and laughing, ran along the ridgepole to the opposite gable. A few feet from the house stood a tree.\\

“Now I’m going to dive,” she cried and jumped right down into the green crown of the tree, caught hold of a branch, swung back and forth a while, and then let herself fall to the ground. Quick as a wink she dashed around to the other side of the house and took away the ladder.\\

The policemen had looked a little foolish when Pippi jumped, but they looked even more so when they had balanced themselves backward along the ridgepole and were about to climb down the ladder. At first they were very angry at Pippi, who stood on the ground looking up at them, and they told her in no uncertain terms to get the ladder and be quick about it, or she would soon get something she wasn’t looking for.\\

“Why are you so cross at me?” asked Pippi reproachfully. “We’re just playing tag, aren’t we?”\\

The policemen thought a while, and at last one of them said, “Oh, come on, won’t you be a good girl and put the ladder back so that we can get down?”\\

“Of course I will,” said Pippi and put the ladder back instantly. “And when you get down we can all drink coffee and have a happy time.”\\

But the policemen were certainly tricky, because the minute they were down on the ground again they pounced on Pippi and cried, “Now you’ll get it, you little brat!” “Oh, no, I’m sorry. I haven’t time to play any longer,” said Pippi. “But it was fun.”\\

Then she took hold of the policemen by their belts and carried them down the garden path, out through the gate, and onto the street. There she set them down, and it was quite some time before they were ready to get up again.\\

“Wait a minute,” she cried and ran into the kitchen and came back with two cookie hearts.\\

“Would you like a taste?” she asked. “It doesn’t matter that they are a little burned, does it?”\\

Then she went back to Tommy and Annika, who stood there wide-eyed and just couldn’t get over what they had seen. And the policemen hurried back to the town and told all the ladies and gentlemen that Pippi wasn’t quite fit for a children’s home. (They didn’t tell that they had been up on the roof.) And the ladies and gentlemen decided that it would be best after all to let Pippi remain in Villa Villekulla, and if she wanted to go to school she could make the arrangements herself.\\

But Pippi and Tommy and Annika had a very pleasant afternoon. They went back to their interrupted coffee party. Pippi stuffed herself with fourteen cookies, and then she said, “They weren’t what I mean by real policemen. No sirree! Altogether too much talk about children’s homes and pluttifikation and Lisbon.”\\

Afterward she lifted the horse down on the ground and they rode on him, all three. At first Annika was afraid and didn’t want to, but when she saw what fun Tommy and Pippi were having, she let Pippi lift her up on the horse’s back. The horse trotted round and round in the garden, and Tommy sang, “Here come the Swedes with a clang and a bang.”\\

When Tommy and Annika had gone to bed that night Tommy said, “Annika, don’t you think it’s good that Pippi moved here?”\\

“Oh, yes,” said Annika.\\

“I don’t even remember what we used to play before she came, do you?”\\

“Oh, sure, we played croquet and things like that,” said Annika. “But it’s lots more fun with Pippi around, I think. And with horses and things.”\\
\newpage

\section{Pippi Goes to School}
Of course Tommy and Annika went to school. Each morning at eight o’clock they trotted off, hand in hand, swinging their schoolbags.\\

At that time Pippi was usually grooming her horse or dressing Mr. Nilsson in his little suit. Or else she was taking her morning exercises, which meant turning forty-three somersaults in a row. Then she would sit down on the kitchen table and, utterly happy, drink a large cup of coffee and eat a piece of bread and cheese.\\

Tommy and Annika always looked longingly toward Villa Villekulla as they started off to school. They would much rather have gone to play with Pippi. If only Pippi had been going to school too; that would have been something else again.\\

“Just think what fun we could have on the way home from school,” said Tommy.\\

“Yes, and on the way to school too,” said Annika.\\

The more they thought about it the worse they felt to think that Pippi did not go to school, and at last they determined to try to persuade her to begin.\\

“You can’t imagine what a nice teacher we have,” said Tommy artfully to Pippi one afternoon when he and Annika had come for a visit at Villa Villekulla after they had finished their homework.\\

“If you only knew what fun it is in school!” Annika added. “I’d die if I couldn’t go to school.”\\

Pippi sat on a hassock, bathing her feet in a tub. She said nothing but just wiggled her toes for a while so that the water splashed around everywhere.\\

“You don’t have to stay so very long.” continued Tommy. “Just until two o’clock.”\\

“Yes, and besides, we get Christmas vacation and Easter vacation and summer vacation,” said Annika.\\

Pippi bit her big toe thoughtfully but still said nothing. Suddenly, as if she had made some decision, she poured all the water out on the kitchen floor, so that Mr. Nilsson, who sat near her playing with a mirror, got his pants absolutely soaked.\\

“It’s not fair!” said Pippi sternly without paying any attention to Mr. Nilsson’s puzzled air about his wet pants. “It is absolutely unfair! I don’t intend to stand it!”\\

“What’s the matter now?” asked Tommy. “In four months it will be Christmas, and then you’ll have Christmas vacation. But I, what’ll I get?” Pippi’s voice sounded sad. “No Christmas vacation, not even the tiniest bit of a Christmas vacation,” she complained. “Something will have to be done about that. Tomorrow morning I’ll begin school.”\\

Tommy and Annika clapped their hands with delight. “Hurrah! We’ll wait for you outside our gate at eight o’clock.”\\

“Oh, no,” said Pippi. “I can’t begin as early as that. And besides, I’m going to ride to school.”\\

And ride she did. Exactly at ten o’clock the next day she lifted her horse off the porch, and a little later all the people in the town ran to their windows to see what horse it was that was running away. That is to say, they thought he was running away, but it was only Pippi in a bit of a hurry to get to school.\\

She galloped wildly into the schoolyard, jumped off the horse, tied him to a tree, and burst into the schoolroom with such a noise and a clatter that Tommy and Annika and all their classmates jumped in their seats.\\

“Hi, there,” cried Pippi, waving her big hat. “Did I get here in time for pluttifikation?”\\

Tommy and Annika had told their teacher that a new girl named Pippi Longstocking was coming, and the teacher had already heard about Pippi in the little town. As she was a very pleasant teacher, she had decided to do all she could to make Pippi happy in school.\\

Pippi threw herself down on a vacant bench without having been invited to do so, but the teacher paid no attention to her heedless way. She simply said in a very friendly voice, “Welcome to school, little Pippi. I hope that you will enjoy yourself here and learn a great deal.”\\

“Yes, and I hope I’ll get some Christmas vacation,” said Pippi. “That is the reason I’ve come. It’s only fair, you know.\\

“If you would first tell me your whole name,” said the teacher, “then I’ll register you in school.”\\

“My name is Pippilotta Delicatessa Window-shade Mackrehnint Efraim’s Daughter Longstocking, daughter of Captain Efraim Longstocking, formerly the Terror of the Sea, now a cannibal king. Pippi is really only a nickname, because Papa thought that Pippilotta was too long to say.”\\

“Indeed!” said the teacher. “Well, then we shall call you Pippi too. But now,” she continued, “suppose we test you a little and see what you know. You are a big girl and no doubt know a great deal already. Let us begin with arithmetic. Pippi, can you tell me what seven and five are?”\\

Pippi, astonished and dismayed, looked at her and said, “Well, if you don’t know that yourself, you needn’t think I’m going to tell you.”\\

All the children stared in horror at Pippi, and the teacher explained that one couldn’t answer that way in school.\\

“I beg your pardon,” said Pippi contritely. “I didn’t know that. I won’t do it again.”\\

“No, let us hope not,” said the teacher. “And now I will tell you that seven and five are twelve.”\\

“See that!” said Pippi. ``You knew it yourself. Why are you asking then?”\\

The teacher decided to act as if nothing unusual were happening and went on with her examination.\\

“Well now, Pippi, how much do you think eight and four are?” “Oh, about sixty-seven,” hazarded Pippi.\\

“Of course not,” said the teacher. “Eight and four are twelve.”\\

“Well now, really, my dear little woman,” said Pippi, “that is carrying things too far. You just said that seven and five are twelve. There should be some rhyme and reason to things even in school. Furthermore, if you are so childishly interested in that foolishness, why don’t you sit down in a corner by yourself and do arithmetic and leave us alone so we can play tag?”\\

The teacher decided there was no point in trying to teach Pippi any more arithmetic. She began to ask the other children the arithmetic questions.\\

“Can Tommy answer this one?” she asked. “If Lisa has seven apples and Axel has nine apples, how many apples do they have together?”\\

“Yes, you tell, Tommy,” Pippi interrupted, “and tell me too, if Lisa gets a stomach-ache and Axel gets more stomach-ache, whose fault is it and where did they get hold of the apples in the first place?”\\

The teacher tried to pretend that she hadn’t heard and turned to Annika. “Now, Annika, here’s an example for you: Gustav was with his schoolmates on a picnic. He had a quarter when he started out and seven cents when he got home. How much did he spend?”\\

“Yes, indeed,” said Pippi, “and I also want to know why he was so extravagant, and if it was pop he bought, and if he washed his ears properly before he left home.”\\

The teacher decided to give up arithmetic altogether. She thought maybe Pippi would prefer to learn to read. So she took out a pretty little card with a picture of an ibex on it. In front of the ibex’s nose was the letter “i”\\

“Now, Pippi,” she said briskly, “you’ll see something jolly. You see here an ibex. And the letter in front of this ibex is called i.”\\

“That I’ll never believe,” said Pippi. “I think it looks exactly like a straight line with a little fly speck over it. But what I’d really like to know is, what has the ibex to do with the fly speck?”\\

The teacher took out another card with a picture of a snake on it and told Pippi that the letter on that was an s.\\

“Speaking of snakes,” said Pippi, “I’ll never, ever forget the time I had a fight with a huge snake in India. You can’t imagine what a dreadful snake it was, fourteen yards long and mad as a hornet, and every day he ate up five Indians and then two little children for dessert, and one time he came and wanted me for dessert, and he wound himself around me -uhhh!- but I’ve been around a bit, I said, and hit him in the head, bang, and then he hissed uiuiuiuiuiuiuiuitch, and then I hit him again, and bingo! he was dead, and, indeed, so that is the letter s- most remarkable!”\\

Pippi had to stop to get her breath. And the teacher, who had now begun to think that Pippi was an unruly and troublesome child, decided that the class should have drawing for a while. Surely Pippi could sit still and be quiet and draw, thought the teacher. She took out paper and pencils and passed them out to the children.\\

“Now you may draw whatever you wish,” she said and sat down at her desk and began to correct homework. In a little while she looked up to see how the drawing was going. All the children sat looking at Pippi, who lay flat on the floor, drawing to her heart’s content.\\

“But, Pippi,” said the teacher impatiently, “why in the world aren’t you drawing on your paper!”\\

“I filled that long ago. There isn’t room enough for my whole horse on that little snip of a paper,” said Pippi. “Just now I’m working on his front legs, but when I get to his tail I guess I’ll have to go out in the hall.”\\

The teacher thought hard for a while. “Suppose instead we all sing a little song,” she suggested.\\

All the children stood up by their seats except Pippi; she stayed where she was on the floor. “You go ahead and sing,” she said. “I’ll rest myself a while. Too much learning breaks even the healthiest.”\\

But now the teacher’s patience came to an end. She told all the children to go out into the yard so she could talk to Pippi alone.\\

When the teacher and Pippi were alone, Pippi got up and walked to the desk. “Do you know what?” she said. “It was awfully jolly to come to school to find out what it was like. But I don’t think I care about going to school any more, Christmas vacation or no Christmas vacation. There’s altogether too many apples and ibexes and snakes and things like that. It makes me dizzy in the head. I hope that you, Teacher, won’t be sorry.”\\

But the teacher said she certainly was sorry, most of all because Pippi wouldn’t behave decently; and that any girl who acted as badly as Pippi did wouldn’t be allowed to go to school even if she wanted to ever so.\\

“Have I behaved badly?” asked Pippi, much astonished. “Goodness, I didn’t know that,” she added and looked very sad. And nobody could look as sad as Pippi when she was sad. She stood silent for a while, and then she said in a trembling voice, “You understand, Teacher, don’t you, that when you have a mother who’s an angel and a father who is a cannibal king, and when you have sailed on the ocean all your whole life, then you don’t know just how to behave in school with all the apples and ibexes.”\\

Then the teacher said she understood and didn’t feel annoyed with Pippi any longer, and maybe Pippi could come back to school when she was a little older. Pippi positively beamed with delight. “I think you are awfully nice, Teacher. And here is something for you.”\\

Out of her pocket Pippi took a lovely little gold watch and laid it on the desk. The teacher said she couldn’t possibly accept such a valuable gift from Pippi, but Pippi replied, “You’ve got to take it; otherwise I’ll come back again tomorrow, and that would be a pretty how-do-you-do.”\\

Then Pippi rushed out to the schoolyard and jumped on her horse. All the children gathered around to pat the horse and see her off.\\

“You ought to know about the schools in Argentina,” said Pippi, looking down at the children. “That’s where you should go. Easter vacation begins three days after Christmas vacation ends, and when Easter vacation is over there are three days and then it’s summer vacation. Summer vacation ends on the first of November, and then you have a tough time until Christmas vacation begins on November 11. But you can stand that because there are at least no lessons. It is strictly against the law to have lessons in Argentina. Once in a while it happens that some Argentine kid sneaks into a closet and sits they’re studying a lesson, but it’s just too bad for him if his mother finds him. Arithmetic they don’t have at all in the schools, and if there is any kid who knows what seven and five are he has to stand in the corner all day--that is, if he’s foolish enough to let the teacher know that he knows. They have reading on Friday, and then only if they have some books, which they never have.”\\

“But what do they do in school?” asked one little boy.\\

“Eat caramels,” said Pippi decidedly. “There is a long pipe that goes from a caramel factory nearby directly into the schoolroom, and caramels keep shooting out of it all day long so the children have all they can do to eat them up.”\\

“Yes, but what does the teacher do?” asked one little girl.\\

“Takes the paper off the caramels for the children, of course,” said Pippi. “You didn’t suppose they did it themselves, did you? Hardly. They don’t even go to school themselves--they send their brothers.” Pippi waved her big hat.\\

“So long, kids,” she cried gaily. “Now you won’t see me for a while. But always remember how many apples Axel had or you’ll be sorry.”\\

With a ringing laugh Pippi rode out through the gate so wildly that the pebbles whirled around the horse’s hoofs and the windowpanes rattled in the schoolhouse.
\newpage

\section{Pippi Sits on the Gate and Climbs a Tree}
Outside Villa Villekulla sat Pippi, Tommy, and Annika. Pippi sat on one gatepost, Annika on the other, and Tommy sat on the gate. It was a warm and beautiful day toward the end of August. A pear tree that grew close to the fence stretched its branches so low down that the children could sit and pick the best little red-gold pears without any trouble at all. They munched and ate and spit pear cores out onto the road.\\

Villa Villekulla stood just at the edge of the little town, where the street turned into a country road. The people in the little town loved to go walking out Villa Villekulla way, for the country out there was so beautiful.\\

As the children were sitting there eating pears, a girl came walking along the road from town.\\

When she saw the children she stopped and asked, “Have you seen my papa go by?” “M-m-m,” said Pippi. “How did he look? Did he have blue eyes?”\\

“Yes,” said the girl.\\

“Medium large, not too tall and not too short?” “Yes,” said the girl.\\

“Black hat and black shoes?”\\

“Yes, exactly,” said the girl eagerly.\\

“No, that one we haven’t seen,” said Pippi decidedly.\\

The girl looked crestfallen and went off without a word.\\

“Wait a minute,” shrieked Pippi after her. “Was he bald-headed?”\\

“No, he certainly was not,” said the girl crossly.\\

“Lucky for him!” said Pippi and spit out a pear core.\\

The girl hurried away, but then Pippi shouted, “Did he have big ears that reached way down to his shoulders?”\\

“No,” said the girl and turned and came running back in amazement. “You don’t mean to say that you have seen a man walk by with such big ears?”\\

“I have never seen anyone who walks with his ears,” said Pippi. “All the people I know walk with their feet.”\\

“Oh, don’t be silly! I mean have you really seen a man who has such big ears?”\\

“No,” said Pippi, “there isn’t anybody with such big ears. It would be ridiculous. How would they look? It isn’t possible to have such big ears. At least not in this country,” she added after a thoughtful pause. “Of course in China it’s a little different. I once saw a Chinese in Shanghai. His ears were so big that he could use them for a cape. When it rained he just crawled in under his ears and was as warm and snug as you please. Of course his ears didn’t have it so good. If it was very bad weather he used to invite his friends to camp under his ears. There they sat and sang sad songs while the rain poured down. They liked him a lot because of his ears. His name was Hai Shang. You should have seen Hai Shang run to work in the morning. He always came dashing in at the last minute because he loved to sleep late, and you can’t imagine how funny he looked, rushing in with his ears flying behind him like two big golden sails.”\\

The girl had stopped and stood open-mouthed listening to Pippi. And Tommy and Annika forgot to eat any more pears, they were so utterly absorbed in the story.\\

“He had more children than he could count, and the littlest one was named Peter,” said Pippi.\\

“Oh, but a Chinese baby can’t be called Peter,” interrupted Tommy.\\

“That’s just what his wife said to him, ‘A Chinese baby can’t be called Peter.’ But Hai Shang was dreadfully stubborn, and he said that the baby should be called Peter or Nothing. And then he sat down in a corner and pulled his ears over his head and howled. And his poor wife had to give in, of course, and the kid was called Peter.”\\

“Really!” said Annika. “It was the hatefulest kid in all Shanghai,” continued Pippi. “Fussy about his food, so that his mother was most unhappy. You know, of course, that they eat swallows’ nests in China? And there sat his mother, with a whole plate full of swallows’ nests, trying to feed him. ‘Now, little Peter, she said, “come, we’ll eat a swallow’s nest for Daddy.” But Peter just shut his mouth tight and shook his head. At last Hai Shang was so angry that he said that no new food should be prepared for Peter until he had eaten a swallow’s nest for Daddy. And when Hai Shang said something, that was that. The same swallow’s nest rode in and out of the kitchen from May until October. On the fourteenth of July his mother begged to be allowed to give Peter a couple of meatballs, but Hai Shang said no.”\\

“Nonsense!” said the girl in the road.\\

“Yes, that’s just what Hai Shang said,” continued Pippi. “Nonsense, " he said, “it’s perfectly plain that the child can eat the swallow’s nest if he’ll only stop being so stubborn.” But Peter kept his mouth shut tight from May to October.\\

“But how could he live?” asked Tommy, astonished.\\

“He couldn’t live,” said Pippi. “He died. Of Plain Common Ordinary Pigheadedness. The eighteenth of October. And was buried the nineteenth. And on the twentieth a swallow flew in through the window and laid an egg in the nest, which was standing on the table. So it came in handy after all. No harm done,” said Pippi happily. Then she looked thoughtfully at the bewildered girl, who still stood in the road.\\

“Why do you look so funny?” asked Pippi.\\

“What’s the matter? You don’t really think that I’m sitting here telling lies, do you? Just tell me if you do,” said Pippi threateningly and rolled up her sleeves.\\

“Oh, no, indeed,” said the girl, terrified. “I don’t really mean that you are lying, but--"\\

“No?” said Pippi. “But it’s just what I’m doing. I’m lying so my tongue is turning black. Do you really think that a child can live without food from May to October? To be sure, I know they can get along without food for three or four months all right. But from May to October? It’s just foolish to think that. You must know that’s a lie. You mustn’t let people fool you so easily.\\

Then the girl left without turning around again.\\

“People will believe anything.” said Pippi to Tommy and Annika. “From May until October! That’s ridiculous!”\\

Then she called after the girl, “No, we haven’t seen your papa. We haven’t seen a single bald- headed person all day. But yesterday seventeen of them went by. Arm in arm.”\\

Pippi’s garden was really lovely. You couldn’t say it was well kept, but there were wonderful grass plots that were never cut, and old rosebushes that were full of white and yellow and pink roses--perhaps not such fine roses, but oh, how sweet they smelled! A good many fruit trees grew there too, and, best of all, several ancient oaks and elms that were excellent for climbing.\\

The trees in Tommy’s and Annika’s garden were not very good for climbing, and besides, their mother was always so afraid they would fall and get hurt that they had never climbed much. But now Pippi said, “Suppose we climb up in the big oak tree?”\\

Tommy jumped down from the gate at once, delighted with the suggestion. Annika was a little hesitant, but when she saw that the trunk had nubbly places to climb on, she too thought it would be fun to try.\\

A few feet above the ground the oak divided into two branches and right there was a place just like a little room. Before long all three children were sitting there. Over their heads the oak spread out its crown like a great green roof.\\

“We could drink coffee here,” said Pippi. “I’ll skip in and make a little.” Tommy and Annika clapped their hands and shouted, “Bravo!”\\

In a little while Pippi had the coffee ready. She had made buns the day before. She came and stood under the oak and began to toss up coffee cups. Tommy and Annika caught them. Only sometimes it was the oak that caught them, and so two cups were broken. Pippi ran in to get new ones. Next it was the buns’ turn, and for a while the air was full of flying buns. At least they didn’t break. At last Pippi climbed up with the coffeepot in one hand. She had cream in a little bottle in her pocket, and sugar in a little box.\\

Tommy and Annika thought coffee had never tasted so good before. They were not allowed to drink it every day -- only when they were at a party. And now they were at a party. Annika spilled a little coffee in her lap. First it was warm and wet, and then it was cold and wet.\\

When they had finished, Pippi threw the cups down on the grass. “I want to see how strong the china they make these days is,” she said. Strangely enough, one cup and three saucers held together, and only the spout of the coffeepot broke off.\\

Presently Pippi decided to climb a little higher.\\

“Can you beat this?” she cried suddenly. “The tree is hollow.”\\

There in the trunk was a big hole, which the leaves had hidden from the children’s sight.\\

“Oh, may I climb up and look too?” called Tommy. But there was no answer. “Pippi, where are you?” he cried, worried.\\

Then they heard Pippi’s voice, not from above but from way down below. It sounded as if it came from under the ground.\\

“I’m inside the tree. It is hollow clear down to the ground. If I peek out through a little crack I can see the coffee pot outside on the grass.”\\

“Oh, how will you get up again?” cried Annika. “I’m never coming up,” said Pippi. “I’m going to stay here until I retire and get a pension. And you’ll have to throw my food down through that hole up there. Five or six times a day.\\

Annika began to cry.\\

“Why be sorry? Why complain?” said Pippi.\\

“You come down here too, and then we can play that we are pining away in a dungeon.”\\

“Never in this world!” said Annika, and to be on the safe side she climbed right down out of the tree.\\

“Annika, I can see you through the crack,” cried Pippi. “Don’t step on the coffee pot; it’s an old well-mannered coffee pot that never did anyone any harm. It can’t help it that it doesn’t have a spout any longer.”\\

Annika went up to the tree trunk, and through a little crack she saw the very tip of Pippi’s finger. This comforted her a good deal, but she was still worried.\\

“Pippi, can’t you really get up?” she asked.\\

Pippi’s finger disappeared, and in less than a minute her face popped out of the hole up in the tree.\\

“Maybe I can if I try very hard,” she said and parted the foliage with her hands.\\

“If it’s as easy as all that to get up,” said Tommy, who was still up in the tree, “then I want to come down and pine away a little too.”\\

“Wait,” said Pippi, “I think well get a ladder.\\

She crawled out of the hole and hurried down the tree. Then she ran after a ladder, pushed it up the tree, and let it down into the hole.\\

Tommy was wild to go down. It was difficult to climb to the hole, because it was so high up, but Tommy was brave. And he wasn’t afraid to climb down into the dark hollow in the trunk. Annika watched him disappear and wondered if she would ever see him again. She peeked in through the crack.\\

“Annika,” came Tommy’s voice. “You can’t imagine how wonderful it is here. You must come in too. It isn’t the least bit dangerous when you have a ladder to climb on. If you only do it once, you’ll never want to do anything else.”\\

“Are you sure?” asked Annika. “Absolutely,” said Tommy.\\

With trembling legs Annika climbed up in the tree again, and Pippi helped her with the last hard bit. She drew back a little when she saw how dark it was in the tree trunk, but Pippi held her hand and kept encouraging her.\\

“Don’t be scared, Annika,” she heard Tommy say from down below. “Now I can see your legs, and I’ll certainly catch you if you fall.”\\

But Annika didn’t fall. She reached Tommy safely, and a moment later Pippi followed.
“Isn’t it grand here?” said Tommy.\\

And Annika had to admit that it was. It wasn’t nearly so dark as she had thought, because light came in through the crack. She peeked through and announced that she too could see the coffee pot outside on the grass.\\

“We’ll have this for our secret hiding place,” said Tommy. “Nobody will know that we are here. And if they should come and hunt around outside for us, we can see them through the crack. And we’ll have a good laugh.”\\

“We can have a little stick and poke it out through the crack and tickle them, and then they’ll think the place is haunted,” said Pippi.\\

At this idea the children were so delighted that they hugged each other, all three. Then they heard the “ding-dong” that meant the bell was ringing for dinner at Tommy’s and Annika’s house.\\

“Oh, bother!” said Tommy. “Now we’ve got to go home. But we’ll come over tomorrow as soon as we get back from school.”\\

“Do that,” said Pippi.\\

And so they climbed up the ladder, first Pippi, then Annika, and Tommy last. And then they climbed down out of the tree, first Pippi, then Annika, and Tommy last.\\
\newpage

\section{Pippi Arranges a Picnic}
“We don’t have any school today because we’re having Scrubbing Vacation,” said Tommy to Pippi.\\

“Scrubbing Vacation? Well, I like that!” said Pippi. “Another injustice! Do I get any Scrubbing Vacation? Indeed I don’t, though goodness knows I need one. Just look at the kitchen floor. But for that matter,” she added, “now I come to think of it, I can scrub without any vacation. And that’s what I intend to do right now, Scrubbing Vacation or no Scrubbing Vacation. I’d like to see anybody stop me! You two sit on the kitchen table, out of the way.”\\

Tommy and Annika obediently climbed up on the kitchen table, and Mr. Nilsson hopped up after them and went to sleep in Annika’s lap.\\

Pippi heated a big kettle of water and without more ado poured it out on the kitchen floor. She took off her big shoes and laid them neatly on the bread plate. She tied two scrubbing brushes on her bare feet and skated over the floor, plowing through the water so that it splashed all around her.\\

“I certainly should have been a skating princess,” she said and kicked her left foot up so high that the scrubbing brush broke a piece out of the overhead light.\\

“Grace and charm I have at least,” she continued and skipped nimbly over a chair standing in her way.\\

“Well, now I guess it’s clean,” she said at last and took off the brushes. “Aren’t you going to dry the floor?” asked Annika.\\

“Oh, no, it can dry in the sun,” answered Pippi. “I don’t think it will catch cold so long as it keeps moving.”\\

Tommy and Annika climbed down from the table and stepped across the floor very carefully so they wouldn’t get wet.\\

Out of doors the sun shone in a clear blue sky. It was one of those radiant September days that make you feel like walking in the woods. Pippi had an idea.\\

“Let’s take Mr. Nilsson and go on a little picnic.” “Oh, yes, let’s,” cried Tommy and Annika.\\

“Run home and ask your mother, then.” said Pippi, “and I’ll be getting the picnic basket ready.\\

Tommy and Annika thought that was a good suggestion. They rushed home and were back again almost immediately, but Pippi was already waiting by the gate with Mr. Nilsson on her shoulder, a walking stick in one hand, and a big basket in the other.\\

The children walked along the road a little way and then turned into a pasture where a pleasant path wound in and out among the thickets of birch and hazel. Presently they came to a gate on the other side of which was an even more beautiful pasture, but right in front of the gate stood a cow who looked as if nothing would persuade her to move. Annika yelled at her, and Tommy bravely went up and tried to push her away, but she just stood there staring at the children with her big cow eyes. To put an end to the matter, Pippi set down her basket and lifted the cow out of the way. The cow, looking very silly, lumbered off into the hazel bushes.\\

“How can cows be so bull-headed,” said Pippi and jumped over the gate.\\

“What a lovely, lovely wood!” cried Annika in delight as she climbed up on all the stones she could see. Tommy had brought along a dagger Pippi had given him, and with it he cut walking sticks for Annika and for himself. He cut his thumb a little too, but that didn’t matter.\\

“Maybe we ought to pick some mushrooms,” said Pippi, and she broke off a pretty, rosy one. “I wonder if it’s possible to eat it?” she continued. “At any rate, it isn’t possible to drink it--that much I know; so there is no choice except to eat it. Maybe it’s possible.”\\

She took a big bite and swallowed it. “It was possible,” she announced, delighted. Yes sirree, we’ll certainly stew the rest of this sometime,” she said and threw it high over the treetops. “What have you got in your basket?” asked Annika. “Is it something good?”\\

“I wouldn’t tell you for a thousand dollars,” said Pippi. “First we must find a good picnic spot.”
The children eagerly began to look for such a place. Annika found a large flat stone that she thought was satisfactory, but it was covered with red ants. “I don’t want to sit with than,” said Pippi, “because I’m not acquainted with them.”\\

“And besides, they bite,” said Tommy. “Do they?” said Pippi. “Bite back then.”\\

Then Tommy found a little clearing among the hazel bushes, and he thought that would be a good place.\\

“Oh, no, that’s not sunny enough for my freckles,” said Pippi, “and I do think freckles are so attractive.”\\

Farther on they came to a hill that was easy to climb. On one side of the hill was a nice sunny rock just like a little balcony, and there they sat down.\\

“Now shut your eyes while I set the table,” said Pippi. Tommy and Annika squeezed their eyes as tightly shut as possible. They heard Pippi opening the basket and rattling paper.\\

“One, two, nineteen--now you may look,” said Pippi at last.\\

They looked, and they squealed with delight when they saw all the good things Pippi had spread on the bare rock. There were good sandwiches with meatballs and ham, a whole pile of sugared pancakes, several little brown sausages, and three pineapple puddings. For, you see, Pippi had learned cooking from the cook on her father’s ship.\\

“Aren’t Scrubbing Vacations grand!” said Tommy with his mouth full of pancakes. “We ought to have them every day.”\\

“No, indeed, I’m not that anxious to scrub,” said Pippi. “It’s fun, to be sure, but not every day. That would be too tiresome.”\\

At last the children were so full they could hardly move, and they sat still in the sunshine and just enjoyed it.\\

“I wonder if it is hard to fly,” said Pippi and looked dreamily over the edge of the rock. The rock sloped down very steeply below them, and it was a long way to the ground.\\

“Down at least one ought to be able to learn to fly,” she continued. “It must be harder to fly up. But you could begin with the easiest way. I do think I’ll try.”\\

“No, Pippi,” cried both Tommy and Annika. “Oh, dear, Pippi, don’t do that!”\\

But Pippi was already standing at the edge.\\

“Fly, you foolish fly, fly, and the foolish fly flew,” she said, and just as she said “flew” she lifted her arms and took off into the air. In half a second there was a thud. It was Pippi hitting the ground. Tommy and Annika lay on their stomachs and looked down at her, terrified.\\

Pippi got up and brushed off her knees. “I forgot to flap,” she said joyfully, “and I guess I had too many pancakes in my stomach.”\\

At that moment the children noticed that Mr. Nilsson had disappeared. He had evidently gone off on a little expedition of his own. They remembered that they had last seen him contentedly chewing the picnic basket to pieces, but during Pippi’s flying experiment they had forgotten him. And now he was gone.\\

Pippi was so angry that she threw her shoe into a big deep pool of water. “You should never take monkeys with you anywhere,” she said. “He should have been left at home to pick fleas off the horse. That would have served him right,” she continued, wading out into the pool to get her shoe. The water reached up to her waist.\\

“I might as well take advantage of this and wash my hair,” said Pippi and ducked her head under the water and kept it there so long that the water began to bubble.\\

“There now, I’ve saved a visit to the hair- dresser,” she said contentedly when at last she came up for air. She stepped out of the pool and put on her shoe. Then they went off to hunt for Mr. Nilsson.\\

“Hear how it squishes when I walk,” laughed Pippi. “It says ‘klafs, klafs’ in my dress and ‘squish, squish’ in my shoes. Isn’t that jolly? I think you ought to try it too,” she said to Annika, who was walking along beside her, with her lovely flaxen hair, pink dress, and little white kid shoes.\\

“Some other time,” said the sensible Annika. They walked on.\\

“Mr. Nilsson certainly can be exasperating,” said Pippi. “He’s always doing things like this. Once in Arabia he ran away from me and took a position as a maidservant to an elderly widow. That last was a lie, of course,” she added after a pause.\\

Tommy suggested they all three go in different directions and hunt. At first Annika didn’t want, but Tommy said, “You aren’t afraid cat, are you?” And, of course, Annika couldn’t tolerate such an insult, so off they all went.\\

Tommy went through a field. Mr. Nilsson he did not find, but he did find something else. A bull! Or to be more exact, the bull found Tommy. And the bull did not like Tommy, for he was a very cross bull who was not at all fond of children. With his head down he charged toward Tommy, bellowing fearfully. Tommy let out a terrified shriek that could be heard all through the woods. Pippi and Annika heard it and came running to see what was the matter. By that time the bull had almost reached Tommy who had fallen head over heels over a stump. !“What a stupid bull!” said Pippi to Annika, who was crying uncontrollably. “He ought to know he can’t act like that. He’ll get Tommy’s white sailor suit all dirty. I’ll have to go and talk some sense into the stupid animal.”\\

And off she started. She ran up and pulled the bull by the tail. “Forgive me for breaking up the party, she said. Since she had given his tail a good hard pull, the bull turned around and saw a new child to catch on his horns.\\

“As I was saying,” went on Pippi, “forgive me for breaking up, and also forgive me for breaking off,” and with that she broke off one of the bull’s horns. It isn’t the style to have two horns this year,” she said. “All the better bulls have just one horn--if any.” And she broke off the other horn too.\\

As bulls have no feeling in their horns, this one didn’t know what she had done. He charged at Pippi, and if she had been any other child there would have been nothing left but a grease spot.\\

“Hey, hey, stop tickling me!” shrieked Pippi.\\

“You can’t imagine how ticklish I am! Hey, stop, stop, or I’ll die laughing!”\\

But the bull didn’t stop, and at last Pippi jumped up on his back to get a little rest. To be sure, she didn’t get much, because the bull didn’t in the least approve of having Pippi on his back. He dodged about madly to get her off, but she clamped her knees and hung on. The bull dashed up and down the field, bellowing so hard that smoke came out of his nostrils. Pippi laughed and shrieked and waved at Tommy and Annika, who stood a little distance away, trembling like aspen leaves. The bull whirled round and round, trying to throw Pippi.\\

“See me dancing with my little friend!” cried Pippi and kept her seat. At last the bull was so tired that he lay down on the ground and wished that he’d never seen such a thing as a child. He had never thought children amounted to much anyway.\\

“Are you going to take a little nap now?” asked Pippi politely. “Then I won’t disturb you.”\\

She got off his back and went over to Tommy and Annika. Tommy had cried a little. He had a cut on one arm, but Annika had bandaged it with her handkerchief so that it no longer hurt.\\

“Oh, Pippi!” cried Annika excitedly.\\

“Sh, sh,” whispered Pippi. “Don’t wake the bull. He’s sleeping. If we wake him he’ll be fussy.\\

But the next minute, without paying any attention to the bull and his nap, she was shrieking at the top of her voice, “Mr. Nilsson, Mr. Nilsson, where are you? We’ve got to go home.”\\

And, believe it or not, there sat Mr. Nilsson up in a pine tree, sucking his tail and looking very lonely. It wasn’t much fun for a little monkey to be left all alone in the woods. He skipped down from the pine and up on Pippi’s shoulder, waving his little straw hat as he always did when he was very happy.\\

“Well, well, so you aren’t going to be a maidservant this time?” said Pippi, stroking his back. “Oh, that was a lie, that’s true,” she continued. “But still, if it’s true, how can it be a lie?” she argued. “You wait and see, it’s going to turn out that he was a maidservant in Arabia after all, and if that’s the case, I know who’s going to make the meatballs at our house hereafter!”\\

And then they strolled home, Pippi’s dress still going klafs, klafs,” and her shoes “squish, squish.”\\

Tommy and Annika thought they had had a wonderful day in spite of the bull, and they sang a song they had learned at school. It was really a summa song, but they thought it fitted very well even if it was now nearly autumn:

\begin{center}
In the jolly summertime\\
Through field and wood we make out way. Nobody’s sad, everyone’s gay.\\
We sing as we go, hol-la, hol-lo!\\
You who are young,\\
Come join in our song.\\
Don’t sit home moping all the daylong. Our song will swell\\
Through wood and dell\\
And up to the mountaintop as well. In the jolly summertime\\
We sing as we go, hol-la, hol-lo!\\
Pippi sang too, but with slightly different words:\\
In the jolly summertime\\
Through field and wood I make my way.\\
I do exactly as I wish,\\
And when I walk it goes squish, squish, Squish, squish. Squish, squish.\\
And my old shoe--\\
It’s really true--\\
Sometimes says “chip” and sometimes “choo. For the shoe is wet.\\
The bull sleeps yet.\\
And I eat all the rice pudding I can get.\\
In the jolly summertime\\
I squish wherever I go. Squish-oh! Squish-oh!\\
\end{center}
\newpage

\section{Pippi Goes to the Circus}
A circus had come to the little town, and all the children were begging their mothers and fathers for permission to go. Of course Tommy and Annika asked to go too, and their kind father immediately gave them some money.\\

Clutching it tightly in their hands, they rushed over to Pippi’s. She was on the porch with her horse, braiding his tail into tiny pigtails and tying each one with red ribbon.\\

“I think it’s his birthday today,” she announced, “so he has to be all dressed up.”\\

“Pippi,” said Tommy, all out of breath because they had been running so fast, “Pippi, do you want to go with us to the circus?”\\

“I can go with you most anywhere,” answered Pippi, “but whether I can go to the surkus or not I don’t know, because I don’t know what a surkus is. Does it hurt?”\\

“Silly! said Tommy. “Of course it doesn’t hurt; it’s fun. Horses and clowns and pretty ladies that walk the tightrope.”\\

“But it costs money,” said Annika, opening her small fist to see if the shiny half-dollar and the quarters were still there.\\

“I’m rich as a troll,” said Pippi, “so I guess I can buy a surkus all right. But it’ll be crowded here if I have more horses. The clowns and the pretty ladies I could keep in the laundry, but it’s harder to know what to do with the horses.”\\

“Oh, don’t be so silly,” said Tommy, “you don’t buy a circus. It costs money to go and look at it see?”\\

“Preserve us!” cried Pippi and shut her eyes tightly. “It costs money to look? And here I go around goggling all day long. Goodness knows how much money I’ve goggled up already!”\\

Then, little by little, she opened one eye very carefully, and it rolled round and round in her head. “Cost what it may,” she said, “I must take a look!”\\

At last Tommy and Annika managed to explain to Pippi what a circus really was, and she took some gold pieces out of her suitcase. Then she put on her hat, which was as big as a millstone, and off they all went.\\

There were crowds of people outside the circus tent and a long line at the ticket window. But at last it was Pippi’s turn. She stuck her head through the window and stared at the dear old lady sitting there.\\

“How much does it cost to look at you?” Pippi asked.\\

But the old lady was a foreigner who did not understand what Pippi meant and answered in broken Swedish.\\

“Little girl, it costs a dollar and a quarter in the grandstand and seventy-five cents on the benches and twenty-five cents for standing room.”\\

Now Tommy interrupted and said that Pippi wanted a seventy-five-cent ticket. Pippi put down a gold piece and the old lady looked suspiciously at it. She bit it too, to see if it was genuine. At last she was convinced that it really was gold and gave Pippi her ticket and a great deal of change in silver.\\

“What would I do with all those nasty little white coins?” asked Pippi disgustedly. “Keep them and then I can look at you twice, in the standing room.\\

As Pippi absolutely refused to accept any change, the lady changed her ticket to one for the grandstand and gave Tommy and Annika grandstand tickets too without their having to pay a single penny. In that way Pippi, Tommy, and Annika came to sit on some beautiful red chairs right next to the ring. Tommy and Annika turned around several times to wave to their schoolmates, who were sitting much farther away.\\

“This is a remarkable place,” said Pippi, looking around in astonishment. “But, see, they’ve spilled sawdust all over the floor! Not that I’m over fussy myself, but that does look careless to me.”\\

Tommy explained that all circuses had sawdust on the floor for the horses to run around in.\\

On a platform nearby, the circus band suddenly began to play a thundering march. Pippi clapped her hands wildly and jumped up and down with delight. “Does it cost money to hear too!” she asked. “Or can you do that for nothing?”\\

At that moment the curtain in front of the performers’ entrance was drawn aside, and the ring-master in a black frock coat, with a whip in his hand, came running in, followed by ten white horses with red plumes on their heads.\\

The ringmaster cracked his whip, and all the horses galloped around the ring. Then he cracked it again, and all the horses stood still with their front feet up on the railing around the ring.\\

One of them had stopped directly in front of the children. Annika didn’t like to have a horse so near her and drew back in her chair as far as she could, but Pippi leaned forward and took the horse’s right foot in her hands.\\

“Hello, there,” she said, “my horse sent you his best wishes. It’s his birthday today too, but he has bows on his tail instead of on his head.”\\

Luckily she dropped the foot before the ring- master cracked his whip again, because then all the horses jumped away from the railing and began to run around the ring.\\

When the act was over, the ringmaster bowed politely and the horses ran out. In an instant the curtain opened again for a coal-black horse. On its back stood a beautiful lady dressed in green silk tights. The program said her name was Miss Carmencita.\\

The horse trotted around in the sawdust, and Miss Carmencita stood calmly on his back and smiled. But then something happened; just as the horse passed Pippi’s seat, something came swishing through the air--and it was none other than Pippi herself. And there she stood on the horse’s back, behind Miss Carmencita. At first Miss Carmencita was so astonished that she nearly fell off the horse. Then she got mad. She began to strike out with her hands behind her back to make Pippi jump off. But that didn’t work.\\

“Take it easy,” said Pippi. “Do you think you’re the only one who can have any fun? Other people have paid too, haven’t they?”\\

Then Miss Carmencita tried to jump off herself, but that didn’t work either, because Pippi was holding her tightly around the waist. At that the audience couldn’t help laughing. They thought it was funny to see the lovely Miss Carmencita held against her will by a little red-headed youngster who stood there on the horse’s back in her enormous shoes and looked as if she had never done anything except perform in a circus.\\

But the ringmaster didn’t laugh. He turned toward an attendant in a red uniform and made a sign to him to go and stop the horse.\\

“Is this act already over,” asked Pippi in a disappointed tone, “just when we were having so much fun?”\\

“Horrible child!” hissed the ringmaster between his teeth. “Get out of here!”\\

Pippi looked at him sadly. “Why are you mad at me?” she asked. “What’s the matter? I thought we were here to have fun.”\\

She skipped off the horse and went back to her seat. But now two huge guards came to throw her out. They took hold of her and tried to lift her up.\\

They couldn’t do it. Pippi sat absolutely still, and it was impossible to budge her although they tried as hard as they could. At last they shrugged their shoulders and went off.\\

Meanwhile the next act had begun. It was Miss Elvira about to walk the tightrope. She wore a pink tulle skirt and carried a pink parasol in her hand. With delicate little steps she ran out on the rope. She swung her legs gracefully in the air and did all sorts of tricks. It looked so pretty. She even showed how she could walk backward on the narrow rope. But when she got back to the little platform at the end of the rope, there stood Pippi.\\

“What are you going to do now?” asked Pippi, delighted when she saw how astonished Miss Elvira looked.\\

Miss Elvira said nothing at all but jumped down from the rope and threw her arms around the ringmaster’s neck, for he was her father. And the ringmaster once more sent for his guards to throw Pippi out. This time he sent for five of them, but all the people shouted, “Let her stay! We want to see the red-headed girl.” And they stamped their feet and clapped their hands.\\

Pippi ran out on the rope, and Miss Elvira’s tricks were as nothing compared with Pippi’s. When she got to the middle of the rope she stretched one leg straight up in the air, and her big shoe spread out like a roof over her head. She bent her foot a little so that she could tickle herself with it back of her ear.\\

The ringmaster was not at all pleased to have Pippi performing in his circus. He wanted to get rid of her, and so he stole up and loosened the mechanism that held the rope taut, thinking surely Pippi would fall down.\\

But Pippi didn’t. She set the rope a-swinging instead. Back and forth it swayed, and Pippi swung faster and faster, until suddenly she leaped out into the air and landed right on the ringmaster. He was so frightened he began to run.\\

“Oh, what a jolly horse!” cried Pippi. “But why don’t you have any pompoms in your hair?”\\

Now Pippi decided it was time to go back to Tommy and Annika. She jumped off the ringmaster and went back to her seat. The next act was about to begin, but there was a brief pause because the ringmaster had to go out and get a drink of water and comb his hair.\\

Then he came in again, bowed to the audience, and said, “Ladies and gentlemen, in a moment you will be privileged to see the Greatest Marvel of all time, the Strongest Man in the World, the Mighty Adolf, whom no one has yet been able to conquer. Here he comes, ladies and gentlemen, Allow me to present to you THE MIGHTY ADOLF.”\\

And into the ring stepped a man who looked as big as a giant. He wore flesh- colored tights and had a leopard skin draped around his stomach. He bowed to the audience and looked very pleased with himself.\\

“Look at these muscles,” said the ringmaster and squeezed the Mighty Adolfs arm where the muscles stood out like balls under the skin.\\

“And now, ladies and gentlemen, I have a very special invitation for you. Who will challenge the Mighty Adolf in a wrestling match? Which of you dares to try his strength against the World’s Strongest Man? A hundred dollars for anyone who can conquer the Mighty Adolf! A hundred dollars, ladies and gentlemen! Think of that! Who will be the first to try?”\\

Nobody came forth.\\

“What did he say?” asked Pippi.\\

“He says that anybody who can lick that big man will get a hundred dollars.” answered Tommy.\\

“I can,” said Pippi, “but I think it would be too bad to, because he looks nice.” “Oh, no, you couldn’t,” said Annika, “he’s the strongest man in the world.”\\

“Man, yes,” said Pippi, “but I am the strongest girl in the world, remember that.” Meanwhile the Mighty Adolf was lifting heavy iron weights and bending thick iron rods in the middle just to show how strong he was.\\

“Oh, come now, ladies and gentlemen,” cried the ringmaster, “is there really nobody here who wants to earn a hundred dollars? Shall I really be forced to keep this myself?” And he waved a bill in the air.\\

“No, that you certainly won’t be forced to do.” said Pippi and stepped over the railing into the ring.
The ringmaster was absolutely wild when he saw her. “Get out of here! I don’t want to see any more of you,” he hissed.\\

“Why do you always have to be so unfriendly?” said Pippi reproachfully. “I just want to fight with Mighty Adolf.”\\

“This is no place for jokes,” said the ringmaster. “Get out of here before the Mighty Adolf hears your impudent nonsense.”\\

But Pippi went right by the ringleader and up to Mighty Adolf. She took his hand and shook it heartily.\\

“Shall we fight a little, you and I?” she asked.\\

Mighty Adolf looked at her but didn’t understand a word. “In one minute I’ll begin,” said Pippi.\\

And begin she did. She grabbed Mighty Adolf around the waist, and before anyone knew what was happening she had thrown him on the mat. Mighty Adolf leaped up, his face absolutely scarlet.\\

“Atta girl, Pippi!” shrieked Tommy and Annika, so loudly that all the people at the circus heard it and began to shriek, “Atta girl, Pippi!” too. The ringmaster sat on the railing, wringing his hands. He was mad, but Mighty Adolf was madder. Never in his life had he experienced anything so humiliating as this. And he certainly intended to show that red-headed girl what kind of a man Mighty Adolf really was. He rushed at Pippi and caught her round the waist, but Pippi stood firm as a rock.\\

“You can do better than that,” she said to encourage him. Then she wriggled out of his grasp, and in the twinkling of an eye Mighty Adolf was on the mat again. Pippi stood beside him, waiting. She didn’t have to wait long. With a roar he was up again, rushing at her.\\

“Tiddelipom and piddeliday,” said Pippi.\\

All the people in the tent stamped their feet and threw their hats in the air and shouted, “Hurrah, Pippi!”\\

When Mighty Adolf came rushing at her for the third time, Pippi lifted him high in the air and, with her arms straight above her, carried him clear around the ring. Then she laid him down on the mat again and held him there.\\

“Now, little fellow,” said she, “I don’t think we’ll bother about this any more. We’ll never have any more fun than we’ve had already.\\

“Pippi is the winner! Pippi is the winner!” cried all the people.\\

Mighty Adolf stole out as fast as he could, and the ringmaster had to go up and hand Pippi the hundred dollars, although he looked as if he’d much prefer to eat her.\\

“Here you are, young lady, here you are,” he said. “One hundred dollars.”\\

“That thing!” said Pippi scornfully. “What would I want with that old piece of paper. Take it and use it to fry herring on if you want to.” And she went back to her seat.\\

“This is certainly a long surkus,” she said to Tommy and Annika. “I think I’ll take a little snooze, but wake me if they need my help with anything else.”\\

And then she lay back in her chair and went to sleep at once. There she lay and snored while the clowns, the sword swallowers, and the snake charmers did their tricks for Tommy and Annika and all the rest of the people at the circus.\\

“Just the same, I think Pippi was best of all,” whispered Tommy to Annika.
\newpage

\section{Pippi Entertains Two Burglars}
After Pippi’s performance at the circus there was not a single person in all the little town who did not know how strong she was. There was even a piece about her in the paper. But people who lived in other places, of course, didn’t know who Pippi was.\\

One dark autumn evening two tramps came walking down the road past Villa Villekulla. They were two bad thieves wandering about the country to see what they could steal. They saw that there was a light in the windows of Villa Villekulla and decided to go in to ask for a sandwich.\\

That evening Pippi had poured out all her gold pieces on the kitchen floor and sat there counting them. To be sure, she couldn’t count very well, but she did it now and then anyway, just to keep everything in order.\\

"…..sixty-five, sixty-six, sixty-seven, sixty- eight, sixty-nine, sixty-ten, sixty- eleven, sixty- twelve, sixty-thirteen, sixty-sixteen-whew, it makes my throat feel like sixty! Goodness, there must be some more numbers in the arithmetic- oh, yes, now I remember--one hundred four, one thousand. That certainly is a lot of money,” said Pippi.\\

There was a loud knock on the door.\\

“Walk in or stay out, whichever you choose!” shouted Pippi. “I never force anyone against his will.”\\

The door opened and the two tramps came in. You can imagine that they opened their eyes when they saw a little red-headed girl sitting all alone on the floor, counting money.\\

“Are you all alone at home?” they asked craftily.\\

“Of course not,” said Pippi. “Mr. Nilsson is at home too.”\\

The thieves couldn’t very well know that Mr. Nilsson was a monkey sleeping in a little green bed with a doll’s quilt around his stomach. They thought the man of the house must be named Mr. Nilsson and they winked at each other. “They can come back a little later” is what they meant, but to Pippi they said, “We just came in to ask what your clock is.”\\

They were so excited that they had forgotten all about the sandwich.\\

“Great, strong men who don’t know what a clock is!” said Pippi. “Where in the world were you brought up? The clock is a little round thingamajig that says ‘tick tack, tick tack,’ and that goes and goes but never gets to the door. Do you know any more riddles? Out with them if you do,” said Pippi encouragingly.\\

The tramps thought Pippi was too little to tell time, so without another word they went out again.\\

“I don’t demand that you say ‘tack’ " [thanks in Swedish], shouted Pippi after them, “but you could at least make an effort and say ‘tick.’ You haven’t even as much sense as a clock has. But by all means go in peace.” And Pippi went back to her counting.\\

No sooner were the tramps outside than they began to rub their hands with delight. “Did you see all that money? Heavenly day!” said one of them.\\

“Yes, once in a while luck is with us,” said the other. “All we need to do is wait until the kid and that Nilsson are asleep. Then we’ll sneak in and grab the dough.”\\

They sat down under an oak tree in the garden to wait. A drizzling rain was falling; they were very hungry, so they were quite uncomfortable, but the thought of all that money kept their spirits up.\\

From time to time lights went out in other houses, but in Villa Villekulla they shone on. It so happened that Pippi was learning to dance the shottische, and she didn’t want to go to bed until she was sure she could do it. At last, however, the lights went out in the windows of Villa Villekulla too.\\

The tramps waited quite a while until they were sure Mr. Nilsson would have gone to sleep. At last they crept quietly up to the kitchen door and prepared to open it with their burglar tools. Meanwhile one of them--his name, as a matter of fact, was Bloom--just happened to feel the doorknob. The door was not locked!\\

“Well, some people are smart!” he whispered to his companion. “the door is open!”\\

“So much the better for us,” answered his companion, a black-haired man called Thunder-Karlsson by those who knew him. Thunder-Karlsson turned on his pocket flashlight, and they crept into the kitchen. There was no one there. In the next room was Pippi’s bed, and there also stood Mr. Nilsson’s little doll bed.\\

Thunder-Karlsson opened the door and looked around carefully. Everything was quiet as he played his flashlight around the room. When the light touched Pippi’s bed the two tramps were amazed to see nothing but a pair of feet on the pillow. Pippi, as usual, had her head under the covers at the foot of the bed.\\

“That must be the girl,” whispered Thunder- Karlsson to Bloom. “And no doubt she sleeps soundly. But where in the world is Nilsson, do you suppose?”\\

“Mr. Nilsson, if you please,” came Pippi’s calm voice from under the covers. “Mr. Nilsson is in the little green doll bed.”\\

The tramps were so startled that they almost rushed out at once, but then it suddenly dawned on them what Pippi had said. That Mr. Nilsson was lying in a dolls bed! And now in the light of the flashlight they could see the little bed and the tiny monkey lying in it.\\

Thunder-Karlsson couldn’t help laughing. “Bloom,” he said, “Mr. Nilsson is a monkey. Can you beat that?”\\

“Well, what did you think he was?” came Pippi’s calm voice from under the covers again. “A lawn mower?”\\

“Aren’t your mother and father at home?” asked Bloom.\\

“No,” said Pippi. “They’re gone. Completely gone.”\\

Thunder-Karlsson and Bloom chuckled with delight.\\

“Listen, little girl,” said Thunder-Karlsson, “come out so we can talk to you.”\\

“No, I’m sleeping,” said Pippi. Is it more riddles you want? If so, answer this one. What is it that goes and goes and never gets to the door?”\\

Now Bloom went over and pulled the covers off Pippi.\\

“Can you dance the schottischel” asked Pippi, looking at him gravely in the eye. “I can.”\\

“You ask too many questions,” said Thunder-Karlsson. “Can we ask a few too? Where, for instance, is the money you had on the floor a little while ago?”\\

“In the suitcase on top of the wardrobe,” answered Pippi truthfully.\\

Thunder-Karlsson and Bloom grinned.\\

“I hope you don’t have anything against our taking it, little friend,” said Thunder-Karlsson.\\

“Certainly not,” said Pippi. “Of course I don’t.” Whereupon Bloom lifted down the suitcase.\\

“I hope you don’t have anything against my taking it back, little friend,” said Pippi, getting out of bed and stepping over to Bloom.\\

Bloom had no idea how it all happened, but suddenly the suitcase was in Pippi’s hand.\\

“Here, quit your fooling!” said Thunder- Karlsson angrily. “Hand over the suitcase.” He took Pippi firmly by the hand and tried to snatch back the booty.\\

“Fooling, fooling, too much fooling,” said Pippi and lifted Thunder-Karlsson up on the wardrobe. A moment later she had Bloom up there too. Then the tramps were frightened; they began to see that Pippi was no ordinary girl. However, the suitcase tempted them so much they forgot their fright.\\

“Come on now, both together,” yelled Thunder- Karlsson, and they jumped down from the wardrobe and threw themselves on Pippi, who had the suitcase in her hand. Pippi gave each one a little poke with her finger, and they shrank away into a corner. Before they had a chance to get up again, Pippi had fetched a rope and quick as a flash had bound the arms and legs of both burglars. Now they sang a different tune.\\

“Please, please, miss,” begged Thunder-Karlsson, “forgive us. We were only joking. Don’t hurt us. We are just two tramps who came in to ask for food.”\\

Bloom even began to cry a bit.\\

Pippi put the suitcase neatly back on the wardrobe. Then she turned to her prisoners. “Can either of you dance the schottischel?”\\

“Why, yes,” said Thunda-Karlsson. “I guess we both can.” “Oh, what fun I” cried Pippi, clapping her hands.\\

“Can’t we dance a little? I’ve just learned, you know.\\

“Well, certainly, by all means,” said Thunder- Karlsson, a bit confused. Pippi took some large scissors and cut the ropes that bound her guests.\\

“But we don’t have any music,” she said in a worried voice. Then she had an idea. “Can’t you blow on a comb?” she said to Bloom. “And I’ll dance with him " She pointed to Thunder-Karlsson.\\

Oh, yes, Bloom could blow on a comb, all right. And blow he did, so that you could hear it all through the house. Mr. Nilsson sat up in bed, wide-awake, just in time to see Pippi whirling around with Thunder-Karlsson. She was dead serious and danced as if her life depended on it.\\

At last Bloom said he couldn’t blow on the comb any longer because it tickled his mouth unmercifully. And Thunder-Karlsson, who had tramped the roads all day, began to feel tired.\\

“Oh, please, just a little longer,” begged Pippi, dancing on, and Bloom and Thunda-Karlsson could do nothing but continue. At three in the morning Pippi said, “I could keep on dancing until Thursday, but maybe you’re tired and hungry.\\

That was exactly what they were, though they hardly dared to say so. Pippi went to the pantry and took out bread and cheese and butter, ham and cold roast and milk; and they sat around the kitchen table--Bloom and Thunder-Karlsson and Pippi--and ate until they were almost four-cornered.\\

Pippi poured a little milk into her ear. “That’s good for earache,” she said.\\

“Poor thing, have you got an earache?” asked Bloom.\\

“No,” said Pippi, “but I might get one.\\

Finally the two tramps got up, thanked Pippi for the food, and begged to be allowed to say good-by.\\

“It was awfully jolly that you came. Do you really have to go so soon?” said Pippi regretfully. “Never have I seen anyone who can dance the schottische the way you do, my sugar pig,” she said to Thunder-Karlsson. And to Bloom, “If you keep on practicing on the comb, you won’t notice the tickling.”\\

As they were going out of the door Pippi came running after them and gave them each a gold piece. “These you have honestly earned,” she said.
\newpage

\section{Pippi Goes to a Coffee Party}
Tommy’s and Annika’s mother had invited a few ladies to a coffee party, and as she had done plenty of baking, she thought Tommy and Annika might invite Pippi over at the same time. The children would entertain each other and give no trouble to anyone.\\

Tommy and Annika were overjoyed when their mother told them and immediately dashed over to Pippi’s to invite her. Pippi was in the garden, watering the few flowers still in bloom with an old rusty watering can. As it was raining cats and dogs that day Tommy told Pippi her watering seemed hardly necessary.\\

“Yes, that’s what you say,” said Pippi grudgingly, “but I’ve lain awake all night thinking what fun it was going to be to get up and water, and I’m not going to let a little rain stand in my way.”\\

Now Annika came forth with the delightful news about the coffee party.\\

“A coffee party! Me!” cried Pippi, and she was so excited that she began to water Tommy instead of the rosebush she intended to sprinkle. “Oh, what will happen? Oh, I’m so nervous. What if I can’t behave myself?”\\

“Of course you can,” said Annika.\\

“Don’t you be too certain about that,” said Pippi. “You can be sure I’ll try, but I have noticed several times that people don’t think I know how to behave even when I’m trying as hard as I can. At sea we were never so fussy about things like that. But I promise that I’ll take special pains today so you won’t have to be ashamed of me.”\\

“Good,” said Tommy, and he and Annika hurried home again in the rain.\\

“This afternoon at three o’clock, don’t forget,” cried Annika, peeking out from under the umbrella.\\

At three o’clock a very stylish young lady walked up the steps of the Settergrens’ house. It was Pippi Longstocking. For this special occasion she had unbraided her pigtails, and her red hair hung like a lion’s mane around her.\\

With red crayon she had painted her mouth fiery red and she had blackened her eyebrows so that she looked almost dangerous. With the crayon she had also painted her fingernails, and she had put big green rosettes on her shoes.\\

“I should imagine I’ll be the most stylish person of all at this party,” she said contentedly to herself as she rang the doorbell.\\

In the Settergrens’ living room sat three fine ladies, and with them Tommy, Annika, and their mother. A wonderful coffee table had been spread, and in the fireplace a fire was burning brightly. The ladies were talking quietly with one another, and Tommy and Annika were sitting on the sofa, looking at an album. Everything was so peaceful.\\

Suddenly the peace was shattered.\\

“Atten-shun!” A piercing cry came from the hall, and the next minute Pippi Longstocking stood in the doorway. She had cried out so loudly and so unexpectedly that the ladies had jumped in their seats.\\

“Forward march!” came the next command, and Pippi, with measured steps, walked up to Mrs. Settergren.\\

“Halt!” She stopped. “Arms forward, one, two, she cried and with both hands gripped one of Mrs. Settagrenns and shook it heartily.\\

“Knees bend!” she shrieked and curtsied prettily. Then she smiled at Mrs. Settergren and said in her ordinary voice, “You see, I am really very shy, so if I didn’t give myself some commands I’d just stand in the hall and not dare to come in.\\

Then she rushed up to the other ladies and kissed them on the cheek.\\

“Charming, charming, upon my honor!” said she, for she had once heard a stylish gentleman say that to a lady. Then she sat down in the best chair she could find. Mrs. Settergren had intended the children to have their party up in Tommy’s and Annika’s room, but Pippi stayed calmly in her chair, slapped herself on the knee, and said, looking at the coffee table, “That certainly looks good. When do we begin?”\\

At that moment Ella, the maid, came in with the coffeepot, and Mrs. Settergren said, “Please come and have some coffee.”\\

“First!” cried Pippi and was up by the table in two skips. She heaped as many cakes as she could onto a plate, threw five lumps of sugar into a coffee cup, emptied half the cream pitcher into her cup, and was back in her chair with her loot even before the ladies had reached the table.\\

Pippi stretched her legs out in front of her and placed the plate of cakes between her toes. Then she merrily clunked cakes in her coffee cup and stuffed so many in her mouth at once that she couldn’t have uttered a word no matter how hard she tried. In the twinkling of an eye she had finished all the cakes on the plate. She got up, struck the plate as if were a tambourine, and went up to the table to see if there were any cakes left. The ladies looked disapprovingly at her, but that didn’t bother her. Chatting gaily, she walked around the table, snatching a cake here and a cake there.\\

“It certainly was nice of you to invite me,” she said. “I’ve never been to a coffee party before.”\\

On the table stood a large cream pie, decorated in the center with a piece of red candy. Pippi stood with her hands behind her back and looked at it. Suddenly she bent down and snatched the candy with her teeth. But she dived down a little too hastily, and when she came up again her whole face was covered with whipped cream.\\

“Goody!” laughed Pippi. “Now we can play blindman’s buff, for we’ve certainly got a blind man all made to order! I can’t see a thing!”\\

She stuck out her tongue and licked away the cream. “This was indeed a dreadful accident,” said she. “And the pie is all ruined now anyway, so I may as well eat it all up at once.\\

She dug into it with the pie server, and in a few minutes the whole pie had disappeared. Pippi patted her stomach contentedly. Mrs. Settergren had gone out into the kitchen so she knew nothing about the accident to the cream pie, but the other ladies looked very sternly at Pippi. No doubt they would have liked a little pie too. Pippi noticed that they looked disappointed and decided to cheer them up.\\

“Now you mustn’t feel bad about such a little accident,” she said comfortingly. “The main thing is that we have our health. And at a coffee party you should have fun.”\\

She then picked up a sugar bowl and tipped all the lump sugar in it out on the door. “Well, my goodness!” she cried. “Now look what I’ve done! How could I make such a mistake? I thought this was the granulated sugar. Bad luck seems to follow me today.”\\

Thereupon she took a sugar spoon out of another bowl and began to sprinkle granulated sugar all over the floor. “I hope you notice,” she said, “that this is the kind of sugar you sprinkle on things. So it’s perfectly all right for me to do this. Because why should there be the kind of sugar to sprinkle on things if somebody doesn’t go and sprinkle it?--that’s what I’d like to know.\\

“Have you ever noticed what fun it is to walk on a floor that has had sugar sprinkled all over it?” she asked the ladies. “Of course it’s even more fun when you’re barefoot,” she added as she pulled off her shoes and stockings. “You ought to try it too, because nothing’s more fun, believe me.”\\

At that moment Mrs. Settergren came in, and when she saw the sugar all over the floor she took Pippi firmly by the arm and led her over to the sofa to Tommy and Annika. Then she went over to the ladies and invited them to have more coffee. That the cream pie had disappeared only made her happy, because she thought the ladies had liked it so much that they had eaten it all.\\

Pippi, Tommy, and Annika sat talking quietly on the sofa. The he crackled on the hearth. The ladies drank their coffee, and all was quiet and peaceful again. And as so often happens at coffee parties, the ladies began to talk about their servant problems. Apparently they had not been able to get very good savants, for they were not at all satisfied with them, and they agreed that it really was better not to have any servants at all. It was much more satisfactory to do things yourself because then you at least knew that things were done right.\\

Pippi sat on the sofa listening, and after the ladies had been talking a while she said, “Once my grandmother had a servant named Malin. She had chilblains on her feet, but otherwise there was nothing wrong with her. The only annoying thing was that as soon as company came she would rush at them and bite their legs. And bark! Oh, how she would bark! You could hear it all through the neighborhood, but it was only because she was playful. Only, of course, strangers didn’t always understand that. The dean’s wife, an elderly woman, came to see Grandmother once soon after Malin first came, and when Malin came dashing at her and bit her in the ankle, the dean’s wife screamed so loudly that it scared Malin, so that her teeth clamped together and she couldn’t get them apart. There she sat, stuck to the dean’s wife’s ankle until Friday. And Grandmother had to peel the potatoes herself. But at least it was well done. She peeled so well that when she was done there were no potatoes left--only peelings. But after that Friday the dean’s wife never came to call on Grandmother again. She just never could take a joke. And poor Malin who was always so good-natured and Happy! Though for that matter she was a little touchy at times, there’s no denying that. Once when Grandmother poked a fork in her ear she howled all day.\\

Pippi looked around and smiled pleasantly.\\

“Yes, that was Malin for you,” she said and twiddled her thumbs.\\

The ladies acted as if they had heard nothing. They continued to talk.\\

“If my Rosa were only clean,” said Mrs. Berggren, “then maybe I could keep her. But she’s a regular pig--"\\

“Say, you ought to have seen Malin,” Pippi interrupted. “Malin was so outrageously dirty that it was a joy to see her, Grandmother said. For the longest time Grandmother thought she had a very dark complexion but, honest and true, it was nothing but dirt that would wash off. And once at a bazaar at the City Hotel she got first prize for the dirt under her nails. Mercy me, how dirty that girl was!” said Pippi happily. Mrs. Settergren looked at her sternly.\\

“Can you imagine!” said Mrs. Granberg. “The other evening when Britta was going out she borrowed my blue satin dress without even asking for it. Isn’t that dreadful?”\\

“Yes, indeed,” said Pippi, “she certainly seems to be cut from the same piece of cloth as Malin, from what you say. Grandmother had a pink undershirt that she was especially fond of. But the worst of it was that Malin liked it too. And every morning Grandmother and Malin argued about who was to wear the undershirt. At last they decided it would be fair to take turns and each wear it every other day. But imagine how tricky Malin could be! Sometimes she’d come running in when it wasn’t her turn at all and say, “No mashed turnip today if I can’t wear the pink woolen undershirt!’ Well, what was Grandmother to do? Mashed turnip was her very favorite dish. There was nothing for it but to give Malin the shirt. As soon as Malin got the shirt she went out into the kitchen as nice as could be and began to mash turnip so that it spattered all over the walls.”\\

There was silence for a little while, and then Mrs. Alexandersson said, “I’m not absolutely certain but I strongly suspect that my Hulda steals. In fact I’ve noticed that things disappear.\\

“Malin,” began Pippi, but Mrs. Settergren interrupted her. “Children,” she said decidedly, “go up to the nursery immediately!”\\

“Yes, but I was only going to tell that Malin stole too,” said Pippi. “Like a raven! Everything she could lay her hands on. She used to get up in the middle of the night and steal; otherwise she couldn’t sleep well, she said. Once she stole Grandmother’s piano and tucked it into her own top bureau drawer. She was very clever with her hands, Grandmother said.”\\

Tommy and Annika took hold of Pippi and pulled her out of the room and up the stairs. The ladies began on their third cups of coffee, and Mrs. Settergren said, “It’s not that I want to complain about my Ella, but she does break the china.”\\

A red head appeared over the stair rail.\\

“Speaking of Malin,” said Pippi, “maybe you are wondering if she used to break any china. Well, she did. She set apart one day a week just to break china. It was Tuesday, Grandmother said. As early as five o’clock on Tuesday morning you could hear this jewel of a maid in the kitchen, breaking china. She began with the coffee cups and glasses and little things like that and then went on to the soup bowls and dinner plates, and she finished up with platters and soup tureens. There was such a crash bang in the kitchen all morning that it was a joy to hear it, Grandmother said. And if Malin had any spare time late in the afternoon, she would go into the drawing room with a little hammer and knock down the antique East Indian plates that were hanging on the walls. Grandmother bought new china every Wednesday,” said Pippi and disappeared up the stairs as quickly as a jack-in-the-box.\\

But now Mrs. Settergen’s patience had come to an end. She ran up the stairs, into the nursery, and up to Pippi, who had just begun to teach Tommy to stand on his head.\\
“You must never come here again,” said Mrs. Settergren, “if you can’t behave any better than this.”\\

Pippi looked at her in astonishment and her eyes slowly filled with tears. “That’s just what I was afraid of,” she said. “That I couldn’t behave properly. It’s no use to try; I’ll never learn. I should have stayed on the ocean.”\\

She curtsied to Mrs. Settergren, said good-by to Tommy and Annika, and went slowly down the stairs.\\

The ladies were now getting ready to go home too. Pippi sat down in the hall near the shelf where rubbers were kept and watched the ladies putting on their hats and coats.\\

“Too bad you don’t like your maids,” said she. “You should have one like Malin. Grandmother always said there was nobody like her. Imagine! One Christmas when Malin was going to serve a little roast pig, do you know what she did? She had read in the cookbook that roast pig must be served with frilled paper in the ears and an apple in the mouth. But poor Malin didn’t understand that it was the pig who was supposed to have the apple. You should have seen her when she came in on Christmas Eve with her best apron on and a big Gravenstein apple in her mouth. ‘Oh, Malin, you’re crazy!’ Grandmother said to her, and Malin couldn’t say a word in her own defense, she could only wiggle her ears until the frilled paper rustled. To be sure, she tried to say something, but it just sounded like blubb, blubb, blubb. And of course she couldn’t bite people in the leg as she usually did, and it would be a day when there was a lot of company. Poor little Malin, it wasn’t a very happy Christmas Eve for her,” said Pippi sadly.\\

The ladies were now dressed and said a last good-by to Mrs. Settagren. And Pippi ran up to her and whispered, “Forgive me because I couldn’t behave myself. Good-by!”\\

Then she put on her large hat and followed the ladies. Outside the gate their ways parted. Pippi went toward Villa Villekulla and the ladies in the other direction.\\

When they had gone a little way they heard someone panting behind them. It was Pippi who had come racing back.\\

“You can imagine that Grandmother mourned when she lost Malin. Just think, one Tuesday morning when Malin had had time to break only about a dozen teacups she ran away and went to sea. And Grandmother had to break the china herself that day. She wasn’t used to it, poor thing, and she got blisters all over her hands. She never saw Malin again. And that was a shame because she was such an excellent maid, Grandmother said.”\\

Pippi left, and the ladies hurried on, but when they had gone a couple of hundred feet, they heard Pippi, from far off, yelling at the top of her lungs, “SHE NEVER SWEPT UNDER THE BEDS!”
\newpage

\section{Pippi Acts as a Life Saver}
One Sunday afternoon Pippi sat wondering what to do. Tommy and Annika had gone to a tea party with their mother and father, so she knew she couldn’t expect a visit from them.\\

The day had been filled with pleasant tasks. She had got up early and served Mr. Nilsson fruit juice and buns in bed. He looked so cute sitting there in his light blue nightshirt, holding the glass in both hands. Then she had fed and groomed the horse and told him a long story of her adventures at sea. Next she had gone into the parlor and painted a large picture on the wall- paper. The picture represented a fat lady in a red dress and a black hat. In one hand she held a yellow flower and in the other a dead rat. Pippi thought it a very beautiful picture; it dressed up the whole room. Then she had sat down in front of her chest and looked at all her birds’ eggs and shells, and thought about the wonderful places where she and her father had collected them, and about all the pleasant little shops all over the world where they had bought the beautiful things that were now in the drawers of her chest. Then she had tried to teach Mr. Nilsson to dance the schottische, but he didn’t want to learn. For a while she had thought of trying to teach the horse, but instead she had crept down into the woodbox and pulled the cover down. She had pretended she was a sardine in a sardine box, and it was a shame Tommy and Annika weren’t there so they could have been sardines too.\\

Now it had begun to grow dark. She pressed her little pug nose against the windowpane and looked out into the autumn dusk. She remembered that she hadn’t been riding for a couple of days and decided to go at once. That would be a nice ending to a pleasant Sunday.\\

Accordingly she put on her big hat, fetched Mr. Nilsson from a corner where he sat playing marbles, saddled the horse, and lifted him down from the porch. And off they went Mr. Nilsson on Pippi and Pippi on the horse.\\

It was quite cold and the roads were frozen, so there was a good crunchy sound as they rode along. Mr. Nilsson sat on Pippi’s shoulder and tried to catch hold of some of the branches of the trees as they went by, but Pippi rode so fast that it was no use. Instead, the branches kept boxing him in the ears, and he had a hard time keeping his straw hat on his head.\\

Pippi rode through the little town, and people pressed arduously up against the walls when she came storming by.\\

The town had a market square, of course. There were several charming old one- story buildings and a little yellow-painted town hall. And there was also an ugly wretch of a building, newly built and three stories high. It was called “The Skyscraper” because it was taller than any of the other houses in town.\\

On a Sunday afternoon the little town was always quiet and peaceful, but suddenly the quiet was broken by loud cries.” The Skyscraper’s burning! Fire! Fire!”\\

People came running excitedly from all directions. The fire engine came clanging down the street, and the little children who usually thought fire engines were such fun now cried from fright because they were sure their own houses would catch fire too. The police had to hold back the crowds of people gathering in the square so that the fire engine could get through. The flames came leaping out of the windows of the Skyscraper, and smoke and sparks enveloped the firemen who were courageously trying to put out the fire. The fire had started on the first floor but was quickly spreading to the upper stories.\\

Suddenly the crowd saw a sight that made them gasp with horror. At the top of the house was a gable, and in the gable window, which a little child’s hand had just opened, stood two little boys calling for help. “We can’t get out because somebody has built a fire on the stairs,” cried the older boy.\\

He was five and his brother a year younger. Their mother had gone out on an errand, and there they stood, all alone. Many of the people in the square began to cry, and the fire chief looked worried. There was, of course, a ladder on the fire truck, but it wouldn’t reach anywhere near to the little boys. To get into the house to save the children was impossible. A wave of despair swept over the crowd in the square when they realized there was no way to help the children. And the poor little things just stood up there and cried. It wouldn’t be long now before the fire reached the attic.\\

In the midst of the crowd in the square sat Pippi on her horse. She looked with great interest at the fire engine and wondered if she should buy one like it. She liked it because it was red and because it made such a fearful noise as it went through the streets. Then she looked at the fire and she thought it was fun when a few sparks fell on her.\\

Presently she noticed the little boys up in the attic. To her astonishment they looked as if they weren’t enjoying the fire at all. That was more than she could understand, and at last she had to ask the crowd around her, “Why are the children crying?”\\

First she got only sobs in answer, but finally a stout gentleman said, “Well, what do you think? Don’t you suppose you’d cry yourself if you were up there and couldn’t get down?”\\

“I never cry,” said Pippi. “But if they want to get down, why doesn’t somebody help them?”\\

“Because it isn’t possible, of course,” said the stout gentleman.\\

Pippi thought for a while. Then she asked, “Can anybody bring me a long rope?”\\

“What good would that do?” asked the stout gentleman. “The children are too small to get down the rope, and, for that matter, how would you ever get the rope up to them?”\\

“Oh, I’ve been around a bit,” said Pippi calmly. “I want a rope.”\\

There was not a single person who thought it would do any good, but somehow or other Pippi got her rope.\\

Not far from the gable of the Skyscraper grew a tall tree. The top of it was almost level with the attic window, but between the tree and the window was a distance of almost three yards. And the trunk of the tree was smooth and had no branches for climbing on. Even Pippi wouldn’t be able to climb it.\\

The fire burned. The children in the window screamed. The people in the square cried.\\

Pippi jumped off the horse and went up to the tree. Then she took the rope and tied it tightly to Mr. Nilsson’s tail.\\

“Now you be Pippi’s good boy,” she said. She put him on the tree trunk and gave him a little push. He understood perfectly what he was supposed to do. And he climbed obediently up the tree trunk. Of course it was no trouble at all for a little monkey to do that.\\

The people in the square held their breath and watched Mr. Nilsson. Soon he had reached the top of the tree. There he sat on a branch and looked down at Pippi. She beckoned to him to come down again. He did so at once, climbing down on the other side of the branch, so that when he reached the ground the rope was looped over the branch and hung down double with both ends on the ground.\\

“Good for you, Mr. Nilsson,” said Pippi. “You’re so smart you can be a professor any time you wish.” She untied the knot that had fastened the rope to Mr. Nilsson’s tail.\\

Nearby, a house was being repaired, and Pippi ran over and got a long board. She took the board in one hand, ran to the tree, grasped the rope in her free hand, and braced her feet against the trunk of the tree. Quickly and nimbly she climbed up the trunk, and the people stopped crying in astonishment. When she reached the top of the tree she placed the board over a stout branch and then carefully pushed it over to the windowsill. And there lay the board like a bridge between the top of the tree and the window.\\

The people down in the square stood absolutely silent. They were so tense they couldn’t say a word. Pippi stepped out on the board. She smiled pleasantly at the two boys in the gabled window. “Why do you look so sad?” she asked. “Have you got a stomach-ache?”\\

She ran across the board and hopped in at the window. “My, it seems warm in here,” she said. “You don’t need to make any more fire in here today, that I can guarantee. And at the most four sticks in the stove tomorrow, I should think.”\\

Then she took one boy under each arm and stepped out on the board again.\\

“Now you’re really going to have some fun,” she said. “It’s almost like walking the tight rope.”\\

When she got to the middle of the board she lifted one leg in the air just as she had done at the circus. The crowd below gasped, and when a little later Pippi lost one of her shoes several old ladies fainted. However, Pippi reached the tree safely with the little boys. Then the crowd cheered so loudly that the dark night was filled with noise and the sound drowned out the crackling of the fire.\\

Pippi hauled up the rope, fastened one end securely to a branch and tied the other around one of the boys. Then she let him down slowly and carefully into the arms of his waiting mother, who was beside herself with joy when she had him safe. She held him close and hugged him, with tears in her eyes.\\

But Pippi yelled, “Untie the rope, for goodness’ sake? There’s another kid up here, and he can’t fly either. So the people helped to untie the rope and free the little boy. Pippi could tie good knots -- she could indeed. She had learned that at sea. She pulled up the rope again, and now it was the second boy’s turn to be let down.\\

Pippi was alone in the tree. She sprang out on the board, and all the people looked at her and wondered what she was going to do. She danced back and forth on the narrow board. She raised and lowered her arms gracefully and sang in a hoarse voice that could barely be heard down in the square:\\

\begin{center}
The Fire is burning,\\
It’s burning so bright,\\
The flumes are leaping and prancing.\\
It’s burning for you,\\
It’s burning for me,\\
It’s burning for all who are dancing\\
\end{center}
\noindent \\

As she sang she danced more and more wildly until many people covered their eyes in horror for they were sure she would fall down and kill herself. Flames came leaping out of the gable window, and in the firelight people could see Pippi plainly. She raised her arms to the night sky, and while a shower of sparks fell over her she cried loudly, “Such a jolly, jolly fire!”\\

She took one leap and caught the rope. “Look out!” she cried and same sliding down the rope like greased lightning.\\

“Three cheers for Pippi Longstocking! Long may she live!” cried the fire chief.\\

“Hip, hip, hurray! Hip, hip, hurray! Hip, hip, hurray!” cried all the people-- three times. But there was one person there who cheered four times.\\

It was Pippi Longstocking.
\newpage

\section{Pippi Celebrates Her Birthday}
One day Tommy and Annika found a letter in their mailbox.
It was addressed to TMMY and ANIKA, and when they opened it they found a card which read:\\

\begin{center}
TMMY AND ANIKA ARE INVITED TO PIPPI’S TOMORO TO HER BERTHDAY PARTY. DRES: WARE WATEVER YOU LIK.
\end{center}
\noindent\\

Tommy and Annika were so happy they began to skip and dance. They understood perfectly well what was printed on the card although the spelling was a little unusual. Pippi had had a great deal of trouble writing it. To be sure, she had not recognized the letter i in school the day she was there, but all the same she could write a little. When she was sailing on the ocean one of the sailors on her father’s ship used to take her up on deck in the evening now and then and try to teach her to write. Unfortunately Pippi was not a very patient pupil. All of a sudden she would say, “No, Fridolf”--that was his name-- “no, Fridolf, bother all this learning! I can’t study any more now because I must climb the mast to see what kind of weather we’re going to have tomorrow.\\

So it was no wonder the writing didn’t go so well now. One whole night she sat struggling with that invitation, and at dawn, just as the stars were paling in the sky over Villa Villekulla, she tiptoed over to Tommy’s and Annika’s house and dropped the letter into their mailbox.\\

As soon as Tommy and Annika came home from school they began to get all dressed up for the party. Annika asked her mother to curl her hair, and her mother did, and tied it with a big pink satin bow. Tommy combed his hair with water so that it would lie all nice and smooth. He certainly didn’t want any curls. Then Annika wanted to put on her very best dress, but her mother thought she’d better not for she was seldom neat and clean when she came home from Pippi’s; so Annika had to be satisfied with her next best dress. Tommy didn’t care what suit he wore so long as he looked nice.\\

Of course they had bought a present for Pippi. They had taken the money out of their own piggy banks, and on the way home from school had run into the toy shop on Main Street and bought a very beautiful--well, what they had bought was a secret for the time being. There it lay, mapped in green paper and tied with a great deal of string, and when they were ready Tommy took the package, and off they went, followed by their mother’s warning to take good care of their clothes. Annika was to carry the package part of the way, and they were both to hold it when they handed it to Pippi--that they had agreed upon.\\

It was already November, and dusk came early. When Tommy and Annika went in through the gate of Villa Villekulla they held each other’s hands tightly, because it was quite dark in Pippi’s garden and the wind sighed mournfully through the bare old trees. “Seems like fall,” said Tommy. It was so much pleasanter to see the lighted windows in Villa Villekulla and to know that they were going to a birthday party.\\

Ordinarily Tommy and Annika rushed in through the kitchen door, but this time they went to the front door. The horse was not on the porch. Tommy gave a lively knock on the door.
From inside came a low voice:\\

\begin{center}
Who comes in the dark night\\
On the road to my house?\\
Is it a ghost or just\\
A poor little mouse?
\end{center}
\noindent \\

“No, no, Pippi, it’s us,” shrieked Annika. “Open the door!”\\

Pippi opened the door. “Oh, Pippi, why did you say that about a ghost? I was so scared,” said Annika and completely forgot to congratulate Pippi.\\

Pippi laughed heartily and opened the door to the kitchen. How good it was to come in where it was light and warm! The birthday party was to be in the kitchen, because that was the pleasantest room in the house. There were only two other rooms on the first floor, the parlor--in which there was only one piece of furniture - and Pippi’s bedroom. The kitchen was large and roomy, and Pippi had scrubbed it until it shone. She had put rugs on the floor and a large new cloth on the table. She had embroidered the cloth herself with flowers that certainly looked most remarkable, but Pippi declared that such flowers grew in Farthest India, so of course that made them all right. The curtains were drawn and the fire burned merrily. On the woodbox sat Mr. Nilsson, banging pot lids together. In a corner stood the horse, for he too had been invited to the party.\\

Now at last Tommy and Annika remembered that they were supposed to congratulate Pippi. Tommy bowed and Annika curtsied, and then they handed Pippi the green package and said, “May we congratulate you and wish you a happy Birthday?” Pippi thanked them and eagerly tore the package open. And there was a music box! Pippi was wild with delight. She patted Tommy and she patted Annika and she patted the music box and she paned the wrapping paper. She wound up the music box, and with much plinking and plonking out came a melody that was probably supposed to be “Ack, du klire Augustin.”\\

Pippi wound and wound and seemed to forget everything else. But suddenly she remembered something. “Oh, my goodness, you must have your birthday presents too!” she said.\\

“But it isn’t our birthday,” said Tommy and Annika.\\

Pippi stared at them in amazement. “No, but it’s my birthday, isn’t it? And so I can give birthday presents too, can’t I? Or does it say in your schoolbooks that such a thing can’t be done? Is it something to do with that old pluttifikation that makes it impossible?”\\

“Oh, of course it’s possible,” said Tommy. “It just isn’t customary. But for my part, I’d be very glad to have a present.”\\

“Me too,” said Annika.\\

Pippi ran into the parlor and brought back two packages from the chest. When Tommy opened his he found a little ivory flute, and in Annika’s package was a lovely brooch shaped like a butterfly. The wings of the butterfly were set with blue and red and green stones.\\

When they had all had their birthday presents it was time to sit down at the table, where there were all sorts of cakes and buns. The cakes were rather peculiar in shape, but Pippi declared they were just the kind of cakes they had in China.\\

Pippi served hot chocolate with whipped cream, and the children were just about to begin their feast when Tommy said, “When Mamma and Papa have a party the gentlemen always get cards telling them what ladies to take in to dinner. I think we ought to have cards too.”\\

“Okay,” said Pippi.\\

“Although it will be kind of hard for us because I’m the only gentleman here,” added Tommy doubtfully.\\

“Fiddlesticks,” said Pippi. “Do you think Mr. Nilsson is a lady, maybe?”\\

“Oh, of course not, I forgot Mr. Nilsson,” said Tommy, and he sat down on the woodbox and wrote on a card:\\

Mr. Settergren will have the pleasure of taking Miss Longstocking in to dinner.\\

“Mr. Settergren, that’s me,” said he with satisfaction and showed Pippi the card. Then he wrote on the next card:\\

\begin{center}
Mr. Nilsson will have the pleasure of\\
taking Miss Settergren in to dinner.
\end{center}
\noindent \\

“Okay, but the horse must have a card too,” said Pippi decidedly, “even if he can’t sit at the table.”\\

So Tommy, at Pippi’s dictation, wrote:\\

The horse will have the pleasure of remaining in the center where he will be served cakes and sugar.\\

Pippi held the card under the horse’s nose and said, “Read this and see what you think of it.”\\

As the horse had no objection to make, Tommy offered Pippi his arm, and they walked to the table. Mr. Nilsson showed no intention of offering his arm to Annika, so she took a firm hold of him and lifted him up to the table. But he didn’t want to sit on a chair; he insisted on sitting right on the table. Nor did he want any chocolate with whipped cream, but when Pippi poured water in his cup he took it in both his hands and drank.\\

Annika and Tommy and Pippi ate and ate, and Annika said that if these cakes were the kind they had in China, then she intended to move to China when she grew up.\\

When Mr. Nilsson had emptied his cup he turned it upside down and put it on his head. When Pippi saw that, she did the same, but as she had not drunk quite all her chocolate a little stream ran down her forehead and over her nose. She caught it with her tongue and lapped it all up.\\

“Waste not, want not,” she said.\\

Tommy and Annika licked their cups clean before they put them on their heads.\\

When everybody had had enough and the horse had had his share, Pippi took hold of all four-corners of the tablecloth and lifted it up so that the cups and plates tumbled over each other as if they were in a sack. Then she stuffed the whole bundle in the woodbox.\\

“I always like to tidy up a little as soon as I have eaten,” she said.\\

Then it was time for games. Pippi suggested that they play a game called “Don’t touch the floor.” It was very simple. The only thing one had to do was walk all around the kitchen without once stepping on the floor. Pippi skipped around in the twinkling of an eye, and even for Tommy and Annika it was quite easy. You began on the drain board, and if you stretched your legs enough it was possible to step onto the back of the stove. From the stove to the woodbox, and from the woodbox to the hat shelf, and down onto the table, and from there across two chairs to the corner cupboard. Between the corner cupboard and the drain board was a distance of several feet, but, luckily, there stood the horse, and if you climbed up on him at the tail end and slid off at the head end, making a quick turn at exactly the right moment, you landed exactly on the drain-board.\\

When they had played this game for a while, and Annika’s dress was no longer her next-best dress but her next-next-next-best one, and Tommy had become as black as a chimney sweep, then they decided to think up something else.\\

“Suppose we go up in the attic and visit the ghosts,” suggested Pippi.\\

Annika gasped. “A-a-are there really ghosts in the attic?” she asked.\\

“Are there ghosts? Millions!” said Pippi. “It’s just swarming with all sorts of ghosts and spirits. You trip over them wherever you walk. Shall we go up?”\\

“Oh, Pippi!” said Annika and looked reproachfully at her.\\

“Mamma says there aren’t any such things as ghosts and goblins,” said Tommy boldly.\\

“And well she might,” said Pippi, “because there aren’t any anywhere else. All the ghosts in the world live in my attic. And it doesn’t pay to try to make them move. But they aren’t dangerous. They just pinch you in the arm so you get black and blue, and they howl, and they play ninepins with their heads.”\\

“Do- do--do they really play n-n-ninepins with their heads?”\\

“Sure, that’s just what they do,” said Pippi. “Come on, let’s go up and talk with them. I’m good at playing n-n-ninepins.”\\

Tommy didn’t want to show that he was frightened, and in a way he really did want to see a ghost. That would be something to tell the boys at school! Besides, he consoled himself with the thought that the ghosts probably wouldn’t dare to hurt Pippi. He decided to go along. Poor Annika didn’t want to go under any circumstances, but then she happened to think that a little tiny ghost might sneak downstairs while she was sitting alone in the kitchen. That decided the matter. Better to be with Pippi and Tommy among thousands of ghosts than alone in the kitchen with even the tiniest little ghost child.\\

Pippi went first. She opened the door to the attic stairs. It was pitch-dark there. Tommy took a firm grip on Pippi, and Annika took an even firmer grip on Tommy, and so they went up. The stairs creaked and squeaked with every step. Tommy began to wonder if it wouldn’t have been better to stay down in the kitchen, and Annika didn’t need to wonder she was sure of it. At last they came to the top of the stairs and stood in the attic. It was pitch-dark there too, except where a little moonbeam shone on the floor. There were sighs and mysterious noises in every comer when the wind blew in through the cracks.\\

“Hi, all you ghosts!” shrieked Pippi.\\

But if there was any ghost there he certainly didn’t answer.\\

“Well, I might have known,” said Pippi, “they’ve gone to a council meeting of the Ghost and Goblin Society.”\\

Annika sighed with relief and hoped that the meeting would last a long time. But just then an awful sound came from one of the comers of the attic.\\

“Whooooooool” it said, and a moment later Tommy saw something come rushing toward him in the dimness. He felt it brush his forehead and saw something disappear through a little window that stood open.\\

He shrieked to high heaven, “A ghost! A ghost!” And Annika shrieked with him.\\

‘That poor thing will be late for the meeting,” said Pippi. “If it was a ghost. And not an owl. For that matter, there aren’t any ghosts,” she continued after a while. “If anybody insists that there are ghosts, I’ll tweak him in the nose.”\\

“Yes, but you said so yourself,” said Annika.\\

“Is that so? Did I?” said Pippi. “Well, then I’ll certainly tweak my own nose.” And she took a firm grip on her nose and tweaked it.\\

After that Tommy and Annika felt a little calmer. In fact they were now so courageous that they ventured to go up to the window and look out over the garden. Big dark clouds sailed through the sky and did their best to hide the moon. And the wind sighed in the trees. Tommy and Annika turned around. But then -- Oh, horrors -- they saw a white figure coming toward them.\\

“A ghost!” shrieked Tommy wildly.\\

Annika was so scared she couldn’t even shriek. The ghost came nearer and nearer. Tommy and Annika hugged each other and shut their eyes.\\

But then they heard the ghost say, “Look what I found! Papa’s nightshirt in an old sea chest over here. If I hem it up around the bottom I can wear it.”\\

Pippi came up to them with the nightshirt dangling around her legs. “Oh, Pippi, I could have died of fright,” said Annika.\\

“But nightshirts aren’t dangerous,” Pippi assured her. “They don’t bite anybody except in self-defense.”\\

Pippi now decided to examine the sea chest thoroughly. She lifted it up and carried it over to the window and opened the cover, so that what little moonlight there was fell on the contents of the chest. There were a great many old clothes, which she threw out on the attic floor. There were a telescope, a few books, three pistols, a sword, and a bag of gold pieces.\\

“Tiddelipom and piddeliday,” said Pippi contentedly. “It’s so exciting!” said, Tommy.\\

Pippi gathered everything in the nightshirt, and down they went into the kitchen again. Annika was perfectly satisfied to leave the attic.\\

“Never let children handle firearms,” said Pippi and took a pistol in each hand and prepared to fire. “Otherwise some accident can easily happen,” she said, shooting off both pistols at once. “That was a good bang,” she announced and looked up in the ceiling. The bullets had made two holes.\\

“Who knows?” she said hopefully. “Perhaps the bullets have gone right through the ceiling and hit some ghosts in the legs. That will teach them to think twice before they set out to scare any innocent little children again. Because even if there aren’t any ghosts, they don’t need to go round scaring folks out of their wits, I should think. Would you each like a pistol?” she asked.\\

Tommy was enchanted, and Annika also very much wanted a pistol, provided it wasn’t loaded.\\

“Now we can organize a robber band if we want to,” said Pippi. She held the telescope up to her eyes. “With this I can almost see the fleas in South America, I think,” she continued. “And it’ll be good to have if we do organize a robber band.”\\

Just then there was a knock at the door. It was Tommy’s and Annika’s father who had come to take them home. It was long past their bedtime, he said. Tommy and Annika hurried to say thank you, bid Pippi good-by, and collect all their belongings, the flute, the brooch, and the pistols.\\

Pippi followed her guests out to the porch and watched them disappear through the garden. They turned around to wave. The light from inside shone on her. There she stood with her stiff red braids, dressed in her father’s nightshirt, which billowed around her feet. In one hand she held a pistol and in the other the sword. She saluted with it.\\

When Tommy and Annika and their father reached the gate they heard her calling. They stopped to listen. The wind whistled through the trees so they could just barely hear what she said.\\

“I’m going to be a pirate when I grow up,” she cried. “Are you?”
\end{document}

### 2015-12-28 chil/texts/watsons.tex
setting up things

\documentclass{standard}
\title{The Watsons Go To Birmingham}
\author{Christopher Paul Curtis}
\date{}
\begin{document}
\maketitle
\newpage

\section{And You Wonder Why We Get Called the Weird Watsons}
It was one of those super-duper-cold Saturdays. One of those days that when you breathed out your breath kind of hung frozen in the air like a hunk of smoke and you could walk along and look exactly like a train blowing out big, fat, white puffs of smoke.\\

It was so cold that if you were stupid enough to go outside your eyes would automatically blink a thousand times all by themselves, probably so the juice inside of them wouldn’t freeze up. It was so cold that if you spit, the slob would be an ice cube before it hit the ground. It was about a zillion degrees below zero.\\

It was even cold inside our house. We put sweaters and hats and scarves and three pairs of socks on and still were cold.The thermostat was turned all the way up and the furnace was banging and sounding like it was about to blow up but it still felt like Jack Frost had moved in with us.\\

All of my family sat real close together on the couch under a blan- ket. Dad said this would generate a little heat but he didn’t have to tell us this, it seemed like the cold automatically made us want to get together and huddle up. My little sister, Joetta, sat in the middle and all you could see were her eyes because she had a scarf wrapped around her head. I was next to her, and on the outside was my mother.\\

Momma was the only one who wasn’t born in Flint so the cold was coldest to her. All you could see were her eyes too, and they were ￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼shooting bad looks at Dad. She always blamed him for bringing her all the way from Alabama to Michigan, a state she called a giant icebox. Dad was bundled up on the other side of Joey, trying to look at any- thing but Momma. Next to Dad, sitting with a little space between them, was my older brother, Byron.\\

Byron had just turned thirteen so he was officially a teenage juvenile delinquent and didn’t think it was “cool” to touch anybody or let anyone touch him, even if it meant he froze to death. Byron had tucked the blanket between him and Dad down into the cushion of the couch to make sure he couldn’t be touched.\\

Dad turned on the TV to try to make us forget how cold we were but all that did was get him in trouble.There was a special news report on Channel 12 telling about how bad the weather was and Dad groaned when the guy said, “If you think it’s cold now, wait until tonight, the temperature is expected to drop into recordlow territory, possibly reaching the negative twenties! In fact, we won’t be seeing anything above zero for the next four to five days!” He was smiling when he said this but none of the Watson family thought it was funny. We all looked over at Dad. He just shook his head and pulled the blanket over his eyes.\\

Then the guy on TV said, “Here’s a little something we can use to brighten our spirits and give us some hope for the future: The temperature in Atlanta, Georgia, is forecast to reach...” Dad coughed real loud and jumped off the couch to turn the TV off but we all heard the weatherman say, “...the mid-seventies!” The guy might as well have tied Dad to a tree and said, “Ready, aim, fire!”\\

Momma said.“That’s a hundred and fifty miles from home!” “Wilona...,” Dad said.\\

“I knew it,”Momma said.“I knew I should have listened to Moses Henderson!”\\

“Who?” I asked.\\

Dad said,“Oh Lord, not that sorry story.You’ve got to let me tell about what happened with him.”\\

Momma said, “There’s not a whole lot to tell, just a story about a young girl who made a bad choice. But if you do tell it, make sure you get all the facts right.”\\

We all huddled as close as we could get because we knew Dad was ￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼going to try to make us forget about being cold by cutting up. Me and Joey started smiling right away, and Byron tried to look cool and bored.
“Kids,” Dad said, “I almost wasn’t your father.You guys came real close to having a clown for a daddy named Hambone Henderson. . . .” “Daniel Watson, you stop right there.You’re the one who started that ‘Hambone’ nonsense. Before you started that everyone called him his Christian name, Moses. And he was a respectable boy too, he was-
n’t a clown at all.”
“But the name stuck, didn’t it? Hambone Henderson. Me and your
granddaddy called him that because the boy had a head shaped just like a hambone, had more knots and bumps on his head than a dinosaur. So as you guys sit here giving me these dirty looks because it’s a little chilly outside ask yourselves if you’d rather be a little cool or go through life being known as the Hambonettes.”
Me and Joey cracked up, Byron kind of chuckled and Momma put her hand over her mouth. She did this whenever she was going to give a smile because she had a great big gap between her front teeth. If Momma thought something was funny, first you’d see her trying to keep her lips together to hide the gap, then, if the smile got to be too strong, you’d see the gap for a hot second before Momma’s hand would come up to cover it, then she’d crack up too.
Laughing only encouraged Dad to cut up more, so when he saw the whole family thinking he was funny he really started putting on a show.
He stood in front of the TV. “Yup, Hambone Henderson proposed to your mother around the same time I did. Fought dirty too, told your momma a pack of lies about me and when she didn’t believe them he told her a pack of lies about Flint.”
Dad started talking Southern-style, imitating this Hambone guy. “Wilona, I heard tell about the weather up that far north in Flint, Mitch-again, heard it’s colder than inside a icebox. Seen a movie about it, think it was made in Flint. Movie called Nanook of the North.Yup, do believe for sure it was made in Flint. Uh-huh, Flint, Mitch-again.
“Folks there live in these things called igloos. According to what I seen in this here movie most the folks in Flint is Chinese. Don’t believe I seen nan one colored person in the whole dang city. You a
￼3
￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼’Bama gal, don’t believe you’d be too happy living in no igloo. Ain’t got nothing against ’em, but don’t believe you’d be too happy living ’mongst a whole slew of Chinese folks. Don’t believe you’d like the food. Only thing them Chinese folks in that movie et was whales and seals. Don’t believe you’d like no whale meat. Don’t taste a lick like chicken. Don’t taste like pork at all.”
Momma pulled her hand away from her mouth. “Daniel Watson, you are one lying man! Only thing you said that was true was that being in Flint is like living in a igloo. I knew I should have listened to Moses. Maybe these babies mighta been born with lumpy heads but at least they’da had warm lumpy heads!
“You know Birmingham is a good place, and I don’t mean just the weather either.The life is slower, the people are friendlier—”
“Oh yeah,” Dad interrupted,“they’re a laugh a minute down there. Let’s see, where was that ‘Coloreds Only’ bathroom downtown?”
“Daniel, you know what I mean, things aren’t perfect but people are more honest about the way they feel”—she took her mean eyes off Dad and put them on Byron—“and folks there do know how to respect their parents.”
Byron rolled his eyes like he didn’t care. All he did was tuck the blanket farther into the couch’s cushion.
Dad didn’t like the direction the conversation was going so he called the landlord for the hundredth time.The phone was still busy.
“That snake in the grass has got his phone off the hook. Well, it’s going to be too cold to stay here tonight, let me call Cydney. She just had that new furnace put in, maybe we can spend the night there.” Aunt Cydney was kind of mean but her house was always warm so we kept our fingers crossed that she was home.
Everyone, even Byron, cheered when Dad got Aunt Cydney and she told us to hurry over before we froze to death.
Dad went out to try and get the Brown Bomber started.That was what we called our car. It was a 1948 Plymouth that was dull brown and real big, Byron said it was turd brown. Uncle Bud gave it to Dad when it was thirteen years old and we’d had it for two years. Me and Dad took real good care of it but some of the time it didn’t like to start up in the winter.
After five minutes Dad came back in huffing and puffing and slap-
￼4
￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼ping his arms across his chest.
“Well, it was touch and go for a while, but the Great Brown One
pulled through again!” Everyone cheered, but me and Byron quit cheering and started frowning right away. By the way Dad smiled at us we knew what was coming next. Dad pulled two ice scrapers out of his pocket and said, “O.K., boys, let’s get out there and knock those windows out.”
We moaned and groaned and put some more coats on and went outside to scrape the car’s windows. I could tell by the way he was pouting that Byron was going to try and get out of doing his share of the work.
“I’m not going to do your part, Byron, you’d better do it and I’m not playing either.”
“Shut up, punk.”
I went over to the Brown Bomber’s passenger side and started hack- ing away at the scab of ice that was all over the windows. I finished Momma’s window and took a break. Scraping ice off of windows when it’s that cold can kill you!
I didn’t hear any sound coming from the other side of the car so I yelled out, “I’m serious, Byron, I’m not doing that side too, and I’m only going to do half the windshield, I don’t care what you do to me.” The windshield on the Bomber wasn’t like the new 1963 cars, it had a big bar running down the middle of it, dividing it in half.
“Shut your stupid mouth, I got something more important to do right now.”
I peeked around the back of the car to see what By was up to.The only thing he’d scraped off was the outside mirror and he was bend- ing down to look at himself in it. He saw me and said, “You know what, square? I must be adopted, there just ain’t no way two folks as ugly as your momma and daddy coulda give birth to someone as sharp as me!”
He was running his hands over his head like he was brushing his hair.
I said, “Forget you,” and went back over to the other side of the car to finish the back window. I had half of the ice off when I had to stop again and catch my breath. I heard Byron mumble my name.
I said, “You think I’m stupid? It’s not going to work this time.” He
￼5
￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼mumbled my name again. It sounded like his mouth was full of some- thing. I knew this was a trick, I knew this was going to be How to Survive a Blizzard, Part Two.
How to Survive a Blizzard, Part One had been last night when I was outside playing in the snow and Byron and his running buddy, Buphead, came walking by. Buphead has officially been a juvenile delinquent even longer than Byron.
“Say, kid,” By had said,“you wanna learn somethin’ that might save your stupid life one day?”
I should have known better, but I was bored and I think maybe the cold weather was making my brain slow, so I said, “What’s that?”
“We gonna teach you how to survive a blizzard.”
“How?”
Byron put his hands in front of his face and said,“This is the most
important thing to remember, O.K.?” “Why?”
“Well, first we gotta show you what it feels like to be trapped in a blizzard.You ready?” He whispered something to Buphead and they both laughed.
“I’m ready.”
I should have known that the only reason Buphead and By would want to play with me was to do something mean.
“O.K.,” By said,“first thing you gotta worry about is high winds.”
Byron and Buphead each grabbed one of my arms and one of my legs and swung me between them going, “Wooo, blizzard warnings! Blizzard warnings! Wooo! Take cover!”
Buphead counted to three and on the third swing they let me go in the air. I landed headfirst in a snowbank.
But that was O.K. because I had on three coats, two sweaters, a T- shirt, three pairs of pants and four socks along with a scarf, a hat and a hood.These guys couldn’t have hurt me if they’d thrown me off the Empire State Building!
After I climbed out of the snowbank they started laughing and so did I.
“Cool, Baby Bruh,” By said,“you passed that part of the test with a B-plus, what you think, Buphead?”
Buphead said, “Yeah, I’d give the little punk a A.”
￼6
￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼They whispered some more and started laughing again.
“O.K.,” By said,“second thing you gotta learn is how to keep your balance in a high wind.You gotta be good at this so you don’t get blowed into no polar bear dens.”
They put me in between them and started making me spin round and round, it seemed like they spun me for about half an hour.When slob started flying out of my mouth they let me stop and I wobbled around for a while before they pushed me back in the same snowbank.
When everything stopped going in circles I got up and we all laughed again.
They whispered some more and then By said, “What you think, Buphead? He kept his balance a good long time, I’m gonna give him a A-minus.”
“I ain’t as hard a grader as you, I’ma give the little punk a double A-minus.”
“O.K., Kenny, now the last part of Surviving a Blizzard, you ready?” “Yup!”
“You passed the wind test and did real good on the balance test but
now we gotta see if you ready to graduate.You remember what we told you was the most important part about survivin’?”
“Yup!”
“O.K., here we go. Buphead, tell him ’bout the final exam.” Buphead turned me around to look at him, putting my back to
Byron. “O.K., square,” he started, “I wanna make sure you ready for this one, you done so good so far I wanna make sure you don’t blow it at graduation time.You think you ready?”
I nodded, getting ready to be thrown in the snowbank real hard this time. I made up my mind I wasn’t going to cry or anything, I made up my mind that no matter how hard they threw me in that snow I was going to get up laughing.
“O.K.,” Buphead said, “everything’s cool, you ’member what your brother said about puttin’ your hands up?”
“Like this?” I covered my face with my gloves.
“Yeah, that’s it!” Buphead looked over my shoulder at Byron and then said, “Wooo! High winds, blowing snow! Wooo! Look out! Blizzard a-comin’! Death around the corner! Look out!”
Byron mumbled my name and I turned around to see why his voice
￼7
￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼sounded so funny.As soon as I looked at him Byron blasted me in the face with a mouthful of snow.
Man! It was hard to believe how much stuff By could put in his mouth! Him and Buphead just about died laughing as I stood there with snow and spit and ice dripping off of my face.
Byron caught his breath and said, “Aww, man, you flunked! You done so good, then you go and flunk the Blowin’ Snow section of How to Survive a Blizzard, you forgot to put your hands up! What you say, Buphead, F?”
“Yeah, double F-minus!”
It was a good thing my face was numb from the cold already or I might have froze to death. I was too embarrassed about getting tricked to tell on them so I went in the house and watched TV.
So as me and By scraped the ice off the Brown Bomber I wasn’t going to get fooled again. I kept on chopping ice off the back win- dow and ignored By’s mumbling voice.
The next time I took a little rest Byron was still calling my name but sounding like he had something in his mouth. He was saying, “Keh-ee! Keh-ee! Hel’ . . . hel’ . . . !”When he started banging on the door of the car I went to take a peek at what was going on.
By was leaned over the outside mirror, looking at something in it real close. Big puffs of steam were coming out of the side of the mirror.
I picked up a big, hard chunk of ice to get ready for Byron’s trick.
“Keh-ee! Keh-ee! Hel’ me! Hel’ me! Go geh Momma! Go geh Mom-ma! Huwwy uh!”
“I’m not playing, Byron! I’m not that stupid! You’d better start doing your side of the car or I’ll tear you up with this iceball.”
He banged his hand against the car harder and started stomping his feet. “Oh, please, Keh-ee! Hel’ me, go geh Mom-ma!”
I raised the ice chunk over my head.“I’m not playing,By,you bet- ter get busy or I’m telling Dad.”
I moved closer and when I got right next to him I could see boogers running out of his nose and tears running down his cheeks. These weren’t tears from the cold either, these were big juicy crybaby tears! I dropped my ice chunk.
“By! What’s wrong?”
“Hel’ me! Keh-ee! Go geh hel’!”
￼8
￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼I moved closer. I couldn’t believe my eyes! Byron’s mouth was frozen on the mirror! He was as stuck as a fly on flypaper!
I could have done a lot of stuff to him. If it had been me with my lips stuck on something like this he’d have tortured me for a couple of days before he got help. Not me, though, I nearly broke my neck try- ing to get into the house to rescue Byron.
As soon as I ran through the front door Momma, Dad and Joey all yelled,“Close that door!”
“Momma, quick! It’s By! He’s froze up outside!”
No one seemed too impressed.
I screamed,“Really! He’s froze to the car! Help! He’s crying!” That shook them up.You could cut Byron’s head off and he prob-
ably wouldn’t cry.
“Kenneth Bernard Watson, what on earth are you talking about?” “Momma, please hurry up!”
Momma, Dad and Joey threw on some extra coats and followed me
to the Brown Bomber.
The fly was still stuck and buzzing. “Oh, Mom-ma! Hel’ me! Geh
me offa ’ere!”
“Oh my Lord!” Momma screamed, and I thought she was going to
do one of those movie-style faints, she even put her hand over her forehead and staggered back a little bit.
Joey, of course, started crying right along with Byron.
Dad was doing his best not to explode laughing. Big puffs of smoke were coming out of his nose and mouth as he tried to squeeze his laughs down. Finally he put his head on his arms and leaned against the car’s hood and howled.
“Byron,” Momma said, gently wiping tears off his cheeks with the end of her scarf,“it’s O.K., sweetheart, how’d this happen?” She sound- ed like she was going to be crying in a minute herself.
Dad raised his head and said, “Why are you asking how it hap- pened? Can’t you tell,Wilona?This little knucklehead was kissing his reflection in the mirror and got his lips stuck!” Dad took a real deep breath.“Is your tongue stuck too?”
“No! Quit teasin’, Da-ee! Hel’! Hel’!”
“Well, at least the boy hadn’t gotten too passionate with himself!” Dad thought that was hilarious and put his head back on his arms.
￼9
￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼Momma didn’t see anything funny. “Daniel Watson! What are we gonna do? What do y’all do when this happens up he-uh?” Momma started talking Southernstyle when she got worried. Instead of saying “here” she said “he-uh” and instead of saying “you all” she said “y’all.”
Dad stopped laughing long enough to say, “Wilona, I’ve lived in Flint all my life, thirty-five years, and I swear this is the first time I’ve ever seen anyone with their lips frozen to a mirror. Honey, I don’t know what to do, wait till he thaws out?”
“Pull him off, Dad,” I suggested. Byron went nuts! He started bang- ing his hands on the Brown Bomber’s doors again and mumbling, “No! No! Mom-ma, doe leh him!”
Joey blubbered out,“This is just like that horrible story Kenny read me about that guy Nar-sissy who stared at himself so long he forgot to eat and starved to death. Mommy, please save him!” She went over and hugged her arms around stupid Byron’s waist.
Momma asked Dad, “What about hot water? Couldn’t we pour enough hot water on the mirror so it would warm up and he could get off?” She kept wiping tears off By’s cheeks and said, “Don’t you worry, Baby, we gonna get you off of this.” But her voice was so shaky and Southern that I wondered if we’d be driving around in the sum- mer with a skeleton dangling from the outside mirror by its lips.
Dad said,“I don’t know, pouring water on him might be the worst thing to do, but it might be our only chance. Why don’t you go get some hot tap water and I’ll stay to wipe his cheeks.”
Joey told By, “Don’t worry, we’ll come right back.” She stood on her tiptoes and gave By a kiss, then she and Momma ran inside. Dad cracked up all over again.
“Well, lover boy, I guess this means no one can call you Hot Lips, can they?”
Dad was killing himself.“Or the Last of the Red Hot Lovers either, huh?” He tugged on Byron’s ear a little, pulling his face back.
By went nuts again. “Doe do dat! Mom-ma! Momma, hel’! Keh- ee, go geh Mom-ma! Huwwy!”
“Hmm, I guess that’s not going to work, is it?”
Every time he wiped away the tears and the little mustache of boogers on Byron’s lip Dad couldn’t help laughing, until a little river of tears was coming out of his eyes too.
￼10
￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼Dad tried to straighten his face out when Momma and Joey came running back with a steaming glass of hot water, but the tears were still running down his cheeks.
Momma tried to pour water on the mirror but her hands were shaking so much, she was splashing it all over the place. Dad tried too, but he couldn’t look at Byron without laughing and shaking.
That meant I had to do it.
I knew that if my lips were frozen on something and everybody was shaking too much to pour water on them except for Byron he’d do some real cruel stuff to me. He probably would have “accidentally” splashed my eyes until they were frozen open or put water in my ears until I couldn’t hear anything, but not me. I gently poured a little stream of water over the mirror.
Dad was right! This was the worst thing we could do! The water made a cracking sound and froze solid as soon as it touched the mir- ror and By’s lips!
Maybe By’s mouth was frozen but his hands sure weren’t and he popped me right in the forehead. Hard! I hate to say it but I started crying too.
It’s no wonder the neighbors called us the Weird Watsons behind our backs.There we were, all five of us standing around a car with the temperature about a million degrees below zero and each and every one of us crying!
“ ’top! ’top!” By yelled.
“Daniel Watson, what’re we gonna do?” Momma went nuts. “You gotta get this boy to the hospital! My baby is gonna die!”
Dad tried to look serious real quick.
“Wilona, how far do you think I’d get driving down the street with this little clown attached to the mirror? What am I supposed to do, have him run beside the car all the way down to the emergency room?”
Momma looked real close at By’s mouth, closed her eyes for a sec- ond like she was praying and finally said, “Daniel, you get in there and call the hospital and see what they say we should do. Joey and Kenny, go with your daddy.”
Dad and Joey went crying into the house. I stayed by the Brown Bomber. I figured Momma was clearing everybody out for something.
￼11
￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼Byron did too and looked at Momma in a real nervous way. Momma put her scarf around Byron’s face and said, “Sweetheart, you know we gotta do something. I’ma try to warm your face up a
little. Just relax.”
“O.K., Mom-ma.”
“You know I love you and wouldn’t do anything to hurt you,
right?” If Momma was trying to make Byron relax she wasn’t doing a real good job at it. All this talk about love and not getting hurt was making him real nervous.
“Wah are you gonna do? Huh? Doe hur’ me! Keh-ee, hel’!”
Momma moved the scarf away and put one hand on Byron’s chin and the other one on his forehead.
“No! Hel’! Hel’ me, Keh-ee!”
Momma gave Byron’s head a good hard snatch and my eyes auto- matically shut and my hands automatically flew up to cover my ears and my mouth automatically flew open and screamed out, “Yeeeowwww!”
I didn’t see it, but I bet Byron’s lips stretched a mile before they finally let go of that mirror. I bet his lips looked like a giant rubber band before they snapped away from that glass!
I didn’t hear it, but I bet Byron’s lips made a sound like a giant piece of paper being ripped in half!
When I opened my eyes Byron was running to the house with his hands over his mouth and Momma following right behind him. I ran over to the mirror to see how much of Byron’s mouth was still stuck there.
The dirty dogs let Byron get away with not doing his share of the windows and I had to do the whole car myself.When we were final- ly going to Aunt Cydney’s house I decided to pay Byron back for punching me in the forehead and getting out of doing his part of the window scraping. Joey was sitting between us so I felt kind of safe. I said to her, loud, “Joetta, guess what. I’m thinking about writing my own comic book.”
“What about?”
“Well, it’s going to be about this real mean criminal who has a ter- rible accident that turns him into a superhero.”
Joey knew I was going to tease Byron so she sat there looking like I should be careful what I said. Finally I asked her, “Do you want to
￼12
￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼
know what I’m going to call this new superhero?” “What?”
“I’m going to call him the Lipless Wonder. All he does is beat up superheroes smaller than him and the only thing he’s afraid of is a cold mirror!”
All the Weird Watsons except Byron cracked up. Momma’s hand even covered her mouth. I was the only one who saw Byron flip me a dirty finger sign and try to whisper without smearing all theVaseline Momma had put on his lips, “You wait, I’m gonna kick your little behind.” Then he made his eyes go crossed, which was his favorite way of teasing me, but I didn’t care, I knew who had won this time!
\end{document}

### 2016-01-02 computer_organization/notes.tex
update several

\documentclass{article}
\usepackage{amssymb,amsmath}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{multicol}
\topmargin=-0.3in
\textheight=9.2in
\textwidth=168mm
\oddsidemargin=-0.2in
\evensidemargin=-0.2in

\begin{document}
\section*{16 Nov 2015}
\hrule
\noindent\\
\subsection*{Chapter 5: Memory}
\begin{multicols}{2}
5 classic components of a computer
\begin{itemize}
\item Datapath
\item Control
\item Memory
\item Input
\item Output
\end{itemize}
Memory is composed of Cache, Main Memory, and Secondary Memory\\
\begin{itemize}
\item Cache memory is composed of Flip-Flops
\item Cache is physically larger than RAM memory (main memory)
\item RAM has one transistor and one capacitor; cache has two AND gates
\end{itemize}
Memory Types\\
\textbf{RAM}
\begin{itemize}
\item Main Memory
\item \textbf{DRAM}: High density, low power, cheap, slow
\item \textbf{SDRAM}: Faster than DRAM
\item \textbf{DDR RAM}: Twice the amount of data in a single bus clock cycle
\item \textbf{SRAM}: Does not refresh and stays as long as there is power.
\begin{itemize}
\item Three levels
\item More complex: 6 - 8 transistors
\item Fast: very close to the CPU
\item Used for cache memory
\end{itemize}
\end{itemize}
Larger capacity memory is usually slower\\\\
Performance
\end{multicols}
\end{document}

### 2016-01-02 computer_organization/review.tex
update several

--- 2015-12-05 comporg/review.tex
@@
 
+\title{Computer Organization Exam Review}
+\date{}
+
 \begin{document}
+\maketitle
 \begin{enumerate}
@@
 \end{align*}
-\item What are the average instruction times for each designs?
-\indent I don't know.
+\item What are the average instruction times for each designs?\\
+The single cycle datapath's instruction time is determined by the longest instruction time, which is the 600ps load word.
+The multicycle and pipelined datapath are dependant on the longest clock cycle in the longest instruction, which is 200ps.
+We then tak that and multiply it by the CPI, which yields 824ps for the multicycle datapath, and 215ps for the pipelined datapath.
 \end{enumerate}
@@
 \end{enumerate}
+\item For the following diagram, explain what happens in each block for the instruction $lw\ \$t3, 100(\$t2)$.
+\begin{center}
+\includegraphics[scale=0.5]{review_01}
+\end{center}
+\begin{itemize}
+\item \underline{Instruction Fetch:} The machine gathers the next instruction as determined by the program counter.
+The PC increments by 4 every cycle, assuming the program does not encouter any exceptions or errors.
+\item \underline{Instruction Decode:} In this step, the instruction is interpreted by the machine to be the load word instruction.
+\item \underline{Register Read:} The appropriate registers are parsed, and the address is decoded.
+\item \underline{ALU Operation:} The ALU adds the offest to the base address, and requests the contents of that location in memory.
+\item \underline {Register Write:} The machine now writes the contents of memory that have been fetched to the $\$t3$ register.
+\item \underline {Instruction Fetch:} The PC is incremented by 4, and the process continues for the next instruction.
+\end{itemize}
+% \item A compiler designer is trying to decide between two code sequences for a particular machine.
+% Based on the hardware implementation, we have the following:
+% \begin{center}
+% \begin{tabular}{|l|c|c|c|}
+% \hline
+% Class & A & B & C\\
+% \hline
+% CPI & 1 & 2 & 3\\
+% \hline
+% \end{tabular}
+% \end{center}
+% The first code sequence has 5 instructions:
+% 2 of A, 1 of B, and 2 of C.
+% The second code sequence has 6 instruction:
+% 4 of A, 1 of B, and 1 of C.
+% \begin{enumerate}
+% \item What is the CPI for each sequence?\\
+% The CPI for the first sequence is
+% \[
+% 2 * 1 + 1 * 2 + 2 * 3 = 10.
+% \]
+% The CPI for the second sequence is
+% \[
+% 4 * 1 + 1 * 2 + 1 * 3 = 9.
+% \]
+% \item Which sequence is faster, and by how much?
+%
+% \end{enumerate}
 \end{enumerate}

### 2016-01-02 partial_differential_equations/cauchy.tex
update several

\section{The Cauchy Problem}
\hrule
\noindent \\\\
\indent Up until now, we have always been given our PDE's with boundary conditions and initial values; but imagine that we did not have a boundary? The traditional method that we have been using would fail at this point. Since we would not be able to find any eigenvalues, we would not be able to form a series solution by using the \textit{Separation of Variables} technique. Let's consider the familiar heat equation:
\[
\begin{cases}
\xi_{t} - k\xi_{xx} = 0\\
\xi(x,0) = f(x)
\end{cases}
\]
Through much careful derivation and theory, we can say that the general solution for this problem will be
\[
\xi(x,t) = \frac{1}{\sqrt{4 \pi k t}}\int_{-\infty}^{\infty}e^{-\frac{(x-y)^{2}}{4kt}}f(y)dy.
\]
The question now should be ``How can we solve this?'' The first step we will take is letting
\[
u = \frac{y-x}{\sqrt{4kt}}.
\]
We are going to perform a \textit{u substitution} to try and put this integral into a nicer form. This yields
\begin{align*}
y &= x + u\sqrt{4kt}\\
dy &= \sqrt{4kt}\ du.
\end{align*}
This changes our original problem into
\[
\xi(x,t) = \frac{1}{\sqrt{\pi}}\int_{-\infty}^{\infty}e^{-u^{2}}f(x+u\sqrt{4kt})du.
\]
Let's look at an example\\\\
\noindent \textbf{\textit{Ex:}} Solve the following PDE
\[
\begin{cases}
\mu_{t} - k\mu_{xx} = 0\\
\mu(x,0) = \cos{x}
\end{cases}
\]
\indent \textbf{\textit{Solution:}} From above, we know what our general solution will look like. All we need to do is substitute in our initial condition. We have the following:
\begin{align*}
\mu(x,t) &= \frac{1}{\sqrt{\pi}}\int_{-\infty}^{\infty}e^{-u^{2}}\cos{(x + u\sqrt{4kt})}du\\
&= \frac{1}{\sqrt{\pi}}\int_{-\infty}^{\infty}e^{-u^{2}}\left[\cos{(x)}\cos{(u\sqrt{4kt})} - \sin{(x)}\sin{(u\sqrt{4kt})}\right]du\\
&=  \frac{2}{\sqrt{\pi}}\cos{(x)}\int_{0}^{\infty}e^{-u^{2}}\cos{(u\sqrt{4kt})}du.
\end{align*}
You may be wondering what happened to the $\sin$ terms. Recall that $\sin$ is an odd function, so it will ultimately be $0$ on the interval $[-\infty,\infty]$; because of this fact, we can remove it now. We also use the face that the $\cos$ is even, so it is okay to halve the interval and multiply by two. Now we need to evaluate this integral. We will substitute $\alpha = \sqrt{4kt}$, and assign a function, $\Upsilon(\alpha)$, to the integral. This gives us the following system:
\[
\begin{cases}
\Upsilon(\alpha) = \int_{0}^{\infty}e^{-u^{2}}\cos{(\alpha u)}du\\
\Upsilon(0) = \int_{0}^{\infty}e^{-u^{2}}du\\
\end{cases}
\]
Believe it or not, we actually know the answer to $\Upsilon(0)$. To find it, we can assign a function to the integral in a similar fashion as we have already done. Calling it $\Psi$, we have
\begin{align*}
\Psi &= \int_{0}^{\infty}e^{-u^{2}}du\\
\Psi^{2} &= \int_{0}^{\infty}e^{-u^{2}}du \int_{0}^{\infty}e^{-v^{2}}dv\\
\Psi^{2} &= \int_{0}^{\infty}\int_{0}^{\infty}e^{-u^{2}-v^{2}}dudv\\
\Psi^{2} &= \int_{0}^{\pi/2}\int_{0}^{\infty}re^{-r^{2}}drd\theta\\
\Psi^{2} &= \frac{\pi}{2}\left[-\frac{1}{2}e^{-r^{2}}\right]\Bigg |_{r=0}^{r=\infty}\\
\Psi^{2} &= \frac{\pi}{4}\\
\Psi &= \frac{\sqrt{\pi}}{2}.
\end{align*}
So we know that $\Upsilon(0) = \frac{\sqrt{\pi}}{2}$. Plugging this into our system gives us
\[
\begin{cases}
\Upsilon(\alpha) = \int_{0}^{\infty}e^{-u^{2}}\cos{(\alpha u)}du\\
\Upsilon(0) = \frac{\sqrt{\pi}}{2}\\
\end{cases}
\]
Now, we will take the derivative of $\Upsilon$. This gives us
\[
\Upsilon'(\alpha) = -\int_{0}^{\infty}ue^{-u^{2}}\sin{(\alpha u)}du.
\]
Integration by parts gives us
\[
\frac{1}{2}e^{-u^{2}}\sin{(\alpha u)}\Big |_{0}^{\infty} - \alpha\int_{0}^{\infty}\frac{1}{2}e^{-u^{2}}\cos{(\alpha u)}du.
\]
The value from $0$ to $\infty$ is zero, since the $\sin$ function odd. This leaves us with following relationship:
\[
\Upsilon'(\alpha) = -\frac{\alpha}{2}\Upsilon(\alpha) \alpha.
\]
We can solve this as an ODE. It's separable, so we get
\begin{align*}
\frac{d\Upsilon}{\Upsilon} &= -\frac{\alpha}{2}d\alpha\\
\ln{(\Upsilon)} &= -\frac{\alpha^{2}}{4} + c_{1}\\
\Upsilon &= c_{2}e^{-\frac{\alpha^{2}}{4}}.
\end{align*}
Applying the initial condition gives us
\[
\Upsilon = \frac{\sqrt{\pi}}{2}e^{-\frac{\alpha^{2}}{4}}.
\]
Substituting back in gives us a final answer of
\begin{align*}
\mu(x,t) &= \frac{2}{\sqrt{\pi}}\cos{x}\Upsilon(\sqrt{4kt})\\
&= \cos{x}e^{-kt},
\end{align*}
and we are done.

\newpage
\indent Now we will move on to the Cauchy wave equation. We will be presented with a similar problem: a wave equation on an unbounded domain. Consider the following problem:
\noindent \\\\
\noindent \textbf{\textit{Ex:}} Solve the following PDE
\[
\begin{cases}
u_{tt} - c^{2}u_{xx} = 0\\
u(x,0) = \sin{x}\\
u_{t}(x,0) = \cos{x}
\end{cases}
\]
\indent \textbf{\textit{Solution:}} Just as with the heat equation, we have a formal solution for the homogeneous wave equation. We will do the derivation, but I will tell you that the formal solution is
\[
u(x,t) = \frac{1}{2}\left[f(x+ct) + f(x-ct)\right] + \frac{1}{2c}\int_{x-ct}^{x+ct}g(s)\mathop{ds}.
\]
Now let's derive the formal solution. We want a solution in the form of
\[
u(x,t) = F(x+ct) + G(x-ct).
\]
Now we can say that $u_{t} = cF'(x+ct) - cG'(x-ct)$, and $u_{t}(x,0) = cF'(x) - cG'(x)$. This means
\begin{align*}
F(x) + G(x) &= f(x)\\
F(x) - G(x) &= \frac{1}{c}\int_{0}^{x}g(s)\mathop{ds} + c_{1}.
\end{align*}
Solving the system yields
\begin{align*}
u(x,t) &= \frac{1}{2} f(x+ct) + \frac{1}{2c}\int_{0}^{x+ct}g(s)\mathop{ds} + \frac{c_{1}}{2} + \frac{1}{2}f(x-ct) -\frac{1}{2c}\int_{0}^{x-ct}g(s)\mathop{ds} - \frac{c_{1}}{2}\\
&= \frac{1}{2}\left[f(x+ct) + f(x-ct)\right] + \frac{1}{2c}\int_{x-ct}^{x+ct}g(s)\mathop{ds},
\end{align*}
and that is the derivation. This problem is simply plugging into the formal solution. We have the following:
\begin{align*}
u(x,t) &= \frac{1}{2}\left[\sin{(x+ct)} + \sin{(x-ct)}\right] + \frac{1}{2c}\int_{x-ct}^{x+ct}\cos{(s)}\mathop{ds}\\
&= \frac{\sin{(x+ct)} + \sin{(x-ct)}}{2} + \frac{\cos{(x)}\sin{(ct)}}{c}\\
&=  \frac{\cos{(x)}\sin{(ct)}}{c} + \cos{(ct)}\sin{(x)},
\end{align*}
and we are done.

\newpage
\indent So far, we have only dealt with homogeneous Cauchy problems. We will now look at the case where we have a non-homogeneous problem. The first variation of this problem is the \textit{half-line} problem. In this problem, we will have a bound on our equation, and we will generalize the solution so that we are dealing with an unbounded Cauchy problem. Let's look at an example.
\noindent\\\\ \textbf{\textit{Ex:}} Solve the following PDE
\[
\begin{cases}
u_{t} - ku_{xx} = 0\\
u(x,0) = \phi(x)\\
u(0,t) = 0
\end{cases}
\]
\indent \textbf{\textit{Solution:}} We have both an initial condition and a boundary condition here. This problem is saying that at $x=0$, there is an impassable boundary. This forces our domain to be $[0,\infty)$. To solve a Cauchy problem, however, we want a domain of $(-\infty,\infty)$. We can force this domain by extending our initial condition, $\phi(x)$, over the boundary. We have the following rule:
\[
\phi_{0}(x) =
\begin{cases}
\phi(x)\quad\quad x>0\\
-\phi(-x)\ x<0.
\end{cases}
\]
Now we are dealing with the system
\[
\begin{cases}
u_{t} - ku_{xx} = 0\\
u(x,0) = \phi_{0}(x),
\end{cases}
\]
which we can solve using the formal solution to the heat equation. Recall that
\[
u(x,t) = \frac{1}{\sqrt{4k\pi t}}\int_{-\infty}^{\infty}e^{-\frac{(x-y)^{2}}{4kt}}\phi_{0}(y)\mathop{dy}.
\]
Substituting into the equation gives us
\begin{align*}
u(x,t) &= \frac{1}{\sqrt{4k\pi t}}\left[\int_{0}^{\infty}e^{-\frac{(x-y)^{2}}{4kt}}\phi(y)\mathop{dy} + \int_{-\infty}^{0}e^{-\frac{(x-y)^{2}}{4kt}}(-\phi(-y))\mathop{dy} \right]\\
&= \frac{1}{\sqrt{4k\pi t}}
\end{align*}

### 2016-01-02 partial_differential_equations/classify.tex
update several

\section{Classification of Partial Differential Equations}
\hrule
\noindent\\\\
\indent We mentioned earlier the topic of classification of PDE's. Since we only
have this and Fourier Series to talk about, we will go ahead and knock out
classification. There's some information that we need to go ahead and review
here. This will be explained in the problem, but it's handy to have it for
reference:
\begin{align*}
b^{2} - ac < 0 &\implies \text{Elliptic with complex roots}\\
b^{2} - ac = 0 &\implies \text{Parabolic with the same root twice}\\
b^{2} - ac > 0 &\implies \text{Hyperbolic with two distinct real roots}
\end{align*}
\noindent Another helpful piece of information is what's known as the
\textit{classical trinity} of PDEs.
\begin{align*}
u_{t} - ku_{xx} &= 0 \qquad \text{The heat equation (parabolic)}\\
u_{tt} - c^{2}u_{xx} &= 0 \qquad \text{The wave equation (hyperbolic)}\\
u_{xx} + u_{yy} &= 0 \qquad \text{The Laplace equation (elliptic)}
\end{align*}
Let's go ahead and look at an example.\\\\
\indent \textbf{Ex. }Classify the following PDE:
\[yu_{xx} + u_{yy} = 0\]
\indent \textbf{\textit{Solution:}} The first thing we need to identify are the
coefficients on the $u_{xx}$, $u_{yy}$, and $u_{xy}$ or $u_{yx}$ terms. Note
that the $u_{xy}$ and $u_{yx}$ are interchangeable. We will assign these
coefficients to $a$, $b$, and $c$ corresponding to $u_{xx}$, $u_{xy}$, and
$u_{yy}$, respectively. Looking at the PDE, we can see that
\[
a = y \qquad b = 0 \qquad c = 1
\]
\noindent Now we look at $b^{2} - ac$. We are really dealing with
$b = \frac{b_{0}}{2}$. In other words, we need to divide our $b$ value by $2$.
In the case of the first problem, we are dealing with $b = 0$. Now we can
classify our PDE as follows:
\[
b^{2} - ac = 0 - y < 0 \implies \text{Elliptic}
\]
Note that we assume $y > 0$. So our PDE is elliptic. Now we need to make it look
like it's corresponding equation from the classical trinity. In this case we
have an Elliptic equation, so we want to make it look like Laplace's equation.
We need to find values for $u_{xx}$ and $u_{yy}$, and then substitute those into
the equation. This is similar to how we solved the first order PDE's; let's look
at our characteristic equation:
\begin{gather*}
am^{2} + bm + c = 0\\
ym^{2} + 1 = 0\\
m = \pm \frac{i}{\sqrt{y}}
\end{gather*}
\noindent So, now we have two solutions, and they are both imaginary as expected.
Next we need to find values for $S$ and $T$ that we can substitute into the
equation. We will do both of them at the same time here. Notice that we will
drop the $i$:
\begin{gather*}
\frac{dy}{dx} = \frac{1}{\sqrt{y}} \qquad\qquad\qquad \frac{dy}{dx} = -\frac{1}{\sqrt{y}}\\
\int\frac{dy}{dx} =\int\frac{1}{\sqrt{y}} \qquad\qquad\qquad \int\frac{dy}{dx} = -\int\frac{1}{\sqrt{y}}\\
S = \frac{2}{3}y^{\left(\frac{3}{2}\right)} - x \qquad\qquad\qquad T = \frac{2}{3}y^{\left(\frac{3}{2}\right)} + x
\end{gather*}
\noindent Now we can substitute into our equation. Unfortunately, we will need
to take the partial derivative twice in order to get the substitutions we need.
Let's start with $u_{xx}$.
\begin{align*}
u_{x} &= u_{S}\frac{\partial S}{\partial x} + u_{T}\frac{\partial T}{\partial x}\\
&= -u_{S} + u_{T}\\
u_{xx} &= \frac{\partial}{\partial x}(-u_{S} + u_{T})\\
&= u_{SS} + u_{TT}
\end{align*}
\noindent We will skip the math for finding $u_{yy}$, which would give us
\[
u_{yy} = yu_{SS} + yu_{TT}
\]
\noindent Now we plug everything in. Remember that we were under the assumption
that $y > 0$:
\begin{gather*}
y(u_{SS} + u_{TT}) + yu_{SS} + yu_{TT} = 0\\
2yu_{SS} + 2yu_{TT} = 0\\
u_{SS} + u_{TT} = 0
\end{gather*}
\noindent And we are finished.

### 2016-01-02 partial_differential_equations/first_order.tex
update several

\section{First Order Partial Differential Equations}
\hrule
\noindent\\\\
\noindent\textbf{\textit{Ex:}} Solve the following PDE
\[
u_{x} + u_{y} + u = x + y
\]
\indent \textbf{\textit{Solution:}} In this case, we have a non-zero term on the
right hand side of the equation. We will use substitution to solve this. For the
time being, we will let $T$ be $x + y$. To find our $S$, we can use a similar
method to the one that we used above. We know that the coefficients on the
$u_{x}$ and $u_{y}$ are $1$ in this case. Now we have:
\begin{align*}
\frac{dy}{dx} &= 1\\
dy &= dx\\
\int dy &= \int dx\\
x &= y + c\\
x - y &= c
\end{align*}
\noindent Now we can say that $S = x - y$; however, we need to check our choice
for $T$ to make sure this substitution will not just yield a trivial solution.
We can do this by Next we have to actually use the substitution. We need to
determine whether the coefficient matrix, or the Jacobian, of our substitutions
has a determinant of zero. If it does, then we need to choose a different $T$.
\[
\left|
\begin{array}{c c}
1 & 1\\
1 & -1
\end{array}
\right| = 2 \neq 0.
\]
Our choices of variables will work here, so now we need to express $u_{x}$ and
$u_{y}$ in terms of $S$ and $T$. The following equations come from the chain
rule; when the $S$ and $T$ terms are differentiated, we get $u_{S}$ and $u_{T}$
left behind.
\begin{alignat*}{3}
u_{x} &= u_{S}\frac{\partial S}{\partial x} + u_{T}\frac{\partial T}{\partial x}
\qquad\qquad &&u_{y} = u_{S}\frac{\partial S}{\partial y} + u_{T}\frac{\partial T}{\partial y}\\
u_{x} &= -u_{S} + u_{T}  \qquad\qquad &&u_{y} = u_{S} + u_{T}
\end{alignat*}
\noindent We can now plug in our values into the original equation. Note that we
didn't do anything with $u$; it gets left alone. In addition, if we needed
$u_{xx}$ or $u_{yy}$, we can just differentiate again. Plugging in yields:
\begin{gather*}
(-u_{S} + u_{T}) + (u_{S} + u_{t}) + u = T\\
2u_{T} + u = T
\end{gather*}
\noindent This is just a first order ODE. Since it's not homogeneous, we need to
solve for the general solution, and then the particular solution. Let's start
with the general.
\begin{gather*}
2y^{'} + y = 0\\
2r + 1 = 0\\
r = -\frac{1}{2}\\
y = ce^{\left(-\frac{1}{2}\right)x}
\end{gather*}
\noindent In this case, we know that our $x$ is really $T$. So we finally have
$y = ce^{\left(-\frac{1}{2}\right)T}$. We will back substitute soon, but let's
find the particular solution before we do that.
\noindent Recall from ODE that we have several cases for the particular
solution. In our case, we have a linear equation on the right hand side of our
ODE. This lets us say:
\begin{gather*}
y = AT + B\\
y^{'} = A\\
\end{gather*}
\noindent Now we can substitute again into our non-homogeneous ODE as follows:
\[
2y^{'} + y = T \equiv 2A + AT + B = T
\]
\noindent And we have that $A = 1$, and $B = -2$. Now we have a final solution of:
\begin{gather*}
y = ce^{\left(-\frac{1}{2}\right)T} + T - 2\\
y = ce^{\left(-\frac{1}{2}\right)(x+y)} + (x + y) - 2
\end{gather*}
\noindent We aren't quite done yet, as we have a solution to an ODE, not a PDE.
It isn't too difficult to switch back; The $c$ is really the only difference. In
our ODE, the $c$ is a constant term, but to our PDE, the $c$ is a function. The
$c$ now becomes $f(s)$. We will go ahead and back substitute, which will give us
$f(y-x)$. Now we have our final solution:
\[
u(x,y) = f(y-x)e^{\left(-\frac{1}{2}\right)(x+y)} + (x + y) - 2
\]
\noindent And we are done.

### 2016-01-02 partial_differential_equations/fourier_series.tex
update several

\section{Fourier Series and Partial Differential Equations}
\hrule
\noindent\\
\subsection{Finding Fourier Series}
\indent It's finally time to talk about Fourier Series. Before we start though,
it may be helpful to look at some common patterns amongst the Fourier Series.\\\\
\centerline{
\bgroup
\def\arraystretch{1.5}
\begin{tabular}{|c|c|c|c|}
\hline
& Cosine & Sine & Full Series \\ \hline
Interval & $[0,L]$ & $[0,L]$ & $[-L,L]$\\ \hline
Basis & $\{1,\cos{\left(\frac{n\pi x}{L}\right)}\}$ &
$\{\sin{\left(\frac{n\pi x}{L}\right)}\}$ &
$\{1,\sin{\left(\frac{n\pi x}{L}\right)},\cos{\left(\frac{n\pi x}{L}\right)}\}$\\ \hline
Series &
$\frac{b_{0}}{2} + \sum_{n=1}^{\infty}[b_{n}\cos{\left(\frac{n\pi x}{L}\right)}]$&
$\sum_{n=1}^{\infty}[a_{n}\sin{\left(\frac{n\pi x}{L}\right)}]$&
$\frac{b_{0}}{2} + \sum_{n=1}^{\infty} [a_{n}\sin{\left(\frac{n\pi x}{L}\right)} + b_{n}\cos{\left(\frac{n\pi x}{L}\right)}]$\\ \hline
\end{tabular}
\egroup
}\\\\
\noindent To solve for the coefficients, we have:
\begin{align*}
Sine:  \qquad\qquad &a_{n} = \frac{2}{L}\int_{0}^{L}f(x)\sin{\left(\frac{n\pi x}{L}\right)}dx\\
Cosine: \qquad\qquad &b_{n} = \frac{2}{L}\int_{0}^{L}f(x)\cos{\left(\frac{n\pi x}{L}\right)}dx\\
&b_{0} =  \frac{1}{L}\int_{0}^{L}f(x)dx\\
Full: \qquad\qquad &a_{n} = \frac{2}{L}\int_{-L}^{L}f(x)\sin{\left(\frac{n\pi x}{L}\right)}dx\\
&b_{n} = \frac{2}{L}\int_{-L}^{L}f(x)\cos{\left(\frac{n\pi x}{L}\right)}dx\\
&b_{0} =  \frac{1}{L}\int_{-L}^{L}f(x)dx
\end{align*}
\indent It will also help us to quickly review the concept of odd and even
functions, as they will help us to simplify the process of computing Fourier
series later. Specifically, let's look at the graphs for $\sin$ and $\cos$ on
$[-\pi,\pi]$, as well as another function that will turn out to be neither even
nor odd.
\begin{center}
\begin{tabular}{c c c}
\includegraphics[scale=0.4]{cos_eo} & \includegraphics[scale=0.4]{sin_eo} & \includegraphics[scale=0.4]{cube_eo}\\
$f(x) = \cos{x}$ & $f(x) = \sin{x}$ & $f(x) = x^{3}-x +1$
\end{tabular}
\end{center}
\noindent\\
\indent We should recall that a function is \textit{even} when $f(x) = f(-x)$
for all $x$ in the domain of $f$. Graphically speaking, an \textit{even}
function is reflected across only the $y$-axis. By looking at the above graphs,
we can see that $f(x) = \cos{x}$ is an even function.\\
\indent A function is \textit{odd} when $-f(x) = f(-x)$ for all x in the domain
of $f$. Graphically speaking, the function is reflected across the $x$-axis and
the $y$-axis. From the above graphs, we can say that $f(x) = \sin{x}$ is an odd
function. Note that we can have a function that is neither odd nor even, but we
won't see much of this in the following problems. The reason that we want to
discuss odd and even functions here is because the integration that we will be
seeing later will be affected by the symmetry.\\\\
\indent The first type of series we will look at is the Full Fourier Series.
Note that this series is just a combination of a Fourier Sine Series and a
Fourier Cosine Series. A typical problem would look something like this:\\\\
\textbf{\textit{Ex:}} Find the corresponding Fourier Series for
\begin{gather*}
f(x)=
\begin{cases*}
L \qquad -L \leq x \leq 0\\
2x \qquad 0 \leq x \leq L
\end{cases*}
\end{gather*}
\indent \textbf{\textit{Solution:}} So we have a piecewise function here, but
that shouldn't throw us off too much; all we need to do is combine the integrals
for each part of the coefficient solution. Let's start by finding the terms for
the $cos$ series.
\begin{align*}
B_{0} &= \frac{1}{2L}\left[\int_{-L}^{0}Ldx + \int_{0}^{L}2xdx\right]\\
&= \frac{1}{2L}[2L^{2}]\\
&= L
\end{align*}
\noindent Now We need to find $B_{n}$, and then we will have our cosine part of
the series. The integration can get pretty messy, so we will skip some steps.
\begin{align*}
B_{n} &= \frac{1}{L}\left[\int_{-L}^{0}L\cos{\left(\frac{n\pi x}{L}\right)}dx + \int_{0}^{L}2x\cos{\left(\frac{n\pi x}{L}\right)}dx\right]\\
&= \frac{1}{L}\left[0 + \left(\frac{2L^{2}}{n^{2}\pi^{2}}\right)((-1)^{n} - 1)\right]\\
&= \frac{2L}{(n\pi)^{2}}((-1)^{n} - 1)
\end{align*}
\noindent Note that the sequence will be 0 for all $n = 2k$. This allows us to
substitute $2k+1$ for $n$, since we don't care about the $0$ terms. For our
cosine part of the series, we now have:
\begin{align*}
B_{n} &= \frac{2L}{(2k+1)^{2}\pi^{2}}((-1)^{2k+1}-1)\\
&= \frac{2L}{(2k+1)^{2}\pi^{2}}(-2)\\
&= \frac{-4L}{(2k+1)^{2}\pi^{2}}
\end{align*}
\noindent Now we need to find the sine part of the series. This is easier,
since we only need to find $A_{n}$.
\begin{align*}
A_{n} &= \frac{1}{L}\left[\int_{-L}^{0}L\sin{\left(\frac{n\pi x}{L}\right)}dx +
\int_{0}^{L}2x\sin{\left(\frac{n\pi x}{L}\right)}dx \right]\\
&= \frac{L}{n\pi}\left[-1-(-1)^{n}\right]
\end{align*}
\noindent Note that we can do a similar substitution here as we did with the
cosine series. This time, however, we will have $0$ for $n = 2k+1$ We now have:
\begin{align*}
A_{n} &= \frac{L}{2k\pi}\left[-1-(-1)^{2k}\right]\\
&= \frac{L}{2k\pi}\left[-1 - 1\right]\\
&= \frac{L}{2k\pi}\left[-2\right]\\
&= \frac{-2L}{2k\pi}
\end{align*}
\noindent Now we have all the parts that we need. Our final answer will be
the sum of all the parts:
\begin{align*}
f(x)&\sim \frac{L}{2} + \sum_{k = 0}^{\infty}\left[\frac{-4L}{(2k+1)^{2}\pi^{2}}
\cos{\left(\frac{(2k + 1)\pi x}{L}\right)} \right] + \sum_{k =
0}^{\infty}\left[\frac{-2L}{2k\pi}\sin{\left(\frac{2k\pi x}{L}\right)} \right]\\
&\sim \frac{L}{2} + \sum_{k = 0}^{\infty}\left[\frac{-4L}{(2k+1)^{2}\pi^{2}}
\cos{\left(\frac{(2k + 1)\pi x}{L}\right)} + \frac{-2L}{2k\pi}\sin{
\left(\frac{2k\pi x}{L}\right)} \right]
\end{align*}
\noindent And we are done.\\

\newpage
\indent Now lets take a look at a slightly more theoretical Fourier Series.\\
\noindent\textbf{\textit{Ex:}} Find the corresponding Fourier Series for\\
\[f(x)=\left|\cos{x}\right|\qquad -\pi\leq x \leq \pi\]
\indent\textbf{\textit{Solution:}} Before we dive into calculations here, let's
analyze the function at hand. We have the absolute value of $\cos{x}$. Usually,
$\cos$ would be an odd function, which would imply that our $\cos$ series
component to the whole Fourier Series would be $0$. However, the absolute value
makes our function even, and as such the $\sin$ component of the series will be
$0$. That's good news for us, because now we don't have to find $a_{n}$; we
actually only have a Fourier Cosine Series here. Recall that we have a formula
for $b_{n}$ and $b_{0}$, and that our general solution will have the form of
\[\frac{b_{0}}{2} + \sum_{n=1}^{\infty}[b_{n}\cos{\left(\frac{n\pi x}{L}\right)}]\]
\noindent Our function is bounded by $\pi$, so we really have $L= \pi$. This gives us:
\[\frac{b_{0}}{2} + \sum_{n=1}^{\infty}[b_{n}\cos{\left(nx\right)}]\]
\noindent Let's go ahead and find our $b_{0}$:
\begin{align*}
b_{0} &= \frac{1}{\pi}\int_{-\pi}^{\pi}\left|\cos{x}\right|dx\\
&= \frac{2}{\pi}\int_{0}^{\pi}\left|\cos{x}\right|dx\\
&= \frac{4}{\pi}\int_{0}^{\frac{\pi}{2}}\left|\cos{x}\right|dx\\
&= \frac{4}{\pi}\left[\sin{\frac{\pi}{2}} - \sin{0}\right]\\
&= \frac{4}{\pi}
\end{align*}
\noindent We know that since the function is even, we can say that the the
integral of the entire range is really twice the integral of half the range.
Since we are dealing with the cosine function, we actually have four times a
quarter of the range. If we take a look at a graph, this may make more sense:\\
\begin{center}
\begin{tabular}{c c}
\includegraphics[scale=0.5]{cos_abs} & \includegraphics[scale=0.5]{cos}\\
$f(x) = |\cos{x}|$ on $[0,2\pi]$ & $f(x) = \cos{x}$ on $[0,2\pi]$
\end{tabular}
\end{center}
\noindent\\ You can see that the section of the graph from
$[\frac{\pi}{2},\frac{3 \pi}{2}]$ is positive for the $|\cos{x}|$, but negative
for $\cos{x}$. If we simply integrated the $|\cos{x}|$ over $[0,\pi]$, the
result would be 0, but graphically this result would not make any sense. Now
let's take a look at the $b_{n}$ term. Before we jump into the integration,
let's think about this function:
\[\left|\cos{x}\right| =
\begin{cases*}
\cos(x) \qquad 0\leq x \leq \frac{\pi}{2}\\
-\cos(x)\quad  \frac{\pi}{2}\ \leq x \leq \pi
\end{cases*}
\]
\noindent Now we can set up the formula for the $b_{n}$ term. We have as follows:
\begin{align*}
b_{n} &= \frac{2}{\pi}\left[ \int_{0}^{\frac{\pi}{2}} \cos{x}\cos{nx}dx - \int_{\frac{\pi}{2}}^{\pi}
\cos{x}\cos{nx}dx \right]\\
&= \frac{2}{\pi} \left[ \frac{-2 \cos{\left(\frac{n\pi}{2}\right)} + \sin{\left(n\pi\right)}}{n^{2} -
1}\right]\\
\end{align*}
\noindent Note that (as we have seen before) the sine term disappears; however,
the more interesting part is what happens with our cosine term. The value of
$\cos{\left(\frac{n\pi}{2}\right)}$ will be 0 at all odd $n$, or for all
$n = 2k + 1$. So, we will let our $n = 2k$. This gives us:
\begin{align*}
b_{n} &= \frac{2}{\pi} \left[ \frac{-2 \cos{\left(k \pi \right)}}{(2k)^{2} - 1}\right]\\
&= -\frac{4}{\pi}\left[\frac{(-1)^{k}}{(2k)^{2}-1}\right]
\end{align*}
\noindent And that leaves us with the final result
\[
f(x)\sim \frac{2}{\pi} - \sum_{k=0}^{\infty}\left[\frac{4}{\pi}\left[\frac{(-1)^{k}}{(2k)^{2}-1}\right]
\cos{\left(\frac{2k\pi}{2}\right)}\right]
\]
And we are done.
\newpage
\subsection{Convergence of Fourier Series}
\indent Up to this point, we have found a Fourier series and left our solutions
at that. We will now look into the convergence of a Fourier series. This topic
includes two questions:
\begin{enumerate}
\item Does the series converge?
\item If the series converges, to what value does it converge?
\end{enumerate}
\noindent Before we jump into convergence, we need to review some ideas.
\begin{enumerate}
\item \textit{\textbf{Jump Discontinuity:}} To say that $f(x)$ has a jump
discontinuity at $x = a$ is to say that
\[
\lim_{x \to a^{-}}f(x) \neq \lim_{x \to a^{+}}f(x)
\]
Where both the limit from the left and the limit from the right exist.
\item \textit{\textbf{Piecewise Smooth:}} A function $f(x)$ is said to be
piecewise smooth if the function can be broken into two distinct intervals, and
on each interval both $f(x)$ and $f^{'}(x)$ are continuous.
\item \textit{\textbf{Periodic Extension:}} The periodic extension of $f(x)$ on
$[-L,L]$ is the repetition of the function on intervals on periods to the left
and right of $[-L,L]$.
\end{enumerate}
\indent With these definitions, we can lay out a general rule for convergence of
a Fourier series. We won't give a proof of these rules for the time being. We
just need them to lay out some groundwork for the rest of the information we
will present on Fourier series. We have the following:
\begin{quote}
Suppose $f(x)$ is a function, and that $f(x)$ is piecewise smooth on the interval
$[-L,L]$. The corresponding Fourier series for $f(x)$ converges to either
\begin{enumerate}
\item the periodic extension of $f(x)$, if $f(x)$ is continuous, or
\item the average of the two one-sided limits, if the periodic extension of
$f(x)$ has a jump discontinuity at $x = a$.
\end{enumerate}
\end{quote}
\noindent We should note that the second condition can be expressed by the formula
\[\frac{1}{2}\left[\lim_{x \to a^{-}}f(x) + \lim_{x \to a^{+}}f(x)\right]\]
which gives us a nice way to calculate the value to which the series converges
if we happen to meet the criteria in (2). Let's take a look at some examples.\\\\

\noindent\textbf{\textit{Ex:}} Find the Fourier series for the following
equation, and determine if the series converges. If it does, tell where the
series will converge.
\[ f(x) =
\begin{cases}
L \qquad\qquad -L \leq x \leq 0\\
2x \qquad\qquad 0 \leq x \leq L
\end{cases}
\]
\indent\textbf{\textit{Solution:}} We will go ahead and skip the details of
finding the Fourier series for this function. If you do want to find it, you
will have to find a full piecewise Fourier series. For this particular $f(x)$,
we have:
\begin{align*}
f(x) &\sim L + \sum_{n=1}^{\infty}\left[\frac{2L}{n^{2}\pi^{2}}([-1]^{n}-1)\cos{\left(\frac{n\pi
x}{L}\right)}\right] - \sum_{n=1}^{\infty}\left[\frac{L}{n\pi}(1 + [-1]^{n})\sin{\left(\frac{n\pi
x}{L}\right)}\right]
\end{align*}
\noindent Notice that we didn't simplify this solution to a nicer form.
If we evaluate this series, we will run across terms where the series is 0, but
that is okay for now. This function is also defined as a piecewise function,
and on both intervals ($[-L,0]$ and $[0,L]$), $f(x)$ is continuous. In other
words, $f(x)$ is piecewise smooth, so we know that this series converges. Now
let's take a look at the graph of $f(x)$ on an arbitrary interval of $[-1,1]$.
\begin{center}
\begin{tabular}{c}
\includegraphics[scale=0.6]{pw_graph_01}\\
$
f(x) =
\begin{cases}
1 \qquad\qquad -1 \leq x \leq 0\\
2x \qquad\qquad 0 \leq x \leq 1
\end{cases}
$
\end{tabular}
\end{center}
\noindent We can see that there is a jump in the graph at $x = 0$. In other
words, there is a jump discontinuity at $x = 0$, and that means that we need to
use (2) in order to find the value of convergence. We will let $\sum$ stand in
for our entire series. We have:
\begin{align*}
\sum &\to \frac{1}{2}\left[\lim_{x\to 0}L + \lim_{x\to0}2x\right]\\
&\to \frac{1}{2}[L + 0]\\
&\to \frac{L}{2}
\end{align*}
\noindent So, at $x=0$, the series converges to $\frac{L}{2}$. We are not done
yet, however. Now, we need to consider the periodic extension of the series. The
function will repeat every $[-L,L]$, so we need to look at what the convergence
of the series looks like at $-L$ and $L$. Let's take a look:
\begin{alignat*}{3}
\sum&\to\frac{1}{2}\left[\lim_{x\to-L^{-}}f(x) + \lim_{x\to-L^{+}}f(x)\right] \qquad\qquad \sum&&\to
\frac{1}{2}\left[\lim_{x\to L^{-}}f(x) + \lim_{x\to L^{+}}f(x)\right]\\
&\to \frac{1}{2}[2L + L] &&\to \frac{1}{2}[2L + L]\\
&\to\frac{3L}{2} &&\to\frac{3L}{2}
\end{alignat*}
\noindent And we are done.

### 2016-01-02 partial_differential_equations/mathe_code.tex
update several

\section{Mathematica and Partial Differential Equations}
\hrule
\noindent \\\\
We will briefly talk about solving PDE's in Mathematica in this section. We can solve PDE's with
the tools available to solve ODE's. Let's consider an example equation:\\\\

\noindent \textbf{\textit{Ex:}} Solve the following PDE
\[
u_{t} + 2u_{x} = 0
\]
\indent \textbf{\textit{Solution:}} We can define both $u_{x}$ and $u_{t}$ in Mathematica by using
the D command.
\begin{minted}{mathematica}
 In[1]:= ClearAll["Global"];
         pde = D[u[x,t], t] + 2 D[u[x,t], x] == 0;
\end{minted}
Here, we have assigned our PDE to the Mathematica variable pde. Note that we never gave a definition
to the function $u(x,t)$. It's not necessary; Mathematica understands what we mean. To solve the
equation, we will use the DSolve command, just as we would with an ODE.
\begin{minted}{mathematica}
 In[2]:= solution = DSolve[pde, u[x,t], {x,t}]
Out[2]:= {{u[x,t] -> C[1][(1/2)(2t-x)]}}
\end{minted}
Mathematica represents the arbitrary function that we get when solving with C[1].
If we want to assign the solution for later use, we can create a new function and assign it to the
solution that we just found.
\begin{minted}{mathematica}
 In[3]:= f[x_,t_] := u[x,t] /. solution[[1]];
\end{minted}
The line above assigns u[x,t] to the function f[x,t]. We can also apply an initial condition by
providing DSolve with a list of equations rather than just the PDE.
\begin{minted}{mathematica}
 In[4]:= solution = DSolve[{pde, u[0,t] == Sin[t]},u[x,t],{x,t}]
Out[4]:= {{u[x,t] -> Sin[(1/2)(2t-x)]}}
\end{minted}
Here we have provided an initial condition of $\sin{t}$. Again, we could assign the result of Out[4] to
a function that we could use later. It may benefit us, for example, to plot the result.
\begin{minted}{mathematica}
 In[5]:= Plot3D[u[x,t]/.solution,{x,-4,4},{t,-4,4}];
\end{minted}
We've supressed the output here, but the result would be the graph of the function in the 3D plane.

### 2016-01-02 partial_differential_equations/second_order.tex
update several

\section{Second Order Partial Differential Equations}
\hrule
\noindent\\
\subsection{The Heat and Wave Equations}
\indent Let's take a look at the second order PDE. We will start with a
homogeneous equation, which we will tackle with the separation of variables
method. The general strategy is that we want to take our PDE and convert it to
an ODE, which we know how to solve. The problems presented here also involve
boundary conditions, which will require us to solve for eigenvalues/functions.
We will also be dealing with series (specifically Fourier Series) to help us
solve these problems. Let's look at an example:\\
\noindent\textbf{\textit{Ex:}} Solve the following boundary value problem:
\[
\begin{cases}
u_{tt} - u_{xx} = 0\\
u(0,t) = 0\qquad u(1,t) = 0\\
u(x,0) = f(x)\qquad u_{t}(x,0) = g(x)
\end{cases}\\
\]
\[f(x) = 2\sin{(\pi x)} + 3\sin{(2\pi x)}\qquad g(x) = 4\sin{(3\pi x)} - 7\sin{(5\pi x)}\]
given the general solution:
\[u(x,t) = \sum_{n=1}^{\infty}\left[\left(a_{n}\cos{[n\pi t]} + b_{n}\sin{[n\pi t]}\right)\sin{[n\pi
x]}\right]\]
\indent\textbf{\textit{Solution:}} There's a lot going on here, so let's break
it down. First of all, notice that we have two initial conditions, one of which
is a derivative. This should make sense since the equation have two second order
partial derivatives included in it. Next notice that we have a Fourier Sine Series,
despite both the $\sin{[n\pi t]}$ and $\cos{[n\pi t]}$ terms. If we had worked
out the entire problem, we would see that our eigenvalue/function for this
equation takes the form of
\[
\{\lambda_{n} = n\pi, \sin{(n\pi x)}\}
\]
\noindent which tells us that the corresponding series is a Fourier Sine Series.
We also see the initial conditions on our equation are already representative
of a series of sine functions, which is not just a conincidence. To apply the
initial conditions, we will need both the general solution and the derivative
with respect to $t$ of the general solution. Let's go ahead and apply the first
initial condition, $u(x,0) = f(x)$. Letting $t= 0$ gives us:
\[
u(x,0) = 2\sin{(\pi x)} + 3\sin{(2\pi x)} = \sum_{n=1}^{\infty}\left[a_{n}\sin{(n\pi x)}\right]
\]
\noindent We see that the term with $b_{n}$ has dropped out, since
$\sin{(0)} = 0$. Now we just have to match up the coefficients on $a_{n}$. We
can see that when we have $n=1$, $a_{n} = 2$, and when $n=2$, $a_{n} = 3$. For
all other $n$, we have $a_{n} = 0$, so we ignore those terms. Next, we can apply
the initial condition for the derivative of the general solution with respect to
$t$. We expect that all the terms related to $a_{n}$ will disappear, and in fact
they will. Let's take a look:
\begin{align*}
u_{t}(x,t) &= \sum_{n=1}^{\infty}\left[(-a_{n}n\pi\sin{[n\pi t]} + b_{n}n\pi\cos{[n\pi t]})\sin{n\pi
x}\right]\\
u_{t}(x,0) &= 4\sin{(3\pi x)} - 7\sin{(5\pi x)} = \sum_{n=1}^{\infty}\left[n\pi b_{n}\sin{(n\pi
x)}\right]
\end{align*}
\noindent We are going to do the same thing here that we did to find our $a_{n}$
terms, but notice that we don't just have $b_{n}$; we actually have $n\pi b_{n}$.
We have the following:
\begin{alignat*}{3}
&3\pi b_{3} = 4 \qquad \qquad \qquad &&5\pi b_{5} = -7\\
&b_{3} = \frac{4}{3\pi} &&b_{5} = -\frac{7}{5\pi}
\end{alignat*}
\noindent Much like the solutions for $a_{n}$, all the other $b_{n}$ terms are 0.
Now we can construct a particular solution. In this case, the answer will not be
in the form of a series, since we have definite terms. Our solution is as follows:
\[
u(x,t) = 2\cos{(\pi t)}\sin{(\pi x)} + 3\cos{(2\pi t)}\sin{(2\pi x)} + \frac{4}{3\pi}\sin{(3\pi
t)}\sin{(3\pi x)} - \frac{7}{5\pi}\sin{(5\pi t)}\sin{(5\pi x)}
\]
\noindent And we are done.
\newpage

\noindent \textbf{\textit{Ex:}} Solve the following PDE
\[
\begin{cases}
\frac{\partial u}{\partial t} = k \frac{\partial^{2} u}{\partial x^{2}}\\
u(x,0) = x\\
u(0,t) = u(L,t) = 0\\
\end{cases}
\]
\indent \textbf{\textit{Solution:}} We will use the separation of variable
technique to solve this problem. We know to use separation of variable because
we are dealing with a homogeneous PDE with homogenous boundary conditions.
Separation of variables tells us that our solution will be in the form of
$u(x,t) = X(x)T(t)$. If we take this fact at face value, we can use the
separation of variables method to solve this. First, we substitute in our solution
into the PDE:
\begin{align*}
\frac{\partial}{\partial t}(X(x)T(t)) &= k \frac{\partial^{2}}{\partial x^{2}}(X(x)T(t))\\
X(x)\frac{dT}{dt} &= k\frac{d^{2}X}{dx^{2}}T(t)\\
\frac{dT}{dt}\frac{1}{T(t)} &= \frac{d^{2}X}{dx^{2}}\frac{k}{X(x)}
\end{align*}
\noindent The only way that our function of x and our function of t can be equal
is if they are equal to a constant; thus we now have:
\[
\frac{dT}{dt}\frac{1}{T(t)} = \frac{d^{2}X}{dx^{2}}\frac{k}{X(x)} = -\lambda
\]
\noindent And now we have a first order ODE for the time and a second order ODE
for the position:
\[
\frac{dT}{dt} = -k \lambda T(t) \qquad\text{and}\qquad \frac{d^{2}X}{dx^{2}} = -\lambda X(x)
\]
\noindent Our next step is to find the Eigenvalues and Eigenfunctions of the
second order ODE; we will come back to the time ODE when we have values for
$\lambda$.\\
\indent To solve for these values, we have to find non-trivial eigenvalues that
correspond to the boundary values of the ODE. In this case, we know that
$X(0) = 0$ and $X(L) = 0$. These boundary conditions come from the BC on the PDE
which states $u(0,t) = u(L,t) = 0$. To find the eigenvalues/functions, we need
to find the characteristic equation corresponding to the ODE in question. Let's
go ahead and find it:
\begin{gather*}
\frac{d^{2}X}{dx^{2}} = -\lambda \frac{X(x)}{k}\\
\frac{d^{2}X}{dx^{2}}  + \lambda \frac{X(x)}{k} = 0\\
r^{2} + \lambda = 0\\
r^{2} = -\lambda\\
r = \pm \sqrt{- \lambda}
\end{gather*}
\indent Now, we can substitute values of $\lambda$ to find out solutions. Let's
start with $\lambda > 0$. Since we have a $-\lambda$ under the radical, we will
have $r = \pm \sqrt{\lambda} \mathrm{i}$. This will yield the solution of
\[
y(x) = c_{1}\cos{\sqrt{\lambda}x} + c_{2}\sin{\sqrt{\lambda}x}
\]
\noindent Applying our boundary conditions gives us:
\[
y(0) = c_{1}\cos{0} + c_{2}\sin{0} = 0
\]
\noindent Now we know that $c_{1} = 0$. This result came from the fact that
$\sin{0} = 0$, so the $c_{2}$ term disappears. Now we are left with $\cos{0} = 1$,
which tells us that $c_{1} = 0$. We can apply this to our second boundary
condition like so:
\begin{gather*}
y(L) = 0 = c_{2}\sin{\sqrt{\lambda}L}\\
\end{gather*}
\noindent It looks like there will only be trivial solutions, but we know this
isn't true due to the Jacobian matrix. Recall that if the determinant of the
Jacobian $\neq 0$, then we only have trivial solutions, and if the determinant
of the Jacobian $= 0$, then we do not have the trivial solution. Let's take a
look at the Jacobian for this case:
\begin{gather*}
\left|
\begin{array}{cc}
1 & 0\\
\cos{\sqrt{\lambda}L} & \sin{\sqrt{\lambda}L}
\end{array}
\right| = 0\\
\sin{\sqrt{\lambda}L} = 0\\
\sqrt{\lambda}L = \arcsin{0}\\
\sqrt{\lambda} = \frac{\arcsin{0}}{L}\\
\lambda = \left(\frac{\arcsin{0}}{L}\right)^{2}\\
\lambda = \left(\frac{n\pi}{L}\right)^{2}
\end{gather*}
\noindent Notice that we replaced the $\arcsin{0}$ term with $n\pi$. Recall from
trig that the $\sin$ function is $0$ at every $n\pi$, starting with $n = 0$.
Next, we just need to find our corresponding eigenfunction. Our $c_{1}$ was $0$,
so we can ignore the entire $\cos$ term. Thus, we have an eigenfunction from
plugging in our $\lambda$ that looks like:
\[
X(x) = \sin{\frac{n\pi x}{L}}
\]
\indent Unfortunately, we aren't done yet. Now we have to find the corresponding
eigenvalues/functions for $\lambda = 0$. This isn't too hard, and can be done in
a few lines if we recall the characteristic equation. Plugging in $0$ for
$\lambda$ gives us that $r = 0$. Using this piece of information, we now know
that $y(x) = c_{1} + xc_{2}$. Applying our boundary conditions gives us
$c_{1} = 0, c_{2} = 0$. We can see that we have only trivial solutions for
$\lambda = 0$.\\
\indent Our final case is for $\lambda < 0$. I'll go ahead and tell you that
this will yield another trivial solution, but we can take a look at the work
behind it. Since $\lambda > 0$, we know that our $r$ from the characteristic
equation will be $r = \pm \sqrt{\lambda}$. From here we are going to do something
 a little unconventional that will help us. Usually, we would have a solution of
the form $y(x) = c_{1}e^{\sqrt{\lambda}x} + c_{2}x e^{-\sqrt{\lambda}x}$. In our
case, it will be more convenient to use hyperbolic trig functions. It's not too
far off from our solution from $\lambda > 0$:
\begin{align*}
y(x) &= c_{1}e^{\sqrt{\lambda}x} + c_{2}e^{-\sqrt{\lambda}x}\\
&= c_{1}\cosh{\sqrt{\lambda}x} + c_{2}\sinh{\sqrt{\lambda}x}
\end{align*}
\noindent If we apply the boundary conditions (remember that
$\sinh{0} = 0$ and $\cosh{0} = 1$), we will find that we have only trivial
solutions.\\
\indent So, we are done with the second order ODE, and now we turn our eye to
the first order time equation. We know our $\lambda$, so it isn't too hard to solve:
\begin{gather*}
\frac{dT}{dt} = -\lambda T(t)\\
T(t) = ce^{-k\left(\frac{n\pi}{L}\right)^{2}t}
\end{gather*}
\indent Finally we can put it all together. Our penultimate solution will be
\[
u_{n}(x,t) = B_{n}\sin{\left(\frac{n\pi x}{L}\right)}e^{-k\left(\frac{n\pi}{L}\right)^{2}t}
\]
\noindent We can analyze this solution and tell where each piece comes from.
The $B_{n}$ is a coefficient that results from our $c's$ in the general solutions
of the differential equations. The $sin$ part results from out eigen vector.
Remember, we get a different value for each $n$, so there may or may not be $0$
terms in the series. Finally, we have the exponential term. This term comes from
the solution to our time equation. When we put it all together, we can see that
this is, of course a series; for every $n$, we get a different solution. From
here we can find the Fourier Series (which will be a Fourier Sine Series). We
won't worry about that yet, but we will come back to it.
\newpage

\indent It will benefit us to take a moment and discuss how boundary conditions
effect our Fourier series. We have three different combinations of boundary
conditions, where each one will produce a different result in our series solution.
\begin{enumerate}
\item \textit{Dirichlet Boundary Conditions}\\
Dirichlet boundary conditions involve the function for which we are looking. In
other words, the boundary conditions are presented in terms of the functions we
are trying to find, rather than the derivative of that function. Typically, this
type of boundary condition will take the form of (in the homogeneous case)
\[u(0,t) = u(L,t) = 0.\]
In regards to our Fourier series, we expect to see $\sin{(\frac{n\pi x}{L})}$.
Our basis will look like
\[
\left\{ \lambda_{n} = \left(\frac{n\pi}{2}\right)^{2}, \phi_{n}(x) = \sin{\left(\frac{n\pi x}{L}\right)}
\right\}
\]
\item \textit{Neumann Boundary Conditions}\\
Neumann boundary conditions involve the derivatives of the function for which we
are solving. Neumann boundary conditions will look like
\[u_{x}(0,t) = u_{x}(L,t) = 0.\]
This will cause our Fourier series to contain $\cos{(\frac{n\pi x}{L})}$. Our
basis will look like
\[
\left\{ \lambda_{n} = \left(\frac{n\pi}{2}\right)^{2}, \phi_{n}(x) = \cos{\left(\frac{n\pi x}{L}\right)}
\right\}
\]
\item \textit{Robin Boundary Conditions}\\
Robin boundary conditions will contain a mixture of the function and it's
derivative. Typically, we will see
\[ u_{x}(0,t) = u(L,t) = 0\qquad \text{or}\qquad u(0,t) = u_{x}(L,0) = 0.\]
These boundary conditions will give us two different results, depending on whether
we have the derivative first or second. We will have
\begin{align*}
u_{x}(0,t) = u(L,t) = 0 &\implies \left\{\lambda_{n} = \left(\left[n +
\frac{1}{2}\right]\frac{\pi}{L}\right)^{2}, \phi_{n}(x) = \cos{\left(\left[n +
\frac{1}{2}\right]\frac{\pi}{L}x\right)}\right\}\\
u(0,t) = u_{x}(L,t) = 0 &\implies \left\{\lambda_{n} = \left(\left[n +
\frac{1}{2}\right]\frac{\pi}{L}\right)^{2}, \phi_{n}(x) =\sin{\left(\left[n +
\frac{1}{2}\right]\frac{\pi}{L}x\right)}\right\}
\end{align*}
\end{enumerate}

\indent Now we can take a look at a slightly harder PDE. We still have the heat
equation, but we will not have homogeneous boundary conditions. Our goal here is
to find a way to transform our boundary conditions into a homogenous form and take
it from there.\\\\
\noindent \textbf{\textit{Ex:}} Solve the following PDE
\[
\begin{cases*}
u_{t} - u_{xx} = 0\\
u(0,t) = 0,\ u(1,t) = t\\
u(x,0) = x^{2}
\end{cases*}
\]
\indent \textbf{\textit{Solution:}} We would like to just jump into this PDE and
solve it, but it's not quite that simple. We have to find a way to express our
boundary conditions in a homogeneous manner. We can do this through a clever
substitution; we will create a new PDE, denoted $\Upsilon$. To find $\Upsilon$,
we use the following formula:
\[ \Upsilon = B_{1} + \frac{x}{l}\Delta B, \]
\noindent where $B$ is a boundary condition. In our case, we have as follows:
\begin{align*}
\Upsilon &= B_{1} + \frac{x}{l}\Delta B\\
&= 0 + \frac{x}{1}(t - 0)\\
&= xt
\end{align*}
\noindent Now, we find a new PDE, denoted by $\upsilon$. We have another formula that tells us
\[ \upsilon(x,t) = u(x,t) - \Upsilon(x,t) \]
\noindent Applying this formula gives us
\begin{align*}
\upsilon(x,t) &= u(x,t) - \Upsilon\\
&= u(x,t) - xt
\end{align*}
\noindent Let's take a moment to analyze the result. We wanted to find a new PDE in terms of our old one
that would allow us to have homogeneous boundary conditions. We found a solution, called $\upsilon$, in
terms of our original PDE, called $u$. Our original PDE was $u_{t} - u_{xx} = 0$, so to properly perform
our substitution, we need to find $\upsilon_{t}$ and $\upsilon_{xx}$. Through some careful
differentiation, we have
\[ \upsilon_{t} = u_{t} - x \qquad\text{and}\qquad \upsilon_{xx} = u_{xx}\]
\noindent To find our new PDE, we simply evaluate.
\[
\upsilon_{t} - \upsilon_{xx} = u_{t} - x - u_{xx} = u_{t} - u_{xx} - x = 0 - x = -x
\]
\noindent So we finally have a new PDE that looks like
\[
\begin{cases*}
\upsilon_{t} - \upsilon_{xx} = -x\\
\upsilon(0,t) = \upsilon(1,t) = 0\\
\upsilon(x,0) = x^{2}
\end{cases*}
\]
\noindent You may wonder how this has made anything simpler, since the PDE itself is no longer
homogeneous. By forcing the boundary conditions to be homogeneous, we can now use a Fourier series to
find our solution. Recall that earlier we discussed the boundary condition and their resulting
eigenvalue/eigenfunction. Here, we have Dirichlet boundary conditions, which tells us that we have a
Fourier sine series. We have that
\begin{align*}
\upsilon(x,t) &= \sum_{n = 1}^{\infty}\left[a_{n}(t)\sin{(n\pi x)}\right]\\
\upsilon_{t}(x,t) &= \sum_{n = 1}^{\infty}\left[a_{n}'(t)\sin{(n\pi x)}\right]\\
\upsilon_{xx}(x,t) &= -\sum_{n = 1}^{\infty}\left[a_{n}(t)(n\pi)^{2}\sin{(n\pi x)}\right]\\
\end{align*}
\noindent Which tells us that
\[
\upsilon_{t} - \upsilon_{xx} = \sum_{n = 1}^{\infty}\left[(a_{n}'(t) + (n\pi)^{2}a_{n}(t))\sin{(n\pi
x)}\right] = -x
\]
\noindent This may look rather complicated, but all we have is a function's Fourier sine series.
Essentially, we have
\[
-x = \sum_{n = 1}^{\infty}\left[ b_{n}\sin{(n\pi x)} \right]
\]
\noindent To find $b_{n}$, we use the formula that was discussed in the earlier section about Fourier
series. We have that
\begin{align*}
a_{n}'(t) + (n\pi)^{2}a_{n}(t) &= \int_{0}^{1}-x\sin{(n\pi x)}dx\\
&= \frac{2x}{n\pi}\cos{(n\pi x)}\bigg\vert_{0}^{1}\\
&= \frac{2}{n\pi}
\end{align*}
\noindent We now have an ODE which we can solve for $a_{n}$. Note that from out PDE, we do have an
initial condition for the ODE, but it is not as simple as just saying $a_{n}(0) = x^{2}$. At $t = 0$, we
have the following series:
\[
\sum_{n=1}^{\infty} a_{n}(0)\sin{(n\pi x)}
\]
\noindent We got this by taking our original $\upsilon(x,t)$ and plugging in $0$ for $t$. We know how to
get $a_{n}(0)$, since this is just a simple Fourier sine series. It is given by
\begin{align*}
a_{n}(0) &= 2\int_{0}^{1}x^{2}\sin{(n\pi x)}dx\\
&= \frac{(-1)^{(n + 1)}}{n\pi} + \frac{2}{(n\pi)^{3}}\left[ (-1)^{n} - 1 \right]
\end{align*}
\noindent We now have all we need to solve the ODE. We will omit the work for the solution, but the
final answer for $a_{n}$ is
\[
a_{n}(t) = \frac{4\left[ (-1)^{n} - 1 \right]}{(n\pi)^{3}}e^{-(n\pi)^{2}t} +
\frac{2(-1)^{n}}{(n\pi)^{3}}
\]
\noindent Now all that's left to do is go back to $u$ from $\upsilon$. This may seem hard, but we saw earlier that $\upsilon = u - \Upsilon$. Solving for $u$ gives us $u = \upsilon + \Upsilon$. Recall that $\Upsilon$ was $xt$, and $\upsilon$ was the series for which we just found $a_{n}$. Putting it all together gives us
\[
u(x,t) = xt + \sum_{n=1}^{\infty}\left[\frac{4\left[ (-1)^{n} - 1 \right]}{(n\pi)^{3}}e^{-(n\pi)^{2}t} + \frac{2(-1)^{n}}{(n\pi)^{3}}\right]\sin{(n\pi x)}
\]
\noindent And we are done.
\newpage

\noindent \textbf{\textit{Ex:}} Solve the following PDE
\[
\begin{cases*}
u_{tt} - u_{xx} = 1 + t\cos{(2\pi x)}\\
u_{x}(0,t) = u_{x}(1,t) = 0\\
u(x,0) = 2\cos{(2\pi x)},\quad u_{t}(x,0) = 1 + \cos{(\pi x)} - \cos{(2\pi x)}
\end{cases*}
\]
\indent \textbf{\textit{Solution:}} This problem is an example of the wave equation with a non-homogeneous solution. We are lucky this time; the boundary conditions are already homogeneous, so there is no need to change our variable. We know that since both of the boundary conditions are derivatives, we are dealing with a Fourier cosine series. This yields the eigenvalue/function of
\[
\left\{\lambda_{n} = (n\pi)^{2}, \cos{(n\pi x)}\right\}
\]
Since we are dealing with a cosine series, we also have a leading term, $a_{0}$. Our full series looks like
\[
u(x,t) = a_{0}(t) + \sum_{n=1}^{\infty} a_{n}(t)\cos{(n\pi x)}
\]
We need $u_{tt}$ and $u_{xx}$, so we have:
\[
u_{tt}(x,t) = a_{0}''(t) + \sum_{n=1}^{\infty} a_{n}''(t)\cos{(n\pi x)}
\qquad \text{and} \qquad
u_{xx}(x,t) = -\sum_{n=1}^{\infty} (n\pi)^{2}a_{n}(t)\cos{(n\pi x)}
\]
All we have to do now is simply plug into our original equation. This gives us
\begin{align*}
a_{0}''(t) + \sum_{n=1}^{\infty} a_{n}''(t)\cos{(n\pi x)} + \sum_{n=1}^{\infty} (n\pi)^{2}a_{n}(t)\cos{(n\pi x)} &= 1 + t\cos{(2\pi x)}\\
a_{0}''(t) + \sum_{n=1}^{\infty}(a_{n}''(t) + (n\pi)^{2}a_{n}(t))\cos{(n\pi x)} &= 1 + t\cos{(2\pi x)}
\end{align*}
Now we can play the matching game. We have a $constant + series\ term$ on the right side, and $constant + series$ on the left, so we can match up the constants. This gives us
\[
a_{0}''(t) = 1
\]
Now we match up the series term. On the right hand side, we have $\cos{(2\pi x)}$, so we know that for $n= 2$, we have
\[
a_{2}''(t) + 4\pi^{2}a_{2}(t) = t
\]
$a_{0}$ and $a_{2}$ are the only terms that appear, so we can assume for now that $\forall n \neq 2\ \text{and}\ 3, a_{n} = 0$. We are left with two ODE's, so we now need to find the corresponding initial conditions. Let's apply the first initial condition to our PDE. This gives us
\[a_{0}(0) + \sum_{n=1}^{\infty}a_{n}(0)\cos{(n\pi x)} = 2\cos{(2\pi t)}\]
Which tells us
\[a_{0} = 0 \qquad\qquad \qquad\qquad a_{2} = 2\]
Now we can apply the second initial condition. After taking the partial derivative of $u$ with respect to $t$, we have
\[
a_{0}'(0) + \sum_{n=1}^{\infty}a_{n}'(0)\cos{(n\pi x)} = 1 + \cos{(\pi x)} - \cos{(2\pi x)}
\]
Some more clever pattern matching give us
\[
a_{0}'(0) = 1 \qquad\qquad a_{1}'(0) = 1 \qquad\qquad a_{2}'(0) = -1
\]
Notice that we have a term $a_{1}'$. We cannot simply ignore this value; rather we have to take it into account when we find our solution. We know that we have an ODE for $a_{1}(t)$, but it may not be evident how to express it. Recall that earlier we said $\forall n \neq 2\ \text{and}\ 3, a_{n} = 0$. This means that $a_{1} = 0$, and since $a_{1}$ is a coefficient in the Fourier series, we know that $a_{1}''(t) + \pi^{2}a_{1}(t) = 0$. We now have three different ODE's that we must solve:
\begin{alignat*}{3}
&\begin{cases}
a_{0}''(t) = 1\\
a_{0}'(0) = 1,\quad a_{0}(0) = 1
\end{cases}
&\begin{cases}
a_{1}''(t) + \pi^{2}a_{1}(t) = 0\\
a_{1}'(0) = 1,\quad a_{1}(0) = 0
\end{cases}
&&\begin{cases}
a_{2}''(t) + 4\pi^{2}a_{2}(t) = t\\
a_{2}'(0) = 1,\quad a_{2}(0) = 2
\end{cases}
\end{alignat*}
Luckily, Mathematica is here to save the day. We will skip the work and let the computer do the thinking for us. We get that
\begin{align*}
a_{0}(t) &= \frac{1}{2}t^{2} + t\\
a_{1}(t) &= \frac{\pi\cos{(t\pi)} + \sin{(t\pi)}}{\pi}\\
a_{2}(t) &= \frac{2\pi t + 16\pi^{3}\cos{(2\pi t)} - \sin{(2\pi t)} + 4\pi^{2}\sin{(2\pi t)}}{8\pi^{3}}
\end{align*}
Now all we need to do is plug back in to our $u$. Our final answer becomes
\[
u(x,t) = \frac{1}{2}t^{2} + t + \frac{\pi\cos{(t\pi)} + \sin{(t\pi)}}{\pi}\cos{(\pi x)} +  \frac{2\pi t + 16\pi^{3}\cos{(2\pi t)} - \sin{(2\pi t)} + 4\pi^{2}\sin{(2\pi t)}}{8\pi^{3}}\cos{(2\pi x)}
\]
Just for interest, the plot of this solution looks like
\begin{center}
\includegraphics[scale=0.6]{pg23_plot}
\end{center}
On $-\pi \leq \theta \leq \pi$ and $-\pi \leq x \leq \pi$; and we are done.
\newpage

\subsection{The Laplace Equation}
\noindent \\
\indent Up to this point, we have only looked at the wave equation and the heat equation. There is, however, one other equation in the classical trinity: the Laplace Equation. We haven't dealt with this equation yet because we have to be very careful in our approach to solving it. When we looked at the homogeneous wave and heat equations, we used the general approach called \textit{Separation of Variables}. In the case of the homogeneous Laplace equation, we may not be able to tackle the problem using separation of variables. The problem arises from the various types of boundaries that effect the domain. We can either have a 'rectangular' bound, which will be separable, or a 'circular' bound which is \textbf{not} separable. Imagine that we have the following two problems:
\begin{align*}
&A =\begin{cases}
u_{xx} + u_{yy} = 0\quad \text{on}\ [0,1]\times[0,2]\\
u(x,0) = u(x,2) = 0\\
u(0,y) = y, u(1,y) = 0
\end{cases}
&B =\begin{cases}
u_{xx} + u_{yy} = 0\quad \text{on}\ x^{2} + y^{2} < a^{2}\\
u(x,y)\ \big |_{x^{2} + y^{2} = a^{2}} = f(x,y)
\end{cases}
\end{align*}
$A$ is an example of a separable Laplace equation. We can see graphically that the domain $[0,1] \times [0,2]$ is rectangular. $B$'s domain is not separable, since it is circular. Let's take a look at $A$.\\\\
\noindent \textbf{\textit{Ex:}} Solve the following PDE
\[
\begin{cases*}
u_{xx} + u_{yy} = 0\quad \text{on}\ [0,1]\times[0,2]\\
u(x,0) = u(x,2) = 0\\
u(0,y) = y, u(1,y) = 0
\end{cases*}
\]
\indent \textbf{\textit{Solution:}} Earlier, we discussed the different boundary conditions we can encounter and their corresponding eigenvalues/functions. In this problem, we have Dirichlet boundary conditions. This tells us that
\[
\left\{\lambda_{n} = \left(\frac{n\pi}{2}\right)^{2}, \phi_{n}(y) = \sin{\left(\frac{n\pi y}{2}\right)}\right\}
\]
Notice that we used $y$ here instead of $x$. We did this because the boundary conditions are defined in terms of $y$, not $x$. Now we can continue to solve our problem. Recall that our separation of variables gives us two second order ODE's,
\[
X'' - \lambda X = 0 \qquad\qquad \text{and} \qquad\qquad \lambda Y - Y'' = 0
\]
We already solved for $Y(y)$ by finding the corresponding eigenvalues/vectors. Now we plug in our $\lambda_{n}$ into our $X$ ODE. Solving the resulting ODE will give us a basis of
\[
\left\{ e^{\frac{n\pi}{2}x}, e^{-\frac{n\pi}{2}x} \right\}
\]
Now we could continue at this point, but if we think ahead, we can see that when we eventually apply our initial conditions, nothing will cancel out. At this point, we want to transform our basis. We have as follows
\[
\left\{ e^{\frac{n\pi}{2}x}, e^{-\frac{n\pi}{2}x} \right\} \implies \left\{ \cosh{\left(\frac{n\pi x}{2}\right)}, \sinh{\left(\frac{n\pi x}{2}\right)} \right\}
\]
All we did here is apply an identity. Now we have everything we need to set up a solution. Our $u(x,y)$ will be
\[
u(x,y) = \sum_{n=1}^{\infty}\left[a_{n}\cosh{\left(\frac{n\pi x}{2}\right)} + b_{n}\sinh{\left(\frac{n\pi x}{2}\right)}\right]\sin{\left(\frac{n\pi y}{2}\right)}
\]
We can apply our initial conditions, which gives us $b_{n}\ \text{and}\ a_{n}$:
\begin{align*}
a_{n} &= \frac{4(-1)^{n+1}}{n\pi}\\
b_{n} &= -\coth{\left(\frac{n \pi}{2}\right)}\frac{4(-1)^{n+1}}{n\pi}
\end{align*}
This yields a final solution of
\[
u(x,y) = \sum_{n=1}^{\infty}\left[ \frac{4(-1)^{n+1}}{n\pi} \cosh{\left(\frac{n\pi x}{2}\right)} -\coth{\left(\frac{n \pi}{2}\right)}\frac{4(-1)^{n+1}}{n\pi}\sinh{\left(\frac{n\pi x}{2}\right)}\right]\sin{\left(\frac{n\pi y}{2}\right)}
\]
And we are done
\newpage

\indent Now we turn our attention to the Laplace Equation with a circular domain. As we discussed earlier, we can't simply say that our solution will be in the form of $X(x)Y(y)$. However, we can force the domain to be rectangular if we use polar coordinates. Let's consider a variation of $B$ from above.\\\\

\noindent \textbf{\textit{Ex:}} Solve the following PDE
\[
\begin{cases*}
u_{xx} + u_{yy} = 0\quad \text{on}\ x^{2} + y^{2} < a^{2}\\
u(a,\theta) = U_{1}\quad \text{on}\ 0 \leq \theta \leq \pi\\
u(a,\theta) = U_{2}\quad \text{on}\ \pi \leq \theta \leq 2\pi
\end{cases*}
\]
\indent \textbf{\textit{Solution:}} We want to express the partial derivatives here in terms of $r\ \text{and}\ \theta$. Recall that we have
\[
x = r\cos{\theta} \qquad\qquad \text{and} \qquad\qquad y = r\sin{\theta}
\]
First, we will find the partial derivatives of $x$ and $y$ with respect to $r$ and $\theta$. Since $x = r\cos{\theta}$,
\[
\frac{\partial x}{\partial r} = \cos{\theta} \qquad\qquad \frac{\partial x}{\partial \theta} = -r\sin{\theta}
\]
We can compute the derivatives for $y$ in a similar fashion:
\[
\frac{\partial y}{\partial r} = \sin{\theta} \qquad\qquad \frac{\partial y}{\partial \theta} = r\cos{\theta}
\]
Now we need to find $u_{rr}$ and $u_{\theta\theta}$. Let's focus on $u_{rr}$ first. We will need to start by finding $u_{r}$. We have as follows
\begin{align*}
\frac{\partial u}{\partial r}& = \frac{\partial u}{\partial x}\frac{\partial x}{\partial r} + \frac{\partial u}{\partial y}\frac{\partial y}{\partial r}\\
&= \cos{\theta}\frac{\partial u}{\partial x} + \sin{\theta}\frac{\partial u}{\partial y}
\end{align*}
We have $u_{r}$, so now we need to compute $u_{rr}$. We have as follows:
\begin{align*}
\frac{\partial^{2} u}{\partial r^{2}} &= \cos{\theta}\frac{\partial }{\partial r}\frac{\partial u}{\partial x} + \sin{\theta}\frac{\partial }{\partial r}\frac{\partial u}{\partial y}\\
&= \cos^{2}{\theta}\frac{\partial^{2} u}{\partial x^{2}} + 2\cos{\theta}\sin{\theta}\frac{\partial^{2}u}{\partial x\partial y} + \sin^{2}{\theta}\frac{\partial^{2} u}{\partial y^{2}}
\end{align*}
We won't delve into finding $u_{\theta\theta}$, as the partials get a little out of hand. However, we end up with
\begin{align*}
\frac{\partial^{2}u}{\partial r^{2}} + \frac{1}{r^{2}}\frac{\partial^{2}u}{\partial\theta^{2}} &= -\frac{1}{r}\frac{\partial u}{\partial r} + \frac{\partial^{2} u}{\partial x^{2}} + \frac{\partial^{2} u}{\partial y^{2}}\\
u_{xx} + u_{yy} &= u_{rr} + \frac{1}{r}u_{r} + \frac{1}{r^{2}}u_{\theta\theta} = 0
\end{align*}
Let's go ahead and formalize it:
\[
\nabla^{2}u = u_{rr} + \frac{1}{r}u_{r} + \frac{1}{r^{2}}u_{\theta\theta} = 0
\]
Now we can treat this as a 2D problem; in other words, we can now use separation of variables to solve for $u$. As before, we know our solution will look like
\[
u(r,\theta) = R(r)T(\theta)
\]
Plugging everything in and realizing by the usual trick that everything s equal to a constant gives us
\[
-\frac{T''}{T} = \frac{R'' + \frac{R'}{r}}{\frac{R}{r^{2}}} = \lambda
\]
Since we are dealing in polar coordinates, we will go ahead and walk through the eigenvalues/functions. We will be solving for $T(\theta)$, since our boundary conditions are in terms of $\theta$. We also should note that the solutions will repeat themselves every $2\pi$, since we are in the polar system. In addition, our $u$ should be finite at $r=0$. Let's take a look at the eigenvalues for the Laplace equation.\\\\
\begin{minipage}[t]{0.32\textwidth}
\underline{$\lambda = 0$}\\
When $\lambda = 0$, we have \[T(\theta) = a\theta + b.\] This also tells us that \[R(r) = c\ln{(r)} + d.\] Putting everything together gives us \[ u(r,\theta) = (a\theta + b)(c\ln{r} + d).\] Since the solution is periodic on $2\pi$, we know that $a=0$. Since the solution has to remain finite at $r=0$, we know that $c = 0$. This gives us that $u = bd$, or that $u$ is equal to a constant.
\end{minipage}
\hspace{0.01\textwidth}
\begin{minipage}[t]{0.32\textwidth}
\underline{$\lambda > 0$}\\
With $\lambda > 0$, we have our characteristic equation $(m = \pm\ \sqrt{\lambda})$ yielding a real valued solution. This tells us that
\[
T(\theta) = c_{1}\cosh{(\sqrt{\lambda}\theta)} + c_{2}\sinh{(\sqrt{\lambda}\theta)}.
\]
The solution is periodic on $2\pi$, so we can say that $c_{1} = c_{2} = 0$. Therefore, we have a trivial solution.
\end{minipage}
\hspace{0.01\textwidth}
\begin{minipage}[t]{0.32\textwidth}
\underline{$\lambda < 0$}\\
$\lambda < 0$ gives us
\[
T(\theta) = c_{1}\cos{(\sqrt{\lambda}\theta)} + c_{2}\sin{(\sqrt{\lambda}\theta)}
\]
This solution is periodic with $2\pi$, and it yields a trivial eigenvector.
\end{minipage}
\noindent\\\\\\
\noindent So we know that the only $\lambda$ that yields eigenvalues is $\lambda = 0$, and we know that this particular eigenvalue yields a constant solution. If we solve the differential equation for $R$, we will find that the only solution we care about is $r^{n}$. We can finally say that our solution will look like
\[
u(r,\theta) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}r^{n}[a_{n}\cos{(n\theta)} + b_{n}\sin{(n\theta)}]
\]
This solution is worth memorizing, as it will generally be the same for all variations of the Laplace equation similar to this one. Now we can apply the conditions for $\theta$. We have two conditions, but it is really just a piecewise function which we will call $f(\theta)$. We also know that our bound on the domain is $a$, or the length of the radius. This is enough information to find all the various variables in our $u(r,\theta)$. We have as follows:
\[
u(a,\theta) = f(\theta) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}a^{n}[a_{n}\cos{(n\theta)} + b_{n}\sin{(n\theta)}]
\]
Note that the $a$ which is the length of the radius is \textit{not} the same $a$ as $a_{n}$. Now we can solve for $a_{0}$ and $a_{n}$.
\begin{alignat*}{3}
a_{0} &= \frac{1}{\pi}\int_{0}^{2\pi}f(\theta)d\theta \qquad\qquad a_{n} &&= \frac{1}{\pi}\left[\int_{0}^{\pi}U_{1}\cos{(n\theta)}d\theta + \int_{\pi}^{2\pi}U_{2}\cos{(n\theta)}d\theta\right]\\
&= U_{1} + U_{2} &&=0
\end{alignat*}
For $b_{n}$, we have
\begin{align*}
a^{n}b_{n} &= \frac{1}{\pi}\left[\int_{0}^{\pi}U_{1}\sin{(n\theta)}d\theta + \int_{\pi}^{2\pi}U_{2}\sin{(n\theta)}d\theta\right]\\
&= \frac{1 - (-1)^{n}}{n\pi}(U_{1} - U_{2})\\
b_{n} &= \frac{\left[1 - (-1)^{n}\right](U_{1} - U_{2})}{n\pi a^{n}}\\
&= \frac{2(U_{1} - U_{2})}{(2k+1)\pi a^{2k+1}}
\end{align*}
Putting it all together gives us a final solution of
\[
u(r,\theta) = \frac{U_{1} + U_{2}}{2} + \sum_{k=0}^{\infty}r^{k}\frac{2(U_{1} - U_{2})}{(2k+1)\pi a^{2k+1}}
\]
And we are done.
\newpage

\indent Now let's look at another Laplace equation in polar coordinates. Again, we will be dealing with Dirichlet boundary conditions, so we will be able to jump straight into the solution. This one won't take us as long to solve, as we have already done a lot of the work. We are asked the following:\\\\
\noindent \textbf{\textit{Ex:}} Solve the following PDE
\[
\begin{cases*}
\nabla^{2}u = 0\\
u(3,\theta) =
\begin{cases*}
1 \qquad\qquad 0 \leq \theta \leq \pi\\
\sin^{2}{\theta}\qquad \pi \leq \theta \leq 2\pi
\end{cases*}
\end{cases*}
\]
\indent \textbf{\textit{Solution:}} We already know what our separable equation will look like:
\[
\frac{r^{2}R'' + rR'}{R} = -\frac{T''}{T} = -\lambda
\]
This yields an eigenvalue of $\lambda = n^{2}$. We know this because our function $T$ must be periodic on $2\pi$, and as such we get the following piecewise function:
\[
T(\theta) =
\begin{cases*}
a_{0}\qquad\qquad\qquad\qquad\qquad\quad\ \text{for}\ n=0\\
a_{n}\cos{(n\theta)} + b_{n}\sin{(n\theta)}\qquad \text{for}\ n > 0
\end{cases*}
\]
The solution to our equation for $R$ yields
\[
R(r) =
\begin{cases*}
c_{1} + c_{2}\ln{r} \qquad\qquad \text{for}\ n=0\\
c_{1}r^{n} + c_{2}r^{-n}\qquad\quad \text{for}\ n > 0
\end{cases*}
\]
We can disregard $c_{2}$ for this equation, since it would cause our solution to go to $\infty$ as $r \to 0$, at the center of the disk. Putting everything together gives us
\begin{align*}
u(r,\theta) &= R(r)T(\theta)\\
&=
\begin{cases*}
a_{0}\qquad\qquad\qquad\qquad\qquad\qquad\quad\ \text{for}\ n=0\\
a_{n}r^{n}\cos{(n\theta)} + b_{n}r^{n}\sin{(n\theta)}\qquad \text{for}\ n > 0
\end{cases*}
\end{align*}
We should probably construct a series out of this, since that is what we usually do. The corresponding series is
\[
u(r,\theta) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}a_{n}r^{n}\cos{(n\theta)} + b_{n}r^{n}\sin{(n\theta)}
\]
And applying the boundary conditions gives us
\[
u(3,\theta) = \frac{a_{0}}{2} + \sum_{n=1}^{\infty}a_{n}3^{n}\cos{(n\theta)} + b_{n}3^{n}\sin{(n\theta)}
\]
For $a_{0}$, we have
\begin{align*}
a_{0} &= \frac{1}{\pi}\left[\int_{0}^{\pi}1 d\theta + \int_{\pi}^{2\pi}sin^{2}{\theta}d\theta\right]\\
&= \frac{3}{2}
\end{align*}
Solving for $a_{n}$ gives us
\begin{align*}
a_{n} &= \frac{1}{3^{n}\pi}\left[\int_{0}^{\pi}\cos{(n\theta)}d\theta + \int_{\pi}^{2\pi}\cos{(n\theta)}\sin^{2}{\theta}d\theta\right]\\
&= \frac{[n^{2} - 2 - 4\cos{(n\pi)}]\sin{(n\pi)}}{3^{n}n(n^{2}-4)\pi}\\
&= 0
\end{align*}
Our $b_{n}$ term is
\begin{align*}
b_{n} &= \frac{1}{3^{n}\pi}\left[\int_{0}^{\pi}\sin{(n\theta)}d\theta + \int_{\pi}^{2\pi}\sin{(n\theta)}\sin^{2}{\theta}d\theta\right]\\
&= \frac{2\left[n^{2} - 6 - 4\cos{(n\pi)}\right]\sin^{2}{\left(\frac{n\pi}{2}\right)}}{3^{n}n\pi(n^{2} - 4)}\\
&= \frac{2\left[n^{2} - 6 - 4(-1)^{n}\right]\sin^{2}{\left(\frac{n\pi}{2}\right)}}{3^{n}n\pi(n^{2} - 4)}\\
&=
\begin{cases*}
0 \qquad\qquad\qquad\qquad\quad\ \ n \in \{ 2k\ |\ k \in \mathbb{Z} \}\\
\frac{1}{3^{n}\pi}\left[\frac{2}{n} + \frac{4}{n(n^{2} - 4)}\right]\qquad n \in \{ 2k + 1\ |\ k \in \mathbb{Z} \}
\end{cases*}
\end{align*}
We essentially have our solution. All we have to do is combine everything:
\[
u(r,\theta) = \frac{3}{4} + \sum_{k=0}^{\infty}\frac{r^{2k + 1}}{3^{2k+1}\pi}\frac{2(4k^{2} + 4k -1)}{(2k + 3)(4k^{2} -1)}\sin{\left[(2k +1)\theta\right]}
\]
Let's take a look at some of the plots for this equation.\\
\begin{center}
\begin{tabular}{c c c}
\includegraphics[scale=0.3]{lap_01_3} & \includegraphics[scale=0.3]{lap_01_30} & \includegraphics[scale=0.3]{lap_01_50}\\
\textit{The first 3 terms} & \textit{The first 30 terms} & \textit{The first 50 terms}
\end{tabular}
\end{center}
Plugging this infinite sum into mathematica tells us that it does not converge; the values are heading towards $\infty$ as the sum grows. We will leave our answer as a series, and we are done.

### 2016-01-02 proofs/pnn.tex
update several

\documentclass{article}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{minted}

\topmargin=-0.3in
\textheight=9.2in
\textwidth=168mm
\oddsidemargin=-0.2in
\evensidemargin=-0.2in

\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\QQ}{\mathbb{Q}}

\pagenumbering{gobble}

\title{3.0 Is Not An Integer}
\date{}
\begin{document}
\maketitle
\newpage

In the following document, we will show that 3.0 is not an integer, or $3.0 \notin \ZZ$. This result
stems from the fundametal idea, which we will prove, that $0.\overline{99} = 1.0$. The notation
$0.\overline{99}$ is a symbolic way to represent $0.99999...$ where the $9$ repeats \textbf{infinitely}.
It is clear that we can accept $0.\overline{99}$ as being an element of the real numbers $\RR$ and not an
element of the set of integers $\ZZ$ or rationals $\QQ$. We will now show that $0.\overline{99} = 1.0$.\\

\noindent\textbf{\textit{Theorem:}} $0.\overline{99} = 1.0$.\\\\
\indent\textbf{\textit{Proof:}} We can represent the number $0.\overline{99}$ as a geometric series
represented as
\begin{align*}
0.\overline{99} &= \frac{9}{10} + \frac{9}{100} + \frac{9}{1000} + \frac{9}{10000} + \cdots\\
&= \frac{9}{10} + \left(\frac{9}{10}\right)\left(\frac{1}{10}\right) +
\left(\frac{9}{10}\right)\left(\frac{1}{100}\right) +
\left(\frac{9}{10}\right)\left(\frac{1}{1000}\right) + \cdots\\
&= \frac{9}{10} + \left(\frac{9}{10}\right)\left(\frac{1}{10}\right)^{1} +
\left(\frac{9}{10}\right)\left(\frac{1}{10}\right)^{2} +
\left(\frac{9}{10}\right)\left(\frac{1}{10}\right)^{3} + \cdots\\
&= \sum_{n=1}^{\infty}\left(\frac{9}{10}\right)\left(\frac{1}{10}\right)^{n-1}.
\end{align*}
The general form for a geometric series is given by
\[
\sum_{n=1}^{\infty}ar^{n-1},
\]
Which leads us to assign the common ratio $r = \frac{1}{10}$ and $a = \frac{9}{10}$. By definition,
a geometric series converges to $\frac{a}{1 - r}$ if $|r| < 1$. This leads us to
\begin{align*}
0.\overline{99} &= \frac{\frac{9}{10}}{1 - \frac{1}{10}}\\
&= \frac{9}{10 -1}\\
&= \frac{9}{9}\\
&= 1.
\end{align*}
$\triangle$
\noindent\\

Now we can extend our argument to the original claim, that $3.0\notin\ZZ$.\\
\noindent\textbf{\textit{Theorem:}} $3.0\notin\ZZ$.\\\\
\indent\textbf{\textit{Proof:}} Assume that $3.0\in\ZZ$. Then $2.0 + 1 \in\ZZ$. We know as a fundamental
truth that if $x\in\RR$ and $y\notin\ZZ$, then $x + y\notin\ZZ$. We have already shown that
$1 = 0.\overline{99}$, so it follows that $2.0 + 1 = 2.0 + 0.\overline{99}$. $0.\overline{99}$ is not an
integer, and therefore $3.0 = 2.0 + 0.\overline{99}\notin\ZZ$.\\
$\triangle$
\newpage

Before we conclude, it should also be noted that from a practical standpoint, 3.0 cannot be considered
an integer. In this day and age, the study of programming is becoming a more and more necessary
skill for people to learn. Programming languages such as Java, C++, Ruby, and countless other
languages that are widely used would never assert that $3.0\in\ZZ$. For example, consider the following
Ruby code:
\begin{minted}{ruby}
def main
    if 3.0 == 3 then
        puts "3.0 = 3"
    else
        puts "3.0 != 3"
end
\end{minted}
This code will execute the else block of code. In the following Java code, the compiler would
automatically cast the value 3.0 to 3, since it was explicitly declared as an integer value:
\begin{minted}{java}
public class ThreePointOh {
    public static void main(String[] args){
        int tpo = 3.0;
        System.out.println(tpo); //Outputs 3, not 3.0
    }
}
\end{minted}
It is simply not practical to continue under the assumption that $3.0\in\ZZ$.
\end{document}

### 2016-01-02 resume/education.tex
update several

\noindent\textbf{\large Education}
\begin{enumerate}[label={\Roman*}.]
\item \textbf{Christopher Newport University $\cdot$ 2013 - Present}
\begin{itemize}
\item Pursuing a B.S. in Mathematics and a B.S. in Computer Science
\item Serving as Secretary for the Society of Physics club
\item Employed by the Center for Academic Success as a tutor
\item Earned Dean's List every semester
\item Relavent coursework includes
\begin{itemize}
\item UNIX Operating Systems, C++ Programming, Java Programming, Android Programming, Programming Language
Concepts, Data Structures, Algorithms, Computer Hardware and Design, Discreet Mathematics, Linear Algebra,
Calculus I - IV, Real Analysis, Partial and Ordinary Differential Equations, Writing in Mathematics using
\hologo{LaTeX}
\end{itemize}
\item Expected to graduate in May 2017
\end{itemize}
\end{enumerate}

### 2016-01-02 resume/work.tex
update several

\noindent\textbf{\large Work Experience}
\begin{enumerate}[label={\Roman*}.]
\item National Aeronautics and Space Administration $\Bigcdot$ November 2015 - Present
\begin{itemize}
\item Developed Android applications for Google Glass, Vuzix, and Recon Jet headsets
\item Used git extensively for version control
\item Wrote code primarily in Java
\end{itemize}
\item Freelance Application Developer $\Bigcdot$ July 2015 - Present
\begin{itemize}
\item Built a location based audio application called Audionce, written for Android and iOS.
\item Created a quiz web-application called WhensIt for the Public History Center at Christopher Newport
University
\item Built websites using JavaScript, HTML5, CSS3, and PHP
\end{itemize}
\item Christopher Newport University $\Bigcdot$ September 2013 - Present
\begin{itemize}
\item IT Customer Service Representative
\begin{itemize}
\item Aided customers with various software related IT issues
\item Assisted technicians with tasks such as installing hardware
\end{itemize}
\item Tutor for the Center for Academic Success
\begin{itemize}
\item Tutored several subjects including Mathematics, Computer Engineering, and Computer Science
\item Participated in both one on one tutoring and group tutoring
\end{itemize}
\end{itemize}
\item Kirk Family YMCA $\Bigcdot$ August 2009 - August 2015
\begin{itemize}
\item Lead Swim Instructor
\begin{itemize}
\item Managed a team of swim instructors that taught and oversaw various camps and lessons
\item Taught swimming lessons to children and adults of all ages
\item Promoted to manager in May 2015
\end{itemize}
\item Lifeguard
\begin{itemize}
\item Attended trainings for CPR and AED certifications, as well as First Aid and lifesaving skills
\item Worked to create and maintain a safe environment for all patrons of the pool by enforcing rules and
policies of the YMCA
\end{itemize}
\end{itemize}
\end{enumerate}

### 2016-01-02 writing_in_mathematics/paper.tex
update several

--- 2015-11-13 wim/paper.tex
@@
 \indent Processing data efficiently has always been a goal of organizations that collect data; companies like Facebook and Google have been collecting and storing data for as long as they have existed.
-One important statistic that companies can use is the number of distinct elements in a large set of data.
-The count-distinct problem poses the question of how many distinct elements are in a given set. 
-There are several algorithms that exist that can solve the count distinct problem; we will consider three of them:
+One important statistic that is useful to companies is the number of distinct elements in a large set of data.
+The question of how many distinct elements there are in a given set is called the \textit{count distinct} problem. 
+Several algorithms exist that can solve the count distinct problem; we will consider three of them:
 the hash table, the Linear Probabilistic Counter, and the HyperLogLog. 
-In order to analyze these algorithms, we will apply them to real world problems, namely (1) The Pok\'emon Problem, which poses the question ``How many unique Pok\'emon encounters will a player face in a given play through of a game from each generation?,'' (2) The Facebook Problem, which asks ``How many unique Facebook app installs are made on any given day?'', and (3) The Twitter Problem, where we will determine how many unique hashtags are created on any given day.
-We can use the hash table, the Linear Probabilistic Counter, and the HyperLogLog to solve each problem, respectively.\\
+To analyze these algorithms, we will apply them to real world problems, namely (1) The Pok\'emon Problem, which poses the question ``How many unique Pok\'emon encounters will a player face in a given play through of a game from each generation?'' (2) The Facebook Problem, which asks ``How many unique Facebook app installs are made on any given day?'' and (3) The Twitter Problem, which asks ``how many unique hashtags are created on any given day?''
+We use the hash table, the Linear Probabilistic Counter, and the HyperLogLog to solve each problem, respectively.\\
 
 \indent The first solution to the count distinct problem is called a hash table. 
-A hash-table is a matrix that is stored in memory where each entry was uniquely obtained from another value, called a key. 
+A hash table is a matrix that is stored in memory where each entry is uniquely obtained from another value, called a key. 
 The algorithm is as follows:
@@
 \end{center}
-Since $\mathbb{V}$ is a bijection, it is surjective. 
-As a result, if we take any $x\in\mathbb{S}\ \text{and}\ y\in\mathbb{S}$, then if $x = y$, they will hash to the same value, and we can ignore it \cite[p. 6]{Maurer}. 
-For example, Say we want to count the number of distinct names in the set $\mathbb{S} = \{\text{``Alice''},\text{``Emily''},\text{``Alice''}\}$. 
-We can create a simple hash function where we map each letter to its corresponding value in the alphabet, sum up total of all the letter values in a name, and mod that value with the number of letters in the name.
-Applying the function on each element of $\mathbb{S}$ yields a set $\mathbb{V} = \{0,4\}$.
-$||\mathbb{V}|| = 2$, so there were two distinct elements in $\mathbb{S}$.
+Since $h$ is a bijection, it is injective. 
+As a result, if we take any $x\in\mathbb{S}\ \text{and}\ y\in\mathbb{S}$, then if $x = y$, they will hash to the same value; since we want to know how many distinct values exist in $\mathbb{V}$, we ignore the duplicate values produced by the hash function \cite[p.
+6]{Maurer}.
+For example, say we want to count the number of distinct names in the set $\mathbb{S} = \{\text{``Alice''},\text{``Emily''},\text{``Alice''}\}$. 
+We create a simple hash function that is bijective for the set $\mathbb{S}$.
+Our function will first translate each letter in a word into its corresponding numerical position in the alphabet.
+We then sum up the total of each of the numbers (call this $n$), and determine the remainder of division of $n$ and the total number of letter in the name.
+Applying our hash function on the first element, ``Alice'', yields $1 + 12 + 9 + 3 + 5 = 30$, which when divided by 5 produces a remainder of 0; therefore ``Alice'' hashes to 0.
+When the hash function is passed all the elements of $\mathbb{S}$, we get a set $\mathbb{V} = \{0,4\}$. 
+The cardinality, or size, of $\mathbb{V}$ (denoted as $||\mathbb{V}||$) is 2, so there were two distinct elements in $\mathbb{S}$. 
 In this example, the hash function we chose was bijective.
-However, this may not always be the case. In fact, it is practically impossible to create a hash function that is truly surjective for huge amounts of data \cite[p. 6]{Maurer}. 
+However, this may not always be the case. In fact, it is practically impossible to create a hash function that is truly injective for huge amounts of data \cite[p. 6]{Maurer}. 
 For smaller sets of data, however, we can create a hash function that is surjective. 
-The pok\'emon problem provides a good example of the benefits of a hash table.\\
+The Pok\'emon  problem provides a good example of the benefits of a hash table.\\
 
 \indent The most beneficial aspect of the hash is that it has a 0\% error rate. 
-In other words, the cardinality of $\mathbb{V}$ will be exactly how many distinct elements are in $\mathbb{S}$, assuming the hash function is surjective. 
+In other words, the cardinality of $\mathbb{V}$ is exactly the number of distinct elements are in $\mathbb{S}$, assuming the hash function is bijective. 
 We know that for very large $\mathbb{S}$, the hash table method is incredibly inefficient, but for the Pok\'emon problem, we are dealing with $\mathbb{S}$ small enough to apply the hash table. 
-In this particular case, 721 pok\'emon can be encountered \cite{Pokemon}, which means that the maximum size needed for our set $\mathbb{V}$ is 721. 
-Typically, a player will encounter in the neighborhood of 1000 wild pok\'emon a game, and with a total of 6 generations of games, we have a set $\mathbb{S}$ of size 6000. 
-The hashing function we created earlier will not be good enough here; instead we can create a new hashing function where we sum up the value of the name as before, then add the pok\'emon's unique id, and finally mod that value with 721.
+In this particular case, 721 Pok\'emon  can be encountered \cite{Pokemon}, which means that the maximum size needed for our set $\mathbb{V}$ is 721. 
+Typically, a player will encounter in the neighborhood of 1000 wild Pok\'emon a game, and with a total of 6 generations of games, we have a set $\mathbb{S}$ of size 6000. 
+The hashing function we created earlier will not be good enough here; instead we can create a new hashing function where we sum up the value of the name as before, then add the Pok\'emon 's unique id, and finally mod that value with 721.
 \begin{floatingfigure}[H!]{8cm}
@@
 \end{floatingfigure}
-\noindent Running the code in \textit{Figure 1} will give us a final answer of 461 unique encounters with wild pok\'emon. 
-In the case of the pok\'emon problem, a simple hash function sufficed, since the only collisions that were encountered were pok\'emon where the name was the same. 
-If we consider a larger data set, however, we may not be able to create a surjective hash function.\\
+\noindent In order to come up with a set $\mathbb{S}$, we can randomly populate an array with 6000 Pok\'emon; we are assuming that a player has an equal chance to run into any Pok\'emon, and that they can encounter all Pok\'emon more than once.
+Running the code in \textit{Figure 1} gives us a final answer of 461 unique encounters with wild Pok\'emon. 
+In the case of the Pok\'emon problem, a simple hash function suffices, since the only collisions encountered are Pok\'emon with the same name; we realistically can test the hash function by passing it all 721 Pok\'emon and determining whether or not there are 721 hashed values.
+If we consider a larger data set, however, we may not be able to create a bijective hash function.\\
 
-\indent Imagine that we are solving the Twitter problem where we want to count the number of distinct hashtags on twitter on a given day. 
+\indent Imagine that we are solving the Twitter problem where we want to count the number of distinct hashtags on Twitter on a given day. 
 Our set $\mathbb{S}$ will now contain every hashtag that was made on a given day. 
-We know that we want a hash function that is surjective, but this is an unrealistic, since we could have some $x \in \mathbb{S}$ and $y\in\mathbb{S}$ where $x \neq y$ but $h(x) = h(y)$. 
+We know that we want a hash function that is bijective, but this is an unrealistic, since we could have some $x \in \mathbb{S}$ and $y\in\mathbb{S}$ where $x \neq y$ but $h(x) = h(y)$. 
 To model this issue, we can consider the following scenario:
@@
 As a result, ``\#cape'' began trending on Twitter.
-On the same day, Target announced that all their \textit{Bob's Burgers} costumes would be on sale for the low price of \$12.95, and ``\#bob'' began trending on twitter.
-\end{quote}  \clearpage
-Our hash function from earlier will not work in this case, since both ``\#bob'' and ``\#cape'' hash to the same value. 
-On a larger scale, such as the Facebook or Twitter problem, we would need to either handle collisions or create a better hash function. 
-Creating a way to handle collisions would not be beneficial, since the data is being processed in real time. 
-We would not be able to tell if we were hashing a value we already have seen or hashing a new value that happens to hash to the same value as another element we have already seen; as a result we would have no way to know which keys were different and which keys were the same when looking at $\mathbb{V}$ \cite[p.
-6]{Maurer}.
+On the same day, Target announced that all their \textit{Bob's Burgers} costumes would be on sale for the low price of \$12.95, and ``\#bob'' began trending on Twitter.
+\end{quote}
+Our previous hash function will not work in this case, since both ``\#bob'' and ``\#cape'' hash to the same value. 
+On a larger scale, such as the Facebook or Twitter problem, we would need to either handle collisions (where two different keys hash to the same value) or create a better hash function. 
+Creating a way to handle collisions would not be beneficial, since the data are being processed in real time. 
+We would not be able to tell if we were hashing a value we already saw or hashing a new value that happened to hash to the same value as another element we already saw. 
+As a result, we would have no way to know which keys were different and which keys were the same when looking at $\mathbb{V}$ \cite[p. 6]{Maurer}. 
 Since the collision policy would not help us, we need to come up with a better hashing function.\\
 
-\indent More complex hash functions, such as SHA-2 (Secure Hash Algorithm 2) or MD5 (Message Digest 5) offer a solution to the collision problem.
-SHA-2 is used by organizations such as the NSA to encrypt data, and has yet to produce any collisions when hashing values \cite[p.
+\indent More complex hash functions, such as SHA-2 (Secure Hash Algorithm 2) or MD5 (Message Digest 5), offer a solution to the collision problem.
+SHA-2 is used by organizations such as the NSA to encrypt data and has yet to produce any collisions when hashing values \cite[p.
 301]{SHA-Algo}.
-However, this is not a viable solution to the count distinct problem either due to the volume of the data being processed.
+However, this is not a viable solution to the count distinct problem either to the volume of the data being processed.
 In both the Twitter and Facebook problems, we are analyzing exabytes of data.
-The SHA-2 function outputs a 512 byte long integer for every input, which means that the resulting size of the set $\mathbb{V}$ would be in the worst case $512 \times ||\mathbb{S}||$.
-All of $\mathbb{V}$ would have to be held in memory, which means that analyzing 50 million hashtags (which is a low estimate)\cite{Twitter} could possibly result in $(50 \times 10^{10}) \times 512$ bytes of data being stored at one time, assuming every hashtag was 10 bytes long.
+The SHA-2 function outputs a 512-byte-long integer for every input, which means that the resulting size of the set $\mathbb{V}$ would be in the worst case $512 \times ||\mathbb{S}||$.
+All of $\mathbb{V}$ would have to be held in memory, which means that analyzing 50 million hashtags (a low estimate)\cite{Twitter} could possibly result in $(50 \times 10^{10}) \times 512$ bytes of data being stored at one time, assuming every hashtag was 10 bytes long.
 A hash table of this size could not be loaded all at once into main memory; it would have to be stored into secondary memory which would exponentially increase the amount of time required to read the data \cite[p.
 209]{Whang}.
-MD5 outputs smaller sized integers (128 in this case), but the collision probability is much higher \cite[p.
+MD5 outputs smaller sized integers (128 bytes long), but the collision probability is much higher \cite[p.
 22 - 23]{Break-MD5}.
-The hash method is not a viable solution to the count distinct problem when we are dealing with a large volume of data, but hash functions are still used in the other solutions to the count-distinct problem.\\
+The hash method is not a viable solution to the count distinct problem when we are dealing with a large volume of data.\\
 
-\indent The hash-table method provided a 0\% error, but was not efficient in handling large sets of data.
+\indent The hash table method provides a 0\% error, but is not efficient in handling large sets of data. 
 However, if we are looking for a \textit{good enough} estimate (within the neighborhood of 1\% to 5\% error) of the number of distinct elements in $\mathbb{S}$, we can use an algorithm called a Linear Probabilistic Counter.
-With the Linear Probabilistic Counter, the user can specify how large they want the error to be.
+With the Linear Probabilistic Counter, users can specify how large they want the error to be.
 This method uses a data structure called a bitmap. 
-bitmaps are a matrix where either a 1 or 0 is stored in a cell, which allows for the map to be incredibly space efficient.
-The Linear Probabilistic Counter also uses a hash, but it only needs to temporarily produce a position and then it can ``forget'' the value it just produced, which in turn reduces the overhead memory needed to store the elements that was required by the hash table method \cite{Whang}.
+Bitmaps are matrices with either a 1 or 0 stored in each cell, which allows for the map to be extremely space efficient. 
+The Linear Probabilistic Counter also uses a hash, but it only needs to temporarily produce a position and then it can ``forget'' the value it just produced, which in turn reduces the overhead memory needed to store the elements required by the hash table method \cite{Whang}.
 The number of distinct elements is determined by the number of 1's in the bitmap when the algorithm is finished.\\
+
 \indent The Linear Probabilistic Counter algorithm is divided into several steps. 
 In the first step, the algorithm creates an arbitrarily sized bitmap in memory, and then hashes the values into the bitmap by determining the address in the bitmap at which the corresponding bit should be changed to a 1.
-In the next step, the algorithm counts the number of 0's in the bitmap, and stores the result.
+In the next step, the algorithm counts the number of 0's in the bitmap and stores the result.
 Finally, the algorithm estimates the cardinality of the unique set by dividing the count of 0's by the size of the bitmap, and then substituting in the result to
@@
 \indent We can use the Linear Probabilistic Counter algorithm to solve the Facebook problem; unfortunately, we have no way of accessing the actual data we need, so we will have to do some guessing. 
-We know that Facebook users make about 7,000,000 app installs a dat \cite{Facebook-2}.
+We know that Facebook users make about 7,000,000 app installs a day \cite{Facebook-2}.
 This means that the size of the set $\mathbb{S}$ will be 7,000,000.
@@
 In other words, we expect a lot of distinct elements in $\mathbb{S}$.\\
-\indent We can model the exponential function by choosing random numbers with a certain weight towards different numbers.
+\indent We can model the exponential function by choosing pseudo-random numbers weighted by a pre defined function.
 We will let the each number correspond to a different application.
@@
 We can use an algorithm (courtesy of GitHub user Danieth5x) that will generate a pseudo-random sample based on a function, which will be an exponential function in our case.
-We can argue that it is not important as to which exponential function we choose, since the distribution will be more or less the same; in our case we will pick $e^{x}$.
+We can argue that it is not important which exponential function we choose, since the distribution will be more or less the same; in our case we pick $e^{x}$.
 Running the pseudo-random generator produces a set $\mathbb{S}$ of size 7,000,000, and we can now apply the Linear Probabilistic Counter to the data set.
-For our purposes, we will use the hash function supplied by Ruby; the only change we need to make is to mod the result of the Ruby hash with the number of spaces in the bitmap, which ensures that we will never produce a  hashed value out of the bounds of our bitmap.
+For our purposes, we will use the hash function supplied by Ruby; the only change we need to make is to mod the result of the Ruby hash with the number of spaces in the bitmap, which ensures that we never produce a  hashed value out of the bounds of our bitmap.
 Finally, running the algorithm tells us that the total number of unique application installs is 3216.\\
@@
 \indent The Linear Probabilistic Counter and the hash table algorithms handled relatively small data sets well, but the Twitter problem deals with a data set of nearly 200,000,000 tweets a day \cite{Twitter-2}.
-As a result, the Linear Probabilistic Counter would be too inefficient to process all the data, but the HyperLogLog algorithm allows for data to be processed more efficiently than the Linear Probabilistic Counter or the hash table.
+The Linear Probabilistic Counter is too inefficient to process all the data, but the HyperLogLog algorithm allows for data to be processed more efficiently than the Linear Probabilistic Counter or the hash table.
 The HyperLogLog algorithm actually uses parts of the hash table and Linear Probabilistic Counter algorithms; values are still hashed to a bitmap held in memory. 
-The hash function hashes values to binary numbers, and the algorithm decides whether or not that value has already been seen the value before by analyzing the leading 0's in the binary number \cite[pp.
-685, 689]{Heule}.
+The hash function hashes values to binary numbers, and the algorithm decides whether that value has already been seen by analyzing the leading 0's in the binary number \cite[pp. 685, 689]{Heule}.
 In other words, the algorithm keeps a record of the first location of a 1 in a binary string, called $\rho$ \cite[pp. 130]{Flaj}.
 If the sequence of 0's is identical to another sequence of 0's, the probability that the original values were the same is very high, and the algorithm will ignore what it finds to be a duplicate value.
-The amount of memory that the HyperLogLog algorithm requires per bitmap is expressed by
+The amount of memory that the HyperLogLog algorithm requires per bitmap is
 \[
@@
 where $M$ is the size of the original set of data \cite[pp. 129]{Flaj}.
-The error amount can be found by
+The error amount is
 \[
@@
 which allows the algorithm to increase in accuracy as the size of the original set grows \cite{Yousra}.\\
+
 \indent We can solve the Twitter problem with the HyperLogLog algorithm.
-We will use the code in \textit{Figure 3} to aggregate a random sampling of the most recent 2000 tweets that contain a ``\#'' every two minutes for 24 hours.
-This will allow for a large random sample of tweets that will closely model the real problem that we are trying to solve.
+We use the code in \textit{Figure 3} to aggregate a random sampling of the most recent 2000 tweets that contain a ``\#'' every two minutes for 24 hours.
+This allows for a large random sample of tweets that will closely model the real problem that we are trying to solve.
 We also need a hashing function that will produce a binary string of 1's and 0's given any key; luckily this hash function already exists.
 A murmur hash will produce a 32 bit binary string, which is exactly what we need.
-Finally, we actually need to implement the algorithm. 
-We will create a class to perform the calculation that is initialized with the amount of memory we are willing to spend on the bitmap for the HyperLogLog algorithm. 
+Finally, we implement the algorithm. 
+We create a class to perform the calculation that is initialized with the amount of memory we are willing to spend on the bitmap for the HyperLogLog algorithm. 
 We know that the HyperLogLog can process nearly 1 billion elements with only 1.5 kilobytes of space to an accuracy of 2\% \cite{Flaj}.
-This is given by the equation we have for memory.
+%This estimate is given by the equation we have for memory.
 Using a similar argument, we can determine the amount of memory we will need.
@@
 \begin{framed}
-\includegraphics[scale=0.3]{twitter_problem/ruby_code_01}
+\includegraphics[scale=0.3]{Twitter_problem/ruby_code_01}
 \caption{An infinite while loop can be useful}
@@
 \end{floatingfigure}
-\noindent If we want the error to be 1\%, then we can calculate the number of spaces required in the bitmap by using the equation we have for the error.
+\noindent If we want the error to be 1\%, then we calculate the number of spaces required in the bitmap by using the equation we have for the error.
 \begin{align*}
 0.01 &= \frac{\sqrt{3\log{(2)} - 1}}{\sqrt{m}}\\
-0.01\,\sqrt{m} &= \sqrt{3\log{(2)} - 1}\\
-\sqrt{m}\, &= \frac{\sqrt{3\log{(2)} - 1}}{0.01}\\
 m &= \frac{\sqrt{3\log{(2)} - 1}}{0.01^{2}}\\
@@
 \end{align*}
-Now we can find the total memory required for our logarithm. If we let the program from \textit{Figure 1} run for 24 hours, at worst it will parse 1,440,000 hashtags.
-We can now calculate the memory required by
+Now we can find the total memory required for our logarithm. If we let the program from \textit{Figure 3} run for 24 hours, at worst it will parse 1,440,000 hashtags.
+We calculate the memory required by
 \begin{align*}
@@
 Multiplying this number by the total number of entires gives us a grand total of 47005 bytes, or 4.7005 kilobytes of required memory.\\
+
 \indent We now have all the necessary information to parse the data.
-We will implement the HyperLogLog in an object oriented manner; that is, we will create a Ruby class called HyperLogLog that we can initialize with the number of bytes per entry.
-We also need a method to calculate the unique elements, and we need a way to actually call this routine.
-\clearpage
+We implement the HyperLogLog in an object oriented manner; that is, we create a Ruby class called HyperLogLog that we initialize with the number of bytes per entry.
+We also need a method to calculate the unique elements, and we need a way to call this routine.
 \begin{center}
 \begin{tabular}{c c c}
-\includegraphics[scale=0.3]{twitter_problem/hll_init}
+\includegraphics[scale=0.3]{Twitter_problem/hll_init}
 &
-\includegraphics[scale=0.4]{twitter_problem/hll_count}
+\includegraphics[scale=0.4]{Twitter_problem/hll_count}
 &
-\includegraphics[scale=0.4]{twitter_problem/hll_imp}
+\includegraphics[scale=0.4]{Twitter_problem/hll_imp}
 \\
@@
 \noindent \\
-Note that we settled with 4 as the size of each entry in the bit-map. This value is close enough to 4.35 to safely say that the error will be just about 1\%.
-Finally, in our main program, we will loop infinitely (it will be up to us as to when to kill the process), and read in $\sim 2000$ tweets every two minutes. 
+Note that we settled with 4 as the size of each entry in the bit-map. This value is close enough to 4.35 to say that the error will be just about 1\%.
+Finally, in our main program, we loop infinitely (it will be up to us as to when to stop the program), and read in $\sim 2000$ tweets every two minutes. 
 After running the algorithm for 24 hours, we will have a set of data that represents the number of unique hashtags found (by using the HyperLogLog algorithm) as the number of tweets being analyzed increases.
-The data grows in a linear fashion; this is expected since we are gathering a random sample from all over the world, which means that at any given time, we expect to see more or less the same amount of unique hashtags.
+The data grow in a linear fashion; this is expected since we are gathering a random sample from all over the world, which means that at any given time, we expect to see about the same number of unique hashtags.
 Plotting the data and applying a best fit line using Mathematica gives us
@@
 \]
-From this data, we can make an estimate that at 200,000,000 tweets, we will have
+From this data, we can make an estimate that at 200,000,000 tweets, we have
 \begin{align*}
@@
 \end{align*}
-which means that we will have roughly 56 million unique hashtags by the HyperLogLog algorithm.\\ 
+which means that we have roughly 56 million unique hashtags by the HyperLogLog algorithm.\\
+
 \indent We have looked at three different approaches to the count distinct problem. 
 We used the HyperLogLog to solve only the Twitter problem, but we could have used the algorithm to solve all of the problems. 
-The count distinct problem is heavily applicable in the real world; companies are able to turn the numbers that are determined by algorithms such as the HyperLogLog to increase profits, target ads, and keep more data.
-Finding an efficient way to count large sets of data may never be competely accurate, but with algorithms like the HyperLogLog and the Linear Probabilistic counter, we can efficiently and with nearly 100\% accuracy solve the count distinct problem.
+The count distinct problem is widely applicable in the real world; companies are able analyze numbers that are determined by algorithms such as the HyperLogLog to increase profits, target ads, and keep more data.
+An efficient way to count large sets of data may never be completely accurate.
+With algorithms like the HyperLogLog and the Linear Probabilistic counter, we can efficiently and with nearly 100\% accuracy solve the count distinct problem.
 
@@
 \begin{thebibliography}{9}
-\bibitem{Twitter} \#numbers. (2011, March 14). Retrieved October 29, 2015, from \url{https://blog.twitter.com/2011/numbers}
+\bibitem{Twitter} \#numbers. (2011, March 14). Retrieved October 29, 2015, from \url{https://blog.Twitter.com/2011/numbers}
 \bibitem{SHA-Algo} Chaves, Ricardo, Kuzmanov, Georgi, Sousa, Leonel \& Vassiliadis, Stamatis (2006). Improving SHA-2 Hardware Implementations. , 4249, 298-310.
@@
 Retrieved November 3, 2015, from \url{http://www.statista.com/statistics/276371/most-popular-facebook-apps-by-monthly-active-users/}
-\bibitem{Twitter-2} Twitter Statistics. (2015, September 15). Retrieved November 3, 2015, from \url{http://www.statisticbrain.com/twitter-statistics/}
+\bibitem{Twitter-2} Twitter Statistics. (2015, September 15). Retrieved November 3, 2015, from \url{http://www.statisticbrain.com/Twitter-statistics/}
 \bibitem{Break-MD5} Wang, Xiaoyun \& Yu, Hongbo (2005). How to Break MD5 and Other Hash Functions. , 3494, 19-35.

### 2016-01-02 writing_in_mathematics/presentation.tex
update several

--- 2015-11-21 wim/presentation.tex
@@
 \mode<presentation>
-\usetheme{Luebeck}
+\usetheme{default}
 
@@
 \pause
-\item Image we have a set called $\mathbb{V}$ that contains a billion elements of the same type.
+\item Imagine we have a set called $\mathbb{S}$ that contains a billion elements of the same type.
 \pause
-\item How many \textit{unique} elements are in $\mathbb{V}$?
+\item How many \textit{unique} elements are in $\mathbb{S}$?
 \end{itemize}
@@
 \begin{frame}{Solving the Pok\'emon Problem}
+\vspace{-2cm}
 \begin{itemize}
@@
 \item Find $m\ mod\ 721$.
-\pause
-\item Results:
-\begin{itemize}
-\pause
-\item On average: 461 unique encounters
-\end{itemize}
 \end{enumerate}
@@
 \begin{frame}[fragile]{Implementation}
+\vspace{-2cm}
 \begin{lstlisting}[basicstyle=\small]
@@
 \end{lstlisting}
+\pause
+On average: 461 unique encounters
 \end{frame}
@@
 \section{The HyperLogLog}
-\begin{frame}{The Algorithm}
+\begin{frame}{The HyperLogLog Algorithm}
 \begin{enumerate}
@@
 \end{align*}
-\pause
+\end{itemize}
+\end{frame}
+
+\begin{frame}{The Math}
+\begin{itemize}
 \item What is the predicted error?
@@
 \begin{frame}{Results}
+\vspace{-2cm}
 \begin{figure}
-\includegraphics[scale=0.7]{twitter_problem/resultspng}
+\includegraphics[scale=0.5]{twitter_problem/resultspng}
 \caption{$0.284356 x + 7361.39$}

### 2016-01-02 writing_in_mathematics/ignore/dum.rb

@@ -1,0 +1,31 @@
+def initialize(log2m, register_set = nil)
+    @log2m = log2m
+    @count = 2 ** log2m
+    @register_set = register_set || RegisterSet.new(@count)
+
+    case log2m
+    when 4
+        @alphaMM = 0.673 * @count * @count
+    when 5
+        @alphaMM = 0.697 * @count * @count
+    when 6
+        @alphaMM = 0.709 * @count * @count
+    else
+        @alphaMM = (0.7213 / (1 + 1.079 / @count)) * @count * @count
+    end
+end
+
+def cardinality
+    register_sum = 0.0
+    zeros = 0.0
+    @register_set.each do |value|
+        register_sum += 1.0 / (1 << value)
+        zeros += 1 if value == 0
+    end
+    estimate = @alphaMM * (1 / register_sum)
+    if estimate <= (5.0 / 2.0) * @count
+        (@count * Math.log(@count / zeros)).round
+    else
+        estimate.round
+    end
+end

### 2016-01-02 writing_in_mathematics/twitter_problem/const.rb

@@ -1,0 +1,22 @@
+class Constants
+	@@oh_one = "JtwwG40EW2O6CKBNp1qeD57IH"
+	@@oh_two = "NUSZ0jPyeWwTKPAlUDYLnoeatyDCsPp5waIJsYnQM6LJ5oDjDJ"
+	@@oh_three = "3494713935-hoqlE0ocEarpGUbSosIEi3YJvHqrJH3j60x7Z74"
+	@@oh_four = "fKTi6UiCESwluB9UbBOsdcBVjHHPfUOxr6PbSWy4UCUi8"
+
+	def self.oh_one
+		@@oh_one
+	end
+
+	def self.oh_two
+		@@oh_two
+	end
+
+	def self.oh_three
+		@@oh_three
+	end
+
+	def self.oh_four
+		@@oh_four
+	end
+end

### 2016-01-13 vector_calc/homework/hw1.tex
starting off the semester right

\documentclass{hw}

%page 7: 7,9,10,14,16,17,21,23

\begin{document}
\makeheader{1}
\begin{enumerate}
\item
Let $A$ be the point with coordinates $(1, 0, 2)$, let $B$ be the point with coordinates $(−3, 3, 1)$, and
let $C$ be the point with coordinates $(2, 1, 5)$.
\begin{enumerate}
\item Describe the vectors $\vv{AB}$ and $\vv{BA}$.
\begin{quote}
$\vv{AB}$ will be the vector from $(1,0,2)$ to $(−3, 3, 1)$, which is the same as the vector $(-4,3,-1)$.
\end{quote}
\item Describe the vectors $\vv{AC},\vv{BC}$, and $\vv{AC} + \vv{CB}$.
\begin{quote}
$\vv{AC}$ will be the vector from $(1,0,2)$ to $(2,1,5)$, which is the same as the vector $(1,1,3)$.\\
$\vv{BC}$ will be the vector from $(-3,3,1)$ to $(2,1,5)$, which is the same as the vector $(5,-2,4)$.\\
$\vv{CB} = (1,2,-4)$, so $\vv{AC} + \vv{CB} = (2,3,-1)$.
\end{quote}
\item Explain with pictures why $\vv{AC} + \vv{CB} = \vv{AB}$.
\vspace{3cm}
\end{enumerate}
\item If $(-12,9,z)+(x,7,-3) = (2,y,5)$, what are $x,y$, and $z$?
\begin{alignat*}{4}
& -12 + x = 2\qquad\qquad && 9 + 7 = y\qquad\qquad &&& z-3= 5\\
& x = 14 && y = 16 &&& z = 8
\end{alignat*}
\item What is the length of the vector $<3,1>$?
\begin{quote}
The length of the vector is $\sqrt{3^{2} + 1^{2}}$, or $\sqrt{10}$.
\end{quote}
\item Find the displacement vectors from $P_{1}$ to $P_{2}$, and sketch $P_{1}, P_{2}$, and
$\vv{P_{1}P_{2}}$.\\\\
\begin{minipage}{0.25\textwidth}
(a) $P_{1}(1,0,2)$, $P_{2}(2,1,7)$
$$\vv{P_{1}P_{2}} = (1,1,5)$$
\end{minipage}
\begin{minipage}{0.25\textwidth}
(b) $P_{1}(1,6,−1),P_{2}(0,4,2)$
$$\vv{P_{1}P_{2}} = (-1,-2,3)$$
\end{minipage}
\begin{minipage}{0.25\textwidth}
(c) $P_{1} (0,4,2), P_{2} (1,6,−1)$
$$\vv{P_{1}P_{2}} = (1,2,-3)$$
\end{minipage}
\begin{minipage}{0.25\textwidth}
(d) $P_{1}(3,1), P_{2}(2,−1)$
$$\vv{P_{1}P_{2}} = (-1,-2)$$
\end{minipage}

\vspace{3cm}
\item If $A$ is the point in $\RR^{3}$ with coordinates $(2,5,-6)$ and the displacement
vector from $A$ to $B$ is $(12,-3,7)$, what are the coordinates of $B$?
\begin{quote}
The vector $\vv{AB} = (12,-3,7) = (x_{B} - 2, y_{B} - 5, z_{B} + 6)$. Setting up a system of
equations yields
\begin{alignat*}{4}
& 12 = x_{B} - 2\qquad\qquad && -3 = y_{B} - 5 \qquad\qquad &&& 7 = z_{B} + 6\\
& x_{B} = 14 && y_{B} = 2 &&& z_{B} = 1
\end{alignat*}
Therefore $B = (14,2,1)$.
\end{quote}
\newpage
\item Suppose that you and your friend are in New York talking on cellular phones. You inform each other
of your own displacement vectors from the Empire State Building to your current position. Explain how
you can use this information to determine the displacement vector from you to your friend.
\begin{quote}
The Empire State Building acts as the origin. If you know your point ($A$) and your friend's point, ($B$),
you can calculate the displacement vector $\vv{AB}$, which will provide you with the displacement vector
from you to your friend by $\vv{AB} = (x_{b} - x_{a}, y_{b} - y_{a})$.
\end{quote}
\item
\begin{enumerate}
\item Let $\vv{a} = (2,0)$ and $\vv{b} = (1,1)$. For $0\leq s\leq 1$ and $0\leq t\leq 1$, consider the
vector $\vv{x} = s\vv{a} + t\vv{b}$. Explain why the vector $\vv{x}$ lies in the parallelogram
determined by $\vv{a}$ and $\vv{b}$.
\begin{quote}
For $\vv{x}$ to be in the parallelogram $\vv{a}\vv{b}$, $\vv{x}$ cannot be greater than $(3,1)$ and
cannot be less than $(0,0)$. $\vv{x}$ is defined as
\begin{align*}
\vv{x} &= s\vv{a} + t\vv{b}\\
&= (2s,0) + (t,t)\\
&= (2s + t, t),
\end{align*}
which at $s = 0$ and $t = 0$ still lies within the bound, and at $s = 1$ and $t=1$ still lies within
the bound.
\end{quote}
\item Now suppose that $\vv{a} = (2,2,1)$ and $\vv{b} = (0,3,2)$. Describe the set of vectors
$\{\vv{x} = s\vv{a} + t\vv{b} | 0\leq s\leq 1, 0\leq t\leq 1\}$.
\begin{quote}
The set of vectors will be in the form of
\[
(2s, 2s+3t, s+2t)
\]
\end{quote}
\end{enumerate}
\item A flea falls onto marked graph paper at the point $(3,2)$. She begins moving from that point with
velocity vector $\vv{v} = (−1, −2)$ (i.e., she moves 1 graph paper unit per minute in the negative
x-direction and 2 graph paper units per minute in the negative y-direction).
\begin{enumerate}
\item What is the speed of the flea?
\begin{align*}
\text{speed} &= \sqrt{1 + 4}\\
&= \sqrt{5}
\end{align*}
\item Where is the flea after 3 minutes?
\begin{quote}
The flea's position can be expressed as $f(t) = (3-t,2-2t)$. Plugging in 3 for $t$ yields $(0,4)$.
\end{quote}
\item How long does it take the flea to get to the point $(-4,-12)$?
\begin{quote}
We have the equation $f(t) = (3-t,2-2t)$. We setting it equal to $(-4,-12)$ and solving for $t$ yields
$t = 7$.
\end{quote}
\item Does the flea reach the point $(−13, −27)$? Why or why not?
\begin{quote}
Using the equation above, we can solve again for $t$. Doing so yields $t=16$ and $t=14.5$, which is
impossible. Therefore the flea does not reach that point.
\end{quote}
\end{enumerate}
\end{enumerate}
\end{document}

### 2016-01-13 vector_calc/homework/hw2.tex
starting off the semester right

\documentclass{hw}

% 11,13,17,19,23,25,27,31,35,37,42

\begin{document}
\makeheader{2}

\begin{enumerate}
\item Let $\vv{a_{1}} = (1,1)$ and $\vv{a_{2}} = (1,-1)$.
\begin{enumerate}
\item Write the vector $\vv{b_{1}} = (3, 1)$ as $c_{1}a_{1} + c_{2}a_{2}$, where
$c_{1}$ and $c_{2}$ are appropriate scalars.
\begin{align*}
(3,1) = (c_{1} + c_{2}, c_{1} - c_{2})\\
\left[
\begin{array}{c c | c}
1 & 1 & 3\\
1 & -1 & 1
\end{array}
\right] \to
\left[
\begin{array}{c c | c}
1 & 0 & 2\\
0 & 1 & 1
\end{array}
\right]\\
\vv{b} = 2(1,1) + 1(1,-1)
\end{align*}

\item Do the same for $(3,-5)$.
\begin{align*}
(3,1) = (c_{1} + c_{2}, c_{1} - c_{2})\\
\left[
\begin{array}{c c | c}
1 & 1 & 3\\
1 & -1 & -5
\end{array}
\right] \to
\left[
\begin{array}{c c | c}
1 & 0 & -1\\
0 & 1 & 4
\end{array}
\right]\\
\vv{b} = -1(1,1) + 4(1,-1)
\end{align*}

\item Show that any vector $\vv{b} = (b1, b2)$ in $\RR^{2}$ may be written in the form $c_{1} a_{1} +
c_{2} a_{2}$ for appropriate choices of the scalars $c_{1}, c_{2}$.
\begin{quote}
Assume that we can express any vector $(x_{1},x_{2})$ as $(c_{1} + c_{2}) + (c_{1} - c_{2})$, where
$x_{1},x_{2}\in\RR$. Then the solution for the corresponding matrix is
\[
\left[
\begin{array}{c c | c}
1 & 0 & \frac{x_{1} + x_{2}}{2}\\
0 & 1 & \frac{x_{1} - x_{2}}{2}
\end{array}
\right].
\]
Since $\frac{x_{1} + x_{2}}{2}$ and $\frac{x_{1} - x_{2}}{2}$ are both in $\RR$, any vector of the form
$(x_{1},x_{2})$ can be written as a linear combination of $\vv{a_{1}}$ and $\vv{a_{2}}$.
\end{quote}
\end{enumerate}

\item Write the following as a set of parametric equations: The line in $\RR^{3}$ through the point $(2,
−1, 5)$ that is parallel to the vector $i + 3j - 6k$.
\[
\begin{cases}
x = t+2\\
y = 3t - 1\\
z = 5 - 6t
\end{cases}
\]
\item Write the following as a set of parametric equations: The line in $\RR^{3}$ through the points
$(1,4,5)$ and $(2,4,1)$.
\[
\begin{cases}
x = t + 1\\
y = 4\\
z = 5 - 6t
\end{cases}
\]
\item Write the following as a set of parametric equations: Write a set of parametric equations for the
line in $\RR^{4}$ through the point $(1, 2, 0, 4)$ and parallel to the vector $(−2, 5, 3, 7)$.
\[
\begin{cases}
x_{1} =1-2t\\
x_{2} =5t+2\\
x_{3} =3t\\
x_{4} =7t+4
\end{cases}
\]

\item Give a symmetric form for the line having parametric equations $x = t + 7$, $y = 3t - 9$, $z = 6 -
8t$.
\[
\begin{cases}
x = t + 7\\
y = 3t - 9\\
z = 6-8t
\end{cases}\to
\begin{cases}
t = x - 7\\
t = \frac{y}{3} + 3\\
t = \frac{6 - z}{8}
\end{cases}
\]

\item Give a set of parametric equations for the line with symmetric form
\[
\frac{x + 5}{3} = \frac{y - 1}{7} = \frac{z + 10}{-2}.
\]
\[
\begin{cases}
t = \frac{x + 5}{3}\\
t = \frac{y - 1}{7}\\
t = \frac{z + 10}{-2}
\end{cases}\to
\begin{cases}
x = 3t - 5\\
y = 7t + 1\\
z = -2t -10
\end{cases}
\]

\item Show that the two sets of equations
\[
\frac{x-2}{3} = \frac{y-1}{7} = \frac{z}{5} \text{ and } \frac{x+1}{-6} = \frac{y+6}{-14} =
\frac{z+5}{10}
\]
represent the same line in $\RR^{3}$.
\begin{quote}
If we move to parametric coordinates, we have
\[
\begin{cases}
x = 3t + 2\\
y = 7t + 1\\
z = 5t
\end{cases}
\text{ and }
\begin{cases}
x = -6t - 1\\
y = -14t - 6\\
z = -10t - 5
\end{cases}.
\]
If we take the second vector and scale it by $-\frac{1}{2}$, we get
\[
\begin{cases}
x = 3t + \frac{1}{2}\\
y = 7t + 3\\
z = 5t = \frac{5}{2}
\end{cases}.
\]
We can now replace $t$ with an arbitrary linear combination of $t$ that we will call $s$. We can now
rewrite the vector in terms of $s$ as such:
\[
\begin{cases}
x = 3s + 2\\
y = 7s + 1\\
z = 5s
\end{cases}.
\]
This vector lies along the same line described by the original equation.
\end{quote}

\item Do the parametric equations $x = 5t^{2} - 1, y = 2t^{2} + 3, z = 1 − t^{2}$ determine a line?
Explain?
\begin{quote}
No, the given parametric equations do not form a line. If we represent the equations in symmetric
form, they will have nonlinear terms.
\end{quote}

\item Find the points of intersection of the line $x = 2t − 3, y = 3t +2, z = 5-t$ with each of the
coordinate planes $x = 0, y = 0$, and $z = 0$.
\begin{quote}
The line intersects the plane at $(\frac{3}{2},-\frac{2}{3},5)$.
\end{quote}

\item Does the line $x = 5-t, y=2t-3, z=7t+1$ intersect the plane $x - 3y+z=1$? Why?
\begin{quote}
Substituting in our values for $x,y,z$ into the equation of the plane yields
\[
(5-t) - 3(2t-3) + (7t+1) = 2.
\]
Solving for $t$ gives us
\[
-3 = 2,
\]
which is impossible. Therefore, the line does not intersect the plane.
\end{quote}

\item Find the point of intersection of the two lines $l_{1}: x = 2t + 3,y = 3t + 3,z = 2t + 1$ and
$l_{2}: x = 15 - 7t, y = t - 2, z = 3t - 7$.
\begin{quote}
To find the point of intersection, we can set the values in each equation equal to each other.
\[
\begin{cases}
2s + 3 = 15 - 7t\\
3s + 3 = t - 2\\
2s + 1 = 3t - 7
\end{cases}.
\]
We only have two unknowns, so we only need two equations:
\begin{align*}
\begin{cases}
2s + 7t &= 12\\
2s - 3t &= 8
\end{cases}.
\end{align*}
Solving the system yields $t = 2$. Therefore the two lines intersect at $t = 2$.
\end{quote}
\end{enumerate}
\end{document}

### 2016-01-13 math_modeling/in_class/babies.R

@@ -1,0 +1,19 @@
+# Example usage of c call
+
+# age <- c(1,3,5,2,11,9,3)
+# weight <- c(4.4,5.3,7.2,5.2,8.5,7.3,6)
+# #linear model
+# linMod <- lm(weight ~ age)
+# print(linMod)
+# #standard deviation
+# print(sd(age))
+# print(sd(weight))
+# #plot age vs. weight
+# plot(age, weight)
+
+#Data frames
+ages <- c(33,45,22,32,19,54)
+diabetic <- c(TRUE,FALSE,FALSE,TRUE,TRUE,FALSE)
+status <- c('poor','improved','excellent','poor','improved','poor')
+frame <- data.frame(ages, diabetic, status)
+print(frame)

### 2016-01-13 math_modeling/in_class/boat_problem.R

@@ -1,0 +1,11 @@
+# Example of problem 2
+# Given a plane 2x + 3y = -4z
+#
+# x	  y     z
+# 1   1   -1.25
+# 1   2     -2
+# 3   2   -2.98
+
+lake <- read.table('lakebed.txt')
+mlm <- lm(lake[,3] ~ lake[,2] + lake[,1])
+print(mlm)

### 2016-01-13 math_modeling/in_class/matrix.R

@@ -1,0 +1,3 @@
+cells <- c(1,26,24,32)
+matrix <- matrix(cells, nrow=2, ncol=2)
+print(matrix)

### 2016-01-14 writing_in_ethics/paper_one.tex
created paper class

--- 2016-01-13 vector_calc/homework/hw2.tex
@@
-\documentclass{hw}
-
-% 11,13,17,19,23,25,27,31,35,37,42
+\documentclass{paper}
 
 \begin{document}
-\makeheader{2}
+\drafth{paper_one.tex}
 
-\begin{enumerate}
-\item Let $\vv{a_{1}} = (1,1)$ and $\vv{a_{2}} = (1,-1)$.
-\begin{enumerate}
-\item Write the vector $\vv{b_{1}} = (3, 1)$ as $c_{1}a_{1} + c_{2}a_{2}$, where
-$c_{1}$ and $c_{2}$ are appropriate scalars.
-\begin{align*}
-(3,1) = (c_{1} + c_{2}, c_{1} - c_{2})\\
-\left[
-\begin{array}{c c | c}
-1 & 1 & 3\\
-1 & -1 & 1
-\end{array}
-\right] \to
-\left[
-\begin{array}{c c | c}
-1 & 0 & 2\\
-0 & 1 & 1
-\end{array}
-\right]\\
-\vv{b} = 2(1,1) + 1(1,-1)
-\end{align*}
-
-\item Do the same for $(3,-5)$.
-\begin{align*}
-(3,1) = (c_{1} + c_{2}, c_{1} - c_{2})\\
-\left[
-\begin{array}{c c | c}
-1 & 1 & 3\\
-1 & -1 & -5
-\end{array}
-\right] \to
-\left[
-\begin{array}{c c | c}
-1 & 0 & -1\\
-0 & 1 & 4
-\end{array}
-\right]\\
-\vv{b} = -1(1,1) + 4(1,-1)
-\end{align*}
-
-\item Show that any vector $\vv{b} = (b1, b2)$ in $\RR^{2}$ may be written in the form $c_{1} a_{1} +
-c_{2} a_{2}$ for appropriate choices of the scalars $c_{1}, c_{2}$.
-\begin{quote}
-Assume that we can express any vector $(x_{1},x_{2})$ as $(c_{1} + c_{2}) + (c_{1} - c_{2})$, where
-$x_{1},x_{2}\in\RR$. Then the solution for the corresponding matrix is
-\[
-\left[
-\begin{array}{c c | c}
-1 & 0 & \frac{x_{1} + x_{2}}{2}\\
-0 & 1 & \frac{x_{1} - x_{2}}{2}
-\end{array}
-\right].
-\]
-Since $\frac{x_{1} + x_{2}}{2}$ and $\frac{x_{1} - x_{2}}{2}$ are both in $\RR$, any vector of the form
-$(x_{1},x_{2})$ can be written as a linear combination of $\vv{a_{1}}$ and $\vv{a_{2}}$.
-\end{quote}
-\end{enumerate}
-
-\item Write the following as a set of parametric equations: The line in $\RR^{3}$ through the point $(2,
-−1, 5)$ that is parallel to the vector $i + 3j - 6k$.
-\[
-\begin{cases}
-x = t+2\\
-y = 3t - 1\\
-z = 5 - 6t
-\end{cases}
-\]
-\item Write the following as a set of parametric equations: The line in $\RR^{3}$ through the points
-$(1,4,5)$ and $(2,4,1)$.
-\[
-\begin{cases}
-x = t + 1\\
-y = 4\\
-z = 5 - 6t
-\end{cases}
-\]
-\item Write the following as a set of parametric equations: Write a set of parametric equations for the
-line in $\RR^{4}$ through the point $(1, 2, 0, 4)$ and parallel to the vector $(−2, 5, 3, 7)$.
-\[
-\begin{cases}
-x_{1} =1-2t\\
-x_{2} =5t+2\\
-x_{3} =3t\\
-x_{4} =7t+4
-\end{cases}
-\]
-
-\item Give a symmetric form for the line having parametric equations $x = t + 7$, $y = 3t - 9$, $z = 6 -
-8t$.
-\[
-\begin{cases}
-x = t + 7\\
-y = 3t - 9\\
-z = 6-8t
-\end{cases}\to
-\begin{cases}
-t = x - 7\\
-t = \frac{y}{3} + 3\\
-t = \frac{6 - z}{8}
-\end{cases}
-\]
-
-\item Give a set of parametric equations for the line with symmetric form
-\[
-\frac{x + 5}{3} = \frac{y - 1}{7} = \frac{z + 10}{-2}.
-\]
-\[
-\begin{cases}
-t = \frac{x + 5}{3}\\
-t = \frac{y - 1}{7}\\
-t = \frac{z + 10}{-2}
-\end{cases}\to
-\begin{cases}
-x = 3t - 5\\
-y = 7t + 1\\
-z = -2t -10
-\end{cases}
-\]
-
-\item Show that the two sets of equations
-\[
-\frac{x-2}{3} = \frac{y-1}{7} = \frac{z}{5} \text{ and } \frac{x+1}{-6} = \frac{y+6}{-14} =
-\frac{z+5}{10}
-\]
-represent the same line in $\RR^{3}$.
-\begin{quote}
-If we move to parametric coordinates, we have
-\[
-\begin{cases}
-x = 3t + 2\\
-y = 7t + 1\\
-z = 5t
-\end{cases}
-\text{ and }
-\begin{cases}
-x = -6t - 1\\
-y = -14t - 6\\
-z = -10t - 5
-\end{cases}.
-\]
-If we take the second vector and scale it by $-\frac{1}{2}$, we get
-\[
-\begin{cases}
-x = 3t + \frac{1}{2}\\
-y = 7t + 3\\
-z = 5t = \frac{5}{2}
-\end{cases}.
-\]
-We can now replace $t$ with an arbitrary linear combination of $t$ that we will call $s$. We can now
-rewrite the vector in terms of $s$ as such:
-\[
-\begin{cases}
-x = 3s + 2\\
-y = 7s + 1\\
-z = 5s
-\end{cases}.
-\]
-This vector lies along the same line described by the original equation.
-\end{quote}
-
-\item Do the parametric equations $x = 5t^{2} - 1, y = 2t^{2} + 3, z = 1 − t^{2}$ determine a line?
-Explain?
-\begin{quote}
-No, the given parametric equations do not form a line. If we represent the equations in symmetric
-form, they will have nonlinear terms.
-\end{quote}
-
-\item Find the points of intersection of the line $x = 2t − 3, y = 3t +2, z = 5-t$ with each of the
-coordinate planes $x = 0, y = 0$, and $z = 0$.
-\begin{quote}
-The line intersects the plane at $(\frac{3}{2},-\frac{2}{3},5)$.
-\end{quote}
-
-\item Does the line $x = 5-t, y=2t-3, z=7t+1$ intersect the plane $x - 3y+z=1$? Why?
-\begin{quote}
-Substituting in our values for $x,y,z$ into the equation of the plane yields
-\[
-(5-t) - 3(2t-3) + (7t+1) = 2.
-\]
-Solving for $t$ gives us
-\[
--3 = 2,
-\]
-which is impossible. Therefore, the line does not intersect the plane.
-\end{quote}
-
-\item Find the point of intersection of the two lines $l_{1}: x = 2t + 3,y = 3t + 3,z = 2t + 1$ and
-$l_{2}: x = 15 - 7t, y = t - 2, z = 3t - 7$.
-\begin{quote}
-To find the point of intersection, we can set the values in each equation equal to each other.
-\[
-\begin{cases}
-2s + 3 = 15 - 7t\\
-3s + 3 = t - 2\\
-2s + 1 = 3t - 7
-\end{cases}.
-\]
-We only have two unknowns, so we only need two equations:
-\begin{align*}
-\begin{cases}
-2s + 7t &= 12\\
-2s - 3t &= 8
-\end{cases}.
-\end{align*}
-Solving the system yields $t = 2$. Therefore the two lines intersect at $t = 2$.
-\end{quote}
-\end{enumerate}
 \end{document}

### 2016-01-14 writing_in_ethics/paper_one.tex
working on paper 1 for ethics

@@
 
+\title{Utilitarianism}
+\date{}
+\author{}
+
 \begin{document}
-\drafth{paper_one.tex}
+\drafth{}
+\inlinetitle
+
+The ethical theory of Utilitarianism can be used to analyze the questions that arise from Alexis'
+situation.\\
+
+The first question that needs to be answered is whether or not Alexis has done anything that can be
+considered morally wrong. According to the principal of utility, it appears that Alexis has not actually
+committed a moral wrong. In Alexis' situation, she has two choices; she either uses the computers, as she
+has chosen, or she does not. If Alexis had chosen to not use the other user's login information, she would
+not have had an outcome that increased her happiness.
 

### 2016-01-14 number_theory/hw/hw1.tex
number theory

\documentclass{hw}

\begin{document}
\makeheader{1}

\begin{enumerate}
\item What are the possible remainders when a perfect square is divided by 3 or by 6?
\begin{quote}
Let $k\in\ZZ$ such that $a = k^{2}$. By the division algorithm, we have that
\[
k = 3q_{1} + r_{1} \qquad\text{and}\qquad k = 6q_{2} + r_{2}.
\]
\underline{case $k = 3q_{1} + r_{1}$}: Squaring $k$ yields
\[
k^{2} = 9q_{1}^{2} + 6q_{1}r_{1} + r_{1}^{2}.
\]
Since $k^{2} = a$, we have that
\begin{align*}
a &= 3(3q_{1}^{2} + 2q_{1}r_{1}) + r_{1}^{2}\\
&= 3q_{0} + r_{0},\text{ where } q_{0} = (3q_{1}^{2} + 2q_{1}r_{1})\text{ and } r_{0} = r_{1}^{2}.
\end{align*}
We know by the division algorithm that $r_{0}\in\{ 0,1,2\}$. If we let $r_{0} = 0$ or $r_{0} = 1$, then
$r_{1}^{2} < 4$ and still in $\{ 0,1,2\}$. Therefore, 0 and 1 are possible remainders. If we let
$r_{0} = 2$, then $r_{1}^{2} = 4$, which is not less than 4. However, we have that
\begin{align*}
a &= 3q_{0} + 2^{2}\\
&= 3q_{0} + 4\\
&= 3q_{0} + 3 + 1\\
&= 3(q_{0} + 1) + 1,
\end{align*}
Which implies that 1 is a valid remainder. Therefore, 0 and 1 are the only valid remainders.\\

\underline{case $k = 6q_{2} + r_{2}$}:
If we square $k$, then we have
\[
k^{2} = 36q_{2}^{2} + 12q_{2}r_{2} + r_{2}^{2}.
\]
Since $a = k^{2}$, we have that
\begin{align*}
a &= 36q_{2}^{2} + 12q_{2}r_{2} + r_{2}^{2}\\
&= 6(6q_{2}^{2} + 2q_{2}r_{2}) + r_{2}^{2}\\
&= 6q_{0} + r_{0},\text{ where } q_{0} = 6q_{2}^{2} + 2q_{2}r_{2} \text{ and } r_{0} = r_{2}^{2}.
\end{align*}
We know that for $r_{0} = 0,1,\text{ and }2$, $r_{2}^{2} \leq 6$, so they are valid remainders. For
$r_{0} = 3,4,\text{ and } 5$, we have
\end{quote}
\begin{minipage}{0.3\textwidth}
\underline{$r_{0} = 3$}
\begin{align*}
a &= 6q_{0} + 9\\
&= 6q_{0} + 6 + 3\\
&= 6(q_{0} + 1) + 3
\end{align*}
$\therefore$ 3 is a valid remainder.
\end{minipage}
\begin{minipage}{0.3\textwidth}
\underline{$r_{0} = 4$}
\begin{align*}
a &= 6q_{0} + 16\\
&= 6q_{0} + 12 + 4\\
&= 6(q_{0} + 2) + 4
\end{align*}
$\therefore$ 4 is a valid remainder.
\end{minipage}
\begin{minipage}{0.3\textwidth}
\underline{$r_{0} = 5$}
\begin{align*}
a &= 6q_{0} + 25\\
&= 6q_{0} + 24 + 1\\
&= 6(q_{0} + 4) + 1
\end{align*}
$\therefore$ 1 is a valid remainder.
\end{minipage}
\begin{quote}
Therefore, the valid remainders are $\{ 0, 1, 2, 3, 4\}$.
\end{quote}
$\triangle$

\item Suppose $a,b,c,d\in\ZZ$ are such that $a|b$ and $c|d$. Prove that $ac|bd$.
\begin{quote}
Since $a|b$, we have that $b = an$ for some $n\in\ZZ$. We also have that $c|d$, so $d=cm$ for some
$m\in\ZZ$. The product $bd = (an)(cm)$, which, after rearranging, yields $bd = (nm)(ac)$. We know that
$nm\in\ZZ$, so we let $j = nm$. Then $bd = jac$, so $ac|bd$.\\
\end{quote}
$\triangle$

\item Suppose $a,b,m\in\ZZ$ and $m\neq 0$. Prove that $a|b$ if and only if $ma|mb$.
\begin{quote}
Suppose $a|b$. Then $b = na$ for some $n\in\ZZ$. Multiplying both sides of the equation yields
$mb = mna$; therefore $ma|mb$.\\
Suppose $ma|mb$, and $m\neq 0$. Then $mb = mna$ for some $n\in\ZZ$. We know that $\frac{m}{m} = 1$, so
dividing both sides by $m$ gives us $b = na$. Therefore $a|b$.
\end{quote}
$\triangle$

\item Suppose $a,b\in\ZZ$ with $b\neq 0$ and $a|b$. Prove that $|a| \leq |b|$.
\end{enumerate}
\end{document}

### 2016-01-14 number_theory/hw/hw1.tex
working on number theory homework

@@
 \documentclass{hw}
+
+\usepackage{cancel}
 
@@
 \item Suppose $a,b\in\ZZ$ with $b\neq 0$ and $a|b$. Prove that $|a| \leq |b|$.
+\begin{quote}
+Since $a|b$, we have that $b=an$ for some $n\in\ZZ$. We first need to show that $n \neq 0$.
+Since $b\neq 0$, $an\neq 0$. Therefore both $a$ and $n$ must not be equal to 0. If we apply
+the absolute value to $b$, we have that $|b| = |an|$. We know that $|an| = |a||n|$. Since
+$n\neq 0$, $|n| \geq 1$. We also know that $|a| \geq 1$, since $a\neq 0$. Therefore
+$|a||n| \geq |a|$, or $|an| \geq |a|$. Since $|b| = |an|$, we have that $|a|\leq|b|$.
+\end{quote}
+$\triangle$
+
+\item Suppose that $a,b,c,d\in\ZZ$ are such that $a|b$ and $c|d$. Does $a+c$ necessarily
+divide $b+d$?
+\begin{quote}
+Assume $a+c$ divides $b+d$. Let $a=1$ and $c=-1$. Both 1 and -1 divide all elements of $\ZZ$, but
+$1 + -1 = 0$, and 0 does not divide anything. Therefore by counterexample it does not.
+\end{quote}
+$\triangle$
+
+\item Suppose that $b\in\ZZ$.
+\begin{enumerate}
+\item If $4|b^{2}$, prove that $2|b$.
+\begin{quote}
+Assume that $4|b^{2}$ but $2\cancel{|}b$. Then $b = 2n + r$ for some $n$ and $r$ in $\ZZ$. We know
+that $r\neq 0$, and that $0 < r < 2$. Therefore $r = 1$. If we square both sides, we get
+\begin{align*}
+b^{2} &= (2n+1)^{2}\\
+&= 4n^{2}+4n+1\\
+&= 4(n^{2} + n) + 1.
+\end{align*}
+Therefore $4\cancel{|}b$, since it has a remainder of 1.
+\end{quote}
+
+\end{enumerate}
 \end{enumerate}

### 2016-01-15 writing_in_ethics/paper_one.tex
working on literally everything

@@
 \author{}
+
+\addbibresource{generic_bib.bib}
 
@@
 has chosen, or she does not. If Alexis had chosen to not use the other user's login information, she would
-not have had an outcome that increased her happiness.
+not have had an outcome that increased her happiness. Even though Alexis did steal the information from
+another student, that student did not experience any decrease in happiness, according to the scenario.\\
+
+Bentham proposes several attributes that should be taken into account when calculating the utility of an
+action. The attributes that apply most directly to Alexis' situation are \textit{Intensity},
+\textit{Duration}, and \textit{Extent} \cite{ethics}.
+
+\newpage
+\printbibliography
 

### 2016-01-15 math_modeling/in_class/matrix.R

@@ -1,3 +1,0 @@
-cells <- c(1,26,24,32)
-matrix <- matrix(cells, nrow=2, ncol=2)
-print(matrix)

### 2016-01-15 math_modeling/in_class/probability.R

@@ -1,0 +1,22 @@
+#!/usr/bin/RScript
+
+# binomial distributions
+#
+# _binom(search_prob,total_number,probability)
+#
+# several prefixes
+# r - random distribution - random plot
+# d - density distribution - finding value of function
+# p - cumulative distribution - area up to a region
+
+# What is the probability of finding one left handed person in a group
+# 17, given that the probability of finding a left handed person is 1.7%.
+bdist <- dbinom(1,1,0.1)
+print(bdist)
+
+# Doing a random distrubution is similar
+rdist <- rbinom(1000,17,0.1)
+print(rdist)
+
+# _norm: normal distrubution
+# _chisq: chi squared distribution

### 2016-01-15 math_modeling/in_class/time_series.R

@@ -1,0 +1,27 @@
+#!/usr/bin/RScript
+
+# Example Business Data
+#
+# 153 189 221 215 302 223 201 173 121 106 86 87 108
+# 133 177 241 228 283 255 238 164 128 108 87 74 95
+# 145 200 187 201 292 220 233 172 119 81 65 76 74
+# 111 170 243 178 248 202 163 139 120 96 95 53 94
+
+sales <- c(153, 189, 221, 215, 302, 223, 201, 173, 121, 106, 86, 87, 108,
+			133, 177, 241, 228, 283, 255, 238, 164, 128, 108, 87, 74, 95,
+			145, 200, 187, 201, 292, 220, 233, 172, 119, 81, 65, 76, 74,
+			111, 170, 243, 178, 248, 202, 163, 139, 120, 96, 95, 53, 94)
+
+# Create a time series out of the sales vector
+# sales = the vector
+# start: the start time (year: 2011, month: 1)
+# end: the end time (year: 2014, month 12)
+# frequency: occurance -- 13 x 4 = 52 in our case
+series <- ts(sales, start=c(2011,1), end=c(2014,13), frequency=13)
+print(series)
+plot(series)
+
+# try to smooth out the data set with a moving average
+plot(ma(series, 3))
+fit <- stl(series, s.window="period")
+plot(fit)

### 2016-01-15 writing_in_ethics/paper_one.tex
working on paper

--- 2016-01-14 writing_in_ethics/paper_one.tex
@@
 
+\title{Utilitarianism}
+\date{}
+\author{}
+
+\addbibresource{generic_bib.bib}
+
 \begin{document}
-\drafth{paper_one.tex}
+\drafth{}
+\inlinetitle
+
+The theory of Utilitarianism can be used to quantitatively analyze various situations that arise in day to day life; it can be applied the scenario presented in \textit{Ethics for the Information Age} that concerns Alexis' education.
+In order to analyze the situation and determine whether or not Alexis' actions were ethically ``good" or ``bad", the total utility change needs to be calculated.
+Alexis' actions can be considered ethically good if the benefit of her actions exceed the harms.
+Alexis' situation can be analyzed using both Rule and Act Utilitarianism.\\
+
+Jeremy Bentham, a famous English philosopher, proposed several attributes that should be taken into account when calculating the utility of an action \cite[75]{ethics}.
+The first attribute, \textit{Intensity}, is concerned with the magnitude of the experience.
+The magnitude of Alexis' experience can be measured in the amount of happiness she experienced.
+She was able to achieve her goal and even positively affected others around her; presumably she was able to save her family and herself a large amount of money while also bettering her own life in the process through a college education.
+Both of these aspects should be taken into account as contributing positively to the magnitude of the situation.\\
+
+The next attribute, \textit{Duration}, applies greatly to Alexis' scenario.
+The duration of the experience extends far beyond the time that Alexis spent in the library on the computer; in fact, the education she received will continue to affect her for the rest of her life.
+However, Alexis stole the information from another student at the university.
+The duration of that action, coupled with the amount of time she spent using the stolen information, is still greatly overshadowed by the overall happiness that Alexis experienced.
+It is reasonable to say that the amount of happiness gained under the \textit{Duration} criterion is far greater than the amount of happiness lost.\\
+
+Bentham mentions the idea of certainty in his criteria for testing actions under Act Utilitarianism.
+There was always a chance that Alexis' end goal could have never been met, in which case the overall unhappiness produced would have far outweighed the happiness.
+However, Alexis did ultimately achieve her goal to be accepted into college; on top of just being accepted into college, she was also provided with a full ride, which positively affected her parents as well.
+As a result, the attribute of \textit{Certainty} greatly adds to the overall happiness gained from Alexis' decision.
+\\
+
+The final attribute that Bentham proposes is concerned with the number of people affected by the action.
+Alexis' actions do in fact affect several other beings, and not all the interactions are not necessarily positive.
+As mentioned before, Alexis' family benefitted greatly from the situation.
+They were already poor; this fact may have greatly affected Alexis' ability to attend college.
+However, due to her actions, Alexis was able to obtain a scholarship that provided her with the funds to attend college for free.
+
+\newpage
+\printbibliography
 

### 2016-01-17 number_theory/hw/hw1.tex
homework and stuff

@@
 \end{quote}
+\item If $9|b^{2}$, prove that $3|b$.
+\begin{quote}
+Assume that $9|b^{2}$, but $3\cancel{|}b$. Then $b=3n+r$ where $n$ and $r$ and in $\ZZ$, and
+$0 < r < 3$. If we let $r=1$, then we have
+\begin{align*}
+b &= 3n + 1\\
+b^{2} &= (3n+1)^{2}\\
+&= 9n^{2} + 6n + 1\\
+&= 3(3n^{2} + 2n) + 1.
+\end{align*}
+Therefore $3\cancel{|}b^{2}$, and since 3 is a factor of 9, $9\cancel{|}b^{2}$.
+\end{quote}
+$\triangle$
+\end{enumerate}
 
-\end{enumerate}
+\item Use the Euclidian Algorithm to calculate $gcd(1485,1745)$.
+\begin{quote}
+\begin{align*}
+1745 &= 1485(1) + 260\\
+1485 &= 260(5) + 185\\
+260 &= 185(1) + 75\\
+185 &= 75(2) + 35\\
+75 &= 35(2) + 5\\
+35 &= 5(7) + 0\\
+\end{align*}
+Therefore $gcd(1485,1745) = gcd(5,0) = 5$.
+\end{quote}
+
+\item Suppose that $a$ and $b$ are positive integers. Prove that $gcd(a,b) = a$ if and
+only if $a|b$.
+\begin{quote}
+Assume that $gcd(a,b) = a$. Then $a$ is a divisor of both $a$ and $b$ by the definition of the $gcd$.
+Therefore $a|b$.
+\end{quote}
 \end{enumerate}

### 2016-01-17 partial_differential_equations/bessel.tex
homework and stuff

\section{Bessel Functions}
\hrule
\noindent\\

The 2D heat equation provided us a way to describe heat flow through a 2D surface, such as a thin rod.
Now imagine that we have a surface that is shaped like a large cylinder. We've now moved into 3D space,
and as such we will have to come up with a way to describe the heat transfer through the cylinder. Let's
start by recalling Laplace's equation,
\[
\nabla^{2}u = 0,
\]
which described a vector composed of the all the partial derivatives of $u$. We will need to keep this
equation in mind as we concern ourselves with the 3D heat equation.\\

Our end goal will be to come up with a formal solution for the equation
\[
u_{t} = k^{2}(u_{xx} + u_{yy}).
\]
This is, of course, the heat equation expanded to a cylinder. Recall that when we dealt with Laplace's
equation on a disk, we moved into polar coordinates. The 3D heat equation contains the laplacian in it,
represented as
\[
u_{xx} + u_{yy}.
\]
Moving into polar coordinates will allow us to have rectangular bounds on our function. Transforming our
heat equation yields
\[
u_{t} = k^{2}(u_{rr} + {1 \over r}u_{r} + {1\over r^{2}}u_{\theta\theta}).
\]
We now have a rectangular bound, which allows us to use the familiar separation of variables technique to
solve the PDE. In this case, we want to separate the spacial variables ($r,\theta$) from the time variable
($t$). We have a general solution of
\[
u = V(r,\theta)T(t) = 0.
\]
Setting them equal to a common value yields
\[
{T \over k^{2}T} = {V_{rr} + {1\over r}V_{r} + {1\over r^{2}}V_{\theta\theta} \over V} = -\lambda^{2}.
\]
As expected, we are left with 2 ODEs. The first ODE, concerning $T$, will be left alone for now.
The remaining ODE can be solved by using separation of variables again. Doing so yields
\begin{gather*}
V = R(r)\nu(\theta)\\
\begin{cases}
\eta^{''} + \gamma\eta = 0\\
r^{2}R'' + rR' + (k^{2}r^{2} - \gamma^{2})R = 0
\end{cases}.
\end{gather*}
Again, we will leave the $\eta$ function for later, and instead turn our attention to the equation
concerning $R$. This leads us to our discussion of Bessel Functions.\\

### 2016-01-17 vector_calc/homework/hw3.tex
homework and stuff

\documentclass{hw}
\begin{document}
\makeheader{3}

\begin{enumerate}
\item Compute $a\cdot b$, $||a||$ and $||b||$ where $\vv{a} = 4i -3j + k$ and $\vv{b}=i+j+k$.
\begin{quote}
\begin{align*}
a\cdot b &= 4 - 3 + 1\\
&= 2\\
||a|| &= \sqrt{16+9+1}\\
&= \sqrt{26}\\
||b|| &= \sqrt{1+1+1}\\
&= \sqrt{3}
\end{align*}
\end{quote}

\item Find the angle between $\vv{a} = \sqrt{3}\ i + j$ and $\vv{a} = -\sqrt{3}\ i + j$.
\begin{align*}
\theta &= \arccos{\left({a\cdot b \over ||a||\ ||b||}\right)}\\
&= \arccos{\left({-2\over 4}\right)}\\
&= \arccos{\left(-{1\over 2}\right)}\\
&= {2\pi \over 3}
\end{align*}

\item Find the angle between $\vv{a} = i + j$ and $\vv{b} = i + j + k$.
\begin{align*}
\theta &= \arccos{\left({a\cdot b \over ||a||\ ||b||}\right)}\\
&= \arccos{\left({2 \over \sqrt{6}}\right)}
\end{align*}

\item Find the angle between $\vv{a} = (1,-2,3)$ and $\vv{b} = (3,-6,-5)$.
\begin{align*}
\theta &= \arccos{\left({a\cdot b \over ||a||\ ||b||}\right)}\\
&= \arccos{(0)}\\
&= {\pi \over 2}
\end{align*}

\item Calculate $proj_{a}b$ where $\vv{a} = i + j$ and $\vv{b} = 2i + 3j - k$.
\begin{align*}
proj_{a}b &= {a\cdot b \over ||a||^{2}}\vv{a}\\
&= {4\over 2}(i + j)\\
&= 2i + 2j
\end{align*}

\item Calculate $proj_{a}b$ where $\vv{a} = i + j + 2k$ and $\vv{b} = 2i - 4j + k$.
\begin{align*}
proj_{a}b &= {a\cdot b \over ||a||^{2}}\vv{a}\\
&= {0\over 4}(i+j+2k)\\
&= 0
\end{align*}

\item Give a vector of length 3 that points in the same direction as the vector $i + j - k$.
\begin{quote}
The length is just a scalar value multiplied by the vector. Therefore we have the unit vector $i + j - k$,
with length 3 will be
\[3i + 3j - 3k.\]
\end{quote}

\item Is there ever a case where $proj_{a}b = proj_{b}a$?
\begin{quote}

\end{quote}
\end{enumerate}
\end{document}

### 2016-01-18 writing_in_ethics/paper_one.tex
working on paper

@@
 \title{Utilitarianism}
-\date{}
-\author{}
 
@@
 The theory of Utilitarianism can be used to quantitatively analyze various situations that arise in day to day life; it can be applied the scenario presented in \textit{Ethics for the Information Age} that concerns Alexis' education.
-In order to analyze the situation and determine whether or not Alexis' actions were ethically ``good" or ``bad", the total utility change needs to be calculated.
+In order to analyze the situation and determine whether or not her actions were ethically ``good" or ``bad", the total utility change needs to be calculated.
 Alexis' actions can be considered ethically good if the benefit of her actions exceed the harms.
-Alexis' situation can be analyzed using both Rule and Act Utilitarianism.\\
+The situation can be analyzed using both Rule and Act Utilitarianism.
 
@@
 She was able to achieve her goal and even positively affected others around her; presumably she was able to save her family and herself a large amount of money while also bettering her own life in the process through a college education.
-Both of these aspects should be taken into account as contributing positively to the magnitude of the situation.\\
+Both of these aspects should be taken into account as contributing positively to the magnitude of the situation.
 
@@
 The duration of that action, coupled with the amount of time she spent using the stolen information, is still greatly overshadowed by the overall happiness that Alexis experienced.
-It is reasonable to say that the amount of happiness gained under the \textit{Duration} criterion is far greater than the amount of happiness lost.\\
+It is reasonable to say that the amount of happiness gained under the \textit{Duration} criterion is far greater than the amount of happiness lost.
 
@@
 As a result, the attribute of \textit{Certainty} greatly adds to the overall happiness gained from Alexis' decision.
-\\
 
@@
 However, due to her actions, Alexis was able to obtain a scholarship that provided her with the funds to attend college for free.
+In this way, Alexis affected, in a positive way, a large number of people that were important to her.
+However, she did also affect the student from whom she stole the login information.
+Her actions had a slight negative affect on that student, but not enough of a negative affect to outweigh the positive affect she had on her family.
+
+Bentham's methods applied to Alexis' actions under the ethical theory of Act Utilitarianism.
+Under Act Utilitarianism, Alexis has made an ethically good decision; the amount of overall happiness produced by her actions far outweighs the unhappiness produced.
+However, many of the calculations are not exact since the certainty of the consequences of Alexis' actions were often assumed \cite[77]{ethics}.
+The theory of Rule Utilitarianism can also provide a viewpoint to determine whether or not Alexis' actions were ethically right or wrong.
+
+Rule Utilitarianism, much like Kantianism, is based in rules that should be morally followed.
+Unlike Kantianism, however, Rule Utilitarianism adopts rules based on the total overall universal happiness that would result from adopting said rule.
+For the most part, the outcome of Alexis' situation is the same under both Rule and Act Utilitarianism.
+The main moral dilemma that Alexis faces is whether or not to steal and use the student's information.
+Rather than calculating the increase in her own happiness, Alexis must now calculate the impact her decision will have if the rule were to be adopted universally.
+Simply put, Alexis must decide whether or not it is universally okay to steal private information for one's own personal gain.
+As a universal rule, it is obvious that theft of private information should not be universally accepted as alright.
+Therefore, Alexis' decision, under Rule Utilitarianism, was morally wrong.
+
+Both Rule and Act Utilitarianism provide interesting views on the ethical decision that Alexis made.
+Under Rule Utilitarianism, Alexis' decision was ethically wrong; under Act Utilitarianism, her decision was ethically right.
+Both of the methods used to analyze the situation provide a method to determine whether or not Alexis made the right choice; nether Rule nor Act Utilitarianism are absolute.
 

### 2016-01-18 writing_in_ethics/bldp.sh

@@ -1,0 +1,8 @@
+if [ -z "$1" ]
+then
+	echo "No File Provided."
+else
+	lualatex --shell-escape "$1"
+	bibtex $(basename $1 .tex)
+	lualatex --shell-escape "$1"
+fi

### 2016-01-18 cleaner.sh
added cleaner

@@ -1,0 +1,14 @@
+recurse () {
+	for type in ./* ; do
+		if [ -d "$type" ]
+		then
+			cd "$type"
+			recurse
+			cd ..
+		else
+			gbbs.sh
+		fi
+	done
+}
+
+recurse

### 2016-01-18 math_modeling/hw/hw1/num_one.R
math modeling work

@@ -1,0 +1,5 @@
+if(require("forecast")){
+	data <- scan("hmwk1.txt")
+	plot(data)
+	plot(ma(data,4))
+}

### 2016-01-18 math_modeling/in_class/batch.R

@@ -1,0 +1,17 @@
+#	function name: batching
+#	parameters:
+#	    k   size of each batch
+#           x   data
+#       return value:
+#           an array of the averages of each batch
+# 	description: this function separates the data x into
+#       floor(length(x)/k) batches and returns the array of
+#       the averages of each batch
+#	To run the code, type y<-batching(k,x) to place the consecutive
+#       batch means of size k, of the data in the vector x, in the vector y
+batching <- function(k,x)  {
+    m <- floor(length(x)/k);
+    result <- rep(0,m);
+    for(i in 1:m) result[i] <- mean(x[(i-1)*k+(1:k)]);
+    return(result);
+}

### 2016-01-18 writing_in_mathematics/ignore/dum.rb
fix ignores

@@ -1,31 +1,0 @@
-def initialize(log2m, register_set = nil)
-    @log2m = log2m
-    @count = 2 ** log2m
-    @register_set = register_set || RegisterSet.new(@count)
-
-    case log2m
-    when 4
-        @alphaMM = 0.673 * @count * @count
-    when 5
-        @alphaMM = 0.697 * @count * @count
-    when 6
-        @alphaMM = 0.709 * @count * @count
-    else
-        @alphaMM = (0.7213 / (1 + 1.079 / @count)) * @count * @count
-    end
-end
-
-def cardinality
-    register_sum = 0.0
-    zeros = 0.0
-    @register_set.each do |value|
-        register_sum += 1.0 / (1 << value)
-        zeros += 1 if value == 0
-    end
-    estimate = @alphaMM * (1 / register_sum)
-    if estimate <= (5.0 / 2.0) * @count
-        (@count * Math.log(@count / zeros)).round
-    else
-        estimate.round
-    end
-end

### 2016-01-18 writing_in_mathematics/twitter_problem/const.rb

@@ -1,22 +1,0 @@
-class Constants
-	@@oh_one = "JtwwG40EW2O6CKBNp1qeD57IH"
-	@@oh_two = "NUSZ0jPyeWwTKPAlUDYLnoeatyDCsPp5waIJsYnQM6LJ5oDjDJ"
-	@@oh_three = "3494713935-hoqlE0ocEarpGUbSosIEi3YJvHqrJH3j60x7Z74"
-	@@oh_four = "fKTi6UiCESwluB9UbBOsdcBVjHHPfUOxr6PbSWy4UCUi8"
-
-	def self.oh_one
-		@@oh_one
-	end
-
-	def self.oh_two
-		@@oh_two
-	end
-
-	def self.oh_three
-		@@oh_three
-	end
-
-	def self.oh_four
-		@@oh_four
-	end
-end

### 2016-01-18 math_modeling/hw/hw1/result.tex
working on modeling

\documentclass{hw}

\usepackage{minted}

\begin{document}
\makeheader{1}

\begin{enumerate}
\item Dr. Martin has emailed you the data set, hmwk1.txt, which gives the time evolution of a certain
quantity of interest to him. Using R, plot the original data. In addition, using R’s ma() function, plot a
smoothed version of the data. What attribute of the original data is hidden by referring only to the
smoothed data?
\begin{quote}
\inputminted{r}{num_one.R}
The output of the above code is\\
\begin{minipage}{0.5\textwidth}
\includegraphics[scale=0.4]{01data_plot}
\end{minipage}
\begin{minipage}{0.5\textwidth}
\includegraphics[scale=0.4]{01data_smoothed}
\end{minipage}
When the data is smoothed, the outlying points are lost.
\end{quote}

\newpage
\item Dr. Martin has emailed to you the data set, co2.txt. The data represents 16 years of collecting
monthly $CO_{2}$ data on the island of Hawaii, with the first year of $CO_{2}$ starting in January of
1958. a) Using R, plot the data. b) Using R’s stl() command, plot the 1. trend, 2. seasonal, and 3.
irregular components of the data.
\begin{quote}
\inputminted{r}{num_two.R}
\begin{minipage}{0.5\textwidth}
\includegraphics[scale=0.4]{02data_plot}
\end{minipage}
\begin{minipage}{0.5\textwidth}
\includegraphics[scale=0.4]{02seasonal}
\end{minipage}
\end{quote}

\newpage
\item 
\end{enumerate}
\end{document}

### 2016-01-18 math_modeling/hw/hw1/num_one.R

@@ -1,5 +1,10 @@
+#!/usr/bin/RScript
 if(require("forecast")){
-	data <- scan("hmwk1.txt")
-	plot(data)
-	plot(ma(data,4))
+    data <- scan("hmwk1.txt")        #read in the file
+    png("01data_plot.png")           #plot to specified png
+    plot(data)                       #plot data
+    dev.off()
+    png("01data_smoothed.png")       #plot to specified png
+    plot(ma(data,4))                 #plot smoothed data
+    dev.off()
 }

### 2016-01-18 math_modeling/hw/hw1/num_two.R

@@ -1,0 +1,12 @@
+#!/usr/bin/RScript
+if(require("forecast")){
+    data <- scan("co2.txt")
+    png("02data_plot.png")
+    plot(data)
+    dev.off()
+    series <- ts(data, start=c(2000,1), end=c(2016,13), frequency=13)
+    png("02seasonal.png")
+    seasonal <- stl(series, s.window="period")
+    plot(seasonal)
+    dev.off()
+}

### 2016-01-19 math_modeling/hw/hw1/result.tex
so much work

@@
 \newpage
-\item 
+\item The following table gives the depth $Z$ of water in feet for surface points with rectangular
+coordinates $X, Y$ in meters. The depth of measurements were made at low tide. Your ship has a draft of 5
+feet. What region should you avoid within the rectangle $[75, 200] \times [−50, 150]$?
+\begin{center}
+\begin{tabular}{c | c | c}
+X & Y & Z\\
+\hline
+129.0 & 7.5 & 4\\
+140.0 & 141.5 & 8\\
+108.5 & 28.0 & 6\\
+88.0 & 147.0 & 8\\
+185.5 & 22.5 & 6\\
+195.0 & 137.5 & 8\\
+105.5 & 85.5 & 8\\
+157.5 & -6.5 & 9\\
+107.5 & -81.0 & 9\\
+77.0 & 3.0 & 8\\
+162.0 & -66.5 & 9\\
+162.0 & 84.0 & 4\\
+117.5 & -38.5 & 9
+\end{tabular}
+\end{center}
+
+\newpage
+\item My full name is James Elder Martin - numerically this corresponds to the vector 10, 1, 13, 5, 19, 5,
+12, 4, 5, 18, 13, 1, 18, 20, 9, 14. Type the numerical representation of your own full name into R as a
+single vector. Then issue the command, source(``batch.R"), and then (instructions for next calling batch.R
+are detailed in the comments of the file, batch.R) produce a vector containing batch averages, using a
+batch size, that is, a k, of 3.
+\begin{quote}
+\inputminted{r}{num_three.R}
+\end{quote}
 \end{enumerate}

### 2016-01-19 vector_calc/homework/hw3.tex
so much work

@@
 
-\item Is there ever a case where $proj_{a}b = proj_{b}a$?
+\item Is there ever a case where $proj_{a}b = proj_{b}a$? If so, under what conditions?
 \begin{quote}
+If we use the explicit formulas equal to each other, we have
+\begin{align*}
+\left({a\cdot b \over a\cdot a}\right)\vv{a} &= \left({a\cdot b \over b\cdot b}\right)\vv{b}\\
+{||a|| \over ||b||}\vv{a} &= \vv{b}
+\end{align*}
+Therefore, the projections are equal only when $\vv{a} = \vv{b}$.
+\end{quote}
 
+\item Prove part 1 of Proposition 3.4.
+\begin{quote}
+We need to show that $||k\vv{a}|| = |k|\ ||\vv{a}||$. Then
+\begin{align*}
+||k\vv{a}|| &= ||(ka_{1},ka_{2},\dots,ka_{n})||\\
+&= \sqrt{k^{2}a_{1}^{2} + k^{2}a_{2}^{2} + \cdots + k^{2}a_{n}^{2}}\\
+&= \sqrt{k^{2}(a_{1}^{2} + a_{2}^{2} + \cdots + a_{n}^{2})}\\
+&= \sqrt{k^{2}}\ \sqrt{a_{1}^{2} + a_{2}^{2} + \cdots + a_{n}^{2}}\\
+&= |k|\ ||\vv{a}||.
+\end{align*}
+$\triangle$
+\end{quote}
+
+\item In physics, when a constant force acts on an object as the object is displaced, the work done by
+the force is the product of the length of the displacement and the component of the force in the
+direction of the displacement. Figure 1.48 depicts an object acted upon by a constant force $F$, which
+displaces it from the point $P$ to the point $Q$. Let $\theta$ denote the angle between $F$ and the
+direction of
+displacement.
+\begin{enumerate}
+\item Show that the work done by $F$ is determined by the formula $F\cdot\vv{PQ}$.
+\begin{quote}
+We can determine the $x$ component of $F$ as $||F||\cos{\theta}$. Then the work can be computed as
+$||F||\ ||\vv{PQ}||\cos{\theta}$, which is $F\cdot\vv{PQ}$ by definition.
+\end{quote}
+\item Find the work done by the (constant) force $F = i + 5j + 2k$ in moving a particle from the point
+$(1, -1, 1)$ to the point $(2, 0, -1)$.
+\begin{quote}
+We need to find the dot product $(1,5,2)\cdot(1,1,-2)$, which is equal to $2$. Therefore, the work done
+is 2 units.
+\end{quote}
+\end{enumerate}
+
+\item Using vectors, prove that the diagonals of a parallelogram are perpendicular if and only if the
+parallelogram is a rhombus. (Note: A rhombus is a parallelogram whose four sides all have the same
+length).
+\begin{quote}
+We need to show that the vectors $\vv{ab}$ and $\vv{ba}$ have a dot product of 0. We know that the
+lengths of $\vv{a}$ and $\vv{b}$ are equal, and that $\vv{ab} = (a_{1} - b_{1}, a_{2} - b_{2})$ and
+$\vv{ba} = (b_{1} + a_{1}, b_{2} + a_{2})$. Therefore
+\begin{align*}
+\vv{ab}\cdot\vv{ba} &= (a_{1} - b_{1})(b_{1} + a_{1}) + (a_{2} - b_{2})(b_{2} + a_{2})\\
+&= a_{1}^{2}+a_{2}^{2}-b_{1}^{2}-b_{2}^{2}\\
+&= b_{1}^{2}+b_{2}^{2}-b_{1}^{2}-b_{2}^{2},\text{ Since the lengths are equal,}\\
+&= 0
+\end{align*}
 \end{quote}

### 2016-01-19 math_modeling/hw/hw1/num_three.R

@@ -1,0 +1,7 @@
+#!/usr/bin/RScript
+source("batch.R")
+lets <- c(19,20,5,22,5,14,1,12,1,14,18,15,19,5,14,4,1,8,12)
+res <- batching(3,lets)
+print(res)
+# Produces
+# [1] 14.666667 13.666667  4.666667 15.666667 12.666667  4.333333

### 2016-01-19 writing_in_ethics/paper_one.tex
so much work
editing paper

@@
 
+% \linenumbers
+
 \begin{document}
@@
 \inlinetitle
-
-The theory of Utilitarianism can be used to quantitatively analyze various situations that arise in day to day life; it can be applied the scenario presented in \textit{Ethics for the Information Age} that concerns Alexis' education.
-In order to analyze the situation and determine whether or not her actions were ethically ``good" or ``bad", the total utility change needs to be calculated.
-Alexis' actions can be considered ethically good if the benefit of her actions exceed the harms.
-The situation can be analyzed using both Rule and Act Utilitarianism.
+\begin{linenumbers}
+The theory of Utilitarianism can be used to quantitatively analyze various situations that arise in day to day life.
+It can be used to analyze the scenario presented in \textit{Ethics for the Information Age} in which a promising high school student, Alexis, steals login information in order to use a computer to which she does not have access.
+Alexis' situation can be analyzed under both Rule and Act Utilitarianism, which both require a numerical value to be computed based on the actions of an individual.
+In Alexis' case, there is no clear unit that can be assigned to the value of her actions; rather, the overall general outcome can be used to gauge the utility of Alexis' actions.
 
@@
 Both of the methods used to analyze the situation provide a method to determine whether or not Alexis made the right choice; nether Rule nor Act Utilitarianism are absolute.
-
+\end{linenumbers}
 \newpage

### 2016-01-20 math_modeling/class/notes/smoothing.tex
math modeling

\documentclass{hw}

\begin{document}
\section{Exponential Smoothing}
\begin{itemize}
\item Used to predict future data trends: For example, weather forecasting
\[
S_{i} = \alpha x_{i} + (1-\alpha)S_{i-1},
\]
Where $\alpha$ and $1-\alpha$ remove the outlying data. $x_{i}$ is the actual data. This is a
recurrence relation, which can be represented by
\begin{align*}
S_{i} = \alpha(x_{i}+(1-\alpha)x_{i-1} + (1-\alpha)^{2}x_{i-2} + \dots)\\
\end{align*}
As the terms increase, the $1-\alpha$ term begins to decrease.
\end{itemize}
\end{document}

### 2016-01-20 math_modeling/hw/hw1/batch.R

@@ -1,14 +1,15 @@
-#	function name: batching
-#	parameters:
-#	    k   size of each batch
-#           x   data
-#       return value:
-#           an array of the averages of each batch
-# 	description: this function separates the data x into
-#       floor(length(x)/k) batches and returns the array of
-#       the averages of each batch
-#	To run the code, type y<-batching(k,x) to place the consecutive
-#       batch means of size k, of the data in the vector x, in the vector y
+# function name: batching
+# parameters:
+#   k   size of each batch
+#   x   data
+#
+# return value:
+#   an array of the averages of each batch
+#
+# description: this function separates the data x into
+#   floor(length(x)/k) batches and returns the array of the averages of each batch
+#   To run the code, type y<-batching(k,x) to place the consecutive batch means of size k, of the data
+#   in the vector x, in the vector y
 batching <- function(k,x)  {
     m <- floor(length(x)/k);
     result <- rep(0,m);

### 2016-01-20 math_modeling/hw/hw1/lakebedm.m

@@ -1,0 +1,8 @@
+plane = 0.004802 y + 0.007480 x + 8.554271 z;
+Print[
+    FindInstance[
+        plane > 5 && -50 <= x && x <= 75 && 150 <= y && y <= 200,
+		{x,y,z},
+		Reals
+    ]
+]

### 2016-01-20 math_modeling/hw/hw1/lakebedr.R

@@ -1,0 +1,3 @@
+data <- read.table('lakebed.txt')
+ml <- lm(data[,3] ~ data[,2] + data[,1])
+print(ml)

### 2016-01-21 algorithms/analysis.tex
added homework for algo

\documentclass{hw}

\usepackage{minted}
\usepackage{mathtools}
% \DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
% \DeclarePairedDelimiter\ceil{\lceil}{\rceil}

\newcommand{\ceil}[1]{\left\lceil #1\right\rceil}

\begin{document}
\makeheader{1}
\noindent\\

Derive the order of magnitude worst-case running time of the following algorithm, and give a clear
explanation of how you arrived at this bound, in the same manner as the CLRS insertion-sort example
from chapter 2 which we did this in class. As I did in class, you can leave out the constants
$c_{i}$ for each line of code; but, where appropriate, you must analyze loops by setting up a sum
over the loop iterations.
Here swap($A, i, j$) means interchange array elements $A[i]$, $A[j]$. Note that swap is not one of
our basic constant-time operations; how will you deal with this?\\

\begin{minted}[linenos]{ruby}
i = 1
j = N
while i < j
    for k = i to j - 1
        if A[k] > A[k + 1]
            swap(A, k, k + 1)
    for k = j - 1 to i + 1 by -1
        if A[k - 1]>A[k]
            swap(A, k - 1, k)
    i = i + 1
    j = j - 1
\end{minted}

We know that the while loop will execute until $i=j$. This means that the sum will execute at most
$\ceil{N \over 2}$. We can analyze both the inner for loops individually in order to formulate a
sum that is representative of the number of instructions for that loop. The first loop executes
$N-1$ times on the first iteration, then $N-3$ times on the second iteration, and so on. This
yields the sum
\[
\sum_{i=0}^{N} N-(2i+1).
\]
Similarly, the second for loop will execute almost the exact same number of times; however, only
the even $i$ terms will be taken into account. As such, we have that the second loop will execute
\[
\sum_{i=0}^{N} N-2i
\]
times. Assuming the worst case scenario, we can say that the comparison in the if statement of both
loops will return true every iteration. As a result, the swap function will be executed on every
iteration. We don't actually know the execution time of the swap function, so we can assign it the
value $\mathcal{S}_{t}$. Combining both of the sums we already have and multiplying them by the
swap time yields
\[
\mathcal{S}_{t}\sum_{i=0}^{N}2N-4i-1.
\]
The only other operations in the while loop are constant time, so we will ignore those.\\

We now need to express the number of times the while loop executes as a sum. We know that every
time the loop executes, we increment $i$ and decrement $j$. As a result, we know that the loop will
only execute $\ceil{N \over 2}$ times. We also know that our $i$ starts at 1. Combining the sums
together gives us
\[
\mathcal{S}_{t}\sum_{i=1}^{\ceil{N \over 2}}\sum_{i=0}^{N}2N-4i-1.
\]
Of course, we don't want to have a sum of sums. Through algebraic manipulation, we have that
\[
\sum_{i=0}^{N}2N-4i-1 = -(1-n),
\]
so we are now dealing with
\[
\mathcal{S}_{t}\sum_{i=1}^{\ceil{N \over 2}}-(1-N).
\]
We have two cases to deal with now: when $N$ is even and when $N$ is odd. If $N$ is even, then we
are dealing with
\[
\mathcal{S}_{t}\sum_{i=1}^{N \over 2}-(1-N).
\]
This can be simplified to
\[
{1\over2}(N-1)N = {N^{2} - N \over 2}
\]
which is $O(N^{2})$. If $N$ is odd, then we have
\[
\mathcal{S}_{t}\sum_{i=1}^{N + 1 \over 2}-(1-N),
\]
which is equivalent to
\[
{N^{2} - 1 \over 2},
\]
and that is still $O(N^{2})$.
\end{document}

### 2016-01-21 math_modeling/hw/hw1/result.tex
math modeling
finished modeling

@@
 \begin{enumerate}
-\item Dr. Martin has emailed you the data set, hmwk1.txt, which gives the time evolution of a certain
-quantity of interest to him. Using R, plot the original data. In addition, using R’s ma() function, plot a
-smoothed version of the data. What attribute of the original data is hidden by referring only to the
-smoothed data?
+\item Dr. Martin has emailed you the data set, hmwk1.txt, which gives the time evolution of a
+certain quantity of interest to him. Using R, plot the original data. In addition, using R’s ma()
+function, plot a smoothed version of the data. What attribute of the original data is hidden by
+referring only to the smoothed data?
 \begin{quote}
@@
 \newpage
-\item Dr. Martin has emailed to you the data set, co2.txt. The data represents 16 years of collecting
-monthly $CO_{2}$ data on the island of Hawaii, with the first year of $CO_{2}$ starting in January of
-1958. a) Using R, plot the data. b) Using R’s stl() command, plot the 1. trend, 2. seasonal, and 3.
-irregular components of the data.
+\item Dr. Martin has emailed to you the data set, co2.txt. The data represents 16 years of
+collecting monthly $CO_{2}$ data on the island of Hawaii, with the first year of $CO_{2}$ starting
+in January of 1958. a) Using R, plot the data. b) Using R’s stl() command, plot the 1. trend, 2.
+seasonal, and 3. irregular components of the data.
 \begin{quote}
@@
 \newpage
+%% data <- Read.table('myfile.txt')
+%% lm(data[,2]~data[,2]+data[,1]) will return equation of lakebed plane
+%% Use Mathematica to predict where the depth is less than 5 on lakebed plane
+%% The equation needs to be such that plane.z < 5
 \item The following table gives the depth $Z$ of water in feet for surface points with rectangular
 coordinates $X, Y$ in meters. The depth of measurements were made at low tide. Your ship has a draft of 5
-feet. What region should you avoid within the rectangle $[75, 200] \times [−50, 150]$?
+feet. What region should you avoid within the rectangle $[75, 200] \times [-50, 150]$?
 \begin{center}
@@
 
+\inputminted{r}{lakebedr.R}
+\includegraphics[scale=0.7]{three_plot}
+\noindent\\
+The entire area appears to be safe. However, the plane is the result of a best fit, so the corner
+of the lakebed should be avoided just to be sure.
+
 \newpage
-\item My full name is James Elder Martin - numerically this corresponds to the vector 10, 1, 13, 5, 19, 5,
-12, 4, 5, 18, 13, 1, 18, 20, 9, 14. Type the numerical representation of your own full name into R as a
-single vector. Then issue the command, source(``batch.R"), and then (instructions for next calling batch.R
-are detailed in the comments of the file, batch.R) produce a vector containing batch averages, using a
-batch size, that is, a k, of 3.
+\item My full name is James Elder Martin - numerically this corresponds to the vector 10, 1, 13, 5,
+19, 5, 12, 4, 5, 18, 13, 1, 18, 20, 9, 14. Type the numerical representation of your own full name
+into R as a single vector. Then issue the command, source(``batch.R"), and then (instructions for
+next calling batch.R are detailed in the comments of the file, batch.R) produce a vector containing
+batch averages, using a batch size, that is, a k, of 3.
 \begin{quote}

### 2016-01-21 math_modeling/hw/hw1/lakebedm.m

@@ -1,7 +1,7 @@
-plane = 0.004802 y + 0.007480 x + 8.554271 z;
+plane = z == 8.554271 + 0.004802 y + 0.007480 x;
 Print[
     FindInstance[
-        plane > 5 && -50 <= x && x <= 75 && 150 <= y && y <= 200,
+        plane && z > 5 && -50 <= x && x <= 75 && 150 <= y && y <= 200,
 		{x,y,z},
 		Reals
     ]

### 2016-01-25 math_modeling/class/notes/ch1.tex
so much work

\documentclass{hw}

\begin{document}
\section{Modeling Free Fall}
\noindent\\

\begin{enumerate}
\item Measure the height of the object above ground
\item Record time in seconds that it takes for object to hit the ground
\item Determine the quantitative relationship between the height and the time (a mathematical
model)
\end{enumerate}
Galileo makes some assumptions before experimenting:
\begin{enumerate}
\item $F = ma$: we already know the mathematical expression for force
\item Gravity is the only force acting on the object
\item Acceleration due to gravity is -32${m\over s^{2}}$.
\begin{gather*}
mg = my''\\
g = y''\\
\iint g = \iint y''\\
-16t^{2} = y + Ct + D\\
y = -16t^{2} + v_{0}t + y_{0}
\end{gather*}
\end{enumerate}
Galileo assumed that gravity was the only force, but what if air resistance is added? We know that
air resistance acts opposite to the velocity. We now have
\begin{gather*}
\kappa v - 32m = mv'\\
v' = 32 - \kappa_{0} v
\end{gather*}

\section{Discreet Modeling}
\noindent\\

We can now consider a discreet case for modeling.
\subsection{Credit Card Balancing}
We can work with the equation
\[
B_{1} = B_{0} - P
\]
as a simple model for a credit card balance. We will eventually find that this model is very
simplistic, and we will need to expand on it later. We have a recurrence relation where
\begin{gather*}
B_{2} = B_{1} - P\\
B_{3} = B_{2} - P\\
\cdots
\end{gather*}
Substituting into the recurrence relation tells us that
\[
B_{n} = B_{0} - nP.
\]
We can reform the model to help use come up with a better representation of the situation.
In the real world, credit cards often come with interest rates. Taking this into account
gives us
\[
B_{\text{new}} = (1+r)B_{\text{old}} - P.
\]
We again have a recurrence relation. We will let $s = (1+r)$:
\begin{align*}
B_{1} &= sB_{0} - P\\
B_{2} &= sB_{1} - P = s^{2}B_{0}-p(1+s)\\
B_{3} &= sB_{2} - P = s^{3}B_{0}-p(1+s+s^{2})\\
B_{4} &= sB_{3} - P = s^{4}B_{0}-p(1+s+s^{2}+s^{3})\\
\cdots
\end{align*}
We can derive a general solution of the form
\begin{align*}
B_{n} &= s^{n}B_{0}-p(1 + s + s^{2} + s^{3} + \cdots + s^{n-1})\\
&= (1+r)^{n}B_{0}-p\left({(1+r)^{n} - 1\over r}\right)
\end{align*}
\subsubsection{Solving With Excel}
We can use Excel to represent the model. We will consider the case where $r=1.5\%$ with an initial
balance of $\$1000$, and $\$10$ per month on payments. In the A1 cell, we want to have
\begin{minted}{text}
=(1.015)^A1 * 1000 - 10*((1.015)^A1-1)/0.015
\end{minted}
\end{document}

### 2016-01-25 number_theory/hw/hw2.tex
so much work

\documentclass{hw}

\begin{document}
\makeheader{2}

\begin{enumerate}
\item Calculate $gcd(5, 7)$ and find two pairs of integers $(u, v)$ such that $gcd(5, 7) = 5u + 7v$.
\begin{quote}
By the definition of the $gcd$, we have that $gcd(5,7) = 1$.
\begin{gather*}
5(3) + 7(-2) = 1\\
5(10) + 7(-7) = 1
\end{gather*}
\end{quote}

\item Show that $c|a$ and $c|b$ if and only if $c | gcd(a,b)$.
\begin{quote}

Assume $c|a$ and $c|b$. Then $c$ is a common factor of $a$ and $b$. We know that $gcd(a,b)$ is also
a common factor of both $a$ and $b$. Since both $gcd(a,b)$ and $c$ are common factors of $a$ and
$b$, then they must also be common factors of each other. Therefore, $c|gcd(a,b)$.\\

Let $d = gcb(a,b)$, and assume that $c|d$. Then $d = cn$ for some $n\in\ZZ$. By the definition of
$gcd$, we know that $d|a$. Therefore, $cn|a$, which means that $a = cnm$ for some $m\in\ZZ$. Then
$a = cj$, where $j = nm\in\ZZ$. Therefore $c|a$. We also know that $d|b$ by the definition of
$gcd$. By the same argument, we know that $c|b$ as well. Therefore $c|b$ and $c|a$.
\end{quote}
$\triangle$

\item Suppose $a_{1},a_{2},\dots,a_{n}$ are integers not all equal to 0. We define
$gcd(a_{1},a_{2},...,a_{n})$ to be the largest integer which divides $a_{k}$ for all $1 \leq k \leq
n$. Prove that $gcd(a_{1},a_{2},\dots,a_{n}) = gcd(gcd(a_{1},a_{2}),a_{3},...,a_{n})$.
\begin{quote}
Let $gcd(a_{1},a_{2},\dots,a_{n}) = d$ and $gcd(gcd(a_{1},a_{2}),a_{3},\dots,a_{n}) = \alpha$. We
know that, by the definition of $gcd$, that $d|a_{k}$ for $1 \leq k \leq n$. We know that $d|a_{1}$
and $d|a_{2}$, so it must also divide $gcd(a_{1},a_{2})$. Therefore, $d|gcd(gcd(a_{1},a_{2})
,a_{3},\dots,a_{n})$, so $d|\alpha$. We also know that $\alpha|gcd(a_{1},a_{2})$, and that
$\alpha|a_{k}$ for $2 < k \leq n$. Since $\alpha$ divides $gcd(a_{1},a_{2})$, it must divide
$a_{1}$ and $a_{2}$. Therefore $\alpha|a_{k}$ for $1 \leq k \leq n$, so
$\alpha|gcd(a_{1},a_{2},\dots,a_{n})$, or $\alpha|d$. We know that if $d|\alpha$ and
$\alpha|d$, then $|d| = |\alpha|$. In this case, both $\alpha$ and $d$ are positive,
so we know that $\alpha = d$.
\end{quote}
$\triangle$

\item Use your answer to the previous problem, along with the Euclidian Algorithm, to determine
$gcd(1092,1155,2002)$ (It's possible that you'll need a calculator to do the arithmetic on this
problem).
\begin{quote}

\end{quote}
\end{enumerate}
\end{document}

### 2016-01-25 vector_calc/homework/hw6.tex
so much work

\documentclass{hw}

\begin{document}
\makeheader{6}

\begin{enumerate}
\item Find the Cartesian coordinates of the point whose polar coordinates are given.
\begin{quote}
\begin{align*}
(\sqrt{2},{\pi\over4}) &= (\sqrt{2}\cos{\left({\pi\over4}\right)},
\sqrt{2}\sin{\left({\pi\over4}\right)})\\
&= (1,1)
\end{align*}
\end{quote}

\item Give a set of polar coordinates for the point whose Cartesian coordinates are given.
\begin{quote}
\begin{align*}
(-2,2) &= (\sqrt{(-2)^{2} + 2^{2}}, \arctan{(-1)})\\
&= (\sqrt{4},-{\pi\over4})\\
&= (2,-{\pi\over4})
\end{align*}
\end{quote}

\item Find the Cartesian coordinates of the point whose cylindrical coordinates are given.
\begin{quote}
\begin{align*}
(1,{2\pi\over3},-2) &=
(\cos{\left({2\pi\over3}\right)},\sin{\left({2\pi\over3}\right)},-2)\\
&= (-{1\over2},{\sqrt{3}\over2},-2)
\end{align*}
\end{quote}

\item Find the rectangular coordinates of the point whose spherical coordinates are given.
\begin{quote}
\begin{align*}
%rho phi theta
(1,{3\pi\over4},{2\pi\over3}) &=
(\cos{{3\pi\over4}}\sin{{2\pi\over3}},\sin{{3\pi\over4}}\sin{{2\pi\over3}},\cos{{3\pi\over4}})\\
&= ({-\sqrt{3}/2\over2},{\sqrt{3}/2\over2},{-1\over\sqrt{2}})
\end{align*}
\end{quote}

\item Find a set of cylindrical coordinates of the point whose Cartesian coordinate is given.
\begin{quote}
\begin{align*}
(-1,\sqrt{3},13) &= (\sqrt{(-1)^2 + (\sqrt{3})^2},\arctan{-{1\over\sqrt{3}}},13)\\
&= (2,-{\pi\over6},13)
\end{align*}
\end{quote}
\end{enumerate}
\end{document}

### 2016-01-25 math_modeling/hw/hw1/lakebedm.m

@@ -1,8 +1,2 @@
-plane = z == 8.554271 + 0.004802 y + 0.007480 x;
-Print[
-    FindInstance[
-        plane && z > 5 && -50 <= x && x <= 75 && 150 <= y && y <= 200,
-		{x,y,z},
-		Reals
-    ]
-]
+(*z == 8.554271 - 0.004802 y - 0.007480 x;*)
+Plot3D[8.554271 - 0.004802 y - 0.007480 x, {x,75,200}, {y,-50,150}]

### 2016-01-25 math_modeling/hw/hw1/lakebedr.R

@@ -1,3 +1,4 @@
+#!/usr/bin/RScript
 data <- read.table('lakebed.txt')
 ml <- lm(data[,3] ~ data[,2] + data[,1])
 print(ml)

### 2016-01-25 number_theory/hw/hw2.tex
number theory

@@
 problem).
+\begin{align*}
+1155 &= 1(1092) + 63\\
+1092 &= 17(63) + 21\\
+63 &= 3(21) + 0\\
+\therefore\gcd{1155}{1092} &= \gcd{21}{0} = 21\\
+2002 &= 95(21) + 7\\
+21 &= 3(7) + 0\\
+\therefore\gcd{21}{2002} &= \gcd{7}{0} = 7
+\end{align*}
+
+\item Let $a_{1},a_{2},\dots,a_{n} \in \NN$ and consider the two following definitions:
+\begin{itemize}
+\item We say that $a_{1},a_{2},\dots,a_{n}$ are relatively prime if
+$gcd(a_{1},a_{2},\dots,a_{n}) = 1$.
+\item We say that $a_{1},a_{2},\dots,a_{n}$ are pairwise relatively prime if
+$gcd(a_{i},a_{j}) = 1$ for all $i\neq j$.
+\end{itemize}
+\begin{enumerate}
+\item If $a_{1}, a_{2},\dots,a_{n}$ are pairwise relatively prime can we conclude that
+$a_{1}, a_{2},\dots,a_{n}$ are relatively prime? Either prove your answer or give a counterexample.
 \begin{quote}
+Let $(a_{n}) = (a_{1},a_{2},\dots,a_{n})$, and assume $(a_{n})$ is pairwise relatively prime.
+Then for any $i,j \leq n$ we have that $\gcd{a_{i},a_{j}} = 1$. This means that if we take any
+to distinct elements in $(a_{n})$, they will have no common factors except for 1. As a result, the
+greatest common factor is 1, or $gcd((a_{n})) = 1$, which is the definition of begin relatively
+prime.
+\end{quote}
+$\triangle$
+\item If $a_{1}, a_{2},\dots,a_{n}$ are relatively prime can we conclude that $a_{1},
+a_{2},\dots,a_{n}$ are pairwise relatively prime? Either prove your answer or give a
+counterexample.
+\begin{quote}
+Consider $(a_{n}) = (2,3,4)$. We know that $gcd((a_{n})) = 1$, but $gcd(2,4)\neq 1$.
+Therefore, we can conclude that $(a_{n})$ being relatively prime does not necessarily
+mean that $(a_{n})$ is pairwise relatively prime.
+\end{quote}
+$\triangle$
+\end{enumerate}
 
+\item Suppose that $a | c$ and $b | c$. Do we necessarily have that $ab | c$? Either prove your
+answer or give a counterexample.
+\begin{quote}
+Let $a,b = 4$ and $c = 8$. We know that $4|8$, but $(4\cdot4)\cancel{|}8$. Therefore $ab$ does
+not divide $c$.
 \end{quote}
+$\triangle$
+
+\item If $a|c$ and $b|c$ prove that $\lcm{a}{b}|c$. Conclude that if $a$ and $b$ are relatively
+prime, then $ab|c$.
+\begin{quote}
+We know that since $a|c$ and $b|c$, $c$ is a common factor of both $a$ and $b$. If we assume
+that $c$ is the smallest common factor of both $a$ and $b$, then $c = \lcm{a}{b}$, and $c|c$.
+Otherwise, let $\delta = \lcm{a}{b}$. Then, by the definition of the $lcm$, we know that
+$\delta$ is a common factor of $a$ and $b$, and that $\delta < c$. Since both $c$ and $\delta$
+are common factors of $a$ and $b$, it follows that $\delta|c$.\\
+
+Since $a$ and $b$ are relatively prime, $\gcd{a}{b} = 1$. We also know that $ab =
+\lcm{a}{b}\gcd{a}{b}$, which means that $ab = \lcm{a}{b}$. Therefore, $ab|c$.
+\end{quote}
+$\triangle$
 \end{enumerate}

### 2016-01-31 algorithms/orders.tex
so much homework done

\documentclass{hw}

\begin{document}
\makeheader{2}

\begin{enumerate}
\item Order these functions from slowest to fastest order of growth, indicating which are the same order.
Logarithms are all base 2.
\begin{center}
\begin{tabular}{l l}
$5n^3-n^2+n$ & $O(n^3)$\\
6 & $O(1)$\\
$2^{n-1}$ & $O(2^{n})$\\
$log(log(n))$ & $O(log(n))$\\
$2^{n}$ & $O(2^n)$\\
$3^{n}$ & $O(3^n)$\\
$(log(n))^2$ & $O(n\ log(n))$\\
$n\ log(n)$ & $O(n\ log(n))$
\end{tabular}
\end{center}
\[
6 \leq log(log(n)) \leq (log(n))^{2} \leq n\ log(n) \leq 5n^3-n^2+n \leq 2^{n-1}\text{ and } 2^n \leq
3^{n}
\]

\item Find two constants $c$, $c'$ such that, for large enough $n$, we have
\[ c\ log(n) \leq log(2n^2) \leq c'\ log(n) \]
\begin{quote}
We can simplify the middle term so that the $n$ is linear for each $log$.
\begin{gather*}
c\ log(n) \leq 2log(2n) \leq c'\ log(n)\\
c\ log(n) \leq 2(log(n) + log(2)) \leq c'\ log(n)\\
c\ log(n) \leq 2 + 2log(n) \leq c'\ log(n)
\end{gather*}
If we let $c = 1$ and $c' = 4$, then for large enough $n$, the inequality will hold.
\end{quote}

\item How large must $n$ be to have $n > 13\sqrt{n}\ $?
\begin{quote}
We can solve the inequality for $n$.
\begin{gather*}
n > 13\sqrt{n}\\
n - 13\sqrt{n} > 0\\
\sqrt{n}\ (\sqrt{n} - 13) > 0
\end{gather*}
We know that $n > 0$ is not a valid, since $n = 1$ breaks the inequality. However, we also have that
$\sqrt{n} - 13 > 0$, which tells us that $n$ must be greater than $13^2$.
\end{quote}
\end{enumerate}
\end{document}

### 2016-01-31 math_modeling/class/notes/classnotes.tex
so much homework done

--- 2016-01-25 math_modeling/class/notes/ch1.tex
@@
 \documentclass{hw}
+
+\usepackage{minted}
 
@@
 \begin{minted}{text}
-=(1.015)^A1 * 1000 - 10*((1.015)^A1-1)/0.015
+=(1.015)^$A1 * 1000 - 10*((1.015)^$A1-1)/0.015
 \end{minted}
+
+\section{Modeling Ordinary Differential Equation}
+Consider the following system of ODE's
+\[
+\left[
+\begin{array}{c}
+Q_{1}^\prime\\
+Q_{2}^\prime\\
+Q_{3}^\prime
+\end{array}
+\right]
+\]
+We will use arbitrary values for the coefficient matrix. We want to solve the system using
+eigenvalues:
+\[
+\left[
+\begin{array}{c c c}
+5-\lambda & 4 & 2\\
+4 & 5-\lambda & 2\\
+2 & 2 & 2-\lambda
+\end{array}
+\right]
+= 0
+\]
+We want to take the determinant of the matrix, but first we can do some row reductions:
+\[
+\left[
+\begin{array}{c c c}
+1-\lambda & 0 & 0\\
+4 & 9-\lambda & 2\\
+2 & 4 & 2-\lambda
+\end{array}
+\right]
+\]
+Taking that determinant gives us
+\[
+\lambda = 1 \qquad\qquad\text{and}\qquad\qquad \lambda = 10.
+\]
+
+\subsection{Modeling War}
+We will model the Nazi party and the Soviet Union military forces. We want to model the rate
+at which Soviet Tanks reduce as Nazi anti-tank guns decrease.
+\begin{gather*}
+\frac{dx}{dt} = -ay\\
+\frac{dy}{dt} = -bx
+\end{gather*}
+In this system, $x'$ represents the rate at which the Soviet tanks decreased, and $y'$ represents
+the rate at which German anti-tank guns decrease. We will consider the battle of Kursk, during
+which the number of Soviet tanks decreased by $50\%$ in the first hour. In this model, $a$ is
+referred to as the anti-tank kill rate, where $b$ is the tank kill rate. We can use separation
+of variables yields
+\begin{gather*}
+{ay^2\over2}+{ay_{0}^{2}\over2} = {bx^2\over2}+{bx_{0}^{2}\over2}\\
+ay^{2}-bx^{2}=C
+\end{gather*}
 \end{document}

### 2016-01-31 vector_calc/homework/hw6.tex
so much homework done

@@
 \begin{document}
-\makeheader{6}
+\makeheader{1.7}
 
@@
 \end{quote}
+
+\item Convert the provided equation into the other two coordinate systems.
+\[
+\rho\sin{\phi}\sin{\theta} = 2
+\]
+\begin{quote}
+In cylindrical coordinates: \[r\sin{\theta} = 2\]
+In cartesian coordinates: \[y=2\]
+\begin{gather*}
+\end{gather*}
+\end{quote}
+
+\item Convert the provided equation into the other two coordinate systems.
+\[
+z^{2} = 2x^{2} + 2y^{2}
+\]
+\begin{quote}
+In cylindrical coordinates:
+\begin{gather*}
+z^{2} = r^{2}\cos^{2}{\theta} + r^{2}\sin^{2}{\theta}\\
+z^{2} = r^{2}
+\end{gather*}
+In spherical coordinates:
+\begin{gather*}
+\rho^{2}\cos^{2}{\phi} = \rho^{2}\sin^{2}{\phi}\cos^{2}{\theta} +
+\rho^{2}\sin^{2}{\phi}\sin^{2}{\theta}\\
+\cos^{2}{\phi} = \sin^{2}{\phi}(\cos^{2}{\theta} + \sin^{2}{\theta})\\
+\cos^{2}{\phi} = \sin^{2}{\phi}
+\end{gather*}
+\end{quote}
+
+\item Sketch the solid whose cylindrical coordinates $(r,\theta,z)$ satisfy the given
+inequalities.
+\[
+0\leq r\leq3,\ 0\leq\theta\leq{\pi\over2},\ -1\leq z\leq2
+\]
+\vspace{3cm}
+
+\item Sketch the solid whose cylindrical coordinates $(r,\theta,z)$ satisfy the given
+inequalities.
+\[
+r\leq z\leq 5,\ 0\leq\theta\leq\pi
+\]
+\vspace{3cm}
+
+\item Sketch the solid whose spherical coordinates satisfy the given inequalities.
+\[
+0\leq\rho\leq1,\ 0\leq\phi\leq{\pi\over2}
+\]
+\vspace{3cm}
+
+\item Sketch the solid whose spherical coordinates satisfy the given inequalities.
+\[
+0\leq\phi\leq{\pi\over4},\ 0\leq\rho\leq2
+\]
+\vspace{3cm}
+
+\item Consider the solid shown below:
+\begin{center}
+\includegraphics[scale=0.4]{hw6_pic}
+\end{center}
+\begin{enumerate}
+\item Describe the solid, using spherical coordinates.
+\[
+0\leq\rho\leq3,\ 0\leq\phi\leq\arctan{1\over\sqrt{8}},\ 0\leq\theta\leq2\pi
+\]
+\item Describe the solid, using cylindrical coordinates.
+\[
+\sqrt{8}\ r\leq z\leq\sqrt{9-r^{2}},\ 0\leq r\leq 1,\ 0\leq\theta\leq2\pi
+\]
+\end{enumerate}
 \end{enumerate}

### 2016-01-31 writing_in_ethics/paper_one.tex
so much homework done

@@
 \begin{document}
-\drafth{}
+\finalh{}
 \inlinetitle
 \begin{linenumbers}
+
 The theory of Utilitarianism can be used to quantitatively analyze various situations that arise in day to day life.
-It can be used to analyze the scenario presented in \textit{Ethics for the Information Age} in which a promising high school student, Alexis, steals login information in order to use a computer to which she does not have access.
+One can analyze the scenario presented in \textit{Ethics for the Information Age} in which a promising high school student, Alexis, steals login information in order to use a computer to which she does not have access.
 Alexis' situation can be analyzed under both Rule and Act Utilitarianism, which both require a numerical value to be computed based on the actions of an individual.
-In Alexis' case, there is no clear unit that can be assigned to the value of her actions; rather, the overall general outcome can be used to gauge the utility of Alexis' actions.
+In Alexis' case, there is no clear unit of measurement that can be assigned to the value of her actions; rather, the overall general outcome can be used to gauge the utility of Alexis' actions.
 
@@
 The magnitude of Alexis' experience can be measured in the amount of happiness she experienced.
-She was able to achieve her goal and even positively affected others around her; presumably she was able to save her family and herself a large amount of money while also bettering her own life in the process through a college education.
-Both of these aspects should be taken into account as contributing positively to the magnitude of the situation.
+She was able to achieve her goal and even positively affect others around her.
+Presumably she was able to save her family and herself a large amount of money while also bettering her own life through a college education.
+Both of these aspects should be taken into account as contributing positively to the overall utility of the situation.
 
 The next attribute, \textit{Duration}, applies greatly to Alexis' scenario.
-The duration of the experience extends far beyond the time that Alexis spent in the library on the computer; in fact, the education she received will continue to affect her for the rest of her life.
+The duration of the experience extends far beyond the time that Alexis spent in the library on the computer.
+In fact, the education she received will continue to affect her for the rest of her life.
 However, Alexis stole the information from another student at the university.
@@
 There was always a chance that Alexis' end goal could have never been met, in which case the overall unhappiness produced would have far outweighed the happiness.
-However, Alexis did ultimately achieve her goal to be accepted into college; on top of just being accepted into college, she was also provided with a full ride, which positively affected her parents as well.
+However, Alexis did ultimately achieve her goal to be accepted into college.
+In addition to just being accepted into college, she was also provided with a full ride, which positively affected her parents as well.
 As a result, the attribute of \textit{Certainty} greatly adds to the overall happiness gained from Alexis' decision.
@@
 The final attribute that Bentham proposes is concerned with the number of people affected by the action.
-Alexis' actions do in fact affect several other beings, and not all the interactions are not necessarily positive.
+Alexis' actions do in fact affect several other beings, and not all the interactions are necessarily positive.
 As mentioned before, Alexis' family benefitted greatly from the situation.
@@
 
-Bentham's methods applied to Alexis' actions under the ethical theory of Act Utilitarianism.
+Bentham's methods also apply to Alexis' actions under the ethical theory of Act Utilitarianism.
 Under Act Utilitarianism, Alexis has made an ethically good decision; the amount of overall happiness produced by her actions far outweighs the unhappiness produced.

### 2016-02-01 math_modeling/class/notes/classnotes.tex
working hard or something

@@
 \end{gather*}
+
+\section{Agents}
+A Schelling Model is used to model things like segregation. Consider the following example:
+\begin{center}
+\begin{tabular}{c c c c c c}
+x1 & x2 &    &    &    &    \\
+x3 & o1 &    & o2 &    &    \\
+x4 & x5 & o3 & o4 & o5 &    \\
+x6 & o6 &    &    & x7 & x8 \\
+   & o7 & o8 & x9 & x10 & x11 \\
+   &    & o9 & o10 & o11 & \\
+\end{tabular}
+\end{center}
+We can consider the case where, in the first iteration, all the members of the grid who do not have
+three directly adjacent neighbors leave the area. In addition, x's want to be with other x's, and
+o's want to be with other o's.
+\begin{center}
+\begin{tabular}{c c c c c c}
+   &    &    &    &    &    \\
+x3 & x6 &    & o1 & o2 &    \\
+x4 & x5 & o3 & o4 &    &    \\
+   & o6 & x2 & x1 & x7 & x8 \\
+o11& o7 & o8 & x9 & x10 & x11 \\
+   &    & o5 & o9 & o10 & \\
+\end{tabular}
+\end{center}
+We can see that as the system changes, there will be a mass exodus of both types. In the first table,
+there was only 1 agent with no neighbors of the opposite type. By the second iteration, there are now 6
+agents with no neighbors of the opposite type.
 \end{document}

### 2016-02-01 number_theory/hw/hw3.tex
so much homework done
working hard or something

--- 2016-01-17 vector_calc/homework/hw3.tex
@@
 \documentclass{hw}
+
 \begin{document}
@@
 \begin{enumerate}
-\item Compute $a\cdot b$, $||a||$ and $||b||$ where $\vv{a} = 4i -3j + k$ and $\vv{b}=i+j+k$.
+\item Suppose that $a,b\in\ZZ$ and define the set $J={ax+by|x,y\in\ZZ}$. Prove that $J=Z$ if and only if
+$gcd(a, b) = 1$.
+
+\item We define the \textit{Fibonacci Sequence} to be the sequence of integers ${x_{0}, x_{1}, x_{2},\dots}$
+satisfying the properties
+\[
+x_{0} = 0,\quad x_{1} = 1,\quad\text{and}\quad x_{n} = x_{n-1}+x_{n}\text{ for all } n\geq2.
+\]
+Prove that $gcd(x_{n},x_{n-1}) = 1$ for all $n \geq 1$. (Hint: Try using induction on $n$.)
+
+\item Let $a, b, x$ be positive integers with $x \geq 2$ and set $d = gcd(a, b)$.
+\begin{enumerate}
+\item Prove that $x^{d} - 1$ divides $gcd(x^{a} - 1, x^{b} - 1)$.
+
+\item Prove that $x^{d}-1$ is a multiple of $gcd(x^{a}-1,x^{b}-1)$ and conclude that
+$x^{d}-1 = gcd(x^{a}-1,x^{b}-1)$. (Hint: We know that there exist integers $u$ and $v$ such that
+$d = au + bv$. Now show that there exist integers $\alpha$ and $\beta$ such that
+$x^{d}-1 = \alpha(x^{a}-1) + \beta(x^{b}-1)$.)
+
+\end{enumerate}
+
+\item Show that the equation $1495x + 50060y = 4$ has no solutions for $x,y\in\ZZ$.
 \begin{quote}
 \begin{align*}
-a\cdot b &= 4 - 3 + 1\\
-&= 2\\
-||a|| &= \sqrt{16+9+1}\\
-&= \sqrt{26}\\
-||b|| &= \sqrt{1+1+1}\\
-&= \sqrt{3}
+50060 &= 1495(33) + 725\\
+1495 &= 725(2) + 45\\
+725 &= 45(16) + 5\\
+45 &= 5(9) + 0\\
+gcd(50060,1495) &= 5
 \end{align*}
+However, $5\cancel{|}4$, so there is no solution.
 \end{quote}
 
-\item Find the angle between $\vv{a} = \sqrt{3}\ i + j$ and $\vv{a} = -\sqrt{3}\ i + j$.
+\item Find all solutions to the equation $1485x + 1745y = 15$ for $x,y\in\ZZ$.
+
+\begin{center}
+\begin{minipage}[t]{0.4\textwidth}
 \begin{align*}
-\theta &= \arccos{\left({a\cdot b \over ||a||\ ||b||}\right)}\\
-&= \arccos{\left({-2\over 4}\right)}\\
-&= \arccos{\left(-{1\over 2}\right)}\\
-&= {2\pi \over 3}
+1745 &= 1485(1) + 260\\
+1485 &= 260(5) + 185\\
+260 &= 185(1) + 75\\
+185 &= 72(2) + 35\\
+75 &= 35(2) + 5\\
+35 &= 7(5) + 0\\
+gcd&(1745,1485) = 5
 \end{align*}
+\end{minipage}
+\begin{minipage}[t]{0.4\textwidth}
+\begin{align*}
+5 &= 75 - 35(2)\\
+5 &= 75 - (185 - 75(2))(2)\\
+&= 75(5) - 185(2)\\
+5 &= (260-185(1))(5)-185(2)\\
+&= 260(5)-187(7)\\
+5 &= 260(5) - (1485 -260(5))(7)\\
+&= 260(40) + 1485(-7)\\
+5 &= (1745 - 1485(1))(40) + 1485(-7)\\
+&= 1745(40) + 1485(-47)
+\end{align*}
+\end{minipage}
+\end{center}
 
-\item Find the angle between $\vv{a} = i + j$ and $\vv{b} = i + j + k$.
+$x = -47 + 349n$ and $y = 40 + 297n$ are solutions to the equation.
+
+\item Suppose you have two small champagne glasses, one holding 8 ounces and another holding 5 ounces. Is
+it possible to fill one of the glasses with exactly 1 ounce of champagne? If so, how can this be done? If
+not, prove that it cannot be done.
+\begin{quote}
+We need to find a solution to the Diophantine Equation
+\[
+8x+5y=1.
+\]
+We know that $gcd(8,5)=1$, and $1|1$, so there is a solution. Working backwards through the Euclidian
+Algorithm gives us
 \begin{align*}
-\theta &= \arccos{\left({a\cdot b \over ||a||\ ||b||}\right)}\\
-&= \arccos{\left({2 \over \sqrt{6}}\right)}
+1 &= 3-2(1)\\
+&= 3(2)-5\\
+&= 8(2)+5(-3).
 \end{align*}
-
-\item Find the angle between $\vv{a} = (1,-2,3)$ and $\vv{b} = (3,-6,-5)$.
-\begin{align*}
-\theta &= \arccos{\left({a\cdot b \over ||a||\ ||b||}\right)}\\
-&= \arccos{(0)}\\
-&= {\pi \over 2}
-\end{align*}
-
-\item Calculate $proj_{a}b$ where $\vv{a} = i + j$ and $\vv{b} = 2i + 3j - k$.
-\begin{align*}
-proj_{a}b &= {a\cdot b \over ||a||^{2}}\vv{a}\\
-&= {4\over 2}(i + j)\\
-&= 2i + 2j
-\end{align*}
-
-\item Calculate $proj_{a}b$ where $\vv{a} = i + j + 2k$ and $\vv{b} = 2i - 4j + k$.
-\begin{align*}
-proj_{a}b &= {a\cdot b \over ||a||^{2}}\vv{a}\\
-&= {0\over 4}(i+j+2k)\\
-&= 0
-\end{align*}
-
-\item Give a vector of length 3 that points in the same direction as the vector $i + j - k$.
-\begin{quote}
-The length is just a scalar value multiplied by the vector. Therefore we have the unit vector $i + j - k$,
-with length 3 will be
-\[3i + 3j - 3k.\]
-\end{quote}
-
-\item Is there ever a case where $proj_{a}b = proj_{b}a$?
-\begin{quote}
-
+Therefore, if we fill the first glass up twice and empty the second glass three times, we will end up with
+1 ounce leftover.
 \end{quote}

### 2016-02-04 math_modeling/class/notes/classnotes.tex
alsdhaklsjdklasjd

@@
 agents with no neighbors of the opposite type.
+
+\section{The Game of Life}
+The game of life is a mathematical model created by John Conway. The rules are as follows:
+\begin{enumerate}
+\item \underline{Birth Rule:} If at time $t$ a cell is dead (empty) and the cell has 3 live neighbors
+in any direction, then at time $t+1$ the cell becomes alive.
+\item \underline{Death Rule:} If at time $t$ a live cell has 0 and 1 neighbors it dies of isolation,
+and if a live cell has 4 or more neighbors, it dies of overcrowding.
+\item \underline{Survival Rule:} If at time $t$ a live cell has 2 or 3 live neighbors, then at time
+$t+1$, the cell is still alive.
+\end{enumerate}
 \end{document}

### 2016-02-04 math_modeling/hw/hw2/agenthw.tex
alsdhaklsjdklasjd

% 1, 2a, 2c, 2f, 2h, 2i and 2k
\documentclass{hw}

\begin{document}
\makeheader{}

\begin{enumerate}
\item One famous example of how different models must be used to solve different problems in the same
system is grocery store checkout queues. If you are a customer deciding which queue to enter, how would
you model the problem? What exact question would your model address? What entities and processes would
be in the model? Now, if instead you are a store manager deciding on how to operate the queues for the
next hour or so, what questions would you model address and what would it look like? Finally, if you
are a store designer and the question is how to design the checkout area so that 100 customers can
check out per hour with the fewest employees, what things would you model?
\begin{quote}
\underline{Customer:} The customer needs to take in to consideration how many queues are open,
how many people are currently in the queue, and how fast the queues are moving. They should also consider
how many items they have, since some lanes may be restricted to a certain amount of items. The question
that needs to be answered in this case is ``Which queue should I enter?"\\\\
\underline{Manager:} The manager also needs to consider how many people are in each queue, and how
quickly the queues are moving. They may also consider the number of people that are not in the queue
yet, since those people will eventually join the queue. Ultimately, the manager is trying to answer
``Are the queues moving slowly, and if so how can I fix them?".\\\\
\underline{Designer:} The designer needs to take in to account the number of expected patrons of the store
on a day to day, and maybe even hour to hour basis. They also need to consider the number of employees
that the store can afford, since that will affect the number of lanes that can be open at a given time.
\end{quote}

\item For the following questions, what should be in a model? what kinds of things should be represented,
what variables should those things have to represent their essential characteristics, and what processes
that change things should be in the model? Should the model be agent-based? If the question is not clear
enough to decide, then reformulate the question to produce one that is sufficiently clear.
\begin{enumerate}
\item How closely together should a farmer plant the trees in a fruit orchard?
\begin{quote}
This model should be agent based where the trees are the agent. If the farmer plants the trees too closely
together, then they will not be able to evenly absorb the nutrients. If they are too far apart, then the
farmer loses land that he could use to plant more trees.
\end{quote}
\item Should a new road have one, two, or three lanes in each direction?
\begin{quote}
This problem is not an agent-model. In this case, the amount of regular traffic in the area would need
to be taken in to account. There may also be the issue of cost for using more asphalt, as well as
the amount of space for the road.
\end{quote}
\item How many trees per year should a timber company harvest?
\begin{quote}
This is an agent-model where the trees are the agents. The timber company should consider the impact that
removing too many trees would have on the area; if they remove too many trees, then they will destroy the
environment around the area.
\end{quote}
\item To maximize profit, how many flights per day should Saxon Airlines schedule between Frankfurt and
Leipzig?
\begin{quote}
This is an agent-model where the flights are the agents. The airline company needs to consider how many
flights are taken between the two cities, and how much each flight costs them versus how much profit they
turn from each flight.
\end{quote}
\item To minimize system wide delays and risk of accidents, how many flights per day should the EAA allow
between Frankfurt and Leipzig?
\begin{quote}
This is not an agent-model. The EAA needs to consider the amount of risk associated with high numbers
of flights.
\end{quote}
\item Any other problems or questions from your studies, research, or experience in general, that might
be the basis of a model or agent-based model.
\begin{quote}
A common model that is not agent-based is modeling heat flow through a surface. In 3D space, this can be
modeled with a PDE that looks like
\[
u_{t} = c^{2}\nabla^{2}u.
\]
This formula assumes an infinite length pipe, but in the real world, this does not exist. If we cut off
the ends of the pipe, we are now dealing with a much more complex system. The heat no longer flows
only through the cylinder, but also through the top and bottoms of the pipe.
\end{quote}
\end{enumerate}
\end{enumerate}
\end{document}

### 2016-02-04 math_modeling/hw/hw3/online.tex
alsdhaklsjdklasjd

% page 610 3,4,5,8,9,10,12-18
\documentclass{hw}
\begin{document}
\makeheader{Chapter 19}

\begin{enumerate}
\setcounter{enumi}{4}
\item
For $30\%$: 72.9\\
For $50\%$: 83.4\\
For $67\%$: 97.0\\
Mean: $84.433\%$\\
Standard Deviation: $12.08318$

\setcounter{enumi}{7}
\item For small values of $\%$-similar wanted, only a few ticks happen before everyone is happy. When
$\%$-similar wanted is between $25\%$ and $26\%$, the $\%$-similar is around $60\%$.

\item In both cases, the model never reaches a state of equilibrium. However, when the value is $70\%$,
it tends to be more segregated. When the value is $80\%$, it tends towards chaos.

\setcounter{enumi}{13}
\item In this case, the Schelling model predicts that for low enough percent similar wanted, the
population
will settle out.

\item The Schelling model predicts that the neighborhood will constantly be changing unless the population
has a very low percent wanted.

\item The Schelling model predicts that the neighborhood will constantly be changing unless the population
has a very low percent wanted.

\setcounter{enumi}{17}
\item The model does not change very drastically by using the Von-Neumann neighborhoods.
\end{enumerate}
\end{document}

### 2016-02-04 number_theory/hw/hw3.tex
alsdhaklsjdklasjd

@@
 $gcd(a, b) = 1$.
+\begin{quote}
+Assume $J=\ZZ$. We know that $1\in\ZZ$, so $1\in J$, and $1\in\{ax+by|x,y\in\ZZ\}$. Therefore
+$1=ax+by$, so $gcd(a,b)=1$ by Bezout's Identity.\\\\
+Assume $gcd(a,b)=1$. Then there exists $u,v\in\ZZ$ such that $1 = au+bv$ by Bezuot's Identity. We know
+that any $n\in\ZZ$, we can represent it as $1\cdot n = a\cdot un + b\cdot vn$. Let $x=un$ and $y=vn$.
+Then we have $n = ax+by$ where $ax+by\in\ZZ$. Since $n$ is any arbitrary integer, we can say that
+$ax+by$ is any arbitrary integer for any $x,y\in\ZZ$, or $\{ax+by|x,y\in\ZZ\}=\ZZ$
+\end{quote}
+$\triangle$
 
-\item We define the \textit{Fibonacci Sequence} to be the sequence of integers ${x_{0}, x_{1}, x_{2},\dots}$
+\item We define the \textit{Fibonacci Sequence} to be the sequence of integers
+${x_{0}, x_{1}, x_{2},\dots}$
 satisfying the properties
 \[
-x_{0} = 0,\quad x_{1} = 1,\quad\text{and}\quad x_{n} = x_{n-1}+x_{n}\text{ for all } n\geq2.
+x_{0} = 0,\quad x_{1} = 1,\quad\text{and}\quad x_{n} = x_{n-1}+x_{n-2}\text{ for all } n\geq2.
 \]
 Prove that $gcd(x_{n},x_{n-1}) = 1$ for all $n \geq 1$. (Hint: Try using induction on $n$.)
+\begin{quote}
+For the base case, $n=2$, we have that
+\begin{align*}
+x_{2} &= x_{1} + x_{0}\\
+&= 1 + 0\\
+&= 1\\
+x_{1} &= 1\\
+gcd&(x_{1},x_{2})= 1.
+\end{align*}
+We can assume that $gcd(x_{n},x_{n-1}) = 1$ for $n\geq2$. We now need to show that
+$gcd(x_{n+1},x_{n}) = 1$ for $n\geq1$. We also have that $x_{n+1}=x_{n}+x_{n-1}$,
+so $gcd(x_{n+1},x_{n}) = gcd(x_{n}+x_{n-1},x_{n})$. By Bezout's Identity, we know that
+$gcd(x_{n}+x_{n-1},x_{n}) = (x_{n}+x_{n-1})u + x_{n}v$ for some $u,v\in\ZZ$. Then
+\begin{align*}
+gcd(x_{n}+x_{n-1},x_{n}) &= x_{n}u+x_{n-1}u + x_{n}v\\
+&= x_{n-1}u + x_{n}(u+v)\\
+&= x_{n-1}u + x_{n}w\\
+&= gcd(x_{n-1},x_{n})\\
+&= 1\text{ by the induction hypothesis}.
+\end{align*}
+Therefore, $gcd(x_{n+1},x_{n}) = 1$.
+\end{quote}
+$\triangle$
 
@@
 \item Prove that $x^{d} - 1$ divides $gcd(x^{a} - 1, x^{b} - 1)$.
-
+\begin{quote}
+We know that $d = gcd(a, b)$, so $d|a$ and $d|b$. We have $a=dn$ and $b=dm$. Then $x^{a}-1=x^{dn}-1$ and
+$x^{b}-1=x^{dm}-1$. We know that $x^{dn}=(x^{d})^{n}$, so we can rewrite $x^{dn}-1$ as $(x^{d})^{n}-1$.
+Now we have that $(x^{d})^{n}-1 = (x^{d}-1)(x^{dn-1}+x^{dn-2}+\cdots+1)$. Recall that
+$(x^{d})^{n}-1 = x^{a} - 1$, so we have $x^{a} - 1 = (x^{d}-1)j$, where
+$j=(x^{dn-1}+x^{dn-2}+\cdots+1)\in\ZZ$. Therefore, $(x^{a} - 1)|(x^{d}-1)$. We can apply the same
+argument for $x^{b}-1$, and we will find that $(x^{b} - 1)|(x^{d}-1)$. Since $(x^{d}-1)$ divides both
+$(x^{b}-1) \text{ and } (x^{a}-1)$, it divides $gcd(x^{a} - 1, x^{b} - 1)$.
+\end{quote}
 \item Prove that $x^{d}-1$ is a multiple of $gcd(x^{a}-1,x^{b}-1)$ and conclude that
@@
 $x^{d}-1 = \alpha(x^{a}-1) + \beta(x^{b}-1)$.)
-
+\begin{quote}
+Since we know $d=au+bv$ for $u,v\in\ZZ$, we can say
+\begin{align*}
+x^{d}-1 &= x^{au+bv}-1\\
+&= x^{au}x^{bv}-1\\
+&= x^{au}x^{bv}-1+x^{au}-x^{au}\\
+&= x^{au}\cdot(x^{bv}-1)+1\cdot(x^{au}-1).
+\end{align*}
+If we let $a=^{au}$ and $b=1$, then we have $x^{d}-1 = a\cdot(x^{bv}-1)+b\cdot(x^{au}-1)$. Therefore,
+by Bezout's Identity, $x^{d}-1 = gcd(x^{a}-1,x^{b}-1)$.
+\end{quote}
+$\triangle$
 \end{enumerate}
@@
 
+\item Find all solutions to the equation $7x+4y=1$ for $x,y\in\ZZ$.
+\begin{center}
+\begin{minipage}[t]{0.4\textwidth}
+\begin{align*}
+7 &= 4\cdot1 + 3\\
+4 &= 3\cdot1 + 1\\
+3 &= 1\cdot3 + 0
+\end{align*}
+\end{minipage}
+\begin{minipage}[t]{0.4\textwidth}
+\begin{align*}
+1 &= 4\cdot1 - 3\cdot1\\
+&= 4\cdot1 - (7-4)\cdot1\\
+&= 4\cdot1 -7\cdot1 + 4\cdot1\\
+&= 4\cdot2-7\cdot1
+\end{align*}
+\end{minipage}
+\end{center}
+$x = 1+7n$ and $y=2+4n$ are solutions to the Diophantine equation.
+
 \item Find all solutions to the equation $1485x + 1745y = 15$ for $x,y\in\ZZ$.
-
 \begin{center}
@@
 5 &= (1745 - 1485(1))(40) + 1485(-7)\\
-&= 1745(40) + 1485(-47)
+&= 1745(40) + 1485(-47)\\
+15 &= 1745(120) + 1485(-141)
 \end{align*}
@@
 
-$x = -47 + 349n$ and $y = 40 + 297n$ are solutions to the equation.
+$x = -141 + {349n\over3}$ and $y = 120 + 99n$ are solutions to the equation.
 

### 2016-02-08 number_theory/defs.tex
440 homework

\documentclass{hw}
\title{Theorems and Definitions}
\date{}
\author{}

\newcommand{\Define}[1]{\item\underline{#1}:}
\newcommand{\prm}{p}

\begin{document}
\maketitle

\begin{enumerate}
\Define{Prime Number} A positive $n\in\ZZ$ such that $n>1$ is prime if its only
divisors are 1 and $n$.

\Define{Fundamental Theorem of Arithmetic} Suppose $n\in\ZZ > 1$. Then
\begin{enumerate}
\item $\exists\alpha_{1},\alpha_{2},\dots,\alpha_{n}\in\NN$ and primes $\prm_{1},
\prm_{2},\dots,\prm_{n}$ such that $n=\prm_{1}^{\alpha_{1}}\prm_{2}^{\alpha_{2}}
\dots\prm_{n}^{\alpha_{n}}$, or $n=\prod_{i=1}^{n}\prm_{i}^{\alpha_{i}}$.
\item The expression $\prod_{i=1}^{n}\prm_{i}^{\alpha_{i}}$ is unique up to
permutation of the primes.

\Define{Lemma} Suppose $\prm$ is prime and $a,b\in\ZZ$.
\begin{enumerate}
\item If $\prm\nmid a$, then $\prm$ and $a$ are relatively prime.
\item If $\prm\mid ab$, then $\prm\mid a$ xor $\prm\mid b$.
\end{enumerate}

\Define{Corollary} If $\prm$ is prime and divides $a_{1},a_{2},\dots,a_{n}$, then
$p|a_{i}$ for some $i\in\{ 1,2,\dots,n \}$.

\Define{Corollary} If $m$ is not a perfect square, then $\sqrt{m}\notin\QQ$.
\end{enumerate}

\Define{Exact Division} If $\prm^{\alpha}\mid n$ and $\prm^{\alpha+1}\nmid n$, then
$\prm^{\alpha}$ exactly divides $n$, noted as $\prm^{\alpha}||n$.

\Define{Theorem} Suppose $a_{1},a_{2},\dots,a_{k}$ are pairwise relatively prime
and are greater than 0. If $a_{1}a_{2}\dots a_{k} = x^{m}$ for some $x\in\NN$ and
$m\geq2$, then $\exists x_{i}\in\NN$ such that $a_{i} = x_{i}^{m}$ for all $i$.

\Define{Theorem} There are infinitely many primes.

\Define{Prime Number Theorem} Let $\pi(x)$ denote the number of primes such that
$\prm_{i}\leq x$. We have that
\begin{enumerate}
\item $\lim_{x\to\infty}\pi(x)=\infty$
\item $\pi(x) < x$ for $x\geq2$.
\end{enumerate}
We can also define the logarithmic integral as
$\mathcal{L}_{i}(x)=\int_{2}^{x}{1\over\log{t}}\mathop{dt}$. Then
\[
\lim_{x\to\infty}{\mathcal{L}_{i}\over\pi(x)}=
\lim_{x\to\infty}{{x\over\log{x}}\over\pi(x)} = 1
\]
\end{enumerate}
\end{document}

### 2016-02-09 number_theory/hw/hw4.tex
Legendre and Number Theory

\documentclass{hw}

\begin{document}
\makeheader{4}

\begin{enumerate}
\item If $p$ is prime and $p\mid a^k$, prove that $p\mid a$. Conclude that $p^k\mid a^k$.
\begin{quote}
Since $p\mid a^k$, we have that $p\mid aa^{k-1}$. Then $p\mid a$ or $p\mid a^{k-1}$, but not
both. If we assume that $p\mid a^{k-1}$, we can say that $p\mid aa^{k-2}$. If $p\mid a^{k-2}$, then we have
that $p\mid aa^{k-3}$. If we repeat this process, we will ultimately have that $p\mid aa^0$. Then $p\mid a$
or $p\mid a^0$. If $p\mid a$, then we are done. If $p\mid a^0$, then $p\mid 1$, which
is a contradiction. Therefore, $p$ must divide $a$.\\\\
Assume $p\mid a$. Then $a=pm$ for some $m\in\ZZ$. Raising both sides to the $k$ power gives $a^k=p^km^k$.
Since $m^k\in\ZZ$, we can say that $m^k=j$. Then $a^k=p^kj$, so $p^k\mid a^k$.
\end{quote}
$\triangle$

\item Suppose that $a$ and $b$ are positive integers.
\begin{enumerate}
\item If $a^2\mid b^2$ prove that $a\mid b$. (Hint: Assuming that $a^2m = b^2$, use unique factorization to
prove that $m$ is a perfect square.)
\begin{quote}
Assume that $a^2\mid b^2$. Then $b^2=a^2m$ for some $m\in\ZZ$. By the Fundamental Theorem of Arithmetic,
we have that
$(p_{1}^{\alpha_1}p_{2}^{\alpha_2}\dots p_{j}^{\alpha_j})^2 =
(p_{1}^{\beta_1}p_{2}^{\beta_2}\dots p_{k}^{\beta_k})^2m$ for $j,k\in\ZZ$ and $\alpha,\beta\in\NN$.
\end{quote}

\item If $n$ is a positive integer such that $a^n \mid b^n$ can we conclude that $a \mid b$? Either prove
your answer or provide a counterexample.
\begin{quote}

\end{quote}
\end{enumerate}

\item Suppose that $a$ and $b$ are positive integers and $p_1, p_2,\dots, p_n$ are primes such that
\[
a = p_{1}^{\alpha_1}p_{2}^{\alpha_2}\dots p_{n}^{\alpha_n}
\qquad\text{and}\qquad
b = p_{1}^{\beta_1}p_{2}^{\beta_2}\dots p_{n}^{\beta_n},
\]
where $\alpha_{i}$ and $\beta_{i}$ are non-negative (possibly equal to 0) integers. Prove that
\[
lcm(a,b)=p_{1}^{max(\alpha_{1},\beta_{1})}p_{2}^{max(\alpha_{2},\beta_{2})}\dots
p_{n}^{max(\alpha_{n},\beta_{n})}.
\]
\begin{quote}
Let $L = lcm(a,b)$. We have by definition that $a\mid L$ and $b\mid L$. If we consider the case where
$a\mid L$, then we have that $L = am$ for some $m\in\ZZ$. We can arrange this as ${L\over a}$, which is
\[
{L\over p_{1}^{\alpha_1}p_{2}^{\alpha_2}\dots p_{n}^{\alpha_n}}
=
{
p_{1}^{\gamma_1}p_{2}^{\gamma_2}\dots p_{n}^{\gamma_n}
\over
p_{1}^{\alpha_1}p_{2}^{\alpha_2}\dots p_{n}^{\alpha_n}
}
\text{ by the FTA.}
\]
If we assume that $\gamma_{i} < \alpha_{i}$ for some $i\leq n$, then $L < a$, and ${L\over a}\notin\ZZ$,
which is a contradiction; we must have that $\gamma_{i} \geq \alpha_{i}$. Similarly, we can express
$b\mid L$ as
\[
{
p_{1}^{\gamma_1}p_{2}^{\gamma_2}\dots p_{n}^{\gamma_n}
\over
p_{1}^{\beta_1}p_{2}^{\beta_2}\dots p_{n}^{\beta_n}
}
= q\in\ZZ.
\]
By the same argument, we must have that we must have that $\gamma_{i} \geq \beta_{i}$. If we take
any arbitrary $\alpha_{i}$ such that $\alpha_{i} < \beta_{i}$, then letting
$\gamma_{i} = \alpha_{i}$ would cause ${L\over b}\notin\ZZ$; we also have that taking
$\gamma_{i} = \beta_{i}$ when $\beta_{i} < \alpha_{i}$ will cause ${L\over a}\notin\ZZ$. Therefore,
$\gamma_{i}$ must be the maximum of $\alpha_{i},\beta_{i}$ for all $i$. We can conclude that
the prime factorization of $L$ must be
$p_{1}^{max(\alpha_{1},\beta_{1})}p_{2}^{max(\alpha_{2},\beta_{2})}\dots
p_{n}^{max(\alpha_{n},\beta_{n})}$.
\end{quote}
$\triangle$

\item Determine whether each of the following statements is true or false. If true, prove it. If false,
provide a counterexample.
\begin{enumerate}
\item If $gcd(a, p^2) = p$ then $gcd(a^2, p^2) = p^2$.
\begin{quote}
We know that $p\mid a$ by the definition of the $gcd$. We can say that $a = pn$ for some $n\in\ZZ$.
Then $a^2=p^2n^2$, so $p^2\mid a^2$. Since $p^2$ is the greatest thing that divides $p^2$ and
$a > p$ so $a^2>p^2$, $gcd(a^2, p^2) = p^2$.
\end{quote}

\item If $gcd(a, p^2) = p$ and $gcd(b, p^2) = p^2$ then $gcd(ab, p^4) = p^3$.
\begin{quote}
If we take the case where $a=p$ and $b=p^3$, then we have $gcd(p, p^2) = p$ and $gcd(p^3, p^2) = p^2$.
However, $gcd(p^4, p^4) \neq p^3$.
\end{quote}

\item If $gcd(a,p^2)=p$ then $gcd(a+p,p^2)=p$.
\begin{quote}
If we take $a,p=2$, then we have that $gcd(2,4)=2$, but $gcd(4,4)\neq 2$.
\end{quote}
\end{enumerate}

\item Prove that every prime $p\neq 3$ has the form $3q + 1$ or $3q + 2$ for some integer $q$. Moreover,
prove that there are infinitely many primes of the form $3q + 2$.
\begin{quote}
If we consider a prime number $p\neq 3$, we will get a remainder of either 1 or 2 when we divide it by
three; if we get a remainder of 0, then that number was a multiple of 3 and therefore not prime. We
also know that all prime numbers above 3 are odd. Therefore, by the division algorithm, we have that
$p=3q+1$ or $p=3q+2$.\\\\
Assume that there are finitely many primes of the form $3q + 2$. Then we can say
\begin{align*}
p_1 &= 3q_1 + 2\\
p_2 &= 3q_2 + 2\\
\dots& \\
p_n &= 3q_n + 2\\
\end{align*}
We will let $m = 3p_1p_2\dots p_k-1$, which we can express as $3p_1p_2\dots p_k-3+2$. By factoring
we have that $m=3(p_1p_2\dots p_k-1)+2$. Then we have a prime $p = 3q+2$ that divides $m$. Since there
are finitely many primes of this form, we have that $p=p_i$ for some $i \leq n$. Without loss of
generality, we will let $i=1$. Then we have
\begin{align*}
1 &= 3(p_1p_2\dots p_n)-m\\
&= p_1(3p_2p_3\dots p_n - {m\over p_1})\\
&= p_1j\text{ for some } j\in\ZZ
\end{align*}
Therefore, $p_1\mid1$, which is a contradiction.
\end{quote}
$\triangle$

\item Two primes $p$ and $q$ with $p<q$ are called twin primes if $p+2=q$. If $p$ and $q$ are twin primes
with $3 < p < q$, prove that $6 \mid p + 1$.
\begin{quote}

\end{quote}
\end{enumerate}
\end{document}

### 2016-02-09 partial_differential_equations/leg.tex
Legendre and Number Theory

\documentclass{hw}

\title{Problems Involving Legendre Polynomials}
\date{}

\begin{document}
\maketitle
\begin{enumerate}
\item Show $\int_{-1}^{1}x^2p_{l}(x)\dx = 0$ for $l\geq 3$.
\begin{quote}
We know that the general form for a Legendre polynomial $p_{l}(x)$ is given by Rodrigues' Formula to be
\[
p_{l}(x)={1\over2^ll!}{d^l\over \dx^l}\left[(x^2-1)^l\right].
\]
We can use integration by parts to perform the integration:\\
\begin{center}
\begin{tabular}{c c c}
$(+)$ & $x^2$ & ${d^{l-1}\over\dx^{l-1}}\left[(x^2-1)^l\right]$\\
& & \\
$(-)$ & $2x$ & ${d^{l-2}\over\dx^{l-2}}\left[(x^2-1)^l\right]$\\
& & \\
$(+)$ & $2$ & ${d^{l-3}\over\dx^{l-3}}\left[(x^2-1)^l\right]$.\\
\end{tabular}
\end{center}
We now have that
\[
\int_{-1}^{1}x^2p_{l}(x)\dx =
x^2{d^{l-1}\over\dx^{l-1}}\left[(x^2-1)^l\right] -
2x{d^{l-2}\over\dx^{l-2}}\left[(x^2-1)^l\right] +
2{d^{l-3}\over\dx^{l-3}}\left[(x^2-1)^l\right].
\]
It is clear that the $l\geq3$ requirement is necessary, as any $l<3$ would lead to a negative
order derivative. We now need to show that this value is equal to 0; to do so, we will start by
substituting $Q(x) = (x^2-1)^l$. We now have
\begin{align*}
&x^2{d^{l-1}\over\dx^{l-1}}\left[(x^2-1)^l\right] -
2x{d^{l-2}\over\dx^{l-2}}\left[(x^2-1)^l\right] +
2{d^{l-3}\over\dx^{l-3}}\left[(x^2-1)^l\right]\\
&= x^2{d^{l-1}\over\dx^{l-1}}Q(x) - 2x{d^{l-2}\over\dx^{l-2}}Q(x) + 2{d^{l-3}\over\dx^{l-3}}Q(x)\\
&= x^2Q^{(l-1)}(x) - 2xQ^{(l-2)}(x) + 2Q^{(l-3)}(x).
\end{align*}
Let's consider the case where $l=3$ to show that this integral is indeed 0. Substituting 3 for $l$ yields
\[
x^2Q^{''}(x) - 2xQ^{'}(x) + 2Q(x).
\]
Through some careful integration, we find that
\begin{align*}
Q &= (x^2-1)\\
Q^{'} &= 6x(x^2-1)^2\\
Q^{''} &= 24x^2(x^2-1)+6(x^2-1)^2.
\end{align*}
Notice that in every term of each polynomial there is an $(x^2-1)^n$ term, where $n\in\NN$. If we factor
this term, we find that we have $(x+1)^n(x-1)^n$. The original integral was over the interval $(-1,1)$,
which yields a 0 in each term when we evaluate the anti-derivative at those bounds. Hence,
\begin{align*}
x^2Q^{''}(x) - 2xQ^{'}(x) + 2Q(x)\Big|_{-1}^{1} &= x^2(0)-2x(0)+2(0)\\
&= 0.
\end{align*}
We have shown that for $l=3$, the value of the integral is 0. However, we need to show that
$\forall l\geq3$, the value of the integral is 0. We can use the Leibniz Rule to express the
$l^{\text{th}}$ derivative in terms of a summation.
\[
(f(x)g(x))^{(n)} = \sum_{k=0}^{n}{n\choose k}f^{(n-k)}(x)g^{(k)}(x).
\]
We are trying to prove that $x^2Q^{(l-1)}(x) - 2xQ^{(l-2)}(x) + 2Q^{(l-3)}(x) = 0$ for all $l\geq3$.
We know $Q=(x^2-1)^l$, so we can take one derivative of $Q$, which gives us $2xl(x^2-1)^{l-1}$. Let
$f(x)=2xl$ and $g(x)=(x^2-1)^{l-1}$. Our goal is to find the $l^{\text{th}}$ derivative of $Q$.
By the Leibniz Rule we have
\[
(fg)^{(l)} = \sum_{k=0}^{l}{l\choose k}f^{(l-k)}g^{(k)}.
\]
Consider the case where we take another derivative. We know that we will have a $(x^2-1)^{(l-2)}$ term.
Again, we can take another derivative, which would provide us with $(x^2-1)^{(l-3)}$. We know that this
term will continue to appear until $k>l$. However, the Leibniz Rule will stop when $k=l$. Therefore,
every product $f^{(l-k)}g^{(k)}$ will have a $(x^2-1)^{n}$ term where $n$ is the $l-k$ power.
We can also rewrite $(x^2-1)^{n}$ as $(x-1)^n(x+1)^n$, so we now know that every term in the summation
will have an $(x-1)^n(x+1)^n$ term. Recall that we are evaluating this term from $-1$ to $1$, so
when we ultimately substitute into the equation, we will either have $(x-1)^n = 0$ or $(x+1)^n = 0$.
Therefore, we can conclude that every term in the summation will be 0. This implies that the
$l^{\text{th}}$ derivative of $Q$ is 0 for any $l \geq 3$. Therefore
$x^2Q^{(l-1)}(x) - 2xQ^{(l-2)}(x) + 2Q^{(l-3)}(x) = 0$.
\end{quote}

\item We want to find $a_{l}$ in the series
\[
f(x) = \sum_{l=0}^{\infty}a_{l}p_{l}(x)
\]
where
\[
f(x)=
\begin{cases}
x,\ 0\leq x < 1\\
0,\ -1 < x \leq 1
\end{cases}.
\]
We know that we can express $a_{l}$ as
\[
a_{l} = {%
\int_{-1}^{1}f(x)p_{l}(x)\dx%
\over%
\int_{-1}^{1}p_{l}^{2}(x)\dx%
}.
\]
Since $f$ is piecewise defined to be $0$ on $-1 < x \leq 1$, we are only dealing with
\[
a_{l} = {%
\int_{0}^{1}xp_{l}(x)\dx%
\over%
\int_{-1}^{1}p_{l}^{2}(x)\dx%
}.
\]
We can use integration by parts to solve the numerator. If we let $u = x$ and
$\dv = p_{l}(x)\dx$, we have that
\[
{%
{x\over2^{l}l!}{d^{l-1}\over\dx^{l-1}}\left[(x^{2}-1)^l\right]\Big|_{0}^{1}%
-\int_{0}^{1}{1\over2^{l}l!}{d^{l-1}\over\dx^{l-1}}\left[(x^{2}-1)^l\right]\dx%
\over%
\int_{-1}^{1}p_{l}^{2}(x)\dx%
}.
\]
Let's consider the first term in the numerator:
${x\over2^{l}l!}{d^{l-1}\over\dx^{l-1}}\left[(x^{2}-1)^l\right]\Big|_{0}^{1}$.
In order to evaluate the differentiated term, we need to express the derivative in a closed form.
If we make the substitution $y=x^{2}$, we have
\[
(y-1)^{l} = (y-1)(y^{l-1}+y^{l-2}+\cdots+1).
\]
By back substitution, we have
\begin{align*}
(x^{2}-1)^{l} &= (x^2-1)(x^{2l-2}+x^{2l-4}+\cdots+1)\\
&= (x^2-1)(\sum_{j=0}^{l}x^{(2l-2j)})\\
&= x^{2l+2}-1
\end{align*}

\end{enumerate}

\end{document}

### 2016-02-09 cleaner.sh

@@ -6,7 +6,7 @@ recurse () {
 			recurse
 			cd ..
 		else
-			gbbs.sh
+			gbbs
 		fi
 	done
 }

### 2016-02-12 number_theory/hw/hw4.tex
more work done

@@
 (p_{1}^{\beta_1}p_{2}^{\beta_2}\dots p_{k}^{\beta_k})^2m$ for $j,k\in\ZZ$ and $\alpha,\beta\in\NN$.
+Then we can say that
+\[
+{
+(p_{1}^{\beta_1}p_{2}^{\beta_2}\dots p_{k}^{\beta_k})^2
+\over
+(p_{1}^{\alpha_1}p_{2}^{\alpha_2}\dots p_{j}^{\alpha_j})^2
+}
+= m\in\ZZ.
+\]
+Therefore, $m$ is a perfect square, so we can take its square root. Let $n=\sqrt{m}$. Then
+$\sqrt{a^2m}=\sqrt{b^2}$, or $an=b$, so $a|b$.
 \end{quote}
@@
 \begin{quote}
-
+By the same argument used above, we can say
+\[
+{b^n\over a^n} =
+{
+(p_{1}^{\beta_1}p_{2}^{\beta_2}\dots p_{k}^{\beta_k})^n
+\over
+(p_{1}^{\alpha_1}p_{2}^{\alpha_2}\dots p_{j}^{\alpha_j})^n
+}
+= m\in\ZZ.
+\]
+Therefore, $m$ is a perfect $n^{\text{th}}$ power, so we can say $j=\sqrt[n]{m}$. Then
+$\sqrt[n]{b^n}=\sqrt[n]{a^nm}$ simplifies to $b=aj$. Therefore, $a|b$.
 \end{quote}
+$\triangle$
 \end{enumerate}
@@
 \begin{quote}
-
+We can take any three consecutive numbers starting with a prime $p$ where $p+2=q$ is also prime. We know
+that any prime number greater than 3 is even, so it is divisible by two. We also know that if we take
+three consecutive numbers, one of them has to be divisible by 3, since it will either have a remainder of
+0, 1, or 2. Since $p$ is prime and $p+2$ is prime, then $p+1$ must be divisible by three. Therefore
+$p+1$ must be divisible by 6, or $6|p+1$.
 \end{quote}

### 2016-02-12 partial_differential_equations/leg.tex
more work done

@@
 
+\newpage
 \item We want to find $a_{l}$ in the series
@@
 \]
-We can use integration by parts to solve the numerator. If we let $u = x$ and
-$\dv = p_{l}(x)\dx$, we have that
+We will begin by analyzing the numerator. We are solving
 \[
-{%
-{x\over2^{l}l!}{d^{l-1}\over\dx^{l-1}}\left[(x^{2}-1)^l\right]\Big|_{0}^{1}%
--\int_{0}^{1}{1\over2^{l}l!}{d^{l-1}\over\dx^{l-1}}\left[(x^{2}-1)^l\right]\dx%
-\over%
-\int_{-1}^{1}p_{l}^{2}(x)\dx%
-}.
+\int_{0}^{1}xp_{l}(x)\dx.
 \]
-Let's consider the first term in the numerator:
-${x\over2^{l}l!}{d^{l-1}\over\dx^{l-1}}\left[(x^{2}-1)^l\right]\Big|_{0}^{1}$.
-In order to evaluate the differentiated term, we need to express the derivative in a closed form.
-If we make the substitution $y=x^{2}$, we have
+Using integration by parts gives us
+\begin{align*}
+&x{1\over2^ll!}{d^{l-1}\over dx^{l-1}}\left[(x^2-1)^l\right]-
+{1\over2^ll!}{d^{l-2}\over dx^{l-2}}\left[(x^2-1)^l\right]\Big|_{x=0}^{x=1}\\
+&= xp_{l-1}(x)-p_{l-2}(x)\Big|_{x=0}^{x=1}\\
+&= \left[1\cdot p_{l-1}(1) - p_{l-2}(1)\right] -\left[0-p_{l-2}(0)\right]\\
+&= p_{l-2}(0).
+\end{align*}
+We can express $p_{l}(0)$ as
 \[
-(y-1)^{l} = (y-1)(y^{l-1}+y^{l-2}+\cdots+1).
+p_{l}(0) =
+\begin{cases}
+{(-1)^{l/2}\over2^l}{l\choose l/2}\quad x\in\mathbb{E}\\
+0\qquad\qquad\quad\ x\in\mathbb{O}
+\end{cases}.
 \]
-By back substitution, we have
+We will consider even $l$ here, since all odd $l$ will produce a 0. We can now express $a_{l}$ as
 \begin{align*}
-(x^{2}-1)^{l} &= (x^2-1)(x^{2l-2}+x^{2l-4}+\cdots+1)\\
-&= (x^2-1)(\sum_{j=0}^{l}x^{(2l-2j)})\\
-&= x^{2l+2}-1
+a_{l} &= {(2l+1)p_{l-2}(0)\over2}\\
+&= {4(2l+1)\over l!}.
 \end{align*}
-
+The Fourier series is now
+\[
+\sum{4(2l+1)\over l!}p_{l}(x)
+\]
+for even $l$.
 \end{enumerate}

### 2016-02-12 writing_in_ethics/paper_two.tex
more work done

\documentclass{paper}

\addbibresource{generic_bib.bib}

\title{Social Patriotism}
\begin{document}
\drafth{}
\inlinetitle
\begin{linenumbers}

Governmental power has drastically increased over the past decade. Much of the growth of power is due to acts of terrorism that have hugely impacted the nation. One of the most devastating acts of terrorism that the United States has seen is the attack on the world trade center on September 11, 2001. As a result, the government passed the USA PATRIOT act, also known as the Patriot Act. Many of the provisions in the Patriot Act grant the government the ability to avoid due process. Under the Patriot Act, the government is given excessive power to bypass the law; Kantian ethics provide an insight that can help determine whether or not use of the liberties granted to the government are ethically sound.

Section 6.6.6 of the text presents a situation in which a British news magazine, known as \textit{The Guardian}, published an article describing how the FBI indirectly requested that Verizon turn over all of its phone records. A key tenant of Kantian ethics is that others should never be used as a means to an end. The FBI's intention was never revealed, so it is hard to say whether or not they were using the phone records for personal gain or not. However, this fact alone is a breach of Kantian ethics; the FBI wanted to have access to private information, but refused to divulge the reason. Under the Categorical Imperative, moral actions should be able to be applied universally. The FBI did not release their private information, so, under the Categorical Imperative, the should not have taken others private information.

The power over communications granted to the FBI is greatly increased by the Patriot Act. Section 6.6.2 discusses the use of National Security Letters (NSLs), namely in regards to the Library Connection of Connecticut. In this scenario, the FBI used a provision of the Patriot Act to demand a user's browser history from a library computer. It follows from the argument made about the FBI requesting Verizon phone records that NSLs are a breach of Kantian ethics; the FBI prohibits individuals and companies that receive NSL's to reveal that the FBI had contacted them in the first place. This scenario differs from the Verizon scenario in that the FBI wanted the information of only one individual. The use of the NSL aside, the FBI's actions are still unethical under Kantian analysis. The agency requested that the Library Connection hand over information about another individual. In doing so, the FBI was using the Library Connection as a means to an end, which violates the Categorical Imperative. In this scenario, as well as the last, it is difficult to completely analyze the situations using Kantian analysis, since the FBI's intention is not actually know.

The third scenario present by the text involves Brandon Mayfield, an attorney from Portland, Oregon. The FBI was presented with a partial fingerprint recovered from a terrorist attack in Spain that was linked back to Mayfield. The FBI then proceeded to investigate Mayfield, going as far as entering his home and seizing private information without proper legal authorization. This scenario differs from others in that the FBI's actions were directly aimed at Mayfield. They did not use anyone as a means to an end, as they did in other scenarios. However, they still violated Kant's Categorical Imperative. Kantianism focuses on the idea that moral laws should be able to be willed into universal laws, and as such, no party involved should be able to opt-out. Following this logic, Mayfield, as well as anyone else who wanted to, would be able to take information without legal basis from the FBI. This scenario differs from the previous two in that the FBI's intention is known. The agency sought to bring a suspected terrorist into custody. The methods used to reach that goal were not ethical however, which makes their actions unethical.

It appears that under Kantian analysis, many of the actions taken by the FBI are unethical. 
\end{linenumbers}
\newpage
\printbibliography
\end{document}

### 2016-02-12 algorithms/projects/proj1/PJOne.java

@@ -1,0 +1,74 @@
+// Test Review
+//
+// Big theta, O, Omega
+// Logs, Exponents, n!
+// Summations
+// Master Theorem
+// Sorting and Searching up to Count Sort
+//      - Running times
+//      - Pseudocode
+//      - Recursion Trees
+// Analyze an algorithm.
+
+
+import static java.lang.System.out;
+
+public class PJOne {
+
+    public static void algoOne(int size, int[] a){
+        int j = 1;
+        int[] b = new int[size];
+        int n = size;
+        while(j <= n - 1){
+            int i = 1;
+            while(i <= n - j){
+                int m = i + j - 1;
+                int r = min(m + j, n);
+                int u = i;
+                for(; u < r; u++){
+                    b[u] = a[u];
+                }
+                u = i;
+                int v = m + 1;
+                int w = i;
+                for(; w < r; w++){
+                    if(u > m || (v <= r && b[v] < b[u])){
+                        a[w] = b[v];
+                        v++;
+                    } else {
+                        a[w] = b[u];
+                        u++;
+                    }
+                }
+                i += 2 * j;
+            }
+            j *= 2;
+        }
+    }
+
+    public static void compute(int[] a, int[] b, int i){
+        if(i > a.length){
+            for(int j = i; j < a.length; j++){
+                if(b[j] > 0){
+                    out.println(b[j]);
+                }
+            }
+            return;
+        }
+        b[i] = 0;
+        compute(a,b,i + 1);
+        b[i] = a[i];
+        compute(a,b,i + 1);
+        b[i] = 0;
+    }
+
+    private static int min(int a, int b){
+        return a < b ? a : b;
+    }
+
+    public static void main(String[] args){
+        int[] a = {1,2,3,4};
+        int[] b = {0,0,0,0};
+        compute(a,b,0);
+    }
+}

### 2016-02-15 writing_in_ethics/paper_two.tex
algorithm work

@@
 \addbibresource{generic_bib.bib}
+\addbibresource{paper_two_bib.bib}
 
@@
 
-The third scenario present by the text involves Brandon Mayfield, an attorney from Portland, Oregon.
+The third scenario presented by the text involves Brandon Mayfield, an attorney from Portland, Oregon.
 The FBI was presented with a partial fingerprint recovered from a terrorist attack in Spain that was linked back to Mayfield.
@@
 It appears that under Kantian analysis, many of the actions taken by the FBI are unethical.
+The Patriot Act allows for the government to interfere with the lives of the American population with little regard for the law, but the legality of an action does not necessarily determine whether that action is ethically right or wrong.
+It is hard to say whether or not the Patriot Act itself is ethical or not, but many of the ways in which law enforcement has interpreted and used it are unethical.
 \end{linenumbers}
 \newpage
+\nocite{*}
 \printbibliography

### 2016-02-15 writing_in_ethics/presentation.tex
algorithm work

\documentclass{beamer}

\usepackage{url}

\title{Microtransactions and DLC}
\author{Steven Rosendahl}
\date{}

\usetheme{metropolis}

\begin{document}
\frame{\titlepage}

\begin{frame}
\frametitle{The Issue}
\begin{itemize}
\pause
\item Are micro-transactions ethical?
\pause
\item Should companies charge money for a product they market as free?
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Examples}
\pause
\begin{enumerate}
\item The Saint's Row Franchise
\pause
\begin{itemize}
\item
\end{itemize}
\end{enumerate}
\end{frame}

\begin{frame}[allowframebreaks]
\frametitle{References}
\bibliographystyle{apalike}
\bibliography{pres_bib.bib}
\end{frame}

\end{document}

### 2016-02-15 algorithms/projects/proj1/PJOne.java

@@ -15,10 +15,10 @@ import static java.lang.System.out;
 
 public class PJOne {
 
-    public static void algoOne(int size, int[] a){
+    public static void algoOne(int[] a){
+        int n = a.length;
         int j = 1;
-        int[] b = new int[size];
-        int n = size;
+        int[] b = new int[n];
         while(j <= n - 1){
             int i = 1;
             while(i <= n - j){
@@ -26,29 +26,29 @@ public class PJOne {
                 int r = min(m + j, n);
                 int u = i;
                 for(; u < r; u++){
-                    b[u] = a[u];
+                    b[u - 1] = a[u - 1];
                 }
                 u = i;
                 int v = m + 1;
                 int w = i;
                 for(; w < r; w++){
-                    if(u > m || (v <= r && b[v] < b[u])){
-                        a[w] = b[v];
+                    if((u > m) || (v <= r && b[v - 1] < b[u - 1])){
+                        a[w - 1] = b[v - 1];
                         v++;
                     } else {
-                        a[w] = b[u];
+                        a[w - 1] = b[u - 1];
                         u++;
                     }
                 }
-                i += 2 * j;
+                i = i + (2 * j);
             }
             j *= 2;
         }
     }
 
     public static void compute(int[] a, int[] b, int i){
-        if(i > a.length){
-            for(int j = i; j < a.length; j++){
+        if(i > a.length - 1){
+            for(int j = i; j < a.length - 1; j++){
                 if(b[j] > 0){
                     out.println(b[j]);
                 }
@@ -67,8 +67,13 @@ public class PJOne {
     }
 
     public static void main(String[] args){
-        int[] a = {1,2,3,4};
-        int[] b = {0,0,0,0};
-        compute(a,b,0);
+        int [] a = new int[]{12,-12,-55,0,-16,-189};
+        for(int aa : a)
+            out.printf("%6d", aa);
+        out.println();
+        algoOne(a);
+        for(int aa : a)
+            out.printf("%6d",aa);
+        out.println();
     }
 }

### 2016-02-15 writing_in_ethics/paper_two.tex
finish paper two

@@
 The Patriot Act allows for the government to interfere with the lives of the American population with little regard for the law, but the legality of an action does not necessarily determine whether that action is ethically right or wrong.
-It is hard to say whether or not the Patriot Act itself is ethical or not, but many of the ways in which law enforcement has interpreted and used it are unethical.
+It is hard to say whether or not the Patriot Act itself is ethical or not; Kantian ethics do however focus on the morality of a decision.
+The Patriot Act was put in place to protect the citizens of the United States, which Kant would argue is a morally right cause.
+Many of the actions taken by the FBI in the name of the Patriot Act were presumably done for the same reason, but there is no way to know for certain.
 \end{linenumbers}

### 2016-02-16 writing_in_ethics/presentation.tex
work on ethics presentation

@@
 \usepackage{url}
+\usepackage{graphicx}
 
-\title{Microtransactions and DLC}
+\makeatletter
+\g@addto@macro{\UrlBreaks}{\UrlOrds}
+\makeatother
+
+\setbeamertemplate{bibliography item}{}
+\setbeamertemplate{bibliography entry title}{}
+\setbeamertemplate{bibliography entry location}{}
+\setbeamertemplate{bibliography entry note}{}
+
+\title{DLC}
 \author{Steven Rosendahl}
@@
 \pause
-\item Are micro-transactions ethical?
+\item Is the concept of Downloadable Content (DLC) ethical?
 \pause
 \item Should companies charge money for a product they market as free?
+\pause
+\item Should companies market an incomplete product and force users to pay more for the rest of the product?
+\pause
+\item When does DLC change from a fun add on to a game to a required purchase for a whole game?
 \end{itemize}
@@
 \frametitle{Examples}
-\pause
+% \begin{columns}
+% \begin{column}{0.5\textwidth}
 \begin{enumerate}
@@
 \begin{itemize}
-\item
+\item Saints Row: The Third released November 15, 2011
+\pause
+\item Saint Row IV released August 20, 2013
+\begin{itemize}
+\pause
+\item A re-skinned version of Saints Row: The Third
 \end{itemize}
+\pause
+\item Saints Row: Gat Outta Hell released January 20, 2015
+\begin{itemize}
+\pause
+\item Marketed as a separate game, but only about $1/8$ as long as Saints Row IV
+\end{itemize}
+\end{itemize}
+\pause
+% \begin{figure}
+% \end{figure}
+\begin{columns}[T]
+\begin{column}{0.5\textwidth}
+\item Train Simulator 2016
+\begin{itemize}
+\item Base game: \$29.99.
+\pause
+\item All DLC: $> \$5,400.00$.
+\end{itemize}
+\end{column}
+
+\hspace{-1.8cm}
+
+\begin{column}{0.5\textwidth}
+\pause
+\includegraphics[scale=0.1]{p3vp4}
+\pause
+
+\vspace{0.2cm}
+
+\includegraphics[scale=0.4]{train_sim_dlc}
+\end{column}
+
+\hspace{-2cm}
+
+\end{columns}
 \end{enumerate}
@@
 
-\begin{frame}[allowframebreaks]
+\begin{frame}
+\frametitle{Kantianism}
+\begin{itemize}
+\item ``Never use people as a means to an end"
+\pause
+\begin{itemize}
+\item Companies no longer care about making games
+\pause
+\item Profit can be maximized by continually making DLC at the expense of the gamer
+\pause
+\item All major game companies follow the DLC pattern, so gamers are forced to buy into it
+\pause
+\item The intention is to make more money off of gamers, which is using them as a means to an end
+\pause
+\end{itemize}
+\end{itemize}
+\begin{center}
+\includegraphics[scale=0.3]{unethical}
+\end{center}
+\end{frame}
+
+\begin{frame}
+\frametitle{Utilitarianism}
+\begin{itemize}
+\item ``The most amount of good for the most amount of people"
+\pause
+\begin{itemize}
+\item Companies turn large profits from DLC
+\pause
+\item Companies use that profit to create more content for gamers
+\pause
+\item Gamers enjoy playing the new games
+\pause
+\item The cycle repeats
+\end{itemize}
+\end{itemize}
+\pause
+\begin{center}
+\includegraphics[scale=0.3]{ethical}
+\end{center}
+\end{frame}
+
+\begin{frame}
+\frametitle{Social Contract}
+\begin{itemize}
+\item ``Rules that are agreed upon by all"
+\pause
+\begin{itemize}
+\item Game Companies vs. Gamers
+\pause
+\item Game companies want to turn a large profit
+\pause
+\item Gamers want complete games at a reasonable price
+\pause
+\item There has to be a compromise that both parties can agree upon
+\pause
+\item Current system favors the game companies, not the gamers
+\end{itemize}
+\end{itemize}
+\begin{center}
+\includegraphics[scale=0.3]{unethical}
+\end{center}
+\end{frame}
+
+\begin{frame}[shrink=20]
 \frametitle{References}
-\bibliographystyle{apalike}
-\bibliography{pres_bib.bib}
+\vspace{1cm}
+\nocite{*}
+\bibliographystyle{alpha}
+\bibliography{pres_bib}
 \end{frame}

### 2016-02-16 writing_in_ethics/presb.sh

@@ -1,0 +1,4 @@
+pdflatex presentation.tex
+bibtex presentation
+pdflatex presentation.tex
+pdflatex presentation.tex

### 2016-02-20 math_modeling/class/notes/classnotes.tex
mucho tarea

@@
 \end{enumerate}
+
+\section{Tumor Growth}
+Tumor growth can be modeled by the differential equation
+\[
+{dV\over dt} = aV^{2/3}-bV'.
+\]
+This equation arises from the logistic equation
+\[
+{dP\over dt} = P(a-bP),
+\]
+which has a solution of
+\[
+P={M\over 1+e^{-\log{be^c}-at}}.
+\]
+As $t\to\infty$, $P$ heads towards $M$, since the exponential decreases. The tumor equation can be solved
+by separation of variables:
+\begin{align*}
+{dV\over aV^{2/3}-bV}&=dt\\
+\int{dV\over aV^{2/3}-bV}&=\int dt\\
+3\int{du\over a-bu}&=t+c.
+\end{align*}
+The previous result arises from the $u$-substitution where $u=V^{1/3}$. We can make another substitution,
+where $w=a-bu$. Then
+\begin{align*}
+3\int{du\over a-bu}&=t+c\\
+{-3\over b}\ln{|a-bu|}&=t+c\\
+{-3\over b}\ln{|a-bV^{1/3}|}&=t+c.
+\end{align*}
+We now will specify the integration constant, $c$, in terms of the the volume of the tumor, $V$.
+\begin{align*}
+c &= t - {3\over b}\ln{|a-bV^{1/3}|}.
+\end{align*}
+The plot of this function can be expressed as
+\[
+t = {3\over b}\ln{|a-bV_{0}^{1/3}\over|a-bV^{1/3}},
+\]
+which simplifies to
+\[
+V = {e^{-bt}\left(-a+bV_{0}^{1/3}+ae^{bt/3}\right)^{3}\over b^{3}}
+\]
+This function's plot acts the same as the logistic equation.
+\subsection{Gompertz Curve}
+Gompertz Curve is an equation that arises from
+\[
+\begin{cases}
+{dV\over dt}=r(t)V\\
+{dr\over dt}=-Kr
+\end{cases}.
+\]
+The model has the same curve that the logistic growth model has, but the equation differs.
+\[
+V=V_{0}^{r_{0}/k}e^{-(r_{0}/k)e^{-kt}}.
+\]
+The model can also be expressed as a single equation:
+\[
+\begin{cases}
+{dV\over dt}=r(t)V\\
+{dr\over dt}=-Kr
+\end{cases}
+\to
+{dV\over dt} = aV-bV\ln{V},
+\]
+which has the solution
+\[
+V(t)=V_{0}e^{r_{0}/t}e^{-r_{0}/ke^{-kt}}
+\]
+
+\section{Voting}
+{\centering
+\begin{tabular}{c c c c c c}
+22\% & 23\% & 15\% & 29\% & 7\% & 4\%\\
+T&T&B&B&I&I\\
+B&I&T&I&B&T\\
+I&B&I&T&T&B
+\end{tabular}
+}
+\noindent\\
+Because of the plurality, T wins the election. If the I candidate is removed, then B wins the election.
+Often in this model, we deal with non-transitive systems. Consider the following voter profile:\\
+{\centering
+\begin{tabular}{c c}
+14 & $A>B>C$\\
+12 & $C>A>B$\\
+10 & $B>C>A$
+\end{tabular}
+}
+\noindent\\
+This profile has the following outcome:\\
+{\centering
+\begin{tabular}{l c c c}
+First & A&C&B\\
+Second & B&A&C\\
+Third & C&B&A\\
+& 14&12&10
+\end{tabular}
+}
+\noindent\\
+In this case, $A>B$, $B>C$, but $C>A$.
+
+\subsection{Pairwise Comparison}
+This is based on combinatorics; if we have 3 candidates, we have
+\[
+{3 \choose 2}
+\]
+possibilities.
+
 \end{document}

### 2016-02-20 math_modeling/hw/hw5/prog.tex
mucho tarea

% page 176, 8,10,12,14,17,18
\documentclass{hw}

\begin{document}

\begin{enumerate}
\item Why does l'H\^opital's rule apply to $\lim_{x\to 0}{V^{x} - 1 \over x}$, and precisely why is this
limit equal to $\lim_{x\to0}{V^{x}\ln{V} - 0\over 1}$?
\begin{quote}
l'H\^opital's rule applies since
\[
\lim_{x\to 0}{V^{x} - 1 \over x} = {V^{0} - 1\over0} = {1-1\over0} = {0\over0}.
\]
Using l'H\^opital's rule yields
\[
\lim_{x\to0}{V^{x}\ln{V} - 0\over 1}.
\]
\end{quote}

\item Use the Gompertz equation in the form ${dV\over dt} = (a-b\ln{V})V$ to explain why
$lim_{t\to\infty}V(t) = e^{a/b}$.
\begin{quote}
We can solve ${dV\over dt} = (a-b\ln{V})V$ for $V(t)$. We get
\[
V(t) = e^{{a\over b}-\left({a\over b}-\ln{V_{0}}\right)e^{-bt}}.
\]
If we take the limit as $t\to\infty$, we get
\begin{align*}
\lim_{t\to\infty}V(t) &= e^{{a\over b}-\left({a\over b}-\ln{V_{0}}\right)(0)}\\
&= e^{{a\over b}}.
\end{align*}
\end{quote}

\item In solving the Gompertz differential equation, we assumed that
$\int{1\over u}\du = \ln{u}+C$ rather than the more formally correct answer
$\int{1\over u}\du = \ln{|u|}+C$. Were we safe in ignoring the absolute value signs?
\begin{quote}
It was okay to ignore the absolute value signs since $u$ will always be positive. When solving the Gompertz
equation, we had that $u=Ce^{-bt}$. The term $e^{-bt}$ is always positive, and our constant $C$ was
determined by $V_{0}$, which is always positive. Therefore, $u > 0$, and $|u| = u$.
\end{quote}

\item Show that the Gompertz curve has a single point of inflection at the time when
$\ln{V} = {a\over b} - 1$.
\begin{quote}
The Gompertz equation has the form
\[
{dV\over dt} = aV-bV\ln{V}.
\]
If we differentiate this again, we get
\[
{d^2V\over dt^2} = a-b-b\ln{V}.
\]
Setting this equal to 0 yields
\begin{align*}
0 &= a-b-b\ln{V}\\
a &= b+b\ln{V}\\
{a\over b} &= 1 + \ln{V}\\
\ln{V} &= {a\over b} - 1.
\end{align*}
\end{quote}

\item With the estimated parameter values for the Gompertz model of chicken growth, show that the predicted
long-range limit to the size is 4.476.
\begin{quote}
From the model, we know that $g(t) = 4.47e^{-3.41e^{-0.251t}}$. If we take the limit as $t\to\infty$, we
have
\[
\lim_{t\to\infty}g(t) = 4.47e^{-3.41e^{0}} = 4.47.
\]
\end{quote}

\item Carry out the details of fitting the logistic model to the chicken weight data to show that
$d = 3.155450907$ and $a = 0.4124054532$.
\begin{quote}

\end{quote}
\end{enumerate}

\end{document}

### 2016-02-20 math_modeling/hw/prehw/q1.tex
mucho tarea

\documentclass{hw}

\usepackage{minted}

\begin{document}
\makeheader{}
\noindent\\

The logistic difference equation, governing population growth, is given by $x_{n+1}=rx_{n}(1-x_{n})$, in which
$r$ is the growth rate parameter. In the following exercises take the initial condition, $x_{1}$, to be
0.5.

\begin{enumerate}
\item Write a compute code (which utilizes either a for loop or a do loop) to analyze the behavior of the
logistic equation for $r=1.5$.

\vspace{0.5cm}

\begin{minipage}{0.5\textwidth}
\begin{minted}{python}
Population[r_, parg_: Automatic] :=
(
  x = Table[n, {n, 1, 200}];
  x[[1]] = 0.5;
  For[n = 2, n <= 200, n++,
   x[[n]] = r*x[[n - 1]]*(1 - x[[n - 1]])];
  ListPlot[x, PlotRange -> parg]
)
Population[1.5]
\end{minted}
\end{minipage}
\begin{minipage}{0.5\textwidth}
\centering
\includegraphics[scale=0.5]{p15}
\end{minipage}

\item Try several other values of the growth rate parameter in the range $1<r<3$. Do your results support
the conjecture that the limiting population always exists and is an increasing function or $r$?

\vspace{0.5cm}

{\centering
\begin{tabular}{c c c}
\includegraphics[scale=0.4]{p11}&
\includegraphics[scale=0.4]{p20}&
\includegraphics[scale=0.4]{p25}\\
$r=1.1$ & $r=2.0$ & $r=2.5$
\end{tabular}

\begin{tabular}{c c}
\includegraphics[scale=0.4]{p275}&
\includegraphics[scale=0.4]{p29}\\
$r=2.75$ & $r=2.9$
\end{tabular}
}

It's not always increasing, but the limiting population does always exist.

\newpage

\item Try values of the growth rate parameter in the range $2.9<r<3.1$ to determine as closely as possible
just where the single limiting population splits into a cycle of period 2.

\vspace{0.5cm}

{\centering
\begin{tabular}{c c c}
\includegraphics[scale=0.4]{p296}&
\includegraphics[scale=0.4]{p297}&
\includegraphics[scale=0.4]{p298}\\
$r=2.96$ & $r=2.97$ & $r=2.98$
\end{tabular}
}

The population appears to begin to split around $2.97<r<2.98$.

\item Verify that a cycle with period 16 is obtained with the growth rate parameter $r=3.565$.

\vspace{0.5cm}

\begin{minipage}{0.5\textwidth}
\begin{minted}{mathematica}
Population[3.565]
\end{minted}
\end{minipage}
\begin{minipage}{0.5\textwidth}
\centering
\includegraphics[scale=0.5]{div16}
\end{minipage}

\end{enumerate}
\end{document}

### 2016-02-20 resume/work.tex
mucho tarea

@@
 \begin{enumerate}[label={\Roman*}.]
+
+\item MyNewLeaf, LLC $\Bigcdot$ January 2016 - Present
+\begin{itemize}
+\item Contracted to develop Android version of iOS application
+\item Collaborated with a large team of designers, managers, and programmers
+\end{itemize}
+
 \item National Aeronautics and Space Administration $\Bigcdot$ November 2015 - Present
@@
 \item Developed Android applications for Google Glass, Vuzix, and Recon Jet headsets
-\item Used git extensively for version control
-\item Wrote code primarily in Java
+\item Worked closely with a university professor to plan and create the application
 \end{itemize}
-\item Freelance Application Developer $\Bigcdot$ July 2015 - Present
+
+\item Application Developer $\Bigcdot$ July 2015 - Present
 \begin{itemize}
+\item Contracted by MyNewLeaf to develop Android application
+\item Worked on Augmented Reality application for NASA
 \item Built a location based audio application called Audionce, written for Android and iOS.
@@
 \end{itemize}
+
 \item Christopher Newport University $\Bigcdot$ September 2013 - Present
@@
 \end{itemize}
-\item Kirk Family YMCA $\Bigcdot$ August 2009 - August 2015
-\begin{itemize}
-\item Lead Swim Instructor
-\begin{itemize}
-\item Managed a team of swim instructors that taught and oversaw various camps and lessons
-\item Taught swimming lessons to children and adults of all ages
-\item Promoted to manager in May 2015
-\end{itemize}
-\item Lifeguard
-\begin{itemize}
-\item Attended trainings for CPR and AED certifications, as well as First Aid and lifesaving skills
-\item Worked to create and maintain a safe environment for all patrons of the pool by enforcing rules and
-policies of the YMCA
-\end{itemize}
-\end{itemize}
+
+% \item Kirk Family YMCA $\Bigcdot$ August 2009 - August 2015
+% \begin{itemize}
+% \item Lead Swim Instructor
+% \begin{itemize}
+% \item Managed a team of swim instructors that taught and oversaw various camps and lessons
+% \item Taught swimming lessons to children and adults of all ages
+% \item Promoted to manager in May 2015
+% \end{itemize}
+% \item Lifeguard
+% \begin{itemize}
+% \item Attended trainings for CPR and AED certifications, as well as First Aid and lifesaving skills
+% \item Worked to create and maintain a safe environment for all patrons of the pool by enforcing rules and
+% policies of the YMCA
+% \end{itemize}
+% \end{itemize}
 \end{enumerate}

### 2016-02-20 writing_in_ethics/paper_two.tex
mucho tarea

@@
 
-\addbibresource{generic_bib.bib}
-\addbibresource{paper_two_bib.bib}
+\addbibresource{generic_bib.bib} \addbibresource{paper_two_bib.bib}
 
-\title{Social Patriotism}
-\begin{document}
-\drafth{}
-\inlinetitle
-\begin{linenumbers}
+\title{Kantian Patriots} \begin{document} \drafth{} \inlinetitle \begin{linenumbers}
 
-Governmental power has drastically increased over the past decade.
-Much of the growth of power is due to acts of terrorism that have hugely impacted the nation.
-One of the most devastating acts of terrorism that the United States has seen is the attack on the world trade center on September 11, 2001.
-As a result, the government passed the USA PATRIOT act, also known as the Patriot Act.
-Many of the provisions in the Patriot Act grant the government the ability to avoid due process.
-Under the Patriot Act, the government is given excessive power to bypass the law; Kantian ethics provide an insight that can help determine whether or not use of the liberties granted to the government are ethically sound.
+Governmental power has drastically increased over the past decade. Much of the growth of power is
+due to acts of terrorism that have greatly impacted the nation. One of the most devastating acts of
+terrorism that the United States has seen is the attack on the World Trade Centers on September 11,
+2001. As a result, the government passed the USA PATRIOT act, also known as the Patriot Act. Many of
+the provisions in the Patriot Act grant the government the ability to avoid due process. Under the
+Patriot Act, the government is given excessive power to bypass the law. Kantian ethics provide an
+insight that can help determine whether or not use of the liberties granted to the government are
+ethically sound.
 
-Section 6.6.6 of the text presents a situation in which a British news magazine, known as \textit{The Guardian}, published an article describing how the FBI indirectly requested that Verizon turn over all of its phone records.
-A key tenant of Kantian ethics is that others should never be used as a means to an end.
-The FBI's intention was never revealed, so it is hard to say whether or not they were using the phone records for personal gain or not.
-However, this fact alone is a breach of Kantian ethics; the FBI wanted to have access to private information, but refused to divulge the reason.
-Under the Categorical Imperative, moral actions should be able to be applied universally.
-The FBI did not release their private information, so, under the Categorical Imperative, the should not have taken others private information.
+Section 6.6.6 of the text presents a situation in which a British news magazine, known as
+\textit{The Guardian}, published an article describing how the FBI indirectly requested that Verizon
+turn over all of its phone records. A key tenant of Kantian ethics is that others should never be
+used as a means to an end. The FBI's intention was never revealed, so it is hard to say whether or
+not they were using the phone records for personal gain. However, this fact alone is a breach of
+Kantian ethics; the FBI wanted to have access to private information, but refused to divulge the
+reason. Under the first Categorical Imperative, moral actions should be able to be applied
+universally. The FBI did not release their private information, so, under the Categorical
+Imperative, they should not have taken others private information.
 
-The power over communications granted to the FBI is greatly increased by the Patriot Act.
-Section 6.6.2 discusses the use of National Security Letters (NSLs), namely in regards to the Library Connection of Connecticut.
-In this scenario, the FBI used a provision of the Patriot Act to demand a user's browser history from a library computer.
-It follows from the argument made about the FBI requesting Verizon phone records that NSLs are a breach of Kantian ethics; the FBI prohibits individuals and companies that receive NSL's to reveal that the FBI had contacted them in the first place.
-This scenario differs from the Verizon scenario in that the FBI wanted the information of only one individual.
-The use of the NSL aside, the FBI's actions are still unethical under Kantian analysis.
-The agency requested that the Library Connection hand over information about another individual.
-In doing so, the FBI was using the Library Connection as a means to an end, which violates the Categorical Imperative.
-In this scenario, as well as the last, it is difficult to completely analyze the situations using Kantian analysis, since the FBI's intention is not actually know.
+The power over communications granted to the FBI is greatly increased by the Patriot Act as well.
+Section 6.6.2 discusses the use of National Security Letters (NSLs), namely in regards to the
+Library Connection of Connecticut. In this scenario, the FBI used a provision of the Patriot Act to
+demand a user's browser history from a library computer. It follows from the argument made about the
+FBI requesting Verizon phone records that NSLs are a breach of Kantian ethics. the FBI prohibits
+individuals and companies that receive NSL's to reveal that the FBI had contacted them in the first
+place. The Categorical Imperative states that actions should be willed to universal laws, so it is
+unethical to demand private information and to demand that their private information not be given
+out. This scenario differs from the Verizon scenario in that the FBI wanted the information of only
+one individual. The use of the NSL aside, the FBI's actions are still unethical under Kantian
+analysis. The agency requested that the Library Connection hand over information about another
+individual. In doing so, the FBI was using the Library Connection as a means to an end, which
+violates the Categorical Imperative. In this scenario, as well as the last, it is difficult to
+completely analyze the situations using Kantian analysis, since the FBI's intention is not actually
+know.
 
-The third scenario presented by the text involves Brandon Mayfield, an attorney from Portland, Oregon.
-The FBI was presented with a partial fingerprint recovered from a terrorist attack in Spain that was linked back to Mayfield.
-The FBI then proceeded to investigate Mayfield, going as far as entering his home and seizing private information without proper legal authorization.
-This scenario differs from others in that the FBI's actions were directly aimed at Mayfield.
-They did not use anyone as a means to an end, as they did in other scenarios.
-However, they still violated Kant's Categorical Imperative.
-Kantianism focuses on the idea that moral laws should be able to be willed into universal laws, and as such, no party involved should be able to opt-out.
-Following this logic, Mayfield, as well as anyone else who wanted to, would be able to take information without legal basis from the FBI.
-This scenario differs from the previous two in that the FBI's intention is known.
-The agency sought to bring a suspected terrorist into custody.
-The methods used to reach that goal were not ethical however, which makes their actions unethical.
+The third scenario presented by the text involves Brandon Mayfield, an attorney from Portland,
+Oregon. The FBI was presented with a partial fingerprint recovered from a terrorist attack in Spain
+that was linked back to Mayfield. The FBI then proceeded to investigate Mayfield, going as far as
+entering his home and seizing private information without proper legal authorization. This scenario
+differs from others in that the FBI's actions were directly aimed at Mayfield. They did not use
+anyone as a means to an end, as they did in other scenarios. However, they still violated Kant's
+Categorical Imperative. Kantianism focuses on the idea that moral laws should be able to be willed
+into universal laws, and as such, no party involved should be able to opt-out. Following this logic,
+Mayfield, as well as anyone else who wanted to, should be able to take information without legal
+authorization as the FBI did. This scenario differs from the previous two in that the FBI's
+intention is known. The agency sought to bring a suspected terrorist into custody. The methods used
+to reach that goal were not ethical however, which makes their actions unethical.
 
-It appears that under Kantian analysis, many of the actions taken by the FBI are unethical.
-The Patriot Act allows for the government to interfere with the lives of the American population with little regard for the law, but the legality of an action does not necessarily determine whether that action is ethically right or wrong.
-It is hard to say whether or not the Patriot Act itself is ethical or not; Kantian ethics do however focus on the morality of a decision.
-The Patriot Act was put in place to protect the citizens of the United States, which Kant would argue is a morally right cause.
-Many of the actions taken by the FBI in the name of the Patriot Act were presumably done for the same reason, but there is no way to know for certain.
-\end{linenumbers}
-\newpage
-\nocite{*}
-\printbibliography
-\end{document}
+It appears that under Kantian analysis, many of the actions taken by the FBI are unethical. The
+Patriot Act allows for the government to interfere with the lives of the American population with
+little regard for the law, but the legality of an action does not necessarily determine whether that
+action is ethically right or wrong. It is hard to say whether or not the Patriot Act itself is
+ethical or not; Kantian ethics do however focus on the morality of a decision. The Patriot Act was
+put in place to protect the citizens of the United States, which Kant would argue is a morally right
+cause. Many of the actions taken by the FBI in the name of the Patriot Act were presumably done for
+the same reason, but there is no way to know for certain. \end{linenumbers} \newpage \nocite{*}
+\printbibliography \end{document}

### 2016-02-22 math_modeling/class/notes/classnotes.tex
presentation for ethics

@@
 
+\subsection{Arrow's Axioms}
+
+Kenneth Arrow has several axioms that must be taken into account when modeling voting situations.
+%page 185-187
+\begin{enumerate}
+\item The model should be transitive
+\item The model should not consider irrelevant alternatives
+\item The voters should be able to vote for whomever they choose
+\item If every voter prefers one alternative to another, so does the whole society
+\item There should be no case that leads to a dictatorship
+\end{enumerate}
+
+Among these axioms, (2) is the most logical of all. This axiom is stating that if we wish to know the
+overall preference of $x$ vs $y$, then we \textbf{only} need to consider $x$ and $y$. In the general
+case, this makes sense. The outcome should not be affected by candidate $z$, so we need not consider that
+candidate.
+
+\begin{quote}
+\textit{\textbf{Example}} Consider the case where we have the following candidates:\\
+\[\text{Rubio: A, Trump: B, Carson: C, Cruz: D, Kaish: E.}\]
+\end{quote}
+We can consider the following setup\\
+\begin{center}
+\begin{tabular}{c c c c c c}
+A & B & C & D & E & E\\
+D & E & B & C& B &C\\
+E&D&E&E&D&D\\
+C&C&D&B&C&B\\
+B&A&A&A&A&A\\
+\hline
+18 & 12 & 10 & 9 & 4 & 2
+\end{tabular}
+\end{center}
+\textbf{Plurality:} Under Plurality, A has the majority vote (18), so A wins.\\\\
+\textbf{Single Runoff:} Under the single runoff model, the two candidates with the most 1st place votes
+have a runoff. The winner of the runoff wins. The two top candidates here are A and B. In a single
+runoff, A has 18 votes and B has 37, so B wins.\\\\
+\textbf{Hare:} Under the Hare system, C wins.\\\\
+\textbf{Bourda:} Under the Bourda method, D wins.\\\\
+\textbf{Pairwise:} Under pairwise comparison, there are ${5 \choose 2} = 10$ comparisons.
+In this case, E wins the election.
+
 \end{document}

### 2016-02-22 pde_two/nonlin.tex
presentation for ethics

\section{Non Linear Equations}
\subsection{Brook's Equation}

Brooks Equation is the simplest form of an equation describing the movement of shockwaves. The equation
is given as follows:
\[
u_{t} - uu_{x} = 0.
\]
We will attempt to solve Brook's equation in a similar fashion to how we solved first order PDE's.
\begin{quote}
\textit{\textbf{Example:}} Find a solution to the equation
\[
\begin{cases}
u_{t} - uu_{x} = 0\\
u(x,0) = x
\end{cases}.
\]
\textit{\textbf{Solution:}} We will begin by noticing that
\[
{dx\over dt} = u(x,t).
\]
This implies that the solution depends on the function itself, and, as a result, we cannot solve this in
the typical way. We can, however, analyze the characteristic equation. Note that $u(x,t)$ is constant. We
know this, since
\[
{d\over dt}u(x(t),t)=
{\partial u\over\partial x}{\partial x \over \partial t} + {\partial u\over \partial t}=
0.
\]
This is not a trivial fact; it actually leads us to several conclusions that we can use:
\begin{enumerate}
\item Any associated characteristic curve is a straight line. We know this since...
\item ...the solution is constant on each characteristic curve. This leads us to...
\item ...the slope of each line is the same as the value of $u(x,t)$ on that line.
\end{enumerate}
We can use (3) to help us here. We will now take our initial condition, $\phi(x)=u(x,0)=x$, into
consideration. We know by (1) that there are infinitely many solutions that are straight lines, so
we can consider the case where we have a solution that passes through $(x,t)$ and $(x_{0},t_{0})$. Note
that $t_{0} = 0$. We can setup a relationship based on the slope $m$ as follows:
\begin{alignat*}{3}
m &= {x-x_{0}\over t-0}\qquad\qquad &&\text{by the definition of a slope}\\
&= {dx\over dt} &&\text{since the slope is constant}\\
&= u(x,t) &&\text{by the characteristic equation}\\
&= u(x_{0},0) &&\text{by (3)}\\
&= \phi(x_{0}) &&\text{by the definition of the initial value}.
\end{alignat*}
Now we have a very nice group of equalities, which we can manipulate as we please. We will take
\[
{x-x_{0}\over t-0} = \phi(x_{0}),
\]
which can be rearranged as
\[
x-x_{0} = t\phi(x_{0}).
\]
We also know that
\[
u(x,t) = \phi(x_{0}),
\]
so we would like to find a value for $\phi(x_{0})$. In our case, $\phi(x) = x$, so
\begin{align*}
&\qquad\ \ x-x_{0} = t\phi(x_{0})\\
&\implies x-x_{0} = tx_{0}\\
&\implies x = 2tx_{0}\\
&\implies x_{0} = {x\over 2t},\quad t\neq0.
\end{align*}
We had that $u(x,t) = \phi(x_{0})$, so
\[
u(x,t) = {x\over2t}.
\]
We are not done yet. Our solution for $u$ is only applicable for $t\neq0$, but we know that the initial
condition applies to $t=0$.
\end{quote}

### 2016-02-22 algorithms/projects/proj1/PJOne.java
final push for presentation

@@ -16,26 +16,34 @@ import static java.lang.System.out;
 public class PJOne {
 
     public static void algoOne(int[] a){
+        int INST_COUT = 0;
+        //Begin counting instructions
         int n = a.length;
         int j = 1;
         int[] b = new int[n];
         while(j <= n - 1){
+            INST_COUT++;
             int i = 1;
             while(i <= n - j){
+                INST_COUT++;
                 int m = i + j - 1;
                 int r = min(m + j, n);
                 int u = i;
                 for(; u < r; u++){
+                    INST_COUT++;
                     b[u - 1] = a[u - 1];
                 }
                 u = i;
                 int v = m + 1;
                 int w = i;
                 for(; w < r; w++){
+                    INST_COUT++;
                     if((u > m) || (v <= r && b[v - 1] < b[u - 1])){
+                        INST_COUT++;
                         a[w - 1] = b[v - 1];
                         v++;
                     } else {
+                        INST_COUT++;
                         a[w - 1] = b[u - 1];
                         u++;
                     }
@@ -44,6 +52,7 @@ public class PJOne {
             }
             j *= 2;
         }
+        out.println(INST_COUT);
     }
 
     public static void compute(int[] a, int[] b, int i){
@@ -67,13 +76,13 @@ public class PJOne {
     }
 
     public static void main(String[] args){
-        int [] a = new int[]{12,-12,-55,0,-16,-189};
-        for(int aa : a)
-            out.printf("%6d", aa);
-        out.println();
+        int [] a = new int[]{-12,-12,-55,0,-16,-189, -42, -98};
+        // for(int aa : a)
+            // out.printf("%6d", aa);
+        // out.println();
         algoOne(a);
-        for(int aa : a)
-            out.printf("%6d",aa);
-        out.println();
+        // for(int aa : a)
+        //     out.printf("%6d",aa);
+        // out.println();
     }
 }

### 2016-02-24 math_modeling/class/notes/gamet.tex
comitting during a tornado. yolo

\documentclass{hw}

\title{Game Theory}
\date{}
\author{}

\begin{document}
\maketitle

\section{Game of Evens and Odds}

The game of even's and odds is a two player game, where one player is the \textit{row player} and the other player is the \textit{column player}. In this game, we can discuss the concept of a \textit{payoff matrix}, which can be used to model the game.\\

Consider the matrix
\[
\begin{array}{c c c}
 & \text{Even} & \text{Odd}\\
\text{Even} & \$1 & -\$1\\
\text{Odd}&-\$1&\$1
\end{array}.
\]
This matrix is representative of a mixed strategy. Both the row and column players have changed their choices between rounds.

Imagine that we are dealing with two competing restaurants. On any given day, restaurant A, the row player, can offer a discount to try and get more customers than restaurant B. In response, B will also offer a discount. We can consider a matrix that is representative of this:
\[
\begin{array}{c c c}
 & 5\% & 30\%\\
10\% & \$5000 & -\$8000\\
20\% & -\$3000 & \$1000
\end{array}.
\]
The $-\$8000$ causes the row player to consider a \textit{max$\to$min} principal. Using this strategy, B will try to keep the best minimum, which is the bottom row $(\$1000)$. A, on the other hand wants to use a \textit{min$\to$max} principal, so A will stick to the second column. Both players want to avoid getting the $-\$8000$. This strategy is different from the mixed strategy used in the even and odd game.

\end{document}

### 2016-02-24 writing_in_ethics/abstract.tex
comitting during a tornado. yolo

\documentclass{paper}

\renewcommand{\standardhead}{%
	\noindent Group B\\
	\noindent Meeting Time: MW 1500\\
	\noindent \today
}

\renewcommand{\finalh}{%
	\begin{flushright}
		\noindent Steven Rosendahl 00885793\\
		\standardhead{}
	\end{flushright}
}

\renewcommand{\inlinetitle}{\vspace{-1.5cm}{\let\newpage\relax\maketitle}\vspace{-1.5cm}}

\addbibresource{pres_bib.bib}

\date{}
\author{}
\title{Abstract}

\begin{document}
\finalh{}
\inlinetitle{}

My presentation is on the use of Downloadable Content (DLC) in order to continue making money off of gamers.
The presentation will provide concrete examples of how companies have exploited gamers using DLC. Three
ethical views, Kantianism, Utilitarianism, and Social Contract, will be used to determine if the use of
DLC is ethical or not in today's society. The presentation will argue that it is not ethical, since game
companies not only try to market incomplete products, but also lie to consumers about having a completed
product.
\nocite{*}

\printbibliography

\end{document}

### 2016-03-01 math_modeling/test_01.tex
test review

% pg 21, #33,34,35
% pg 61, #29
% pg 100, #8,11,1,2
% pg 138, #6,14,29
% pg 176, #4,5,6,13
% pg 224, #3,4
\documentclass{hw}

\begin{document}

\section{Mathematical Models}
\begin{enumerate}
\item A bank charges 18\% annual interest on credit card balances. What is the value of an original
balance of \$1,000 after 1 year if the bank computes new balances
\begin{enumerate}
\item once a year?
\begin{quote}
We can use the equation $B_{n} = (1+r)B_{n-1}$ to calculate the balance. The original balance was \$1000,
so after 1 year, we have $B_{1} = (1+0.18)(\$1000)$, so $B_{1}=\$1180$.
\end{quote}
\item twice a year?
\begin{quote}
Similarly, we can solve the problem using the same equation as above. We know that after the first 6
months, the balance is \$1180. If we calculate again, we find that the new balance is $B_{2}=\$1392.40$.
\end{quote}
\item every three months?
\begin{quote}
After the first 8 months, the balance is \$1392.40. Calculating again yields $B_{3}=\$1643.03$.
\end{quote}
\item every week?
\begin{quote}
In this case, it helps to come up with a more generalized model, since there are 52 weeks in a year.
We can abstract the model into $B_{n}=(1+r)^nB_{0}$, so we have that $B_{52}=(1+0.18)^{52}(\$1000)$,
so $B_{52} = \$5468451.69$.
\end{quote}
\item every day?
\begin{quote}
We can again use our abstracted model, so we have $B_{365} = (1+0.18)^{365}(\$1000)$, so
$B_{365} = \$1.73\times10^{29}$.
\end{quote}
\end{enumerate}
\end{enumerate}

\section{Stable and Unstable Arms Races}
\begin{enumerate}
\item Determine the outcome of an arms race governed by the Richardson model:
\[
\begin{cases}
{dx\over dt}=10y-14x-12\\
{dy\over dt}=8x-4y-24
\end{cases}
\]
when the initial level is
\begin{enumerate}
\item (4,4)
\item (13,6)
\end{enumerate}
\begin{quote}
We will begin by finding a general solution to the provided system of ODE's. As it stands now, we
need to find a substitution that will allow us to solve the system. We can do so by finding the stable
point of the system.
\[
\left[
\begin{array}{c c | c}
-14 & 10 & 12\\
8 & -4 & 24
\end{array}
\right]
\to
\left[
\begin{array}{c c | c}
1 & 0 & 12\\
0 & 1 & 18
\end{array}
\right].
\]
We can make the substitution of $\alpha = x - 12$ and $\gamma = y - 18$. If we substitute in terms
of $\alpha$ and $\gamma$ we get the system
\[
\begin{cases}
\alpha' = 10\gamma - 14\alpha\\
\gamma' = 8\alpha - 12\gamma
\end{cases}.
\]
We can solve this. Our $A$ matrix is
\[
\left[
\begin{array}{c c}
-14 & 10\\
8 & -12
\end{array}
\right]
\implies
\left[
\begin{array}{c c}
-14-\lambda & 10\\
8 & -12-\lambda
\end{array}
\right].
\]
Our eigenvalues can be found by
\begin{gather*}
\lambda^2 -(a+d)\lambda+(ad-bc) = 0\\
\lambda^2 + 26\lambda+88 = 0\\
\lambda = -22\qquad \lambda = -4.
\end{gather*}
Our first eigenvector is given by
\[
\left[
\begin{array}{c c}
-14+22 & 10\\
8 & -12+22
\end{array}
\right]
\to
\left[
\begin{array}{c c}
8 & 10\\
8 & 10
\end{array}
\right]
\to
\left[
\begin{array}{c c}
1 & {5\over 4}\\
0 & 0
\end{array}
\right]
\implies
\vv{\lambda}_{1} =
\left[
\begin{array}{c}
1\\
-{4\over5}
\end{array}
\right].
\]
We can find the second eigenvector in a similar fashion:
\[
\vv{\lambda}_{2} =
\left[
\begin{array}{c}
1\\
1
\end{array}
\right].
\]
We now have solutions in terms of $\alpha$ and $\gamma$.
\[
\begin{cases}
\alpha = c_{1}e^{-22t}+c_{2}e^{-4t}\\
\gamma = -{4\over5}c_{1}e^{-22t} + c_{2}e^{-4t}
\end{cases}.
\]
We can back substitute, and we find that $x = \alpha + 12$ and $y = \gamma + 18$, so
\[
\begin{cases}
x = c_{1}e^{-22t}+c_{2}e^{-4t} + 12\\
y = -{4\over5}c_{1}e^{-22t} + c_{2}e^{-4t}+18
\end{cases}.
\]
Finally, we can apply our initial conditions, as specified by the problem. With the initial vector
$(4,4)$, we have that $c_{1} = 10/3$ and $c_{2} = -34/3$.
\end{quote}
\end{enumerate}

\section{Single Species Ecological Models}
\begin{enumerate}
\item The rate of growth of a certain population of bacteria in a culture is directly proportional
to the size of the population. If an experiment begins with 1,000 bacteria and one hour later
the count is 1,500 bacteria, then how many bacteria are present at the end of 24 hours?
\begin{quote}
We can model population growth with the equation $P = P_{0}e^{rt}$. To find the rate, we can solve
for $r$, which gives us
\[
r = {\log{P\over P_{0}}\over t}.
\]
Using the information provided, we have that
\[
r = {\log{1500\over 1000}\over 1} = 0.4054.
\]
After 24 hours, we will have
\[
P = 1000e^{(0.4054)(24)} = 16834112.19.
\]
\end{quote}

\item Suppose that 20 years ago the population of a town was 2,000, and that the population increased
continuously at a rate proportional to the existing population. If the population of the town is now
6,000, what has been the rate of growth?
\begin{quote}
If we take 20 years ago to be $t=0$, then we have $P_{0} = 2000$. We can similarly find the rate
as above. Doing so yields $r = 0.0549$.
\end{quote}

\item If the population of a country is undergoing exponential growth at a rate of r percent per
year, show that the population doubles every $(\log{2})/r$ years. This number is called the
``doubling time." Compute the doubling time if r = 2.
\begin{quote}
We know that population growth can be modeled by the equation $P = P_{0}e^{rt}$. We can divide and
take the natural log of both sides to get $\log{(P/P_{0})} = rt$. $P$ is the current population,
so if it double the original, $P = 2P_{0}$. This yields $\log{(2)} = rt$, or $(\log{(2)})/r = t$.
When $r=2$, we have $t \approx 0.34$.

\item Assume that the U.S. population has grown exponentially. Estimate the growth rate using
each of the following years in place of the year 1830 as done in the text.
\begin{enumerate}
\item 1800
\begin{quote}
From the text, we have that the population of the United States can be modeled by the equation
\[
P(t) = 3.929e^{0.029643(t - 1790)}.
\]
For the year 1800, we have that $P(t) = 5.28$ million.
\end{quote}
\item 1850
\[
P(t) = 23.25\text{ million}.
\]
\item 1900
\[
P(t) = 102.32\text{ million}.
\]
\item 1970
\[
P(t) = 814.44\text{ million}.
\]
\end{enumerate}
\end{quote}
\end{enumerate}

\end{document}

### 2016-03-01 number_theory/hw/hw5.tex
test review

\documentclass{hw}

\begin{document}
\begin{enumerate}
\item Use modular arithmetic to determine the day of the week 1000 days after Friday, March 11, 2016.
\begin{quote}
\[
1000 = (142)(7) + 6 \implies 1000\equiv6\text{ mod }7
\]
It will be 6 days after Friday, which is Thursday.
\end{quote}

\item Use modular arithmetic to determine the day of the week on March 11, 2036.
\begin{quote}
There are $(36-16)(365)+5=7305$ days between March 11, 2016 and March 11, 2036. We know
$7305\equiv4\text{ mod }7$, so it will be $4$ days after Friday, which is Tuesday.
\end{quote}

\item Can modular arithmetic be used to help us determine what month it is exactly 5000 days after Friday,
March 11, 2016?  If so, what month is it?  Otherwise, explain why modular arithmetic provides no help
in solving this problem.
\begin{quote}

\end{quote}

\item Without using a calculator, find the remainder when $8^{10}$ is divided by 11 and when $5^9$
is divided by 13.

\begin{minipage}[b]{0.5\textwidth}
\begin{align*}
10 &= 5 \cdot 2\\
8^{10} &= 8^{5\cdot2}\\
64^5&\equiv2^5\text{ mod }11\\
&\equiv32\text{ mod }11\\
&\equiv1\text{ mod }11
\end{align*}
The remainder is 1.
\end{minipage}
\begin{minipage}[b]{0.5\textwidth}
\begin{align*}
5^9 &= 5^3\cdot5^6\\
5^9 &\equiv 125\cdot5^6\text{ mod }13\\
&\equiv -8\cdot5^6\text{ mod }13\\
&\equiv 5\cdot5^6\text{ mod }13\\
&\equiv 5\cdot125\cdot125\text{ mod }13\\
&\equiv 125\text{ mod }13\\
&\equiv 5\text{ mod }13
\end{align*}
The remainder is 5.
\end{minipage}

\item If $a\in\ZZ$, prove that $a^3\not\equiv2\text{ mod }4$.
\begin{quote}
Since $a\in\ZZ$, we have that $a\equiv0\text{ mod }4$, $a\equiv1\text{ mod }4$, $a\equiv2\text{ mod }4$, or
$a\equiv3\text{ mod }4$.
If $a\equiv0\text{ mod }4$, then $a^3\equiv0^3\text{ mod }4$, which is equivalent to $0\text{ mod }4$.
If $a\equiv1\text{ mod }4$, then $a^3\equiv1^3\text{ mod }4$, which is equivalent to $1\text{ mod }4$.
If $a\equiv2\text{ mod }4$, then $a^3\equiv2^3\text{ mod }4,$ which is equivalent to
$8\text{ mod }4$,or $0\text{ mod }4$.
If $a\equiv3\text{ mod }4$, then $a^3\equiv3^3\text{ mod }4$, which is equivalent to
$27\text{ mod }4$, or $1\text{ mod }4$.
Therefore, it is impossible that $a^3\equiv2\text{ mod }4$.
\end{quote}
$\triangle$

\item Suppose that $m,n\in\ZZ$, with $m,n\geq2$ and write $l=lcm(m,n)$.
\begin{enumerate}
\item If $a$ and $b$ are integers such that $a\equiv b\text{ mod }l$, prove that
$a\equiv b\text{ mod }m$ and $a\equiv b\text{ mod }n$.
\begin{quote}
Suppose $a\equiv b\text{ mod }l$. Then $l\mid(a-b)$, or $a-b=lj,\ j\in\ZZ$. Since $l=lcm(m,n)$, we know that
both the prime factorization of $m$ and $n$ appear in the prime factorization of $l$. Suppose, without loss
of generality, that we factor out the prime factorization of
$m=p_{1}^{\alpha_{1}}p_{2}^{\alpha_{2}}\dots p_{k}^{\alpha_{k}},\ k\in\ZZ$ out of the prime factorization of
$l = p_{1}^{\gamma_{1}}p_{2}^{\gamma_{2}}\dots p_{q}^{\gamma_{q}},\ q\in\ZZ$. Then we are left with
$l=m\cdot{l\over m}$, where ${l\over m}\in\ZZ$. Then we have $a-b=m\cdot{l\over m}\cdot j$, which is $m$
times an integer $r$. Therefore $a-b=mr$, so $m\mid(a-b)$, and $a\equiv b\text{ mod }m$.
\end{quote}
$\triangle$

\item If $a$ and $b$ are integers such that $a\equiv b\text{ mod }m$ and $a\equiv b\text{ mod }n$, can
we conclude that $a\equiv b\text{ mod }l$? Either prove your answer or provide a counterexample.
\end{enumerate}

\item If $a,b,n\in\ZZ$, we adopt the notation that
\[
a+n\ZZ = \{a+nz\ :\ z\in\ZZ\}.
\]
If $(a+n\ZZ)\cap(b+n\ZZ)\neq\emptyset$, prove that $a+n\ZZ = b+n\ZZ$.
\end{enumerate}
\end{document}

### 2016-03-01 algorithms/projects/proj1/PJOne.java

@@ -15,13 +15,15 @@ import static java.lang.System.out;
 
 public class PJOne {
 
+    private static int COUNT_TWO = 0;
+
     public static void algoOne(int[] a){
         int INST_COUT = 0;
         //Begin counting instructions
         int n = a.length;
         int j = 1;
         int[] b = new int[n];
-        while(j <= n - 1){
+        while(j <= n){
             INST_COUT++;
             int i = 1;
             while(i <= n - j){
@@ -31,20 +33,20 @@ public class PJOne {
                 int u = i;
                 for(; u < r; u++){
                     INST_COUT++;
-                    b[u - 1] = a[u - 1];
+                    b[u] = a[u];
                 }
                 u = i;
                 int v = m + 1;
                 int w = i;
                 for(; w < r; w++){
                     INST_COUT++;
-                    if((u > m) || (v <= r && b[v - 1] < b[u - 1])){
+                    if((u > m) || (v <= r && b[v-1] < b[u-1])){
                         INST_COUT++;
-                        a[w - 1] = b[v - 1];
+                        a[w] = b[v];
                         v++;
                     } else {
                         INST_COUT++;
-                        a[w - 1] = b[u - 1];
+                        a[w] = b[u];
                         u++;
                     }
                 }
@@ -56,10 +58,13 @@ public class PJOne {
     }
 
     public static void compute(int[] a, int[] b, int i){
+        COUNT_TWO++;
         if(i > a.length - 1){
             for(int j = i; j < a.length - 1; j++){
+                COUNT_TWO++;
                 if(b[j] > 0){
                     out.println(b[j]);
+                    COUNT_TWO++;
                 }
             }
             return;
@@ -76,13 +81,18 @@ public class PJOne {
     }
 
     public static void main(String[] args){
-        int [] a = new int[]{-12,-12,-55,0,-16,-189, -42, -98};
-        // for(int aa : a)
-            // out.printf("%6d", aa);
-        // out.println();
+        int [] a = new int[]{1,2,3,4};
+        int [] b = new int[]{1,2,3,4,5,6,7,8};
+        int [] c = new int[]{-1,-2,-3,0,1,2,3,4,5,6,7,8};
         algoOne(a);
-        // for(int aa : a)
-        //     out.printf("%6d",aa);
-        // out.println();
[…]

### 2016-03-01 algorithms/projects/proj1/POne.java

@@ -1,0 +1,128 @@
+/**
+ * 420 - Project One
+ * 
+ * @author Taliahicks
+ * @author StevenR
+ * 
+ * @version 2/24/16
+ *
+ */
+
+
+public class POne {
+	
+	private static int alg2operations = 0;
+
+	public static void algOne(int[] A) {
+		int i;
+		int m;
+		int r;
+		int u;
+		int v;
+
+		// added to count operations
+		int operations = 0;
+
+		int j = 1;
+		int n = A.length - 1;
+		int[] B = new int[n];
+
+		operations += 3; // operation count
+
+		while (j <= (n - 1)) {
+			i = 1;
+			operations++; // operation count
+			while (i < (n - j)) {
+				m = i + j - 1;
+
+				operations++; // operation count
+
+				r = min(m + j, n);
+				operations++; // operation count
+				for (u = i; u < r; u++) {
+					B[u] = A[u];
+					operations++; // operation count
+				}
+
+				u = i;
+				v = m + 1;
+				operations+=2; // operation count
+
+				for (int w = i; w < r; w++) {
+					if ((u > m) || ((v < r) && B[v] < B[u])) {
+						A[w] = B[v];
+						v++;
+						operations++; // operation count
+					} 
+					else {
+						A[w] = B[u];
+						u++;
+						operations++; // operation count
+					}
+					operations++; // operation count
+				}
+				i += (2 * j);
+				operations++; // operation count
+			}
+			j = 2 * j;
+			operations++; // operation count
+		}
+
+		System.out.println("Approximate Operations: " + operations);
+	}
+
+	public static int min(int a, int b) {
+		return a < b ? a : b;
+	}
+
+	public static void compute(int[] A, int[] B, int i) {
+		int N = A.length - 1;
+		alg2operations++; //operation count
+		
+		if (i > N) {
+			for (int j = 1; j < N; j++) {
+				alg2operations++; //operation count
+				if (B[j] > 0) {
+					System.out.println(B[j]);
+					alg2operations++; //operation count
+				} // end if
+			} // end for
+			alg2operations++; //operation count
+			return;
+		} // end if
+
+		B[i] = 0;
+		compute(A, B, i + 1);
+
+		B[i] = A[i];
+		compute(A, B, i + 1);
+		alg2operations+=2; //operation count
+
+		B[i] = 0;
+		
+		
+	}
+
[…]

### 2016-03-07 math_modeling/test_01.tex
math modeling test

@@
 
+\usepackage{cancel}
+
 \begin{document}
+
+\section{Formulas/Methods}
+\begin{enumerate}
+\item Population Growth: $P = P_{0}e^{at}$
+\item Logistic Population Growth: $P^{'}=bP({a\over b}-P)$
+\item Account Balance: $B_{n}=(1+r)^nB_{0}$
+\end{enumerate}
+
+\newpage
 
@@
 
+\newpage
 \section{Stable and Unstable Arms Races}
@@
 
+\newpage
 \section{Single Species Ecological Models}
@@
 
+\newpage
+\section{Interacting Species Ecological Models}
+\begin{enumerate}
+\item Show that $y=0$, $x =e^{at}$ is a solution of both predator-prey and competitive hunters models. What
+do the orbits look like? In which direction are they traced out? Answer the same questions if it is
+specified that $x = 0$ is one of the functions in a solution.
+\begin{quote}
+The predator-prey model can be written as
+\[
+\begin{cases}
+{dx\over dt} = ax-bxy\\
+{dy\over dt} = mxy - ny
+\end{cases}.
+\]
+We can simply plug into the solutions into the equations to check if they are valid. Since $y=0$,
+$y'=0$, and therefore $0=mx(0)-n(y) = 0$. We also have that $x'=ae^{at}$, so we have
+\begin{gather*}
+ae^{at} = ae^{at} - b(e^{at})(0)\\
+ae^{at} = ae^{at}.
+\end{gather*}
+We can do the same procedure for $x=0$, and we will find that the answer is still the same.
+\end{quote}
+\end{enumerate}
+
+% \newpage
+% \section{Tumor Growth Models}
+% \begin{enumerate}
+% \item
+% \end{enumerate}
+
+\newpage
+\section{Voting Methods}
+\begin{enumerate}
+\item Consider the following distribution of votes among 55 voters.\\
+\begin{center}
+\begin{tabular}{c | c}
+Number of Voters & Ranking\\
+\hline
+18 & $t>m>h>k>c$\\
+\hline
+12 & $c>h>m>k>t$\\
+\hline
+10 & $k>c>h>m>t$\\
+\hline
+9 & $m>k>h>c>t$\\
+\hline
+4 & $h>c>m>k>t$\\
+\hline
+2 & $h>k>m>c>t$
+\end{tabular}
+\end{center}
+Determine the outcome of the election using the Plurality, Pairwise Comparison, Borda, and Hare methods.
+\begin{quote}
+\begin{enumerate}
+\item \textit{Plurality}: In this case, $t$ has the most votes, so $t$ wins the election.\\
+\item \textit{Pairwise}: We have to compare every possible combination of matches and determine the
+outcome from there. We have
+\begin{align*}
+&t\ vs.\ m:\textbf{ m} \qquad &&t\ vs.\ h:\textbf{ h} \qquad &&&t\ vs.\ k:\textbf{ k} \qquad &&&&t\ vs.\ c: \textbf{ c}\\
+&m\ vs.\ h:\textbf{ h}  &&m\ vs.\ k:\textbf{ m}  &&&m\ vs.\ c: \textbf{ m} &&&&\\
+&h\ vs.\ k:\textbf{ h}  &&h\ vs. c:\textbf{ h} &&& &&&&\\
+&k\ vs.\ c:\textbf{ k} && &&& &&&&
+\end{align*}
+In this case, $h$ has 4 votes, $m$ has 3 votes, $k$ has 2 votes, $c$ has 1 vote, and $t$ has 0 votes,
+so $h$ is the winner.\\
+\item \textit{Borda}: Under the Borda method, we consider how many points each candidate receives based
+on their ranking in each row. The first place candidate in each row receives 4 times the number of votes
+points, the next receives 3 times the number of votes points, and so on to the last candidate, who receives
+0 points.\\
+\begin{center}
+\begin{tabular}{c | c | c | c | c | c}
+& $t$ & $m$ & $h$ & $k$ & $c$\\
+\hline
+18 $\times$ & 4 & 3 & 2 & 1 & 0\\
+\hline
+12 $\times$ & 0 & 2 & 3 & 1 & 4\\
+\hline
+10 $\times$ & 0 & 1 & 2 & 4 & 3\\
+\hline
+9 $\times$ & 0 & 4 & 2 & 3 & 1\\
+\hline
+4 $\times$ & 0 & 2 & 4 & 1 & 3\\
+\hline
+2 $\times$ & 0 & 2 & 4 & 3 & 1\\
+\hline
+\textbf{Total: }& 72 & 136 & 134 & 107 & 101
+\end{tabular}
+\end{center}
+\noindent\\
+\noindent In this case, $m$ had the most votes at 136, so $m$ wins the election.\\
+\item \textit{Hare}: The Hare method focuses on finding the majority vote needed to win the election.
+There were 55 voters in this example, so the majority would be 28 votes. In the Hare method, the candidates
+have runoffs until someone reaches a majority vote.\\
+\begin{center}
+\begin{tabular}{c | c | c | c | c}
+& 1 & 2 & 3 & 4\\
+\hline
+$t$ & 18 & 18 & 18 & 18\\
+\hline
+$c$ & 12 & 16 & $\cancel{16}$ & \\
+\hline
+$m$ & 9 & $\cancel{9}$ & & \\
+\hline
+$h$ & $\cancel{6}$ & & & \\
+\hline
+$k$ & 10 & 12 & 21 & 22\\
+\hline
+\end{tabular}
+\end{center}
+\noindent\\
+\noindent Here, neither party reached 28, but $k$ has more votes than $t$, so $k$ wins.
+\end{enumerate}
+\end{quote}
+
+\end{enumerate}
+
 \end{document}

### 2016-03-08 number_theory/hw/hw5.tex
preparing paper three for ethics

@@
 \begin{quote}
-
+Modular arithmetic can be used to help solve the problem, but it cannot be used alone. Since months vary
+in the number of days, we need to know how many days are in each month. We can use mod arithmetic to
+calculate how many years go by.
+\[
+5000 = 365(13)+255\implies 13\text{ days from now}.
+\]
+March 24 and three leap years leaves us with March 27 and 252 days. From here, we can use a calendar
+to determine that the day should be sometime in December.
 \end{quote}
@@
 we conclude that $a\equiv b\text{ mod }l$? Either prove your answer or provide a counterexample.
+\begin{quote}
+Since $a\equiv b\text{ mod }m$, we know that $m\mid (a-b)$, and since $a\equiv b\text{ mod }n$, we know
+that $n\mid (a-b)$. This implies that $a-b$ is a common factor of both $m$ and $n$, so $l=lcm(m,n)$ is
+also divisible by $a-b$.
+\end{quote}
+$\triangle$
 \end{enumerate}
@@
 If $(a+n\ZZ)\cap(b+n\ZZ)\neq\emptyset$, prove that $a+n\ZZ = b+n\ZZ$.
+\begin{quote}
+Suppose we have a $w\in(a+n\ZZ)\cap(b+n\ZZ)$. Then, by definition of $a+n\ZZ$, $w=a+nz_{1}$, and,
+by definition of $b+n\ZZ$, $w=b+nz_{2}$. We can say that $a+nz_{1} = b+nz_{2}$, or that
+$a-b=nz_{2}-nz_{1}$. Consider the case where we take an $x\in a+n\ZZ$. Then, by definition of $a+n\ZZ$,
+we have that $x=a+nz_{3}$. We can rearrange $a-b=nz_{2}-nz_{1}$ in terms of $a$ so that
+$a=b+nz_{2}-nz_{1}$. Then $x=b+nz_{2}-nz_{1}+nz_{3}$, or $x=b+n(z_{2}-z_{1}+z_{3})$. Then $x\in b+n\ZZ$.
+Similarly, we can take a $y\in b+n\ZZ$. Then $y=b+nz_{4}$ and $b=a+nz_{1}-nz_{2}$, so
+$y=a+nz_{1}-nz_{2}+nz_{4}$. Therefore $y=a+n(z_{1}-z_{2}+z_{4})$ which implies that $y\in a+n\ZZ$.
+\end{quote}
+$\triangle$
+
+\item If $a, b, n\in\ZZ$ with $n\geq 2$, prove that the following statements are equivalent.
+\begin{enumerate}[i]
+\item $a\equiv b\text{ mod }n$
+\item $a\in b+n\ZZ$
+\item $b\in a+n\ZZ$
+\end{enumerate}
+Conclude that $a+n\ZZ=\{x\in\ZZ\ :\ a\equiv x\text{ mod }n\}$.
+\begin{quote}
+\underline{$(i)\implies(ii)$} Suppose $a\equiv b\text{ mod }n$. Then $n\mid(a-b)$, so $a-b=nk,\ k\in\ZZ$.
+Then $a=b+nk$, so $a\in b+n\ZZ$ by definition.\\\\
+\underline{$(ii)\implies(iii)$} Suppose $a\in b+n\ZZ$. Then $a=b+nz,\ z\in\ZZ$, and $a-b=nz$. Then
+$b=a+n(-z)$, so $b\in a+n\ZZ$.\\\\
+\underline{$(iii)\implies(i)$} Suppose $b\in a+n\ZZ$. Then $b=a+nz,\ z\in\ZZ$, so $a-b=-nz$, or
+$a-b=n(-z)$. Then $n\mid(a-b)$, and $a\equiv b\text{ mod }n$ by definition.\\\\
+By the transitive property of implication, we are done.\\\\
+Let $x\in a+n\ZZ$. Then by properties $(iii)$ and $(i)$, $a\equiv x\text{ mod }n$. Let
+$y\in\{ x\in\ZZ\ :\ a\equiv x\text{ mod }n \}$. Then by properties $(i)$ and $(iii)$, we have that
+$y\in a+n\ZZ$.
+\end{quote}
+$\triangle$
 \end{enumerate}

### 2016-03-08 writing_in_ethics/paper_three.tex
preparing paper three for ethics

\documentclass{paper}
\title{Title}

\addbibresource{generic_bib.bib}

\begin{document}
\drafth{}
\inlinetitle

\begin{linenumbers}

\end{linenumbers}
\newpage

\printbibliography

\end{document}

### 2016-03-09 math_modeling/class/notes/liquids.tex
algo homework

\section{Liquids}
Fluids can be modeled using several different techniques. The classical way involves using
partial differential equations. However, they can also be modeled with an agent based model,
known as DSMC (Direct Simulation Monte-Carlo).
\subsection{Probability}
\textit{\textbf} Find the probability of getting a 5 in 20 rolls of a die.
\begin{quote}
Our probability of not getting a 5 is $5/6$, and we only want one 5, which mean that the chance of not
getting a 5 is $(5/6)^{19}$. The chance of getting a 5 is $1/6$, and we make 20 rolls, so we have
\[
P = \left({5\over 6}\right)^{19}\cdot20\cdot\left({1\over6}\right)
\]
\end{quote}
This problem is easy to simulate on a TI-84 calculator:\\
\begin{quote}
\begin{verbatim}
randInt(1,6,60)->L1
\end{verbatim}
\end{quote}
\textit{\textbf} Given an answer key of A B B C D D A C B A, model how a student who has not studied
at all will do.
\begin{quote}

\end{quote}

### 2016-03-09 math_modeling/hw/simulation/hw.tex
algo homework

\documentclass{hw}

% \renewcommand{\arraystretch}{1.5}

\begin{document}
\begin{enumerate}
\item Simulate 36 rolls of a fair die. Give the relative frequency and the corresponding theoretical
probability of each of the outcomes and compare them.
\begin{quote}
With 36 rolls, the following numbers are produced:\\
\[\{2,6,3,3,4,5,2,1,1,2,1,1,5,5,4,3,5,4,1,1,4,3,1,1,3,3,2,6,3,1,2,5,5,2,4,4\}\]
\begin{center}
\begin{tabular}{c | c | c}
Dice Number & Frequency & Theoretical Probability\\
\hline
1 & 9 & ${1/4}$\\

2 & 6 & ${1/6}$\\

3 & 7 & ${7/36}$\\

4 & 6 & ${1/6}$\\

5 & 6 & ${1/6}$\\

6 & 2 & ${1/18}$\\
\end{tabular}
\end{center}
\end{quote}

\item Simulate 96 rolls of a pair of dice where the sum is observed. Give the relative frequency and the
corresponding theoretical probability of each of the outcomes. Make a table showing your results. Repeat
the experiment 6 times and consider the total as if 576 rolls were simulated.
\begin{quote}
Rolling two die 96 times produces the matrix
\[
\left[
\begin{array}{c c c c c c c c}
7 & 3 & 6 & 12 & 9 & 7 & 2 & 6\\
10 & 7 & 9 & 6 & 10 & 9 & 7 & 7\\
5 & 8 & 4 & 7 & 3 & 11 & 7 & 11\\
10 & 6 & 10 & 8 & 4 & 8 & 4 & 9\\
10 & 9 & 6 & 11 & 7 & 7 & 9 & 7\\
5 & 8 & 5 & 10 & 7 & 4 & 12 & 6\\
5 & 10 & 6 & 8 & 9 & 9 & 12 & 4\\
2 & 7 & 5 & 9 & 6 & 4 & 7 & 5\\
9 & 6 & 4 & 10 & 6 & 6 & 6 & 11\\
8 & 7 & 6 & 8 & 9 & 9 & 8 & 9\\
5 & 6 & 9 & 12 & 5 & 8 & 7 & 7\\
8 & 6 & 12 & 3 & 10 & 8 & 6 & 6
\end{array}
\right]
\implies
\begin{array}{c | c | c}
\text{Value} & \text{Frequency} & \text{Probability}\\
\hline
1 & 0 & 0\\

2 & 2 & {1/48}\\

3 & 3 & {1/32}\\

4 & 7 & {7/96}\\

5 & 8 & {1/12}\\

6 & 17 & {17/96}\\

7 & 16 & {1/6}\\

8 & 11 & {11/96}\\

9 & 14 & {7/48}\\

10 & 9 & {3/ 32}\\

11 & 4 & {1/ 24}\\

12 & 5 & {5/96}\\
\end{array}
\]
Running the experiment 6 times produces
\begin{center}
\begin{tabular}{c | c | c | c | c | c | c | c}
Value & Run 1 & Run 2 & Run 3 & Run 4 & Run 5 & Run 6 & Total\\
\hline
1 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
2 & 1.5 & 1 & 1 & 1 & 1 & 2.5 & 8\\
3 & 2.33 & 3 & 1.66 & 2.33 & 3.33 & 1.66 & 14.33\\
4 & 2.5 & 1.75 & 1.75 & 3 & 1.5 & 1.75 & 12.25\\
5 & 1.6 & 2.2 & 2.6 & 1.6 & 2 & 2.2 & 12.2\\
6 & 2.5 & 3.16 & 2.33 & 2 & 1 & 1.5 & 12.5\\
7 & 2 & 2.286 & 2.142 & 2 & 2.571 & 2.571 & 13.571\\
8 & 2.5 & 1.25 & 1.5 & 1.75 & 2 & 1.625 & 10.625\\
9 & 0.33 & 1.33 & 0.88 & 1.44 & 1.11 & 1.11 & 6.22\\
10 & 1 & 0.4 & 1.1 & 0.7 & 0.8 & 1 & 5\\
11 & 0.27 & 0.18 & 0.54 & 0.54 & 0.63 & 0.45 & 2.63\\
12 & 0.25 & 0.33 & 0.25 & 0.0833 & 0.25 & 0.25 & 1.4166\\
\end{tabular}
\end{center}
\end{quote}

\newpage
\item Simulate 10 free-throws for Kobe Bryant, whose free-throw average in 2008 was 81\%. How many
of the shots were successful.
\begin{quote}
We can let 1 represent a successful free-throw, and 2 represent a miss. Running one test yields
\begin{center}
\begin{tabular}{c | c}
Shot & Result\\
1 & 1\\
2 & 1\\
3 & 1\\
4 & 1\\
5 & 1\\
6 & 1\\
7 & 1\\
8 & 2\\
9 & 2\\
10 & 1
\end{tabular}
\end{center}
From the table, we can see that he made 8 shots and missed two.
\end{quote}

\item A baseball player is a 0.331 hitter. Simulate 10 at-bats for this player and tell how many hits he
gets.
\begin{quote}
We can let 0 represent a miss, and 1 represent a hit.
\begin{center}
\begin{tabular}{c | c}
Shot & Result\\
\hline
1 & 1\\
2 & 0\\
3 & 1\\
4 & 1\\
5 & 0\\
6 & 0\\
7 & 1\\
8 & 0\\
9 & 0\\
10 & 1
\end{tabular}
\end{center}
Based on this run, the batter hits 50\% of the time.
\end{quote}

\newpage
\item A student who has not studied for a 10 question multiple choice test with 4 choices among the answers
(a, b, c, d) for each question, decides to simulate such a test and answer the questions according to a
simulation in which each choice has the same probability. Assume the correct answers are
\[\text{(a, b, b, c, d, d, a, c, b, a).}\]
What is the student's score?
\begin{center}
\begin{tabular}{c | c | c}
Student & Answer & Correct\\
\hline
a & a & Yes\\
d & b & No\\
a & b & No\\
a & c & No\\
b & d & No\\
c & d & No\\
a & a & Yes\\
d & c & No\\
d & b & No\\
a & a & No\\
\hline
& & \textbf{Score:} 20\%
\end{tabular}
\end{center}

\item In sampling 4 balls at random from an urn containing 30 balls, \textit{without replacing} after each
draw, we consider the balls as numbered 1 to 30. In selecting random whole number from 1 to 30, we ignore
any number that has already been selected and continue the selection until we obtain a sample size of 4.
Assume there are 20 red balls and 10 green balls in the urn. Draw 10 samples of size 4 and tabulate the
number of red balls in each sample. Compare your results with the theoretical probability.
\begin{quote}

\end{quote}

\newpage
\item Students are queued up at the registrars office with the registration windows open at 8. There are
four open windows; students approach the first open window as they advance to the front of the queue.
Assume that 10\% of the students require 5 minutes of service time, 30\% require 7 minutes, 40\% require 10
minutes, and 20\% require 15 minutes. Simulate the service of the first 20 students in a random queue. Show
the schedule of service at the four windows (A, B, C, D), and determine how long it takes to process these
students and give the average time from 8 to leaving the service window.
\begin{quote}
One result of running the simulation yields
\[5,10,7,7,10,10,5,15,15,7,15,15,10,10,7,7,10,7,15,15\]
as the queue. Assuming that students prefer A to B to C to D, we have the following process
\begin{center}
\begin{tabular}{c | c | c | c}
A & B & C & D\\
\hline
5 & 10 & 7 & 7\\
10 & 15 & 10 & 5\\
7 & 15 & 15 & 15\\
15 & 7 & 10 & 10\\
7 & 15 & 7 & 10\\
\hline
44 & 62 & 49 & 47
\end{tabular}
\end{center}
The average time is 8.8 minutes for A, 12.4 for B, 9.8 for C, and 9.4 for D. The overall average time
is 10.1 minutes.
\end{quote}

\item Simulate the following bank queue, and give the average time needed to process each customer.
There are four bank tellers, and 40\% of customers need 3 minutes, 50\% need 5, and 10\% need 8. Assume
there are 20 customers in line.
\begin{quote}
One run of the simulation yields
\[3,5,3,3,5,5,3,5,5,3,5,5,5,5,3,3,5,3,5,8\]
as the queue. On possible outcome of the tellers is
\begin{center}
\begin{tabular}{c | c | c | c}
Teller 1 & Teller 2 & Teller 3 & Teller 4\\
\hline
3 & 5 & 3 & 3\\
5 & 3 & 5 & 5\\
5 & 3 & 5 & 5\\
5 & 5 & 3 & 3\\
8 & 5 & 3 & 5\\
\hline
26 & 21 & 19 & 21
\end{tabular}
\end{center}
The average time is 5.2 for Teller 1, 4.2 for Teller 2, 3.8 for Teller 3, and 4.2 for Teller 4. The
overall average is 4.35 minutes for each customer.
\end{quote}

\newpage
\item A gas station with four self-serve pumps has determined that 80\% of all customers completely fill
their gas tanks and the remaining 20\% fill their tank with a fixed dollar amount's worth of fuel. Suppose
that it takes an average of 5 minutes for a complete fill up and 3 minutes for a partial fill up, and that
for an hour customers arrive steadily. Simulate this process for 30 customers.
\begin{quote}
One outcome of the simulation is
\[5,5,5,5,5,5,5,3,3,5,3,3,5,5,5,5,5,5,3,3,5,5,3,3,5,5,5,5,5,3.\]
This yields the setup:
\begin{center}
\begin{tabular}{c|c|c|c}
Pump 1 & Pump 2 & Pump 3 & Pump 4\\
\hline
5 & 5 & 5 & 5\\
5 & 5 & 5 & 3\\
5 & 3 & 3 & 3\\
5 & 3 & 3 & 5\\
3 & 5 & 5 & 3\\
5 & 5 & 5 & 5\\
  & 5 & 3 &  \\
\hline
28 & 31 & 29 & 26
\end{tabular}
\end{center}
The overall average time is 4.4 minutes per customer.
\end{quote}

\item Simulate 108 rolls of three dice and show the frequency of each possible sum of the faces.
\begin{quote}
Here is the outcome of one such run
\[
\left[
\begin{array}{c c c c c c c c c}
12 & 6 & 12 & 7 & 10 & 13 & 13 & 5 & 13\\
8 & 10 & 10 & 14 & 11 & 5 & 8 & 14 & 9\\
12 & 5 & 10 & 6 & 14 & 5 & 6 & 11 & 16\\
14 & 13 & 12 & 7 & 9 & 11 & 13 & 12 & 7\\
15 & 8 & 11 & 13 & 8 & 10 & 7 & 14 & 13\\
12 & 8 & 12 & 10 & 7 & 15 & 13 & 12 & 13\\
15 & 12 & 11 & 9 & 7 & 8 & 11 & 4 & 14\\
7 & 12 & 14 & 10 & 16 & 6 & 12 & 12 & 7\\
14 & 9 & 14 & 3 & 14 & 14 & 8 & 10 & 10\\
8 & 9 & 12 & 5 & 9 & 8 & 13 & 9 & 9\\
5 & 18 & 13 & 14 & 9 & 11 & 7 & 12 & 15\\
11 & 16 & 14 & 16 & 8 & 6 & 12 & 10 & 10\\
\end{array}
\right]
\implies
\begin{array}{c | c}
\text{Dice Number} & \text{Frequency}\\
\hline
3 & 1\\
4 & 1\\
5 & 6\\
6 & 5\\
7 & 9\\
8 & 10\\
9 & 9\\
10 & 11\\
11 & 8\\
12 & 15\\
13 & 11\\
14 & 13\\
15 & 4\\
16 & 4\\
17 & 0\\
18 & 1\\
\end{array}
\]
\end{quote}
\end{enumerate}
\end{document}

### 2016-03-09 algorithms/homework/CutRod.java

@@ -1,0 +1,39 @@
+/*
+ Compute maximum revenue using dynamic programming
+ */
+
+public class CutRod {
+	//array of prices
+	static int [] p = {0,1,5,8,9,10,17,17,20,24,30};
+    //cost per cut
+    static final int COST = 1;
+
+	static int cut_rod(int n){
+        //initially, there are no cuts
+        int numCuts = 0;
+        int oq = 0;
+		//r[i] is maximum possible revenue from rod of length i
+		int [] r = new int[n+1];
+		r[0] = 0; //trivial initial condition
+		int q;
+		//use r[0] to get r[1]; use r[0..1] to get r[2], use r[0..2] to get r[3] ......
+		for (int j = 1; j <= n; j++){
+			q = -1;
+			for(int i = 1; i <= j; i++){
+                oq = q;
+				q = Math.max(q, p[i] + r[j-i]); //maximize p[i] + r[j-i]
+                if(q != oq)
+                    numCuts++;
+			}
+			r[j] = q;
+		}
+		return r[n] - (numCuts * COST);
+	}
+
+	public static void main(String[] args){
+		int n = 9;
+		System.out.println("Maximum revenue: " + cut_rod( n ));
+	}
+
+
+}

### 2016-03-10 number_theory/hw/hw6.tex
finished algorithms

\documentclass{hw}

\begin{document}
\makeheader{6}

\begin{enumerate}
\item For each of the following linear congruences of the form $ax\equiv c\mod n$, determine whether a
solution exists. If so, find a formula for all solutions and determine how many solutions there are in
$\ZZ_{n}$.
\begin{enumerate}
\item $3x\equiv 5\mod7$
\begin{quote}
The GCD of $a$ and $n$ is 1, so there is a solution since $1\mid5$, and there is only one solution.
If we solve the Diophantine equation $3x+7y=5$, we get a general solution for x as $x=3+7n$. The solution
we want is 3, since all other $n$ give a solution equal to 3 in $\ZZ_{7}$.
\end{quote}
\item $4x\equiv 9\mod12$
\begin{quote}
This congruence has no solution since $gcd(4,12)=4$, but $4\nmid9$.
\end{quote}
\item $18x\equiv 27\mod45$
\begin{quote}
The $gcd(18,45)=9$, and $9\mid27$, so there is a solution, and in fact there are 9 solutions. If we
solve the Diophantine equation $18x_{0}+45y_{0}=27$, we get that $x_{0}=-1+5n$. The nine solutions can
be found by starting with $n=1$ to $n=9$, and are $4,9,14,19,24,29,34,39,44$.
\end{quote}
\end{enumerate}

\item Let $S$ denote the number of solution to the linear congruence $ax\equiv c\mod20$. Prove that
$S\in\{ 0,1,2,3,4,5,10,20\}$.
\end{enumerate}
\end{document}

### 2016-03-10 algorithms/homework/CutRod.java

@@ -6,32 +6,30 @@ public class CutRod {
 	//array of prices
 	static int [] p = {0,1,5,8,9,10,17,17,20,24,30};
     //cost per cut
-    static final int COST = 1;
+    static int cost = 1;
 
 	static int cut_rod(int n){
         //initially, there are no cuts
         int numCuts = 0;
-        int oq = 0;
 		//r[i] is maximum possible revenue from rod of length i
 		int [] r = new int[n+1];
 		r[0] = 0; //trivial initial condition
 		int q;
 		//use r[0] to get r[1]; use r[0..1] to get r[2], use r[0..2] to get r[3] ......
 		for (int j = 1; j <= n; j++){
-			q = -1;
+            q = p[j];
 			for(int i = 1; i <= j; i++){
-                oq = q;
-				q = Math.max(q, p[i] + r[j-i]); //maximize p[i] + r[j-i]
-                if(q != oq)
-                    numCuts++;
+				q = Math.max(q, p[i] + r[j - i] - cost); //maximize p[i] + r[j-i]
 			}
 			r[j] = q;
 		}
-		return r[n] - (numCuts * COST);
+        return r[n];
 	}
 
 	public static void main(String[] args){
-		int n = 9;
+		int n = (args.length > 0) ? Integer.parseInt(args[0]) : 3;
+        //For n = 3, we expect the max revenue to be $8
+        //For n = 4, we expect the max revenue to be $9
 		System.out.println("Maximum revenue: " + cut_rod( n ));
 	}
 

### 2016-03-11 pde_two/elliptic.tex
ethics paper needs to change

\documentclass[notitlepage]{hw}
\title{Elliptic Functions}
\date{}
\author{}

\begin{document}
\maketitle

We will begin the discussion of Elliptic functions by analyzing the familiar trig functions, $\sin$ and
$\cos$. Consider the following system of ODEs:
\[
\begin{cases}
\frac{\dx}{\dt} = y\\
\frac{\dy}{\dt} = -x
\end{cases}
\]
This system has the solution $x(t)=\cos{t}$ and $y(t)=\sin{t}$. We will use this definition to formulate
the Elliptic functions.\\

We define a system of ODE's by letting $k\in(0,1)$ and $t\in\RR$ representative of time. We can define
three functions, $sn(t,k),cn(t,k),dn(t,k)$, as solutions to the system
\[
\begin{cases}
\frac{\dx}{\dt} = yz\\
\frac{\dy}{\dt} = -zx\\
{\dz\over\dt}=-k^{2}xy
\end{cases}
\implies
\begin{cases}
{d\over\dt}sn(t,k)=cn(t,k)dn(t,k)\\
{d\over\dt}cn(t,k)=-dn(t,k)sn(t,k)\\
{d\over\dt}dn(t,k)=-k^{2}sn(t,k)cn(t,k)
\end{cases}
\]
The functions are not arbitrary; they have special properties. The first one we notice is that
\[
\lim_{k\to0^{+}}sn(t,k)=\sin{t}
\qquad\lim_{k\to0^{+}}cn(t,k)=\cos{t}
\qquad\lim_{k\to0^{+}}dn(t,k)=1.
\]
In addition, we have that
\[
\lim_{k\to1^{-}}sn(t,k)=\tanh{t}
\qquad\lim_{k\to1^{-}}cn(t,k)=\sech{t}
\qquad\lim_{k\to1^{-}}dn(t,k)=\sech{t}.
\]

\end{document}

### 2016-03-11 writing_in_ethics/paper_three.tex
ethics paper needs to change

@@
 \documentclass{paper}
-\title{Title}
+\title{Catfish}
 
-\addbibresource{generic_bib.bib}
+\addbibresource{pt_bib.bib}
 
@@
 
+The television show, Catfish, is a show where the hosts, Nev and Ariel, attempt to help others find out the truth behind online relationships.
+Often times, the subject of the show is a victim of ``Catfishing," or being deceived by an anonymous person online who claims they are someone who they are not.
+The ``Catfishers" are people who create fake Facebook or Twitter profiles, use other's property, such as pictures, and hide behind the anonymity of the internet to deceive others.
+While it is unethical to lie to others under most ethical theories, a Utilitarian perspective provides insight into the ethics of many of the methods that Catfishs use to Catfish others.
+
+The key focus of Act Utilitarianism is that actions are determined ethical or unethical based on the quantitative outcome of the action \cite[74]{Ethics}.
+In almost every situation presented in the show, the victim only communicates to the Catfish over the internet, usually Facebook or a similar platform \cite{wiki}.
+Since Facebook requires users to use actual names, many Catfishers steal other's real names.
+Under Utilitarianism, this may not actually be an inherently unethical action.
+For example, a Catfish may steal a name without any intention of causing harm to the person from whom they stole the name.
+In one episode of the show, a woman undergoing Transgender surgery assumed the name of a close friend of the victim because she found the name interesting \cite{ctfs}.
+The outcome of this action is just that the Catfish goes under another name, rather than his or her own name, which has no negative effect.
+In other situations, Catfishers have been known to assume the identity of another person to harm that person.
+The negative effect of this action far outweighs the neutrality of simply picking the name.
+In that respect, stealing another's name is certainly unethical.
+
+Technology, especially the internet, has made it very easy for people to take other people's property, such as pictures and other artistic works.
+In many episodes of Catfish, Catfishers will steal the profile pictures of others.
+As Nev stated in one episode of the show, ``...that's when you know [the catfishing] is bad...
+when they use someone else's profile picture" \cite{ctfs}.
+Taking pictures from the internet is not necessarily unethical or even illegal.
+People's Facebook profile pictures are easy to download, and there is no Facebook policy that prevents users from taking other's profile pictures.
+Since Utilitarianism does not focus on the intention of the action, it may not even be unethical for a Catfish to take another's profile pictures and use them as their own.
+In order to determine the ethical implications, one must consider the amount of good and bad produced by the action.
+Posing as someone else only does harm if the Catfish is attempting to make the victim look bad.
+If the Catfish does not defame or boost the victim's character, then the Catfish has not done anything unethical.
+Moreover, if the Catfish makes the original picture's owner look better, then the amount of good done actually outweighs the bad.
+
+Catfish often try to remain anonymous online.
+
+Not every episode of Catfish ends with deceit.
+In some cases, the Catfish is struggling with other issues that end up misinterpreted by the other party.
+One episode in particular resulted in the marriage of two lovers after they met following an eight year online relationship.
+
 \end{linenumbers}

### 2016-03-11 algorithms/homework/CutRod.java

@@ -1,12 +1,13 @@
 /*
- Compute maximum revenue using dynamic programming
- */
+Steven Rosendahl
+Compute maximum revenue using dynamic programming
+*/
 
 public class CutRod {
 	//array of prices
 	static int [] p = {0,1,5,8,9,10,17,17,20,24,30};
     //cost per cut
-    static int cost = 1;
+    static int cost = 2;
 
 	static int cut_rod(int n){
         //initially, there are no cuts

### 2016-03-13 writing_in_ethics/paper_three.tex
finished paper three draft

--- 2016-03-08 writing_in_ethics/paper_three.tex
@@
 \documentclass{paper}
-\title{Title}
+\title{Apple Security}
 
-\addbibresource{generic_bib.bib}
+\addbibresource{pt_bib.bib}
 
@@
 
+Apple has always prided itself in the security of its products, even to the point that the majority of apple users feel safe using using apple computers without a virus protection software.
+As the technology has become more entangled, the need to keep users safe has increased.
+When Apple released their new operating system, OS X El Capitan, they decided to remove user access to the system's most important files; the current security system does not allow non-Apple verified products to install anything in the top level directories of the operating system.
+Under Utilitarian ethics, it is unethical for Apple to take away a user's access to his or her own operating system.
+
+Apple's removal of top level permissions appears at first to be a smart choice on their part.
+Apple reasoned that they wanted to protect their users by not allowing malicious programs masquerading as friendly applications.
+Under Utilitarianism, there is a lot of good produced by doing this.
+Apple is preventing common users from acquiring viruses due to not being computer literate.
+This action produces a lot of good for both Apple, who gains more support from their users, and the users themselves, who are secure.
+However, not all Apple users are developers; in fact the majority of programmers use Apple machines \cite{wiki}.
+This is due to the accessibility of programming tools available for Apple.
+In addition, OS X is a UNIX based architecture.
+UNIX based architectures are meant to allow users to have complete control over the operating system.
+By removing the access to the top level directories, developers are forced to forgo installing many  open source programs that are not approved by Apple.
+This hurts developers who produce new software and do not want to buy in to the Apple Developer's Program.
+In addition, many tools that developers rely on are blocked from installation, which affects the productivity of developers.
+Since Apple's user base is more developers than common user, this produces a lot of bad.
+
+Apple's implementation of the security features included in El Capitan was not widely known when the new operating system was announced \cite{sec}.
+Apple forces developers who want to use the Apple Store to distribute applications to sign up for a developers account.
+The account is not free; it costs developers \$100 per year.
+It appears that Apple did not want to publicize the security features since it would drive developers away from using the new operating system, and therefore depriving apple of developers who could be forced to pay to use the Apple Store.
+Under Act Utilitarianism, this is not necessarily a bad action.
+From an individual perspective, Apple does need to make a profit, and they do distribute, for free, the operating system and the majority of the tools developers need.
+Under an Act Utilitarianism, the goods provided by Apple to the developers contributes more good than overall harm done by the forced security measures.
+Rule Utilitarianism, however, focuses on the ethical actions being willed into universal rules \cite[75]{Ethics}.
+Under Rule Utilitarianism, Apple withheld important information about its operating system from both users and developers alike.
+If all the important information about an operating system was omitted, then no one would use that operating system.
+Overall, Apple is not only hurting itself, but also its all of its consumers.
+
+The security measures implemented in El Capitan have been attributed to Apple becoming more irrelevant as other competitor companies have grown \cite{guard}; some developers have claimed that Apple's new security measures are a ploy to convince many users to continue to buy from the technology giant \cite{verge}.
+Under Utilitarianism, Apple's actions may be considered ethical if the quantity being measured is money.
+If Apple cannot keep users, then the company will go out of business, and users will be forced to completely change operating systems.
+This will produce a lot of bad, but not as much as a company lying to its users.
+
+Apple's decision to block access to root files has caused many developers to completely rework their products in order to allow users to continue to use them.
+Under both Act and Rule Utilitarianism, this can be seen as unethical, largely due to the fact that Apple did not publicize the new security measures included in the operating system.
+Apple, however, has promised a way to remove these features as the operating system becomes more popular.
+While this may help Apple make amends with the developers, the damage has already been done, and many developers have already designed their programs to work with the new security updates.
+
 \end{linenumbers}

### 2016-03-13 math_modeling/class/c/mosq.c

@@ -1,0 +1,26 @@
+/**
+This program attempts to solve the probability that a person lost in a forest will
+return to a river bank whle doing a ranndom walk
+*/
+#include <stdio.h>
+#include <stdlib.h>
+#include <math.h>
+
+#define PI 3.14159265358979323846
+
+int main(){
+    //Generate random numbers with rand();
+    //In order to return to the river, the man needs to follow
+    // l * sin(theta1)
+    // The next day is
+    // l * sin(theta2)
+    // The sum
+    // lsin(theta1) + lsin(theta2) <= 0
+    // sin(theta1) + sin(theta2) <= 0
+
+    double theta1 = rand(); //* PI;
+    double theta2 = rand(); //* 2 * PI;
+    printf("theta1 = %f, theta2 = %f\n", theta1, theta2);
+
+    return 0;
+}

### 2016-03-15 math_modeling/hw/chw/hw.tex
number theory

\documentclass{hw}

\usepackage{minted}

\begin{document}
\begin{enumerate}
\item The following equation for the temperature $T=T(t)$ represents a spherical thermocouple with
convective conditions and includes radiation exchange with its surrounding walls
\[
c_{2}{dT\over\dt}=-(T-T_{\infty}+c_{1}(T^{4}-T^{4}_{sur})).
\]
In this equation, $c_{1}=1.27575\times10^{-10}K^{-3}$ and $c_{2}=0.991667s$. If it is assumed that
$T_{\infty}=473.15$, $T_{sur}=673.15K$, and $T(0)=298.15K$, determine the time $t_{s}$ at which
$T(t_{s})=490.85K$.
\end{enumerate}
\end{document}

### 2016-03-15 number_theory/hw/hw6.tex
number theory

@@
 \begin{quote}
-The GCD of $a$ and $n$ is 1, so there is a solution since $1\mid5$, and there is only one solution.
-If we solve the Diophantine equation $3x+7y=5$, we get a general solution for x as $x=3+7n$. The solution
-we want is 3, since all other $n$ give a solution equal to 3 in $\ZZ_{7}$.
+The $gcd$ of $a$ and $n$ is 1, so there is a solution since $1\mid5$, and there is only one solution.
+If we solve the Diophantine equation $3x+7y=5$, we get a general solution for x as $x=-3+7n$. The
+solution we want is 4, since all other $n$ give a solution equal to 4 in $\ZZ_{7}$.
 \end{quote}
@@
 \item Let $S$ denote the number of solution to the linear congruence $ax\equiv c\mod20$. Prove that
-$S\in\{ 0,1,2,3,4,5,10,20\}$.
+$S\in\{ 0,1,2,4,5,10,20\}$.
+\begin{quote}
+We know that there are either 0 or $gcd(a,20)$ solutions to the congruence $ax\equiv c\mod20$.
+Suppose we have the set $A=\{0,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20\}$. We know that for
+these numbers, we will have a unique $gcd$ with 20, and in fact all possible $gcd's$ will be
+represented in this set. If we find the $gcd$ of a relatively prime number $r$ and 20, then
+$gcd(r,20)=1$, so it is possible to have only one solution. In addition, $gcd(1,20)=1$, so we can
+eliminate this number as well. Now we can consider the set
+$B=A\backslash\{3,7,9,11,13,17,19\}=\{0,2,4,5,6,8,10,12,14,15,16,18,20\}$. For both 0 and 20,
+$gcd(0,20)=gcd(20,20)=20$, so it is possible to have 20 solutions. Now consider the set
+$C=B\backslash\{0,20\}=\{2,4,5,6,8,10,12,14,15,16,18\}$. We can find all the elements of $C$ such that
+$gcd(20,c)=2$, which are $\{2,6,14,18\}$, which leaves us with the set $D=\{4,5,8,10,12,15,16\}$.
+Now we can consider all the elements with which give us a $gcd$ of 4, which are $\{4,8,12,16\}$.
+This leaves us with the set $E=\{5,15\}$. If we take any element of $E$ and find $gcd(e,20)$, we
+will get 5. Now we are only left with $gcd(10,20)=10$, which provides us with 10.
+\end{quote}
+
+\item Suppose that $a,n\in\ZZ$ with $n\geq3$ and $gcd(a,n)>1$. Prove that there exist at least two
+non-zero points $c\in\ZZ_{n}$ such that $ax\equiv c\mod n$ has no solutions.
+\begin{quote}
+Suppose $ax\equiv c\mod n$. Then there are either $gcd(a,n)$ solutions, or 0 solutions. If there are
+0 solutions, then there are $n$ number of non-solutions $c\in\ZZ_{n}$. Now suppose we have $gcd(a,n)$
+solutions. We know that $gcd(a,n)>1$, so we can let $c_{1}=1$. In this case, the only thing that
+divides $c_{1}$ is 1, but $gcd(a,n)>1$, so there is no solution. We can also choose $c_{2}=n-1$. In
+this case, $n-1\equiv-1\mod n$, and the only thing that divides $-1$ is 1. Since $gcd(a,n)>1$, it does
+not divide $-1$, and therefore there are no solutions.
+\end{quote}
+
+\item Determine whether each given point is a unit in the given $\ZZ_{n}$. If so, find its
+multiplicative inverse. If not, explain why it fails to be a unit.
+\begin{enumerate}
+\item $3\in\ZZ_{6}$
+\begin{quote}
+$3\equiv1\mod6$ implies that $3x+6y=1$. However, $gcd(3,6)=3$, which does not divide 1. Therefore, 3
+is not a unit.
+\end{quote}
+\item $7\in\ZZ_{12}$
+\begin{quote}
+The greatest common divisor of 7 and 12 is 1, so 7 is a unit in $\ZZ_{12}$. We have the relationship
+that $7x\equiv1\mod12$, so $7x+12y=1$. One solution to this Diophantine equation is $x_{0}=-5$, so all
+solutions can be expressed as $x=-5+12n,\ n\in\ZZ$. When $n=1$, $x=7$, which means that $7$ is its own
+inverse.
+\end{quote}
+\item $13\in\ZZ_{18}$
+\begin{quote}
+$gcd(13,18)=1$, so 13 is a unit in $\ZZ_{18}$. We can form the relationship $13\equiv1\mod18$, so we
+have that $13x+18y=1$. We have one solution, $x_{0}=7$, so all solutions $x$ can be expressed as
+$x=7+18n,\ n\in\ZZ$. If we choose $n=0$, then we have $x=7$, so 7 is the inverse of 13.
+\end{quote}
+\end{enumerate}
+
+\item Suppose that $p$ is prime.
+\begin{enumerate}
+\item Prove that the set $\{0,1,2,\dots,p^{2}-2,p^{2}-1\}$ contains exactly $p(p-1)$ elements which
+are relatively prime to $p$. Conclude that $\ZZ_{p^{2}}$ contains exactly $p(p-1)$ units.
+\begin{quote}
+Suppose we consider the set containing all non-units of $\ZZ_{p^{2}}$. This set contains elements of
+the form $\{0,p,2p,3p,\dots,p(p-1)\}$. We know this set contains $p$ elements, since it contains
+$p-1$ multiples of $p$, and 0. If we subtract the number of elements in the $\ZZ_{p^{2}}$ from the
+number of non-units, we get $p^{2}-p$, or $p(p-1)$.
+\end{quote}
+\item How many units does $\ZZ_{p^{n}}$ have? Prove your answer.
+\begin{quote}
+We can consider the set of non-units in $\ZZ_{p^{n}}$. This yields the set
+\[\{0,p,2p,\dots,p^2,2p^{2},\dots,p^{3},\dots,p^{n-1}(p-1)\}.\] We know this set has size $p^{n-1}$,
+since it contains all multiples of powers of $p$ up to the $n-1$ power. Again, we can subtract the
+sizes of the entire set and the set of non-units, and we get $p^{n}-p^{n-1}=p^{n-1}(p-1)$.
+\end{quote}
+\end{enumerate}
+
+\item Suppose $p$ and $q$ are distinct primes. Prove that $\ZZ_{pq}$ contains exactly $(p-1)(q-1)$
+units.
+\begin{quote}
+We can consider the set of non-units of $\ZZ_{pq}$, which takes the form
+\[\{0,p,q,\dots,np,mq\},\ n<p,\ m<q.\]
+If we can determine the size of the set of non-units, then we can find the size of the units. We know
+that the non-units contain all multiples of $p$, $np$ for some $n\in\ZZ$, and the set also contains
+all multiples of $q$, $mq$ for some $m\in\ZZ$. We know we have up to $p$ multiples of $p$, and $q$
+multiples of $q$, so there are $q+p$ multiples in total. However, when $n=m$, we have a duplicate
+multiple, so the set of non-units contains $q+p-1$ elements. If we subtract the total size of the
+set from the size of the set of non-units, we have $pq-p-q+1=(p-1)(q-1)$.
+% Suppose $a$ is a unit of $\ZZ_{pq}$. We have that $ax\equiv1mod(pq)$, so $pq\mid(ax-1)$. By definition
+% of prime division, we have that $p\mid(ax-1)$ and $q\mid(ax-1)$. It follows that $ax\equiv1\mod p$,
+% so $gcd(a,p)=1$ for all $p\in\ZZ_{p}$. This means that every element of $\ZZ_{p}$ is a unit, so there
+% are $p-1$ units. Similarly, we have that there are $q-1$ units in $\ZZ_{q}$. Therefore there are
+% $(p-1)(q-1)$ units in $\ZZ_{pq}$.
+\end{quote}
 \end{enumerate}

### 2016-03-15 algorithms/projects/proj2/StringSplit.java

@@ -1,0 +1,22 @@
+public class StringSplit {
+
+    public static int splitString(String s, int[] splits, int cost){
+        // int[] costs = new int[splits.length]; //the array of costs
+        if (cost == 1)
+            return 0;
+        else {
+            int option = Integer.MAX_VALUE;
+
+            option = Math.min()
+        }
+    }
+
+    // private static int splitStringRec(String s, int splits){
+    //
+    // }
+
+    public static void main(String[] args){
+
+    }
+
+}

### 2016-03-15 math_modeling/class/c/mosq.c

@@ -18,9 +18,10 @@ int main(){
     // lsin(theta1) + lsin(theta2) <= 0
     // sin(theta1) + sin(theta2) <= 0
 
-    double theta1 = rand(); //* PI;
-    double theta2 = rand(); //* 2 * PI;
-    printf("theta1 = %f, theta2 = %f\n", theta1, theta2);
+    double theta1 = rand() * PI;
+    double theta2 = rand() * 2 * PI;
+
+    
 
     return 0;
 }

### 2016-03-15 math_modeling/hw/chw/math.c

@@ -1,0 +1,7 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <math.h>
+
+int main(int argc, const char* argv[]){
+    
+}

### 2016-03-16 math_modeling/class/c/crofton.c
nothing like some fun C code

@@ -1,0 +1,33 @@
+/* Heres a comment for Dr. Martin's Math440 students!!!!!!!!!!!!!!!!!! */
+/* Math440 student's -- u must compile with the line gcc crofton.c -lm */
+/* Please notice the -lm in the above line                             */
+/***********************************************************************/
+
+#include <math.h>
+#include <stdio.h>
+#include <stdlib.h>
+
+void main()
+{
+  double pi;
+  double theta1;
+  double theta2;
+  int i;
+  int total = 0;
+  double frandinc, frandexc;
+  
+  pi = 4.*atan(1.);
+  
+  for (i=1; i<=100000; i++){
+    frandinc = (double)rand()/(double)RAND_MAX;
+    frandexc = (double)rand()/((double)(RAND_MAX)+1);
+
+    theta1 = pi * frandinc;
+    theta2 = 2.*pi * frandexc;
+
+    if( sin(theta1) + sin(theta2) <= 0)
+      total = total + 1;
+  }
+
+  printf("probability is %f", total/100000.);
+}

### 2016-03-16 math_modeling/class/c/montecarlo.c

@@ -1,0 +1,24 @@
+#include <stdio.h>
+#include <stdlib.h>
+
+double myfunc (double x){
+    return x*x;
+}
+
+void main(){
+    int i;
+    int total=0;
+    double a = 4.;
+    double b = 7.;
+    double m = 100.;
+    double frandinc1,frandinc2;
+
+    for(i = 1; i <= 100000; i++){
+        frandinc1 = a + (b - a) * (double)rand()/(double)RAND_MAX;
+        frandinc2 = m * (double)rand()/(double)RAND_MAX;
+        if(frandinc2 <= myfunc(frandinc1))
+            total = total + 1;
+    }
+
+    printf("value is %f ", total*(b-a)*m/100000);
+}

### 2016-03-16 math_modeling/class/c/mosq.c

@@ -20,8 +20,8 @@ int main(){
 
     double theta1 = rand() * PI;
     double theta2 = rand() * 2 * PI;
-
     
+
 
     return 0;
 }

### 2016-03-16 math_modeling/hw/chw/math.c

@@ -1,7 +1,50 @@
 #include <stdio.h>
 #include <stdlib.h>
 #include <math.h>
+#include <unistd.h>
 
 int main(int argc, const char* argv[]){
-    
+
+    float temps [10000];
+    temps[0] = 298.15;
+
+    float starttime = 0;
+    float deltat;
+    if(argc > 1)
+        deltat = atof(argv[1]);
+    else
+        deltat = 0.001;
+    float c2 = 0.991667;
+    float c1 = 1.27575 * powf(10,-10);
+    float tinf = 473.15;
+    float tsur = powf(673.15,4);
+
+    FILE* fp = fopen("data.txt","w+");
+
+    int i = 0;
+    for(i = 0; i < 10000 && temps[i] < 490.85; i++){
+        if(i % 10 == 0){
+            fprintf(fp, "%f\n", temps[i]);
+        }
+        temps[i + 1] = (-1) * (deltat/c2) * (temps[i] - tinf +
+            c1 * (pow(temps[i],4) - tsur)) + temps[i];
+        starttime += deltat;
+    }
+
+    fclose(fp);
+
+    int status;
+    int pid = fork();
+    if(pid == 0){
+        execl("/usr/local/bin/RScript", "/usr/local/bin/RScript",
+            "/Users/sjyn/Documents/LahTech/math_modeling/hw/chw/plot.R", (char *)NULL);
+    } else if(pid > 0) {
+        printf("The temperature at time %f is %f\n", starttime, temps[i]);
+    } else {
+        exit(2);
+    }
+    if(waitpid(pid, &status, 0) != pid)
+        status = -1;
+
+    return 0;
 }

### 2016-03-16 math_modeling/hw/chw/plot.R

@@ -1,0 +1,3 @@
+#!/usr/bin/env RScript
+invisible(data <- scan("data.txt"))
+plot(data, type="b")

### 2016-03-16 writing_in_ethics/paper_three.tex
ethics paper

@@
 
-Apple has always prided itself in the security of its products, even to the point that the majority of apple users feel safe using using apple computers without a virus protection software.
+Apple has always prided itself in the security of its products, even to the point that the majority of Apple users feel safe using using Apple computers without a virus protection software.
 As the technology has become more entangled, the need to keep users safe has increased.
@@
 Apple's removal of top level permissions appears at first to be a smart choice on their part.
-Apple reasoned that they wanted to protect their users by not allowing malicious programs masquerading as friendly applications.
+Apple reasoned that they wanted to protect their users by not allowing malicious programs masquerading as friendly applications from installing themselves in the heart of the operating system.
 Under Utilitarianism, there is a lot of good produced by doing this.
@@
 In addition, many tools that developers rely on are blocked from installation, which affects the productivity of developers.
-Since Apple's user base is more developers than common user, this produces a lot of bad.
+Since Apple's user base is more developers than common user, this produces a lot of harm.
+
+Rule and Act Utilitarianism appear to favor the developer's who use OS X, but developers are not the only consumers of Apple products.
+To a general consumer, Apple has made a safer operating system; but has also denied users access to critical files of the operating system.
+Apple computers, though less likely to contract viruses, are still able to be infected.
+Just because a computer has locked a user out of a directory does not mean that a virus cannot find its way in.
+In this case, it is significantly harder for businesses that repair Apple computers to fix the problem.
+While employees at the Apple store may be able to handle such a situation, not everyone has access to an Apple store.
+Therefore, this action actually produces more harm then good.
 
@@
 The account is not free; it costs developers \$100 per year.
-It appears that Apple did not want to publicize the security features since it would drive developers away from using the new operating system, and therefore depriving apple of developers who could be forced to pay to use the Apple Store.
+It appears that Apple did not want to publicize the security features since it would drive developers away from using the new operating system, and therefore depriving Apple of developers who could be forced to pay to use the Apple Store.
 Under Act Utilitarianism, this is not necessarily a bad action.
@@
 
-The security measures implemented in El Capitan have been attributed to Apple becoming more irrelevant as other competitor companies have grown \cite{guard}; some developers have claimed that Apple's new security measures are a ploy to convince many users to continue to buy from the technology giant \cite{verge}.
+% The security measures implemented in El Capitan have been attributed to Apple becoming more irrelevant as other competitor companies have grown \cite{guard}; some developers have claimed that Apple's new security measures are a ploy to convince many users to continue to buy from the technology giant \cite{verge}.
 Under Utilitarianism, Apple's actions may be considered ethical if the quantity being measured is money.

### 2016-03-16 math_modeling/hw/chw/plot.R

@@ -1,3 +1,3 @@
 #!/usr/bin/env RScript
-invisible(data <- scan("data.txt"))
+data <- scan("data.txt")
 plot(data, type="b")

### 2016-03-20 pde_two/high_dimen.tex
working on a lot

\section{High Order PDEs}

We can begin by talking about the heat equation in 2D space. We have the following
\[
u_{t}=ku_{xx},
\]
which, under certain conditions, yields the general solution
\[
u(x,t) = \sum_{n=0}^{\infty}b_{n}e^{-k\left({n\pi\over l}\right)^{2}t}\sin{\left({n\pi x\over l}\right)}.
\]
If we consider the laplacian, $\nabla^{2}u$, we can express the heat equation as
\begin{align*}
u_{t}&=k\nabla^{2}u\\
&=k(u_{xx} + u_{yy})\text{ in }\RR^{3}.
\end{align*}
We can now begin to solve a three-dimensional heat equation.
\begin{quote}
\textbf{\textit{Example:}} Solve the following heat equation for $u(x,t)$:
\[
\begin{cases}
u_{t} = a^{2}(u_{xx} + u_{yy})\\
u(x,t) = f(x,y)\\
u_{2D} = 0\qquad\text{(there is no external source of energy in the system)}
\end{cases}.
\]
\textit{\textbf{Solution:}} We begin by noticing that this equation is taking place on a cylinder.
much like the process when solving Laplace's equation on a disk, we can transform our PDE from
cartesian coordinates to polar coordinates. The time variable, $t$ is not affected by this
transformation, and we know
\[
\nabla^{2} u=u_{xx} + u_{yy} = u_{rr} + {1\over r}u_{r} + {1\over r^{2}}u_{\theta\theta}.
\]
Applying our transformation yields the new PDE
\[
\begin{cases}
u_{t}=a^{2}\left(u_{rr} + {1\over r}u_{r} + {1\over r^{2}}u_{\theta\theta}\right)\\
u_{t=0}=f(r,\theta)\\
u_{r=r}=0
\end{cases}.
\]
We will now have a general solution of $u(r,\theta,t)$, but we cannot jump immediately to a Fourier
series solution. Instead, we will use separation of variables to provide us with a solution. We have
\begin{gather*}
u = V(r,\theta)T(t)\\
{T^{'}\over a^{2}T}= {V_{rr} + {1\over r}V_{r}+{1\over r^{2}}V_{\theta\theta}\over V} = -k^{2}.
\end{gather*}
We now have two ODEs to deal with:
\[
T'+a^{2}k^{2}T=0
\qquad\qquad\text{and}\qquad\qquad
V_{rr} + {1\over r}V_{r}+{1\over r^{2}}V_{\theta\theta}+k^{2}V=0
\]
We can again apply separation of variables to $V(r,\theta)$, and we find that we have two more resulting
ODEs.
\[
\begin{cases}
\theta''+\mu\theta=0\\
r^{2}R''+rR'+(k^{2}r^{2}-\mu)R=0
\end{cases}.
\]
The equation involving $\theta$ is a simple ODE; the equation involving $R$ will take a little more
work to solve. We will discover a method to solve this ODE in the next section.
\end{quote}

\subsection{Power Series Solutions to ODEs}

From the previous problem, we saw that we need to develop a method to solve
\begin{gather*}
x^{2}y''+xy'+(x^{2}-n^{2})y = 0\\
p(x)y''+q(x)y'+r(x)y = 0.
\end{gather*}
We will attempt to solve this using power series solutions to ODEs. There is, however, an issue that
we will face. Since $p(x)=x^{2}$, we cannot recover $y''$ near zero ($p(0)=0$). The regular power
series expansion we would use would take the form
\[
y(x) = \sum_{n=0}^{\infty}a_{n}x^{n},
\]
but we cannot use this method since $y''$ is lost near zero. We can, however modify our sequence
in such a way as to preserve $y''$. Doing so gives us the sequence
\[
y(x)=x^{\alpha}\sum_{n=0}^{\infty}a_{n}x^{n},
\]
where $\alpha$ is called the characteristic value. The strategy here will be to determine the values
of $\alpha$, and then determine the value of $a_{n}$.
\begin{quote}
\textbf{\textit{Example:}} Determine the solution to the following ODE
\[
x^{2}y''+x^{2}y'+ry=0.
\]
\textit{\textbf{Solution:}} We start by saying
\[
y(x) = \sum_{n=0}^{\infty}a_{n}x^{n+\alpha},
\]
so
\begin{align*}
y'(x) &= \sum_{n=0}^{\infty}(n+\alpha)a_{n}x^{n+\alpha-1}\\
y''(x) &= \sum_{n=0}^{\infty}(n+\alpha)(n+\alpha-1)a_{n}x^{n+\alpha-2}.
\end{align*}
We now have enough to substitute back into our original ODE.
\begin{gather*}
x^{2}\sum_{n=0}^{\infty}(n+\alpha)(n+\alpha-1)a_{n}x^{n+\alpha-2} +
x^{2}\sum_{n=0}^{\infty}(n+\alpha)a_{n}x^{n+\alpha-1}+
r\sum_{n=0}^{\infty}a_{n}x^{n+\alpha} = 0\\
\end{gather*}
\end{quote}

### 2016-03-20 algorithms/projects/proj2/StringBreaking.java

@@ -1,0 +1,95 @@
+import java.io.File;
+import java.io.IOException;
+import java.util.Scanner;
+
+/**
+ * This algorithm finds the optimal solution to a string breaking problem
+ *
+ * @author Talia Hicks
+ * @author Steven Rosendahl
+ * @version 03.15.2016
+ *
+ */
+public class StringBreaking {
+
+	public static int findBreakPoint(int n, int[] breakpoints) {
+		if (breakpoints.length == 1)
+			return 0;
+
+		else {
+			int median = 0;
+			int minDiff = n / 2 - breakpoints[0];
+			for (int i = 1; i < breakpoints.length; i++) {
+				if ((n / 2 - breakpoints[i]) < minDiff) {
+					minDiff = n / 2 - breakpoints[i];
+					median = i;
+				} // end if
+			} // end for
+			return median;
+		}
+	}
+
+	public static int stringBreakingAlg(String s, int[] breakpoints) {
+		int optsolution = 0;
+		int n = s.length(); // size of string
+		int breaks = breakpoints.length; // number of breakpoints
+
+		int x = 0;
+
+		int breakpoint = findBreakPoint(n, breakpoints);
+		System.out.println("Breakpoint: " + breakpoints[breakpoint]);
+		System.out.println("Cost of break: " + n);
+		optsolution += n;
+		System.out.println("Current total optimal cost: " + optsolution);
+		breaks--;
+
+		int breakpoint1 = breakpoint;
+		int breakpoint2 = breakpoint;
+
+		while (x <= breaks)
+		{
+
+			int[] brokenArray1 = new int[breakpoint1 + 1];
+			for (int i = 0; i < brokenArray1.length; i++) {
+				brokenArray1[i] = breakpoints[i];
+			}//end for
+
+			breakpoint1 = findBreakPoint(brokenArray1.length, brokenArray1);
+
+			System.out.println("Breakpoint: " + brokenArray1[breakpoint]);
+			System.out.println("Cost of break: " + brokenArray1.length);
+
+			optsolution += brokenArray1.length;
+			System.out.println("Current total optimal cost: " + optsolution);
+
+			breaks--;
+
+			int[] brokenArray2 = new int[n - (breakpoint2 + 1)];
+            // System.out.printf("SIZE OF ARRAY: %d\n", brokenArray2.length);
+			for (int i = 0; i < brokenArray2.length; i++) {
[…]

### 2016-03-20 algorithms/projects/proj2/StringSplit.java

@@ -1,22 +1,81 @@
+import java.util.Arrays;
+
+
 public class StringSplit {
 
-    public static int splitString(String s, int[] splits, int cost){
-        // int[] costs = new int[splits.length]; //the array of costs
-        if (cost == 1)
-            return 0;
-        else {
-            int option = Integer.MAX_VALUE;
+    static int[] optcuts;
 
-            option = Math.min()
-        }
-    }
-
-    // private static int splitStringRec(String s, int splits){
-    //
+    // public static int splitString(String s, ArrayList<Integer> splits){
+    //     if(splits.size() == 0) {
+    //         return s.length();
+    //     } else {
+    //         // int mid = splits.get(splits.size() / 2);
+    //         int mid = splits.get(splits.size() - 1);
+    //         // int mid = splits.get(0);
+    //         System.out.println("Splitting " + s + " at index " + mid);
+    //         ArrayList<Integer> less = new ArrayList<>();
+    //         for(Integer i : splits){
+    //             if(i < mid)
+    //                 less.add(Math.abs(i - mid));
+    //         }
+    //         ArrayList<Integer> max = new ArrayList<>();
+    //         for(Integer i : splits){
+    //             if(i > mid)
+    //                 max.add(i);
+    //         }
+    //         return s.length() + Math.min(
+    //             splitString(s.substring(0, mid), less),
+    //             splitString(s.substring(mid, s.length() - 1), max)
+    //         );
+    //         // return min;
+    //     }
     // }
 
-    public static void main(String[] args){
+    public static int splitString(String s, int[] cuts){
+        Arrays.sort(cuts);
+        optcuts = new int[cuts.length];
+        int cost = s.length() * cuts.length;
+        for(int i = 0; i < cuts.length; i++){
+            cost = Math.min(splitStringRec(cuts, s.length(), i), cost);
+        }
+        return cost;
+    }
 
+    public static int splitStringRec(int[] cuts, int stringlen, int init){
+        if(cuts.length == 1){
+            return stringlen;
+        }
+        int cutlen = 0;
+        int cutr = 0;
+        if (init > 0){
+            //run as normal
+            cutlen = Integer.MAX_VALUE;
+            int[] cpy = Arrays.copyOfRange(cuts, 0, init);
+            int first = cuts[init];
+            for(int j = 0; j < cpy.length; j++){
[…]

### 2016-03-20 math_modeling/hw/chw/hw.tex
nothing like some fun C code
initial net logo

@@
 $T(t_{s})=490.85K$.
+\begin{quote}
+We can start by expressing our derivative in a discreet form. This gives the equation
+\[
+c_{2}\left({T(n+1)-T(n)\over\Delta t}\right)=-(T-T_{\infty}+c_{1}(T^{4}-T^{4}_{sur})).
+\]
+Solving this for $T(n+1)$ gives
+\[
+T(n+1)=-{\Delta t\over c_{2}}(T-T_{\infty}+c_{1}(T^{4}-T^{4}_{sur})) + T.
+\]
+\end{quote}
+We can model this simply in C:
+
+\begin{minted}[fontsize=\footnotesize,linenos]{c}
+#include <stdio.h>
+#include <stdlib.h>
+#include <math.h>
+#include <unistd.h>
+
+int main(int argc, const char* argv[]){
+    float temps [10000];
+    temps[0] = 298.15;
+    float starttime = 0;
+    float deltat;
+    if(argc > 1)
+        deltat = atof(argv[1]);
+    else
+        deltat = 0.001;
+    float c2 = 0.991667;
+    float c1 = 1.27575 * powf(10,-10);
+    float tinf = 473.15;
+    float tsur = powf(673.15,4);
+    FILE* fp = fopen("data.txt","w+");
+    int i = 0;
+    for(i = 0; i < 10000 && temps[i] < 490.85; i++){
+        if(i % 10 == 0){
+            fprintf(fp, "%f\n", temps[i]);
+        }
+        temps[i + 1] = (-1) * (deltat/c2) * (temps[i] - tinf + c1 * (pow(temps[i],4) - tsur)) + temps[i];
+        starttime += deltat;
+    }
+    fclose(fp);
+}
+\end{minted}
+\includegraphics[scale=0.3]{Rplots}
+
 \end{enumerate}

### 2016-03-20 math_modeling/final_project/twitch_agent.rb
initial net logo

@@ -1,0 +1,3 @@
+#!/usr/bin/env ruby
+
+require 'twitch'

### 2016-03-21 math_modeling/class/notes/popvsoda.tex
update c guide

\documentclass{hw}

\begin{document}
Suppose we have the following table representative of how each state refers to soda.

\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
Time Zone & Pop & Soda & Coke\\
\hline
EST (0.4) & 0.2 & 0.3 & 0.5\\
\hline
CST (0.2) & 0.5 & 0.3 & 0.2\\
\hline
MST (0.1) & 0.333 & 0.333 & 0.333\\
\hline
PST (0.3) & 0.1 & 0.2 & 0.7\\
\hline
\end{tabular}
\end{center}

By the law of total probability, we can determine that the probability of someone calling it ``pop"
is
\[
P(``pop")=P(pop|EST)\cdot P(EST) + P(``pop"|CST)\cdot P(CST) +
P(``pop"|MST)\cdot P(MST) + P(``pop"|PST)\cdot P(PST).
\]
More interestingly, we can find the probability that someone is from the central US by Bayes Theorem.
To do so, we need to come up with the value of $P(``pop")$. Recall that
$P(A|B)={P(A\cap B)\over P(B)}$.
\[
P(``pop") = 0.2(0.4) + 0.5(0.2) + 0.3333(0.1) + 0.1(0.3) = 0.243
\]
We can now calcuate the probability that one is from CST given that they say ``pop".
\[
P(CST|``pop")={P(CST\cap``pop")\over P(``pop")}={0.2\cdot 0.5\over 0.243} = 0.41.
\]
\end{document}

### 2016-03-21 math_modeling/final_project/const.rb

@@ -1,0 +1,12 @@
+class Const
+    @@CLIENT_ID = "y21r4bnd7hwbg2p22lw43xib0r6yng"
+    @@CLIENT_SECRET = "sw18eyd8xwdop46h3ahzas6rvr52qi8"
+
+    def self.cid
+        @@CLIENT_ID
+    end
+
+    def self.secret_key
+        @@CLIENT_SECRET
+    end
+end

### 2016-03-21 math_modeling/final_project/twitch_agent.rb

@@ -1,3 +1,26 @@
 #!/usr/bin/env ruby
 
 require 'twitch'
+require 'pp'
+require 'json'
+require 'net/http'
+
+require './const.rb'
+
+def main
+    twitch = Twitch.new({
+        client_id: Const.cid,
+        secret_key: Const.secret_key,
+        redirect_uri: "http://localhost",
+        scope: ["user_read"]
+    })
+
+    twitch.link
+
+    user = twitch.channel 'capcomfighters'
+    userhash = Net::HTTP.get(URI(user[:body].parsed_response["_links"]['chat']))
+    userself = JSON.parse(userhash)['_links']
+    pp userself
+end
+
+main

### 2016-03-22 number_theory/hw/hw7.tex
updated many things

\documentclass{hw}

\begin{document}
\makeheader{7}

\begin{enumerate}
\item If $p$ is prime and $x,y\in\ZZ_{p}$, prove that $(x-y)^{p}\equiv x^{p}-y^{p}\mod p$.
\begin{quote}
\begin{align*}
(x-y)^{p} &\equiv (x-y)\mod p \qquad\text{ by the corollary to FLT}\\
&\equiv (x\mod p) - (y\mod p)\\
&\equiv (x^{p}\mod p) - (y^{p}\mod p)\qquad\text{ by the corollary to FLT}\\
&\equiv x^{p} - y^{p} \mod p
\end{align*}
\end{quote}

\item Suppose that $p$ is prime and $k$ is a positive integer, and $x\in\ZZ_{p}$.
\begin{enumerate}
\item Prove that $x^{p^{k}}\equiv x\mod p$.

\item Prove that $x^{p^{k}-1}\equiv 1\mod p$ if and only if $\gcd{(x,p)}=1$.
\end{enumerate}

\item Let $p,n$, and $r$ be non-negative integers with $p$ prime. Further, assume that $r$ is the
remainder when $n$ is divided by $p-1$. Prove that $a^{n}\equiv a^{r}\mod p$.

\item Assume that $n$ is a positive integer. Prove that $1^{n}+2^{n}+3^{n}+4^{n}$ is divisible by
5 if and only if $n$ is not divisible by 4.

\item Determine whether there exists a solution to each of the following systems of congruences.  If
there is a solution, find all solutions to the system by writing the solution set as a single residue
class modulo $n$ for some $n\geq 2$. If there is no solution, prove that there is no solution.
\begin{enumerate}
\item $x\equiv5\mod7$\\
$x\equiv0\mod4$
\begin{align*}
x_{0}&=a_{1}c_{1}d_{1}+1_{2}c_{2}d_{2}\\
&= 5c_{1}d_{1} + 0c_{2}d_{2}\\
&= 5\cdot4\cdot d_{1}\\
&= 5\cdot4\cdot2\\
&= 40\\
&\equiv 12\mod28
\end{align*}

\item $x\equiv 5\mod7$\\
$x\equiv1\mod4$\\
$x\equiv0\mod5$
\begin{quote}
The $\gcd{(7,4,5)}=1$, so by the Chinese remainder theorem, we have a solution.
\begin{align*}
x_{0}&=a_{1}c_{1}d_{1}+a_{2}c_{2}d_{2}+a_{3}c_{3}d_{3}\\
&=5\cdot c_{1}\cdot d_{1} + 1\cdot c_{2}\cdot d_{2} + 0\\
&=5\cdot20\cdot d_{1}+1\cdot35\cdot d_{2}\\
&=5\cdot20\cdot-1 + 1\cdot35\cdot3\\
&=-100+105\\
&=5\\
&\equiv5\mod140
\end{align*}
\end{quote}

\item $x\equiv5\mod6$\\
$x\equiv2\mod4$
\begin{quote}
We cannot use Chinese remainder theorem here since 6 and 4 are not coprime. Suppose, however, that
there is a solution. We can form a new system by determining the prime factorization of 6.
\[
\begin{cases}
x\equiv5\mod2\\
x\equiv5\mod3\\
x\equiv2\mod4
\end{cases}
\to
\begin{cases}
x\equiv1\mod2\\
x\equiv2\mod3\\
x\equiv2\mod4
\end{cases}
\]
If this is the case, then $x\equiv1\mod2$ implies that the solution is odd, and $x\equiv2\mod4$ implies
the solution is even. This is not possible; therefore there is no solution.
\end{quote}

\item $3x\equiv1\mod10$\\
$5x\equiv2\mod7$
\begin{quote}
We cannot initially use Chinese remainder theorem to solve this problem. If we find $3^{-1}$ in
$\ZZ_{10}$, then we can multiply both sides of the congruence by that value to produce a new
congruence. We have $3x\equiv1\mod10$, which is satisfied by $x=7=3^{-1}$. We can rewrite this
congruence as $x\equiv7\mod10$. Similarly, we can find $5^{-1}$ in $\ZZ_{7}$. We have that
$5x\equiv1\mod7$, so $x=3=5^{-1}$. Multiplying both sides of the congruence yields the new system
\[
\begin{cases}
x\equiv7\mod10\\
x\equiv3\mod7
\end{cases}.
\]
We know this has a solution by the Chinese remainder theorem, since 10 and 7 are coprime.
\begin{align*}
x_{0}&=a_{1}c_{1}d_{1}+a_{2}c_{2}d_{2}\\
&=7\cdot7\cdot d_{1}+3\cdot10\cdot d_{2}\\
&=7\cdot7\cdot3+3\cdot10\cdot5\\
&=147+150\\
&=297\\
&\equiv13\mod70
\end{align*}
\end{quote}
\end{enumerate}

\item Find all solutions to the congruence $97x\equiv301\mod315$. It may be helpful to note that
$315=3^{2}\cdot4\cdot7$.
\begin{quote}
We can split this congruence into several parts.
\[
\begin{cases}
97x\equiv301\mod9\\
97x\equiv301\mod5\\
97x\equiv301\mod7
\end{cases}
\to
\begin{cases}
7x\equiv4\mod9\\
2x\equiv1\mod5\\
6x\equiv0\mod7
\end{cases}
\to
\begin{cases}
x\equiv7\mod9\\
x\equiv3\mod5\\
x\equiv0\mod7
\end{cases}
\]
By the Chinese remainder theorem, which we can use since 9, 7, and 5 are coprime, we have that
\begin{align*}
x_{0}&=7c_{1}d_{1}+3c_{2}d_{2}+0c_{3}d_{3}\\
&=7\cdot35\cdot d_{1}+3\cdot63\cdot d_{2} + 0\\
&=7\cdot35\cdot -1+3\cdot63\cdot 2 + 0\\
&=-245+378\\
&=133\\
&\equiv133\mod315.
\end{align*}
\end{quote}

\item Find all solutions to the congruence $x^{1000}\equiv1\mod10$.
\begin{quote}
By the prime factorization of 10, we have that
\[
5\mid(x^{1000}-1)\qquad\qquad\text{and}\qquad\qquad2\mid(x^{1000}-1).
\]
Since $2\mid(x^{1000}-1)$, we have that $x^{1000}\equiv1\mod2$, implying that $x$ is odd. This leaves
us with two possibilities in $\ZZ_{5}$, namely $\{1,3\}$. We can express 1000 as $5\cdot5\cdot5\cdot8$,
so we have
$((((x)^{8})^{5})^{5})^{5}\equiv(((x)^{8})^{5})^{5}\equiv((x)^{8})^{5}\equiv x^{8}\equiv1\mod5$.
We can express 8 as $4\cdot2$, so we have $(x^{2})^{4}\equiv1\mod5$, which means $x^{2}\equiv1\mod5$
by Fermat's Little Theorem. We know $x$ is either 1 or 3, so we can test the values. If we try 3, we
get that $9\equiv1\mod5$, which is not true, whereas if we try 1, we get $1\equiv1\mod5$, which is true.
Therefore, $x=1$.
\end{quote}
\end{enumerate}
\end{document}

### 2016-03-22 pde_two/elliptic.tex
updated many things

@@
 
-We will begin the discussion of Elliptic functions by analyzing the familiar trig functions, $\sin$ and
-$\cos$. Consider the following system of ODEs:
+\section{Introduction}
+
+The Jacobi Elliptic functions come from the result of integrating certain algebraic functions. The
+functions arose due to the lack on any elementary antiderivative to
+\[
+u=\int_{0}^{\phi}{d\theta\over\sqrt{1-k^{2}\sin^{2}{\theta}}}.
+\]
+The Jacobi Elliptic functions provide a way to express the solution to this integral via the properties
+\begin{align*}
+sn(u)&=sn\left(\int_{0}^{\phi}{d\theta\over\sqrt{1-k^{2}\sin^{2}{\theta}}}\right)=\sin{\phi}\\
+cn(u)&=cn\left(\int_{0}^{\phi}{d\theta\over\sqrt{1-k^{2}\sin^{2}{\theta}}}\right)=\cos{\phi}\\
+dn(u)&=dn\left(\int_{0}^{\phi}{d\theta\over\sqrt{1-k^{2}\sin^{2}{\theta}}}\right)=
+\sqrt{1-k^{2}\sin^{2}{\phi}}.
+\end{align*}
+
+The elliptic functions end up being applicable to more than just finding the antiderivative of $u$.
+We can express the solutions to several differential equations in terms of these functions, much like
+we can with $\sin{}$ and $\cos{}$. For example, the system
 \[
 \begin{cases}
-\frac{\dx}{\dt} = y\\
-\frac{\dy}{\dt} = -x
+\dot{x} = yz\\
+\dot{y} = -zx\\
+\dot{z} = -k^{2}xy
+\end{cases}
+\qquad
+\text{where}
+\qquad
+\begin{cases}
+x(0)=0\\
+y(0)=1\\
+z(0)=1
 \end{cases}
 \]
-This system has the solution $x(t)=\cos{t}$ and $y(t)=\sin{t}$. We will use this definition to formulate
-the Elliptic functions.\\
+can be solved by these elliptic functions, when subject to the initial conditions. More interestingly,
+we can find the solutions for several ODEs via the elliptic functions.
 
-We define a system of ODE's by letting $k\in(0,1)$ and $t\in\RR$ representative of time. We can define
-three functions, $sn(t,k),cn(t,k),dn(t,k)$, as solutions to the system
+\subsection{Applications to ODEs}
+Consider the second order ODE
 \[
 \begin{cases}
-\frac{\dx}{\dt} = yz\\
-\frac{\dy}{\dt} = -zx\\
-{\dz\over\dt}=-k^{2}xy
-\end{cases}
-\implies
-\begin{cases}
-{d\over\dt}sn(t,k)=cn(t,k)dn(t,k)\\
-{d\over\dt}cn(t,k)=-dn(t,k)sn(t,k)\\
-{d\over\dt}dn(t,k)=-k^{2}sn(t,k)cn(t,k)
-\end{cases}
+\ddot{x} = (1-x^{2})(1-k^{2}x^{2})\\
+\dot{x}(0)=1\\
+x(0)=0
+\end{cases}.
 \]
-The functions are not arbitrary; they have special properties. The first one we notice is that
+Recall that we defined the Jacobi Elliptic functions as solutions to a prior system of ODEs. We had
+the relationship that $\dot{x}=yz$, where $y$ and $z$ were functions of $t$. Taking another derivative
+yields $\ddot{x}=\dot{y}z+y\dot{z}$. We additionally have values for both $\dot{y}$ and $\dot{z}$.
+We can further substitute, which leads us to
 \[
-\lim_{k\to0^{+}}sn(t,k)=\sin{t}
-\qquad\lim_{k\to0^{+}}cn(t,k)=\cos{t}
-\qquad\lim_{k\to0^{+}}dn(t,k)=1.
-\]
-In addition, we have that
-\[
-\lim_{k\to1^{-}}sn(t,k)=\tanh{t}
-\qquad\lim_{k\to1^{-}}cn(t,k)=\sech{t}
-\qquad\lim_{k\to1^{-}}dn(t,k)=\sech{t}.
+\ddot{x}=\dot{y}z+y\dot{z}= -xz^{2}-k^{2}xy^{2}.
 \]

### 2016-03-22 resume/education.tex
updated many things

@@
 \begin{enumerate}[label={\Roman*}.]
-\item \textbf{Christopher Newport University $\cdot$ 2013 - Present}
+\item \textbf{Christopher Newport University $\cdot$ 2013 - Present}\\
+Bachelors of Science $-$ Mathematics $\cdot$ Bachelors of Science $-$ Computer Science
 \begin{itemize}
-\item Pursuing a B.S. in Mathematics and a B.S. in Computer Science
-\item Serving as Secretary for the Society of Physics club
-\item Employed by the Center for Academic Success as a tutor
-\item Earned Dean's List every semester
-\item Relavent coursework includes
+\item Community Involvement and Accomplishments
 \begin{itemize}
-\item UNIX Operating Systems, C++ Programming, Java Programming, Android Programming, Programming Language
-Concepts, Data Structures, Algorithms, Computer Hardware and Design, Discreet Mathematics, Linear Algebra,
-Calculus I - IV, Real Analysis, Partial and Ordinary Differential Equations, Writing in Mathematics using
-\hologo{LaTeX}
+\item Employed by the Center for Academic Success as a tutor in Mathematics and Computer Science
+\item Inducted into the Virginia Mu chapter of Pi Mu Epsilon
+\item Served as Secretary for the Society of Physics club
+\item Earned Dean's List consistently
 \end{itemize}
-\item Expected to graduate in May 2017
+% \newpage
+\item Relevant coursework includes
+\begin{itemize}
+\item Mathematical Education
+\begin{quote}
+Calculus I - IV, Linear Algebra, Real Analysis, Partial and Ordinary Differential Equations, Writing in
+Mathematics using \hologo{LaTeX}, Proofs and Discreet Mathematics, Number Theory, Independent Study in
+Partial Differential Equations
+\end{quote}
+\item Computer Science Education
+\begin{quote}
+UNIX Operating Systems, C++ Programming, Java Programming, Android Programming, Programming Language
+Concepts, Data Structures, Algorithms, Computer Hardware and Design, Discreet Mathematics
+\end{quote}
+\end{itemize}
 \end{itemize}

### 2016-03-22 resume/work.tex
updated many things

--- 2016-01-02 resume/work.tex
@@
 \begin{enumerate}[label={\Roman*}.]
-\item National Aeronautics and Space Administration $\Bigcdot$ November 2015 - Present
+
+\item MyNewLeaf, LLC $\Bigcdot$ January 2016 - Present
+\begin{itemize}
+\item Contracted to develop Android version of iOS application
+\item Collaborated with a large team of designers, managers, and programmers
+\item Served as lead programmer for Android version of application
+\end{itemize}
+
+\item National Aeronautics and Space Administration (NASA) $\Bigcdot$ November 2015 - Present
 \begin{itemize}
 \item Developed Android applications for Google Glass, Vuzix, and Recon Jet headsets
-\item Used git extensively for version control
-\item Wrote code primarily in Java
+\item Worked closely with a university professor to plan and create the application
+\item Participated in laser safety training
 \end{itemize}
-\item Freelance Application Developer $\Bigcdot$ July 2015 - Present
+
+\item Application Developer $\Bigcdot$ July 2015 - Present
 \begin{itemize}
-\item Built a location based audio application called Audionce, written for Android and iOS.
+\item Built a location based audio application called Audionce, written for Android and iOS
 \item Created a quiz web-application called WhensIt for the Public History Center at Christopher Newport
@@
 \begin{itemize}
-\item Tutored several subjects including Mathematics, Computer Engineering, and Computer Science
+\item Tutored several subjects including Mathematics (mostly Calculus), Computer Engineering, and
+Computer Science
 \item Participated in both one on one tutoring and group tutoring
+\item Created various resources for other Calculus tutors to use
 \end{itemize}
 \end{itemize}
+
+\newpage
 \item Kirk Family YMCA $\Bigcdot$ August 2009 - August 2015

### 2016-03-22 algorithms/projects/proj2/StringBreaking.java

@@ -1,95 +1,0 @@
-import java.io.File;
-import java.io.IOException;
-import java.util.Scanner;
-
-/**
- * This algorithm finds the optimal solution to a string breaking problem
- *
- * @author Talia Hicks
- * @author Steven Rosendahl
- * @version 03.15.2016
- *
- */
-public class StringBreaking {
-
-	public static int findBreakPoint(int n, int[] breakpoints) {
-		if (breakpoints.length == 1)
-			return 0;
-
-		else {
-			int median = 0;
-			int minDiff = n / 2 - breakpoints[0];
-			for (int i = 1; i < breakpoints.length; i++) {
-				if ((n / 2 - breakpoints[i]) < minDiff) {
-					minDiff = n / 2 - breakpoints[i];
-					median = i;
-				} // end if
-			} // end for
-			return median;
-		}
-	}
-
-	public static int stringBreakingAlg(String s, int[] breakpoints) {
-		int optsolution = 0;
-		int n = s.length(); // size of string
-		int breaks = breakpoints.length; // number of breakpoints
-
-		int x = 0;
-
-		int breakpoint = findBreakPoint(n, breakpoints);
-		System.out.println("Breakpoint: " + breakpoints[breakpoint]);
-		System.out.println("Cost of break: " + n);
-		optsolution += n;
-		System.out.println("Current total optimal cost: " + optsolution);
-		breaks--;
-
-		int breakpoint1 = breakpoint;
-		int breakpoint2 = breakpoint;
-
-		while (x <= breaks)
-		{
-
-			int[] brokenArray1 = new int[breakpoint1 + 1];
-			for (int i = 0; i < brokenArray1.length; i++) {
-				brokenArray1[i] = breakpoints[i];
-			}//end for
-
-			breakpoint1 = findBreakPoint(brokenArray1.length, brokenArray1);
-
-			System.out.println("Breakpoint: " + brokenArray1[breakpoint]);
-			System.out.println("Cost of break: " + brokenArray1.length);
-
-			optsolution += brokenArray1.length;
-			System.out.println("Current total optimal cost: " + optsolution);
-
-			breaks--;
-
-			int[] brokenArray2 = new int[n - (breakpoint2 + 1)];
-            // System.out.printf("SIZE OF ARRAY: %d\n", brokenArray2.length);
-			for (int i = 0; i < brokenArray2.length; i++) {
-                // System.out.printf("i: %d\n", i);
[…]

### 2016-03-22 algorithms/projects/proj2/StringSplit.java

@@ -1,42 +1,46 @@
-import java.util.Arrays;
+/**
+* Talia Hicks and Steven Rosendahl
+*
+*/
 
+import java.util.Arrays;
+import java.io.File;
+import java.io.IOException;
+import java.util.Collections;
+import java.util.Scanner;
 
 public class StringSplit {
 
-    static int[] optcuts;
+    static int stringlength;
+	static int[] optcuts;
+	static int[] breakpoints;
+    static int optCutsCount;
 
-    // public static int splitString(String s, ArrayList<Integer> splits){
-    //     if(splits.size() == 0) {
-    //         return s.length();
-    //     } else {
-    //         // int mid = splits.get(splits.size() / 2);
-    //         int mid = splits.get(splits.size() - 1);
-    //         // int mid = splits.get(0);
-    //         System.out.println("Splitting " + s + " at index " + mid);
-    //         ArrayList<Integer> less = new ArrayList<>();
-    //         for(Integer i : splits){
-    //             if(i < mid)
-    //                 less.add(Math.abs(i - mid));
-    //         }
-    //         ArrayList<Integer> max = new ArrayList<>();
-    //         for(Integer i : splits){
-    //             if(i > mid)
-    //                 max.add(i);
-    //         }
-    //         return s.length() + Math.min(
-    //             splitString(s.substring(0, mid), less),
-    //             splitString(s.substring(mid, s.length() - 1), max)
-    //         );
-    //         // return min;
-    //     }
-    // }
+    public static void readFile(File input) {
+		try {
+			Scanner scanner = new Scanner(input); // read in file
+			stringlength = scanner.nextInt(); // set string size
+			int breaks = scanner.nextInt(); // set number of breakpoints
+			breakpoints = new int[breaks]; // Initialize size of array
+			for (int i = 0; i < breakpoints.length; i++) {
+				breakpoints[i] = scanner.nextInt(); // set break points
+			} // end while
+			scanner.close();
+		}
+		catch (IOException e) {
+			e.printStackTrace();
+		}
+	}
 
-    public static int splitString(String s, int[] cuts){
+    public static int splitString(int n, int[] cuts){
         Arrays.sort(cuts);
         optcuts = new int[cuts.length];
-        int cost = s.length() * cuts.length;
+        optCutsCount = 0;
+        int cost = n * cuts.length;
+        // int prevcost = 0;
+        int nval = 0;
         for(int i = 0; i < cuts.length; i++){
[…]

### 2016-03-22 pde_two/elliptic.tex
working on elliptic

--- 2016-03-11 pde_two/elliptic.tex
@@
 
+\usepackage{pgf,tikz}
+\usepackage{mathrsfs}
+\usetikzlibrary{arrows}
+
 \begin{document}
@@
 
-We will begin the discussion of Elliptic functions by analyzing the familiar trig functions, $\sin$ and
-$\cos$. Consider the following system of ODEs:
+\section{Introduction}
+
+The Jacobi Elliptic functions come from non-elementary antiderivatives that arise when integrating
+over an ellipse. We can define these functions in several ways. One such way is to define the
+functions, $sn$, $cn$, and $dn$ as solutions to a system of ordinary differential equations.
+
+\section{Jacobi Functions as solutions to ODEs}
+
+We will begin by making a comparison of JEFs to Trig functions, with which we are familiar. Consider
+$\sin{}$ and $\cos{}$. There are many definitions of these two functions, but we will consider a
+definition that relates the two functions to ODEs, namely the system
 \[
 \begin{cases}
-\frac{\dx}{\dt} = y\\
-\frac{\dy}{\dt} = -x
-\end{cases}
+\dot{x}=y\\
+\dot{y}=-x
+\end{cases}.
 \]
-This system has the solution $x(t)=\cos{t}$ and $y(t)=\sin{t}$. We will use this definition to formulate
-the Elliptic functions.\\
+Under certain initial conditions, this leads to a basis composed of $\sin{}$ and $\cos{}$. It is with
+this analogy that we can consider the JEFs.
 
-We define a system of ODE's by letting $k\in(0,1)$ and $t\in\RR$ representative of time. We can define
-three functions, $sn(t,k),cn(t,k),dn(t,k)$, as solutions to the system
+Suppose we have a system of ODEs as follows
 \[
 \begin{cases}
-\frac{\dx}{\dt} = yz\\
-\frac{\dy}{\dt} = -zx\\
-{\dz\over\dt}=-k^{2}xy
-\end{cases}
-\implies
+\dot{x}=yz\\
+\dot{y}=-zx\\
+\dot{z}=-k^{2}xy
+\end{cases}.
+\]
+These may seem like arbitrary selections for the system, but, as we will see, they are not. We need
+to define $k$ before continuing; it is known as the modulus. We will also want to define the
+complimentary modulus, $\kappa=\sqrt{1-k^{2}}$. If we return to our ODEs, we can define the JEFs
+as solutions to the system under the following initial conditions:
+\[
+\begin{cases}
+sn(0,k)=x(0)=0\\
+cn(0,k)=y(0)=1\\
+dn(0,k)=z(0)=1\\
+\end{cases}.
+\]
+
+We can draw a relationship between $x$, $y$, and $z$ by forming a basis, much like we did with the
+ODE definitions of $\sin{}$ and $\cos{}$. Namely, we have that
+\[
+\{ x,y,z \}\to\{ sn(t,k),cn(t,k),dn(t,k) \}.
+\]
+This gives rise to the immediate result that
+\[
 \begin{cases}
@@
 {d\over\dt}dn(t,k)=-k^{2}sn(t,k)cn(t,k)
+\end{cases}.
+\]
+
+With these properties in mind, we will consider a slightly more geometric view of JEFs.
+
+\section{Jacobi Functions in a Geometric Sense}
+
+We will begin our discussion by considering an ellipse centered about the origin in the cartesian plane.
+Such an ellipse can be given by the equation
+\[
+\left({x^{2}\over a^{2}}\right) + \left({y^{2}\over b^{2}}\right) = 1.
+\]
+It will benefit us to consider the case where $b=1$, which gives the equation
+\[
+\left({x^{2}\over a^{2}}\right) + \left(y^{2}\right) = 1.
+\]
+We can take a given radius $r$ at the point $Q$ on the ellipse, and a given point $P$ at $a$. In
+addition, we can call the angle between the the $x$ axis and the radius $\theta$. Rather than consider
+a point $(x,y)$, we will actually consider a point $Q=(x(u),y(u))$, where $u$ is a function that we will
+define later. We can define the JEFs in terms of this ellipse; we have
+\[
+\begin{cases}
+cn(u,k)=y\\
+sn(u,k)={x\over a}\\
+dn(u,k)={r\over a}
 \end{cases}
 \]
-The functions are not arbitrary; they have special properties. The first one we notice is that
+\begin{center}
+\includegraphics[scale=0.5]{ellipsis}
+\end{center}
+We can establish some useful identities from this relationship, such as
 \[
-\lim_{k\to0^{+}}sn(t,k)=\sin{t}
-\qquad\lim_{k\to0^{+}}cn(t,k)=\cos{t}
-\qquad\lim_{k\to0^{+}}dn(t,k)=1.
+sn^{2}(u,k)+cn^{2}(u,k)=\left({x\over a}\right)^{2}+y^{2} = 1.
 \]
-In addition, we have that
+This is similar to our trig identity of $\sin^{2}{x}+\cos^{2}{x}=1$. We also have that
 \[
-\lim_{k\to1^{-}}sn(t,k)=\tanh{t}
-\qquad\lim_{k\to1^{-}}cn(t,k)=\sech{t}
-\qquad\lim_{k\to1^{-}}dn(t,k)=\sech{t}.
+k^{2}sn^{2}(u,k)+dn^{2}(u,k)=1,
+\]
+which may not be as familiar, but it still holds true. Interestingly enough, we can formulate the
+special (or maybe not so special) system of ODEs from the previous section using this new definition.
+
+\section{Jacobi Functions as an Integral}
+
+There is a third definition of JEFs that we can consider; that is, the form they take from an integral.
+Consider the ODE
+\[
+{\dx\over\dt}=\sqrt{(1-x^{2})(1-k^{2}x^{2})}.
+\]
+We can attempt separation of variables as an approach to solve this. Doing so yields
+\[
+\int_{0}^{sn(t,k)}{\dx\over\sqrt{(1-x^{2})(1-k^{2}x^{2})}}=t.
+\]
+The upper limit comes from the comparison to the integral definition of $\sin$:
+\[
+t=\int_{0}^{\sin{t}}{\dx\over\sqrt{1-x^{2}}}.
 \]

### 2016-03-22 pde_two/elliptic.tex
finished elliptic

@@
 \documentclass[notitlepage]{hw}
-\title{Elliptic Functions}
+\title{Jacobi Elliptic Functions in $\RR\times\RR$}
 \date{}
 \author{}
-
-\usepackage{pgf,tikz}
-\usepackage{mathrsfs}
-\usetikzlibrary{arrows}
 
@@
 define later. We can define the JEFs in terms of this ellipse; we have
+
+\vspace{0.5cm}
+
+\begin{minipage}{0.5\textwidth}
 \[
@@
 \]
+\end{minipage}
+\begin{minipage}{0.5\textwidth}
 \begin{center}
-\includegraphics[scale=0.5]{ellipsis}
+\includegraphics[scale=0.5]{ellipsis}\\
+
+A generalized ellipse
 \end{center}
+\end{minipage}
+
+\vspace{0.5cm}
+
 We can establish some useful identities from this relationship, such as
@@
 
+\section{Properties of Jacobi Functions}
+
+Before we continue, we will want to outline some fundamental properties of the JEFs. Recall that we
+already saw the analog to $\sin^{2}{x} + \cos^{2}{x} = 1$, which is
+\begin{gather}
+sn^{2}(u,k)+cn^{2}(u,k)= 1.
+\end{gather}
+In addition, we have
+\begin{gather}
+k^{2}sn^{2}(u,k)+dn^{2}(u,k)=1.
+\end{gather}
+We can also related the derivatives of the JEFs by
+\begin{align}
+{d\over\dt}sn(t,k)&=cn(t,k)dn(t,k)\\
+{d\over\dt}cn(t,k)&=-dn(t,k)sn(t,k)\\
+{d\over\dt}dn(t,k)&=-k^{2}sn(t,k)cn(t,k).
+\end{align}
+Since the functions are differentiable, we can talk about the limits of JEFs. If we let $k\to0$, we
+can see that our defining system of ODEs becomes
+\[
+\begin{cases}
+\dot{x}=yz\\
+\dot{y}=-zx\\
+\dot{z}=0
+\end{cases},
+\]
+which yields a basis of $\{\sin{t},\cos{t},1\}$, implying that
+\begin{align}
+\lim_{k\to0^{+}}sn(t,k)&=\sin{t}\\
+\lim_{k\to0^{+}}cn(t,k)&=\cos{t}\\
+\lim_{k\to0^{+}}dn(t,k)&=1.
+\end{align}
+We can also see what happens a $k\to1$:
+\begin{align}
+\lim_{k\to1^{-}}sn(t,k)&=\tanh{t}\\
+\lim_{k\to1^{-}}cn(t,k)&=\sech{t}\\
+\lim_{k\to1^{-}}dn(t,k)&=\sech{t}.
+\end{align}
+These are enough
+
 \section{Jacobi Functions as an Integral}
 
-There is a third definition of JEFs that we can consider; that is, the form they take from an integral.
-Consider the ODE
+There is a third definition of JEFs that we can consider; the form they take from an integral. As we
+saw previously, JEFs served as solutions to the system
 \[
-{\dx\over\dt}=\sqrt{(1-x^{2})(1-k^{2}x^{2})}.
+\begin{cases}
+\dot{x}=yz\\
+\dot{y}=-zx\\
+\dot{z}=-k^{2}xy
+\end{cases},
 \]
-We can attempt separation of variables as an approach to solve this. Doing so yields
+but this is not the only system for which JEFs provide a solution. Consider the following system
 \[
-\int_{0}^{sn(t,k)}{\dx\over\sqrt{(1-x^{2})(1-k^{2}x^{2})}}=t.
+\begin{cases}
+\dot{x}^{2}=(1-x^{2})(1-k^{2}x^{2})\\
+\dot{y}^{2}=(1-y^{2})(\kappa^{2}+k^{2}y^{2})\\
+\dot{z}^{2}=(1-z^{2})(z^{2}-\kappa^{2})
+\end{cases}.
 \]
-The upper limit comes from the comparison to the integral definition of $\sin$:
+We can transform the first ODE into
 \[
-t=\int_{0}^{\sin{t}}{\dx\over\sqrt{1-x^{2}}}.
+{\dx\over\dt}=\sqrt{(1-x^{2})(1-k^{2}x^{2})},
 \]
+which is a separable equation. We can rewrite this as
+\[
+\int{\dx\over\sqrt{(1-x^{2})(1-k^{2}x^{2})}}=t,
+\]
+Which we may notice as an Elliptic Integral of the first kind. If we make the substitution
+$x=\sin{u}$, then we have
+\[
+\int{\du\over\sqrt{1-k^{2}\sin^{2}{u}}}.
+\]
+This integral is not solvable by elementary means; rather we define the solution to be dependent on
+$sn(u,k)$. There are similar integral definitions corresponding to $cn(u,k)$ and $dn(u,k)$.
 
+\section{Applications to Elliptic Integrals}
+
+We saw previously that we cannot integrate
+\[
+\int{\dt\over\sqrt{1-k^{2}\sin^{2}{t}}}
+\]
+using elementary methods. We saw that this integral came from a substitution on
+\[
+\int{\dx\over\sqrt{(1-x^{2})(1-k^{2}x^{2})}}.
+\]
+Recall that by property $(3)$, we have
+\[
+{d\over\dt}sn(t,k)=cn(t,k)dn(t,k)=\sqrt{(1-sn^{2}(t,k))(1-k^{2}sn^{2}(t,k))}
+\]
+by property $(1)$. This leads us to the result that
+\[
+\int{d\over\dt}sn(t,k)=sn(t,k)=\int\sqrt{(1-sn^{2}(t,k))(1-k^{2}sn^{2}(t,k))}\dt.
+\]
+Note that we can invert the $sn(t,k)$, which leads us to
+\[
+sn^{-1}(t,k)=\int{\dt\over\sqrt{(1-t^{2})(1-k^{2}t^{2})}}.
+\]
+We can similarly formulate definitions for $cn^{-1}(t,k)$ and $dn^{-1}(t,k)$.
 \end{document}

### 2016-03-24 algorithms/projects/proj2/rep.tex
number theory

\documentclass{hw}

\usepackage{minted}

\begin{document}
\section*{String Splitting Algorithm}
\noindent Talia Hicks and Steven Rosendahl

\vspace{1cm}

We chose to implement the string breaking algorithm. To do this, we built a recursive algorithm
that looks down the left and right side of the string, using the $i^{\text{th}}$ breakpoint as a
starting index. The program loops through all possible starting indexes, and takes the least of them.

\begin{minted}{java}
for(int i = 0; i < cuts.length; i++){
    cost = Math.min(nval = splitStringRec(cuts, n, i), cost);
}
\end{minted}

This loop runs $n$ times, where $n$ is the size of the list of breakpoints. The function
\textit{splitStringRec} is a recursive function that runs down the left and right sides of the
string, trying out different breaks.

\begin{center}
\begin{minted}{java}
//Looping down the left
for(int j = 0; j < cpy.length; j++){
    cutlen = Math.min(splitStringRec(cpy, first, j), cutlen);
}

//Looping down the right
for(int i = 0; i < cpy.length; i++){
    cutr = Math.min(splitStringRec(cpy, newcost, i), cutr);
}
\end{minted}
\end{center}

The loop that runs through the left breakpoints runs once the first time through, then twice on the
second time through, and so on until it has run $n-1$ times. Similarly, the loop that runs through the
right breakpoints runs $n-1$ times, then $n-2$, $n-3$, and so on until it only runs once. Since each
loop takes the minimum of the cost, we have that only the minimum cuts will be preserved, while the
rest are thrown away. The runtime of such an algorithm is close to
\begin{align*}
\sum_{n=0}^{i}\left(
\sum_{k=0}^{n}(n-k) +
\sum_{k=0}^{n}(k+n)
\right)
&=\sum_{n=0}^{i}\left(O(n^{2}) + O(n^{2})\right)\\
&=\sum_{n=0}^{i}O(n^{2})\\
&=O(n^{3})
\end{align*}

To state the recurrence relation, let $n$ be the length of the string, $C$ be the array of cuts,
and $i=|C|$. We want to build a definition such that we loop down the left half of the string, making
cuts, which results in a new string of size $n-c$. Similarly, we want to move down the right side of
the string, making cuts which result in strings of size $c-n$. We also want to ``remove" the cuts that
we have already made, so we create a new array that contains all the remaining cuts without the cut we
just made (i.e. $C\backslash c_{0}$ where $c_{0}$ is the cut we just made).
\[
\text{min\_cost}(n, C)=
\begin{cases}
n,\hspace{6.2cm} i = 1\\
n + \text{min}
\begin{cases}
\text{min\_cost}(n - c_{0},\ C\backslash c_{0})\\
\text{min\_cost}(c_{0} - n,\ C\backslash c_{0})
\end{cases},
\hspace{0.8cm} i > 1
\end{cases}
\]

\end{document}

### 2016-03-24 math_modeling/class/notes/stochastics.tex
number theory

\documentclass{hw}

\begin{document}
\section*{Stochastic Modeling}

A stochastic matrix is a square matrix whose elements are probabilities and where each row sums to
one. One special example of a stochastic matrix is a matrix of transition probabilities.

\begin{quote}
\textit{\textbf{Example}}\\
Consider the case where a student enters a college as a one major, but changes to another.

\begin{center}
\begin{tabular}{c | c c c}
& STEM & Humanities & Business\\
\hline
STEM & 0.6 & 0.3 & 0.1\\
Hum. & 0.05 & 0.7 & 0.25\\
Bus. & 0.2 & 0.3 & 0.5
\end{tabular}
\end{center}

In this matrix, each $a_{ij}$ entry represents the probability that a student will change to that
major (or keep that major if it is the same). Suppose we have 1000 freshmen that enter the college
\[
x_{0}=[300,500,200]
\]
in the order STEM, Humanities, and Business. If we want to know the number of STEM majors in the
class's sophomore year, we can multiply our initial vector by the STEM column of our matrix:
\begin{align*}
\text{STEM Majors}&=(0.6)(300)+(0.05)(500)+(0.2)(200)\\
&=245\text{ students}\\
\text{Hum. Majors}&=(0.3)(300) + (0.7)(500) + (0.3)(200)\\
&= 500\text{ students}\\
\text{Bus. Majors}&=(0.1)(300) + (0.25)(500) + (0.5)(200)\\
&=255\text{ students}
\end{align*}
We can continue this process for each year that the students attend the college.
\end{quote}

This process for creating a sequence of vectors is called a Markov Chain. In general, we can
start with a vector $\bf{x}_{0}$ and a matrix $P$, and we get that
\begin{alignat*}{3}
\bf{x}_{1}&=\bf{x}_{0}\cdot P\\
\bf{x}_{2}&=\bf{x}_{1}\cdot P &&= \bf{x}_{0}\cdot P\cdot P\\
\bf{x}_{3}&=\bf{x}_{2}\cdot P &&= \bf{x}_{0}\cdot P\cdot P\cdot P,\\
\end{alignat*}
which yields the recurrence relation
\[
\bf{x}_{n}=\bf{x}_{0}P^{n}.
\]
$P^{n}$ is called the $n$-step transition matrix. In this matrix, the $a_{ij}$ element gives the
probability of going from state $i$ to state $j$.

\begin{quote}
\textit{\textbf{Example}}\\
Consider the movement of people from the city to the suburb.
\begin{center}
\begin{tabular}{c | c c}
& Cities & Suburbs\\
\hline
Cities & 0.96 & 0.04\\
Suburbs & 0.1 & 0.99
\end{tabular}
\end{center}
with the initial vector
\[
\bf{x}_{0}=[60,125].
\]
The current model assumes that population stabilizes, but this is not always the case. 
\end{quote}

\end{document}

### 2016-03-24 number_theory/hw/hw7.tex
number theory

@@
 \item Prove that $x^{p^{k}}\equiv x\mod p$.
-
+\begin{quote}
+We can rewrite $x^{p^{k}}$ as $(((x^{p})^{p})^{p}...)^{p}$, where there are $k\ p'\text{s}$. We
+can mod out by $p$, and we get $(((x^{p})^{p})^{p}...)^{p}$, where there are $k-1\ p'\text{s}$. If
+we mod out $k-2$ more times, we get $x^{p}\equiv x\mod p$, which by the corollary to Fermat's
+Little Theorem means $x^{p^{k}}\equiv x\mod p$.
+\end{quote}
 \item Prove that $x^{p^{k}-1}\equiv 1\mod p$ if and only if $\gcd{(x,p)}=1$.
+\begin{quote}
+$\Rightarrow)$ Assume that $gcd(x,p)\neq1$. Then $x\equiv0\mod p$, since the only thing not coprime
+with $p$ is a multiple of $p$. We can raise both sides of the congruence to the ${p^{k}-1}$
+power to produce
+\[
+x^{p^{k}-1}\equiv 0^{p^{k}-1} \equiv 0 \ \cancel{\equiv}\ 1 \mod p.
+\]
+\noindent\\
+$\Leftarrow)$ Suppose $gcd(x,p)=1$. Then $x$ has an inverse in $\ZZ_{p}$. We know by part $(a)$
+that $x^{p^{k}}\equiv x\mod p$, so $x^{-1}\cdot x^{p^{k}}\equiv x\cdot x^{-1}\equiv 1\mod p$.
+\end{quote}
 \end{enumerate}
@@
 remainder when $n$ is divided by $p-1$. Prove that $a^{n}\equiv a^{r}\mod p$.
+\begin{quote}
+Since $r$ is the remainder, we have that $n = q(p-1)+r$ by the division algorithm. Since
+$p,n$, and $r$ are non-negative, we can say $a^{n}=a^{q(p-1) + r}$, which is the same as
+$a^{n}=a^{q(p-1)}a^{r}$. By Fermat's Little Theorem, we have that $a^{p-1}=1\mod p$, so
+$a^{n}\equiv 1^{q}a^{r}\mod p$, or $a^{n}\equiv a^{r}\mod p$.
+\end{quote}
 
@@
 5 if and only if $n$ is not divisible by 4.
+\begin{quote}
+$\Rightarrow)$ Suppose $4\mid n$. Then $n=4k,\ k\in\ZZ$, so we have
+\begin{align*}
+1^{n}+2^{n}+3^{n}+4^{n}&\equiv 1^{4k}+2^{4k}+3^{4k}+4^{4k}\\
+&\equiv 1^{4^k}+2^{4^k}+3^{4^k}+4^{4^k}\\
+&\equiv 1^k+1^k+1^k+1^k\mod5\\
+&\ \cancel{\equiv}\ 0\mod5.
+\end{align*}
+\noindent\\
+$\Leftarrow)$ Suppose $4\nmid n$. Then $n\ \cancel{\equiv}\ 0\mod4$. Suppose that $n\equiv1\mod4$.
+We now have $n=4k+1$ for some $k\in\ZZ$. By $(3)$, we know that $a^{n}\equiv a^{r}\mod p$.
+Similarly, we can say $a^{n} + b^{n} + c^{n} + d^{n}\equiv a^{r}+b^{r} + c^{r} + d^{r}\mod p$.
+Then, we have $1^{n}+2^{n}+3^{n}+4^{n}\equiv 1 + 2 + 3 + 4\mod5\equiv0\mod5$. Suppose $n\equiv2\mod4$.
+Then we have $1^{n}+2^{n}+3^{n}+4^{n}\equiv 1 + 4 + 9 + 16\equiv1-1+1-1\mod5\equiv0\mod5$. Finally,
+suppose $n\equiv3\mod4$. Then $1^{n}+2^{n}+3^{n}+4^{n}\equiv1+8+27+64\equiv1+2+3-1\equiv0\mod5$.
+\end{quote}
 
@@
 $x\equiv0\mod4$
+\begin{quote}
+Since $gcd(7,4)=1$, we can use Chinese remainder theorem, which tells us
 \begin{align*}
@@
 \end{align*}
+\end{quote}
 

### 2016-03-24 pde_two/elliptic.tex
number theory

@@
 We will begin by making a comparison of JEFs to Trig functions, with which we are familiar. Consider
-$\sin{}$ and $\cos{}$. There are many definitions of these two functions, but we will consider a
+$\sin{t}$ and $\cos{t}$. There are many definitions of these two functions, but we will consider a
 definition that relates the two functions to ODEs, namely the system
@@
 \]
-Under certain initial conditions, this leads to a basis composed of $\sin{}$ and $\cos{}$. It is with
+Under certain initial conditions, this leads to a basis composed of $\sin{t}$ and $\cos{t}$. It is with
 this analogy that we can consider the JEFs.
@@
 These may seem like arbitrary selections for the system, but, as we will see, they are not. We need
-to define $k$ before continuing; it is known as the modulus. We will also want to define the
-complimentary modulus, $\kappa=\sqrt{1-k^{2}}$. If we return to our ODEs, we can define the JEFs
-as solutions to the system under the following initial conditions:
+to define $k$ before continuing; it is known as the modulus. On an ellipse, the value of $k$ is given
+by
+\[
+k=\sqrt{1-{a^{2}\over b^{2}}},
+\]
+where $a$ is the major axis and $b$ is the semi-major axis.
+We can also define $\kappa=\sqrt{1-k^{2}}$, known as the complimentary modulus. If we return to
+our system, we can define the JEFs as solutions to the system under the following initial conditions:
 \[
@@
 We can draw a relationship between $x$, $y$, and $z$ by forming a basis, much like we did with the
-ODE definitions of $\sin{}$ and $\cos{}$. Namely, we have that
+ODE definitions for $\sin{t}$ and $\cos{t}$. Namely, we have that
 \[
@@
 \[
-\begin{cases}
+\begin{dcases*}
 {d\over\dt}sn(t,k)=cn(t,k)dn(t,k)\\
@@
 {d\over\dt}dn(t,k)=-k^{2}sn(t,k)cn(t,k)
-\end{cases}.
+\end{dcases*}.
 \]
@@
 \[
-\begin{cases}
-cn(u,k)=y\\
+\begin{dcases*}
+cn(u,k)=y\vphantom{\frac{0}{0}}\\
 sn(u,k)={x\over a}\\
 dn(u,k)={r\over a}
-\end{cases}
+\end{dcases*}
 \]

### 2016-03-24 algorithms/projects/proj2/StringSplit.java

@@ -32,12 +32,12 @@ public class StringSplit {
 		}
 	}
 
+    // Call this method to start the process
     public static int splitString(int n, int[] cuts){
         Arrays.sort(cuts);
         optcuts = new int[cuts.length];
         optCutsCount = 0;
         int cost = n * cuts.length;
-        // int prevcost = 0;
         int nval = 0;
         for(int i = 0; i < cuts.length; i++){
             cost = Math.min(nval = splitStringRec(cuts, n, i), cost);
@@ -45,6 +45,7 @@ public class StringSplit {
         return cost;
     }
 
+    // don't call this method directly.
     public static int splitStringRec(int[] cuts, int stringlen, int init){
         if(cuts.length == 1){
             return stringlen;
@@ -52,15 +53,12 @@ public class StringSplit {
         int cutlen = 0;
         int cutr = 0;
         if (init > 0){
-            //run as normal
             cutlen = Integer.MAX_VALUE;
-            int[] cpy = Arrays.copyOfRange(cuts, 0, init);
+            int[] cpy = copyArray(cuts, 0, init);
             int first = cuts[init];
             int pcl;
             for(int j = 0; j < cpy.length; j++){
-                // System.out.println("Trying to make a cut " + cutlen);
                 cutlen = Math.min(pcl = splitStringRec(cpy, first, j), cutlen);
-                // System.out.println("Making a cut of cost " + cutlen);
                 if(cutlen <= stringlen){
                     System.out.println("Making a cut on the left side of cost " + cutlen);
                 }
@@ -68,7 +66,7 @@ public class StringSplit {
         }
         if (init < cuts.length - 1){
             cutr = Integer.MAX_VALUE;
-            int[] cpy = Arrays.copyOfRange(cuts, init + 1, cuts.length);
+            int[] cpy = copyArray(cuts, init + 1, cuts.length);
             int newcost = stringlen - cuts[init];
             for(int i = 0; i < cpy.length; i++){
                 cpy[i] = cpy[i] - cuts[init];
@@ -81,6 +79,16 @@ public class StringSplit {
             }
         }
         return stringlen + cutlen + cutr;
+    }
+
+    // Sometimes we need to copy an array
+    public static int[] copyArray(int[] src, int startIndex, int endIndex){
+        int[] ret = new int[endIndex - startIndex];
+        for(int i = 0; i < endIndex; i++){
+            if(i >= startIndex)
+                ret[i - startIndex] = src[i];
+        }
+        return ret;
     }
 
     public static void main(String[] args){

### 2016-03-24 math_modeling/hw/chw/math.c

@@ -26,8 +26,7 @@ int main(int argc, const char* argv[]){
         if(i % 10 == 0){
             fprintf(fp, "%f\n", temps[i]);
         }
-        temps[i + 1] = (-1) * (deltat/c2) * (temps[i] - tinf +
-            c1 * (pow(temps[i],4) - tsur)) + temps[i];
+        temps[i + 1] = (-1) * (deltat/c2) * (temps[i] - tinf + c1 * (pow(temps[i],4) - tsur)) + temps[i];
         starttime += deltat;
     }
 

### 2016-03-26 math_modeling/class/notes/stochastics.tex
polynomials are fun

@@
 \]
-The current model assumes that population stabilizes, but this is not always the case. 
+We can introduce some ``noise" into the system by discussing the eigenvalues of this system. In the
+usual case, $\lambda=1$. We can show this by solving the system.
+\begin{align*}
+(\bf{P}-\lambda)\bf{x}&\to
+\left[
+\begin{array}{c c c}
+-0.4 & 0.4 & 0\\
+0.1 & 0.1 & 0
+\end{array}
+\right]\\
+&\to
+\left[
+\begin{array}{c c c}
+1 & -1 & 0\\
+0 & 0  & 0
+\end{array}
+\right],
+\end{align*}
+so $x=y=1$.
 \end{quote}
 
+Often in Markov processes, we come across matrices raised to a power. To help us calculate these
+products, we can come up with a new matrix, $\bf{S}$. We can then calculate
+$\bf{S}^{-1}\bf{D}\bf{S}$, where $D$ is diagonalized matrix of the eigenvalues of $\bf{P}$. We know
+that we can diagonalize $\bf{P}$ by the following theorem:
+\begin{quote}
+An $n\times n$ matrix $\bf{P}$ is diagonalizable if and only if it has $n$ linearly independent
+eigenvectors.  
+\end{quote}
 \end{document}

### 2016-03-26 number_theory/hw/hw7.tex
polynomials are fun

@@
 by Fermat's Little Theorem. We know $x$ is either 1 or 3, so we can test the values. If we try 3, we
-get that $9\equiv1\mod5$, which is not true, whereas if we try 1, we get $1\equiv1\mod5$, which is true.
-Therefore, $x=1$.
+get that $9^{4}\equiv1\mod5$, so $(-1)^{4}\equiv1\mod5$ and $1\equiv1\mod5$. If we try 1, we
+get $1\equiv1\mod5$, which is true. Therefore, $x=1$ and $x=3$ are solutions. We can represent all
+solutions as $1+10n,\ n\in\ZZ$ and $3+10k,\ k\in\ZZ$.
+% If we consider the two
+% other odd options in $\ZZ_{10}$, we have that
+% $(7^{2})^{4}\equiv((-3)^{2})^{4}\equiv9^{4}\equiv(-1)^{4}\equiv1\mod10$. Similarly, if we allow
 \end{quote}

### 2016-03-26 number_theory/hw/hw8.tex
polynomials are fun

\documentclass{hw}

\begin{document}
\makeheader{8}

\begin{enumerate}
\item Determine whether each of the following polynomials has a zero in the given $\ZZ_{p}$. Either
find a zero or prove that no zero exists.
\begin{enumerate}
\item $f(x)=x^3+x^2+x+1$ in $\ZZ_{2}$.
\begin{quote}
We can factor $x^3+x^2+x+1$ into $(x^{2}+1)(x+1)$. The only two options for a root in $\ZZ_{2}$ are
1 and 0. If we plug in zero we get $(0+1)(0+1)\equiv1\mod2$, so 1 is not a zero of the polynomial.
If we try 1, we get $(1+1)(1+1)\equiv(2)(2)\equiv(0)(0)\equiv0\mod2$, so 1 is a zero of the
polynomial.
\end{quote}
\item $f(x)=x^3+x^2+x+1$ in $\ZZ_{3}$.
\begin{quote}
$f(x)$ factors into $(x^{2}+1)(x+1)\equiv(x^{2}-2)(x-2)\mod3$. We have one zero at $x=2$. If we test
0, we get $(0+1)(0+1)\equiv1\mod3$, so 0 is not a zero of the polynomial, and if we test 1 we get
$(-1)(-1)\equiv1\mod3$, so 1 is not a zero either.
\end{quote}
\item $f(x)=x^{2}+2x+3$ in $\ZZ_{5}$.
\begin{quote}
The polynomial $f(x)=x^{2}+2x+3$ cannot be factored in $\ZZ_{5}$. Therefore, it has no zeros since
we cannot express it in the form $g(x)(x-\alpha)$.
\end{quote}
\end{enumerate}

\item Let $f(x)=x^{2}+1$. Prove that $f$ has a zero in $\ZZ_{5}$, but not in $\ZZ_{7}$.
\begin{quote}
If we consider the polynomial in $\ZZ_{5}$, we have that
\[x^{2}+1\equiv x^{2}-4\equiv(x+2)(x-2)\equiv(x-3)(x-2)\mod5.\]
Since $f$ was reducible in $\ZZ_{5}$, we know it has at least one zero, and in this case it has
two zeros, $x=3$ and $x=5$. In $\ZZ_{7}$, we have
\[x^{2}+1\equiv x^{2}-6\mod7.\]
The polynomial $x^{2}-6$ is irreducible in $\ZZ_{7}$, so there are no roots.
\end{quote}
\end{enumerate}
\end{document}

### 2016-03-31 math_modeling/class/notes/stochastics.tex
working on number theory

@@
 An $n\times n$ matrix $\bf{P}$ is diagonalizable if and only if it has $n$ linearly independent
-eigenvectors.  
+eigenvectors.\\\\
+
+\textit{\textbf{Example}}\\
+Consider three types of detergent that yield the Markov matrix
+\begin{center}
+\begin{tabular}{c | c c c}
+& Tide & Bounty & Bounce\\
+\hline
+Tide & .8 & .1 & .1\\
+Bounty & .2 & .6 & .2\\
+Bounce & .3 & .3 & .4
+\end{tabular}
+\end{center}
+where the entries indicate the probability that a user will stay with that product. Rather than
+finding the next state, we will consider a vector $\pmb{\pi}$, called the limiting state probability
+vector. This vector will be representative of the probabilities of users staying with a brand.
+Similar to the previous examples, we have
+\[\pmb{\pi}\bf{P}=\pmb{\pi}.\]
+From this equation, we get
+\[
+\begin{cases}
+.8\pi_{1}+.2\pi_{2}+.3\pi_{3}=\pi_{1}\\
+.1\pi_{1}+.6\pi_{2}+.3\pi_{3}=\pi_{2}\\
+.1\pi_{1}+.2\pi_{2}+.4\pi_{3}=\pi_{3}
+\end{cases}
+\]
+The third equation in this case is a linear combination of the first two, so we cannot user it here.
+However, we can replace it with a linearly independent row:
+\[
+\begin{cases}
+.8\pi_{1}+.2\pi_{2}+.3\pi_{3}=\pi_{1}\\
+.1\pi_{1}+.6\pi_{2}+.3\pi_{3}=\pi_{2}\\
+\pi_{1}+\pi_{2}+\pi_{3}=1
+\end{cases}
+\to
+\left[
+\begin{array}{c c c | c}
+-.2 & .2 & .3 & 0\\
+.1 & -.4 & .3 & 0\\
+1 & 1 & 1 & 0
+\end{array}
+\right]
+\to
+\pmb{\pi}=
+\left[
+\begin{array}{c}
+{6/11}\\
+{3/11}\\
+{2/11}
+\end{array}
+\right].
+\]
 \end{quote}
+
 \end{document}

### 2016-03-31 number_theory/hw/hw8.tex
working on number theory

@@
 \end{quote}
+
+\item Suppose that $p$ is prime, $k$ is a non-zero element of $\ZZ_{p}$, and $f_{k}(x)=x^{2}-k$.
+\begin{enumerate}
+\item Prove that $s$ is a zero of $f_{k}$ if and only if $-s$ is also a zero of $f_{k}$.
+\begin{quote}
+$\Rightarrow)$ Suppose $s$ is a zero of $f_{k}$. Then $s^{2}-k\equiv0\mod p$. If $-s$ is not a zero,
+then we have $(-s)^{2}-k\ \cancel{\equiv}\ 0\mod p$, so $s^{2}-k\ \cancel{\equiv}\ 0\mod p$, which
+is a contradiction.\\\\
+$\Leftarrow)$ Suppose $-s$ is a zero of $f_{k}$. Then $(-s)^{2}-k\equiv0\mod p$. Then
+$s^{2}-k\equiv0\mod p$, so $s$ is a zero of $f_{k}$.
+\end{quote}
+\item Further assuming that $p>2$ prove that $f_{k}$ either has no zeros in $\ZZ_{p}$ or exactly
+two zeros in $\ZZ_{p}$.
+\begin{quote}
+Suppose there was only one zero in $\ZZ_{p}$. We know this cannot be the case, since by $(a)$ we
+showed that if $s$ is a zero, then $-s$ is also a zero. If we suppose there are more than 2 zeros,
+then we can consider another zero, $\sigma$. By $(a)$, we know that $-\sigma$ is also a zero. Since
+$deg(f)=2$, we know that there are at most 2 zeros in $\ZZ_{p}$. There cannot be just one zero, so
+there must be only two zeros, or no zeros.
+% We
+% know that, without loss of generality, that if we consider the first root $\sigma$, then we can
+% say
+% \begin{align*}
+% f_{k}(x)&=(x-\sigma)g(x)\\
+% x^{2}-k&=(x-\sigma)g(x)\\
+% {x^{2}-k\over x-\sigma}&=g(x).
+% \end{align*}
+% If we perform the division, we find that
+% \[g(x)=x+\sigma+{\sigma^{2}-k\over x-\sigma}=x+\sigma,\]
+% meaning that
+\end{quote}
+\end{enumerate}
+
+\item Let $p$ be a prime with $p>2$ and assume that $a,b,c,r\in\ZZ_{p}$ with $a\nequiv0\mod p$.
+Further, we define the polynomial $f(x)=ax^{2}+bx+c\in\ZZ_{p}[x]$.
+\begin{enumerate}
+\item Prove that $r$ is a zero of $f$ in $\ZZ_{p}$ if and only if
+$(2ar+b)^{2}\equiv b^{2}-4ac\mod p$.
+\begin{quote}
+$\Rightarrow)$ Since $r$ is a zero of $f(x)$ and $f$ is a quadratic polynomial, we can create the
+expression
+\begin{align*}
+ar^{2}+br+c&\equiv0\mod p\\
+ar^{2}+br&\equiv-c\mod p.
+\end{align*}
+Consider the term $(2ar+b)^{2}$. Then
+\begin{align*}
+(2ar+b)^{2}&= 4a^{2}r^{2}+4bar+b^{2}\\
+&=4a(ar^{2}+br)+b^{2}\\
+&=4a(-c)+b^{2}\\
+&=b^{2}-4ac\\
+&\equiv b^{2}-4ac\mod p
+\end{align*}
+\noindent\\
+$\Leftarrow)$ Suppose $(2ar+b)^{2}\equiv b^{2}-4ac\mod p$. Then
+\begin{align*}
+&4a^{2}r^{2}+4arb+b^{2}\equiv b^{2}-4ac\mod p\\
+&4a^{2}r^{2}+4arb\equiv-4ac\mod p\\
+&4a^{2}r^{2}+4arb+4ac\equiv0\mod p\\
+&4a(ar^{2}+br+c)\equiv0\mod p.
+\end{align*}
+Since we are in $\ZZ_{p}$, we know all elements have an inverse. We can multiply both sides of the
+congruence by $a^{-1}4^{-1}$, which gives us $ar^{2}+br+c\equiv0\mod p$. Therefore, $r$ is a root
+of $f$.
+\end{quote}
+\item A point $y\in\ZZ_{p}$ is called a \textit{perfect square} if there exists $z\in\ZZ_{p}$
+such that $z^{2}=y$.
+\begin{enumerate}
+\item If $f$ has at least one zero in $\ZZ_{p}$, prove that $b^{2}-4ac$ is a perfect square.
+\begin{quote}
+Suppose $r$ is a zero of $f$. Then by $(a)$ we have
+\[
+(2ar+b)^{2}\equiv b^{2}-4ac\mod p,
+\]
+which implies $b^{2}-4ac$ is a perfect square.
+\end{quote}
+\item If $b^{2}-4ac\equiv0\mod p$, prove that $f$ has a uniqe zero in $f$. Find a formula for that
+zero in terms of $a$ and $b$.
+\begin{quote}
+Suppose $b^{2}-4ac\equiv0\mod p$. We know by $(a)$ that $r$ is root, and therefore
+$(2ar+b)^2\equiv b^2-4ac \mod p$. Suppose $\rho$ is a root of $f$. then
+$(2a\rho+b)^2\equiv b^2-4ac \mod p$, so $(2ar+b)^2\equiv(2a\rho+b)^2$. Since $b^2-4ac$ is a perfect
+square, then $2ar+b\equiv2a\rho+b$. Then $2ar\equiv2a\rho$, and multiplying by $2^{-1}a^{-1}$ yields
+$r\equiv\rho$. To find an expression for $r$, we know that $(2ar+b)^2=0$, so if we multiply by
+$(2ar+b)^{-1}$, we have $2ar+b=0$. Then $r=-b2^{-1}a^{-2}$.
+\end{quote}
+\item If $b^{2}-4ac\nequiv0\mod p$ and $b^{2}-4ac$ is a perfect square, prove that $f$ has exactly
+two distinct zeros in $\ZZ_{p}$.
+\begin{quote}
+We know that we have the equivalence $(2ar+b)^2\equiv b^2-4ac$. Let $x=(2ar+b)^2$ and
+$k=b^2-4ac$. Then we can rearrange the congruence to say $x^2-k\equiv0\mod p$. We know
+$x^2-k$ has either 0 or 2 zeros. Since $k$ is a perfect square, we can let $k=j^2$ for some
+$j\in\ZZ_{p}$. Then $x^2-k=x^2-j^2=(x+j)(x-j)$, which has two distinct roots in $\ZZ_{p}$.
+\end{quote}
+\end{enumerate}
+\end{enumerate}
+
+\item Suppose that $g(x)=x^2+1\in\ZZ_{3}[x]$ and let $\phi$ be a zero of $g$.
+\begin{enumerate}
+\item Prove that $\phi\notin\ZZ_{3}$.
+\begin{quote}
+
+\end{quote}
+\item Define the set $\FF_{9}=\{ 0,1,2,\phi,\phi+1,\phi+2,2\phi,2\phi+1,2\phi+2 \}$. Assuming that
+the multiplication and addition in $\FF_{9}$ obeys the distributive law, prove that every element of
+$\FF_{9}$ has a multiplicative inverse.
+\begin{quote}
+
+\end{quote}
+\end{enumerate}
+
+\item If $p>2$ is prime, prove that $\ZZ_{p}$ contains exactly $(p+1)/2$ perfect squares.
+\begin{quote}
+
+\end{quote}
 \end{enumerate}

### 2016-03-31 math_modeling/final_project/twitch_agent.rb

@@ -4,12 +4,9 @@ require 'rubygems'
 require 'twitch'
 require 'pp'
 require 'json'
-# require 'twitch-chat'
-# require 'nokogiri'
 require 'open-uri'
 require 'open_uri_redirections'
 require 'thread_safe'
-# require 'cinch'
 
 require './const.rb'
 

### 2016-04-01 number_theory/hw/hw8.tex
working on presentation

@@
 $\Leftarrow)$ Suppose $-s$ is a zero of $f_{k}$. Then $(-s)^{2}-k\equiv0\mod p$. Then
-$s^{2}-k\equiv0\mod p$, so $s$ is a zero of $f_{k}$.
+$s^{2}-k\equiv0\mod p$, and $s$ is a zero of $f_{k}$, so $-s$ is a zero of $f$.
 \end{quote}
@@
 $r\equiv\rho$. To find an expression for $r$, we know that $(2ar+b)^2=0$, so if we multiply by
-$(2ar+b)^{-1}$, we have $2ar+b=0$. Then $r=-b2^{-1}a^{-2}$.
+$(2ar+b)^{-1}$, we have $2ar+b=0$. Then $r=-b2^{-1}a^{-1}$.
 \end{quote}
@@
 \begin{quote}
-
+Suppose that there is a zero of the polynomial $r\in\ZZ_{3}$. If $r=0$, then we have
+$g=1\nequiv0\mod3$, so $0$ is not a zero of the polynomial. If $r=1$, then we have
+$g=2\nequiv0\mod3$. If $r=2$, we have $2\equiv-1\mod3$, so $g=(-1)^2+1\equiv2\nequiv0\mod3$, so there
+are no zeros in $\ZZ_{3}$.
 \end{quote}
 \item Define the set $\FF_{9}=\{ 0,1,2,\phi,\phi+1,\phi+2,2\phi,2\phi+1,2\phi+2 \}$. Assuming that
-the multiplication and addition in $\FF_{9}$ obeys the distributive law, prove that every element of
-$\FF_{9}$ has a multiplicative inverse.
+the multiplication and addition in $\FF_{9}$ obeys the distributive law, prove that every non-zero
+element of $\FF_{9}$ has a multiplicative inverse.
 \begin{quote}
-
+We know that $\phi$ is a zero of $g$, so we have that $\phi^2+1=0$, and $\phi^{2}=2$. Subtracting
+1 from both sides yields $\phi^{2}-1=1$, so $(\phi+1)(\phi-1)=1$. Therefore the inverse of
+$(\phi+1)$ is $(\phi-1)$. If we consider $\phi(2\phi)$, we get $2\phi^2=4=1$, so
+$\phi$ and $2\phi$ are inverses. If we consider $(2\phi+1)(2\phi+2)$, we get
+$4\phi^2+4\phi+2\phi+2=(1)(2)+\phi-\phi-1=2-1=1$, so $(2\phi+1)$ and $(2\phi+2)$ are inverses.
+We also know that $(1)(1)=1$, so 1 is its own inverse. Finally, $(2)(2)=4=1$, so 2 is its own inverse
+as well.
 \end{quote}
@@
 \begin{quote}
-
+Consider the set $S=\{ \{x,-x\}|x\in\ZZ_{p}\text{ and }x\nequiv0\mod p \}$. We first want to show
+that $x\nequiv-x\mod p$.
+\begin{quote}
+Suppose $x\equiv-x\mod p$. Then $x\equiv(p-1)x\equiv px-x\mod p$, so $2x\equiv xp\mod p$. We can
+multiply by $x^{-1}$, which yields $2\equiv p\mod p$, so $p\equiv2\mod p$, which is a contradiction
+since $p>2$. Therefore $x\nequiv-x\mod p$.
+\end{quote}
+We now need to show that every set $A\in S$ is disjoint with another set $B\in S$.
+\begin{quote}
+Suppose $A=\{x,-x\}$ and $B=\{y,-y\}$ with $A\cap B\neq\emptyset$. Take $z\in A\cap B$.
+Then $z\in A$ and $z\in B$. Suppose without loss of generality that $z\equiv x\mod p$ and
+$z\equiv y\mod p$. Then $x\equiv y\mod p$, so they are the same element.
+\end{quote}
+We can now take every subset of $S$ and square the elements within that set. Doing so gives for
+any subset $\{ x^{2},x^{2} \}$ which contains two equivalent perfect squares. We know that for
+every $p$ there are $(p-1)/2$ of these sets, and that there is a trivial set $\{0,0\}$, so there are
+$((p-1)/2)+1$, or $(p+1)/2$ sets of perfect squares.
 \end{quote}

### 2016-04-03 math_modeling/hw/hw6/hw6.tex
ethics final paper

\documentclass{hw}

\usepackage{tikz}
\usetikzlibrary{automata,positioning}

\begin{document}

\begin{enumerate}
\item Two competing companies, Pollution Products and Environmental Hazards, simultaneously introduce
new enzyme laundry detergents. Market tests indicate that during a year, Pollution keeps 60\% of
its customers and loses 40\% of its customers to Environmental. On the other hand, Environmental
keeps half of its customers and loses the other half to Pollution. Set up this process as a Markov
chain. Determine the transition matrix and sketch a state diagram.

\begin{minipage}{0.5\textwidth}
\[
\begin{array}{c | c c}
& \text{Pollution} & \text{Environmental}\\
\hline
\text{Pollution} & .6 & .4\\
\text{Environmental} & .5 & .5
\end{array}
\]
\end{minipage}
\begin{minipage}{0.5\textwidth}
\begin{center}
\begin{tikzpicture}[shorten >=1pt,node distance=2cm,on grid,auto]
   \node[state] (q_0) {Pol};
   \node[state] (q_1) [right=of q_0] {Env};
   % \node[state] (q_2) [below right=of q_0] {$q_2$};
   % \node[state,accepting](q_3) [below right=of q_1] {$q_3$};
    \path[->]
    (q_0) edge [bend left] node {0.5} (q_1)
          edge [loop left] node {0.6} ()
    (q_1) edge [bend left] node {0.4} (q_0)
          edge [loop right] node {0.5} ();
\end{tikzpicture}
\end{center}
\end{minipage}

\item Abigail spends her entire weekly allowance on either candy or toys. If she buys candy 1 week,
she is 60\% sure to buy toys the next week. The probability that she buys toys in two successive
weeks is 1/5. Set up this process as a Markov chain. Determine the transition matrix and sketch a
state diagram.

\begin{minipage}{0.5\textwidth}
\[
\begin{array}{c | c c}
& \text{Candy} & \text{Toys}\\
\hline
\text{Candy} & .4 & .6\\
\text{Toys} & .8 & .2
\end{array}
\]
\end{minipage}
\begin{minipage}{0.5\textwidth}
\begin{center}
\begin{tikzpicture}[shorten >=1pt,node distance=2cm,on grid,auto]
   \node[state] (q_0) {Candy};
   \node[state] (q_1) [right=of q_0] {Toys};
   % \node[state] (q_2) [below right=of q_0] {$q_2$};
   % \node[state,accepting](q_3) [below right=of q_1] {$q_3$};
    \path[->]
    (q_0) edge [bend left] node {0.8} (q_1)
          edge [loop left] node {0.4} ()
    (q_1) edge [bend left] node {0.6} (q_0)
          edge [loop right] node {0.2} ();
\end{tikzpicture}
\end{center}
\end{minipage}

\item A political scientist in Canada discovered that of the children of Conservatives, 80\% vote
Conservative and the rest vote Labor; of the sons and daughters of Labor supporters, 60\% vote
Labor, 20\% vote Conservative, and 20\% vote for the New Democratic Party (NDP); and of the
offspring of NDP followers, 75\% vote NDP, 15\% vote Labor, and 10\% vote Conservative.
\begin{enumerate}
\item What is the probability that the grandchild of a Conservative will vote for the NDP?
\begin{quote}
The probability that a grandchild of a conservative will vote for the NDP is $(.2)(.2)$
\end{quote}
\item Set up this process as a Markov chain, with steps corresponding to successive generations.
Determine the transition matrix and sketch the state diagram.

\begin{minipage}{0.5\textwidth}
\[
\begin{array}{c | c c c}
& \text{L} & \text{C} & \text{N}\\
\hline
\text{L} & .6 & .2 & .15\\
\text{C} & .2 & .8 & .1\\
\text{N} & .2 & 0 & .75
\end{array}
\]
\end{minipage}
\begin{minipage}{0.5\textwidth}
\begin{center}
\begin{tikzpicture}[shorten >=1pt,node distance=2cm,on grid,auto]
   \node[state] (q_0) {L};
   \node[state] (q_1) [right=of q_0] {C};
   \node[state] (q_2) [right=of q_1] {N};
   % \node[state] (q_2) [below right=of q_0] {$q_2$};
   % \node[state,accepting](q_3) [below right=of q_1] {$q_3$};
    \path[->]
    (q_0)   edge [bend left] node {0.2} (q_1)
            edge [bend left=70] node {0.2} (q_2)
            edge [loop left] node {0.6} ()
    (q_1)   edge [bend left] node {0.2} (q_0)
            edge [loop above] node {0.8} ()
    (q_2)   edge [bend left=70] node {.15} (q_0)
            edge [bend left] node {.1} (q_1)
            edge [loop right] node {.75} ();
\end{tikzpicture}
\end{center}
\end{minipage}
\end{enumerate}

\item A secret CIA report gives the following analysis of the arms race between India and Pakistan:
There are four possible states: War, Total Disarmament, Escalating Arms Race, and De-escalating
Arms Race. It is not possible to change the situation if War or Total Disarmament is occurring this
year. If there is an Escalating Arms Race this year, the probability of continued escalation next
year is .6, of de-escalation next year is .2, and of War next year is .2. If there is a
De-escalating Arms Race this year, the probability of continued de-escalation next year is .7, of
escalation next year is .1, and of total disarmament next year is .2. Set up this process as a
Markov chain. Determine the transition matrix and sketch the state diagram

% \begin{minipage}{0.5\textwidth}
\[
\begin{array}{l | c c c c}
& \text{War} & \text{Disarmed} & \text{Escalating} & \text{De-Escalating}\\
\hline
\text{War}              & 1 & 0 & 0 & 0\\
\text{Disarmed}         & 0 & 1 & 0 & 0\\
\text{Escalating}       & 0.2 & 0 & 0.6 & 0.2\\
\text{De-Escalating}    & 0 & 0.2 & 0.1 & 0.7
\end{array}
\]
% \end{minipage}
% \begin{minipage}{0.5\textwidth}
\begin{center}
\begin{tikzpicture}[shorten >=1pt,node distance=2cm,on grid,auto]
   \node[state] (q_0) {War};
   \node[state] (q_1) [above right=of q_0] {Dis};
   \node[state] (q_2) [below right=of q_0] {Esc};
   \node[state] (q_3) [below right=of q_1] {De-Esc};
   \path[->]
    (q_0)   edge [loop left] node {1} ()
    (q_1)   edge [loop above] node {1} ()
    (q_2)   edge [bend left=70] node {.2} (q_0)
            edge [loop below] node {.6} ()
            edge [bend right] node {.2} (q_3)
    (q_3)   edge [bend left] node {.2} (q_1)
            edge [bend right=-45] node {.1} (q_2)
            edge [loop right] node {.7} ();
\end{tikzpicture}
\end{center}
% \end{minipage}

\item The National League and American League used to alternate as hosts for the opening game of
each year’s World Series. Show that this process can be set up as a Markov process with transition
matrix
\[
P=
\left(
\begin{array}{c c}
0 & 1\\
1 & 0
\end{array}
\right)
\]
\begin{quote}
Since the venues are alternating, there is a 100\% chance that the venue will change, and a 0\%
chance it will stay the same.
\end{quote}

\item A particle moves along a line from an initial position 2 feet to the right of the origin.
Each minute it moves one foot to the right with probability 1/2 or 1 foot to the left. There
are barriers at the origin and 4 feet to the right of the origin; if the particle hits a barrier,
it remains there. Show that this process can be set up as a Markov process with five states.
Determine the transition matrix and draw the state diagram.

\begin{minipage}{0.5\textwidth}
\[
\begin{array}{c | c c c c c}
& \text{L} & \text{R} & \text{S} & \text{RB} & \text{LB}\\
\hline
\text{L} & .25 & .5 & 0 & 0 & .25\\
\text{R} & .5 & .25 & 0 & .25 & 0\\
\text{S} & 0 & 0 & 1 & 0 & 0\\
\text{RB} & 0 & 0 & 1 & 0 & 0\\
\text{LB} & 0 & 0 & 1 & 0 & 0
\end{array}
\]
\end{minipage}
\begin{minipage}{0.5\textwidth}
\begin{center}
\begin{tikzpicture}[shorten >=1pt,node distance=2cm,on grid,auto]
   \node[state] (left) {L};
   \node[state] (right) [right=of left] {R};
   \node[state] (rb) [below=of right] {RB};
   \node[state] (lb) [below=of left] {LB};
   \node[state] (stop) [right=of right] {S};
   \path[->]
        (left)  edge [loop above] node {.25} ()
                edge [bend right, swap] node {.5} (right)
                edge [bend right] node {.25} (lb)
        (right) edge [loop above] node {.25} ()
                edge [bend right] node {.25} (rb)
                edge [bend right, swap] node {.5} (left)
        (stop)  edge [loop right] node {1} ()
        (lb)    edge [bend right=70, swap] node {1} (stop)
        (rb)    edge [bend right] node {1} (stop);
\end{tikzpicture}
\end{center}
\end{minipage}

\item The random walk of Exercise 6 is modified so that if the particle reaches the barrier at the
origin, it must move one foot to the right in the next minute, while if it hits the other barrier,
it must move one foot to the left in the next minute. Determine the transition matrix for the
associated Markov chain.
\[
\begin{array}{c | c c c c c}
& \text{L} & \text{R} & \text{S} & \text{RB} & \text{LB}\\
\hline
\text{L} & .25 & .5 & 0 & 0 & .25\\
\text{R} & .5 & .25 & 0 & .25 & 0\\
\text{S} & 0 & 0 & 1 & 0 & 0\\
\text{RB} & 1 & 0 & 0 & 0 & 0\\
\text{LB} & 0 & 1 & 0 & 0 & 0
\end{array}
\]

\item Find the probability that a woman whose birthweight was average has a granddaughter with an
average birthweight.
\[
\begin{array}{c | c c}
& \text{Average} & \text{Not Average}\\
\hline
\text{Average} & .5 & .5\\
\text{Not Average} & .5 & .5
\end{array}
\]
\begin{quote}
The probability that a granddaughter will have average weight is $(.5)(.5)=.25$.
\end{quote}

\item Sketch the state diagram for the Lower Pine Cone College example.
\begin{center}
\begin{tikzpicture}[shorten >=1pt,node distance=2cm,on grid,auto]
   \node[state] (q_0) {F};
   \node[state] (q_1) [above right=of q_0] {T};
   \node[state] (q_2) [below right=of q_0] {R};
   \node[state] (q_3) [below right=of q_1] {P};
   \path[->]
    (q_0)   edge [loop left] node {1} ()
    (q_1)   edge [loop above] node {1} ()
    (q_2)   edge [loop below] node {$c$} ()
            edge [bend right] node {$g$} (q_3)
    (q_3)   edge [bend right=-45] node {$r$} (q_2);
\end{tikzpicture}
\end{center}

\setcounter{enumi}{10}

\item Using probabilistic considerations only, show that the square of a stochastic matrix is also a
stochastic matrix.
\begin{quote}

\end{quote}
\end{enumerate}

\end{document}

### 2016-04-03 number_theory/hw/hw9.tex
ethics final paper

\documentclass{hw}

\begin{document}
\makeheader{9}

\begin{enumerate}
\item Directly calculate $\displaystyle\sum\limits_{d|12}\phi(d)$ and verify that you obtain 12 as your
answer.
\begin{align*}
\sum_{d|12}\phi(d) &= \phi(1) + \phi(2) + \phi(3) + \phi(4) + \phi(6) + \phi(12)\\
&= 1 + 1 + 2 + 2 + 2 + 4\\
&= 12
\end{align*}
\end{enumerate}

\end{document}

### 2016-04-03 writing_in_ethics/paper_four.tex
algo homework
working on number theory
ethics final paper

--- 2016-03-08 writing_in_ethics/paper_three.tex
@@
 \documentclass{paper}
-\title{Title}
+\title{The Code of Ethics}
 
-\addbibresource{generic_bib.bib}
+\addbibresource{pfour.bib}
 
 \begin{document}
-\drafth{}
+\finalh{}
 \inlinetitle
@@
 
+The ACM code of ethics provides those in technical industries with guidelines to follow when making tough ethical decisions.
+One such ethical decision arose in January of 1986 when NASA's Challenger space shuttle exploded above Cape Canaveral due to the neglect of suggestions from various engineers who were part of the team.
+One engineer, Roger Boisjoly, testified against NASA and the the engineering company Morton Thiokol during the trial, despite the recommendations given Morton Thiokol's lawyers.
+Although Boisjoly broke some of the guidelines of the ACM code of Ethics, his actions were not unethical.
+
+The explosion of the Challenger space shuttle could have been prevented.
+Boisjoly attempted to warn NASA and the other engineers multiple times.
+The Code of Ethics explicitly states that one should avoid harm to others.
+In Boisjoly's letter to the vice president of engineering, he only noted that the O-ring failure would result in damages to the shuttle and the launch pad, but not the loss of human life.
+It is possible to argue that Boisjoly did not stress the seriousness of the situation enough to convince his managers to take action.
+In this regard, Boisjoly did not completely comply with the Code of Ethics.
+The Code notes
+\begin{quote}
+Well-intended actions...may lead to harm unexpectedly.
+One way to avoid unintentional harm is to carefully consider potential impacts on all those affected by decisions made during design and implementation.
+\end{quote}
+By only stressing the damage that could be done to the property, and not to human life, Boisjoly violated the code of ethics.
+
+Despite this pseudo-violation, Boisjoly did uphold many of the other guidelines of the Code of Ethics.
+Arguably one of the most important ideas outlined by the Code is that one should be honest and trustworthy.
+Despite the pressure from his peers who did not listen or acknowledge his warnings, he continued to warn NASA and Morton Thiokol about the dangers of launching the shuttle.
+By not accepting the faulty device, Boisjoly was upholding the ACM Code.
+In addition, Boisjoly could have easily gone to the press and released details about the O-ring and NASA's carelessness, but he did not.
+Instead, he worked to tell others inside the company about what was happening.
+The ACM Code states that one should honor confidentiality, which Boisjoly did by not whistle-blowing on the company until the time came for him to testify in court.
+One can argue that it would have actually been ethical to make the information public, which would have put pressure on NASA and Morton Thiokol to reconsider the launch.
+Boisjoly's warnings may have been in vain regardless, as the issues were not heavily considered until the night before the launch.
+In addition, NASA prevented the communication of the issues to anyone despite the situation, so it may not have been possible for Boisjoly to make the issue public even if he had gone to an outside source.
+
+When Boisjoly testified in court, he told the jury about the faulty O-ring inside the rocket.
+This action can be seen as ethical under the ACM Code.
+If Boisjoly had omitted the details and obeyed the orders given to him by Morton Thiokol's lawyers, he would have been violating the Code in several ways.
+One aspect of the Code of Ethics is that one should provide appropriate professional review.
+Boisjoly had already done this numerous times before the Challenger shuttle was destroyed.
+During his testimony, he presented the same information that was already known by the managers at NASA and Morton Thiokol.
+In addition, if Boisjoly had not told the panel what he knew, then he would have been omitting the truth.
+The Code of Ethics states that a professional must be honest about circumstances that may cause a conflict of interest.
+Omission of the truth, especially when dealing with loss of human life, would have certainly been a conflict of interest.
+One can argue that by not following the advice of Morton Thiokol's lawyers was in violation of the Code of Ethics, since he did not keep engineering work confidential.
+The Code, however, states that the obligation to keep work confidential can be changed when the law or other sections of the Code require that the information be divulged.
+
+Overall, Boisjoly's actions were not completely in violation of the ACM Code of Ethics.
+While he did not fully consider the implications of his actions, he did everything he could to act ethically given the circumstances.
+Ultimately, Boisjoly left his job at Morton Thiokol due to pressure related to his actions during the trial, but his actions ensured that a disaster of this magnitude could never take place again.
+By following the ACM Code of Ethics, Boisjoly was able to act ethically and make decisions that will help others in the future.
 \end{linenumbers}
@@
 
+\nocite{*}
 \printbibliography

### 2016-04-05 math_modeling/hw/hw6/hw6.tex
working on modeling
working on number theory

@@
+%% Quiz 4 on page 22/23
+
 \documentclass{hw}
+
+\usepackage{multicol}
 
@@
 \hline
-\text{L} & .6 & .2 & .15\\
-\text{C} & .2 & .8 & .1\\
-\text{N} & .2 & 0 & .75
+\text{L} & .6 & .2 & .2\\
+\text{C} & .2 & .8 & 0\\
+\text{N} & .15 & .1 & .75
 \end{array}
@@
 \begin{quote}
+If it was not a stochastic matrix, then there would be no way to represent the next state in the
+system. The total probability of all the transitions must sum to 1, since there is a 100\% chance
+of a transition occurring in every step of the chain.
+\end{quote}
 
+\item A stochastic matrix is doubly stochastic if the sum of the entries in each column is 1. If $A$
+and $B$ are doubly stochastic square matrices, is $A^{2}$ doubly stochastic? Is $AB$?
+\begin{quote}
+Yes, the square of a doubly stochastic matrix is still a doubly stochastic matrix. $AB$ is also
+still doubly stochastic.
+\end{quote}
+
+\item Write out inductive proofs for Theorems 2 and 3.
+
+\item Use matrix multiplication to solve Exercise 8.
+\[
+\left(
+\begin{array}{c c}
+.5 & .5\\
+.5 & .5
+\end{array}
+\right)
+\cdot
+\left(
+\begin{array}{c c}
+.5 & .5\\
+.5 & .5
+\end{array}
+\right)
+=
+\]
+
+\item Abigail bought a toy with her allowance this week (see Exercise 2). Find the probability that
+she will buy a toy 4 weeks from now.
+
+\item Find the distribution of birth weights (Example 1) after one generation if the initial
+probability distribution is (.4, .3, .3).
+
+\item Suppose the distribution of birth weights of a generation of daughters is
+$p^{(1)}=7(0.31,0.45,0.24)$. Can you find the distribution of birth weights of the mothers?
+
+\item Use the state diagram in Fig. 11.3 to find the probability that the process reaches state
+$S_{3}$ in two steps if it starts in state $S_{1}$.
+
+\item Let $P$ be the transition matrix of Eq. (4). Compute $P^{2}$. What can you say about the
+first column of $P^{3}$ and $P^{n}$?
+
+\setcounter{enumi}{20}
+\item The matrices determined in Exercises 1-7 are all transition matrices for Markov processes.
+Which ones are regular?
+\begin{quote}
+Exercises 1 and 2 are regular matrices, since they have all positive entries.
+\end{quote}
+
+\item Find, if possible, a fixed-point vector for each of the transition matrices of Exercises 1-7.
+\begin{multicols}{2}
+\begin{enumerate}
+\item $\displaystyle \mathbf{w}=\left\{{5\over9}, {4\over9}\right\}$
+\item $\displaystyle \mathbf{w}=\left\{{4\over7}, {3\over7}\right\}$
+\item $\displaystyle \mathbf{w}=\left\{{23\over70}, {43\over70}, {2\over35}\right\}$
+\item This is not a regular matrix, so no $\bf{w}$ exists.
+\item This is not a regular matrix, so no $\bf{w}$ exists.
+\item This is not a regular matrix, so no $\bf{w}$ exists.
+\item This is not a regular matrix, so no $\bf{w}$ exists.
+\end{enumerate}
+\end{multicols}
+
+\item Does a matrix of the form
+$\displaystyle A=\left(\begin{array}{c c}a & b \\ -b & d\end{array}\right)$ have a fixed-point vector?
+\begin{quote}
+If it has a fixed-point vector, then that vector is given by
+\[
+\mathbf{w}=\left\{ {c\over b+c},{b\over b+c} \right\}
+= \left\{ {-b\over b-b},{b\over b-b}\right\}
+= \left\{ {-b\over 0},{b\over 0}\right\}.
+\]
+Therefore, there is no fixed-point vector for the matrix.
+\end{quote}
+
+\setcounter{enumi}{31}
+\item Which of the transition matrices of Exercises 1-7 represent absorbing Markov processes?
+\begin{quote}
+Exercises 4 and 6 represent absorbing Markov processes, since they have nodes which are not
+escapable.
 \end{quote}

### 2016-04-05 number_theory/hw/hw9.tex
working on number theory

@@
 \end{align*}
+
+\item Suppose that $p_{1},p_{2},\dots,p_{N}$ are distinct primes. Prove that
+\[
+{\phi(p_{1}p_{2}\dots p_{N}) \over p_{1}p_{2}\dots p_{N}}
+=
+\prod_{n=1}^{N}\left(1-{1\over p_{n}}\right)
+\]
+\begin{align*}
+{\phi(p_{1}p_{2}\dots p_{N}) \over p_{1}p_{2}\dots p_{N}} &=
+{\phi(p_{1})\phi(p_{2})\dots \phi(p_{N}) \over p_{1}p_{2}\dots p_{N}}\\
+&= \prod_{n=1}^{N}{\phi(p_{n})\over p_{n}}\\
+&= \prod_{n=1}^{N}{p_{n} -1 \over p_{n}}\\
+&= \prod_{n=1}^{N}\left(1-{1\over p_{n}}\right)
+\end{align*}
+
+\item Find a value of $n$ such that $\phi(n)/n < 1/4$. What do you think is a good strategy for
+choosing $n$ so that $\phi(n)/n$ is close to zero?
+\begin{quote}
+One value for which this holds true is $n=210$. $\phi(210)$ is 48, and $48/210 = 8/35 < 1/4$. One
+strategy for finding these numbers would be to
+\end{quote}
+
+\item Suppose that $p$ is prime and $m$ and $n$ are non-negative integers.
+\begin{enumerate}
+\item Prove that $\phi(p^{m+n})\geq\phi(p^{m})\phi(p^{n})$.
+\begin{quote}
+We can consider $\phi(p^{m+n})$. If we let $m+n=j$, then we have $\phi(p^{j})$, which can be
+expressed as $p^{j}-p^{j-1}$. If we consider $\phi(p^{m})\phi(p^{n})$, we have
+\begin{align*}
+\phi(p^{m})\phi(p^{n})&= (p^{m}-p^{m-1})(p^{n}-p^{n-1})\\
+&= p^{m+n} - 2p^{m+n-1} + p^{m+n-2}\\
+&= p^{j}-2p^{j-1}+p^{j-2}.
+\end{align*}
+If we compare the two values, we get
+\begin{align*}
+p^{j}-p^{j-1} &\stackrel{?}{\geq} p^{j}-2p^{j-1}+p^{j-2}\\
+p^{j-1} &\geq p^{j-2}.
+\end{align*}
+We know this is true since $m,n>0$, so $j>0$.
+\end{quote}
+
+\item Under what additional assumptions on $m$ and $n$ do we obtain
+$\phi(p^{m+n})=\phi(p^{m})\phi(p^{n})$?
 \end{enumerate}
 
+\end{enumerate}
 \end{document}

### 2016-04-05 algorithms/HELLO.CPP

@@ -1,0 +1,298 @@
+#include <fstream>
+// #include "apstring.h"
+// #include "apstring.cpp"
+#include <string>
+#include <iostream>
+#include <iomanip>
+
+using namespace std; 	//introduces namespace std
+
+const int maxVotes=100;
+const int states=50;
+ofstream dataout;
+
+struct state
+{
+	string name;
+	int votes;
+	double power;
+	int population;
+};
+
+int voteSum(int votes[states]);
+int sum(int pass[states]);
+void calculateBanzhaf(int votetype[maxVotes], int required, double power[states]);
+void calculateBanzhafOveride(int votetype[maxVotes], int required, double power[maxVotes]);
+int pass(int vote[states]);
+void printState(state a);
+void printStateFile(state a);
+void printHeader();
+void printHeaderFile();
+double combin(int n, int r);
+int worst(state UnitedStates[states]);
+
+int main( void )
+{
+	const string stateNames[states] = {
+    "Alabama","Alaska","Arizona","Arkansas","California","Colorado","Connecticut","Delaware","Florida",
+    "Georgia","Hawaii","Idaho","Illinois","Indiana","Iowa","Kansas","Kentucky","Louisiana","Maine",
+    "Maryland","Massachusetts","Michigan","Minnesota","Mississippi","Missouri","Montana","Nebraska","Nevada","New Hampshire",
+    "New Jersey","New Mexico","New York","North Carolina","North Dakota","Ohio","Oklahoma","Oregon","Pennsylvania","Rhode Island",
+    "South Carolina","South Dakota","Tennessee","Texas","Utah","Vermont","Virginia","Washington","West Virginia","Wisconsin",
+    "Wyoming"};
+	int totalVotes;
+	int required;
+	int worstRep;
+	const long populations[states]={4447100, 626932, 5130632, 2673400, 33871648, 4301261, 3405565, 783600, 15982378,
+	8186453, 1211537, 1293953, 12419293, 6080485, 2926324, 2688418, 4041769, 4468976, 1274923,
+	5296486, 6349097, 9938444, 4919479, 2844658, 5595211, 902195, 1711263, 1998257, 1235786,
[…]

### 2016-04-07 math_modeling/class/notes/stochastics.tex
number theory

@@
 
+\section*{Irreducible Matrices}
+We attempted to find the $\pi$ vector in the previous examples. We can use the following theorem to
+discuss special Markov Chains.
+\begin{quote}
+Let $P$ be the transition matrix for an irreducible and aperiodic Markov Chain. Then there exists
+a unique probability vector $\pi$ such that
+\[
+\pi\cdot P=\pi.
+\]
+\end{quote}
+It follows that if $P$ is a regular matrix, then we can also find the $\pi$ vector. A more
+interesting phenomenon occurs when we do not have a regular $P$ matrix. A typical model of this can
+be seen in the Drunkard's walk without reflection.
+
+\subsection*{The Drunkard's Walk}
+The Drunkard's walk represents a random walk with several states. At some states, the drunk man may
+either end up in his bed or on the street. In the reflective version of this walk, it is possible that
+after going into his bed, the drunkard goes back out. In this example, we want to consider absorbing
+states, which are states that are not able to be escaped.
+
+In this walk, we want to model the drunkard by saying that if he falls asleep in the street or he
+goes to his bed, then the walk ends. Those conditions are representative of absorbing states.
+\[
+P=
+\left(
+\begin{array}{c c c c c}
+1 & 0 & 0 & 0 & 0\\
+1/2 & 0 & 1/2 & 0 & 0\\
+0 & 1/2 & 0 & 1/2 & 0\\
+0 & 0 & 1/2 & 0 & 1/2\\
+0 & 0 & 0 & 0 & 1
+\end{array}
+\right).
+\]
+More generally, an absorbing matrix takes the form of an $r\times r$ matrix with $k$ absorbing
+states. It is a fact that ultimately everything will end up in the absorbed state. Markov Chains
+are only absorbing when all absorbing state can be reached. If a state is not absorbing, it is
+called a transient state.
+
+For the drunkard's walk, we can partition the matrix into four parts:
+\[
+P=
+\left(
+\begin{array}{c c c c c}
+1 & 0 & 0 & 0 & 0\\
+1/2 & 0 & 1/2 & 0 & 0\\
+0 & 1/2 & 0 & 1/2 & 0\\
+0 & 0 & 1/2 & 0 & 1/2\\
+0 & 1/2 & 0 & 1/2 & 0
+\end{array}
+\right)
+=
+\left(
+\begin{array}{c c c c c}
+1 & 0 & 0 & 0 & 0\\
+0 & 1 & 0 & 0 & 0\\
+1/2 & 0 & 0 & 1/2 & 0\\
+0 & 0 & 1/2 & 0 & 1/2\\
+0 & 0 & 0 & 0 & 1
+\end{array}
+\right)
+\]
+which allows us to calculate a new matrix $N=(I-Q)^{-1}$. Every $n_{ij}$ entry in the $N$ matrix
+tells us the average number of times that the process is in the $j$ state if it starts in the $i$
+state.
+
 \end{document}

### 2016-04-07 number_theory/hw/hw9.tex
number theory

@@
 $\phi(p^{m+n})=\phi(p^{m})\phi(p^{n})$?
+\begin{quote}
+If either $m$ or $n$, but not both $m$ and $n$ are zero, then we have
+$\phi(p^{m+0})=\phi(p^{m})\phi(p^{0})=\phi(p^{m})$ or
+$\phi(p^{0+n})=\phi(p^{0})\phi(p^{n})=\phi(p^{n})$.
+\end{quote}
 \end{enumerate}
 
+\item Suppose that $a$ and $b$ are positive integers.
+\begin{enumerate}
+\item Prove that $\phi(ab)\geq\phi(a)\phi(b)$.
+\begin{quote}
+Suppose that $a$ and $b$ are not relatively prime and consider the product $\phi(a)\phi(b)$. Then
+\begin{align*}
+\phi(a)\phi(b) &<
+\phi(p_{1}^{\alpha_{1}}p_{2}^{\alpha_{2}}\dots p_{r}^{\alpha_{r}})
+\phi(p_{1}^{\beta_{1}}p_{2}^{\beta_{2}}\dots p_{r}^{\beta_{r}})\\
+&= \phi(p_{1}^{\alpha_{1}})\phi(p_{2}^{\alpha_{2}})\dots \phi(p_{r}^{\alpha_{r}})
+\phi(p_{1}^{\beta_{1}})\phi(p_{2}^{\beta_{2}})\phi(p_{r}^{\beta_{r}})\\
+&= \prod_{i=1}^{r}\left(p_{i}^{\alpha_{i}}-p_{i}^{\alpha_{i}-1}\right)
+\prod_{i=1}^{r}\left(p_{i}^{\beta_{i}}-p_{i}^{\beta_{i}-1}\right)\\
+&= \prod_{i=1}^{r}p_{i}^{\alpha_{i}}\left(1-{1\over p_{i}}\right)
+\prod_{i=1}^{r}p_{i}^{\beta_{i}}\left(1-{1\over p_{i}}\right)\\
+&= ab\prod_{i=1}^{r}\left(1-{1\over p_{i}}\right)^{2}.
+\end{align*}
+We can also consider the function $\phi(ab)$.
+\begin{align*}
+\phi(ab) &= \phi(p_{1}^{\alpha_{1}}p_{2}^{\alpha_{2}}\dots p_{r}^{\alpha_{r}}
+p_{1}^{\beta_{1}}p_{2}^{\beta_{2}}\dots p_{r}^{\beta_{r}})\\
+&= \phi(p_{1}^{\alpha_{1}+\beta_{1}}p_{2}^{\alpha_{2}+\beta_{2}}\dots p_{r}^{\alpha_{r}+\beta_{r}})\\
+&= \phi(p_{1}^{\alpha_{1}+\beta_{1}})\phi(p_{2}^{\alpha_{2}+\beta_{2}})\dots
+\phi(p_{r}^{\alpha_{r}+\beta_{r}})\\
+&= \prod_{i=1}^{r}p_{i}^{\alpha_{i}+\beta_{i}}-p_{i}^{\alpha_{i}+\beta_{i}}{1\over p}\\
+&= \prod_{i=1}^{r}p_{i}^{\alpha_{i}+\beta_{i}}\left(1-{1\over p_{i}}\right)\\
+&= ab\prod_{i=1}^{r}\left(1-{1\over p_{i}}\right).
+\end{align*}
+Since
+\[
+\phi(ab)=ab\prod_{i=1}^{r}\left(1-{1\over p_{i}}\right) >
+ab\prod_{i=1}^{r}\left(1-{1\over p_{i}}\right)^{2}>\phi(a)\phi(b),
+\]
+we have
+\[
+\phi(ab) > \phi(a)\phi(b).
+\]
+If $gcd(a,b)=1$, then we can say $\phi(ab) = \phi(a)\phi(b)$. Therefore $\phi(ab)\geq\phi(a)\phi(b)$.
+\end{quote}
+
+\item Prove that $\phi(ab) = \phi(a)\phi(b)$ if and only if $gcd(a,b)=1$.
+\begin{quote}
+$\Rightarrow)$ Suppose $\phi(ab) = \phi(a)\phi(b)$ implies $gcd(a,b)>1$. We know from $(a)$ that
+when $a$ and $b$ are not relatively prime, $\phi(ab) > \phi(a)\phi(b)$, which is a direct contradiction
+to $\phi(ab) = \phi(a)\phi(b)$. Therefore $\phi(ab) = \phi(a)\phi(b)$ implies $gcd(a,b)=1$.
+\noindent\\\\
+$\Leftarrow)$ Suppose $gcd(a,b)=1$. Then $a$ and $b$ are relatively prime, so
+$\phi(ab)=\phi(a)\phi(b)$ by the theorem.
+\end{quote}
+\end{enumerate}
+
+\item Suppose that $n$ is a positive integer and $k$ is any integer.
+\begin{enumerate}
+\item Prove that $gcd(n,k)=1$ if and only if $gcd(n,n-k)=1$.
+\begin{quote}
+$\Leftarrow)$ Suppose that $gcd(n,k)>1$. Then $gcd(n,k)=q$, so $q\mid n$ and
+$q\mid k$, $n=qa$ and $k=qb$ for some $a,b\in\ZZ$. Then $n-k=pa-pb=p(a-b)$, so $p\mid n-k$, which
+implies that $gcd(n,n-k)>1$. Therefore, by contrapositive, $gcd(n,k)=1$ implies $gcd(n,n-k)=1$.
+\noindent\\\\
+$\Rightarrow)$ Suppose $gcd(n,n-k)>1$. Then there exists $p\in\ZZ$ such that $p\mid n$ and $p\mid k$.
+Then $n=pa$ and $k=pb$ for some $a,b\in\ZZ$. We can express $n-k=pa-pb=p(a-b)$, so $p\mid(n-k)$ and
+$p\mid n$, and $gcd(n,n-k)>1$.
+\end{quote}
+\item Prove that $\phi(n)$ is an even integer for all $n\geq3$.
+\begin{quote}
+By definition, the totient function counts the number of units in $\ZZ_{n}$. Suppose that
+$n\geq3$. If we take any element $k$ from $\ZZ_{n}$, we know by $(a)$ that if $k$ is relatively
+prime to $n$, then $n-k$ is relatively prime to $n$. Therefore, if $k$ is a unit in $\ZZ_{n}$, then
+$n-k$ is also a unit in $\ZZ_{n}$. Then for every unit $k$ in $\ZZ_{n}$, we can find another unit
+$n-k$ also in $\ZZ_{n}$, which means there are an even number of units in $\ZZ_{n}$, so
+$\phi(n)\in\{ 2j\ |\ j\in\ZZ \}$.
+\end{quote}
+\end{enumerate}
+
+\item If $n$ is a positive integer prove that $\phi(n)=2$ if and only if $n\in\{ 3,4,6\}$.
+\begin{quote}
+$\Rightarrow)$
+\noindent\\\\
+$\Leftarrow)$ Suppose $n\in\{ 3,4,6\}$. Then $\phi(3)=2$, $\phi(4)=2$, and $\phi(6)=2$.
+\end{quote}
 \end{enumerate}

### 2016-04-07 number_theory/hw/hw9.tex
finished pres for now

@@
 prime to $n$, then $n-k$ is relatively prime to $n$. Therefore, if $k$ is a unit in $\ZZ_{n}$, then
-$n-k$ is also a unit in $\ZZ_{n}$. Then for every unit $k$ in $\ZZ_{n}$, we can find another unit
-$n-k$ also in $\ZZ_{n}$, which means there are an even number of units in $\ZZ_{n}$, so
-$\phi(n)\in\{ 2j\ |\ j\in\ZZ \}$.
+$n-k$ is also a unit in $\ZZ_{n}$. If $n$ is odd, then we know that there will not be any situation
+where $n-k=k$, since that would imply that $n=2k$, or $n$ is even which is a contradiction. Then for
+every unit $k$ in $\ZZ_{n}$, we can find another unit $n-k$ also in $\ZZ_{n}$, which means there are
+an even number of units in $\ZZ_{n}$, so $\phi(n)\in\{ 2j\ |\ j\in\ZZ \}$ when $n$ is odd. If $n$
+is even, however, there may be a unit $k$ in $\ZZ_{n}$ such that $k=n-k$. If this is the case,
+then this unit will be the same as $n/2$. Since it is a unit, then $gcd(n/2,n)=1$. We saw that
+$n$ was even, so we can say $n=2l,\ l\in\ZZ$. Then $gcd(2l/2,2l)=gcd(l,2l)\neq1$, so there cannot
+be a unit $k$ in an even $\ZZ_{n}$ such that $k=n-k$.
 \end{quote}

### 2016-04-07 pde_two/threshold.tex
finished pres for now

\documentclass{hw}
\usepackage{amsmath,amssymb}

\title{Threshold Phenomenon in a Nonlinear Heat Equation}
\author{}
\date{}

\begin{document}
\maketitle
\section{Introduction}

This paper concerns the behavior of the following non-linear heat equation given by
\begin{gather}
u_{t}=u_{xx}+u^{3},\ 0<x<\pi,\ t>0\\
u(0,t)=u(\pi,t)=0,\ t>0\\
u(x,0)=u_{0}(x),\ 0<x<\pi.
\end{gather}

This system corresponds to several physical situation, including exothermic reactions between
chemicals. We know that this equation has a solution $u(x,t)$, but it may help us to consider the
\textit{steady state} equation corresponding to this system. The steady state arises from holding
time constant; thus, our differential equation becomes dependent only on $x$.

\section{Steady State Analysis}
The steady state equation of $(1)$ is given by
\begin{gather}
u_{xx}+u^{3}=0\\
u(0)=u(\pi)=0.
\end{gather}
We want to find a solution $u_{s}(x)$ that satisfies the following properties
\begin{enumerate}
\item If $u_{0}(x)<u_{s}(x)$, then $u(x,t)\to u_{s}(x)$ as $t\to\infty$.
\[
\lim_{t\to \infty}u(x,t)=u_{s}(x).
\]
In this case, the solution to $(1)$ would independent of time. If this property is fulfilled, then
there is a \textit{global solution} $u(x,t)$ to the boundary value problem described in $(1)$.
\item If $u_{0}(x)>u_{s}(x)$, then $u(x,t)$ blows up in a finite time. This means that the solution
tends towards infinity in a finite time $T$.
\[
\lim_{t\to T}u(x,t)=\infty.
\]
This case if often indicative of a singularity in the solution $u(x,t)$. The solution $u$ of $(1)$
would not be global in this case.
\end{enumerate}
We can now attempt to solve this steady state problem.

\subsection{Solution to the Steady State Problem}
We are attempting to solve the ordinary differential equation
\[
u''=-u^{3}.
\]
Multiplying both sides of the equation gives us
\begin{align*}
u''u'&=-u'u^{3}\\
\int u''u'&=\int -u'u^{3}\\
{1\over2}(u')^{2}&= -{1\over4}u^{4}+c\\
2(u')^{2}&=-u^{4}+c_{1}.
\end{align*}
This first order equation does not lend itself to separation of variables as readily as we would like
(although it is possible if we want to deal with imaginary numbers). Instead, we will use a technique
called the \textit{shooting method} in order to determine a solution. The shooting method provides
us with a way to relax boundary conditions into initial conditions. Recall that $(5)$ provides us
with boundary conditions that arose from the steady state of $(1)$. We will leave the first boundary
condition alone and change the second boundary condition so that

\end{document}

### 2016-04-08 number_theory/hw/hw9.tex
finished number theory

@@
 One value for which this holds true is $n=210$. $\phi(210)$ is 48, and $48/210 = 8/35 < 1/4$. One
-strategy for finding these numbers would be to
+strategy for finding these numbers would be to pick values for which $\phi(n)$ is small but has
+many unique prime factors.
 \end{quote}
@@
 \begin{quote}
-$\Rightarrow)$
+$\Rightarrow)$ Suppose $\phi(n)=2$ and let $k\in\ZZ > 2$ and $l\in\ZZ > 1$. We want to consider the
+case where $n=2^k3^l$, so $\phi(n)=\phi(2^k3^l)=(2^k-2^{k-1})(3^l-3^{l-1})$. This can be further
+simplified to $2^k(1/2)3^l(2/3)=2^k3^l(1/3)>2$ for $l>1$ and $k>2$. Therefore we only have
+$\{ 1,2,3,4,6,12 \}$ to consider. Since $\phi(1)=\phi(2)=1$ and $\phi(12)=4$, we can eliminate
+those values. However, $\phi(3)=\phi(4)=\phi(6)=2$, so $\phi(n)=2$ when $n\in\{ 3,4,6\}$.
 \noindent\\\\

### 2016-04-12 math_modeling/class/notes/stochastics.tex
number theory and modeling

@@
 \documentclass{hw}
+
+\usepackage{tikz}
+\usetikzlibrary{automata,positioning}
 
@@
 
+\section*{Poisson Probabilities}
+
+Suppose that $f(x,t)$ is the probability of getting $x$ successes during a time interval of length
+$t$. Assume the prob of success during a very small time internal from $t$ to $t+\delta t$ is $\mu$,
+the probability of more than one success during a time interval is negligible, and the probability
+of a success during such a time interval does not depend on what happened prior to time $t$. The
+Poisson distribution is given by
+\[
+f(x,t+\Delta t)=f(x,t)\cdot[1-\mu\Delta t]+f(x-1,t)\mu\Delta t = {e^{-\mu}\mu^{x}\over x!}
+\]
+The variable $x$ is indicative of the number of successes we want in the next time interval
+$t+\delta t$. One example of the use of Poisson distributions is to calculate the probability that
+an earthquake will occur again in a certain area.\\\\
+
+\begin{quote}
+\textit{\textbf{Example}}\\
+Suppose that in southern California, 5 earthquakes occur every 100 years on average. What is the
+probability that 2 earthquakes will occur in the next century?\\\\
+
+The answer is very straightforward:
+\[
+P={e^{-5}5^{2}\over 2!}
+\]
+If we wanted to know the probability in the next 10 years, we can calculate
+\[
+P={e^{-0.5}0.5^{2}\over 2!}.
+\]
+\end{quote}
+
+\section*{Markov Chains and Continuity}
+We can form a relationship between Markov Chains and differential equations.
+\begin{quote}
+\textit{\textbf{Example}}\\
+In the summer, an AC unit is in one of 3 states: (0) off, (1) low, or (2) high. While off,
+transitions to low occur after an exponential time with an expected time of 3 minutes.
+\begin{center}
+\begin{tikzpicture}[shorten >=1pt,node distance=2cm,on grid,auto]
+    \node[state] (off) {off};
+    \node[state] (low)[right=of off] {low};
+    \node[state] (high)[right=of low] {high};
+    \path[->]
+    (off) edge[bend right,swap] node {$1/3$} (low)
+    (low) edge[bend right,swap] node {$1/4$} (off)
+          edge[bend right,swap] node {$1/4$} (high)
+    (high) edge[bend right,swap] node {$1/3$} (low);
+\end{tikzpicture}
+\end{center}
+We assign a $\lambda$ value to each state where $\lambda\in(0,\infty)$. The density function is
+given by
+\[
+f(t)=\lambda e^{-\lambda t}
+\]
+for $t>0$. We can call the waiting time $T_{n}$, where $n$ is the state where we want to move. If
+we have $T_{1}$ and $T_{2}$, the smaller $T$ is the state to which we move.
+\end{quote}
+
+\section*{Various Distributions}
+\begin{center}
+\bgroup
+\def\arraystretch{3.5}
+\begin{tabular}{|c | c | c | c|}
+\hline
+Binomial Distribution & $B(x;n,p)$ & $\displaystyle \binom{n}{x}p^{x}q^{n-x}$ & Discreet\\
+\hline
+Poisson Distribution & $\rho(x,\mu)$ & $\displaystyle {e^{-\mu}\mu^{x}\over x!}$ & Discreet\\
+\hline
+Geometric Distribution & $G(x;p)$ & $\displaystyle p(1-p)^{x-1}$ & Discreet, Memoryless\\
+\hline
+Bell Curve & $B(\sigma,x,\mu)$ & $\displaystyle \int_{a}^{b}{1\over \sigma\sqrt{2\pi}}e^{-0.5\left({x-\mu\over\sigma}\right)^{2}}\dx$ & Continuous\\
+\hline
+Exponential Distribution & $E(t,\theta)$ & $\displaystyle\int_{a}^{b}{1\over\theta}e^{-t/\theta}\dt$ & Continuous, Memoryless\\
+\hline
+\end{tabular}
+\egroup
+\end{center}
+
 \end{document}

### 2016-04-12 math_modeling/hw/hw6/hw6.tex
finished pres for now
number theory and modeling

@@
 \begin{center}
-\begin{tikzpicture}[shorten >=1pt,node distance=2cm,on grid,auto]
+\begin{tikzpicture}%[shorten >=1pt,node distance=15cm,on grid,auto]
    \node[state] (q_0) {L};
@@
 \begin{quote}
-The probability that a granddaughter will have average weight is $(.5)(.5)=.25$.
+The probability that a granddaughter will have average weight is $(.5)(.5) + (.5)(.5)=.5$.
 \end{quote}
@@
 \item Write out inductive proofs for Theorems 2 and 3.
+\begin{quote}
+\textbf{Theorem 2}: From the text, we know
+\[
+p^{(0)}=(p_{1}^{(0)},p_{2}^{(0)},\dots p_{r}^{(0)})=
+(p_{1},p_{2},\dots p_{r})=p^{0}\cdot p.
+\]
+We can assume that $p^{(n)}=p^{n}$, and we want to show that $p^{(n+1)}=p^{n+1}$. We know that
+\begin{align*}
+p^{(n+1)}&=(p_{1}^{(n+1)},p_{2}^{(n+1)},\dots p_{r}^{(n+1)})\\
+&= (p_{1}^{(n)}p_{1}^{(1)},p_{2}^{(n)}p_{2}^{(1)},\dots p_{r}^{(n)}p_{r}^{(1)})\\
+&= p^{n+1}.
+\end{align*}
+\textbf{Theorem 3}: We know that $p^{(n)}=p^{(0)}p^{n}$. We want to show $p^{(n+1)}=p^{(0)}p^{n+1}$.
+From theorem 2, we have
+\begin{align*}
+p^{(n+1)}&=(p_{1}^{(n+1)},p_{2}^{(n+1)},\dots p_{r}^{(n+1)})\\
+&= (p_{1}^{(n)}p_{1}^{(0)},p_{2}^{(n)}p_{2}^{(0)},\dots p_{r}^{(n)}p_{r}^{(0)})\\
+&= p^{n+1}p^{(0)}.
+\end{align*}
+\end{quote}
 
 \item Use matrix multiplication to solve Exercise 8.
+\begin{quote}
 \[
@@
 =
+\left(
+\begin{array}{c c}
+.5 & .5\\
+.5 & .5
+\end{array}
+\right)
 \]
+Therefore, the probability is .5.
+\end{quote}
 
@@
 she will buy a toy 4 weeks from now.
+\begin{quote}
+The probability that she will buy a toy can be found by raising the matrix to the fourth power.
+\[
+\left(
+\begin{array}{c c}
+.4 & .6\\
+.8 & .2
+\end{array}
+\right)^{4}
+=
+\left(
+\begin{array}{c c}
+.5824 & .4176\\
+.5568 & .4432
+\end{array}
+\right).
+\]
+Therefore, the probability is $0.4432$.
+\end{quote}
 
 \item Find the distribution of birth weights (Example 1) after one generation if the initial
-probability distribution is (.4, .3, .3).
+probability distribution is $(.4, .3, .3)$.
+\[
+(.4,.3,.3)\cdot
+\left(
+\begin{array}{c c c}
+.5 & .45 & .05\\
+.25 & .5 & .25\\
+.3 & .3 & .4
+\end{array}
+\right)
+= (.365,.42,.215).
+\]
 
@@
 $p^{(1)}=7(0.31,0.45,0.24)$. Can you find the distribution of birth weights of the mothers?
+\begin{quote}
+We need to solve the system
+\[
+(x,y,z)
+\left(
+\begin{array}{c c c}
+.5 & .45 & .05\\
+.25 & .5 & .25\\
+.3 & .3 & .4
+\end{array}
+\right)
+= 7(.31,.45,.24).
+\]
+Doing so yields
+\[
+(x,y,z)=(1.4,4.2,1.4).
+\]
+\end{quote}
 
@@
 $S_{3}$ in two steps if it starts in state $S_{1}$.
+\begin{quote}
+To move from $S_{1}$ to $S_{2}$ has a probability of $a$, and from $S_{2}$ to $S_{3}$ has a
+probability $(1-b)$, so the total probability is $a(1-b)$.
+\end{quote}
 
@@
 first column of $P^{3}$ and $P^{n}$?
+\begin{quote}
+\[
+P^{2}=
+\left(
+\begin{array}{c c c}
+0 & a^2 & (1-a)^2\\
+0 & b^2 & (1-b)^2\\
+0 & c^2 & (1-c)^2
+\end{array}
+\right).
+\]
+The first column of $P^{n}$ will always be 0.
+\end{quote}
 

### 2016-04-12 number_theory/hw/hw10.tex
number theory and modeling

\documentclass{hw}

\begin{document}
\makeheader{10}
\begin{enumerate}
\item Let $\phi$ denote Euler's Function.
\begin{enumerate}
\item Calculate $\phi(22)$.
\[
\phi(22) = \phi(11)\phi(2)=10\cdot1 = 10.
\]
\item Does there exist $l\in\ZZ$ such that $a^{7l}\equiv a\mod 22$ for all units $a\in\ZZ_{22}$? If
so, find $l$. Otherwise, explain why no such $l$ exists.
\begin{quote}
We know that we can find an $l$ if $7l\equiv1\mod\phi(22)$. Since $gcd(7,10)=1$, there is an inverse
of 7 in $\ZZ_{10}$, which is 3. Therefore, $a^{21}\equiv a\mod22$ for all units $a\in\ZZ_{22}$.
\end{quote}
\end{enumerate}

\item Suppose that $a\in\ZZ$ is such that $gcd(a,15)=1$. If $gcd(k,8)=1$, prove that
$a^{k^{2}}\equiv a\mod15$.
\begin{quote}
Suppose $gcd(k,8)=1$, and define the set $U_{8}$ to be the set of units in $\ZZ_{8}$. We want to find
$k^{-1}$ such that $k^{-1}k\equiv1\mod\phi(15)$, which allows us to say $a^{k^{-1}k}\equiv a\mod15$.
We know that $U_{8}=\{ 1,3,5,7 \}$. Since $gcd(k,8)=1$, $k\in U_{8}$. Suppose that $k=1$. Then
$k^{-1}=1$, so $a^{k^{2}}\equiv a^{1}\equiv a\mod15$. If $k=3$, $k^{-1}=3$, so $a^{k^{2}}\equiv
a^{9}\equiv a\mod15$. If $k=5$, then $k^{-1}=5$, and $a^{k^{2}}\equiv a^{25}\equiv a\mod15$. If $k=7$,
then $k^{-1}=7$, so $a^{k^{2}}\equiv a^{49}\equiv a\mod15$.
\end{quote}

\item Let $\Omega$ denote the prime counting map and $\lambda$ the Liouville function. Prove that
$\lambda(x)=\lambda(y)$ if and only if $\Omega(x)\equiv\Omega(y)\mod2$.
\begin{quote}
$\Rightarrow)$ Suppose $\lambda(x)=\lambda(y)$. We know
\[
\lambda(a)=
\begin{cases}
1,\ \ \qquad\Omega(a)\in\{2k|k\in\ZZ\}\\
-1,\qquad\Omega(a)\in\{2k+1|k\in\ZZ\}
\end{cases}.
\]
Therefore, the only way $\lambda(x)=\lambda(y)$ is if the parity of $\Omega(x)$ is the same
as the parity of $\Omega(y)$. Assume that $\Omega(x),\Omega(y)\in\{ 2k|k\in\ZZ\}$. Then
$\Omega(x)\equiv\Omega(y)\equiv0\mod2$. Assume $\Omega(x),\Omega(y)\in\{ 2k+1|k\in\ZZ\}$. Then
$\Omega(x)\equiv\Omega(y)\equiv1\mod2$. Therefore, if $\lambda(x)=\lambda(y)$,
$\Omega(x)\equiv\Omega(y)\mod2$.\\\\
$\Leftarrow)$ Suppose that $\Omega(x)\equiv\Omega(y)\mod2$. We know that $\lambda(a)$ is defined
to be $(-1)^{\Omega(a)}$, so suppose without loss of generality that
$\Omega(x),\Omega(y)\in\{ 2k|k\in\ZZ\}$. Then $\lambda(x)=(-1)^{2k}=1$, and $\lambda(y)=(-1)^{2k}=1$,
so $\lambda(x)=\lambda(y)$.
\end{quote}

\item Let $i=\sqrt{-1}\in\CC$ and define $\rho:\NN\to\CC$ by $\rho(x)=i^{\Omega(x)}$. Prove that
$\rho(x)=\rho(y)$ if and only if $\Omega(x)\equiv\Omega(y)\mod4$.
\begin{quote}
$\Rightarrow)$ Suppose $\rho(x)=\rho(y)$. We know that $i$ to a power is either $1,-1,i,$ or $-i$.
If $\rho(x)$ is to equal $\rho(y)$, then both must equal one of the four variations of $i$ to a
power. Suppose $\rho(x)=\rho(y)=1$. Then $i^{\Omega(x)}=i^{\Omega(y)}=1$, so $\Omega(x)$ and
$\Omega(y)$ must be a multiple of 4. Then $\Omega(x)\equiv\Omega(y)\equiv0\mod4$. If
$\rho(x)=\rho(y)=-1$, then $\Omega(x)$ and $\Omega(y)$ must be a multiple of 2 but not a multiple
of 4, since being a multiple of 4 would cause $\rho(x)=\rho(y)=1$. If $\Omega(x)$ and $\Omega(y)$
are multiples of 2 and not multiples of 4, then $\Omega(x)\equiv\Omega(y)\equiv2\mod4$. Suppose that
$\rho(x)=\rho(y)=i$. Then $i^{\Omega(x)}=i^{\Omega(y)}=i$. We know that $\Omega(x)$ and $\Omega(y)$
must be in the set $\{ 4k+1|k\in\ZZ\}$, or $\{ 4k+3|k\in\ZZ\}$. Suppose
$\Omega(x),\Omega(y)\in\{ 4k+1|k\in\ZZ\}$. Then $i^{\Omega(x)}=i^{\Omega(y)}=i^{4k+1}=i^{4k}i=i$.
In this case, we have $\Omega(x)\equiv\Omega(y)\equiv1\mod4$. Finally, suppose that
$\Omega(x),\Omega(y)\in\{ 4k+3|k\in\ZZ\}$. Then $i^{\Omega(x)}=i^{\Omega(y)}=i^{4k+3}=i^{4k}(-i)=-i$.
In this case, $\Omega(x)\equiv\Omega(y)\equiv3\mod4$. Therefore, if $\rho(x)=\rho(y)$,
$\Omega(x)\equiv\Omega(y)\mod4$.\\\\

$\Leftarrow)$ Suppose $\Omega(x)\equiv\Omega(y)\mod4$. Then $\Omega(x),\Omega(y)\in\{0,1,2,3\}$.
If $\Omega(y)=0$, then $\Omega(x)=0$, and $\rho(x)=i^{0}=1$ and $\rho(y)=i^{0}=1$, so $\rho(x)=\rho(y)$.
If $\Omega(x)=\Omega(y)=1$, then  $\rho(x)=i=\rho(y)$. If $\Omega(x)=\Omega(y)=2$, then
$\rho(x)=-1=\rho(y)$. If $\Omega(x)=\Omega(y)=3$, then $\rho(x)=-i=\rho(y)$.
\end{quote}

\item Suppose that $f:\NN\to\CC$ is a function such that $f(xy)=f(x)+f(y)$ for all $x,y\in\NN$. If
$f(p)=1$ for all primes $p$, prove that $f(x)=\Omega(x)$ for all $x\in\NN$.

\item For an ordered pair $(a,b)\in\NN\times\NN$ we define the \textit{Generalized Prime Counting Map}
by \[\overline{\Omega}(a,b)=\Omega(a)-\Omega(b).\] If $(a,b),(c,d)\in\NN\times\NN$ are such that
$ad=bc$, prove that $\overline{\Omega}(a,b)=\overline{\Omega}(c,d)$.

\item Let $\mu$ be the M\"{o}bius Function.
\begin{enumerate}
\item If $p$ and $q$ are distinct primes, prove that $\mu(pq)=1$.
\begin{quote}
Since $\mu$ is multiplicative, we have that $\mu(xy)=\mu(x)\mu(y)$ when $gcd(x,y)=1$. Since $p$ and
$q$ are distinct, we know that $\mu(pq)=\mu(p)\mu(q)$. By definition of the M\"{o}bius function, for
a prime $\rho$ we have
\[
\mu(\rho)=-\sum_{\substack{d\mid \rho\\d\neq\rho}}\mu(\rho)=-\mu(1)=-1,
\]
so $\mu(p)\mu(q)=(-1)(-1)=1$.
\end{quote}
\item If $p$ is prime, prove that $\mu(p^{2})=0$.
\begin{quote}
From $(a)$, we know that $\mu(\rho)=1$ for some prime $\rho$. If we consider $p^2$, we know that the
only divisors are $1,p,$ and $p^2$. By the definition of the M\"{o}bius function, we have
\[
\mu(p^2)=-\sum_{\substack{d\mid p^2\\d\neq p^2}}\mu(d)=-(\mu(1) + \mu(p))=-1 +1 = 0.
\]
\end{quote}
\item If $p$ is prime and $n$ is a positive integer, find a formula for the value of $\mu(p^{n})$.
Prove your answer.
\begin{quote}
\[
\mu(p^n)=
\begin{dcases*}
-1,\qquad n=1\\
0,\ \ \qquad n>1
\end{dcases*}
\]
Consider the value of $\mu(p^{n})$. We have already shown that $\mu(p)=-1$ for any prime $p$ and
that $\mu(p^2)=0$ for any prime. We can suppose that $\mu(p^{n})=0$ for $n\geq2$, and
$\mu(p^{n})=-1$ for $n=1$. We want to show that $\mu(p^{n+1})=0$ for $n\geq2$. Then
\begin{align*}
\mu(p^{n+1})&=-\sum_{\substack{d\mid p^{n+1}\\ d\neq p^{n+1}}}\mu(d)\\
&= -(\mu(1)+\mu(p)+\mu(p^{2})+\mu(p^{3})+\dots+\mu(p^{n}))\\
&= -(1+(-1)+0+0+\dots+0)\text{ by the inductive hypothesis}\\
&= -(0+0+0+\dots+0)\\
&= -0\\
&= 0.
\end{align*}
%If $n$ is odd, then
%we have $p^{n}=p^{2k+1}=p^{2k}p$ for some $k\in\ZZ$.
\end{quote}
\end{enumerate}

\item If $\phi$ denotes Euler's Function and $s>2$, prove that
\[
{\zeta(s-1)\over\zeta(s)}=\sum_{n=1}^{\infty}{\phi(n)\over n^{s}}.
\]
\end{enumerate}
\end{document}

### 2016-04-12 pde_two/threshold.tex
number theory and modeling

@@
 with boundary conditions that arose from the steady state of $(1)$. We will leave the first boundary
-condition alone and change the second boundary condition so that
+condition alone and change the second boundary condition so that $2u'(0)^{2}=a^{4}$. This will allow
+us to factor the resulting term:
+\begin{align*}
+u'(x)^2&={1\over2}(a^{4}-u^{4})\\
+{\du\over\dx}&= \sqrt{{1\over2}(a^{2}+u^{2})(a^{2}-u^{2})}\\
+{\dx\over\du}&= {1\over\sqrt{{1\over2}(a^{2}+u^{2})(a^{2}-u^{2})}}.
+\end{align*}
+Integrating both sides and substituting $s$ for $u$ gives us
+\[x=\int_{0}^{u}{1\over\sqrt{{1\over2}(a^{2}+s^{2})(a^{2}-s^{2})}}\mathop{ds}.\]
+We can now make another substitution for $s$ in terms of $t$. Letting
+\[
+s=\sqrt{{a^2t^2\over2(1-t^2/2)}}
+\]
+yields
+\[
+a^2+s^2={2a^2\over2(1-t^2/2)}\qquad\qquad\text{and}\qquad\qquad
+a^2-s^2={a^2(1-t^2)\over(1-t^2/2)}.
+\]
+We also have
+\[
+\mathop{ds}={a(t-1)(t-2)\over2\sqrt{2}(1-t^2/2)^{3/2}}\dt,
+\]
+which gives us the new integral
+\[
+ax=\int_{0}^{\sqrt{\frac{\displaystyle2u}{\displaystyle a^2+u^2}}}{1\over\sqrt{(1-t^2)(1-t^2)}}.
+\]
+We may notice that this integral is one of the ways to define $sn(x,k)$. Rewriting it in terms
+of our Jacobi elliptic function gives us
+\[
+x=\int_{0}^{sn(x,k)}{1\over\sqrt{(1-t^2)(1-t^2/2)}}.
+\]
 

### 2016-04-13 math_modeling/test_02.tex
another one on number

% how to compute the fundamental matrix?
% how to determine communication classes?
% what is the point of computing the entries?
% find ultimate vector
% generalized eigenvalues
\documentclass{hw}
\usepackage{tikz}
\usetikzlibrary{automata,positioning}

\begin{document}
\section*{Bayes Theorem}
\begin{enumerate}
% See the pop vs. soda question
\item Suppose there are 2 cookie jars. Jar A has 10 chocolate chip and 30 plain cookies while jar
B has 20 of each type. We pick a cookie and it turns out to be plain. What is the probability that
we picked out of the jar A?
\begin{quote}
We are trying to find the probability that you chose a cookie from jar A given that you picked a
plain cookie.
\[
P(\text{A}|\text{plain}).
\]
We can find this value by
\[
P(\text{A}|\text{plain})=
{P(\text{A and plain})\over P(\text{plain})}=
{(1/2)(3/4)\over(1/2)(3/4)+(1/2)(20/40)}.
\]
\end{quote}
\item The blue M\&M was introduced in 1995. Before 1995, the ratios of colors was
\begin{gather*}
30\%\text{ brown}\qquad 20\%\text{ yellow}\qquad20\%\text{ red}\\
10\%\text{ green}\qquad 10\%\text{ orange}\qquad10\%\text{ tan}
\end{gather*}
After 1995, the ratios were
\begin{gather*}
24\%\text{ blue}\qquad 20\%\text{ green}\qquad16\%\text{ orange}\\
14\%\text{ yellow}\qquad 13\%\text{ red}\qquad13\%\text{ brown}
\end{gather*}
If we pick an M\&M that turns out to be yellow, what is the probability that the bag came before
1995?
\begin{quote}
We want to find $P(1994|\text{yellow})$. By Bayes theorem, we can find this value by
\[
P(1994|\text{yellow})=
{P(1994\text{ and yellow})\over P(\text{yellow})}=
{(.5)(.2)\over (.5)(.2)+(.5)(.25)}
\]
\end{quote}
\end{enumerate}

\section*{Communication Classes}
\begin{enumerate}
\item What are the communication classes of
\begin{center}
\begin{tikzpicture}[shorten >=1pt,node distance=2cm,on grid,auto]
    \node[state] (one) {1};
    \node[state] (two) [below=of one] {2};
    \node[state] (three) [right=of one] {3};
    \node[state] (four) [right=of three] {4};
    \node[state] (five) [right=of four] {5};
    \node[state] (six) [below=of five] {6};
    \node[state] (seven) [below left=of six] {7};
    \node[state] (eight) [below=of three] {8};

    \path[->]
    (one) edge [bend right] node {} (two)
          edge [bend right] node {} (three)
    (two) edge [bend right] node {} (one)
          edge [loop below] node {} ()
    (three) edge [bend right] node {} (four)
            edge [bend right] node {} (eight)
    (four) edge [bend right] node {} (three)
           edge [bend right] node {} (five)
    (five) edge [bend right] node {} (six)
    (six) edge [bend right] node {} (seven)
    (seven) edge [bend right] node {} (six)
            edge [bend left] node {} (eight)
    (eight) edge [bend left] node {} (six);
\end{tikzpicture}
\end{center}
\begin{quote}
We can find the communication classes by determining if we can move from state $i$ to $j$ and from
state $j$ to $i$.
\end{quote}

\item Consider the matrix
\[
\left(
\begin{array}{c c}
.6 & .4\\
.3 & .7
\end{array}
\right)
\]
\begin{enumerate}
\item How many communication classes are there?
\begin{quote}
There is 1 communication class, since every state is accessible from ever other state.
\end{quote}
\item List the communication class(es).
\begin{quote}
Since there are two states, with one communication class, the communication class is $\{A,B\}$.
\end{quote}
\item Find $\mathbf{\pi}$.
\begin{quote}
To find the $\mathbf{\pi}$ vector, we can solve the system
\[
\begin{cases}
.6\pi_{1}+.3\pi_{2}=\pi_{1}\\
\pi_{1}+\pi_{2}=1
\end{cases}
\]
\end{quote}
\end{enumerate}
\item Consider the model of moving from the city to the suburbs with an initial state.
\[
\begin{array}{c | c c}
& \text{Sub} & \text{City}\\
\hline
\text{Sub} & .95 & .05\\
\text{City} & .03 & .97
\end{array}
\]
\begin{enumerate}
\item What is the probability that a person in 3 years transfers from city to suburb?
\item What number of people in the city will be in the suburbs next year?
\item What is the ultimate number of people in the suburbs? The cities?
\end{enumerate}
\end{enumerate}
\end{document}

### 2016-04-13 number_theory/hw/hw10.tex
another one on number

@@
 $ad=bc$, prove that $\overline{\Omega}(a,b)=\overline{\Omega}(c,d)$.
+\begin{quote}
+Suppose $ad=bc$. We can write out the prime factorizations of $a,b,c$, and $d$ as
+\begin{align*}
+a &= p_{1}^{\alpha_{1}}p_{2}^{\alpha_{2}}\dots p_{n}^{\alpha_{n}}\\
+b &= p_{1}^{\beta_{1}}p_{2}^{\beta_{2}}\dots p_{n}^{\beta_{n}}\\
+c &= p_{1}^{\gamma_{1}}p_{2}^{\gamma_{2}}\dots p_{n}^{\gamma_{n}}\\
+d &= p_{1}^{\delta_{1}}p_{2}^{\delta_{2}}\dots p_{n}^{\delta_{n}}.
+\end{align*}
+Then
+\begin{align*}
+ad&=p_{1}^{\alpha_{1}+\delta_{1}}p_{2}^{\alpha_{2}+\delta_{2}}\dots p_{n}^{\alpha_{n}+\delta_{n}}\\
+bc&=p_{1}^{\beta_{1}+\gamma_{1}}p_{2}^{\beta_{2}+\gamma_{2}}\dots p_{n}^{\beta_{n}+\gamma_{n}}.
+\end{align*}
+We can consider $\overline{\Omega}(ad,bc)=\Omega(ad)-\Omega(bc)=0$. We know
+$\Omega(ad)-\Omega(bc)=0$ since $ad=bc$ implies they will have the same prime factorization up to
+permutation, so the powers of the prime factors of $ad$ will be equivalent to the powers of the prime
+factors of $bc$. We now have
+\begin{align*}
+\Omega(ad)-\Omega(bc)&=(\alpha_{1}+\delta_{1})+(\alpha_{2}+\delta_{2})+\dots+(\alpha_{n}+\delta_{n})
+- (\beta_{1}+\gamma_{1})-(\beta_{2}+\gamma_{2})-\dots-(\beta_{n}+\gamma_{n})\\
+&= (\alpha_{1}+\alpha_{2}+\dots+\alpha_{n})-(\beta_{1}+\beta_{2}+\dots+\beta_{n})+
+(\delta_{1}+\delta_{2}+\dots+\delta_{n})-(\gamma_{1}+\gamma_{2}+\dots+\gamma_{n})\\
+&=\Omega(a)-\Omega(b)-\Omega(c)+\Omega(d)\\
+&= 0.
+\end{align*}
+Then $\Omega(a)-\Omega(b)=\Omega(c)-\Omega(d)$, which by definition is
+$\overline{\Omega}(a,b)=\overline{\Omega}(c,d)$.
+\end{quote}
 

### 2016-04-13 pde_two/threshold.tex
update paper

@@
 \usepackage{amsmath,amssymb}
+\usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
 \]
+%% TODO -- Expand on this information
+Finally, we have a solution $u(x)$
+\[
+u(x)={a\ sn(ax,\sqrt{1/2})\over\sqrt{2-sn^{2}(ax,\sqrt{1/2})}}.
+\]
+Recall that the first positive zero of $sn(x,k)$ is given by
+\[
+K=\int_{0}^{1}{1\over\sqrt{(1-t^2)(1-{t^2\over2})}}\dt.
+\]
+Then $u(x)=u_{s}(x)$ provided that $a=2K/\pi$. We can see the results via these Mathematica plots:
+\begin{figure}[h]
+\centering
+\begin{subfigure}{0.45\textwidth}
+\includegraphics[scale=0.5]{u_of_1_x}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\textwidth}
+\includegraphics[scale=0.5]{u_of_2_x}
+\end{subfigure}
+\caption{The graphs of $u_{s}(x)$ for $a=1$ and $a=2$}
+\end{figure}
+
+\noindent It appears that, for $a=1$, $u_{s}(x)$ behaves similar to $\sin{x}$.
+
+\begin{figure}[h]
+\centering
+\begin{subfigure}{0.45\textwidth}
+\includegraphics[scale=0.5]{u_of_1_x}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.45\textwidth}
+\includegraphics[scale=0.5]{sinx}
+\end{subfigure}
+\caption{The graphs of $u_{s}(x)$ for $a=1$ and $\sin{x}$}
+\end{figure}
+
+\subsection{Average of the Steady State Solution}
+
+If we can find the average temperature of the steady state, then we know that our function $u(x)$
+has a global solution (otherwise there would be points in the range of $u$ for which $u$ blows up).
+We will want to show that
+\[
+{1\over\pi}\int_{0}^{\pi}u(x)\dx={\sqrt{2}\over2}.
+\]
+To begin, let
+\[
+y=\int_{0}^{x}u(t)\dt
+\]
+and $v(y)=u^{2}(x)$ where $x=x(y)=y^{-1}(x)$.
 

### 2016-04-13 writing_in_ethics/paper_four.tex
update paper

@@
 One engineer, Roger Boisjoly, testified against NASA and the the engineering company Morton Thiokol during the trial, despite the recommendations given Morton Thiokol's lawyers.
-Although Boisjoly broke some of the guidelines of the ACM code of Ethics, his actions were not unethical.
+Although Boisjoly broke some of the guidelines of the ACM code of Ethics, his actions were not ultimately unethical since he took steps to prevent the launch and remained honest throughout the resulting court trials.
 
@@
 Omission of the truth, especially when dealing with loss of human life, would have certainly been a conflict of interest.
-One can argue that by not following the advice of Morton Thiokol's lawyers was in violation of the Code of Ethics, since he did not keep engineering work confidential.
+One can argue that by not following the advice of Morton Thiokol's lawyers was a violation of the Code of Ethics, since he did not keep engineering work confidential.
 The Code, however, states that the obligation to keep work confidential can be changed when the law or other sections of the Code require that the information be divulged.

### 2016-04-14 algorithms/projects/proj3/Project3.java
yolo

@@ -1,0 +1,17 @@
+public class Project3 {
+
+    public static void main(String[] args){
+
+    }
+
+    public static void worstFit(){
+
+    }
+
+    public static void bestFit(){
+
+    }
+
+    
+
+}

### 2016-04-16 number_theory/hw/hw10.tex
yolo
oops
presentation

@@
 $f(p)=1$ for all primes $p$, prove that $f(x)=\Omega(x)$ for all $x\in\NN$.
+\begin{quote}
+We know that we can express $x$ and $y$ as
+\begin{gather*}
+x=p_{1}^{\alpha_{1}}p_{2}^{\alpha_{2}}+\dots+p_{n}^{\alpha_{n}}\\
+y=p_{1}^{\beta_{1}}p_{2}^{\beta_{2}}+\dots+p_{n}^{\beta_{n}}.
+\end{gather*}
+We can express $f(xy)-f(y)$ as $f(x)+f(y)-f(y)=f(x)$, and we also know
+\begin{align*}
+f(xy)&=f(p_{1}^{\alpha_{1}+\beta_{1}}p_{2}^{\alpha_{2}+\beta_{2}}+\dots+p_{n}^{\alpha_{n}+\beta_{n}})\\
+%
+&= f(p_{1}^{\alpha_{1}}) + f(p_{1}^{\beta_{1}}) + f(p_{2}^{\alpha_{2}}) + f(p_{2}^{\beta_{2}})+
+\dots + f(p_{n}^{\alpha_{n}}) + f(p_{n}^{\beta_{n}})\\
+%
+f(xy)-f(y)&=f(x)\\
+&=f(p_{1}^{\alpha_{1}}p_{2}^{\alpha_{2}}\dots p_{n}^{\alpha_{n}})\\
+%
+&=f(p_{1}^{\alpha_{1}})+f(p_{2}^{\alpha_{2}})+\dots+f(p_{n}^{\alpha_{n}})\\
+%
+&=\sum_{k=1}^{\alpha_{1}}f(p_{1}) + \sum_{k=1}^{\alpha_{2}}f(p_{2}) + \dots +
+\sum_{k=1}^{\alpha_{n}}f(p_{n})\\
+%
+&= \sum_{k=1}^{\alpha_{1}}1 + \sum_{k=1}^{\alpha_{2}}1\dots + \sum_{k=1}^{\alpha_{n}}1\\
+&= \alpha_{1} + \alpha_{2} + \dots + \alpha_{n}\\
+&=\Omega(x)
+\end{align*}
+\end{quote}
 
@@
 
+\newpage
 \item Let $\mu$ be the M\"{o}bius Function.
@@
 \begin{quote}
-From $(a)$, we know that $\mu(\rho)=1$ for some prime $\rho$. If we consider $p^2$, we know that the
+From $(a)$, we know that $\mu(\rho)=-1$ for some prime $\rho$. If we consider $p^2$, we know that the
 only divisors are $1,p,$ and $p^2$. By the definition of the M\"{o}bius function, we have
@@
 \end{align*}
-%If $n$ is odd, then
-%we have $p^{n}=p^{2k+1}=p^{2k}p$ for some $k\in\ZZ$.
 \end{quote}
@@
 
-\item If $\phi$ denotes Euler's Function and $s>2$, prove that
-\[
-{\zeta(s-1)\over\zeta(s)}=\sum_{n=1}^{\infty}{\phi(n)\over n^{s}}.
-\]
+% \item If $\phi$ denotes Euler's Function and $s>2$, prove that
+% \[
+% {\zeta(s-1)\over\zeta(s)}=\sum_{n=1}^{\infty}{\phi(n)\over n^{s}}.
+% \]
 \end{enumerate}

### 2016-04-16 math_modeling/final_project/graphs.R

@@ -1,0 +1,2 @@
+mc <- c(890,1062,1172,1440,2080,2897,7328,8859,10492,15829,37324)
+plot(mc,type="s")

### 2016-04-19 algorithms/projects/proj3/Project3.java
project 3 in algo

@@ -1,17 +1,70 @@
+/**
+Steven Rosendahl and Talia Hicks
+
+You can use this program by passing in command line arguments which will test an array of that size.
+If you don't pass in command line arguments, it defaults to an array of size 100.
+
+With command line args:
+java Project3 50
+
+Without command line args:
+java Project3
+*/
+
+import static java.lang.Math.*;
+import static java.util.Arrays.*;
+import static java.lang.Integer.*;
+
 public class Project3 {
 
     public static void main(String[] args){
-
+        int arraySize = (args.length > 0) ? parseInt(args[0]) : 100;
+        double[] sizes = new double[arraySize];
+        for(int i = 0; i < arraySize; i++){
+            sizes[i] = (random());
+        }
+        firstFit(sizes);
+        sortedFirstFit(sizes, false);
+        sortedFirstFit(sizes, true);
     }
 
-    public static void worstFit(){
-
+    public static void firstFit(double[] sizes){
+        int n = sizes.length;
+        double[] used = new double[sizes.length];
+        for(int i = 0; i < n; i++)
+            used[i] = 0.0;
+        int binsUsed = 0;
+        for(int i = 0; i < n; i++){
+            int binLocation = 0;
+            while(used[binLocation] + sizes[i] > 1)
+                binLocation++;
+            used[binLocation] += sizes[i];
+            binsUsed = Math.max(binsUsed, binLocation);
+        }
+        System.out.printf("I used %d bins on an array of size %d\n", binsUsed, n);
     }
 
-    public static void bestFit(){
-
+    public static void sortedFirstFit(double[] sizes, boolean ascending){
+        sort(sizes);
+        if(!ascending)
+            reverseArray(sizes);
+        firstFit(sizes);
     }
 
-    
+    private static void reverseArray(double[] array){
+        for(int i = 0; i < array.length / 2; i++){
+            double temp = array[i];
+            array[i] = array[array.length - i - 1];
+            array[array.length - i - 1] = temp;
+        }
+    }
+
+    private static void printArray(double[] array){
+        System.out.print("[");
+        for(double i : array){
+            System.out.print(" " + i + " ");
+        }
+        System.out.println("]");
+    }
 
 }

### 2016-04-22 pde_two/threshold.tex
yolo
almost done :D

@@
 \begin{gather}
-u_{t}=u_{xx}+u^{3},\ 0<x<\pi,\ t>0\\
+u_{t}=u_{xx}+u^{3},\ 0<x<\pi,\ t>0\label{nlinh}\\
 u(0,t)=u(\pi,t)=0,\ t>0\\
@@
 \begin{gather}
-u_{xx}+u^{3}=0\\
+u_{xx}+u^{3}=0\label{ssone}\\
 u(0)=u(\pi)=0.
@@
 \[
-y=\int_{0}^{x}u(t)\dt
+y=\int_{0}^{x}u(t)\dt \implies {\dy\over\dx} = u(x)
 \]
-and $v(y)=u^{2}(x)$ where $x=x(y)=y^{-1}(x)$.
+and $v(y)=u^{2}(x)$ where $x=x(y)=y^{-1}(x)$. Finally, we will let
+\[
+R=\int_{0}^{\pi}u(x)\dx.
+\]
+We now can set up an ordinary differential equation in terms of $v$.
+\begin{align*}
+v'(y)&=2u(x)u'(x){\dx\over\dy}\\
+&= 2{\dy\over\dx}u'(x){\dx\over\dy}\\
+&= 2u'(x)\\
+v''(y)&= 2u''(x){\dx\over\dy}\\
+&= {2u'(x)\over u(x)}.
+\end{align*}
+This implies that $v$ satisfies the following linear ordinary differential equation:
+\begin{gather}\label{vode}
+v''(y)+2v(y)=0
+\end{gather}
+on the interval $0<y<R$ where $v(0)=v(R)=0$. The differential equation in \eqref{vode} has the
+characteristic equation
+\[
+r^{2}+2=0,
+\]
+implying that the basis of our solution is the in the form $\{\sin{\sqrt{2}y},\cos{\sqrt{2}y}\}$.
+Applying our initial conditions leaves us with
+\[
+v(y)=c\sin{\sqrt{2}y},
+\]
+which yields $\sqrt{2}R=\pi$. Then
+\[
+{1\over\pi}R={1\over\pi}\int_{0}^{\pi}u(x)\dx={\sqrt{2}\over2},
+\]
+which was the desired result.
 
+\section{Non-Linear Analysis}
+
+We have shown that $u_{s}$ is a unique solution for for the system described in \eqref{ssone}. We now
+want to relate $u_{s}$ to \eqref{nlinh}. We make the observation that $u_{s}(x)$ is an upper bound
+of the solution to \eqref{nlinh}, called $u(x,t)$ when the initial condition $u_{0}(x)<u_{s}(x)$; by
+comparison we have that the solution to \eqref{nlinh} exists globally, and
+$0\leq u(x,t)\leq u_{s}(x)$. This solution, however is only stable when the initial condition
+satisfies
+\begin{equation}\label{bound}
+\int_{0}^{\pi}u_{0}(x)\sin{x}\dx\leq2.
+\end{equation}
+If $u_{0}(x)$ is too large, the solution blows up as $t\to\infty$. We will now try to show that the
+solution exits when \eqref{bound} holds.
+
+\subsection{Analysis of the Initial Conditions}
+We begin by letting
+\[
+F(t)=\int_{0}^{\pi}u(x,t)\sin{x}\dx.
+\]
+Differentiating with respect to $t$ yields
+\begin{align*}
+F'(t)&=\int_{0}^{\pi}u_{t}\sin{x}\dx\\
+&=\int_{0}^{\pi}(u_{xx}+u^{3})\sin{x}\dx\\
+&=
+\end{align*}
 \end{document}

### 2016-04-22 algorithms/projects/proj3/Project3.java

@@ -17,15 +17,29 @@ import static java.lang.Integer.*;
 
 public class Project3 {
 
+    private static double usAvg, sAvg;
+    private static final int TRIALS = 800;
+    // Change this to false if you want to see it sort the other way
+    private static final boolean ASCENDING = false;
+
     public static void main(String[] args){
-        int arraySize = (args.length > 0) ? parseInt(args[0]) : 100;
-        double[] sizes = new double[arraySize];
-        for(int i = 0; i < arraySize; i++){
-            sizes[i] = (random());
+        usAvg = sAvg = 0.0;
+        int cSortedBetter = 0;
+        int arraySize = (args.length > 0) ? (parseInt(args[0]) > 0 ? parseInt(args[0]) : 100) : 100;
+        for(int k = 0; k < TRIALS; k++){
+            double aa = 0;
+            double bb = 0;
+            double[] sizes = new double[arraySize];
+            for(int i = 0; i < arraySize; i++)
+                sizes[i] = (random() * random());
+            aa = bestFit(sizes);
+            usAvg += aa;
+            bb = sortedBestFit(sizes,ASCENDING);
+            sAvg += bb;
+            if(aa > bb)
+                cSortedBetter++;
         }
-        firstFit(sizes);
-        sortedFirstFit(sizes, false);
-        sortedFirstFit(sizes, true);
+        System.out.printf("On average, unsorted best fit used %.2f bins, sorted average used %.2f bins,\nand sorted was better than unsorted %d times out of %d runs\n", usAvg / TRIALS, sAvg / TRIALS, cSortedBetter, TRIALS);
     }
 
     public static void firstFit(double[] sizes){
@@ -44,11 +58,50 @@ public class Project3 {
         System.out.printf("I used %d bins on an array of size %d\n", binsUsed, n);
     }
 
+    public static int bestFit(double[] sizes){
+        int n = sizes.length;
+        double[] used = new double[sizes.length];
+        for(int i = 0; i < n; i++)
+            used[i] = 0.0;
+        int binsUsed = 0;
+        for(int i = 0; i < n; i++){
+            //loop through the array of sizes
+            double leftoverSpace = 1;
+            int bestBin = 0;
+            for(int j = 0; j < n; j++){
+                // determine the leftoever space if you put
+                // that element into the bin. If you find a bin with a smaller
+                // leftover space, use that bin instead
+                if(used[j] + sizes[i] < 1){
+                    double oldLS = leftoverSpace;
[…]

### 2016-08-23 number_theory/defs.tex
starting new semester

@@
 \newcommand{\prm}{p}
+\newcommand{\set}[1]{\mathcal{#1}}
+\newcommand{\setA}{\mathcal{A}}
+\newcommand{\setB}{\mathcal{B}}
+\newcommand{\setU}{\set{U}}
 
@@
 
+\section{Definitions}
+
+\subsection{Sets and Logic}
 \begin{enumerate}
-\Define{Prime Number} A positive $n\in\ZZ$ such that $n>1$ is prime if its only
-divisors are 1 and $n$.
-
-\Define{Fundamental Theorem of Arithmetic} Suppose $n\in\ZZ > 1$. Then
-\begin{enumerate}
-\item $\exists\alpha_{1},\alpha_{2},\dots,\alpha_{n}\in\NN$ and primes $\prm_{1},
-\prm_{2},\dots,\prm_{n}$ such that $n=\prm_{1}^{\alpha_{1}}\prm_{2}^{\alpha_{2}}
-\dots\prm_{n}^{\alpha_{n}}$, or $n=\prod_{i=1}^{n}\prm_{i}^{\alpha_{i}}$.
-\item The expression $\prod_{i=1}^{n}\prm_{i}^{\alpha_{i}}$ is unique up to
-permutation of the primes.
-
-\Define{Lemma} Suppose $\prm$ is prime and $a,b\in\ZZ$.
-\begin{enumerate}
-\item If $\prm\nmid a$, then $\prm$ and $a$ are relatively prime.
-\item If $\prm\mid ab$, then $\prm\mid a$ xor $\prm\mid b$.
+\Define{Logical Operator} A logical operator is a symbol that acts on a logical statement. The operators act as follows
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline
+& & Negation & Disjunction & Conjunction & Implication\\
+\hline
+P & Q & $\neg$ P & P $\wedge$ Q & P $\vee$ Q & P $\implies$ Q \\
+\hline
+1 & 1 & 0 & 1 & 1 & 1\\
+1 & 0 & 0 & 0 & 1 & 0\\
+0 & 1 & 1 & 0 & 1 & 1\\
+0 & 0 & 1 & 0 & 0 & 1\\
+\hline
+\end{tabular}
+\end{center}
+A statement that is always true (i.e. P $\vee\ \neg$P) is called a Tautology. A statement that is always false (i.e. P $\wedge\ \neg$P) is called a contradiction.\\
+An implication P $\implies$ Q is made of two parts: the hypothesis (P) and the conclusion (Q).\\
+Additionally, statements can be quantified using the two quantifiers, the universal quantifier, for all ($\forall$) and the existential quantifier, there exists ($\exists$).
+\Define{Set} A set is a collection of objects. The objects in a set are called elements.
+\Define{Empty Set} The empty set, symbolized by $\emptyset$ is the set with no elements.
+\Define{Subset} A set $\setA$ is a subset of a set $\setB$ (noted $\setA\subset\setB$) if for all $x\in\setA$, $x\in\setB$.
+\Define{Power Set} The power set of a set $\setA$ (noted $\set{P}(\setA)$) is the set of all subsets of $\setA$.
+\Define{The Universal Set} The universal set, $\setU$ is the set of which all sets are subsets.
+\Define{Intersection} For sets $\setA$ and $\setB$, the intersection ($\setA\cap\setB$) is the set of elements in both $\setA$ and $\setB$.
+\[\setA\cap\setB = \{x|x\in\setA\wedge x\in\setB\}.\]
+To say that $\setA$ and $\setB$ have a trivial intersection means that $\setA\cap\setB=\emptyset$. This is equivalent to saying $\setA$ and $\setB$ are disjoint.
+\Define{Union} For sets $\setA$ and $\setB$, the union ($\setA\cup\setB$) is the set of elements in either $\setA$ or $\setB$.
+\[\setA\cup\setB = \{x|x\in\setA\vee x\in\setB\}.\]
+\Define{Set Difference} The set difference of a set $\setA$ and a set $\setB$ ($\setA-\setB$) is the set of all elements in $\setA$ that are not in $\setB$.
+\[\setA-\setB=\{x|x\in\setA\wedge x\notin\setB\}.\]
+The complement of a set $\setA$ in regards to $\setU$ is $\setU-\setA$.
+\Define{Cartesian Product} For sets $\setA$ and $\setB$, the cartesian product ($\setA\times\setB$) is defined to be the set of ordered pairs $(a,b)$ such that $a\in\setA$ and $b\in\setB$.
+\[\setA\times\setB=\{(a,b)|a\in\setA\wedge b\in\setB\}.\]
+\Define{Axiom} An axiom is a statement whose truth value is accepted without proof.
+\Define{Theorem} A theorem is a mathematical statement whose truth value can be verified through proof.
+\Define{Lemma} A Lemma is a mathematical result that is used to prove other results.
+\Define{Corollary} A corollary is a mathematical result that follows from another result.
 \end{enumerate}
 
-\Define{Corollary} If $\prm$ is prime and divides $a_{1},a_{2},\dots,a_{n}$, then
-$p|a_{i}$ for some $i\in\{ 1,2,\dots,n \}$.
-
-\Define{Corollary} If $m$ is not a perfect square, then $\sqrt{m}\notin\QQ$.
+\newpage
+\subsection{Number Theory}
+\begin{enumerate}
+\Define{Division} To say that $a$ divides $b$ ($a\mid b$) implies that $\exists x\in\ZZ$ such that $b=ax$.
+\Define{Relation} A relation $\mathcal{R}$ from $\setA$ to $\setB$ is a subset of $\setA\times\setB$.
+$\setA$ is related to $\setB$ ($a\mathcal{R}b$) if $(a,b)\in\mathcal{R}$ for all $a\in\setA$ and
+$b\in\setB$. The domain of $\mathcal{R}$ is the set $\{x\ |\ (x,y)\in\mathcal{R}\}$. The range of
+$\mathcal{R}$ is the set $\{y\ |\ (x,y)\in\mathcal{R}\}$. If $\mathcal{R}$ has an inverse
+$\mathcal{R}^{-1}$, then it is $\{(y,x)\ |\ (x,y)\in\mathcal{R}\}$. A relation between two elements
+$a\in\setA$ and $b\in\setB$ is denoted by $a\sim b$.
+\Define{Reflexive} A relation $\mathcal{R}$ is reflexive if $x\mathcal{R}x$ for all $x\in\setA$.
+\Define{Symmetric} A relation $\mathcal{R}$ is symmetric if $x\mathcal{R}y$ and $y\mathcal{R}x$ for all
+$x\in\setA$ and $y\in\setB$.
+\Define{Transitive} A relation $\mathcal{R}$ is transitive if $x\mathcal{R}y$ and $y\mathcal{R}z$ implies
+$x\mathcal{R}z$ for all $x\in\setA$, $y\in\setB$, and $z\in\set{C}$.
+\Define{Equivalence Relation} An equivalence relation is a relation that is reflexive, symmetric, and
+transitive.
+\Define{Equivalence Class} For a non-empty set $\setA$ containing elements $a$ and $b$, the equivalence
+class of $a$, noted $[a]$ is the set $\{b\ |\ b\sim a\}$.
+\Define{Partition} A partition of a non-empty set $\setA$ is the set of subsets where
+\begin{enumerate}
+\item The union of all sets in the partition of $\setA$ is $\setA$
+\item The intersection of any two different sets in the partition of $\setA$ is not equivalent to
+$\emptyset$.
+\end{enumerate}
+\Define{Function} A function $f$ from $\setA\to\setB$ is a relation from $\setA$ to $\setB$ that satisfies
+\begin{enumerate}
+\item $(a,b)\wedge(a,c)\in f\implies b=c$.
+\item $\forall a\in\setA, \exists b\in\setB$ such that $(a,b)\in f$.
+\end{enumerate}
+\Define{Image} The image of a function $f:\setA\to\setB$ is $\{f(a)\ |\ a\in\setA\}$.
+\Define{Inverse Image} For a function $f:\setA\to\setB$ and sets $\setB$ and $\set{D}$ where
+$\set{D}\subset\setB$, the inverse image of $\set{D}$, $f^{-1}(\set{D})$, is defined to be
+$\{a\in\setA\ |\ f(a)\in\set{D}\}$.
+\Define{Injection} A function $f$ is injective if $\forall (a,b)\in\set{X},\ f(a)=f(b)\implies a=b$.
+\Define{Surjection} A function $f$ is surjective if $\forall y\in\set{Y},\ \exists x\in\set{X}$ such that
+$f(x)=y$.
+\Define{Bijection} A function $f$ is a bijection if it is an injection and a surjection.
 \end{enumerate}
 
-\Define{Exact Division} If $\prm^{\alpha}\mid n$ and $\prm^{\alpha+1}\nmid n$, then
-$\prm^{\alpha}$ exactly divides $n$, noted as $\prm^{\alpha}||n$.
-
-\Define{Theorem} Suppose $a_{1},a_{2},\dots,a_{k}$ are pairwise relatively prime
-and are greater than 0. If $a_{1}a_{2}\dots a_{k} = x^{m}$ for some $x\in\NN$ and
-$m\geq2$, then $\exists x_{i}\in\NN$ such that $a_{i} = x_{i}^{m}$ for all $i$.
-
-\Define{Theorem} There are infinitely many primes.
-
-\Define{Prime Number Theorem} Let $\pi(x)$ denote the number of primes such that
-$\prm_{i}\leq x$. We have that
+\newpage
+\section{Theorems and Important Ideas}
 \begin{enumerate}
-\item $\lim_{x\to\infty}\pi(x)=\infty$
-\item $\pi(x) < x$ for $x\geq2$.
-\end{enumerate}
-We can also define the logarithmic integral as
-$\mathcal{L}_{i}(x)=\int_{2}^{x}{1\over\log{t}}\mathop{dt}$. Then
+\Define{Division Algorithm} Suppose $a,b\in\ZZ$ and assume $b>0$. Then $\exists q,r\in\ZZ$ such that
+$a=qb+r$ and $0\leq r\leq b$. Moreover, $q$ and $r$ are the only integers satisfying this property.
+The integer $q$ is commonly called the quotient and $r$ is commonly called the remainder.
+\begin{quote}
+\textit{\textbf{Proof}}: Let $\set{S}=\{a-nb\ |\ n\in\ZZ\}$. We know that
 \[
-\lim_{x\to\infty}{\mathcal{L}_{i}\over\pi(x)}=
-\lim_{x\to\infty}{{x\over\log{x}}\over\pi(x)} = 1
+\lim_{n\to-\infty}(a-nb)\to+\infty,
 \]
+so $\set{S}$ must contain at least one positive integer. Let $r$ be the smallest element of $\set{S}$
+that is greater than zero, called the remainder. Then there is a $q\in\ZZ$ such that $r=a-qb\geq0$, or
+$a=qb+r$ for $r\geq0$. We have shown that this $r$ exists.\\
+Assume $r\geq b$. Then $r-b\geq0$, and we know $r=a-qb$, so $a-qb-b\geq 0$. Then $a-b(q+1)\in\set{S}$
+is greater than zero but less than $r$, which is a contradiction since $r$ is the smallest positive
+integer in $\set{S}$. Therefore $a=qr+b$ and $0\leq r\leq b$.\\
+Assume $a=bq+r$ and $a=bg+h$ where $0\leq r\leq b$ and $0\leq h\leq g$.
+\end{quote}
 \end{enumerate}

### 2016-08-23 pde_two/threshold.tex
starting new semester

@@
 F'(t)&=\int_{0}^{\pi}u_{t}\sin{x}\dx\\
-&=\int_{0}^{\pi}(u_{xx}+u^{3})\sin{x}\dx\\
-&=
+&=\int_{0}^{\pi}(u_{xx}+u^{3})\sin{x}\dx
 \end{align*}
+Integrating leaves us with
+\begin{equation}\label{fprime}
+F'(t)=-F(t)+\int_{0}^{\pi}u^{3}\sin{x}\dx.
+\end{equation}
+
+We now want to determine bound on $F(t)$. Recall H\"older's inequality when concerned with the
+Lebesgue measure:
+\[
+\int_{D}|f(x)g(x)|\dx\leq
+\left(\int_{D}|f(x)|^{p}\dx\right)^{1\over p}
+\left(\int_{D}|g(x)|^{q}\dx\right)^{1\over q}
+\]
+where $1/p + 1/q=1$. Letting $p=3$ and $q=3/2$, we have
+\begin{align*}
+F(t)&=\int_{0}^{\pi}\sin^{1/3}{(x)}u\sin^{2/3}{(x)}\dx\\
+&\leq
+\left(\int_{0}^{\pi}(\sin^{2/3}{x})^{3/2}\dx\right)^{2/3}
+\left(\int_{0}^{\pi}(u\sin^{1/3}{x})^{3}\dx\right)^{1/3}\\
+&=\left(\int_{0}^{\pi}\sin{x}\dx\right)^{2/3}
+\left(\int_{0}^{\pi}u^{3}\sin{x}\dx\right)^{1/3}\\
+&= 2^{2/3}\left(\int_{0}^{\pi}u^{3}\sin{x}\dx\right)^{1/3}.
+\end{align*}
+Cubing both sides of the inequality yields
+\begin{align*}
+F^{3}(t)&\leq4\int_{0}^{\pi}u^{3}\sin{x}\dx\\
+{1\over4}F^{3}(t)&\leq\int_{0}^{\pi}u^{3}\sin{x}\dx\\
+\end{align*}
+By \eqref{fprime}, we have that
+\begin{equation}\label{fpineq}
+F'(t)\geq-F(t)+{1\over4}F^{3}(t).
+\end{equation}
+
+We now need to show that the solution to \eqref{nlinh} blows up as $F(t)\to\infty$. We begin by
+letting $G(t) = F^{-2}(t)$. Then
+\begin{align*}
+G'(t)&=-2F^{-1}(t)F'(t)\\
+-{1\over2}G'(t)F(t)&=F'(t).
+\end{align*}
+We can substitute this into the inequality in \eqref{fpineq}.
+\begin{align*}
+-{1\over2}G'(t)F(t)&\geq-F(t)+{1\over4}F^{3}(t)\\
+G'(t)&\leq2-{1\over2}F^{2}(t)\\
+G'(t)&\leq2-{1\over2G(t)} %% TODO -- fix this :(
+\end{align*}
+Solving this differential inequality for $G(t)$ yields
+\[
+G(t)\leq{1\over4}+\left(G(0)-{1\over4}\right)e^{2t}.
+\]
+Back substituting $G(t)=F^{-2}(t)$ gives us
+\begin{align}
+F^{-2}(t)&\leq{1\over4}+\left({1\over F^{2}(0)}-{1\over4}\right)e^{2t}\\
+F(t)&\geq\left[{1\over4}+\left({1\over F^{2}(0)}-{1\over4}\right)e^{2t}\right]^{-1/2}\label{final_ineq}.
+\end{align}
+
 \end{document}

### 2016-08-23 religions_of_the_west/paper.tex
starting new semester

\documentclass[12pt]{hw}

\usepackage{setspace}
\usepackage[margin=1in]{geometry}

\title{Taoism}
\date{}

\newcommand{\inlinetitle}{\vspace{-1.5cm}{\let\newpage\relax\maketitle}\vspace{-1.5cm}}
\doublespacing

\begin{document}
\noindent Steven Rosendahl\\
\noindent Religions of the West\\
\inlinetitle

Taoism has been an incredibly influential religion in Chinese culture over the last several centuries. It's reach has been broad; several revolutions have been sparked by Taoist ideals, and many Chinese dynasties have claimed to share the lineage of Laozi, the Chinese philosopher responsible for writing down the Tao Te Ching. Taoism is unique in that it combines religion, philosophy, and myth. the Tao is considered to be the ultimate goal of Taoism, and texts such as the Tao Te Ching and the Chuang-tzu were written to be moral guides that teach the way to reach the Tao, while much of the history of Taoism involves tales of magic and immortality that are widely considered to be myth. The history of has been shaped by China just as much as Taoism has shaped the history of China.

Laozi is considered to be a sort of prophet, in the loosest sense of the word. Born at the age of 61, Laozi lived in a time when the Chou dynasty was crumbling, and much of China was divided into small territories all vying for power. During this period of turmoil, Laozi decided to leave the civilized area of China and live as a hermit in the wild. It was during this time that he wrote the Tao Teh Ching while trying to leave China. According to legend, Laozi was exiting China through the western gate when a soldier stopped him. The soldier felt that Laozi had much to teach, and refused to let him exit until he had written his teachings down. Laozi saw no way out of the situation, and agreed to produce the Tao Teh Ching.

The Tao Teh Ching, a set of eighty-one cryptic verses, is considered to be the most popular text in Taoism. Many of the verses focus on how to better one's self, so that they may follow the Tao, which literally translates into ``Path". Several very famous saying have originated from the Tao Teh Ching, such as:
\begin{quote}
We make a vessel from a lump of clay;\\
It is the empty space within the vessel that makes it useful.\\
We make doors and windows for a room;\\
But it is these empty spaces that make the room livable.\\
\textbf{$-$ Laozi, Tao Teh Ching, Ch. 11}\\\\
He who knows does not speak.\\
He who speaks does not know.\\
\textbf{$-$ Laozi, Tao Teh Ching, Ch. 56}
\end{quote}
The ideas in the Tao Teh Ching reflect Laozi's feelings towards the state of China at the time. The above quotation emphasizes the usefulness of emptiness, while many of the feudal lords sought to have as much power and land as possible. The beginning lines of chapter 29 emphasize that mankind is not the ruler of the nature.
\begin{quote}
Does anyone want to take the world and do what he wants with it?\\
I do not see how he can succeed.\\
\textbf{$-$ Laozi, Tao Teh Ching, Ch. 29}
\end{quote}
Laozi's writings primarily focus on the importance of nature and the temporality of life; he believed that power and possessions were not what the people of China should be focused on. The Tao Teh Ching is not the only sacred writing in the Taoist canon, however.

Another very famous set of writings is the Chuang-tzu, named after it's author. Chuang-tzu lived in the time of the Chou dynasty as well, although he was born after Laozi. During his younger years, Chuang-tzu experienced the horrors of the wars that were tearing apart China. Many of his writings were in response to the death and pain that he saw around him. One legend surrounding Chuang-tzu describes how, upon his wife's passing, he was found playing music and happily singing. When questioned, Chuang-tzu replied that he was happy for his wife's passing, as she had returned to nature and her death was only a loss from the human perspective. Chuang-tzu believed that the wants of humans were petty and ultimately ignored by the universe. The teachings of Laozi and Chuang-tzu are considered to be two sides to the same coin; both philosophers had a very different view on the Tao, yet both taught the same fundamental ideas.

Laozi and Chuang-tzu, while being incredibly influential in Taoist teaching, never considered themselves to be religious teachers; In Taoist mythology, Laozi was revered as a deity and Chuang-tzu was considered to be immortal. The first true religious teacher of Taoism is considered to be Chang Tao-Ling. According to legend, Tao-Ling had mastered the Tao Teh Ching as well as the Five Classics of Confucianism by the time he was seven. Taoist lore states that Tao-Ling was a magician, capable of turning invisible, teleporting, and cloning himself. When he came of age, Tao-Ling decided that he would search for the Tao; he abandoned the teaching of Confucius and set out to find the Tao. During his time in the mountains of China, he encountered the deified form of Laozi, who gave him the secret to immortality. Tao-Ling, however, saw this as a chance to help all of the people of China. He is credited with founding the first religious organization for Taoism, as well as being the first true teacher of the religion. Many of his teaching are still widely followed today.

One of the key ideas of Taoism is the relationship that humans have with nature. According the the teachings of Taoism, all humans came from nature, and, as such, should be one with nature during one's life. The Tao is meant to guide humans in being one with nature; being in tune with the Tao is being in tune with nature. Some very well known symbols, such as Yin-Yang and parts of the Chinese zodiac originate from Taoism. One must be well-balanced to stay with the Tao, which is reflected in the Yin-Yang symbol. The Tao itself is the central focus of Taoism. There is no one deity to which Taoists must bow; instead, they simply must seek out the Tao. In the Tao Teh Ching, Laozi describes the Tao by saying ``I know not its name, so I style it the way. I give it the makeshift name of the great." Chuang-tzu described the Tao as an force that ``gave spirits and rulers their spiritual powers." Interestingly, the Tao, while being the creator of all, is not immediately present in the world. While many religions present gods who involve themselves in mankind, the Tao is considered to just be; it does not impose itself on humanity, since, according to Chuang-tzu, humanity is too petty for it. Laozi explains this behavior by saying that ``The Tao nurtures her children by allowing them to be themselves, so that when they are grown, they will be able to stand on their own." In other words, the Tao creates everything to be independent, without forcing its creations to follow it.

The concept of balance is very important to Taoism. Taoists believe that there is always another side to every event that transpires in the universe. For example, Taoists believe that the human world and the non-human world are connected so that events in one world impact the other. Many political leaders of China followed this principal and would attempt to please the non-human world in order to bring good fortune to their lives. The Tao is also considered to be ``self-balancing", meaning that if the Tao was brought off balance, it would bring balance back by whatever means necessary. To avoid drastic changes in balance, followers of Taoism would seek out balance in life, and avoid doing anything considered to be extreme. The ideas of nature and balance are reflected in the teaching of the Five Elements.

The Five Elements are a combination of balance, nature, and humankind. They are Wood, Fire, Earth, Metal, and Water, and each one is related to every other element. For example, Fire can destroy Metal, help the Earth, gain strength from Wood, and be weakened by Water. While the Yin-Yang symbol shows the duality of actions, Taoist belief states that it is not always evenly divided into one side or the other.

\end{document}

### 2016-08-23 resume/work.tex
starting new semester

@@
 \begin{itemize}
-\item Developed Android applications for Google Glass, Vuzix, and Recon Jet headsets
+\item Developed Android applications for Google Glass, Vuzix M100, and Recon Jet headsets
 \item Worked closely with a university professor to plan and create the application
-\item Participated in laser safety training
+\item Participated in laser safety training and received laser operator certification
+
 \end{itemize}
@@
 \begin{itemize}
-\item IT Customer Service Representative
-\begin{itemize}
-\item Aided customers with various software related IT issues
-\item Assisted technicians with tasks such as installing hardware
-\end{itemize}
-\item Tutor for the Center for Academic Success
+\item Tutor for the Center for Academic Success (Fall 2015 - Present)
 \begin{itemize}
@@
 \item Created various resources for other Calculus tutors to use
+\end{itemize}
+\item IT Customer Service Representative (Fall 2014 - Spring 2016)
+\begin{itemize}
+\item Aided customers with various software related IT issues
+\item Assisted technicians with tasks such as installing hardware
 \end{itemize}

### 2016-08-23 software_design/madlibs/js/main.js

@@ -1,0 +1,86 @@
+var globalMap = {};
+var madlib = "";
+
+/**
+* loads the file at the given filename (relative to the caller)
+*
+* filename -- the name of the file
+* onFileReady -- a callback that's passed the contents of the file when the read is finished
+* onFileErr -- a callback that's called when an error occurs
+*/
+function loadFileAt(filename, onFileReady, onFileErr) {
+    var fileHandle = new XMLHttpRequest();
+    fileHandle.open("GET", filename, false);
+    fileHandle.onreadystatechange = function(){
+        if(fileHandle.readyState === 4){
+            if(fileHandle.status === 200 || fileHandle.status === 0){
+                onFileReady(fileHandle.responseText);
+            } else {
+                onFileErr();
+            }
+        }
+    }
+    fileHandle.send(null);
+}
+
+/**
+* loads a random madlib
+*
+* return -- the string contents of the madlib
+*/
+function getRandomMadlib(){
+    var madlibs = ["madlib_01.txt","madlib_02.txt","madlib_03.txt","madlib_04.txt"];
+    var random = Math.floor(Math.random() * madlibs.length);
+
+    var chosenMadlib = "./mlibs/" + madlibs[random];
+    globalMap = {};
+    document.getElementById('madlib_table').innerHTML = "";
+    // var chosenMadlib = "./mlibs/madlib_01.txt";
+
+    loadFileAt(chosenMadlib, parseMadLib, function(){
+        console.log("error :(");
+    });
+}
+
+/**
+* parses the madlib and poplates the DOM with the appropriate fields
+*
+* madlib -- The raw madlib content
+*/
+function parseMadLib(ml){
+    madlib = ml;
+    var table = document.getElementById('madlib_table');
+    var madlibVarsRx = /\$[a-zA-Z0-9_ ']*\$\$/g;
+    var madlibFields = madlib.match(madlibVarsRx);
+
+    console.log(madlibFields);
+    var rowCount = 0;
+    madlibFields.forEach(function(currentValue, index, array){
+        var fieldName = currentValue.substring(1, currentValue.length - 5);
+        var fieldID = currentValue.substring(currentValue.length - 4, currentValue.length -2);
+        if(!(fieldID in globalMap)){
+            var row = table.insertRow(rowCount++);
+            var cell1 = row.insertCell(0);
+            var cell2 = row.insertCell(1);
+            globalMap[fieldID] = "";
[…]

### 2016-08-30 padics/hw1.tex
working on p-adics

\documentclass{article}

\usepackage{geometry}
\usepackage{amssymb,amsmath}
\geometry{margin=1in}

\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\NN}{\mathbb{N}}

\begin{document}
\begin{enumerate}
\item If $F$ is a field in which 0 has a mutiplicative inverse, show $|F|=1$.
\begin{quote}
\textbf{\textit{Proof:}} Suppose $|F|>1$, and let $x\in F$ such that $x=0^{-1}$. Then $0x=1=0$. We know that under multiplication, any element in $F$ can be expressed as $1\times y\ \forall y\in F$, and since $1=0$, $1\times y = 0 \times y = 0$. Therefore the only element in $F$ is 0, and $|F|=1$
\end{quote}

\item Suppose that $n>1$ is an integer and let $\ZZ_{n}$ be equipped with addition and multiplication modulo $n$. Prove that $\ZZ_{n}$ is a field if and only if $n$ is prime.
\begin{quote}
In order to be a field, $\ZZ_{n}$ must fulfill the following axioms:
\begin{enumerate}
\item[\textbf{(A1)}] Addition is commutative on $F$.
\item[\textbf{(A2)}] Addition is associative on $F$.
\item[\textbf{(A3)}] There is a uniqe additive identity, called 0.
\item[\textbf{(A4)}] There is an additive inverse $-a$ for all $a\in F$.
\item[\textbf{(M1)}] Multiplication is commutative on $F$.
\item[\textbf{(M2)}] Multiplication is associative on $F$.
\item[\textbf{(M3)}] There is a unique multiplicative identity called 1.
\item[\textbf{(M4)}] There is an multiplicative inverse element $a^{-1}$ for every $a\in F$.
\item[\textbf{(D)}] For all $x,y,z\in F$, $x\cdot(y+z)=(x\cdot y)+(x\cdot z)$.
\item[\textbf{(ZO)}] The additive and multiplicative identity are distinct.
\end{enumerate}

\textbf{\textit{Proof:}} Suppose $n$ is prime. Then every element in $\ZZ_{n}$ is a unit, so $\ZZ_{n}$ is equipped with a multiplicative inverse. Since $\ZZ_{n}$ has addition and multiplication, we know it is commutative and associateve for both those operations. We know that $\ZZ_{n}$ contains 0, so it has the additive inverse. Addition and multiplication form the distributive law, so $\ZZ_{n}$ is equipped with the distributive law. Since $n>1$, $\ZZ_{n}$ $0\neq1$. Therefore, by defintion, $\ZZ_{n}$ is a field since it fulfills all the field axioms.\\\\
Suppose $\ZZ_{n}$ is a field with $n$ not prime. Since $n$ is not prime, then we can find an element $x\in\ZZ_{n}$ such that $gcd(n,x)\neq1$. Therefore there is an element in $\ZZ_{n}$ that is not a unit, so $\ZZ_{n}$ is not a field, since there is an element without an inverse.
\end{quote}

\item Suppose that $F$ is a finite field and $x\in F\setminus\{0\}$. Prove that there exists $n\in\NN$ such that $x^{n}=1$.
\end{enumerate}
\end{document}

### 2016-09-01 padics/hw1.tex
pushy push

@@
 \item Suppose that $F$ is a finite field and $x\in F\setminus\{0\}$. Prove that there exists $n\in\NN$ such that $x^{n}=1$.
+\begin{quote}
+\textit{\textbf{Proof:}} Consider $x^{r}=x^{s}$ for $s\neq r\in\NN$ where $s>r$ without loss of generality.
+We know that for every element in $F$, there is a multiplicative inverse of that element, so we define $(x^{r})^{-1}=x^{-r}$.
+Then $x^{-r}x^{r}=x^{-r}x^{s}$ and $1=x^{s-r}$.
+Since $s>r$, $s-r\in\NN$, so we have found the $n$ we were looking for.
+\end{quote}
 \end{enumerate}

### 2016-09-01 misc/gpa.py
hopefully

@@ -1,0 +1,11 @@
+# A program that calculates your GPA
+# @author Steven Rosendahl
+
+print("This program calculates grade point average.")
+i = 0
+avg = 0.0
+while(i < 5):
+    print("Enter course %d grade->" % (i+1))
+    avg = avg + float(input())
+    i = i + 1
+print("Your GPA for last semester was: %.2f" % (avg / 5))

### 2016-09-01 misc/lottery.py

@@ -1,0 +1,16 @@
+# A sassy little lucky number generator.
+# @author Steven Rosendahl
+
+from random import randint
+
+print("Enter the low score--->")
+low = int(float(input()))
+print("Enter the high score--->")
+high = int(float(input()))
+
+if low > high:
+    print("You daft bimbo! Your low score can't be higher than your high score!")
+elif low == high:
+    print("You didn't really give me much choice... take a wild guess what your lucky number is.")
+else:
+    print("Your lucky lottery number is %d" % randint(low,high))

### 2016-09-01 software_design/inclass/j_range.js

@@ -1,0 +1,9 @@
+function onSubmitPressed(){
+    var upper = parseInt(document.getElementById('upper-input').value);
+    var lower = parseInt(document.getElementById('lower-input').value);
+
+    var val = Math.random() * (upper - lower + 1);
+    val += lower;
+
+    document.getElementById('result-field').innerHTML = val;
+}

### 2016-09-01 software_design/madlibs/js/main.js

@@ -45,15 +45,13 @@ function getRandomMadlib(){
 /**
 * parses the madlib and poplates the DOM with the appropriate fields
 *
-* madlib -- The raw madlib content
+* ml -- The raw madlib content
 */
 function parseMadLib(ml){
     madlib = ml;
     var table = document.getElementById('madlib_table');
     var madlibVarsRx = /\$[a-zA-Z0-9_ ']*\$\$/g;
     var madlibFields = madlib.match(madlibVarsRx);
-
-    console.log(madlibFields);
     var rowCount = 0;
     madlibFields.forEach(function(currentValue, index, array){
         var fieldName = currentValue.substring(1, currentValue.length - 5);

### 2016-09-01 misc/report.tex
finishing assignments all the time

\documentclass{hw}

\begin{document}
\noindent\textbf{Steven Rosendahl}\\
\noindent\textbf{CPSC 125 Lab I}\\
\section*{Careers in CS and CE}
\begin{enumerate}
\item What are five professions within the fields of computer science or computer engineering? Provide a short job description for each of them. What is the income potential for these types of jobs? Give numbers. How do they compare to other professions? Give a few examples. What are the employment projections for the next 5 to 10 years for these five professions?
\begin{enumerate}
\item Cybersecurity - Writing code to protect people on the web. The pay is an average of \$100,000 a year. The job market for Cybersecurity is very understaffed and will continue growing in the next 5-10 years.
\item Mobile Development - Designing and building applications for mobile devices (tables, phones, wearables, etc...). The pay is an average of \$72,000 annually. Nearly every website I visited suggested that the job market would continue growing for this job.
\item Web Developer - Building webpages and web-apps on the internet and sometimes working on back-end code. The average pay is about \$55,000. The demand for web developers is projected to grow 27\% in the next several years.
\item Developer Operations Specialist - A devops is someone who suffers at the hands of the developers. A devops is in charge of managing systems and teams when developing software. The average salary for a devops specialist is \$100,000 a year. The field is projected to grow, as well as provide room for people to move into management. This is one of the more higher-paying jobs that I looked up.
\item Database Manager - A database manager is someone who manages databases, usually for a large company. The average pay is around \$70,000 per year. The market is projected to grow around 11\% in the next few years.
\end{enumerate}
\item Which CS/CE specialization does your team consider the most interesting one? Why?
\begin{quote}
I think that Mobile Development is one of the more interesting fields, since mobile devices are becoming more and more powerful and widespread every day. The market has a lot of promise and doesn't seem to be in high demand.
\end{quote}
\item What are some other interesting facts about CS/CE careers that you came across during your research?
\begin{quote}
I searched for CE jobs a few times and most of the CE jobs were programming based, which is not something that I expected. Typically, I think of CE people as designing circuitry and working more with hardware.
\end{quote}
\end{enumerate}

\section*{Interviewing Another Student}

I interviewed Matt McCarthy for this part of the assignment.
\begin{enumerate}
\item What is your major and why did you choose it?
\begin{quote}
Computer Science, since CNU doesn't have a math masters
\end{quote}
\item What do you like/dislike about your major?
\begin{quote}
I like that it gives me an excuse to do math/focus on the theoretical side of math. I don't like having to run simulations and experiments.
\end{quote}
\item What was your favorite class you took?
\begin{quote}
Parallel Processing (a 600 level class)
\end{quote}
\item What's your favorite proof? Your least favorite proof?
\begin{quote}
Favorite: Fundamental Homomorphism Proof. Least Favorite: Anything that involves heavy calculation.
\end{quote}
\item Do you have any advice for students in the major?
\begin{quote}
Don't be afraid of math.
\end{quote}
\end{enumerate}
\end{document}

### 2016-09-01 padics/hw1.tex
hopefully
finishing assignments all the time

@@
-\documentclass{article}
-
-\usepackage{geometry}
+\documentclass{hw}
 \usepackage{amssymb,amsmath}
-\geometry{margin=1in}
-
-\newcommand{\ZZ}{\mathbb{Z}}
-\newcommand{\NN}{\mathbb{N}}
 
 \begin{document}
+\makeheader{1}
+
 \begin{enumerate}
-\item If $F$ is a field in which 0 has a mutiplicative inverse, show $|F|=1$.
+\item If $F$ is a field in which 0 has a multiplicative inverse, show $|F|=1$.
 \begin{quote}
@@
 \item[\textbf{(A2)}] Addition is associative on $F$.
-\item[\textbf{(A3)}] There is a uniqe additive identity, called 0.
+\item[\textbf{(A3)}] There is a unique additive identity, called 0.
 \item[\textbf{(A4)}] There is an additive inverse $-a$ for all $a\in F$.
@@
 Then every element in $\ZZ_{n}$ is a unit, so $\ZZ_{n}$ is equipped with a multiplicative inverse.
-Since $\ZZ_{n}$ has addition and multiplication, we know it is commutative and associateve for both those operations.
+Since $\ZZ_{n}$ has addition and multiplication, we know it is commutative and associative for both those operations.
 We know that $\ZZ_{n}$ contains 0, so it has the additive inverse.
@@
 Since $n>1$, $\ZZ_{n}$ $0\neq1$.
-Therefore, by defintion, $\ZZ_{n}$ is a field since it fulfills all the field axioms.\\\\
+Therefore, by definition, $\ZZ_{n}$ is a field since it fulfills all the field axioms.\\\\
 Suppose $\ZZ_{n}$ is a field with $n$ not prime.

### 2016-09-04 padics/hw2.tex
more padic work

\documentclass{hw}

\begin{document}
\makeheader{2}

\begin{enumerate}
\item If $F$ is a finite field, prove that the only absolute value on $F$ is the trivial absolute
value.
\begin{quote}
\textit{\textbf{Proof:}} By definition, if we take any $x\in F$ such that $x\neq0$, then there is some
integer $m\in\ZZ$ such that $x^{m}=1$, meaning that $|x|=1$. If $x=0$, then $|x|=0$, so the absolute
value is trivial.
\end{quote}

\item Suppose $p$ is prime. Prove that $|p^{n}|_{p}$ tends to 0 as $n\to\infty$.
\begin{quote}
\textit{\textbf{Proof:}} We know that $v_{p}(x)>0$ for all $x\in\ZZ$, and that $v_{p}(x)$ counts the
number of factors of $p$ in $x$. If we consider $p^n$, there are $n$ factors of $p$, meaning that
$v_{p}(p^{n})=n$. Then $|p^{n}|_{p}=p^{-v_{p}(p^{n})}=p^{-n}$, so
\[
\lim_{n\to\infty}p^{-n}=0.
\]
\end{quote}

\item Suppose that $F$ is a field and $|\ |$ is a non-Archemedian absolute value on $F$. if $c\in F$
and $r$ is a positive real number, we define the \textit{open ball (or disk) centered at c of radius r}
by
\[
B(c,r)=\{x\in F:|x-c|<r\}.
\]
If $d\in B(c,r)$, prove that $B(c,r)=B(d,r)$.
\end{enumerate}
\end{document}

### 2016-09-08 padics/hw2.tex
updating hw

@@
 \documentclass{hw}
+
+\geometry{margin=1in}
 
@@
 
-\item Suppose that $F$ is a field and $|\ |$ is a non-Archemedian absolute value on $F$. if $c\in F$
+	\item Suppose that $F$ is a field and $|\ |$ is a non-Archemedian absolute value on $F$. If $c\in F$
 	      and $r$ is a positive real number, we define the \textit{open ball (or disk) centered at c of radius r}
@@
 	      If $d\in B(c,r)$, prove that $B(c,r)=B(d,r)$.
+	      \begin{quote}
+	      	\textit{\textbf{Proof:}} Suppose $x\in B(d,r)$. Then
+			\begin{align*}
+				|x-d| &= |x-d-c+c|\\
+				&= |x-c + c-d|\\
+				&\leq \text{max}\{|x-c|,|c-d|\}\\
+				&= \text{max}\{|x-c|,|d-c|\}\\
+				&\leq\text{max}\{|x-c|,r\}.
+			\end{align*}
+			Suppose $|x-d|\leq |x-c|$. Then the value $|x-d|$ is always within the disk $B(c,r)$.\\\\
+			Suppose $x\in B(c,r)$. Then
+			\begin{align*}
+				|x-c| &= |x-c-d+d|\\
+				&= |x-d + d-c|\\
+				&\leq \text{max}\{|x-d|,|d-c|\}\\
+				&<\text{max}\{|x-d|,r\}.
+			\end{align*}
+			Suppose max$\{|x-d|,r\} = |x-d|$. Then $|x-c|$ is always within the disk $B(d,r)$.
+	      \end{quote}
 \end{enumerate}

### 2016-09-08 math_modeling/hw/chw/math.c

@@ -5,45 +5,45 @@
 
 int main(int argc, const char* argv[]){
 
-    float temps [10000];
-    temps[0] = 298.15;
+        float temps [10000];
+        temps[0] = 298.15;
 
-    float starttime = 0;
-    float deltat;
-    if(argc > 1)
-        deltat = atof(argv[1]);
-    else
-        deltat = 0.001;
-    float c2 = 0.991667;
-    float c1 = 1.27575 * powf(10,-10);
-    float tinf = 473.15;
-    float tsur = powf(673.15,4);
+        float starttime = 0;
+        float deltat;
+        if(argc > 1)
+                deltat = atof(argv[1]);
+        else
+                deltat = 0.001;
+        float c2 = 0.991667;
+        float c1 = 1.27575 * powf(10,-10);
+        float tinf = 473.15;
+        float tsur = powf(673.15,4);
 
-    FILE* fp = fopen("data.txt","w+");
+        FILE* fp = fopen("data.txt","w+");
 
-    int i = 0;
-    for(i = 0; i < 10000 && temps[i] < 490.85; i++){
-        if(i % 10 == 0){
-            fprintf(fp, "%f\n", temps[i]);
+        int i = 0;
+        for(i = 0; i < 10000 && temps[i] < 490.85; i++) {
+                if(i % 10 == 0) {
+                        fprintf(fp, "%f\n", temps[i]);
+                }
+                temps[i + 1] = (-1) * (deltat/c2) * (temps[i] - tinf + c1 * (pow(temps[i],4) - tsur)) + temps[i];
+                starttime += deltat;
         }
-        temps[i + 1] = (-1) * (deltat/c2) * (temps[i] - tinf + c1 * (pow(temps[i],4) - tsur)) + temps[i];
-        starttime += deltat;
-    }
 
-    fclose(fp);
+        fclose(fp);
 
-    int status;
-    int pid = fork();
-    if(pid == 0){
-        execl("/usr/local/bin/RScript", "/usr/local/bin/RScript",
-            "/Users/sjyn/Documents/LahTech/math_modeling/hw/chw/plot.R", (char *)NULL);
-    } else if(pid > 0) {
-        printf("The temperature at time %f is %f\n", starttime, temps[i]);
-    } else {
-        exit(2);
-    }
-    if(waitpid(pid, &status, 0) != pid)
-        status = -1;
+        int status;
+        int pid = fork();
+        if(pid == 0) {
[…]

### 2016-09-11 padics/hw3.tex
uggggg

\documentclass{hw}

\begin{document}
    \makeheader{3}
    \begin{enumerate}
        \item Suppose $F$ is a field with a non-Archemedian absolute value $|\cdot|$ and $\IRing{F}$ is the ring of
        integers of $F$. Prove that a point $x\in\IRing{F}$ is a unit if and only if $|x|=1.$
        \begin{quote}
            Suppose $x\in\IRing{F}$ is a unit and consider $x^{-1}\in\IRing{F}$ as the inverse of $x$. Since $x\in\IRing{F}$, we know
            \begin{align*}
                |x| &\leq 1\\
                |x^{-1}||x| &\leq |x^{-1}|\leq1 \qquad\text{(since $x^{-1}\in\IRing{F}$)}\\
                |x^{-1}x| &\leq |x^{-1}|\leq1\\
                |1| &\leq |x^{-1}| \leq1\\
                1 &\leq |x^{-1}| \leq 1.
            \end{align*}
            The only choice we have for $x^{-1}$ is 1, so
            \begin{align*}
                x^{-1} &= 1\\
                xx^{-1}&= x\\
                x&=1\\
                |x|&=1.
            \end{align*}
        \end{quote}

        \item If $n$ is a non-negative integer a $p$ is prime, define $\mathcal{I}_{n}=\{x\in\IRing{p}:|x|_{p}\leq p^{-n}\}$.
        \begin{enumerate}
            \item If $x,y\in\mathcal{I}_{n}$ prove that $x+y\in\mathcal{I}_{n}$.
            \begin{quote}
                Suppose $x,y\in\mathcal{I}_{n}$, and consider $|x+y|_{p}$. Then
                \begin{align*}
                    |x+y|_{p} &\leq \mmax{|x|_{p}}{|y|_{p}}\\
                    &\leq \mmax{p^{-n}}{p^{-n}}\\
                    &= p^{-n}
                \end{align*}
                Therefore $|x+y|_{p}\leq p^{-n}$, so $x+y\in\mathcal{I}_{n}$.
            \end{quote}
        \end{enumerate}
    \end{enumerate}
\end{document}

### 2016-09-14 padics/hw3.tex
updating padcis
excel sheet

@@
             \end{align*}
+
+            Suppose $|x| = 1$.
         \end{quote}
 
-        \item If $n$ is a non-negative integer a $p$ is prime, define $\mathcal{I}_{n}=\{x\in\IRing{p}:|x|_{p}\leq p^{-n}\}$.
+        \item If $n$ is a non-negative integer a $p$ is prime, define
+        $\mathcal{I}_{n}=\{x\in\IRing{p}:|x|_{p}\leq p^{-n}\}$ where\\ $\IRing{p}=\{x\in\QQ : |x|_{p}\leq1\}$.
         \begin{enumerate}
@@
             \end{quote}
+
+            \item If $x\in\mathcal{I}_{n}$ and $r\in\IRing{p}$ prove that $rx\in\mathcal{I}_{n}$.
+            \begin{quote}
+                Let $|\cdot|$ be the p-adic absolute value and consider $|rx| = |r|\cdot|x|$.
+Since $|x|\leq p^{-n}$ and $|r|\leq1$, we have $|r|\cdot|x|\leq 1\cdot p^{-n} = p^{-n}$, so $|rx|\leq p^{-n}$ which implies that $rx\in\mathcal{I}_{n}$.
+            \end{quote}
+
+            \item What is another way to describe the set $\ZZ\cap\mathcal{I}_{n}$?
         \end{enumerate}

### 2016-09-14 software_design/excel/script.js

@@ -1,0 +1,90 @@
+var app = angular.module("excel_app",[]);
+app.controller('excel_controller', function($scope){
+    $scope.sheet = [];
+    init($scope);
+});
+
+function init($scope){
+    $scope.sheet = [
+        {
+            b1: 0,
+            b2: 0
+        },
+        {
+            b1: 0,
+            b2: 0
+        },
+        {
+            b1: 0,
+            b2: 0
+        },
+        {
+            b1: 0,
+            b2: 0
+        }
+    ];
+
+    $scope.totals = [0,0,0,0,0];
+    $scope.reorders = [false,false,false,false,false];
+
+    $scope.reorderN1 = function (){
+        $scope.countInd = 2;
+        $scope.sheet.sort(function(obj1, obj2){
+            var diff = obj1.b1 - obj2.b1;
+            return diff;
+            // var ud = $scope.reorders[0];
+            // $scope.reorders[0] = !$scope.reorders[0];
+            // if(ud){
+            //     return diff;
+            // } else {
+            //     return (-1) * diff;
+            // }
+        });
+    };
+    $scope.reorderN2 = function(){
+        $scope.countInd = 2;
+        $scope.sheet.sort(function(o1,o2){
+            return o1.b2 - o2.b2;
+        });
+    };
+    $scope.reorderSum = function(){
+        $scope.countInd = 2;
+        $scope.sheet.sort(function(o1, o2){
+            return (o1.b1 + o1.b2) - (o2.b1 + o2.b2);
+        });
+    };
+    $scope.reorderProd = function(){
+        $scope.countInd = 2;
+        $scope.sheet.sort(function(o1, o2){
+            return (o1.b1 * o1.b2) - (o2.b1 * o2.b2);
+        })
+    };
+    $scope.reorderAvg = function(){
+        $scope.countInd = 2;
+        $scope.sheet.sort(function(o1, o2){
+            return ((o1.b1 + o1.b2) / 2) - ((o2.b1 + o2.b2) / 2);
+        })
+    };
+
+    $scope.addItem = function(){
+        $scope.sheet.push({
+            b1: 0,
+            b2: 0
+        });
+        $scope.totals.push(0);
+    };
+
+    $scope.recalcTotals = function(){
+        for(var i = 0; i < $scope.totals.length; i++){
+            $scope.totals[i] = 0;
+        }
+        for(var i = 0; i < $scope.sheet.length; i++){
[…]

### 2016-09-20 padics/cont_fcn.tex
update proofs

\documentclass{hw}

\begin{document}

    \noindent We can define a function $f:\RR\to\RR$ by
    \[
        f(x)=
        \begin{cases}
            0, \qquad x\in\QQ\\
            1, \qquad x\in\RR\setminus\QQ
        \end{cases}.
    \]
    We know this function to be dicontinuous under the usual absolute value. In $\RR$ under the usual absolute value,
    we define a function $h$ to be continuous at a point $a$ in the domain of $h$ by
    \[
        \forall\epsilon > 0,\ \exists\delta > 0 \text{ such that } 0<|x-a|<\delta\implies|h(x)-h(a)|<\epsilon.
    \]
    We can reformulate this definition in the p-adic sense by
    \[
        \forall\epsilon > 0,\ \exists\delta > 0 \text{ such that } 0<|x-a|_{p}<\delta\implies|h(x)-h(a)|_{p}<\epsilon.
    \]
    If we refomulate our original function, then we can show that under the p-adic absolute value, the function is
    continuous at every point. We will define $g:\QQ_{p}\to\QQ_{p}$ as
    \[
        g(x)=
        \begin{cases}
            0, \qquad x\in\ZZ\\
            1, \qquad x\in\QQ_{p}\setminus\ZZ
        \end{cases}
    \]
    as an analog to the funtion $f$ we defined earlier. We will show that this function is indeed continuous.

    \begin{quote}
        \textit{\textbf{Proof:}} Let $\epsilon>0$ be given. If $g$ is continuous at a point $c$, then $\exists\delta>0$
        such that $\forall\epsilon >0$, $|x-c|_{p}<\delta\implies|g(x)-g(c)|_{p}<\epsilon$. If we let $\delta=1$, then we have

    \end{quote}
\end{document}

### 2016-09-20 padics/hw3.tex
update proofs

@@
         \begin{quote}
-            Suppose $x\in\IRing{F}$ is a unit and consider $x^{-1}\in\IRing{F}$ as the inverse of $x$. Since $x\in\IRing{F}$, we know
+            Suppose $x\in\IRing{F}$ is a unit and consider $x^{-1}\in\IRing{F}$ as the inverse of $x$. We know
+            $x^{-1}\in\IRing{F}$ since $x$ is a unit in $\IRing{F}$. Since $x\in\IRing{F}$, we know
             \begin{align*}
@@
 
-            Suppose $|x| = 1$.
+            Suppose $|x| = 1$. We want to show that $x^{-1}\in\IRing{F}$. Consider $|x^{-1}|$ where $x^{-1}\in F$.
+            Then
+            \begin{align*}
+                |x|&=1\\
+                |x||x^{-1}| &= |x^{-1}|\\
+                |xx^{-1}| &= |x^{-1}|\\
+                |1| &= |x^{-1}|\\
+                1 &= |x^{-1}|.
+            \end{align*}
+            Therefore $|x^{-1}|=1$, so $x^{-1}\in\IRing{F}$ and therefore $x$ is a unit. 
         \end{quote}
@@
             \item What is another way to describe the set $\ZZ\cap\mathcal{I}_{n}$?
+            \begin{quote}
+                We know by definition of the p-adic absolute value that for all $x\in\ZZ$, $|x|_{p}\leq1$. This implies
+                $\ZZ\subset\IRing{p}$. We want to find all $x\in\ZZ$ where $|x|_{p}\leq p^{-n}$ for some $n\in\NN$.
+                We cannot consider any $x$ where $x$ is a multiple of $p$, since those values would be in $\QQ$. This
+                leaves only 1 (when $x$ does not share any factors of $p$) or 0 (when $x=0$). For any value $p$,
+                $0 < p^{-n} < 1$, so the only value we have is $x=0$. Therefore, $\ZZ\cap\mathcal{I}_{n}=\{0\}$.
+            \end{quote}
         \end{enumerate}

### 2016-09-20 padics/lemma_pf.tex
update proofs

\documentclass{hw}

\usepackage{amsthm}
\newcommand{\FF}{\mathbb{F}}
\newtheorem{theorem}{Lemma}

\begin{document}
% \begin{center}
\section*{Equality of Absolute Values}

\begin{theorem}
	Let $|\cdot|_{1}$ and $|\cdot|_{2}$ be absolute values on a field $\FF$. Then the following are
    equivalent:
	\begin{quote}
		\begin{enumerate}
			\item $|\cdot|_{1}$ and $|\cdot|_{2}$ are equivalent (i.e. their topologies are the same).
			\item For any $x\in\FF$, we have $|x|_{1}<1$ if and only if $|x|_{2} < 1$.
			\item There is an $\alpha\in\RR$ such that for all $x\in\FF$, $|x|_{1} = |x|_{2}^{\alpha}$.
		\end{enumerate}
	\end{quote}
\end{theorem}

\begin{proof}
    We will start by showing 1 $\implies$ 2. Suppose $|\cdot|_{1}$ and $|\cdot|_{2}$ are equivalent. Since the topologies are the same, we have that any sequence $a_{n}$ that converges under $|\cdot|_{1}$ will also converge under $|\cdot|_{2}$. If we consider any $x\in\FF$, then we have
    \[
        \lim_{n\to\infty}x^{n}=0
    \]
    exactly when $x<1$, so $|x|_{1}<1$ if and only if $|x|_{2} < 1$.\\\\

    We must now show that 2 $\implies$ 3. Suppose that $x\in\FF$ and that $|x|_{1}<1$ if and only if $|x|_{2} < 1$. Our goal is to find an $\alpha$ such that for all $x\in\FF$, $|x|_{1} = |x|_{2}^{\alpha}$. We can find one such $\alpha$ rather easily. Consider an arbitrary $y\in\FF$ where $|y|_{1}<1|$. Then
    \begin{align*}
        |y|_{1} &= |y|_{2}^{\alpha}\\
        \log{|y|_{1}} &= \log{|y|_{2}^{\alpha}}\\
        \log{|y|_{1}} &= \alpha\log{|y|_{2}}\\
        \alpha &= { \log{|y|_{1}} \over \log{|y|_{2}} }.
    \end{align*}
    We have asserted that such an $\alpha$ exists for $y$, but we want to show that this $\alpha$ does indeed work for any $x\in\FF$ (ideally that $\alpha$ does not depend on our choice of $x$).
    We will choose another element of $\FF\setminus\{0\}$, $z\neq x$. We know that if $|x|_{1}=|y|_{1}$ then $|x|_{2}=|y|_{2}$ since $x/y$ or $y/x$ would have $|\cdot|_{2} < 1$ which would violate the assumption.
\end{proof}

% \end{center}
\end{document}

### 2016-09-20 software_design/inclass/class_stuff.py

@@ -1,0 +1,10 @@
+from PIL import Image, ImageFilter
+
+image = Image.open("./Harambe.jpg")
+
+box = (100, 100, 400, 400)
+region = image.crop(box)
+region = region.transpose(Image.ROTATE_90)
+image.paste(region, box)
+
+image.filter(ImageFilter.DETAIL).show()

### 2016-09-27 religions_of_the_west/test_01.tex
updating hella

\documentclass[8pt]{article}
\usepackage{extsizes}
\usepackage{enumitem}
\usepackage{amsmath}
\setlist{nosep}
\usepackage[margin=0.2in]{geometry}

\begin{document}
    \section*{Test 1 Review}
    \begin{enumerate}
        \item What is the Qur'an? What are the Hadiths?
            \begin{itemize}
                \item Qur'an is the holy book of Islam
                \begin{itemize}
                    \item Infallable revaltion of Muhammad
                    \item Every chapter begins by exahlting Allah
                    \item Accepts many of the prophets from the Old Testament, but is considered to be the true word of
                    God (smaller than New Testament)
                    \item Does not vereate Mary like Christian Bible, does not consider Jesus to be Messiah
                    \item Grants Mohammad special privledges (cannot be cursed, was allowed to be polygamist)
                    \item Text emphasizes wrath (Jihad, fighting is obligatory, idolitors should be killed)
                    \item Discusses final judgment
                \end{itemize}
                \item Hadiths are stories about Muhammad and how he acted in certain situations
                \begin{itemize}
                    \item Allows for clairification of certain situation not specified in Qur'an
                    \item Six collections
                \end{itemize}
            \end{itemize}

        \item What are the five pillars of Islam?
            \begin{enumerate}
                \item \textbf{Confession (Shahada)}: Must confess there is no God other than Allah
                \item \textbf{Prayer (Salat)}
                \begin{itemize}
                    \item Prayer is performed face down to the ground (submission)
                    \item Performed as a community five times a day (towards Mecca)
                \end{itemize}
                \item \textbf{Almsgiving (Zakat)}: Taking care of the poor is emphasized in the Qur'an.
                \item \textbf{Fasting (Sawm)}
                \begin{itemize}
                    \item Takes place during month of Rahmadan (when Qur'an was given to Mohammad)
                    \item Fasting during daylight hours
                \end{itemize}
                \item \textbf{Pilgrammage (Hajj)}: Journey to Mecca to circumnavigate the Ka'ba
            \end{enumerate}

        \item Differentiate the Sunnis, Shiites, Sufis, and Wahhabis
            \begin{enumerate}
                \item \textbf{Sunni}: Majority of Muslims
                \begin{itemize}
                    \item Take pride in the Muslim empire and believe that Islam can become world-dominating
                    \item Interpret scripture through reason and believe in pre-destination
                \end{itemize}
                \item \textbf{Shia}: Iran/Iraq Muslims
                \begin{itemize}
                    \item Believe Mahdi will reveal himself as messianic figure
                    \item Authority is given to clerics
                \end{itemize}
                \item \textbf{Sufi}: Mystical Muslims
                \begin{itemize}
                    \item Look for mystical meaning in the scripture
                    \item Believe Allah is one with humans (very controversial view)
                    \item Meditation on Mohammad's night journey
                    \item Pacifist; not looking for worldly Islam
                \end{itemize}
                \item \textbf{Wahhabi}: Terrorist ideology
                \begin{itemize}
                    \item Radical views of Islam
                    \item Will kill others in the name of Allah (even other Muslims)
                \end{itemize}
            \end{enumerate}

        \item Explain the war-like nature of Islam
            \begin{itemize}
                \item Allah gave Mohammad permission to fight against those who resisted Islam
                \item Mohammad was a warrior; took Mecca back after massing an army in Medina
                \item Qur'an says fighting is necessary (Jihad), but most Muslims consider it to be defensive
                \item War against infidels (Christians, Jews, etc...)
                \item Mujuhadeen (holy warriors) are promised gifts upon death
            \end{itemize}

        \item Relate Judaism and Islam
            \begin{itemize}
                \item Judaism follows from Isaac, Islam follows from Ishmael
                \item Both recognize the Old Testament (Islam sees Bible as sacred but flawed)
                \item Both tend to follow laws very strictly (Torah and Sharia)
                \item Jews did not accept Mohammad as a prophet
            \end{itemize}

        \item Who are the Black Muslims?
            \begin{itemize}
                \item Response to state of blacks in soceity
                \item Edward Byden - thought Islam could strengthen blacks place in society
                \item Timothy Drew - thought blacks had divine origins
                \item Elijah Mohammad
                \begin{itemize}
                    \item Believed he was a prophet of Allah and only he could interpret Qur'an
                    \item Blacks originated from tribe of Shabazz (ties to Arab culture)
                    \item White men were created by scientists wishing to defy Allah $\implies$ white men are evil
                    \item Blacks need to segregate themselves from whites
                \end{itemize}
            \end{itemize}

        \item Describe the process and problems of Hellenism among the Jewish people
            \begin{itemize}
                \item 330 B.C. - Alexander the Great
                \item Greek culture spread into the Isrealite lands
                \item Old Testament was translated into Greek
                \item Many Jews adopted the hellenistic ideas
                \item Some Jews saw Greek culture as too secular (adopting Greek culture meant abandoning the Torah)
            \end{itemize}

\newpage

        \item Who are the Pharisees, Sadducees, Zealots, and Essenes?
            \begin{itemize}
                \item \textbf{Pharisees}: The lower preisthood
                \begin{itemize}
                    \item Viewed Torah as an everchanging set of laws (adaptation to modern life)
                    \item Accepted oral tradition and adapted the Torah to fit it
                    \item Believed hellenistic/Romanized world created by Sadducees was doomed
                \end{itemize}
                \item \textbf{Sadducees}: The high priests
                \begin{itemize}
                    \item Civic and religious heads of society
                    \item Concerned with economic affairs - collected taxes and became very wealthy
                    \item Very conservative religious ideas (rejected angles/apocrypha and emphasized law of Moses)
                    \item Very liberal in secular matters (called Hellenizers) and compromised with many Roman demands
                \end{itemize}
                \item \textbf{Zealots}: Supporters of rebellion against Rome
                \begin{itemize}
                    \item Believed submission to Romans meant forsaking God
                    \item Fought many battles against Roman infantry
                \end{itemize}
                \item \textbf{Essenes}: Watchers of the Messiah
                \begin{itemize}
                    \item Removed themselves from society to await the Messiah
                    \item Small remote communities
                    \item Responsible for the Dead Sea Scrolls
                \end{itemize}
            \end{itemize}

        \item Differentiate Orthodox, Reform, and Conservative Judaism
            \begin{itemize}
                \item \textbf{Orthodox}: By the book Judaism
                \begin{itemize}
                    \item Believe that the written law is the only law
                    \item Very traditional and unbending to modern culture
                \end{itemize}
                \item \textbf{Reform}: Flexible tradition
                \begin{itemize}
                    \item Moral laws are eternal, written laws are not
                    \item Influenced by French Enlightenment
                    \item Did not look for return of the Messiah
                \end{itemize}
                \item \textbf{Conservative}: American Judaism
                \begin{itemize}
                    \item Tradition as a guide but not as the only choice
                    \item Middle ground between Orthodox and Reform
                \end{itemize}
            \end{itemize}

        \item Discuss the following concepts in Judaism: Jew, Synagogue, Kosher, and Sabbath
            \begin{itemize}
                \item \textbf{Jew}
                \begin{itemize}
                    \item Someone descended from the tribe of Isreal
                    \item A chosen person of God
                    \item Someone with a Jewish mother
                    \item Proselytes (converted Jews) must undergo Circumcision (males only) and Baptism
                \end{itemize}
                \item \textbf{Synagogue}
                \begin{itemize}
                    \item Jewish place of worship (\textbf{not} a temple)
                    \item Need 10 adults (orthodox - must be men) to form one
                    \item Arc containing Torah is the centerpiece (reding is very important)
                    \item Usually face towards Jerusalem
                    \item Prayer is very focused done together
                \end{itemize}
                \item \textbf{Kosher}
                \begin{itemize}
                    \item Food is sacred to Jewish people
                    \item Kosher laws are outlined in Deuteronomy/Leviticus
                    \item Keeps Jews separate from other culture
                \end{itemize}
                \item \textbf{Sabbath}
                \begin{itemize}
                    \item Holy day of worship (Saturday)
                    \item No work is to be done on Sabbath
                \end{itemize}
            \end{itemize}

        \item Relate the history of Anti-Semitism
            \begin{itemize}
                \item Begins with book of Ester - Persians attempt to anihillate Jewish people
                \item Theodosionus (Middle Ages) declares Judaism to be illegal (no conversion, Jews can't hold office)
                \item John Chrysostom - Jews are ``Christ Killers"
                \item 1096: Crusades kill many Jews, and church condemns them for usuary
                \item Jews blamed for Black Plague (poisoning wells)
                \item Many killed during Spanish inquisition
                \item Martin Luther - Just before death, writes ``On The Jews and Their Lies"
                \item Voltaire - Hated the church and the Jews (background of Church)
                \item Germany adopts influential anti-semetic ideas
                \item Hitler blames Jewish people for defeat in WWI (Jews wouldn't finance war)
                \item Holocaust of Jewish people during WWII
            \end{itemize}
    \end{enumerate}
\end{document}

### 2016-09-29 abstract_algebra/test_review.tex
updatearoo

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\newcommand{\Group}{\mathcal{G}}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}

\begin{document}
    \theoremstyle{definition}
    \begin{definition}
        A \underline{Binary Operation} on a set $\mathcal{S}$ is a mapping $*$ from
        $\mathcal{S}\times\mathcal{S}\to\mathcal{S}$ that associates elements in $\mathcal{S}\times\mathcal{S}$ to an
        element in $\mathcal{S}$.
    \end{definition}

    \theoremstyle{definition}
    \begin{definition}
        A \underline{Group} is a pairing of a set $\Group$ and a binary operator $*$ $(\Group,*)$ such that
        \begin{enumerate}
            \item $*$ is associative, i.e. $a*(b*c) = (a*b)*c$.
            \item $\exists e\in\Group \backepsilon x*e=e*x=x$. This is called the identity.
            \item $\forall x\in\Group, \exists y\in\Group \backepsilon x*y = y*x =e$. $y$ is called the inverse of $x$.
        \end{enumerate}
        If $*$ is commutative, then $\Group$ is called \underline{abelian}.
    \end{definition}

    \begin{theorem}
        Let $*$ be a binary operation on $\mathcal{S}$. If there is an identity, it is unique.
    \end{theorem}
    \begin{proof}
        Let $e,f$ be identities in $(\mathcal{S},*)$. Then $e*f=e$ and $e*f=f$, so $e=f$.
    \end{proof}

    \begin{theorem}
        If $\Group$ is a group, then the inverse of an element is unique.
    \end{theorem}
    \begin{proof}
        Let $a\in\Group$ with $a''$ and $a'$    
    \end{proof}

\end{document}

### 2016-09-29 cs_foundations/box.py

@@ -1,0 +1,6 @@
+#sTeven Rosendahl
+
+def drawRow():
+    print("***")
+for x in range(0,3):
+    drawRow()

### 2016-09-29 cs_foundations/grade.py

@@ -1,0 +1,13 @@
+#Steven Rosendahl
+
+def getGrade(num):
+    if num <= 50:
+        return 'F'
+    else:
+        letter = 'F'
+        for x in range(50,num):
+            if x % 10 == 0:
+                letter = chr(ord(letter) - 1)
+        return letter
+print("Enter your number grade:")
+print("Your Grade is %s" % getGrade(int(input())))

### 2016-09-29 cs_foundations/lettertonbr.py

@@ -1,0 +1,15 @@
+#Steven Rosendahl
+#Partner: python.org
+
+diction = {'A':4.0, 'A-':3.7, 'B+':3.3, 'B':3.0, 'B-':2.7, 'C+':2.3, 'C':2.0, 'C-':1.7, 'D+':1.3, 'D':1.0, 'D-':0.7, 'F':0.3}
+
+scores = []
+i = 0
+total = 0
+while i < 5:
+    print("Enter your letter grade:")
+    scores.append(input())
+    i = i + 1
+for score in scores:
+    total = total + diction[score.upper()]
+print("Your GPA is %.2f" % (total / (i)))

### 2016-09-29 cs_foundations/lettertonbr_tk.py

@@ -1,0 +1,31 @@
+from tkinter import Tk, RIGHT, BOTH, RAISED
+from tkinter.ttk import Frame, Style, Button
+
+class GPACalc(Frame):
+
+    def __init__(self, parent):
+        Frame.__init__(self, parent)
+        self.parent = parent
+        self.initUI()
+
+    def onPressCalc():
+        print("Hello world")
+
+    def initUI(self):
+        self.parent.title("GPA Calculator")
+        self.style = Style()
+        self.style.theme_use("default")
+
+        frame = Frame(self, relief=RAISED, borderwidth=1)
+        frame.pack(fill=BOTH, expand=True)
+        self.pack(fill=BOTH, expand=True)
+
+        calculateButton = Button(self, text="Calculate", command=lambda: self.onPressCalc)
+        calculateButton.pack(side=RIGHT, padx=5, pady=5)
+
+def main():
+    root = Tk()
+    root.geometry("300x200+300+300")
+    app = GPACalc(root)
+    root.mainloop()
+main()

### 2016-09-29 cs_foundations/printnbrs.py

@@ -1,0 +1,8 @@
+#Steven Rosendahl
+
+print("Enter a postivie integer")
+num = int(input())
+i = 0
+while i <= num:
+    print(i)
+    i = i + 1

### 2016-09-29 cs_foundations/smallest.py

@@ -1,0 +1,6 @@
+#Steven Rosendahl
+
+def getSmall(a,b,c):
+    return min(a,b,c)
+print("Enter 3 ints")
+print("The smallest number is %d" % getSmall(int(input()),int(input()),int(input())))

### 2016-10-08 religions_of_the_west/paper.tex
working on religions paper

@@
-\documentclass[12pt]{hw}
+\documentclass[12pt]{article}
 
@@
 Taoism is unique in that it combines religion, philosophy, and myth.
-the Tao is considered to be the ultimate goal of Taoism, and texts such as the Tao Te Ching and the Chuang-tzu were written to be moral guides that teach the way to reach the Tao, while much of the history of Taoism involves tales of magic and immortality that are widely considered to be myth.
-The history of has been shaped by China just as much as Taoism has shaped the history of China.
+the Tao is considered to be the ultimate goal of Taoism, and texts such as the Tao Te Ching and the Chuang-tzu were written to be moral guides that teach the way to reach the Tao.
+Much of the history of Taoism involves tales of magic and immortality that are widely considered to be myth, and the religion does not focus on God in the way western religions do.
+The history of Taoism has been shaped by China just as much as Taoism has shaped the history of China.
 
@@
 During this period of turmoil, Laozi decided to leave the civilized area of China and live as a hermit in the wild.
-It was during this time that he wrote the Tao Teh Ching while trying to leave China.
-According to legend, Laozi was exiting China through the western gate when a soldier stopped him.
+It was during this time that he wrote the Tao Teh Ching.
+According to legend, Laozi was leaving China through the western gate when a soldier stopped him.
 The soldier felt that Laozi had much to teach, and refused to let him exit until he had written his teachings down.
@@
 The Tao Teh Ching, a set of eighty-one cryptic verses, is considered to be the most popular text in Taoism.
-Many of the verses focus on how to better one's self, so that they may follow the Tao, which literally translates into ``Path".
+Many of the verses focus on how to better one's self, so that they may follow the Tao, which literally translates into \textit{Path}.
 Several very famous saying have originated from the Tao Teh Ching, such as:
@@
 But it is these empty spaces that make the room livable.\\
-\textbf{$-$ Laozi, Tao Teh Ching, Ch. 11}\\\\
+\textbf{$-$ Laozi, Tao Teh Ching, Ch. 11}\\
 He who knows does not speak.\\
@@
 The ideas in the Tao Teh Ching reflect Laozi's feelings towards the state of China at the time.
-The above quotation emphasizes the usefulness of emptiness, while many of the feudal lords sought to have as much power and land as possible.
+The above quotation emphasizes the usefulness of emptiness during a time in which many of the feudal lords sought to have as much power and land as possible.
 The beginning lines of chapter 29 emphasize that mankind is not the ruler of the nature.
@@
 \end{quote}
-Laozi's writings primarily focus on the importance of nature and the temporality of life; he believed that power and possessions were not what the people of China should be focused on.
+Laozi's writings primarily focus on the importance of nature and the temporariness of life; he believed that power and possessions were not what the people of China should be focused on.
 The Tao Teh Ching is not the only sacred writing in the Taoist canon, however.
@@
 There is no one deity to which Taoists must bow; instead, they simply must seek out the Tao.
-In the Tao Teh Ching, Laozi describes the Tao by saying ``I know not its name, so I style it the way.
-I give it the makeshift name of the great." Chuang-tzu described the Tao as an force that ``gave spirits and rulers their spiritual powers." Interestingly, the Tao, while being the creator of all, is not immediately present in the world.
+In the Tao Teh Ching, Laozi describes the Tao by saying
+``I know not its name, so I style it the way. I give it the makeshift name of the great."
+Chuang-tzu described the Tao as an force that
+``gave spirits and rulers their spiritual powers."
+Interestingly, the Tao, while being the creator of all, is not immediately present in the world.
 While many religions present gods who involve themselves in mankind, the Tao is considered to just be; it does not impose itself on humanity, since, according to Chuang-tzu, humanity is too petty for it.
-Laozi explains this behavior by saying that ``The Tao nurtures her children by allowing them to be themselves, so that when they are grown, they will be able to stand on their own." In other words, the Tao creates everything to be independent, without forcing its creations to follow it.
+Laozi explains this behavior by saying that ``The Tao nurtures her children by allowing them to be themselves, so that when they are grown, they will be able to stand on their own."
+In other words, the Tao creates everything to be independent, without forcing its creations to follow it.
 
@@
 Many political leaders of China followed this principal and would attempt to please the non-human world in order to bring good fortune to their lives.
-The Tao is also considered to be ``self-balancing", meaning that if the Tao was brought off balance, it would bring balance back by whatever means necessary.
+The Tao is also considered to be \textit{self-balancing}, meaning that if the Tao was brought off balance, it would bring balance back by whatever means necessary.
 To avoid drastic changes in balance, followers of Taoism would seek out balance in life, and avoid doing anything considered to be extreme.
-The ideas of nature and balance are reflected in the teaching of the Five Elements.
+%The ideas of nature and balance are reflected in the teaching of the Five Elements.
 
-The Five Elements are a combination of balance, nature, and humankind.
+%The Five Elements are a combination of balance, nature, and humankind.
 They are Wood, Fire, Earth, Metal, and Water, and each one is related to every other element.
@@
 While the Yin-Yang symbol shows the duality of actions, Taoist belief states that it is not always evenly divided into one side or the other.
+Taoism is a uniquely different religion from many of the religions of the west and the east.
+The Tao is not a God in the traditional sense; it is surrounded in myth, and does not interfere with the lives of humans.
+Humans can become closer to the Tao by searching for peace and balance in their lives.
+Taoism is not a religion of rules and guidelines, but rather a religion that focuses on bettering one's self, which will in turn lead to a better society.
+In this way, Taoism can almost be considered a philosophy by which to conduct one's self.
 

### 2016-10-13 padics/hw4.tex
paper and padics

\documentclass{hw}
\begin{document}
\makeheader{4}

\begin{enumerate}
    \item Recall that $f(x)=x^2 + 1$ has two zeros $\alpha,\beta\in\ZZ_{5}$ satisfying $|\alpha-2|_{5}\leq 1/5$ and
    $|\beta-3|_{5}\leq 1/5$. Find the first three terms $b_{0}, b_{1}$, and $b_{3}$ of the \textit{p-adic} expansion
    of $\beta$.

    \item User Hensel's Lemma to verify that $f(x)=x^3+1$ has a zero $\alpha\in\ZZ_{7}$. Find an integer $n$ such that
    $|\alpha-n|_{7} \leq 1/7$.
    \begin{quote}
        If we let $\alpha=3$, then $f(\alpha)=27+1\equiv 0\mod 7$. We also have that
        $f'(\alpha)=3(9)=27\not\equiv0\mod7$, so Hensel's Lemma tells us that there is a zero of the function. 
    \end{quote}

    \item Show that Hensel's Lemma fails to apply to the polynomial $f(x)=x^3+1$ in $\ZZ_{3}$. Is this sufficient to
    conclude that $f$ has no zeros in $\ZZ_{5}$?
    \begin{quote}
        Hensel's Lemma requires that $f'(x)\not\equiv 0 \mod p$. In this case, $f'(x)=3x^{2}$, and
        $3x^{2}\equiv0\mod 3$, so we cannot use it. This is not enough to say that $f$ has no zeros. Hensel's Lemma only tells us about the properties of a zero, but it does not guarantee the existence of a zero.
    \end{quote}
\end{enumerate}
\end{document}

### 2016-10-15 resume/education.tex
working on elliptic
starting new semester
updated resume

@@
 \begin{enumerate}[label={\Roman*}.]
-\item \textbf{Christopher Newport University $\cdot$ 2013 - Present}\\
-Bachelors of Science $-$ Mathematics $\cdot$ Bachelors of Science $-$ Computer Science
+	\item \textbf{Christopher Newport University $\Bigcdot$ August 2013 - May 2017}\\
+	      Bachelors of Science $-$ Mathematics $\cdot$ Bachelors of Science $-$ Computer Science $\cdot$ 3.67 GPA
 	      \begin{itemize}
@@
 	      	      	\item Served as Secretary for the Society of Physics club
-\item Earned Dean's List consistently
+	      	      	\item Served as Secretary for Math club
+	      	      	\item Earned Dean's List since Fall 2013
 	      	      \end{itemize}
-% \newpage
 	      	\item Relevant coursework includes
 	      	      \begin{itemize}
-\item Mathematical Education
+	      	      	\item Mathematics Education
 	      	      	      \begin{quote}
-Calculus I - IV, Linear Algebra, Real Analysis, Partial and Ordinary Differential Equations, Writing in
-Mathematics using \hologo{LaTeX}, Proofs and Discreet Mathematics, Number Theory, Independent Study in
-Partial Differential Equations
+	      	      	      	\def\arraystretch{1.1}
+	      	      	      	\begin{tabular}{l l l}
+	      	      	      		Calculus I - IV                            & Linear Algebra         & Real Analysis                         \\
+	      	      	      		Group Theory                               & Differential Equations & Technical Writing with \hologo{LaTeX} \\
+	      	      	      		Proofs and Discreet Mathematics            & Number Theory          & \textit{p-adic} Analysis              \\
+	      	      	      		Research in Partial Differential Equations & Mathematical Modeling  & Programming in Mathematica
+	      	      	      	\end{tabular}
 	      	      	      \end{quote}
@@
 	      	      	      \begin{quote}
-UNIX Operating Systems, C++ Programming, Java Programming, Android Programming, Programming Language
-Concepts, Data Structures, Algorithms, Computer Hardware and Design, Discreet Mathematics
+	      	      	      	\def\arraystretch{1.1}
+	      	      	      	\begin{tabular}{l l l}
+	      	      	      		UNIX Operating Systems & C++ Programming & Java Programming              \\
+	      	      	      		Android Programming    & Software Design & Programming Language Concepts \\
+	      	      	      		Data Structures        & Algorithms      & Computer Hardware and Design  \\
+	      	      	      		Discreet Mathematics & Writing and Computer Ethics
+	      	      	      	\end{tabular}
+	      	      	      \end{quote}
+	      	      	\item Programming Language and Framework Experience
+	      	      	      \begin{quote}
+	      	      	      	\def\arraystretch{1.1}
+	      	      	      	\begin{tabular}{c  c  c}
+	      	      	      		\textbf{Proficient} & \textbf{Advanced}     & \textbf{Intermediate} \\
+	      	      	      		\hline
+	      	      	      		Java/Android        & PHP                   & C\#/Unity             \\
+	      	      	      		\LaTeX              & Javascript/Angular.js & Swift/iOS             \\
+	      	      	      		Ruby                & C/C++                 & Objective-C           \\
+	      	      	      		Lua                 & Python                &                       \\
+	      	      	      		                    & Mathematica           &
+	      	      	      	\end{tabular}
 	      	      	      \end{quote}

### 2016-10-15 resume/work.tex
updated resume

@@
 
-\item MyNewLeaf, LLC $\Bigcdot$ January 2016 - Present
+\item \textbf{MyNewLeaf, LLC $\Bigcdot$ January 2016 - Present}
 \begin{itemize}
@@
 \item Served as lead programmer for Android version of application
+\item Contributed to UI design process
 \end{itemize}
 
-\item National Aeronautics and Space Administration (NASA) $\Bigcdot$ November 2015 - Present
+\item \textbf{National Aeronautics and Space Administration (NASA) $\Bigcdot$ November 2015 - Present}
 \begin{itemize}
 \item Developed Android applications for Google Glass, Vuzix M100, and Recon Jet headsets
-\item Worked closely with a university professor to plan and create the application
+\item Worked closely with a university professor to plan, create, and design the application
+\item Created universal application for a wide range of Android powered devices
 \item Participated in laser safety training and received laser operator certification
@@
 
-\item Application Developer $\Bigcdot$ July 2015 - Present
-\begin{itemize}
-\item Built a location based audio application called Audionce, written for Android and iOS
-\item Created a quiz web-application called WhensIt for the Public History Center at Christopher Newport
-University
-\item Built websites using JavaScript, HTML5, CSS3, and PHP
-\end{itemize}
+% \item \textbf{Application Developer $\Bigcdot$ July 2015 - Present}
+% \begin{itemize}
+% \item Built a location based audio application called Audionce, written for Android and iOS
+% \item Created a quiz web-application called WhensIt for the Public History Center at Christopher Newport
+% University
+% \item Built websites using JavaScript, HTML5, CSS3, and PHP
+% \end{itemize}
 
-\item Christopher Newport University $\Bigcdot$ September 2013 - Present
+\item \textbf{Christopher Newport University $\Bigcdot$ September 2014 - Present}
 \begin{itemize}
@@
 \begin{itemize}
-\item Tutored several subjects including Mathematics (mostly Calculus), Computer Engineering, and
-Computer Science
+\item Tutored several subjects including Mathematics, Computer Engineering, and Computer Science
 \item Participated in both one on one tutoring and group tutoring
-\item Created various resources for other Calculus tutors to use
+\item Created various resources for other tutors
+\item Participated in group training exercises to receive tutor certification 
 \end{itemize}
@@
 \newpage
-\item Kirk Family YMCA $\Bigcdot$ August 2009 - August 2015
+\item \textbf{Kirk Family YMCA $\Bigcdot$ August 2009 - August 2015}
 \begin{itemize}

### 2016-10-20 ode_two/paper.tex
finished hw

\documentclass{article}
\usepackage{amsmath, amssymb}
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\usepackage{graphicx}

\doublespacing

\title{The R\"{o}ssler System}
\author{Steven Rosendahl}
\date{}

\begin{document}
\maketitle

In 1976, Otto R\"{o}ssler proposed a system of nonlinear ordinary differential equations that illustrated the simplest possible strange attractor. An attractor is a set of values towards which a system moves when initial conditions are \textit{near} the attractor; to call an attractor strange means that the attractor exhibits fractal behavior. Often, strange attractors are associated with chaotic systems. R\"{o}ssler's strange attractor is a chaotic attractor that solves his proposed system
\begin{align}
    \dot{x} &= -y-z\\
    \dot{y} &= x+ay\\
    \dot{z} &= b+z(x-c)
\end{align}
where $a,b,$ and $c$ are arbitrary values. R\"{o}ssler studied the effects that small (i.e. less than 1) $a$ and $b$ paired with a relatively large $c$ had on the system's chaotic behavior.

\end{document}

### 2016-10-20 padics/hw4.tex
finished hw

@@
     \item Recall that $f(x)=x^2 + 1$ has two zeros $\alpha,\beta\in\ZZ_{5}$ satisfying $|\alpha-2|_{5}\leq 1/5$ and
-    $|\beta-3|_{5}\leq 1/5$. Find the first three terms $b_{0}, b_{1}$, and $b_{3}$ of the \textit{p-adic} expansion
+    $|\beta-3|_{5}\leq 1/5$. Find the first three terms $b_{0}, b_{1}$, and $b_{2}$ of the \textit{p-adic} expansion
     of $\beta$.
+    \begin{quote}
+        Since $b=3$ satisfies $f(x)=0$, we know $b_{0}=3$. We can consider $\beta$ of the form
+        \[
+            \beta = b_{0} + 5b_{1} + 5^{2}b_{2} + \dots,
+        \]
+        where $0\leq b_{n} \leq 5$. To find $b_{1}$, we look at $f(\beta)\equiv 0\mod 5^{2}$. Then
+        \begin{gather*}
+            (3+5b_{1})^{2} + 1 \equiv 0\mod 25\\
+            9 + 5b_{1} + 25b_{2}^{2} + 1 \equiv 0\mod 25\\
+            10 + 5b_{1} \equiv 0\mod 25\\
+            5b_{1}\equiv 15\mod 25\\
+            \implies b_{1}=3.
+        \end{gather*}
+        Similarly, for $b_{2}$, we have
+        \begin{gather*}
+            (18+25b_{2})^{2} + 1 \equiv 0\mod 125\\
+            74 + 25b_{2} + 1 \equiv 0 \mod 125\\
+            25b_{2} \equiv 50\mod 125\\
+            5b_{2}\equiv 10\mod 25\\
+            \implies b_{2} = 2.
+        \end{gather*}
+        Therefore $\beta \approx 3 + 5\cdot3 + 5^{2}\cdot2$.
+    \end{quote}
 
@@
         If we let $\alpha=3$, then $f(\alpha)=27+1\equiv 0\mod 7$. We also have that
-        $f'(\alpha)=3(9)=27\not\equiv0\mod7$, so Hensel's Lemma tells us that there is a zero of the function. 
+        $f'(\alpha)=3(9)=27\not\equiv0\mod7$, so Hensel's Lemma tells us that there is a zero of the function. Let $n  10$. Then $|3-10|_{7}=|-7|_{7}=|7|_{7}=1/7 \leq 1/7$.
     \end{quote}
@@
     \item Show that Hensel's Lemma fails to apply to the polynomial $f(x)=x^3+1$ in $\ZZ_{3}$. Is this sufficient to
-    conclude that $f$ has no zeros in $\ZZ_{5}$?
+    conclude that $f$ has no zeros in $\ZZ_{3}$?
     \begin{quote}

### 2016-10-25 ode_two/paper.tex
update paper

@@
 \usepackage{graphicx}
+\usepackage{multicol}
+\usepackage[backend=bibtex]{biblatex}
+
+\addbibresource{mbib.bib}
 
@@
 In 1976, Otto R\"{o}ssler proposed a system of nonlinear ordinary differential equations that illustrated the simplest possible strange attractor.
-An attractor is a set of values towards which a system moves when initial conditions are \textit{near} the attractor; to call an attractor strange means that the attractor exhibits fractal behavior.
+An attractor is a set of values towards which a system moves when initial conditions are \textit{near} the attractor; to call an attractor \textit{strange} means that the attractor exhibits fractal behavior.
 Often, strange attractors are associated with chaotic systems.
@@
 \begin{align}
-    \dot{x} &= -y-z\\
-    \dot{y} &= x+ay\\
-    \dot{z} &= b+z(x-c)
+	f(x,y,z) = \dot{x} & = -y-z     \\
+	g(x,y,z) = \dot{y} & = x+ay     \\
+	h(x,y,z) = \dot{z} & = b+z(x-c)
 \end{align}
-where $a,b,$ and $c$ are arbitrary values. R\"{o}ssler studied the effects that small (i.e. less than 1) $a$ and $b$ paired with a relatively large $c$ had on the system's chaotic behavior.
+where $a,b,$ and $c$ are arbitrary values.
+R\"{o}ssler studied the effects that small (i.e.
+less than 1) $a$ and $b$ paired with a relatively large $c$ had on the system's chaotic behavior.
+One such plot is shown in figure \ref{fig:3dsys_01}.
+
+We first want to analyze the stability of the critical points of the system. The Jacobian for this system can be found by constructing a matrix of partial derivatives for $f,g,$ and $h$.
+\[
+	J(x,y,z)=
+	\left[
+	\begin{array}{c c c}
+		\frac{\partial f}{\partial x} & \frac{\partial f}{\partial y} & \frac{\partial f}{\partial z} \\
+		\frac{\partial g}{\partial x} & \frac{\partial g}{\partial y} & \frac{\partial g}{\partial z} \\
+		\frac{\partial h}{\partial x} & \frac{\partial h}{\partial y} & \frac{\partial h}{\partial z} \\
+	\end{array}
+	\right]
+	=
+	\left[
+	\begin{array}{c c c}
+		0 & -1 & -1  \\
+		1 & a  & 0   \\
+		z & 0  & x-c
+	\end{array}
+	\right].
+\]
+We now need to find the critical points of the system. We can accomplish this by solving $f=0,g=0,$ and $h=0$. Through the use of Mathematica's \texttt{Solve} command, we get
+\begin{center}
+	\texttt{Solve[f[x, y, z] == 0 \&\& g[x, y, z] == 0 \&\& h[x, y, z] == 0, \{x, y, z\}]}
+\end{center}
+\begin{gather}
+	\left[
+	x=\frac{1}{2}\left(c-\sqrt{c^{2}-4ab}\right),
+	y=\frac{1}{2}\left(\frac{\sqrt{c^{2}-4ab}}{a}-\frac{c}{a}\right),
+	z=\frac{c-\sqrt{c^{2}-4ab}}{2a}
+	\right]\label{eq:csln1}\\
+	\left[
+	x=\frac{1}{2}\left(c+\sqrt{c^{2}-4ab}\right),
+	y=\frac{1}{2}\left(-\frac{\sqrt{c^{2}-4ab}}{a}-\frac{c}{a}\right),
+	z=\frac{c+\sqrt{c^{2}-4ab}}{2a}
+	\right]\label{eq:csln2}.
+\end{gather}
+
+R\"{o}ssler's analysis focused on the behavior of the system for small $a$ and $b$.
+For simplicity, we can consider the solutions to (\ref{eq:csln1}) and (\ref{eq:csln2}) for very small $a$ and $b$.
+We now have
+\begin{gather}
+	\left[
+	x=0,
+	y=0,
+	z=0
+	\right]\label{eq:csln1sm}\\
+	\left[
+	x=c,
+	y=-\frac{c}{a},
+	z=\frac{c}{a}
+	\right]\label{eq:csln2sm}.
+\end{gather}
+To find the behavior around the critical points given by (\ref{eq:csln1sm}), we evaluate the Jacobian at the appropriate points and find the eigenvalues of the system.
+\[
+	|J(0,0,0)-\lambda I| =
+	\left|
+	\left[
+	\begin{array}{c c c}
+		-\lambda & -1       & -1         \\
+		1        & -\lambda & 0          \\
+		0        & 0        & -c-\lambda
+	\end{array}
+	\right]
+	\right| =
+	(c-\lambda)(\lambda-1)(\lambda + 1)=0.
+\]
+Hence, the system has three eigenvalues: $\lambda=1, \lambda=-1,$ and $\lambda = c$ (we will assume here that $c > 1$).
+This implies that the system has an unstable saddle at the origin \cite{equilibria}.
+Figure \ref{fig:plot_cuts} shows various cuts of the system for various $z$ values; the system spirals in the $x-y$ plane, while moving towards and then away from the origin as $z$ moves from $-\infty$ to $\infty$.
+This represents only one of the equilibrium points of the system, however.
+We now need to analyze the linearized system near (\ref{eq:csln2sm}).
+\[
+	\left|J\left(c,-\frac{c}{a},\frac{c}{a}\right) -\lambda I\right| =
+	\left|
+	\left[
+	\begin{array}{c c c}
+		-\lambda    & -1          & -1       \\
+		1           & a - \lambda & 0        \\
+		\frac{c}{a} & 0           & -\lambda
+	\end{array}
+	\right]
+	\right| =
+	\frac{ ac - a\lambda - c\lambda + a^{2}\lambda^{2} -a\lambda^{3}}{a}.
+\]
+Setting the determinant to $0$ and solving in Mathematica yields three distinct (and incredibly verbose) eigenvalues with $\lambda_{1}\in\mathbb{R}$ the other two complex conjugates of each other.
+Such behavior lends itself towards what is known as a \textit{saddle focus}, which is always unstable \cite{equilibria}.
+Figure \ref{fig:3dsys_02} displays the behavior of both equilibrium points.
+The spiral behavior can be seen in the $x-y$ plane, and the saddle behavior can be seen predominantly in the $z-x$ plane.
+
+We now know that both equilibrium points are unstable when we ignore $a$ and $b$ in comparison to $c$.
+In fact, it turns out that this stability is preserved even when we reintroduce $a$ and $b$ back into the system.
+Figure \ref{fig:raw_ev} shows us what one of the raw eigenvalues looks like, according to Mathematica.
+The three eigenvalues resulting from the unsimplified system still exhibit the saddle focus behavior that we saw when we allowed $a$ and $b$ to drift towards zero.
+We now want to turn our attention towards the actions of the system when we change the values for $a$, $b$, and $c$.
+
+Attractors are often associated with chaotic systems; as such, we will now examine various bifurcations on the R\"{o}ssler system.
+We have three variables that contribute to the chaos of the system: $a$, $b$, and $c$.
+R\"{o}ssler analyzed the system for relatively small $a$ and $b$, and we analyzed the stability of critical points when we neglected $a$ and $b$.
+We will first analyze the effect that small changes to $a$ have on the system.
+
+The $a$ parameter affects our $g(x,y,z)$. For simplicity, we notice that $g$ is only dependent on $x$ and $y$, so we will let $g(x,y,z)=\gamma(x,y)$.
+
+\newpage
+\clearpage
+\section*{Plots and Graphics}
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.5]{vp_01.png}
+	\caption{The R\"{o}ssler system with $a=0.2$, $b=0.1$, and $c=2.3$}
+	\label{fig:3dsys_01}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.7]{attractor_a0p14_b0p2_c8p8}
+	\caption{The R\"{o}ssler attractor with $a=0.14$, $b=0.2$, and $c=8.8$, courtesy of \cite{rossler_wf}}
+	\label{fig:3dsys_02}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.5]{messy_output}
+	\begin{gather*}
+		\{l\to -4.98425\}\\
+		\{l\to 0.0961307\, -0.995341 i\}\\
+		\{l\to 0.0961307\, +0.995341 i\}
+	\end{gather*}
+	\caption{Example of one raw eigenvalue and its corresponding evaluation at $a=0.2$, $b=0.2$, and $c=0.2$.}
+	\label{fig:raw_ev}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+	\begin{tabular}{c c}
+		\includegraphics[scale=0.5]{spz-2} & \includegraphics[scale=0.5]{spz-1} \\
+		$z=-2$                             & $z=-1$                             \\
+		\multicolumn{2}{c}{\includegraphics[scale=0.5]{spz0}}\\
+		\multicolumn{2}{c}{$z=0$}\\
+		\includegraphics[scale=0.5]{spz1}  & \includegraphics[scale=0.5]{spz2}  \\
+		$z=1$                              & $z=2$
+	\end{tabular}
+	\caption{Several cuts of the stream plot given by Mathematica.}
+	\label{fig:plot_cuts}
+\end{figure}
+
+\newpage
+\clearpage
+\nocite{*}
+\printbibliography
 

### 2016-10-25 religions_of_the_west/paper.tex
paper and padics
update paper

@@
 \usepackage[margin=1in]{geometry}
+\usepackage[american]{babel}
+\usepackage{hyperref}
+\usepackage[backend=bibtex,style=mla,showmedium=false]{biblatex}
 
@@
 \date{}
+
+\addbibresource{paper_bib.bib}
 
@@
 The Tao is also considered to be \textit{self-balancing}, meaning that if the Tao was brought off balance, it would bring balance back by whatever means necessary.
+This idea lends itself back to the idea that the Tao is not concerned with human affairs.
 To avoid drastic changes in balance, followers of Taoism would seek out balance in life, and avoid doing anything considered to be extreme.
-%The ideas of nature and balance are reflected in the teaching of the Five Elements.
 
-%The Five Elements are a combination of balance, nature, and humankind.
-They are Wood, Fire, Earth, Metal, and Water, and each one is related to every other element.
-For example, Fire can destroy Metal, help the Earth, gain strength from Wood, and be weakened by Water.
-While the Yin-Yang symbol shows the duality of actions, Taoist belief states that it is not always evenly divided into one side or the other.
 Taoism is a uniquely different religion from many of the religions of the west and the east.
 The Tao is not a God in the traditional sense; it is surrounded in myth, and does not interfere with the lives of humans.
-Humans can become closer to the Tao by searching for peace and balance in their lives.
+Humans can become closer to the Tao by searching for peace and balance in their lives by analyzing and following the nature of the Tao.
 Taoism is not a religion of rules and guidelines, but rather a religion that focuses on bettering one's self, which will in turn lead to a better society.
+It does not have ``holy books" like many other religions, but rather abstract sayings that are free to interpretation.
 In this way, Taoism can almost be considered a philosophy by which to conduct one's self.
 
+\newpage
+
+\nocite{*}
+\printbibliography
+
 \end{document}

### 2016-10-28 padics/hw5.tex
padics

\documentclass{hw}

\begin{document}
\makeheader{5}
\begin{enumerate}
	\item Let $f(x,y)=x^2 + y^2$. Observe that $(x,y)=(0,0)$ defines a solution to the equation $f(x,y)=0$. Do the remainder of this problem, we shall call $(0,0)$ the trivial solution to $f(x,y)=0$.
	      \begin{enumerate}
	      	\item Prove that $f(x,y)=0$ has no non-trivial solutions with $x,y\in\QQ$.
	      	\item Prove that there exists a prime $p$ such that $f(x,y)=0$ has a nontrivial solution with $x,y\in\QQ_{p}$.
	      	\item Verify that $f$ satisfies the Hasse Principle without referring to the Hasse-Minkowski Theorem.
	      \end{enumerate}

	\item Suppose that $\{ a_{n} \}_{n=0}^{\infty}$ is a sequence of points in $\ZZ_{p}$. Prove that the series
	      \[
	      	\sum_{n=0}^{\infty}a_{n}p^{n}
	      \]
	      converges in $\QQ_{p}$.
	      \begin{quote}
	      		\textit{\textbf{Proof:}} We know that a series will converge if and only if its sequence of partial sums converges. We can test for convergence by calculating
				\[
					\lim_{n\to\infty}\left|a_{n+1}p^{n+1}-a_{n}p^{n}\right|_{p}.
				\]
				We can say
				\begin{align*}
					\left|a_{n+1}p^{n+1}-a_{n}p^{n}\right|_{p} &= \max\left\{%
						\left|a_{n+1}p^{n+1}\right|_{p},\left|a_{n}p^{n}\right|_{p}%
					\right\}\\
					&= \max\left\{%
						\frac{1}{p^{v_{p}(a_{n+1}) + n + 1}}, \frac{1}{p^{n+v_{p}(a_{n})}}%
					\right\}.\\
					\lim_{n\to\infty}\max\left\{%
						\frac{1}{p^{v_{p}(a_{n+1}) + n + 1}}, \frac{1}{p^{n+v_{p}(a_{n})}}%
					\right\} &= \max\left\{0,0\right\}\\
					&= 0.
				\end{align*}
				Therefore, the sequence of partial sums is Cauchy, and thus convergent, so the series converges.
	      \end{quote}

	\item Suppose that $k\in\NN$.
	      \begin{enumerate}
	      	\item Prove that the series
	      	      \[
	      	      	\sum_{n=1}^{\infty}\frac{1}{n^{k}}
	      	      \]
	      	      does not converge in $\QQ_{p}$ for any prime $p\neq\infty$.

	      	\item Prove that the series
	      	      \[
	      	      	\sum_{n=1}^{\infty}n^{k}
	      	      \]
	      	      does not converge in $\QQ_{p}$ for any prime $p\neq\infty$.
	      \end{enumerate}
\end{enumerate}
\end{document}

### 2016-10-28 software_design/test1/p3.py

@@ -1,0 +1,58 @@
+csvFile = open('mid2prob3_results.csv',"r")
+lc = 0
+
+colnames = []
+finalArray = []
+tmpNameArray = []
+
+# for line in csvFile:
+    # print line
+
+lc = 0
+
+for line in csvFile:
+    if lc < 2:
+        lc = lc + 1
+        continue
+    csved = line.split(',')
+    mArray = []
+    for val in csved:
+        mArray.append(val.strip())
+    finalArray.append(mArray)
+
+currRow = 0
+for array in finalArray:
+    count = 0
+    avg = 0.0
+    total = 0.0
+    wStr = ""
+    # if currRow < 2:
+    #     currRow = currRow + 1
+    #     continue
+    for val in array:
+        if count == 0 or count == 1:
+            wStr = wStr + " " + val
+        else:
+            if val != '':
+                total = total + float(val)
+            else:
+                total = total + 0
+        count = count + 1
+    print (wStr + " average: " + str(total / (count - 2))).strip()
+# print finalArray
+print("")
+print("Column Averages:")
+numRows = len(finalArray)
+numCols = len(finalArray[0])
+
+currRow = 0
+count = 0
+for i in range(2,numCols - 1):
+    total = 0.0
+    avg = 0.0
+    cName = "G" + str(i - 1)
+    for array in finalArray:
+        if array[i] != '':
+            total = total + float(array[i])
+        count = count + 1
+    print(cName + " " +str(total / count))

### 2016-10-30 ode_two/paper.tex
working on rossler system

@@
 	\right| =
-	(c-\lambda)(\lambda-1)(\lambda + 1)=0.
+	(c-\lambda)(\lambda + 1)^{2}=0.
 \]
-Hence, the system has three eigenvalues: $\lambda=1, \lambda=-1,$ and $\lambda = c$ (we will assume here that $c > 1$).
-This implies that the system has an unstable saddle at the origin \cite{equilibria}.
+Hence, the system has three eigenvalues: $\lambda=-i, \lambda=-i,$ and $\lambda = -c$ (we will assume here that $c > 1$).
+This implies that the system has an unstable focus node at the origin \cite{equilibria}.
 Figure \ref{fig:plot_cuts} shows various cuts of the system for various $z$ values; the system spirals in the $x-y$ plane, while moving towards and then away from the origin as $z$ moves from $-\infty$ to $\infty$.
+We can solve this linearized system by looking at the equation
+\[
+	\mathbf{\dot{x}} =
+	\left[
+	\begin{array}{c c c}
+		0 & -1 & -1 \\
+		1 & 0  & 0  \\
+		0 & 0  & -c
+	\end{array}
+	\right]
+	\mathbf{x}.
+\]
+Since we've already found the eigenvalues of the system, we can find the eigenvectors immediately.
+
+\begin{minipage}{0.3\textwidth}
+	\begin{gather*}
+		\text{\underline{$\lambda=-i$}}\\
+		\mathbf{\lambda_{1}}=\left[
+		\begin{array}{c}
+			-i \\
+			1  \\
+			0
+		\end{array}
+		\right].
+	\end{gather*}
+\end{minipage}
+\begin{minipage}{0.3\textwidth}
+	\begin{gather*}
+		\text{\underline{$\lambda=i$}}\\
+		\mathbf{\lambda_{2}}=\left[
+		\begin{array}{c}
+			i \\
+			1 \\
+			0
+		\end{array}
+		\right].
+	\end{gather*}
+\end{minipage}
+\begin{minipage}{0.3\textwidth}
+	\begin{gather*}
+		\text{\underline{$\lambda=-c$}}\\
+		\mathbf{\lambda_{3}}=\left[
+		\begin{array}{c}
+			\frac{c}{1+c^{2}}  \\
+			-\frac{1}{1+c^{2}} \\
+			1
+		\end{array}
+		\right].
+	\end{gather*}
+\end{minipage}
+
+\noindent\\
+\noindent Letting $\mathbf{\gamma}$ stand in for our constants, we get
+\[
+	\mathbf{x}(t) =
+	\gamma_{1}e^{-it}\left[
+	\begin{array}{c}
+		-i \\
+		1  \\
+		0
+	\end{array}
+	\right] +
+	\gamma_{2}e^{it}
+	\left[
+	\begin{array}{c}
+		i \\
+		1 \\
+		0
+	\end{array}
+	\right] +
+	\gamma_{3}e^{-ct}\left[
+	\begin{array}{c}
+		\frac{c}{1+c^{2}}  \\
+		-\frac{1}{1+c^{2}} \\
+		1
+	\end{array}
+	\right].
+\]
+Note that we choose to leave the $e^{it}$ in this form, since \textbf{${\lambda_{1}}$} and $\mathbf{\lambda_{2}}$ are linearly independent of each other.
+Figure \ref{fig:zero_sln} shows the behavior of this linearized system with two different sets of initial conditions.
+If we compare this to figure \ref{fig:3dsys_02}, we see similar behavior in the area of the graph with large $z$ values.
+
 This represents only one of the equilibrium points of the system, however.
 We now need to analyze the linearized system near (\ref{eq:csln2sm}).
+In this case, we cannot let $a$ tend towards zero as we did in (\ref{eq:csln1sm}) since our roots rely on division by $a$.
 \[
@@
 The spiral behavior can be seen in the $x-y$ plane, and the saddle behavior can be seen predominantly in the $z-x$ plane.
+
+Again, we can solve the linearized system and analyze the behavior as we did with the system centered around $(0,0,0)$.
+Due to the complexity of the eigenvalues of this system, we let Mathematica determine the behavior.
+The command
+\begin{center}
+	\texttt{Eigenvectors[jcbn[x, y, z] /. \{x -> c, y -> -(c/a), z -> (c/a)\}]}
+\end{center}
+yields an unfriendly expression involving the \texttt{Root} command; in this case we will want to give concrete values to $c$ and $a$.
+Figure \ref{fig:3dsys_02} involves the behavior of the system at $a=0.14$ and $c=8.8$, so we will examine the resulting eigenvalues at those points.
+\begin{center}
+	\texttt{Eigenvectors[jcbn[x, y, z] /. \{x -> c, y -> -(c/a), z -> (c/a)\}]/.\{a -> 0.14, c -> 8.8\}}
+\end{center}
+results in the following three eigenvectors:
+
+\begin{minipage}{0.3\textwidth}
+	\begin{gather*}
+		% \text{\underline{$\lambda=-i$}}\\
+		\mathbf{\lambda_{1}}=\left[
+		\begin{array}{c}
+			-0.0022 \\
+			-1.0003 \\
+			1.0000
+		\end{array}
+		\right]
+	\end{gather*}
+\end{minipage}
+\begin{minipage}{0.3\textwidth}
+	\begin{gather*}
+		% \text{\underline{$\lambda=i$}}\\
+		\mathbf{\lambda_{2}}=\left[
+		\begin{array}{c}
+			0.0000174 - 0.12713i \\
+			0.0159 + 0.00028i    \\
+			1
+		\end{array}
+		\right]
+	\end{gather*}
+\end{minipage}
+\begin{minipage}{0.3\textwidth}
+	\begin{gather*}
+		% \text{\underline{$\lambda=-c$}}\\
+		\mathbf{\lambda_{3}}=\left[
+		\begin{array}{c}
+			0.0000174 - 0.12713i \\
+			0.0159 + 0.00028i    \\
+			1
+		\end{array}
+		\right].
+	\end{gather*}
+\end{minipage}
+
+\noindent\\
+\noindent We now have enough to for a general solution with $\mathbf{\gamma}$ in place of our constants.
+\begin{gather*}
+	\mathbf{x}(t) =
+	\gamma_{1}e^{0.138t}\left[
+	\begin{array}{c}
+		-0.0022 \\
+		-1.0003 \\
+		1
+	\end{array}
+	\right] +
+	\gamma_{2}e^{(0.00109-7.99i)t}\left[
+	\begin{array}{c}
+		0.0000174 - 0.12713i \\
+		0.0159 + 0.00028i    \\
+		1
+	\end{array}
+	\right] +\\
+	\gamma_{3}e^{(0.00109+7.99i)t}\left[
+	\begin{array}{c}
+		0.0000174 + 0.12713i \\
+		0.0159 - 0.00028i    \\
+		1
+	\end{array}
+	\right].
+\end{gather*}
+Figure \ref{fig:nzero_sln} displays two parametric plots of the solution curves for different initial values.
+The curves exhibit a center-like behavior that looks like a spring.
+We can compare this to the actual curve seen in \ref{fig:3dsys_02}.
+This behavior is seen in the purple spiral in the $x-y$ plane of the graph.
 
@@
 	\label{fig:3dsys_02}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+	\begin{tabular}{c c}
+		\includegraphics[scale=0.5]{gen_sln_01_01} & \includegraphics[scale=0.5]{gen_sln_01_02} \\
+		$\gamma_{1}=-0.7$                          & $\gamma_{1}=-0.7$                          \\
+		$\gamma_{2}=-0.7$                          & $\gamma_{2}=-0.7$                          \\
+		$\gamma_{3}=-1$                            & $\gamma_{3}=1$                             \\
+	\end{tabular}
+	\caption{Two parametric plots for the linearized system around $(0,0,0)$ with different initial values}
+	\label{fig:zero_sln}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+	\begin{tabular}{l r}
+		\includegraphics[scale=0.5]{gen_sln_02_01} & \includegraphics[scale=0.5]{gen_sln_02_02} \\
+		$\gamma_{1}=1$                             & $\gamma_{1}=9$                             \\
+		$\gamma_{2}=1$                             & $\gamma_{2}=-3$                             \\
+		$\gamma_{3}=1$                            & $\gamma_{3}=-3$                             \\
+	\end{tabular}
+	\caption{Two parametric plots for the linearized system around $(c,-\frac{c}{a}, \frac{c}{a})$ with different initial values}
+	\label{fig:nzero_sln}
 \end{figure}

### 2016-10-30 padics/hw5.tex
working on rossler system

@@
 	      	      does not converge in $\QQ_{p}$ for any prime $p\neq\infty$.
+				  \begin{quote}
+					  \textit{\textbf{Proof:}} Suppose otherwise. Then
+					  \[
+					  	\lim_{n\to\infty}\left|%
+							\frac{1}{(n+1)^{k}}-\frac{1}{n^{k}}%
+						\right|_{p} = 0.
+					  \]
+					  We can analyze the $p$-adic absolute value inside the limit:
+					  \begin{align*}
+						  \left|\frac{1}{(n+1)^{k}}-\frac{1}{n^{k}}\right|_{p} &\leq
+						  \max\left\{%
+						  	p^{-v_{p}\left(1/(n+1)^{k}\right)}, p^{-v_{p}\left(1/n^{k}\right)}%
+						  \right\}
+					  \end{align*}
+					  We can solve for the valuation in the first term:
+					  \[
+					  	v_{p}\left(
+							\frac{1}{(n+1)^{k}}
+						\right) =
+						-\frac{
+							k\log{(n+1)}+\log{m}
+						}{
+							\log{p}
+						},
+					  \]
+					  where $m\nmid p$.
+If we take the limit as $n\to\infty$ of this expression, we see that it tends towards negative infinity, so the $p$-adic absolute value tends towards infinity.
+Similarly, we can evaluate the limit of $|1/n^{k}|_{p}$, and we will see that it too tends towards infinity.
+Therefore,
+					  \[
+					  	\lim_{n\to\infty}\left|%
+							\frac{1}{(n+1)^{k}}-\frac{1}{n^{k}}%
+						\right|_{p} \neq 0,
+					  \]
+					  so the series does not converge since its sequence of partial sums does not converge.
+				  \end{quote}
 
@@
 	      	      does not converge in $\QQ_{p}$ for any prime $p\neq\infty$.
+				  \begin{quote}
+					  \textit{\textbf{Proof:}} We want to show that
+					  \[
+					  	\lim_{n\to\infty}\left|(n+1)^{k} - n^{k}\right|_{p}\neq 0.
+					  \]
+					  We can use a similar trick as above:
+					  \begin{align*}
+						  \left|(n+1)^{k} - n^{k}\right|_{p} &\leq \max\left\{
+
+						  \right\}
+					  \end{align*}
+				  \end{quote}
 	      \end{enumerate}

### 2016-10-30 padics/hw5.tex
workign on padics

--- 2016-10-28 padics/hw5.tex
@@
 	      	\begin{align*}
-					\left|a_{n+1}p^{n+1}-a_{n}p^{n}\right|_{p} &= \max\left\{%
+	      		\left|a_{n+1}p^{n+1}-a_{n}p^{n}\right|_{p} &\leq \max\left\{%
 	      		\left|a_{n+1}p^{n+1}\right|_{p},\left|a_{n}p^{n}\right|_{p}%
@@
 	      		\frac{1}{p^{v_{p}(a_{n+1}) + n + 1}}, \frac{1}{p^{n+v_{p}(a_{n})}}%
-					\right\} &= \max\left\{0,0\right\}\\
-					&= 0.
+	      		\right\}                                   & = \max\left\{0,0\right\} \\
+	      		                                           & = 0.
 	      	\end{align*}
+			Since the $p$-adic absolute value is greater than or equal to 0, we know
+			\[
+			\lim_{n\to\infty}\left|a_{n+1}p^{n+1}-a_{n}p^{n}\right|_{p} = 0.
+			\]
 	      	Therefore, the sequence of partial sums is Cauchy, and thus convergent, so the series converges.
@@
 	      	      does not converge in $\QQ_{p}$ for any prime $p\neq\infty$.
+	      	      \begin{quote}
+	      	      	\textit{\textbf{Proof:}} If we can show that the sequence of partial sums does not converge, then the series will not converge. We want to evaluate (and eventually take the limit of)
+	      	      	\[
+	      	      		\left|
+	      	      		\frac{1}{(n+1)^{k}}-\frac{1}{n^{k}}
+	      	      		\right|_{p}.
+	      	      	\]
+	      	      	We know by definition that
+	      	      	\begin{align*}
+	      	      		\left|
+	      	      		\frac{1}{(n+1)^{k}}-\frac{1}{n^{k}}
+	      	      		\right|_{p} &=
+						\left|
+							\frac{n^{k}-(n+1)^{k}}{n^{k}(n+1)^{k}}
+						\right|_{p}\\
+	      	      		&= p^{-
+	      	      			\left[
+	      	      			v_{p}(n^{k}-(n+1)^{k}) - v_{p}(n^{k})-v_{p}((n+1)^{k})
+	      	      			\right]
+	      	      		}.
+	      	      	\end{align*}
+					Through some tedious calculation, we can solve for the valuation of each term in the exponent.
+					\begin{gather*}
+						v_{p}\left(n^k-(n+1)^k\right) =
+						\frac{
+							\log{(n^k-(n+1)^k) - \log(m_{0})}
+						}{
+							\log{(p)}
+						}\\
+						v_{p}(n^{k}) =
+						\frac{
+							k\log{(n)}-\log{(m_{1})}
+						}{\log{(p)}}\\
+						v_{p}((n+1)^k) =
+						\frac{
+							k\log{(n+1)}-\log{(m_{2})}
+						}{
+							\log{(p)}
+						},
+					\end{gather*}
+					where $m_{0},m_{1}$, and $m_{2}$ do not divide $p$. Combining these valuations as required yields
+					\[
+					-\left[v_{p}(n^{k}-(n+1)^{k}) - v_{p}(n^{k})-v_{p}((n+1)^{k})\right]=
+					\frac{
+					\log{\left(
+						\frac{n^{k}(n+1)^{k}}{n^k-(n+1)^k}
+					\right)}
+					}{
+						\log{p}
+					} + c
+					\]
+					where $c$ is some constant number independant of $n$. We can now evaluate the real valued limit as $n\to\infty$.
+					\[
+						\lim_{n\to\infty}\frac{
+						\log{\left(
+							\frac{n^{k}(n+1)^{k}}{n^k-(n+1)^k}
+						\right)}}{\log{p}} + c = c+\frac{1}{\log{p}}\lim_{n\to\infty}\log{\left(
+							\frac{n^{k}(n+1)^{k}}{n^k-(n+1)^k}
+						\right)}.
+					\]
+					The function on the inside of the limit is continuous for all $k\in\NN$, so we can determine the limit as follows:
+					\begin{align*}
+						\lim_{n\to\infty}\log{\left(
+							\frac{n^{k}(n+1)^{k}}{n^k-(n+1)^k}
+						\right)} &= L\\
+						e^{
+						\lim_{n\to\infty}\log{\left(
+							\frac{n^{k}(n+1)^{k}}{n^k-(n+1)^k}
+						\right)}
+						} &= e^{L}\\
+						\lim_{n\to\infty}e^{\log{\left(
+							\frac{n^{k}(n+1)^{k}}{n^k-(n+1)^k}
+						\right)}
+						} &= e^{L}\\
+						\lim_{n\to\infty}\left(
+							\frac{n^{k}(n+1)^{k}}{n^k-(n+1)^k}
+						\right) &= e^{L}.
+					\end{align*}
+					For the fraction inside the limit, notice that after multiplication, the highest order term in the numerator will be $n^{2k}$, while the highest order term in the denominator will be $n^{k}$, so the limit behaves like
+					\[
+					\lim_{n\to\infty}\frac{n^{2k}}{n^k}=\lim_{n\to\infty}n^{k}=\infty = e^{L}.
+					\]
+					Taking the logarithm of both sides changes nothing, nor does dividing by $\log{p}$ where $p\neq\infty$, or adding $c$. Hence, we have that
+					\[
+					\lim_{n\to\infty}
+					\left|
+					\frac{1}{(n+1)^{k}}-\frac{1}{n^{k}}
+					\right|_{p} = \infty,
+					\]
+					so the sequence of partial sums does \textbf{not} converge, and therefore the series does not converge.
+	      	      \end{quote}
 
@@
 	      	      does not converge in $\QQ_{p}$ for any prime $p\neq\infty$.
+	      	      \begin{quote}
+	      	      	\textit{\textbf{Proof:}} We want to show that
+	      	      	\[
+	      	      		\lim_{n\to\infty}\left|(n+1)^{k} - n^{k}\right|_{p}\neq 0.
+	      	      	\]
+					Using the binomial theorem, we find that
+					\[
+					\left|(n+1)^{k} - n^{k}\right|_{p} =
+					\left|\sum_{j=0}^{k}\binom{k}{j}n^{j} - n^{k}\right|_{p} =
+					\left|\sum_{j=0}^{k-1}\binom{k}{j}n^{j}\right|_{p}
+					\]
+	      	      \end{quote}
 	      \end{enumerate}

### 2016-11-01 ode_two/paper.tex
finished hw

@@
 R\"{o}ssler analyzed the system for relatively small $a$ and $b$, and we analyzed the stability of critical points when we neglected $a$ and $b$.
+The bifurcations will reveal behavior such as the periodicity and chaotic nature of the system.
 We will first analyze the effect that small changes to $a$ have on the system.
 
-The $a$ parameter affects our $g(x,y,z)$. For simplicity, we notice that $g$ is only dependent on $x$ and $y$, so we will let $g(x,y,z)=\gamma(x,y)$.
+In order to analyze the effect of $a$ on the system, we need to hold $b$ and $c$ constant.
+We will let $b=0.1$ and $c=8.8$.
+Figure \ref{fig:a_bifur_01} shows the effect that varying $a$ has on the system in the $x-y$ plane.
+As $a$ increases from $-\infty$, the system moves from a spiral sink to a center and finally to a saddle, moving from stable to semi-stable to unstable.
+
+When we examine the stability changes caused by $a$, we notice that the
 
@@
 	\centering
+	\begin{tabular}{c c c}
+		\includegraphics[scale=0.4]{a_bifur_n0p2} &
+		\includegraphics[scale=0.4]{a_bifur_n1p52656} &
+		\includegraphics[scale=0.4]{a_bifur_2p1}\\
+		$a=-0.2$ & $a=-1.52656$ & $a=2.1$
+	\end{tabular}
+	\caption{Three bifurcations for $a$ when $z=0$}
+	\label{fig:a_bifur_01}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
 	\begin{tabular}{c c}

### 2016-11-01 ode_two/presentation.tex
finished hw

\documentclass{beamer}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{graphicx}

\title{The R\"{o}ssler System}
\author{Steven Rosendahl}
\date{}

\usetheme{Rochester}
\usecolortheme{whale}

\begin{document}

\begin{frame}
    \titlepage
\end{frame}

\begin{frame}{Introduction}
    \begin{itemize}
        \item Created by Otto R\"{o}ssler in 1976
        \pause
        \item System constructed to display simplest possible \textit{Strange Attractor}
        \pause
    \end{itemize}
    \begin{definition}
        An \alert{attractor} is a set of values towards which a system moves when initial conditions are \textit{near} the attractor.
        \pause
        An attractor is called \alert{strange} if it exhibits fractal behavior.
    \end{definition}
    \pause
    \begin{itemize}
        \item Strange attractors are often associated with chaotic systems.
    \end{itemize}
\end{frame}

\begin{frame}{System}
    \begin{align*}
        \dot{x}&= -y-z\\
        \dot{y}&= x+ay\\
        \dot{z}&= b+z(x-c)
    \end{align*}
    \pause
    \begin{itemize}
        \item $a,b,c$ are responsible for attractive behavior
        \pause
        \item Want to analyze the effect small $a$ and $b$ have on system
    \end{itemize}
\end{frame}

\end{document}

### 2016-11-01 padics/hw5.tex
finished hw

--- 2016-10-28 padics/hw5.tex
@@
 \begin{enumerate}
-	\item Let $f(x,y)=x^2 + y^2$. Observe that $(x,y)=(0,0)$ defines a solution to the equation $f(x,y)=0$. Do the remainder of this problem, we shall call $(0,0)$ the trivial solution to $f(x,y)=0$.
+	\item Let $f(x,y)=x^2 + y^2$. Observe that $(x,y)=(0,0)$ defines a solution to the equation $f(x,y)=0$. For the remainder of this problem, we shall call $(0,0)$ the trivial solution to $f(x,y)=0$.
 	      \begin{enumerate}
 	      	\item Prove that $f(x,y)=0$ has no non-trivial solutions with $x,y\in\QQ$.
+	      	      \begin{quote}
+	      	      	Suppose that $f(x,y)$ has a non-trivial zero in $\QQ$.
+Then $x^{2}+y^{2} = 0$ for some pair $(x,y)\neq(0,0)$.
+For any choice of $x\neq0$ or $y\neq0$, we know that $x^{2}$ and $y^{2}$ are both positive.
+The sum of two positive numbers is still positive, so $x^{2}+y^{2} > 0$ for $x\neq0$ and $y\neq0$.
+Therefore there cannot be any other zeros besides $(0,0)$.
+	      	      \end{quote}
 	      	\item Prove that there exists a prime $p$ such that $f(x,y)=0$ has a nontrivial solution with $x,y\in\QQ_{p}$.
+	      	      \begin{quote}
+	      	      	Let $p=2$ and $y=1$. We want to show by Hensel's Lemma that $f(x,1)=g(x)=x^{2}+1$ has a solution near $x=1$. We can see that the first property of Hensel's Lemma is satisfied since
+	      	      	\[
+	      	      		g(1)=1^{2}+1=2\equiv0\mod2.
+	      	      	\]
+	      	      	Similarly, the second property is satisfied, since
+	      	      	\[
+	      	      		g'(1)=1 \not\equiv 0\mod2.
+	      	      	\]
+	      	      	Hence, there is a non-trivial zero of $f$ at $(x,y)=(1,1)$.
+	      	      \end{quote}
 	      	\item Verify that $f$ satisfies the Hasse Principle without referring to the Hasse-Minkowski Theorem.
+	      	      \begin{quote}
+	      	      	We have shown that $f$ has has no non-trivial solutions in $\QQ$ and that $f$ has a non-trivial solution in $\QQ_{2}$, so it satisfies the negation of the Hasse Principle.
+	      	      \end{quote}
 	      \end{enumerate}
@@
 	      	\begin{align*}
-					\left|a_{n+1}p^{n+1}-a_{n}p^{n}\right|_{p} &= \max\left\{%
+	      		\left|a_{n+1}p^{n+1}-a_{n}p^{n}\right|_{p} & \leq \max\left\{%
 	      		\left|a_{n+1}p^{n+1}\right|_{p},\left|a_{n}p^{n}\right|_{p}%
 	      		\right\}\\
-					&= \max\left\{%
+	      		                                           & = \max\left\{%
 	      		\frac{1}{p^{v_{p}(a_{n+1}) + n + 1}}, \frac{1}{p^{n+v_{p}(a_{n})}}%
@@
 	      		\frac{1}{p^{v_{p}(a_{n+1}) + n + 1}}, \frac{1}{p^{n+v_{p}(a_{n})}}%
-					\right\} &= \max\left\{0,0\right\}\\
-					&= 0.
+	      		\right\}                                   & = \max\left\{0,0\right\} \\
+	      		                                           & = 0.
 	      	\end{align*}
+	      	Since the $p$-adic absolute value is greater than or equal to 0, we know
+	      	\[
+	      		\lim_{n\to\infty}\left|a_{n+1}p^{n+1}-a_{n}p^{n}\right|_{p} = 0.
+	      	\]
 	      	Therefore, the sequence of partial sums is Cauchy, and thus convergent, so the series converges.
@@
 	      	      does not converge in $\QQ_{p}$ for any prime $p\neq\infty$.
+	      	      \begin{quote}
+	      	      	\textit{\textbf{Proof:}} If we can show that the sequence of partial sums does not converge, then the series will not converge. We want to evaluate (and eventually take the limit of)
+	      	      	\[
+	      	      		\left|
+	      	      		\frac{1}{(n+1)^{k}}-\frac{1}{n^{k}}
+	      	      		\right|_{p}.
+	      	      	\]
+	      	      	We know by definition that
+	      	      	\begin{align*}
+	      	      		\left|
+	      	      		\frac{1}{(n+1)^{k}}-\frac{1}{n^{k}}
+	      	      		\right|_{p} & =
+	      	      		\left|
+	      	      		\frac{n^{k}-(n+1)^{k}}{n^{k}(n+1)^{k}}
+	      	      		\right|_{p}\\
+	      	      		            & = p^{-
+	      	      		\left[
+	      	      		v_{p}(n^{k}-(n+1)^{k}) - v_{p}(n^{k})-v_{p}((n+1)^{k})
+	      	      		\right]
+	      	      		}.
+	      	      	\end{align*}
+	      	      	Through some tedious calculation, we can solve for the valuation of each term in the exponent.
+	      	      	\begin{gather*}
+	      	      		v_{p}\left(n^k-(n+1)^k\right) =
+	      	      		\frac{
+	      	      			\log{(n^k-(n+1)^k) - \log(m_{0})}
+	      	      			}{
+	      	      			\log{(p)}
+	      	      			}\\
+	      	      		v_{p}(n^{k}) =
+	      	      		\frac{
+	      	      			k\log{(n)}-\log{(m_{1})}
+	      	      			}{\log{(p)}}\\
+	      	      		v_{p}((n+1)^k) =
+	      	      		\frac{
+	      	      			k\log{(n+1)}-\log{(m_{2})}
+	      	      			}{
+	      	      			\log{(p)}
+	      	      		},
+	      	      	\end{gather*}
+	      	      	where $m_{0},m_{1}$, and $m_{2}$ do not divide $p$. Combining these valuations as required yields
+	      	      	\[
+	      	      		-\left[v_{p}(n^{k}-(n+1)^{k}) - v_{p}(n^{k})-v_{p}((n+1)^{k})\right]=
+	      	      		\frac{
+	      	      			\log{\left(
+	      	      				\frac{n^{k}(n+1)^{k}}{n^k-(n+1)^k}
+	      	      				\right)}
+	      	      			}{
+	      	      			\log{p}
+	      	      		} + c
+	      	      	\]
+	      	      	where $c$ is some constant number independant of $n$. We can now evaluate the real valued limit as $n\to\infty$.
+	      	      	\[
+	      	      		\lim_{n\to\infty}\frac{
+	      	      			\log{\left(
+	      	      				\frac{n^{k}(n+1)^{k}}{n^k-(n+1)^k}
+	      	      				\right)}}{\log{p}} + c = c+\frac{1}{\log{p}}\lim_{n\to\infty}\log{\left(
+	      	      		\frac{n^{k}(n+1)^{k}}{n^k-(n+1)^k}
+	      	      		\right)}.
+	      	      	\]
+	      	      	The function on the inside of the limit is continuous for all $k\in\NN$, so we can determine the limit as follows:
+	      	      	\begin{align*}
+	      	      		\lim_{n\to\infty}\log{\left(
+	      	      		\frac{n^{k}(n+1)^{k}}{n^k-(n+1)^k}
+	      	      		\right)} & = L      \\
+	      	      		e^{
+	      	      		\lim_{n\to\infty}\log{\left(
+	      	      		\frac{n^{k}(n+1)^{k}}{n^k-(n+1)^k}
+	      	      		\right)}
+	      	      		}        & = e^{L}  \\
+	      	      		\lim_{n\to\infty}e^{\log{\left(
+	      	      		\frac{n^{k}(n+1)^{k}}{n^k-(n+1)^k}
+	      	      		\right)}
+	      	      		}        & = e^{L}  \\
+	      	      		\lim_{n\to\infty}\left(
+	      	      		\frac{n^{k}(n+1)^{k}}{n^k-(n+1)^k}
+	      	      		\right)  & = e^{L}.
+	      	      	\end{align*}
+	      	      	For the fraction inside the limit, notice that after multiplication, the highest order term in the numerator will be $n^{2k}$, while the highest order term in the denominator will be $n^{k}$, so the limit behaves like
+	      	      	\[
+	      	      		\lim_{n\to\infty}\frac{n^{2k}}{n^k}=\lim_{n\to\infty}n^{k}=\infty = e^{L}.
+	      	      	\]
+	      	      	Taking the logarithm of both sides changes nothing, nor does dividing by $\log{p}$ where $p\neq\infty$, or adding $c$. Hence, we have that
+	      	      	\[
+	      	      		\lim_{n\to\infty}
+	      	      		\left|
+	      	      		\frac{1}{(n+1)^{k}}-\frac{1}{n^{k}}
+	      	      		\right|_{p} = \infty,
+	      	      	\]
+	      	      	so the sequence of partial sums does \textbf{not} converge, and therefore the series does not converge.
+	      	      \end{quote}
 
@@
 	      	      does not converge in $\QQ_{p}$ for any prime $p\neq\infty$.
+	      	      \begin{quote}
+	      	      	\textit{\textbf{Proof:}} We want to show that the individual terms in the series do not tend towards 0 $p$-adically.
+Consider all the elements of $\ZZ/p\ZZ$.
+We know that all multiples of $p$ will live in $\overline{0}$ and as such will have a $p$-adic absolute value less than 1.
+All other elements will have a $p$-adic absolute value of 1, since they contain a factor of $p^{0}$.
+Hence, the sum tends towards infinity, so the series does not converge.
+	      	      \end{quote}
 	      \end{enumerate}

### 2016-11-15 ode_two/paper.tex
finished some p-adic stuff

@@
 
-We now know that both equilibrium points are unstable when we ignore $a$ and $b$ in comparison to $c$.
-In fact, it turns out that this stability is preserved even when we reintroduce $a$ and $b$ back into the system.
-Figure \ref{fig:raw_ev} shows us what one of the raw eigenvalues looks like, according to Mathematica.
-The three eigenvalues resulting from the unsimplified system still exhibit the saddle focus behavior that we saw when we allowed $a$ and $b$ to drift towards zero.
-We now want to turn our attention towards the actions of the system when we change the values for $a$, $b$, and $c$.
+We now know that both equilibrium points are unstable when we ignore $a$ and $b$ in comparison to $c$, but that behavior seems odd for an attractor.
+We analyzed the simple roots so that we would not have to try to symbolically analyze the eigenvalues of the system (see Figure \ref{fig:raw_ev} which gives one of the three eigenvalues of the linearized system when $a$ and $b$ are not ignored).
+We will now see the effects of the individual parameters on the system; to do this, we will have to hold two of the values constant.
+Throughout this discussion, we will consider $a=0.1, b=0.1$, and $c=8.8$ for the respective constant values.
 
@@
 
-In order to analyze the effect of $a$ on the system, we need to hold $b$ and $c$ constant.
-We will let $b=0.1$ and $c=8.8$.
+In order to analyze the effect of $a$ on the system, we need to hold $b$ and $c$ constant while changing $a$.
+The general strategy here will be to analyze the change on the eigenvalues of the linearized system, which will in turn allow us to .
 Figure \ref{fig:a_bifur_01} shows the effect that varying $a$ has on the system in the $x-y$ plane.
@@
 
-When we examine the stability changes caused by $a$, we notice that the
+To see the change  in the eigenvalues, we can use the roots we found in (\ref{eq:csln1}) and (\ref{eq:csln2}).
+Replacing $a,b,$ and $c$ with our constant values yields the new Jacobians the we need to analyze.
+\begin{equation}
+	J_{\text{roots at (\ref{eq:csln1})}}=
+	\left[
+	\begin{array}{c c c}
+		0                                  & -1 & -1                        \\
+		1                                  & a  & 0                         \\
+		\frac{8.8-\sqrt{77.44 - 0.4a}}{2a} & 0  & -4.4-0.5\sqrt{77.44-0.4a}
+	\end{array}
+	\right]\label{eq:a_mat_01}
+\end{equation}
+\noindent and
+\begin{equation}
+	J_{\text{roots at (\ref{eq:csln2})}} =
+	\left[
+	\begin{array}{c c c}
+		0                                   & -1 & -1                          \\
+		1                                   & a  & 0                           \\
+		\frac{0.5(8.8\sqrt{77.44-0.4a})}{a} & 0  & 0.5(-8.8+\sqrt{77.44-0.4a})
+	\end{array}
+	\right]\label{eq:a_mat_02}
+\end{equation}
+\noindent yield different stability based on the small changes in $a$. For the Jacobian given in (\ref{eq:a_mat_01}), the eigenvalues change around $a=2$.
+\begin{center}
+	\begin{tabular}{l | l}
+		$a=1.8$                      & $a=2.1$              \\
+		\hline
+		$\lambda_{1}=-8.77$          & $\lambda_{1}=-8.775$ \\
+		$\lambda_{2}=0.899-0.435i$   & $\lambda_{2}=0.727$  \\
+		$\lambda_{3}=0.899 + 0.435i$ & $\lambda_{3}=1.371$
+	\end{tabular}
+\end{center}
+\noindent\\
+\noindent We can clearly see that the stability changes from a stable focus node to an unstable saddle.
+As $a$ increases past $2.1$, the system continues to be unstable.
+In Figure \ref{fig:3dsys_02}, the purple section of the graph represents the portion of the system that we are analyzing.
+When $a>2$, the system is unstable at this point, causing trajectories to head away from the spiral and towards the orange section of the graph.
+We can also see what happens near the other fixed point.
+The eigenvalues of (\ref{eq:a_mat_02}) appear to change around $a=58.5$ and again around $a=193.6$.
+\begin{center}
+	\begin{tabular}{l | l | l}
+		$a=58.5$                      & $a=58.6$             & $a=198.4$                   \\
+		\hline
+		$\lambda_{1}=58.483$          & $\lambda_{1}=-0.366$ & $\lambda_{1}=193.695$       \\
+		$\lambda_{2}=-0.354 - 0.025i$ & $\lambda_{2}=-0.343$ & $\lambda_{2}=-4.39+0.1i $   \\
+		$\lambda_{3}=-0.354 + 0.025i$ & $\lambda_{3}=58.58$  & $\lambda_{3}=-.0002-0.001i$
+	\end{tabular}
+\end{center}
+\noindent\\
+\noindent Here we see an unstable focus node changing into an unstable saddle.
+As $a$ increases, the system cycles between the a focus node and and unstable saddle, which is indicative of periodic behavior of the system.
+
+We can now analyze the effect that changes in $b$ have on the system.
+For the fixed point in (\ref{eq:csln1}), we see a change in stability around $b=94.2$ and then again around $b=193.8$.
+At this fixed point, the eigenvalues change from a stable focus node to a stable node, and then into a set of three complex eigenvalues with non-zero real part.
+The final case is not a named type of equilibrium, and in fact it is not hyperbolic, so it does not tell us anything useful about the nonlinear system.
+It is unstable though, since at least one of the eigenvalues has a positive real part.
+\begin{center}
+	\begin{tabular}{l | l | l}
+		$b=94.2$                           & $b=100$                 & $b=195$                           \\
+		\hline
+		$\lambda_{1}=-5.27985$             & $\lambda_{1}=-4.75948$  & $\lambda_{1}=-2.51539 - 6.51211i$ \\
+		$\lambda_{2}=-1.08805 - 0.107786i$ & $\lambda_{2}=-1.93582$  & $\lambda_{2}=-1.78605 + 6.15461i$ \\
+		$\lambda_{3}=-1.08805 + 0.107786i$ & $\lambda_{3}=-0.664116$ & $\lambda_{3}=0.00144 - 0.0167i$
+	\end{tabular}
+\end{center}
+\noindent\\
+\noindent Similarly, we can analyze the effect that perturbations in $b$ have on the fixed point given in (\ref{eq:csln2}).
+When varying $b$ from $0.1$ to 250, we do not see any change in the stability of the fixed point.
+It remains an unstable saddle focus, so we need not analyze this any further.
+It should be noted that we are considering fixed values for $a,b$, and $c$.
+If we chose other values and repeat this process, we may find that the changes to $b$ in this case actually do affect the system.
+We will not consider these cases, since the resulting work would take several more pages.
+
+Finally, we can analyze the effect that $c$ has on the system.
+Similar to the effect of $b$ on (\ref{eq:csln2}), varying $c$ from $8.8$ to 2250 does not appear to change the stability of (\ref{eq:csln1}) and (\ref{eq:csln2}).
+In the first case, the fixed point remains an unstable saddle focus; in the latter case, the system also remains an unstable saddle focus.
+While $c$ may not change the stability of the fixed points, small changes in $c$ can yield both chaotic and periodic behavior.
+Figure \ref{fig:c_chaos} shows some of the values for which the attractor is periodic and chaotic.
+The figure considers the system with $a=b=0.1$, as we have done in our analysis.
+Notice that chaotic behavior creates infinite periods and that the system is subject to very small changes in $c$ (i.e.
+12.6 to 13).
+
+Overall, the R\"{o}ssler system is globally stable, but exhibits local instability based on initial conditions.
+This is a property of attractors; any trajectories that start outside of the attractor will eventually head towards the attractor.
+We can see this behavior in Figure \ref{fig:attractive}.
+The pink trajectory moves towards the attractor regardless of the local stability of the attractor itself (hence the name \textit{attractor}).
+This behavior allows us to conclude that the attractor will be globally stable.
+
+The R\"{o}ssler system illustrates a simple attractor; as a result, it is not a useful system to model practical situations.
+Viewing the system as an illustration for more complex system provides insight to more complex examples.
+We have seen the effect that small changed in $a,b$, and $c$ have on the system, as well as the effect of initial conditions on solutions to the linear system.
+Luckily, the fixed points tend to remain hyperbolic as $a,b$, and $c$ change, which allows for easier analysis of the system.
+The R\"{o}ssler system provides an elegant view of a complex structure that would otherwise be very complicated.
 
@@
 
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.5]{varying_c}
+	\caption{The chaotic behavior and periodicity of the system based on changed in $c$, courtesy of \cite{rossler_bifur}}
+	\label{fig:c_chaos}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[scale=0.6]{attractive}
+	\caption{An example of the global stability of the system. The pink trajectory heads towards the attractor and will remain there as $t\to\infty$ \cite{rossler_bifur} }
+	\label{fig:attractive}
+\end{figure}
+
 \newpage

### 2016-11-15 ode_two/presentation.tex
finished some p-adic stuff

@@
 \usepackage{graphicx}
+
+\newcommand{\RR}{\mathbb{R}}
 
@@
 	\begin{align*}
-        \dot{x}&= -y-z\\
-        \dot{y}&= x+ay\\
-        \dot{z}&= b+z(x-c)
+		\dot{x} & = -y-z     \\
+		\dot{y} & = x+ay     \\
+		\dot{z} & = b+z(x-c)
 	\end{align*}
@@
 		\item Want to analyze the effect small $a$ and $b$ have on system
+		      \pause
+		      \begin{itemize}
+		      	\item Is the system globally stable? Locally?
+		      	      \pause
+		      	\item When does chaotic behavior occur?
+		      \end{itemize}
 	\end{itemize}
@@
 
+\begin{frame}{First Glance}
+	\begin{itemize}
+		\item Finding fixed points
+	    \pause
+	\end{itemize}
+	\begin{center}
+		\scalebox{0.75}{
+			\texttt{Solve[f[x, y, z] == 0 \&\& g[x, y, z] == 0 \&\& h[x, y, z] == 0, \{x, y, z\}]}
+		}
+	\end{center}
+	\pause
+	\begin{columns}[T]
+		\begin{column}{0.48\textwidth}
+			\[
+				\begin{cases}
+					x=\frac{1}{2}\left(c-\sqrt{c^{2}-4ab}\right)                     \\
+					y=\frac{1}{2}\left(\frac{\sqrt{c^{2}-4ab}}{a}-\frac{c}{a}\right) \\
+					z=\frac{c-\sqrt{c^{2}-4ab}}{2a}
+				\end{cases}
+			\]
+		\end{column}
+		\begin{column}{0.48\textwidth}
+			\[
+				\begin{cases}
+					x=\frac{1}{2}\left(c+\sqrt{c^{2}-4ab}\right)                      \\
+					y=\frac{1}{2}\left(-\frac{\sqrt{c^{2}-4ab}}{a}-\frac{c}{a}\right) \\
+					z=\frac{c+\sqrt{c^{2}-4ab}}{2a}
+				\end{cases}
+			\]
+		\end{column}
+	\end{columns}
+\end{frame}
+
+\begin{frame}{Simplification}
+    \begin{itemize}
+        \item Let $a,b$ tend to zero
+        \pause
+        \item Cannot actually be zero (division by 0 at fixed points)
+        \pause
+    \end{itemize}
+    \[\left\{x = 0\qquad y = 0\qquad z = 0\right\}\]
+    \pause
+    \[\left\{x = c\qquad y = -\frac{c}{a}\qquad z = \frac{c}{a}\right\}\]
+    \pause
+    \begin{itemize}
+        \item Much easier to analyze stability of system
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Linearization}
+    \begin{itemize}
+        \item Can take Jacobian of system to analyze stability
+        \pause
+        \item Which stabilities are hyperbolic in a 3D system?
+        \pause
+    \end{itemize}
+    % \begin{definition}
+    %     When $\lambda_{1}$, $\lambda_{2}$, and $\lambda_{3}$ are in $\RR$ and they all have the same sign, the fixed point is a \alert{node}.
+    % \end{definition}
+    % \pause
+    % \begin{definition}
+    %     When $\lambda_{1}$, $\lambda_{2}$, and $\lambda_{3}$ are in $\RR$ and at least one is positive and one is negative, the fixed point is a \alert{saddle}.
+    % \end{definition}
+    \begin{definition}
+        When $\lambda_{1}\in\RR$ and $\lambda_{2},\lambda_{3}$ are complex conjugates \textit{and} the real part of $\lambda_{1},\lambda_{2}$, and $\lambda_{3}$ all have the same sign, the fixed point is called a \alert{Focus Node}
+    \end{definition}
+    \pause
+    \begin{definition}
+        When $\lambda_{1}\in\RR$ and $\lambda_{2},\lambda_{3}$ are complex conjugates, then if the sign of $\lambda_{1}$ is the opposite of the sign of the real part of $\lambda_{2}$ and $\lambda_{3}$, the fixed point is called a \alert{Saddle Focus}
+    \end{definition}
+    \pause
+    \begin{itemize}
+        \item Other stability exists, but are often named after the bifurcations that produce them
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Linearization}
+    \begin{itemize}
+        \item Jacobian of system will tell us about behavior of fixed points
+    \end{itemize}
+
+\end{frame}
+
 \end{document}

### 2016-11-15 padics/hw6.tex
finished some p-adic stuff

\documentclass{hw}

\begin{document}
\makeheader{6}
\begin{enumerate}
	\item Find the radius of convergence of the series
	      \[
	      	f(x)=\sum_{i=0}^{\infty}\left(
	      	\frac{x}{7}
	      	\right)^{n}
	      \]
	      in $\QQ_{p}$.
	      \begin{quote}
	      	We will consider the case in $\QQ_{7}$ first. We need to find
	      	\begin{align*}
	      		r & = \left(\limsup_{n\to\infty}\left(
	      		\left|\frac{1}{7^n}\right|_{7}\right)^{1/n}
	      		\right)^{-1}\\
	      		  & = \left(
	      		\limsup_{n\to\infty}(7^{n})^{1/n}
	      		\right)^{-1}\\
	      		  & = \left(
	      		\limsup_{n\to\infty}7
	      		\right)^{-1}\\
	      		  & = \frac{1}{7}.
	      	\end{align*}
	      	By the $p$-adic root test, this means that the series converges for all $x\in\QQ_{p}$ such that $|x|_{p}<1/7$. To check convergence in $\QQ_{p}$ when $p\neq 7$, we can consider
	      	\begin{align*}
	      		\left|\frac{1}{7^{n}}\right| & = p^{-v_{p}(1/7^{n})}                                                              \\
	      		                             & = p^{-v_{p}(1) + v_{p}(7^{n})}                                                     \\
	      		                             & = p^{0+0} \quad\text{since there will be no factors of $p$ in $7^n$ when $p\neq7$} \\
	      		                             & = 1,
	      	\end{align*}
	      	so
	      	\[
	      		r = \frac{1}{\limsup_{n\to\infty}\left|\frac{1}{7^{n}}\right|_{p}^{1/n}} =
	      		\frac{1}{\limsup_{n\to\infty}1} = 1,
	      	\]
	      	and the radius of convergence is 1. Therefore, $f(x)$ converges on $\{x : |x|_{p}<1\}$, also known as $\ZZ_{p}$.
	      \end{quote}

	\item Suppose that $\{a_{n}\}$ is the sequence of points in $\{1,2,3,4,5\}$ defined so that $a_{n}\equiv n+1 \mod 5$. That is, $\{a_{n}\}$ is given by
	      \[
	      	1,2,3,4,5,1,2,3,4,5,1,2,3,4,5,\dots
	      \]
	      Find the radius of convergence $r$ of the series
	      \[
	      	g(x)=\sum_{n=0}^{\infty}\left(\frac{x}{a_{n}}\right)^{n}
	      \]
	      in $\QQ_{5}$ and in $\QQ_{3}$. In each case, determine whether the series converges when $|x|_{p}=r$.
	      \begin{quote}
	      	Consider the case in $\QQ_{5}$. We have that
	      	\begin{align*}
	      		r & = \frac{1}{\limsup_{n\to\infty}\left|\frac{1}{a_{n}^{n}}\right|_{5}^{1/n}}     \\
	      		  & = \frac{1}{\limsup_{n\to\infty}\left(\left|a_{n}\right|_{5}^{-n}\right)^{1/n}} \\
	      		  & = \frac{1}{\limsup_{n\to\infty}\left|a_{n}\right|_{5}^{-1}}                    \\
	      		  & = \frac{1}{\limsup_{n\to\infty}\left|\frac{1}{a_{n}}\right|_{5}}               \\
	      		  & = \frac{1}{5}.
	      	\end{align*}

	      	We know that $1/5$ is the right choice, since the 5-adic absolute value of $1/5$ is the largest value that appears in the sequence 5-adically. Therefore, the radius of convergence of the sequence is $1/5$. We need a value that has $|x|_{5}=r$, so we will let $x=5$. Evaluating $g(5)$ gives us
	      	\[
	      		\sum_{n=0}^{\infty}\left(\frac{5}{a_{n}}\right)^{n},
	      	\]
	      	which we can analyze for convergence by looking at
	      	\begin{align*}
	      		\lim_{n\to\infty}\left|\left(\frac{5}{a_{n}}\right)^{n}\right|_{5}
	      		  & =\lim_{n\to\infty}5^{-v_{5}(5^n)+v_{5}(a_{n}^n)}  \\
	      		  & =\lim_{n\to\infty}\frac{5^{nv_{5}(a_{n})}}{5^{n}}
	      	\end{align*}
	      	If we consider the subsequence of $a_{n}$ consisting of $\{5,5,5,5,\dots\}$, then the limit tends to $1$ since
	      	\[
	      		\lim_{n\to\infty}\frac{5^{nv_{5}(a_{n})}}{5^{n}}=\lim_{n\to\infty}\frac{5^{n}}{5^{n}}=1
	      	\]
	      	Similarly, for the subsequence defined by $\{1,2,3,4,1,2,3,4,\dots\}$, we have
	      	\[
	      		\lim_{n\to\infty}\frac{5^{nv_{5}(a_{n})}}{5^{n}}=\lim_{n\to\infty}\frac{5^{0}}{5^{n}}=0,
	      	\]
	      	So the limit does not exist. Therefore, the sequence does not converge on the boundary.

	      	\noindent Consider the case in $\QQ_{3}$. Then we can say that
	      	\[
	      		\frac{1}{\limsup_{n\to\infty}\left(\left|a_{n}\right|_{3}^{-n}\right)^{1/n}} =
	      		\frac{1}{\limsup_{n\to\infty}\left|\frac{1}{a_{n}}\right|_{3}} = \frac{1}{3}.
	      	\]
	      	We need to pick a point on the boundary (i.e. $|x|_{3}=1/3$), so let $x=3$. Then
	      	\begin{align*}
	      		\lim_{n\to\infty}\left|\frac{3^{n}}{a_{n}^{n}}\right|_{3}
	      		  & = \lim_{n\to\infty}3^{-v_{3}(3^{n})+v_{3}(a_{n}^{n})} \\
	      		  & = \lim_{n\to\infty}\frac{3^{nv_{3}(a_{n})}}{3^{n}}.
	      	\end{align*}
	      	If we consider the subsequence of $a_{n}$ given by $\{3,3,3,3,\dots\}$, we see that
	      	\[
	      		\lim_{n\to\infty}\frac{3^{nv_{3}(a_{n})}}{3^{n}}=
	      		\lim_{n\to\infty}\frac{3^{n}}{3^{n}}=1,
	      	\]
	      	and if we consider the subsequence given by $\{1,2,4,5,1,2,4,5,\dots\}$, then we find that
	      	\[
	      		\lim_{n\to\infty}\frac{3^{nv_{3}(a_{n})}}{3^{n}}=
	      		\lim_{n\to\infty}\frac{3^{0}}{3^{n}}=0.
	      	\]
	      	Therefore, the limit does not exist, so the series diverges on the boundary of the region of convergence.

	      \end{quote}

	\item Find the radius of convergence of the power series
	      \[
	      	h(x)=\sum_{n=0}^{\infty}n!x^{n}
	      \]
	      in $\QQ_{p}$.
		  \begin{quote}
			  To begin, we want a formula for $v_{p}(n!)$. Luckily, there' a nice theorem that can help us here.
			  \begin{quote}
				  \textbf{Legendre's Theorem}:
				  \[
				  	v_{p}(n!)=\sum_{k=1}^{\infty}\left\lfloor\frac{n}{p^{k}}\right\rfloor.
				  \]
				%   \textbf{\textit{Proof:}} We claim that there are $\lfloor n/p \rfloor$ integers below $n$ that have a factor of $p$ in them. $n!$ is the product of integers $1\dots n$, so we get at least one factor of $p$ in $n!$ for each multiple of $p$ in $\{1,2,3,\dots,n\}$. Similarly, each multiple of $p^2$ contributes a factor of $p$, and each multiple of $p^3$ contributes a factor of $p$, and so on.
			  \end{quote}
			  In order to determine the radius of convergence, we need to evaluate
			  \[
			  	\frac{1}{\limsup_{n\to\infty}(|n!|_{p})^{1/n}}.
			  \]
			  Consider the quantity inside the $\limsup$ expression. We need to calculate $(p^{-v_{p}(n!)})^{(1/n)}$. Note that
			  \begin{align*}
				  \sum_{k=1}^{\infty}\left\lfloor\frac{n}{p^{k}}\right\rfloor &\leq \sum_{k=1}^{\infty}\frac{n}{p^{k}}\\
				  -\sum_{k=1}^{\infty}\left\lfloor\frac{n}{p^{k}}\right\rfloor &\geq -\sum_{k=1}^{\infty}\frac{n}{p^{k}}\\
				  p^{-\sum_{k=1}^{\infty}\left\lfloor\frac{n}{p^{k}}\right\rfloor} &\geq p^{-\sum_{k=1}^{\infty}\frac{n}{p^{k}}}\\
				  \frac{1}{p^{-\sum_{k=1}^{\infty}\left\lfloor\frac{n}{p^{k}}\right\rfloor}} &\leq \frac{1}{p^{-\sum_{k=1}^{\infty}\frac{n}{p^{k}}}},
			  \end{align*}
			  so
			  \begin{align*}
				  \frac{1}{\limsup_{n\to\infty}|n!|_{p}^{1/n}} &=
				  \frac{1}{\limsup_{n\to\infty}p^{\left(-\sum_{k=1}^{\infty}\left\lfloor\frac{n}{p^{k}}\right\rfloor\right)^{1/n}}}\\
				  &\leq \frac{1}{\limsup_{n\to\infty}p^{\left(-\sum_{k=1}^{\infty}\frac{n}{p^{k}}\right)^{1/n}}}\\
				  &= \frac{1}{\limsup_{n\to\infty}p^{-\sum_{k=1}^{\infty}\frac{1}{p^{k}}}}.
			  \end{align*}
			  By the geometric series test, we find that
			  \[
			  	\sum_{k=1}^{\infty}\frac{1}{p^k}=\frac{1}{1-p},
			  \]
			  so
			  \[
			  	\frac{1}{\limsup_{n\to\infty}|n!|_{p}^{1/n}} \leq p^{\frac{1}{p-1}}
			  \]
			  To determine a lower bound for the expression, consider
			  \[
			  	\sum_{k=1}^{\infty}\left\lfloor\frac{n}{p^{k}}\right\rfloor \geq \sum_{k=1}^{\infty}\frac{n}{p^{k}}-1.
			  \]
			  Using a similar argument as above, we find that
			  \[
			  	\frac{1}{p^{-\sum_{k=1}^{\infty}\left\lfloor\frac{n}{p^{k}}\right\rfloor}} \geq
				\frac{1}{p^{\left(-\sum_{k=1}^{\infty}\frac{n}{p^{k}}-1\right)}}
			  \]
			  so that
			  \begin{align*}
				\frac{1}{\limsup_{n\to\infty}\left(p^{-\sum_{k}\left\lfloor\frac{n}{p^k}\right\rfloor}\right)^{1/n}} &\geq
				\frac{1}{\limsup_{n\to\infty}\left(p^{-\sum_{k}\left(\frac{n}{p^k}-1\right)}\right)^{1/n}}\\
				&=\frac{1}{\limsup_{n\to\infty}p^{-\sum_{k}\frac{1}{p^k}-\frac{1}{n}}}\\
				&= \frac{1}{p^{-\sum_{k=1}^{\infty}\frac{1}{p^{k}}}}\\
				&= \frac{1}{p^{-\frac{1}{p-1}}}\\
				&= p^{\frac{1}{p-1}}
			  \end{align*}
			  Hence
			  \[
			  	p^{\frac{1}{p-1}}\leq
				\frac{1}{\limsup_{n\to\infty}(|n!|_{p})^{1/n}}
				\leq p^{\frac{1}{p-1}}\implies
				\frac{1}{\limsup_{n\to\infty}(|n!|_{p})^{1/n}} = p^{\frac{1}{p-1}}
			  \]
		  \end{quote}
\end{enumerate}
\end{document}

### 2016-11-18 software_design/datareader/main.py
update a lot

@@ -1,0 +1,72 @@
+from clint.textui import colored
+from csv import reader
+
+CRASH_LOCATION_INDEX = 2
+
+homs = open('./crashes.csv','r')
+table = []
+lc = 0
+for line in reader(homs):
+    if lc == 0:
+        lc = lc + 1
+        continue
+    table.append(line)
+
+def mostPopular():
+    undoc = "undocumented"
+    crashSites = {}
+    crashSites[undoc] = 0
+    for array in table:
+        placeOfCrash = array[CRASH_LOCATION_INDEX].strip().lower()
+        actualPlace = placeOfCrash
+        if not placeOfCrash == "x":
+            if not actualPlace in crashSites:
+                crashSites[actualPlace] = 1
+            else:
+                crashSites[actualPlace] = crashSites[actualPlace] + 1
+        else:
+            crashSites[undoc] = crashSites[undoc] + 1
+    sortedList = sorted(crashSites, key=crashSites.__getitem__)
+    tenBestVals = sortedList[-10:]
+    for place in tenBestVals:
+        pstrA = ""
+        amnt = crashSites[place]
+        for i in range(0,amnt):
+            pstrA = pstrA + colored.red("[]")
+        print("%-20s\t%s" % (place, pstrA))
+
+def amIAtRisk():
+    totalCrashes = len(table)
+    print("What country/state do you live in?")
+    location = input().lower().strip()
+    myChance = 0
+    for array in table:
+        placeOfCrash = array[CRASH_LOCATION_INDEX].strip().lower()
+        if placeOfCrash != "x":
+            if (location in placeOfCrash) or ((location + '"') in placeOfCrash):
+                myChance = myChance + 1
+    avg = myChance / totalCrashes
+    print("There have been an average of " + str(avg) + " crashes in you area.")
+    print("The total number of crashes in " + location + " since 1929 is " + str(myChance))
+    if(avg > 0.03):
+        print(colored.red("You are gonna get Donnie-Darkoed"))
+    else:
+        print(colored.green("You are not at risk"))
+
+while True:
+    print("What would you like to know?")
+    print("1. Am I at risk for being hit by an airplane?")
+    print("2. Most popular places for crashes")
+    print("Q: Exit")
+    choice = input().strip().lower()
+
+    if choice == "q":
+        break
+    elif choice == "1":
+        amIAtRisk()
+    elif choice == "2":
+        mostPopular()
+    else:
+        print("Unknown Choice")
+    print("Press Enter to continue...")
+    input()

### 2016-11-18 ode_two/paper.tex
finished prezzie

@@
 	\right| =
-	(c-\lambda)(\lambda + 1)^{2}=0.
+	(-c-\lambda)(\lambda^{2} + 1)=0.
 \]
-Hence, the system has three eigenvalues: $\lambda=-i, \lambda=-i,$ and $\lambda = -c$ (we will assume here that $c > 1$).
-This implies that the system has an unstable focus node at the origin \cite{equilibria}.
+Hence, the system has three eigenvalues: $\lambda=-i,\ \lambda=i,$ and $\lambda = -c$ (we will assume here that $c > 1$).
+This implies that the linear system has a center at the origin \cite{equilibria}.
+Since centers are \textbf{not} hyperbolic, we can look at various stream plots of the system.
 Figure \ref{fig:plot_cuts} shows various cuts of the system for various $z$ values; the system spirals in the $x-y$ plane, while moving towards and then away from the origin as $z$ moves from $-\infty$ to $\infty$.
+Hence, the system appears to actually have a spiral at the origin.
 We can solve this linearized system by looking at the equation
@@
 		\begin{array}{c}
-			i \\
+			-i \\
 			1 \\
@@
 	\begin{array}{c}
-		-i \\
+		-1 \\
 		1  \\
@@
 	\begin{array}{c}
-		i \\
+		-1 \\
 		1 \\

### 2016-11-18 ode_two/presentation.tex
finished prezzie

@@
         \item Jacobian of system will tell us about behavior of fixed points
+		\pause
     \end{itemize}
+	\[
+		|J(0,0,0)-\lambda I| =
+		\left|
+		\left[
+		\begin{array}{c c c}
+			-\lambda & -1       & -1         \\
+			1        & -\lambda & 0          \\
+			0        & 0        & -c-\lambda
+		\end{array}
+		\right]
+		\right| =
+		(c-\lambda)(\lambda + 1)^{2}=0.
+	\]
+	\pause
+	\begin{itemize}
+		\item System has three eigenvalues
+		\pause
+	\end{itemize}
+	\[
+		\lambda=-i\qquad\pause \lambda=-i\qquad\pause\lambda = -c
+	\]
+	\pause
+	\begin{definition}
+		If two of the eigenvalues of the system are pure imaginary, then the system has a \alert{center} at the fixed point. This is \textbf{not} a hyperbolic equilibrium.
+    \end{definition}
+	\pause
+	\begin{itemize}
+		\item $\therefore (0,0,0)$ is a center of the linear system
+	\end{itemize}
+\end{frame}
 
+\begin{frame}{Analyzing the Center}
+	\begin{itemize}
+		\item How to analyze the center of linear system?
+		\pause
+		\item Hold $z$ constant
+		\pause
+		\item Look at behavior of $(0,0)$ in $x-y$ plane
+	\end{itemize}
+	\begin{center}
+		\begin{tabular}{c c}
+			\includegraphics[scale=0.3]{spz-2} & \includegraphics[scale=0.3]{spz2}\\
+			$z=-2$ & $z=2$
+		\end{tabular}
+	\end{center}
+	\pause
+	\begin{itemize}
+		\item Actually appears to be a spiral
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Linearization}
+	\begin{itemize}
+		\item Can analyze the other fixed point
+		\pause
+	\end{itemize}
+	\begin{gather*}
+		\left|J\left(c,-\frac{c}{a},\frac{c}{a}\right) -\lambda I\right| =
+		\frac{ ac - a\lambda - c\lambda + a^{2}\lambda^{2} -a\lambda^{3}}{a}.
+	\end{gather*}
+	\pause
+	\begin{itemize}
+		\item Solving with Mathematica yields bad values
+		\pause
+		\item Recall system was meant to be analyzed with small $a,b$
+		\pause
+		\item Let $a\to0.14$, $c\to8.8$
+	\end{itemize}
+	\pause
+	\[
+		\lambda =0.001-7.99i\pause\qquad \lambda=0.001+7.99i\pause\qquad\lambda=0.13
+	\]
+	\pause
+	\begin{itemize}
+		\item $\therefore$ fixed point is an unstable focus node
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Chaos}
+	\begin{center}
+		\includegraphics[scale=0.2]{bane}
+	\end{center}
+	\pause
+	\begin{itemize}
+		\item Mathematica demonstration
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Chaos}
+	\begin{itemize}
+		\item Small changes in the system yield drastic changes in behavior
+		\pause
+		\item Periodicity changes to chaotic behavior
+	\end{itemize}
+	\pause
+	\begin{center}
+		\includegraphics[scale=0.3]{varying_c}
+	\end{center}
+\end{frame}
+
+%can talk about atrractiveness if need more time
+\begin{frame}{References}
+	\nocite{*}
+	\bibliographystyle{alpha}
+	\bibliography{mbib}
 \end{frame}

### 2016-11-30 religions_of_the_west/test_02.tex
working on final stuff

\documentclass[8pt]{article}
\usepackage{extsizes}
\usepackage{enumitem}
\usepackage{amsmath}
\setlist{nosep}
\usepackage[margin=0.2in]{geometry}

\begin{document}
\section*{Test 2 Review}
\begin{enumerate}
    \item Who is Ahura Mazda? Who is Angra Mainyu?
    \begin{enumerate}
        \item Ahura Mazda
        \begin{itemize}
            \item Creator God in Zoroastrianism
            \item Sole God of the religion
            \item Followers worship his attributes (\textit{Amesha Spentas})
            \item Had evil twin spirit named Angra Mainyu
            \item Truth teller
        \end{itemize}
        \item Angra Mainyu
        \begin{itemize}
            \item Evil counterpart to Ahura Mazda
            \item Can also create things
            \item Lier and deceiver (similar to Satan)
        \end{itemize}
    \end{enumerate}

    \item List some similarities between Zoroastrianism, Christianity, and post-exilic Judaism.
    \begin{itemize}
        \item Central God; Zoroastrianism: Ahura Mazda, Judaism/Christianity: Yahweh
        \item Some sects of Judaism look for Messiah; Christianity has Jesus; Zoroastrianism looked for Saoshyant\
        \item Similar eschatology (final judgment, raising of the dead)
        \item Devil figure (Angra Mainyu/Satan)
    \end{itemize}

    \item Discuss what the church said about Jesus Christ at Nicea and Chalcedon.
    \begin{enumerate}
        \item Nicea
        \begin{itemize}
            \item First ecumenical council of the Church
            \item Needed to establish doctrine to unite Christians
            \item Hosted by Constantine
            \item Athanasius proposed homoousios
            \begin{itemize}
                \item Father and son have same nature but are not the same
                \item Father and son share the same essence.
            \end{itemize}
            \item Trinity idea proposed (later rejected by enlightenment Christians)
        \end{itemize}
        \item Chalcedon
        \begin{itemize}
            \item Council met to settle disputes over Jesus' nature
            \item Greek ideas (God is separate) clashing with Hebrew ideas (God is with humanity)
            \item Leo's Tome: Jesus has two full natures in one person (Orthodox position)
            \item Established Mary's position in the church (Humans < Mary < God)
            \item Roman bishop gains authority at this council
        \end{itemize}
    \end{enumerate}

    \item Describe the evolution of the Papacy.
    \begin{itemize}
        \item Peter considered to be head of Church after Jesus' death
        \item Bishops are established throughout the known world
        \item Chalcedon: Bishop of Rome gains authority in establishing Doctrine
        \item Bishop of Rome becomes authority of the Church (i.e. Pope)
        \item Pope reigns above kings in the middle ages
        \item Church of England splits from the Catholic church over disputes (power of Pope weakened)
        \item Other theologians begin to reject Church's teaching (Martin Luther, John Calvin, etc.) further weakening the Pope's authority
    \end{itemize}

    \item What do Pelagius, Augustine, Erasmus, and Calvin believe about God's sovereignty and human freedon?
    \begin{enumerate}
        \item Pelagius
        \begin{itemize}
            \item Humans have free will
            \item Good deeds will lead you to heaven
            \item Did not believe in God's saving grace
        \end{itemize}
        \item Augustine
        \begin{itemize}
            \item Humans were predestined
            \item God selects certain people to receive his grace
            \item Only those who have been given the grace of God can follow
            \item Works on Earth do not matter
        \end{itemize}
        \item Erasmus
        \begin{itemize}
            \item Grace of God and works lead to salvation
            \item Humans have free will to follow or reject God
            \item Strong focus on working with God
            \item Followed by Catholic church
        \end{itemize}
        \item Calvin
        \begin{itemize}
            \item Humans have no free will
            \item Salvation comes \textbf{only} from the grace of God
            \item Humans are inherently tainted with sin
            \item Some are chosen for heaven; others for hell
            \item Strong emphasis on God's role in human life
        \end{itemize}
    \end{enumerate}

    \item Explain the following theories of atonement: Penal Substitution, Classical Redemption, Acceptation Theory, and Moral Influence
    \begin{enumerate}
        \item Penal Substitution
        \begin{itemize}
            \item Jesus was punished for sins of humanity
            \item God forgives all sins through this sacrifice
        \end{itemize}
        \item Classical Redemption
        \begin{itemize}
            \item
        \end{itemize}
        \item Acceptation Theory
        \begin{itemize}
            \item
        \end{itemize}
        \item Moral Influence
    \end{enumerate}

    \item Differentiate the Catholic, Lutheran, Reformed, and Anabaptist theories of the sacraments (baptism and eucharist)
    \begin{enumerate}
        \item Catholic
        \begin{itemize}
            \item Baptism of infants
            \item Original Sin is forgiven through baptism
            \item Eucharist is \textit{Transubstantiation}: transforming into literal body and blood
            \item Eucharist is central focus of the mass (sacrifice)
        \end{itemize}
        \item Lutheran
        \begin{itemize}
            \item Baptism of infants
            \item Original Sin is forgiven through baptism
            \item Eucharist is \textit{Consubstantiation}: elements are the same but Christ is present in them
        \end{itemize}
        \item Reformed
        \begin{itemize}
            \item Baptism of infants
            \item Baptism is an oath to bring up child in name of Christ
            \item Baptism is Symbol of the Covenant
            \item Eucharist is symbolic of Christ; used to remember his sacrifice
        \end{itemize}
        \item Anabaptist
        \begin{itemize}
            \item Baptism of adults
            \item Baptism is symbolic of conversion and acceptance to Christ
            \item Eucharist is symbolic of Christ; used to remember his sacrifice
        \end{itemize}
    \end{enumerate}

    \item Discuss the religious and secular heritage of this country
    \begin{itemize}
        \item Puritans had democratic structure in churches

    \end{itemize}

    \item Explain the following: Mass, The Council of Trent, Franciscans, Jesuits, Immaculate Conception, Papal Infallibility, and Vatican II
    \begin{enumerate}
        \item Mass
        \begin{itemize}
            \item Religious service for Catholics
            \item Focused mainly on Eucharist
            \item Was traditionally performed in Latin or Greek; later moved to vernacular
        \end{itemize}
        \item Council of Trent
        \begin{itemize}
            \item Council to make certain practices of the church into official dogma
            \item Cleared up disputes between different sects of Catholicism
        \end{itemize}
        \item Franciscans
        \begin{itemize}
            \item Order of monks founded by Francis of Assisi
            \item Live very simple lives
        \end{itemize}
        \item Jesuits
        \begin{itemize}
            \item
        \end{itemize}
        \item Immaculate Conception
        \begin{itemize}
            \item Involves Mary's conception of Jesus
            \item Catholics believe this leaves Mary sinless
        \end{itemize}
        \item Papal Infallibility
        \begin{itemize}
            \item Pope is direct line to God
            \item Pope's decrees are final
        \end{itemize}
        \item Vatican II
        \begin{itemize}
            \item Sought to bring the Church in to the modern age
            \item Discussed relationships with other religions
        \end{itemize}
    \end{enumerate}

    \item Who are the Methodists, Mormons, Quakers, and Fundamentalists?
    \begin{enumerate}
        \item Methodists
        \begin{itemize}
            \item Founded by John Wesley
            \item Theology focuses on the relationship between faith and character
            \item Focus on works that lead to salvation
        \end{itemize}
        \item Mormons
        \begin{itemize}
            \item Founded by Joseph Smith
            \item Believe their leader is a prophet of God
            \item Believe that Jesus came to America
        \end{itemize}
        \item Quakers
        \begin{itemize}
            \item Focus on finding God in each person
            \item Live removed lives
            \item Very simple dress/lifestyle
        \end{itemize}
        \item Fundamentalists
        \begin{itemize}
            \item Reaction to modernization of the church
            \item Believe that many theologians had strayed from the true teaching
            \item Sought to reform teachings by referring to literal text in the Bible
        \end{itemize}
    \end{enumerate}

    \item Explain Hinduism and Taoism
    \begin{enumerate}
        \item Hinduism
        \begin{itemize}
            \item 
        \end{itemize}
    \end{enumerate}
\end{enumerate}
\end{document}

### 2016-11-30 resume/education.tex
padics
working on final stuff

@@
 	      	      	      		Group Theory                               & Differential Equations & Technical Writing with \hologo{LaTeX} \\
-	      	      	      		Proofs and Discreet Mathematics            & Number Theory          & \textit{p-adic} Analysis              \\
+	      	      	      		Proofs and Discrete Mathematics            & Number Theory          & \textit{p-adic} Analysis              \\
 	      	      	      		Research in Partial Differential Equations & Mathematical Modeling  & Programming in Mathematica
@@
 	      	      	      		Data Structures        & Algorithms      & Computer Hardware and Design  \\
-	      	      	      		Discreet Mathematics & Writing and Computer Ethics
+	      	      	      		Discrete Mathematics & Writing and Computer Ethics
 	      	      	      	\end{tabular}
@@
 	      	      	      	\def\arraystretch{1.1}
-	      	      	      	\begin{tabular}{c  c  c}
-	      	      	      		\textbf{Proficient} & \textbf{Advanced}     & \textbf{Intermediate} \\
-	      	      	      		\hline
+	      	      	      	\begin{tabular}{l l l}
+	      	      	      		%       	      		\textbf{Proficient} & \textbf{Advanced}     & \textbf{Intermediate} \\
+	      	      	      		%       	      		\hline
 	      	      	      		Java/Android & PHP                   & C\#/Unity   \\

### 2016-11-30 software_design/report.tex
working on final stuff

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage{indentfirst}
\usepackage{marvosym}

\begin{document}
\section*{Open Source Project}

I began this semester with a few open source projects of my own that I've been maintaining. I had had some open source experience before; my capstone project is completely open sourced as well as a repository full of work I've done throughout my college career. Much of the open source work I've done, however, has not been collaborative. At the beginning of the semester, we read over \textit{The Mythical Man Month} which outlined various open source ideas. One of the ideas perpetuated by the author is that open source projects fall under Brook's Law, which states adding more manpower to a project delays the ultimate outcome of the project. In the open source world, however, this phenomenon seems to be absent. The project that we worked on, Habitica, is a prime example of an application making great strides towards completion while having hundreds of contributors. One of the reasons that open source projects grow so quickly towards their goals might actually be due to the lack of communication. Large groups having to communicate often leads to miscommunication; additionally, it takes a lot of time to meet, talk about what needs to be done, and ultimately get around to doing it.\\

Midway through the semester, my Apple laptop died, so I was forced to buy a new laptop. Since the newest models of MacBook Pros have only PCIe SSD's in them, I went with a Windows machine; since I don't care for windows, I decided to install Linux; since I was in the middle of the semester and didn't have the time to configure a version like Arch, I decided to install Ubuntu. I've noticed subtle bugs within the Ubuntu distribution that I have, such as the Dock getting stuck and the system not always obeying the Super key. I've been considering looking into the Ubuntu source to see if these issues are fixable by someone with my experience. Linux overall is an incredible example of open source in its prime. The idea that an entire operating system can be driven by an open source community is extraordinary. The main reason that I haven't looked into the source of Ubuntu or Linux is that I wouldn't have anywhere to test my changes and the source is not in a neat place such as Github (the kernal is on Github, but I'm nowhere near experienced enough to mess with it). The Linux operating system is not very different from OS X, so making the switch was not too hard. Originally, I wanted to use a fork of Ubuntu called Elementary OS. The user interface of Elementary OS is very close to that of OS X; ultimately I was not able to make it past the installation onto my computer.\\

Habitica was a very large project; I had never dealt with a project on the magnitude of it before. It was hard to find a bug that I though I could easily fix, much less actually fix it. I ultimately settled on an issue with notifications when purchasing gems.
\begin{quote}
    If you set your language to some of the non-English languages and then go to the Market to buy a gem with gold, you receive two notifications for \texttt{+1 gem}, one in your chosen language and then after a short delay, another one in English:
    \begin{center}
        \includegraphics[scale=0.4]{issue}
    \end{center}
    I've tested a few languages and it happens with only those that have a translation for the \texttt{plusOneGem} text that is different than the English version. For example, it happens with German and French, but not with \texttt{en\_GB}, \texttt{en\MVAt pirate}, or \texttt{zh\_TW} all of which still use the default text: \texttt{"plusOneGem": "+1 Gem"}
\end{quote}
To fix this bug we had to first reproduce it. In the local build for Habitica, there was already a button that would grant the user 10 gems, so I managed to wire it to make a purchase without authentication. The goal was to have the notification fire. Once we reproduced the bug, we needed to locate the place where the notification was being fired twice; it turned out to be in the \texttt{userServices.js} file. The way that Habitica determined whether or not a notification had fired was to compare the contents of the notification that should have fired to the contents of the notification that was being created. If they were the same string value, then the program would not create and fire the second notification; if they were different, however, then the second notification would fire. This was a problem only in languages that had translations for the word \textit{Gem} within Habitica. Since, for example, the French word \textit{Gemme} is not the same as the English translation \textit{Gem}, the notification fired again with \textit{Gem}. To fix this, we simply added a boolean that checked whether or not a notification had already been fired. If it had, then we wouldn't fire the second one; otherwise we would.
\end{document}

### 2016-12-01 religions_of_the_west/test_02.tex
done with 480

@@
             \item Creator God in Zoroastrianism
-            \item Sole God of the religion
+            \item Sole God of the religion (Zoroaster believed in one God)
             \item Followers worship his attributes (\textit{Amesha Spentas})
+            \item \textit{Amesha Spentas} are good spirits
             \item Had evil twin spirit named Angra Mainyu
-            \item Truth teller
+            \item Truth teller; truth was very important to Zoroastrianism
         \end{itemize}
@@
             \item Evil counterpart to Ahura Mazda
-            \item Can also create things
+            \item created Devas - evil spirits that paralleled \textit{Amesha Spentas}
+            \item Attempted to lead Zoroaster astray
+            \item Can also create things (evil is a creative force)
             \item Lier and deceiver (similar to Satan)
@@
 
-    \item List some similarities between Zoroastrianism, Christianity, and post-exilic Judaism.
+    \item List some similarities between Zoroastrianism, Christianity, and post-exile Judaism.
     \begin{itemize}
         \item Central God; Zoroastrianism: Ahura Mazda, Judaism/Christianity: Yahweh
-        \item Some sects of Judaism look for Messiah; Christianity has Jesus; Zoroastrianism looked for Saoshyant\
+        \item Morality comes from God
+        \item Dualism - contest between good an evil (only some sects of Judaism)
+        \item Some sects of Judaism look for Messiah; Christianity has Jesus; Zoroastrianism looked for Saoshyant
+        \item Coming of God - 1000 years
         \item Similar eschatology (final judgment, raising of the dead)
@@
             \begin{itemize}
+                \item Nature of Father and Son was very important topic
                 \item Father and son have same nature but are not the same
@@
             \end{itemize}
+            \item Arius: Jesus was separate from God
+            \begin{itemize}
+                \item Jesus was first creation of God
+                \item Jesus was not divine
+                \item Rejected by council
+            \end{itemize}
+            \item Difference between \textit{made} (Arius) and \textit{begotten} (Athanasius)
             \item Trinity idea proposed (later rejected by enlightenment Christians)
@@
         \begin{itemize}
-            \item Council met to settle disputes over Jesus' nature
+            \item Council met to settle disputes over Jesus' nature (divine vs. human)
             \item Greek ideas (God is separate) clashing with Hebrew ideas (God is with humanity)
             \item Leo's Tome: Jesus has two full natures in one person (Orthodox position)
+            \item Solution proposed by Leo (bishop of Rome)
+            \begin{itemize}
+                \item Decided Jesus was fully God an fully man
+                \item United in one person
+                \item Hypostatic union
+                \item Gave Leo position of bishop of bishops
+            \end{itemize}
             \item Established Mary's position in the church (Humans < Mary < God)
-            \item Roman bishop gains authority at this council
+            \item Gave Mary the status of mother of God
         \end{itemize}
@@
         \item Bishops are established throughout the known world
-        \item Chalcedon: Bishop of Rome gains authority in establishing Doctrine
-        \item Bishop of Rome becomes authority of the Church (i.e. Pope)
+        \item Chalcedon: Bishop of Rome (Leo) gains authority in establishing Doctrine
+        \item Gregory the Great is the first recognized Pope
+        % \item Innocent III and Gregory VII
+        % Innocent III and Gregory VII, nation states, loss of papal power, Vatican I and II
         \item Pope reigns above kings in the middle ages
+        \item Holy Roman empire formed by Constantine
         \item Church of England splits from the Catholic church over disputes (power of Pope weakened)
@@
 
-    \item What do Pelagius, Augustine, Erasmus, and Calvin believe about God's sovereignty and human freedon?
+    \item What do Pelagius, Augustine, Erasmus, and Calvin believe about God's sovereignty and human freedom?
     \begin{enumerate}
@@
             \item Good deeds will lead you to heaven
+            \item Possibility to serve God
             \item Did not believe in God's saving grace
+            \item Reacted against Augustine's teachings
         \end{itemize}
@@
             \item Humans were predestined
+            \item Give what you command and command what you will
             \item God selects certain people to receive his grace
+            \begin{itemize}
+                \item Not all who receive grace are selected (can lose grace)
+            \end{itemize}
+            \item Human beings are totally depraved (only God's grace can save)
             \item Only those who have been given the grace of God can follow
@@
             \item Grace of God and works lead to salvation
+            \item God's grace is far above the good works (do good through God's grace)
             \item Humans have free will to follow or reject God
@@
             \item Followed by Catholic church
+            \item Salvation can be lost/reacquired
         \end{itemize}
@@
         \begin{itemize}
-            \item Humans have no free will
+            \item Humans have no free will (impunes sovereignty of God)
+            \begin{itemize}
+                \item Double predestination
+                \item Some are chosen for Heaven, others for Hell
+            \end{itemize}
             \item Salvation comes \textbf{only} from the grace of God
-            \item Humans are inherently tainted with sin
-            \item Some are chosen for heaven; others for hell
+            \item Humans are inherently tainted with sin (total depravity)
+            \begin{itemize}
+                \item Every aspect is tainted with iniquity
+            \end{itemize}
             \item Strong emphasis on God's role in human life

### 2016-12-01 software_design/report.tex
done with 480

@@
 \section*{Open Source Project}
+\subsection*{Steven Rosendahl}
 
@@
 Large groups having to communicate often leads to miscommunication; additionally, it takes a lot of time to meet, talk about what needs to be done, and ultimately get around to doing it.\\
+
+One of my least favorite things we did this semester was using Angular.js.
+I've never been a fan of JavaScript in the first place.
+Not ever knowing what type of Object you have as well as creating Objects on the fly with no schema makes the process of debugging and contributing very difficult.
+I also dislike frameworks/languages that force you into a certain design pattern; this is one of the reasons I've never been big into iOS programming.
+With so many JavaScript frameworks out there, Angular.js is a small fish in a big pond.
+Python was easier to use due to its structure over JavaScript, but much of the open source world is stuck in Python 2.7 whereas the standard is slowly becoming Python 3.
+This is a little forgivable, since the tools to convert back and forth between the versions are virtually non-existent.
+This is definitely a problem that could be fixed by the open source world, but it doesn't appear to be a major focus of the majority of projects, especially the scientific ones for which Python is best known.
+This behavior seems to appeal to the idea that we shouldn't fix what isn't broken, and being a major version number behind doesn't qualify a project as broken, so it is understandable why many large open source projects are not focusing on migration to Python 3.\\
 
@@
 
+About midway through the semester, I encountered a problem with an application that I was writing for my job.
+The issue involved an \texttt{SSLException} that was being thrown by versions of Android greater than 6.0.
+In my search for the answer to why this was happening, I came across the open Android project, which is the open source code for the Android operating system.
+I knew my issue was arising from Android's native implementation of \texttt{okhttp}, so I spent a while trying to pinpoint my error.
+I was able to narrow it down to something happening in the handshake, but I couldn't go any further since the stack trace given by LogCat was not specific enough to tell me exactly what step in the handshake process was causing an error.
+After talking with Dr.
+Koehl, he recommended that I look into the server logs for the particular server that was hosting the data I needed.
+When I went to our server manager, he noted that the server was rejecting the certificate that my phone was using to connect; his best suggestion was to look into the source code of the program that was running on the server.
+He directed me to the appropriate repository, and I began looking through to see if I could find anything about certificate validation and why it would be rejecting the requests made from Android 7.0.
+Ultimately, I was not able to solve this issue (it's still ongoing).
+Both the Android OS and the code for the server are open source, so it may be possible that others in the community will be able to fix this issue.\\
+
 Habitica was a very large project; I had never dealt with a project on the magnitude of it before.
+The first step was to get Habitica installed.
+Since my computer had just recently been setup with Ubuntu and I'm not a JavaScript developer, I had to install Node.js, Bower, and all of the dependencies.
+The developer's wiki helped marginally, but I was able to get around many of my troubles by performing the installations essentially backwards.
+This all had to be run under \texttt{sudo}, since Ubuntu is sensitive to touching anything above the \texttt{{\raise.17ex\hbox{$\scriptstyle\sim$}}/} directory.
+Others in my group found it a little more difficult and had to resort to using Vagrant/Docker to get things up and running since \texttt{npm} wouldn't play nicely.
 It was hard to find a bug that I though I could easily fix, much less actually fix it.
-I ultimately settled on an issue with notifications when purchasing gems.
+The first few bugs we looked at involved documenting the Habitica API.
+Unfortunately, this was near impossible without having used Habitica extensively \textbf{and} having knowledge of the public APIs that were provided.
+We abandoned that idea an moved on to try and fix an issue with notification not appearing on certain critical hits.
+I began to look extensively into this issue, but I could not find the place in code where notifications were fired for critical hits.
+We posted in the developer's chat on Habitica's website, but their suggestions did not help very much.
+Finally, we found another issue which Matthew promptly claimed.
+The issue we ultimately settled on involved problems with notifications when purchasing gems.
 \begin{quote}
@@
 If it had, then we wouldn't fire the second one; otherwise we would.
+This appears to have fixed the problem; the only thing left to do now is test that the problem is truly fixed.
+It might be the case that the second check to see if a notification was fired is necessary; as such, there may be another issue that arises by just ignoring it.\\
+
+Overall, I think open source projects are one of the greatest developments in software engineering.
+Being able to root though code yourself and fix your own issues is truly a homage to the idea that \textit{if I don't do it, nobody will}.
+I've come across a lot of bugs in a lot of places, such as wrestling with a seemingly impossible \texttt{SSLException} in Android to the minor annoyance of broken syntax highlighting in Github's Atom with \LaTeX.
+The open source community is so influential that many companies such as Microsoft, Apple, and Google support and encourage the use and creation of open source software within their own company.
+Even if I don't have the skills to fix the problem (or the want to fix the problem), chances are someone out there does.
+That is the true beauty of open source.
 \end{document}

### 2016-12-01 software_design/datareader/main.py

@@ -1,3 +1,6 @@
+#Steven rosendahl
+#Donnie Darko Simulator
+
 from clint.textui import colored
 from csv import reader
 

### 2016-12-04 religions_of_the_west/test_02.tex
finished rstd

@@
 
-    \item Describe the evolution of the Papacy.
-    \begin{itemize}
-        \item Peter considered to be head of Church after Jesus' death
-        \item Bishops are established throughout the known world
-        \item Chalcedon: Bishop of Rome (Leo) gains authority in establishing Doctrine
-        \item Gregory the Great is the first recognized Pope
-        % \item Innocent III and Gregory VII
-        % Innocent III and Gregory VII, nation states, loss of papal power, Vatican I and II
-        \item Pope reigns above kings in the middle ages
-        \item Holy Roman empire formed by Constantine
-        \item Church of England splits from the Catholic church over disputes (power of Pope weakened)
-        \item Other theologians begin to reject Church's teaching (Martin Luther, John Calvin, etc.) further weakening the Pope's authority
-    \end{itemize}
+    % \item Describe the evolution of the Papacy.
+    % \begin{itemize}
+    %     \item Peter considered to be head of Church after Jesus' death
+    %     \item Bishops are established throughout the known world
+    %     \item Chalcedon: Bishop of Rome (Leo) gains authority in establishing Doctrine
+    %     \item Gregory the Great is the first recognized Pope
+    %     % \item Innocent III and Gregory VII
+    %     % Innocent III and Gregory VII, nation states, loss of papal power, Vatican I and II
+    %     \item Pope reigns above kings in the middle ages
+    %     \item Holy Roman empire formed by Constantine
+    %     \item Church of England splits from the Catholic church over disputes (power of Pope weakened)
+    %     \item Other theologians begin to reject Church's teaching (Martin Luther, John Calvin, etc.) further weakening the Pope's authority
+    % \end{itemize}
 
-    \item What do Pelagius, Augustine, Erasmus, and Calvin believe about God's sovereignty and human freedom?
-    \begin{enumerate}
-        \item Pelagius
-        \begin{itemize}
-            \item Humans have free will
-            \item Good deeds will lead you to heaven
-            \item Possibility to serve God
-            \item Did not believe in God's saving grace
-            \item Reacted against Augustine's teachings
-        \end{itemize}
-        \item Augustine
-        \begin{itemize}
-            \item Humans were predestined
-            \item Give what you command and command what you will
-            \item God selects certain people to receive his grace
-            \begin{itemize}
-                \item Not all who receive grace are selected (can lose grace)
-            \end{itemize}
-            \item Human beings are totally depraved (only God's grace can save)
-            \item Only those who have been given the grace of God can follow
-            \item Works on Earth do not matter
-        \end{itemize}
-        \item Erasmus
-        \begin{itemize}
-            \item Grace of God and works lead to salvation
-            \item God's grace is far above the good works (do good through God's grace)
-            \item Humans have free will to follow or reject God
-            \item Strong focus on working with God
-            \item Followed by Catholic church
-            \item Salvation can be lost/reacquired
-        \end{itemize}
-        \item Calvin
-        \begin{itemize}
-            \item Humans have no free will (impunes sovereignty of God)
-            \begin{itemize}
-                \item Double predestination
-                \item Some are chosen for Heaven, others for Hell
-            \end{itemize}
-            \item Salvation comes \textbf{only} from the grace of God
-            \item Humans are inherently tainted with sin (total depravity)
-            \begin{itemize}
-                \item Every aspect is tainted with iniquity
-            \end{itemize}
-            \item Strong emphasis on God's role in human life
-        \end{itemize}
-    \end{enumerate}
+    % \item What do Pelagius, Augustine, Erasmus, and Calvin believe about God's sovereignty and human freedom?
+    % \begin{enumerate}
+    %     \item Pelagius
+    %     \begin{itemize}
+    %         \item Humans have free will
+    %         \item Good deeds will lead you to heaven
+    %         \item Possibility to serve God
+    %         \item Did not believe in God's saving grace
+    %         \item Reacted against Augustine's teachings
+    %     \end{itemize}
+    %     \item Augustine
+    %     \begin{itemize}
+    %         \item Humans were predestined
+    %         \item Give what you command and command what you will
+    %         \item God selects certain people to receive his grace
+    %         \begin{itemize}
+    %             \item Not all who receive grace are selected (can lose grace)
+    %         \end{itemize}
+    %         \item Human beings are totally depraved (only God's grace can save)
+    %         \item Only those who have been given the grace of God can follow
+    %         \item Works on Earth do not matter
+    %     \end{itemize}
+    %     \item Erasmus
+    %     \begin{itemize}
+    %         \item Grace of God and works lead to salvation
+    %         \item God's grace is far above the good works (do good through God's grace)
+    %         \item Humans have free will to follow or reject God
+    %         \item Strong focus on working with God
+    %         \item Followed by Catholic church
+    %         \item Salvation can be lost/reacquired
+    %     \end{itemize}
+    %     \item Calvin
+    %     \begin{itemize}
+    %         \item Humans have no free will (impunes sovereignty of God)
+    %         \begin{itemize}
+    %             \item Double predestination
+    %             \item Some are chosen for Heaven, others for Hell
+    %         \end{itemize}
+    %         \item Salvation comes \textbf{only} from the grace of God
+    %         \item Humans are inherently tainted with sin (total depravity)
+    %         \begin{itemize}
+    %             \item Every aspect is tainted with iniquity
+    %         \end{itemize}
+    %         \item Strong emphasis on God's role in human life
+    %     \end{itemize}
+    % \end{enumerate}
 
@@
             \item Jesus was punished for sins of humanity
+            \item Christ died in place of other people
+            \item Righteous and wrathful God
             \item God forgives all sins through this sacrifice
@@
         \begin{itemize}
-            \item
+            \item Emphasizes resurrection over the suffering
+            \item God gains victory over devil
+            \item Christ conquers death in the realm of the devil through resurrection
         \end{itemize}
@@
         \begin{itemize}
+            \item Calvinists identify with this position
             \item
@@
         \item Moral Influence
+        \begin{itemize}
+            \item Jesus' purpose was to bring moral teaching to humanity
+            \item Sets moral example for self-sacrificing love
+            \item Most popular theory
+        \end{itemize}
     \end{enumerate}
 
-    \item Differentiate the Catholic, Lutheran, Reformed, and Anabaptist theories of the sacraments (baptism and eucharist)
+    \item Differentiate the Catholic, Lutheran, Reformed, and Anabaptist theories of the sacraments of Baptism and the Eucharist
     \begin{enumerate}
@@
             \item Eucharist is \textit{Transubstantiation}: transforming into literal body and blood
+            \item Eucharist is re-sacrificing of Christ
             \item Eucharist is central focus of the mass (sacrifice)
@@
             \item Original Sin is forgiven through baptism
-            \item Eucharist is \textit{Consubstantiation}: elements are the same but Christ is present in them
+            \item Eucharist is \textit{Consubstantiation}: elements are the same but Christ is present in, with, and under them
         \end{itemize}
@@
             \item Baptism of infants
-            \item Baptism is an oath to bring up child in name of Christ
+            \item Baptism is an oath by the congregation to bring up child in name of Christ (covenant)
             \item Baptism is Symbol of the Covenant
+            \item Do not believe in baptismal regeneration
             \item Eucharist is symbolic of Christ; used to remember his sacrifice
@@
         \begin{itemize}
-            \item Baptism of adults
+            \item Not a sacramental church
+            \item Baptism of adults (child is innocent until adulthood)
             \item Baptism is symbolic of conversion and acceptance to Christ
@@
     \begin{itemize}
+        %chapters 1 and 3 of purple book
+        \item Enlightenment philosophers followed Puritan ideals
+        \begin{itemize}
+            \item God's will revealed in nature
+            \item Will is known by all humans regardless of position in life
+            \item Preservation of the truth must be governments main job
+            \item Natural laws $\to$ rights (Life, liberty, possessions)
+        \end{itemize}
+        \item Puritans came to new world to avoid religious persecution
+        \begin{itemize}
+            \item Locke believed in ideas of religious tolerance
+            \item Government should protect all rights of the people
+        \end{itemize}
         \item Puritans had democratic structure in churches
-
+        \begin{itemize}
+            \item Congregation was self sufficient
+            \item Bound by its own covenant
+            \item Federal government is a covenant by the people
+            \item Breaking of the covenant leads to revolution
+        \end{itemize}
+        \item Puritans focused on equality
+        \begin{itemize}
+            \item Following example of Jesus
+            \item Followers of Jesus were simple, blue collar
+            \item Lead to ideas of equality in America
+        \end{itemize}
+        \item Puritans held property
+        \begin{itemize}
+            \item Goal of life was not to become rich, but to better society through contributions
+            \item Puritans were very hard workers
+            \item Lead to capitalistic ideas (Benjamin Franklin)
+        \end{itemize}
+        \item Various Enlightenment philosophers liked the ideas but not the religion
+        \begin{itemize}
+            \item Saw Church as oppressive
+            \item Wanted to move away from hierarchy
+        \end{itemize}
     \end{itemize}
@@
             \item Religious service for Catholics
-            \item Focused mainly on Eucharist
+            \item Focused mainly on Eucharist (transubstantiation)
             \item Was traditionally performed in Latin or Greek; later moved to vernacular
@@
         \begin{itemize}
+            \item Condemned certain beliefs of protestants
             \item Council to make certain practices of the church into official dogma
+            \begin{itemize}
+                \item Transubstantiation, Baptismal regeneration
+            \end{itemize}
+            \item Catholics follow scripture \textit{and} tradition
+            \item Latin Vulgate is the accepted Bible
+            \item Catholic church has sole right of scriptural interpretation
+            \item Justification rests on faith, but not faith alone
             \item Cleared up disputes between different sects of Catholicism
@@
             \item Order of monks founded by Francis of Assisi
-            \item Live very simple lives
+            \item Live very simple lives (no possessions)
+            \item Administer to the poor
+            \item Famed for animal husbandry
         \end{itemize}
@@
         \begin{itemize}
-            \item
+            \item Founded by Ignacius Loyola
+            \item Felt that other orders were tainted
+            \item Tend to follow the pope directly (loyalty to Rome)
+            \item Soldiers of Christ
         \end{itemize}
@@
         \begin{itemize}
-            \item Involves Mary's conception of Jesus
+            \item Mary was born without the taint of original sin
             \item Catholics believe this leaves Mary sinless
@@
         \begin{itemize}
-            \item Pope is direct line to God
+            \item Pope is direct line to God when speaking from his chair (ex cathedra)
             \item Pope's decrees are final
+            \item Does not mean the pope is sinless
         \end{itemize}
@@
         \begin{itemize}
-            \item Sought to bring the Church in to the modern age
+            \item John 23rd - Paul in the 1960's
+            \item Sought to bring the church in to the modern age
+            \item Liberalized the church
+            \item Promoted vernacular in services as opposed to Latin
             \item Discussed relationships with other religions
+            \begin{itemize}
+                \item Accepted other religions
+                \item Did away with antisemitism
+            \end{itemize}
+            \item Less authority in the papacy
         \end{itemize}
@@
             \item Founded by John Wesley
+            \item Assurance of salvation
+            \item Doctrine is not as important as living a good life
             \item Theology focuses on the relationship between faith and character
@@
             \item Founded by Joseph Smith
+            \begin{itemize}
+                \item Says he experienced visitation from God
+                \item Saw corruption of church and bible
+                \item Received his revelation from angel Maroni
+            \end{itemize}
+            \item Church of Latter Day Saints
+            \item Book of Mormon
             \item Believe their leader is a prophet of God
@@
         \begin{itemize}
+            \item Founded during puritan revolution by George Fox
+            \begin{itemize}
+                \item Outgrowth of puritan theology
+            \end{itemize}
+            \item William Penn is major figure
+            \item Focus on pacifism
             \item Focus on finding God in each person
+            \item Focus on equality (men and women, white and black)
             \item Live removed lives
@@
             \item Reaction to modernization of the church
+            \item Very conservative views
+            \begin{itemize}
+                \item Liberals began to reject liter aspects of the Bible
+                \item Proposed 5 fundamentals
+                \begin{enumerate}
+                    \item Scripture is infallible
+                    \item Jesus was born of a virgin
+                    \item Christ's death was an atonement of sin
+                    \item Jesus was raised from the dead
+                    \item Jesus actually performed miracles
+                \end{enumerate}
+            \end{itemize}
             \item Believe that many theologians had strayed from the true teaching
@@
 
-    \item Explain Hinduism and Taoism
-    \begin{enumerate}
-        \item Hinduism
+    \item Explain Taoism
     \begin{itemize}
-            \item
+        \item Founded by Lao Tzu
+        \begin{itemize}
+            \item Saw the ethical followers of Confucianism as too proud
+            \item Believed that ethics lead to people who believed they were better than others
+            \item Believed that society had forced people to become aggressive
+            \item Thought that people were naturally free
+        \end{itemize}
+        \item Wu We: \textit{not doing}
+        \begin{itemize}
+            \item Taoism focuses on letting things go
+            \item Letting people have their own lives is very important
+        \end{itemize}
+        \item The Dao is the driving force in life
+        \begin{itemize}
+            \item Separate from humans
+            \item Dao is balanced - Humans should try to be balanced
+            \item Yin and Yang is symbol of the balance of life
+        \end{itemize}
     \end{itemize}
 \end{enumerate}
-\end{enumerate}
 \end{document}
